### 2016-09-26 outline.rmd
adding ideas for talk.

# 1. Consumption and shopping
- largest supermarket chain in UK, famous for Clubcard starting in 1991; often cited as early example of Big Data Analytics
-

# 2. The web and the shopping basket
- online commerce -- transforming shopping;

>Tesco is the clear winner in the online grocery market, in fact it takes almost 50p of every £1 spent on food shopping on the internet. 
(https://econsultancy.com/blog/64841-how-seo-helps-tesco-to-dominate-the-online-grocery-market/)[https://econsultancy.com/blog/64841-how-seo-helps-tesco-to-dominate-the-online-grocery-market/]

# 3. From the rules-based list to a 'relevancy model'
How would I characterise this shift? Use it to emphasize why attention to the algorithm is not enough -- it is the model that counts
A little history of recommender systems -- uses Varian 1997 for a start. But has no real reference list. Not sure  
Also can use Schafer -- back to Association Rule Mining from early 1990s

# 4. Hack? Customers go straight to the checkout
people inhabit lists and work with them; no matter how they are made; cf ethnomethodology

# 5. Data: 52 weeks of data for each customer
- history of data accumulations;
- shift from data mining to personalization?

# 6. Gain: use good features to build models

# 7. Output 20 recommendations from 200,000 products
The problem of too many things to list -- this must be happening in many settings -- videos, social networks, products, etc.

# 8. Value: Balancing conversion vs spend
The value of a production

# 9. Metric: precision-at-K

# 10. Experiment: A/B allocation
Could use the Google Infrastructure of experiment or Facebook Flow stuff here -- 1 million models

# 11. Infrastructure: realtime engine serves up top 5
Only in certain settings are things realtime; or only in limited ways;
Realtime has its own economies and cultures

# 12. Similarity, substitution and complementarity
- interesting contrasts here between the relations of things in a list

# 13. Complementarity drives conversion 
- the problem of action

### 2016-09-27 outline.rmd
starting proper writing

@@
-# 1. Consumption and shopping
-- largest supermarket chain in UK, famous for Clubcard starting in 1991; often cited as early example of Big Data Analytics
--
+# 'Have you forgotten'
 
-# 2. The web and the shopping basket
-- online commerce -- transforming shopping;
+# Introduction
+
+Presentation presents another presentation I saw recently about online grocery shopping at Tesco, a large supermarket chain based in the UK.
+The lists in this setting are recommendations for further purchases under the title of 'Have you forgotten?'.
+I regard this list of just five suggestions given to Tesco's online customers as symptomatic of a major contemporary form of list, the recommendation list, that we see in many  forms of news, media, entertainment, marketing, advertising.
+Shopping lists are quite prosaic -- a dozen lines scrawled on the back of a used envelope stuck on the refrigerator door -- and sublime -- a million products ranked in some kind of order for tens or hundreds of millions of people.
+If we are interested in lists from various angles -- asignifying semiotics as Maurizio Lazzarato calls it in his _Signs and Machines_ [@Lazzarato_2014] or 'a new order of spatio-temporal continuity for forms of economic, political and cultural life' as Celia Lury terms in her account of topological turn [@TBA], or even in terms of different cultural practices of listing [@Jullien_TBA] -- then contemporary lists practices such as recommendations will probably not only objects of interest, but raise the question of how we institute our approach to them.
+While lists such as recommendations intersect strongly with meaning and sense-making practices, they are also somewhere refractory to methods of humanities and social science research.
+For on the one hand, lists are highly associative. They can be interpreted or decoded semiotically, but they also exhibit great variations in textuality that limit semiotic readings.
+For on the other hand, as operational entities, they can be treated ethnomethodologically, as the production of ('the order of social settings through ...
+shared sense making practices.')[https://en.wikipedia.org/wiki/Ethnomethodology]  The question for me here is whether contemporary lists can still be understood as produced through 'shared sense-making practices.'
+All lists are forms of order, since even random ordering is operationally important, but what counts as order might historically transformable, in which case the order of social settings or the problem of social order might be deeply historically contingent.
+The question of order or the problem of order lies at the core of lists making, and our relation to lists can be seen as a test of the way that order operates more generally. 
+Much of my paper, and my responses to the presentation I saw from a data scientist  will focus on these two related questions of order and new forms of social order, and participation and its contemporary forms.
+I respond to the presentation in two ways -- one by recounting its step as an example of the sense-making done by data scientists and customer analytics teams in many places, and secondly as an occasion to intervene and participate in the list-making practices they engage in.
+
+# 1. The presentation - Consumption and shopping
+- largest supermarket chain in UK, famous for Clubcard starting in 1991; often cited as a leading early example of Big Data Analytics. 
+- the problems of scale: Tesco is like Amazon or Netflix in this respect: too many people and too many products
 
@@
 
-# 3. From the rules-based list to a 'relevancy model'
+# 2. The web and the shopping basket: how it was reconstituted in the 1990s
+- online commerce -- transforming shopping; 
+- personalization and recommendation as the killer apps for social media as a commercial process, but also in government etc.
+
 How would I characterise this shift? Use it to emphasize why attention to the algorithm is not enough -- it is the model that counts
@@
 Also can use Schafer -- back to Association Rule Mining from early 1990s
+
+DunnHumby -- Edwina Dunn and Clive Humby -- are famous for having convinced the CEO of Tesco sometime in 1991 that a loyalty card program could change the supermarket chain's relationship to its customers.
+[@Humby_1984]
+
+![Humby, 1984]( figure/Humby_young0R_1994.png )
+
+When commentators first began talking about recommender systems in the 1990s, they highlighted the shift from demographic-driven market research techniques to personalized recommendations.
+For instance, writing in 1997, Paul Resnick and Hal Varian (at that time Dean of Information Sciences at UC Berkeley, but somehow now Chief Economist at Google), made much of this difference [@Resnick_1997].
+
+# 3. From the rules-based list to a 'relevancy model'
+
+Shreena described her work at DunnHumby in terms of a shift in models from a 'rules-based list' to a 'relevancy model.' 
+This was a significant change in how the 'Have You Forgotten' list was constructed.
+
+I assume that the rules-based model derived from the early 1990s work on the `apriori` algorithm [@Agrawal_1993TBA]
+
+```{r arules, echo=TRUE}
+library(arules)
+library(arulesViz)
+library(datasets)
+data(Groceries)
+itemFrequencyPlot(Groceries,topN=20,type="absolute")
+# Get the rules
+rules <- apriori(Groceries, parameter = list(supp = 0.001, conf = 0.8))
+
+# Show the top 5 rules, but only 2 digits
+options(digits=2)
+inspect(rules[1:5])
+summary(rules)
+
+rules<-sort(rules, by="confidence", decreasing=TRUE)
+inspect(rules[1:5])
+rules<-apriori(data=Groceries, parameter=list(supp=0.001,conf = 0.15,minlen=2), 
+           appearance = list(default="rhs",lhs="whole milk"),
+           control = list(verbose=F))
+rules<-sort(rules, decreasing=TRUE,by="confidence")
+inspect(rules[1:5])  
+plot(rules,method="graph",interactive=TRUE,shading=NA)
+```
 

### 2016-09-28 outline.rmd
trying to run the set argument

@@
-# 'Have you forgotten'
+# 'Have you forgotten?'
 
 # Introduction
-
 Presentation presents another presentation I saw recently about online grocery shopping at Tesco, a large supermarket chain based in the UK.
 The lists in this setting are recommendations for further purchases under the title of 'Have you forgotten?'.
-I regard this list of just five suggestions given to Tesco's online customers as symptomatic of a major contemporary form of list, the recommendation list, that we see in many  forms of news, media, entertainment, marketing, advertising.
+I regard this list of just five suggestions given to Tesco's online customers as symptomatic of a major contemporary form of ordering, the recommendation list, that we see in many  forms of news, media, entertainment, marketing, advertising.
 Shopping lists are quite prosaic -- a dozen lines scrawled on the back of a used envelope stuck on the refrigerator door -- and sublime -- a million products ranked in some kind of order for tens or hundreds of millions of people.
-If we are interested in lists from various angles -- asignifying semiotics as Maurizio Lazzarato calls it in his _Signs and Machines_ [@Lazzarato_2014] or 'a new order of spatio-temporal continuity for forms of economic, political and cultural life' as Celia Lury terms in her account of topological turn [@TBA], or even in terms of different cultural practices of listing [@Jullien_TBA] -- then contemporary lists practices such as recommendations will probably not only objects of interest, but raise the question of how we institute our approach to them.
+If we are interested in lists from various angles -- asignifying semiotics as Maurizio Lazzarato calls it in his _Signs and Machines_ [@Lazzarato_2014] or 'a new order of spatio-temporal continuity for forms of economic, political and cultural life' as Celia Lury terms in her account of topological turn [@TBA], or even in terms of different cultural practices of listing [@Jullien_TBA] -- then contemporary lists practices such as recommendations figure as sites of intersection between different forms.
+Lists also raise the question of how we institute our approach to them.
 While lists such as recommendations intersect strongly with meaning and sense-making practices, they are also somewhere refractory to methods of humanities and social science research.
@@
 All lists are forms of order, since even random ordering is operationally important, but what counts as order might historically transformable, in which case the order of social settings or the problem of social order might be deeply historically contingent.
-The question of order or the problem of order lies at the core of lists making, and our relation to lists can be seen as a test of the way that order operates more generally. 
-Much of my paper, and my responses to the presentation I saw from a data scientist  will focus on these two related questions of order and new forms of social order, and participation and its contemporary forms.
-I respond to the presentation in two ways -- one by recounting its step as an example of the sense-making done by data scientists and customer analytics teams in many places, and secondly as an occasion to intervene and participate in the list-making practices they engage in.
+The problem of order lies at the core of lists making, and our relation to lists can be seen as a test of the way that order operates more generally. 
+Much of my paper, and my responses to the presentation I saw from a data scientist  will focus on how the ordering that is done in a recommendation lists relates to the problem of social order more generally.
+- The main narrative of the presentation concerned a shift from a set-based model for making lists to a predictive, probabilistic way of making lists in almost realtime and dealing with lists on a scale that is often difficult to grasp concretely.
+I respond to the presentation in two ways -- firstly by recounting, contextualising and commenting on some of the main steps as a example of the sense-making done by data scientists and customer analytics teams in many places; and secondly, by trying to step into the list-making practices  as an occasion to intervene and participate in the list-making practices they engage in.
 
-# 1. The presentation - Consumption and shopping
-- largest supermarket chain in UK, famous for Clubcard starting in 1991; often cited as a leading early example of Big Data Analytics. 
+# 1. The context of the presentation - Consumption and shopping
+- presentation at an academic event with speakers from industry and government on statistics and operations research in industry. Speakers came from Shell, Disney, GCHQ, and various universities. 
+- Shreena Patel works as a data scientist for DunnHumby focusing on online grocery shopping at Tesco.
+Tesco is the largest supermarket chain in UK, famous for Clubcard starting in 1991; often cited as a leading early example of Big Data Analytics. 
+DunnHumby -- Edwina Dunn and Clive Humby -- are famous for having convinced the CEO of Tesco sometime in 1991 that a loyalty card program could change the supermarket chain's relationship to its customers.
+[@Humby_1984]
+
+![Humby, 1984]( figure/Humby_young0R_1994.png )
 - the problems of scale: Tesco is like Amazon or Netflix in this respect: too many people and too many products
@@
 
-DunnHumby -- Edwina Dunn and Clive Humby -- are famous for having convinced the CEO of Tesco sometime in 1991 that a loyalty card program could change the supermarket chain's relationship to its customers.
-[@Humby_1984]
-
-![Humby, 1984]( figure/Humby_young0R_1994.png )
-
 When commentators first began talking about recommender systems in the 1990s, they highlighted the shift from demographic-driven market research techniques to personalized recommendations.
@@
 This was a significant change in how the 'Have You Forgotten' list was constructed. 
-
-I assume that the rules-based model derived from the early 1990s work on the `apriori` algorithm [@Agrawal_1993TBA]
+I assume that the rules-based model derived from the early 1990s work on the `apriori` algorithm by Rakesh Agrawal and Ramakrishnan Srikanti at IBM Research Alameda [@Agrawal_1994].
+The `apriori` algorithm finds sets of items that commonly occur in transactions. 
+These commonly occurring sets are expressed as association rules. 
+For instance, using the `Groceries` dataset in the `R` package `arules,` we can readily see how Walmart might have discovered  -- but didn't -- the fabled association between diapers and beer.
+[http://www.dssresources.com/newsletters/66.php]
+It is a key algorithm in 'market basket analysis' and attempts to re-structure shopping lists.
 
@@
 plot(rules,method="graph",interactive=TRUE,shading=NA)
+items = Groceries@itemInfo
+head(items)
+nrow(items)
 ```
 
+The association rule-based model for the `Groceries` dataset suggests some of the characteristic ordering practices of data-mining approaches to lists. 
+The association between items is expressed in the form of _rules_. These rules, a term derived from decision support literature, define sets of associations between items. 
+As a set theory-based practice, the association rules define likely sets of items or shopping baskets in terms of decreasing probability.   
+`r inspect(rules[1])` is a stronger association than `r inspect(rules[2])`.
+So any recommendation in the 'Have You Forgotten' list should probably put them in that order. 
+Association rules and the `apriori` algorithm present social order in terms of sets of common associations. 
+This means that that the algorithm can be understood in terms of the set-based multiples proposed by Alain Badiou. 
+Badiou in his account of multiples highlight the contrast between two different ways of grasping a multiple -- in terms of the 'numerical one' and the 'vital One' [@Badiou_2004, 72], between the 'closed' and the 'open' respectively.
+The `apriori` algorithm, might be seen as an attempt to deal with the excess of things in a Tesco supermarket, and the lack of interest or engagement that people have with them by identifying subsets whose membership is stronger by virtue of the other sets of associations that people bring with them.
+Association is stronger when it derives from an excess of different ways of being in sets.  
+Badiou's argument is that being part of a set differs from belonging to a set.
+An open set arises from when we dwell within a set by 'taking the intersection of two elements.' 
+What might this mean in terms of ordering?
+If we take the itemset from the `Groceries` dataset [@Hahsler_2006], and then run it through the Tesco online groceries site, each of the 169 items in the dataset yields dozens and sometimes thousands of results from Tesco online.
+Each of the items in the `Groceries` dataset proliferates in the Tesco site, into a new list of branded variations.
+
 # 4. Hack? Customers go straight to the checkout
-people inhabit lists and work with them; no matter how they are made; cf ethnomethodology
 
-# 5. Data: 52 weeks of data for each customer
+Patel observed that the 'Have you forgotten' lists matter to people.
+Some people log on and go straight to the checkout, allowing the recommendation system to supply them with the suggestions that they then move to the shopping basket.
+This is the online equivalent of using an old shopping list and only adding new items when needed. 
+It is also a strategy to avoid having to view too many lists, since the only way Tesco can show people what they might buy is by showing them lists of things (e.g.
+chicken products), but these lists are too long to work with.
+People inhabit lists and work with them no matter how they are made; thus vindicating ethnomethodologists assertion that social order is the product of ongoing sense-making practices.
+
+As Badiou might say,
+
+# 5. Data: 52 weeks of data for each customer: how is order constructed -- from history?
 - history of data accumulations;
@@
 # 6. Gain: use good features to build models 
+- what counts as a good feature? See Domingos on this
 
-# 7. Output 20 recommendations from 200,000 products
+# 7. Sparsity: output 20 recommendations from 200,000 products
 The problem of too many things to list -- this must be happening in many settings -- videos, social networks, products, etc.
@@
 # 8. Value: Balancing conversion vs spend
-The value of a production
+The value of an ordering
 
@@
 - the problem of action
+
+# Conclusion

### 2016-09-29 outline.rmd
I went shopping ...

@@
 # Introduction
+
 Presentation presents another presentation I saw recently about online grocery shopping at Tesco, a large supermarket chain based in the UK.
 The lists in this setting are recommendations for further purchases under the title of 'Have you forgotten?'.
+When you shop for groceries online, the 'Have you forgotten' recommendations appear when at the checkout stage.
+While the title 'have you forgotten' implies that you will be shown items that could have been or are usually on a shopping list, as we will see, the items on this list are recommendations.
+They are only a few of them: five.
+
+Shopping lists are quite prosaic -- a dozen lines scrawled on the back of a used envelope stuck on the refrigerator door ![shopping_list](figure/1_shopping.png ) -- and sublime -- a million products ranked in some kind of order for tens or hundreds of millions of people.
 I regard this list of just five suggestions given to Tesco's online customers as symptomatic of a major contemporary form of ordering, the recommendation list, that we see in many  forms of news, media, entertainment, marketing, advertising.
-Shopping lists are quite prosaic -- a dozen lines scrawled on the back of a used envelope stuck on the refrigerator door -- and sublime -- a million products ranked in some kind of order for tens or hundreds of millions of people.
+
+The shift from a handwritten list to 'Have you forgotten?' is perhaps a development in memory practices.
+It reminds me of the game 'I went shopping and I bought ...
+' that I play with children sometimes at the dinner table to settle them down.
+The game is about remembering a well-ordered list of items - a is apple, b is for banana, c is for carrot.
+The point of the game is to use the well-ordered set of labels provided by the sequence of letters in the alphabet to remember a random and perhaps very diverse list of items. 
+The longer the game continues, the harder it is to remember since the shopping list grows longer.
+At the same time, the many repetitions of this list, the fact that each player recites the whole list, helps memorisation.
+One trick in playing this game is to choose items that are quite unexpected.
+Rather than `ham`, list `halva`; rather than `fish`, add `face cream.` This is especially useful for adults in beating children.
+The children have better memories for items they know, but adults know more unexpected items.
+
+Something similar applies to recommendation systems. The system can remember very many things.
+It may know everything you bought in the last few years on that site, and it can easily see what you would normally buy and add those as suggestions.
+But a recommendation system finds it much harder to suggest items that you have not forgotten but may have never bought.
+
 If we are interested in lists from various angles -- asignifying semiotics as Maurizio Lazzarato calls it in his _Signs and Machines_ [@Lazzarato_2014] or 'a new order of spatio-temporal continuity for forms of economic, political and cultural life' as Celia Lury terms in her account of topological turn [@TBA], or even in terms of different cultural practices of listing [@Jullien_TBA] -- then contemporary lists practices such as recommendations figure as sites of intersection between different forms.
+
 Lists also raise the question of how we institute our approach to them.
@@
 - presentation at an academic event with speakers from industry and government on statistics and operations research in industry. Speakers came from Shell, Disney, GCHQ, and various universities. 
-- Shreena Patel works as a data scientist for DunnHumby focusing on online grocery shopping at Tesco.
+- Shreena Patel works as a data scientist for DunnHumby, a well-known [ customer science company ](https://www.dunnhumby.com/)  focusing on online grocery shopping at Tesco.
 Tesco is the largest supermarket chain in UK, famous for Clubcard starting in 1991; often cited as a leading early example of Big Data Analytics. 
@@
 [@Humby_1984]
-
+Clive Humby's academic work is hard to track down, but an early paper of his given at the Conference of Your Operational Researchers in Nottingham in 1984 suggests the direction that he and later Tesco would take in constructing lists [@Humby_1984]
 ![Humby, 1984]( figure/Humby_young0R_1994.png )
@@
 
-As Badiou might say,
-
 # 5. Data: 52 weeks of data for each customer: how is order constructed -- from history?

### 2016-09-29 outline.rmd
hansen

@@
 
-# Introduction
+# Introduction: the presentation of a presentation
 
@@
 But a recommendation system finds it much harder to suggest items that you have not forgotten but may have never bought.
+If you do buy recommended items, then the recommender system has been effective.
+It has led to a 'conversion.'
 
-If we are interested in lists from various angles -- asignifying semiotics as Maurizio Lazzarato calls it in his _Signs and Machines_ [@Lazzarato_2014] or 'a new order of spatio-temporal continuity for forms of economic, political and cultural life' as Celia Lury terms in her account of topological turn [@TBA], or even in terms of different cultural practices of listing [@Jullien_TBA] -- then contemporary lists practices such as recommendations figure as sites of intersection between different forms.
+If we are interested in lists from various angles -- as asignifying semiotics as Maurizio Lazzarato calls it in his _Signs and Machines_ [@Lazzarato_2014] or as elements in 'a new order of spatio-temporal continuity for forms of economic, political and cultural life' as Celia Lury terms in her account of topological turn [@TBA], or even in terms of different cultural practices of listing [@Jullien_TBA] -- then contemporary list practices such as recommendations figure as sites of intersection between different forms.
 
 Lists also raise the question of how we institute our approach to them.
-While lists such as recommendations intersect strongly with meaning and sense-making practices, they are also somewhere refractory to methods of humanities and social science research.
-For on the one hand, lists are highly associative. They can be interpreted or decoded semiotically, but they also exhibit great variations in textuality that limit semiotic readings.
+While lists such as recommendations intersect strongly with other meaning and sense-making practices in everyday life and popular culture, they are also somewhere refractory to methods of humanities and social science research.
+For on the one hand, lists are highly associative.
+They can be interpreted or decoded semiotically, but they also exhibit great variations in textuality that limit semiotic readings.
 For on the other hand, as operational entities, they can be treated ethnomethodologically, as the production of ('the order of social settings through ...
-shared sense making practices.')[https://en.wikipedia.org/wiki/Ethnomethodology]  The question for me here is whether contemporary lists can still be understood as produced through 'shared sense-making practices.'
-All lists are forms of order, since even random ordering is operationally important, but what counts as order might historically transformable, in which case the order of social settings or the problem of social order might be deeply historically contingent.
+shared sense making practices.')[https://en.wikipedia.org/wiki/Ethnomethodology]
+
+The question for me here is whether contemporary lists can still be understood as produced through 'shared sense-making practices.' All lists are forms of order, since even random ordering is operationally important, but what counts as order might historically transformable, in which case the order of social settings or the problem of social order might be deeply historically contingent.
 The problem of order lies at the core of lists making, and our relation to lists can be seen as a test of the way that order operates more generally.
+
 Much of my paper, and my responses to the presentation I saw from a data scientist  will focus on how the ordering that is done in a recommendation lists relates to the problem of social order more generally.
-- The main narrative of the presentation concerned a shift from a set-based model for making lists to a predictive, probabilistic way of making lists in almost realtime and dealing with lists on a scale that is often difficult to grasp concretely.
+The main narrative of the presentation concerned a shift from a set-based model for making lists to a predictive, probabilistic way of making lists in almost realtime and dealing with lists on a scale that is often difficult to grasp concretely.
+The change in list-making associated with learned models is increasingly widespread, and
 I respond to the presentation in two ways -- firstly by recounting, contextualising and commenting on some of the main steps as a example of the sense-making done by data scientists and customer analytics teams in many places; and secondly, by trying to step into the list-making practices  as an occasion to intervene and participate in the list-making practices they engage in.
+This is no mere methodological problem if, as Mark Hansen has recently suggested,
+
+> predictive analytics are discoveries of micrological propensities that are not directly correlated with human understanding and affectivity and that do not by themselves cohere into clearly identifiable events [@Hansen_2015b, 111-112]
+
+The question exercising much contemporary work on data, on media, platforms and finance pivots on the 'micrological propensities' that might be re-ordering lists.
+Hansen's argument here is that these propensities make probabilities into something real, some actually in things:
+
+>Whatever explanatory and causal value predictive analytics of large datasets have is, I suggest, ultimately rooted in this ontological transformation whereby probabilities are understood to be expressions of the actual propensity of things [@Hansen_2015b, 120]
+
+As the quote shows, he presents this as an 'ontological transformation.' The practical and indeed experimental question for me here is whether such transformations can be detected and articulated in the prosaic setting of shopping lists and  shopping baskets.
 

### 2016-09-29 outline.rmd
I went shopping ...

@@
 
-The shift from a handwritten list to 'Have you forgotten?' is perhaps a development in memory practices.
-It reminds me of the game 'I went shopping and I bought ...
+The shift from a handwritten list to 'Have you forgotten?' is perhaps a **development in memory practices.** It reminds me of the game 'I went shopping and I bought ...
 ' that I play with children sometimes at the dinner table to settle them down.
@@
 At the same time, the many repetitions of this list, the fact that each player recites the whole list, helps memorisation.
-One trick in playing this game is to choose items that are quite unexpected.
+One trick in playing this game is to choose items that are quite unexpected or increasingly random.
 Rather than `ham`, list `halva`; rather than `fish`, add `face cream.` This is especially useful for adults in beating children.
 The children have better memories for items they know, but adults know more unexpected items.
+On the other hand, very unexpected items are quite easy to remember and tend to even the score in the favour of the children.
+Balancing between the improbable and memorable ...
 
@@
 If you do buy recommended items, then the recommender system has been effective.
-It has led to a 'conversion.'
+It has led to a **'conversion.'** From a supermarket perspective, the conversion should be high-value.
 

### 2016-09-30 outline.rmd
adding stuff on tesco and on the combinatorial problems

@@
-# 'Have you forgotten?'
+# 'Have you forgotten?' Conversion experiences in list worlds
 
@@
 
+The question: which five items and in what order?
+
 Shopping lists are quite prosaic -- a dozen lines scrawled on the back of a used envelope stuck on the refrigerator door ![shopping_list](figure/1_shopping.png ) -- and sublime -- a million products ranked in some kind of order for tens or hundreds of millions of people.
@@
 ' that I play with children sometimes at the dinner table to settle them down.
-The game is about remembering a well-ordered list of items - a is apple, b is for banana, c is for carrot.
+The game is about remembering a well-ordered list of items - 'a' is apple, 'b' is for banana, 'c' is for carrot.
 The point of the game is to use the well-ordered set of labels provided by the sequence of letters in the alphabet to remember a random and perhaps very diverse list of items.
@@
 
-Something similar applies to recommendation systems. The system can remember very many things.
+Something similar applies to recommendation systems.
+The system can remember very many things.
 It may know everything you bought in the last few years on that site, and it can easily see what you would normally buy and add those as suggestions.
@@
 
-> predictive analytics are discoveries of micrological propensities that are not directly correlated with human understanding and affectivity and that do not by themselves cohere into clearly identifiable events [@Hansen_2015b, 111-112]
+> predictive analytics are discoveries of micrological propensities that are not directly correlated with human understanding and affectivity and that do not by themselves cohere into clearly identifiable events [@Hansen_2015a, 111-112]
 
@@
 
->Whatever explanatory and causal value predictive analytics of large datasets have is, I suggest, ultimately rooted in this ontological transformation whereby probabilities are understood to be expressions of the actual propensity of things [@Hansen_2015b, 120]
+>Whatever explanatory and causal value predictive analytics of large datasets have is, I suggest, ultimately rooted in this ontological transformation whereby probabilities are understood to be expressions of the actual propensity of things [@Hansen_2015a, 120]
 
@@
 # 1. The context of the presentation - Consumption and shopping
-- presentation at an academic event with speakers from industry and government on statistics and operations research in industry. Speakers came from Shell, Disney, GCHQ, and various universities. 
+- presentation at an academic event with speakers from industry and government on statistics and operations research in industry.
+Speakers came from Shell, Disney, GCHQ, and various universities.
+Because I've done a lot of work on machine learning and data science more generally, I found the sessions both familiar and occasionally exciting.
 - Shreena Patel works as a data scientist for DunnHumby, a well-known [ customer science company ](https://www.dunnhumby.com/)  focusing on online grocery shopping at Tesco.
+I found her presentation on shopping lists to be very practical since it was focused quite a lot on implementation of recommendations, and symptomatic in various ways.
 Tesco is the largest supermarket chain in UK, famous for Clubcard starting in 1991; often cited as a leading early example of Big Data Analytics. 
@@
 Clive Humby's academic work is hard to track down, but an early paper of his given at the Conference of Your Operational Researchers in Nottingham in 1984 suggests the direction that he and later Tesco would take in constructing lists [@Humby_1984]
-![Humby, 1984]( figure/Humby_young0R_1994.png )
+![Humby, 1984]( figure/2_humby_young_org_1994.png)
 - the problems of scale: Tesco is like Amazon or Netflix in this respect: too many people and too many products
@@
 Shreena described her work at DunnHumby in terms of a shift in models from a 'rules-based list' to a 'relevancy model.' 
-This was a significant change in how the 'Have You Forgotten' list was constructed. 
+This was a significant change in how the 'Have You Forgotten' list was constructed from one kind of model to another.
+For me, the shift in models matters since it amounts to a much more probabilistic structuring of lists.
 I assume that the rules-based model derived from the early 1990s work on the `apriori` algorithm by Rakesh Agrawal and Ramakrishnan Srikanti at IBM Research Alameda [@Agrawal_1994].
@@
 If we take the itemset from the `Groceries` dataset [@Hahsler_2006], and then run it through the Tesco online groceries site, each of the 169 items in the dataset yields dozens and sometimes thousands of results from Tesco online.
+
+```{r tesco, cache=TRUE, }
+    library(ggplot2)
+    tesco = read.csv('data/tesco_groceries.csv')
+    total = sum(tesco$actual, na.rm=TRUE) 
+    tesco = na.omit(tesco)
+    tesco$actual = as.integer(tesco$actual)
+
+    ggplot(tesco[tesco$actual>50,], aes(x=labels, y=actual, fill=labels)) +geom_bar(stat='identity') + coord_flip() + ggtitle('Tesco grocery items with more than 50 products')
+```
+
 Each of the items in the `Groceries` dataset proliferates in the Tesco site, into a new list of branded variations.
+The 169 items of the `Groceries` dataset, itself from a local German or Austrian supermarket [@REF_TBA],  expand into roughly `r total` possible items.
 
@@
 
-Patel observed that the 'Have you forgotten' lists matter to people.
-Some people log on and go straight to the checkout, allowing the recommendation system to supply them with the suggestions that they then move to the shopping basket.
+Shreena observed that the 'Have you forgotten' lists matter to people, and we can see why in the light of this proliferation of things in the list.
+A `r total` list of items is quite hard to list let alone read.
+Shreena said that some people log on and go straight to the checkout, allowing the recommendation system to supply them with the suggestions that they then move to the shopping basket.
 This is the online equivalent of using an old shopping list and only adding new items when needed. 
 It is also a strategy to avoid having to view too many lists, since the only way Tesco can show people what they might buy is by showing them lists of things (e.g.
-chicken products), but these lists are too long to work with.
+chicken products), but these lists are often too long to work with.
+For instance, Tesco has more than 1000 chicken products on its site.
 People inhabit lists and work with them no matter how they are made; thus vindicating ethnomethodologists assertion that social order is the product of ongoing sense-making practices. 
+But it also suggests a transformation in the way that lists exist in the world.
 
 # 5. Data: 52 weeks of data for each customer: how is order constructed -- from history?
+
+Shreena described the data they on grocery purchases. 'We have 52 weeks of purchase data for each customer.' 
+We know that the construction of lists has been transformed by recommendation systems.
+These transformations have been widely discussed in the context of big data, where questions of who has how much data are usually most prominent.
+In terms of the ways that lists order the world, the amount of data is not so amount as the way in which things in the world take on a list-like existence. 
+I want to dwell on this idea for a while. We have seen two different ways in which lists might be constructed in a recommender system. 
+The first uses history and existence of previous lists, including the customer's own list, to populate a new list. 
+The second uses many shopping lists to identify associations between things and then uses the strongest associations to recommend additions to a current list. 
+The 'relevancy model,' however, tries to do something different. 
+The problem that the proliferation of things on the shelves of supermarket or grocery warehouse confronts is partly combinatorial.
+With  `r total` products (actually Shreena mentions 200,000 products) can be combined in truly astronomical numbers of combinations.
+If a typical shopping list has 20 items, then there are `r choose(total, 20)` possible combinations.
+Some of this vast number of possible shopping lists is highly unlikely in terms of shopping lists and histories of shopping, but potentially highly interesting and value as recommendation.   
+In some ways, this resembles the Badiouean ontology of the multiple a HERE
+
+> This means, to put it slightly differently, that surrounding any delimited predictive system is a larger field of data—­what I elsewhere call a “surplus of sensibility”—­that, viewed speculatively, indexes the causal efficacy of the total situation within which this delimited system operates.
+(Effectively, the latter gains its reliability from closing off this larger surplus of sensibility, thereby transforming an always excessive propensity into a [provisionally] closed dataset.) Because it affords data that exceed whatever any given predictive system might include, the data of the world’s causal efficacy—­the data constituting its real potentiality—­always and in principle facilitates knowledge that cannot be restricted to any particular agenda [@Hansen_2015, 126-7]
+
 - history of data accumulations;
+
 - shift from data mining to personalization? 
+- use some of the Varian, Linden, Resnick, Schaffer literature
 
@@
 - what counts as a good feature? See Domingos on this
+- move from Dunnhumby's ClubCard focus to much more general world of predictive models (e.g. Facebook use a million models, etc.)
 
@@
 # Conclusion
+- lists have complex temporal structure
+- listing practices very closely entangled with problems of social order
+- lists undergoing a 'conversion' experience as they become more probabilistic

### 2016-09-30 scrape_tesco.py
forgot to add this script

@@ -1,0 +1,25 @@
+import BeautifulSoup as bs
+import requests
+import pandas as pd
+
+groceries = pd.read_csv('data/grocery_items.csv')
+base = 'http://www.tesco.com/groceries/product/search/default.aspx?notepad=ITEM&newSort=true'
+# page = requests.get(base)
+# items = bs.BeautifulSoup(str(page.content))
+# results = items.find('div', attrs={'class':'headerContent'}).getText()
+urls = [base.replace('ITEM', i) for i in groceries['labels']]
+groceries['url'] = urls
+
+def getItems(url):
+    page = requests.get(url)
+    items = bs.BeautifulSoup(str(page.content))
+    results = items.find('div', attrs={'class':'headerContent'}).getText()
+    print results
+    return results
+
+groceries['product_count'] = [getItems(u) for u in urls]
+r = groceries.Product_count.str.extract('(\d+)')
+r[pd.isnull(r) != True].astype('int').sum()
+groceries['actual_count'] = r
+groceries.to_csv('data/tesco_groceries.csv')
+

### 2016-10-05 outline.rmd
trying to get relevance and basket similarity into the argument

@@
 When you shop for groceries online, the 'Have you forgotten' recommendations appear when at the checkout stage.
-While the title 'have you forgotten' implies that you will be shown items that could have been or are usually on a shopping list, as we will see, the items on this list are recommendations.
-They are only a few of them: five.
+While the title 'have you forgotten' implies that you will be shown items that could have been or are usually on a shopping list.
+The working assumption is a temporal one -- the list is incomplete either because you have forgotten to list something, or you didn't know what should be on your list, perhaps because you have never even thought of putting that thing on your list.
 
-The question: which five items and in what order?
+![Tesco ClubCard](figure/tesco-clubcard.jpg )
 
-Shopping lists are quite prosaic -- a dozen lines scrawled on the back of a used envelope stuck on the refrigerator door ![shopping_list](figure/1_shopping.png ) -- and sublime -- a million products ranked in some kind of order for tens or hundreds of millions of people.
+As we will see, the items on this list are recommendations. They are only a few of them: five. The question: which five items and in what order?
+
+![shopping_list](figure/1_shopping.png )
+
+Shopping lists are quite prosaic -- a dozen lines scrawled on the back of a used envelope stuck on the refrigerator door  -- and sublime -- a million products ranked in some kind of order for tens or hundreds of millions of people.
 I regard this list of just five suggestions given to Tesco's online customers as symptomatic of a major contemporary form of ordering, the recommendation list, that we see in many  forms of news, media, entertainment, marketing, advertising.
 
-The shift from a handwritten list to 'Have you forgotten?' is perhaps a **development in memory practices.** It reminds me of the game 'I went shopping and I bought ...
-' that I play with children sometimes at the dinner table to settle them down.
+The problem of incomplete lists reminds me of the game 'I went shopping and I bought ...
+' that adults sometimes play with children.
 The game is about remembering a well-ordered list of items - 'a' is apple, 'b' is for banana, 'c' is for carrot.
@@
 The longer the game continues, the harder it is to remember since the shopping list grows longer.
-At the same time, the many repetitions of this list, the fact that each player recites the whole list, helps memorisation.
+At the same time, each player repeats the list, and this  helps memorisation.
 One trick in playing this game is to choose items that are quite unexpected or increasingly random.
@@
 On the other hand, very unexpected items are quite easy to remember and tend to even the score in the favour of the children.
-Balancing between the improbable and memorable ...
+Balancing between the improbable and memorable, between a predictable list like 'apples, bananas, carrots' and an unusual list such as 'agave syrup, boccadilla,cilantro' is complicated.
 
@@
 It has led to a **'conversion.'** From a supermarket perspective, the conversion should be high-value.
+There is little value in recommending a whole lot of small and cheap items such as rubber bands or toothpicks.
+So it might need to choose amongst different recommendations and weight high value items differently to low value items.
+In either case, 'conversion'  is the valued event.
 
@@
 
-Lists also raise the question of how we institute our approach to them.
-While lists such as recommendations intersect strongly with other meaning and sense-making practices in everyday life and popular culture, they are also somewhere refractory to methods of humanities and social science research.
+**How do we approach lists as objects of research?** While lists such as recommendations intersect strongly with other meaning and sense-making practices in everyday life and popular culture, they are also somewhere refractory to discourse analysis and textual analysis methods of humanities and social science research.
 For on the one hand, lists are highly associative.
-They can be interpreted or decoded semiotically, but they also exhibit great variations in textuality that limit semiotic readings.
+They can be interpreted or decoded semiotically, but they also exhibit great variations in textuality that multiply semiotic readings.
 For on the other hand, as operational entities, they can be treated ethnomethodologically, as the production of ('the order of social settings through ...
@@
 The change in list-making associated with learned models is increasingly widespread, and
-I respond to the presentation in two ways -- firstly by recounting, contextualising and commenting on some of the main steps as a example of the sense-making done by data scientists and customer analytics teams in many places; and secondly, by trying to step into the list-making practices  as an occasion to intervene and participate in the list-making practices they engage in.
+I respond to the presentation in two ways -- firstly by recounting, contextualising and commenting on some of the main steps in making the 'Have your forgotten?' list  as a example of the sense-making done by data scientists and customer analytics teams in many places; and secondly, by trying to step into the list-making practices  as an occasion to intervene and participate in the ontological conversion practices they engage in.
 This is no mere methodological problem if, as Mark Hansen has recently suggested,
@@
 The question exercising much contemporary work on data, on media, platforms and finance pivots on the 'micrological propensities' that might be re-ordering lists.
-Hansen's argument here is that these propensities make probabilities into something real, some actually in things:
+Hansen's argues that these propensities make probabilities into something real, some actually in things:
 
@@
 As the quote shows, he presents this as an 'ontological transformation.' The practical and indeed experimental question for me here is whether such transformations can be detected and articulated in the prosaic setting of shopping lists and  shopping baskets.
+I will suggest that 'conversion' might be a better way of mapping these changes than 'transformation.'
 
@@
 Clive Humby's academic work is hard to track down, but an early paper of his given at the Conference of Your Operational Researchers in Nottingham in 1984 suggests the direction that he and later Tesco would take in constructing lists [@Humby_1984]
+
 ![Humby, 1984]( figure/2_humby_young_org_1994.png)
+
 - the problems of scale: Tesco is like Amazon or Netflix in this respect: too many people and too many products
@@
 inspect(rules[1:5])  
-plot(rules,method="graph",interactive=TRUE,shading=NA)
+#plot(rules,method="graph",interactive=TRUE,shading=NA)
 items = Groceries@itemInfo
-head(items)
-nrow(items)
+#head(items)
+#nrow(items)
 ```
@@
 This means that that the algorithm can be understood in terms of the set-based multiples proposed by Alain Badiou.
+
+> Multiplicity will be called  ‘closed’ when grasped by the numerical one, and ‘open’ when grasped by the vital One.
+Every multiplicity is the joint effectuation of the closed and open, but its ‘veritable’ multiple-being lies on the side of the open.
+[@Badiou_2004, 72]
+
 Badiou in his account of multiples highlight the contrast between two different ways of grasping a multiple -- in terms of the 'numerical one' and the 'vital One' [@Badiou_2004, 72], between the 'closed' and the 'open' respectively.
@@
 
-    ggplot(tesco[tesco$actual>50,], aes(x=labels, y=actual, fill=labels)) +geom_bar(stat='identity') + coord_flip() + ggtitle('Tesco grocery items with more than 50 products')
+    ggplot(tesco[tesco$actual>50,], aes(x=labels, y=actual, fill=labels)) +geom_bar(stat='identity') + coord_flip() + ggtitle('Tesco grocery items with more than 50 products') + theme(legend.position="none")
 ```
@@
 With  `r total` products (actually Shreena mentions 200,000 products) can be combined in truly astronomical numbers of combinations.
-If a typical shopping list has 20 items, then there are `r choose(total, 20)` possible combinations.
+If a typical shopping list has 20 items, then there are `r format(choose(total, 20), 2)` possible combinations.
 Some of this vast number of possible shopping lists is highly unlikely in terms of shopping lists and histories of shopping, but potentially highly interesting and value as recommendation.
-In some ways, this resembles the Badiouean ontology of the multiple a HERE
 
-> This means, to put it slightly differently, that surrounding any delimited predictive system is a larger field of data—­what I elsewhere call a “surplus of sensibility”—­that, viewed speculatively, indexes the causal efficacy of the total situation within which this delimited system operates.
-(Effectively, the latter gains its reliability from closing off this larger surplus of sensibility, thereby transforming an always excessive propensity into a [provisionally] closed dataset.) Because it affords data that exceed whatever any given predictive system might include, the data of the world’s causal efficacy—­the data constituting its real potentiality—­always and in principle facilitates knowledge that cannot be restricted to any particular agenda [@Hansen_2015, 126-7]
+> This means, to put it slightly differently, that surrounding any delimited predictive system is a larger field of data -- what I elsewhere call a “surplus of sensibility” -- that, viewed speculatively, indexes the causal efficacy of the total situation within which this delimited system operates.
+(Effectively, the latter gains its reliability from closing off this larger surplus of sensibility, thereby transforming an always excessive propensity into a [provisionally] closed dataset.) Because it affords data that exceed whatever any given predictive system might include, the data of the world’s causal efficacy -- the data constituting its real potentiality -- always and in principle facilitates knowledge that cannot be restricted to any particular agenda [@Hansen_2015, 126-7]
+
+In some ways, Hansen's account resembles the Badiouean ontology of the multiple in which a closed set is surrounded and conditioned by an open set that is full of many more potentials.
+However, if we think of the  new model that Shreena describes as the delimited (or delimiting?) system within   the larger surplus of sensibility, how that help us understand what the 'Have you forgotten?' recommendations do?
 
@@
 - move from Dunnhumby's ClubCard focus to much more general world of predictive models (e.g. Facebook use a million models, etc.)
+- the model counts for surprisingly little -- at least, they all seem similar ...
+rather what counts is all that stuff on working with substitutions, similiarities and complementarities for each person.
+
+It was surprising that Shreena spoke relatively briefly about the new models.
+She showed a graph that compared the old rule-based model with some of the new alternatives -- logistic regression, random forests, and a few others.
+She dismissed most of the models quite quickly and focused only on one, the logistic regression model.
+Or more accurately, she just said that logistic regression did better than other, and there was little more to discusson on that front.
+
+Why did she say little about the model, which after all, produces the probabilistic predictions on which everything in the recommendation list depends?
+
+Her interest lay in the relevance of the recommendations, and in particular, in how 'basket similarity affects relevance of recommendations.' 
+HERE
 
@@
 - interesting contrasts here between the relations of things in a list
+- whereas the `arule` approach compares lists looking for similiarities and ways of completing lists, the more recent approaches look for micrological propensities in the list that relate to a particular time and place, amidst a continuously differentiable pathways between things ...
 

### 2016-10-05 scrape_tesco.py

@@ -17,9 +17,10 @@ def getItems(url):
     print results
     return results
 
+
 groceries['product_count'] = [getItems(u) for u in urls]
 r = groceries.Product_count.str.extract('(\d+)')
-r[pd.isnull(r) != True].astype('int').sum()
-groceries['actual_count'] = r
+total = r[pd.isnull(r) != True].astype('int').sum()
+groceries['actual_count'] = r.astype('int')
 groceries.to_csv('data/tesco_groceries.csv')
 

### 2016-10-06 outline.rmd
adding api stuff

@@
 Or more accurately, she just said that logistic regression did better than other, and there was little more to discusson on that front.
+The measure of doing better here is called precision, and it measures the proportion of the recommended products that are actually purchased.
 
-Why did she say little about the model, which after all, produces the probabilistic predictions on which everything in the recommendation list depends?
+Why did she say little about the model, which after all, produces the probabilistic predictions on which everything in the recommendation list depends and hence the conversion experience?
 
-Her interest lay in the relevance of the recommendations, and in particular, in how 'basket similarity affects relevance of recommendations.' 
-HERE
+![Models](figure/patel_recommendation1.png)
+
+The graph compares models in terms of their predictions and how those predictions came to pass.
+The two columns on the left show earlier models -- the global model, which is not personalized, the association rules model for individual customers, and then the new models (logistic regression, random forest, etc) for customers.
+An increase in precision means more sales.
+Her interest lay in the relevance of the recommendations, and in particular, in how 'most of the gain comes from thinking of good features.'
 
 # 7. Sparsity: output 20 recommendations from 200,000 products
-The problem of too many things to list -- this must be happening in many settings -- videos, social networks, products, etc.
+
+What counts as a good feature in a world of predictive lists?
+Some things in the data that increases the precision of the model.
+The problem that Tesco faces, however, is that there are too many possible recommendations to be tested.
+They sell around 200,000 products and no customer wants to buy any more than a fraction of a per cent of those products.
+The recommendation list needs to very short.
+
+Predictive models, we might say, encounter their own surplus of sensibility in the data as they search for the most accurate predictions.
+A  very brief comment from Shreena -- 'we have lots of zeros' -- points to the problem here.
+Each of the millions of Tesco online grocery shoppers has bought some selection of the 200,000 products.
+But the product-customer matrix, the basic vector-space in which all recommendation systems operate, remain very sparse and unpopulated.
+
+Say Tesco has 1 million online customers.
+A data matrix for just whether they ever bought a particular product or not will be `r format(1e6 * 2e5)` in size.
+Given that any one customer is like to only have bought 100 different products, the data matrix will be `r format(100-(100/2e5)*100, 2)`% empty.
+
+```{r sparsity, cache=TRUE}
+
+library(ggplot2)
+customers = seq(1000000,1)
+products = seq(200000,1)
+format(1000000 * 200000, 3)
+
+m = matrix(p, nrow=900, byrow=TRUE)
+df = as.dataframe(m)
+#TODO: plot here
+
+```
+
+Having too many things to list -- this must be happening in many settings -- videos, social networks, products, etc.
+-- is not the real problem.
+The problem is that 'massively unbalanced data,' which means that different classes of products are not equally represented.
+Since so many people buy milk, the recommendation system might end up always recommending milk products.
+So the data needs to be 'corrected' by, in this case, random under-sampling or removal of some of the data for common purchases.
+Secondly, even if a recommendation system finds uncommon items that are good recommendations, they should neither be too cheap ('low spend'), nor, as we will see too similar to what the customer has already ordered.
+
+So the raw data -- the 52 weeks of purchases for each customer -- can be aggregated in various ways to increase the density of the data in ways that helps the model.
+Shreena mentions some of the 'features' used in the new model --
+HERE
+In Hansen's or Badiou's terms, this is one way in which the dataset is closed.
+While 200,000 products is still a closed dataset since it has a defined number of items, undersampling alters the frequency of occurrence of very common items so that the predictive models are not distracted by them.
 

### 2016-10-07 simulate_baskets.r
adding a lot of data

@@ -1,0 +1,25 @@
+library(dplyr)
+library(ggplot2)
+library(readr)
+
+dir('data/')
+csvs = list.files('data', pattern = '*.csv',full.names=TRUE)
+df = bind_rows(sapply(csvs, read_tsv))
+
+str(df)
+colnames(df)
+ggplot(df, aes(x=price,y=ContentsQuantity)) + geom_point() 
+ggplot(df, aes(x=price)) + geom_histogram(bins=200)
+fivenum(df$price)
+df[df$price>=20,'name']
+
+#sample some baskets
+
+sample_n(df[order(df$tpnb),  'name'], 100) 
+
+# randomise quantity
+
+# work out total cost
+
+# work out total weight
+df %>% select(name, price) %>% filter(

### 2016-10-07 tesco_fetch_product_images.sh

@@ -1,0 +1,6 @@
+#!/bin/bash
+
+cut -f11 $1|xargs wget -P $2 
+cd $2
+montage  -mode concatenate -tile 40x IDShot_90x90.jpg.* out.jpg
+eog out.jpg

### 2016-10-07 tesco_get_groceries.py

@@ -1,0 +1,68 @@
+#! /usr/bin/env python
+
+import httplib, urllib
+import sys
+import pandas as pd
+import json
+
+# read API key
+keyfile = open('tesco_api_key.txt')
+key = keyfile.readline().strip()
+
+def get_groceries(q,off,lmt):
+    # get first 100 and total number of results
+
+    df_all = pd.DataFrame()
+    total = 0
+    params = urllib.urlencode({ })
+    headers = {
+        # Request headers
+        'Ocp-Apim-Subscription-Key':  key,
+    }
+    try:
+        conn = httplib.HTTPSConnection('dev.tescolabs.com')
+        conn.request("GET",
+                        "/grocery/products/?query=" + q + "&offset=" +
+                        str(off) + "&limit=" + lmt + "&%s" % params, "{body}",
+                        headers)
+        response = conn.getresponse()
+        data = response.read()
+        j = json.loads(data)
+        df = pd.DataFrame(j['uk']['ghs']['products']['results'])
+        df_all = df_all.append(df)
+        total = j['uk']['ghs']['products']['totals']['all']
+        print "There are {}  {} products".format(str(total), q)
+        conn.close()
+    except Exception as e:
+        print(e)
+
+    remain = total
+
+    while remain > 0:
+        try:
+            off += 100
+            conn = httplib.HTTPSConnection('dev.tescolabs.com')
+            conn.request("GET",
+                        "/grocery/products/?query=" + q + "&offset=" +
+                        str(off) + "&limit=" + lmt + "&%s" % params, "{body}",
+                        headers)
+            response = conn.getresponse()
+            data = response.read()
+            j = json.loads(data)
+            df = pd.DataFrame(j['uk']['ghs']['products']['results'])
+            df_all = df_all.append(df)
+            print "{} products remaining to fetch".format(remain)
+            remain -= 100
+            conn.close()
+        except Exception as e:
+            print(e)
+
+    return df_all
+
+
+args = sys.argv
+q = args[1]
+off = 100
+lmt = '100'
+df_all = get_groceries(q,off,lmt)
+df_all.to_csv('data/' + q + '.csv', encoding='utf-8', sep='\t')

### 2016-10-07 lists_conversion_experience.rmd
renaming

--- 2016-10-06 outline.rmd
@@
 
+In this respect, the presentation was not really surprising or exceptional.
+Its ordinariness and prosaic character stands out as the most significant feature.
+The presentation itself is a working example for an argument about our experience of predictive ordering.
+
 The problem of incomplete lists reminds me of the game 'I went shopping and I bought ...
@@
 It may know everything you bought in the last few years on that site, and it can easily see what you would normally buy and add those as suggestions.
+
 But a recommendation system finds it much harder to suggest items that you have not forgotten but may have never bought.
@@
 
-The question for me here is whether contemporary lists can still be understood as produced through 'shared sense-making practices.' All lists are forms of order, since even random ordering is operationally important, but what counts as order might historically transformable, in which case the order of social settings or the problem of social order might be deeply historically contingent.
+The question for me here is **how** contemporary lists can still be understood as produced through 'shared sense-making practices.' All lists are forms of order, since even random ordering is operationally important, but what counts as order might historically transformable, in which case the order of social settings or the problem of social order might be deeply historically contingent.
 The problem of order lies at the core of lists making, and our relation to lists can be seen as a test of the way that order operates more generally.
+
+In this context, the fact that the lists and indeed supermarkets more generally are driven by database logics suggests that working with list data could well entail interacting with databases.
+
+To this end, I will be seeking to simulate the workings of the recommendation system by generating queries, and exploring results from  a very large supermarket chain database.
+
+```{r simulate_basket, echo=FALSE, cache=TRUE}
+library(dplyr)
+library(ggplot2)
+library(readr)
+
+dir('data/')
+csvs = list.files('data', pattern = '*.csv',full.names=TRUE)
+df = bind_rows(sapply(csvs, read_tsv))
+
+str(df)
+colnames(df)
+ggplot(df, aes(x=price,y=ContentsQuantity)) + geom_point() 
+ggplot(df, aes(x=price)) + geom_histogram(bins=200)
+fivenum(df$price)
+df[df$price>=20,'name']
+```
 
@@
 The change in list-making associated with learned models is increasingly widespread, and
+
 I respond to the presentation in two ways -- firstly by recounting, contextualising and commenting on some of the main steps in making the 'Have your forgotten?' list  as a example of the sense-making done by data scientists and customer analytics teams in many places; and secondly, by trying to step into the list-making practices  as an occasion to intervene and participate in the ontological conversion practices they engage in.
@@
 Because I've done a lot of work on machine learning and data science more generally, I found the sessions both familiar and occasionally exciting.
+
+![patel_notes](figure/patel_overall.png)
+
 - Shreena Patel works as a data scientist for DunnHumby, a well-known [ customer science company ](https://www.dunnhumby.com/)  focusing on online grocery shopping at Tesco.
@@
 inspect(rules[1:5])  
-#plot(rules,method="graph",interactive=TRUE,shading=NA)
 items = Groceries@itemInfo
-#head(items)
-#nrow(items)
 ```
@@
 Association is stronger when it derives from an excess of different ways of being in sets.
+
 Badiou's argument is that being part of a set differs from belonging to a set.
-An open set arises from when we dwell within a set by 'taking the intersection of two elements.' 
-What might this mean in terms of ordering?
+An open set arises from when we dwell within a set by 'taking the intersection of two elements.' What might this mean in terms of ordering?
+
 If we take the itemset from the `Groceries` dataset [@Hahsler_2006], and then run it through the Tesco online groceries site, each of the 169 items in the dataset yields dozens and sometimes thousands of results from Tesco online.
@@
     library(ggplot2)
-    tesco = read.csv('data/tesco_groceries.csv')
+    tesco = read.csv('data/reference_data/tesco_groceries.csv')
     total = sum(tesco$actual, na.rm=TRUE) 
@@
 format(1000000 * 200000, 3)
-
+p = rbinom(2e5, n=1,prob=0.5)
 m = matrix(p, nrow=900, byrow=TRUE)
-df = as.dataframe(m)
-#TODO: plot here
+df = as.data.frame(m)
+#TODO: plot here, could possibly use the images from tescoe
 
@@
 
-So the raw data -- the 52 weeks of purchases for each customer -- can be aggregated in various ways to increase the density of the data in ways that helps the model.
-Shreena mentions some of the 'features' used in the new model --
-HERE
 In Hansen's or Badiou's terms, this is one way in which the dataset is closed.
 While 200,000 products is still a closed dataset since it has a defined number of items, undersampling alters the frequency of occurrence of very common items so that the predictive models are not distracted by them.
+
+So the raw data -- the 52 weeks of purchases for each customer -- can be aggregated to increase the density of the data in ways that helps the model.
+Shreena mentions some of the 'features' used in the new model but somewhat strangely at this point, her presentation turns toward the difficulties of measuring the success of the model and the infrastructural limitations on making recommendations in ways that are more alive to what is happening right now.
 
@@
 
+The discussion of imperfection was brief - it incorporated the notion of recommendations weighted by price and the notion of precision (or the proportion of products actually purchased.).
+Shreena simply said that this metric was imperfect, but without elaborating why.
+Instead she moved to describe some of the ways in which the recommendation system worked in practice.
+
 # 10. Experiment: A/B allocation
-Could use the Google Infrastructure of experiment or Facebook Flow stuff here -- 1 million models
+
+![A/B testing](figure/a_b_tests.png)
+
+First of all, the new model was deployed experimentally in a random A/B controlled trial.
+This means that all customers are allocated to one of four categories as shown in the figure.
+For an identical shopping list, or even for the same person, it is possible to receive different recommendations.
+
+This randomising of the application of the model, drawing on protocols for randomised clinical trials first developed in the 1960s, is widely used in social media platforms and hence in all recommender systems.
+It is difficult to gauge how much.
+Facebook reports that it has more than 1 million models operation in its infrastructure.
+[TBA: ref for FBflow]   We could also discuss the integration of experiments into the operational infrastructures at Google.[REF tba]
+
+Random allocation of customers to the four categories (Test A, Test B, Control A, Control B) adds another layer of probability to the recommendation system.
+This is done in the name of statistical validation of the effects or the uplift of the model.
 

### 2016-10-07 simulate_basket_test.r
simulate script

@@ -1,0 +1,90 @@
+    library(ggplot2)
+    tesco = read.csv('data/reference_data/tesco_groceries.csv')
+    total = sum(tesco$actual, na.rm=TRUE) 
+    tesco = read.csv('data/reference_data/tesco_groceries.csv')
+dir()
+setwd('~/lists')
+    tesco = read.csv('data/reference_data/tesco_groceries.csv')
+    total = sum(tesco$actual, na.rm=TRUE) 
+    tesco = na.omit(tesco)
+    tesco$actual = as.integer(tesco$actual)
+    ggplot(tesco[tesco$actual>50,], aes(x=labels, y=actual, fill=labels)) +geom_bar(stat='identity') + coord_flip() + ggtitle('Tesco grocery items with more than 50 products') + theme(legend.position="none")
+```
+head(df)
+library(dplyr)
+library(ggplot2)
+library(readr)
+dir('data/')
+csvs = list.files('data', pattern = '*.csv',full.names=TRUE)
+df = bind_rows(sapply(csvs, read_tsv))
+str(df)
+colnames(df)
+ggplot(df, aes(x=price,y=ContentsQuantity)) + geom_point() 
+ggplot(df, aes(x=price)) + geom_histogram(bins=200)
+fivenum(df$price)
+df[df$price>=20,'name']
+df[1:3, 4:10]
+df[1:3, 4:15]
+df[1:3, 10:15]
+df[1:3, 12:15]
+df[1:3, c('image', 'tpnb')]
+df$tpnb
+sort(df$tpnb)
+unique(sort(df$tpnb))
+head(unique(sort(df$tpnb)))
+diff(unique(sort(df$tpnb)))
+hist(diff(unique(sort(df$tpnb))))
+unique(sort(df$tpnb))
+unique(order(df$tpnb))
+df$name[(order(df$tpnb))]
+head(df$name[(order(df$tpnb))])
+unique(head(df$name[(order(df$tpnb))]))
+head(unique(df$name[(order(df$tpnb))]))
+head(unique(df$name[(order(df$tpnb))]),30)
+head(unique(df[order(df$tpnb), c('tpnb', 'name')]),30)
+head(unique(df[order(df$tpnb), c('tpnb', 'name')]),30)
+head(unique(df[order(df$tpnb), c('tpnb', 'name')]),50)
+df[order(df$tpnb), c('tpnb', 'name')]
[…]

### 2016-10-10 lists_conversion_experience.rmd
got to infrastructure

@@
 
-# Introduction: the presentation of a presentation
+## Introduction: the presentation of a presentation
 
@@
 The working assumption is a temporal one -- the list is incomplete either because you have forgotten to list something, or you didn't know what should be on your list, perhaps because you have never even thought of putting that thing on your list.
-
-![Tesco ClubCard](figure/tesco-clubcard.jpg )
 
@@
 The presentation itself is a working example for an argument about our experience of predictive ordering.
+In this paper, I re-present the presentation using a mixture of theoretical and practical resources.
+The theory is largely recent media theory, and the practical resources are that of coding and predictive models.
 
@@
 
-If we are interested in lists from various angles -- as asignifying semiotics as Maurizio Lazzarato calls it in his _Signs and Machines_ [@Lazzarato_2014] or as elements in 'a new order of spatio-temporal continuity for forms of economic, political and cultural life' as Celia Lury terms in her account of topological turn [@TBA], or even in terms of different cultural practices of listing [@Jullien_TBA] -- then contemporary list practices such as recommendations figure as sites of intersection between different forms.
+If we are interested in lists from various angles -- as asignifying semiotics as Maurizio Lazzarato calls it in his _Signs and Machines_ [@Lazzarato_2014] or as elements in 'a new order of spatio-temporal continuity for forms of economic, political and cultural life' as Celia Lury terms in her account of topological turn [@TBA], or in terms of the predictive conditions of contemporary experience [@Hansen_2015a] -- then contemporary list practices such as recommendations figure as sites of intersection between different forms.
 
@@
 
-# 1. The context of the presentation - Consumption and shopping
-- presentation at an academic event with speakers from industry and government on statistics and operations research in industry.
+## 1. The context of the presentation - Consumption and shopping
+
+I heard this presentation at an academic event with speakers from industry and government on statistics and operations research in industry.
 Speakers came from Shell, Disney, GCHQ, and various universities.
@@
 
-- Shreena Patel works as a data scientist for DunnHumby, a well-known [ customer science company ](https://www.dunnhumby.com/)  focusing on online grocery shopping at Tesco.
-I found her presentation on shopping lists to be very practical since it was focused quite a lot on implementation of recommendations, and symptomatic in various ways.
+Shreena Patel works as a data scientist for DunnHumby, a well-known [ customer science company ](https://www.dunnhumby.com/)  focusing on online grocery shopping at Tesco.
+I found her presentation on shopping lists to be  practical and evocative since it was focused quite a lot on implementation of recommendations, and symptomatic in various ways.
+
+![Tesco ClubCard](figure/tesco-clubcard.jpg )
+
 Tesco is the largest supermarket chain in UK, famous for Clubcard starting in 1991; often cited as a leading early example of Big Data Analytics. 
@@
 
+While Tesco has been very successful in data-mining its customers, the online shopping situation is a bit different.
+It is no longer organised around campaigns involving special offers or redemption of points, but on continuous up-selling and cross-selling.
+
 - the problems of scale: Tesco is like Amazon or Netflix in this respect: too many people and too many products
@@
 
-# 2. The web and the shopping basket: how it was reconstituted in the 1990s
+## 2. The web and the shopping basket: how it was reconstituted in the 1990s
+
+![People add to list](figure/online_lists_in_practice.png)
+
 - online commerce -- transforming shopping; 
@@
 
-# 3. From the rules-based list to a 'relevancy model'
+## 3. From the rules-based list to a 'relevancy model'
 
@@
 
-# 4. Hack? Customers go straight to the checkout
+## 4. Hack? Customers go straight to the checkout
 
@@
 
-# 5. Data: 52 weeks of data for each customer: how is order constructed -- from history?
+## 5. Data: 52 weeks of data for each customer: how is order constructed -- from history?
 
@@
 
-# 6. Gain: use good features to build models 
+## 6. Gain: use good features to build models 
 - what counts as a good feature? See Domingos on this
@@
 
-# 7. Sparsity: output 20 recommendations from 200,000 products
+## 7. Sparsity: output 20 recommendations from 200,000 products
 
@@
 
-# 8. Value: Balancing conversion vs spend
+## 8. Value: Balancing conversion vs spend
 The value of an ordering
 
-# 9. Metric: precision-at-K
+## 9. Metric: precision-at-K
 
@@
 
-# 10. Experiment: A/B allocation
+## 10. Experimental settings: A/B allocation and uplift
 
@@
 This is done in the name of statistical validation of the effects or the uplift of the model.
+It also suggests that the effects of the model cannot be known in advance -- they can only be observed experimentally.
+Does this experimentality of the list point more to the propensities that Hansen says saturate experience?
 
-# 11. Infrastructure: realtime engine serves up top 5
-Only in certain settings are things realtime; or only in limited ways;
-Realtime has its own economies and cultures
+It is perhaps significant that this random allocation of customers to control and test groups occurs without any relation to the particular profile or propensities of the customer.
+It seeks to statistically validate the effects -- the *uplift* -- of the model on  conversion events by allowing the effects of the model on what people do to be measured.
+The lift or uplift refers to a change in events associated with the same groups of people.
+This effectively is an experiment in inhabiting two different worlds.
+This sets up a feedback look between the predictive model, the logistic regression, and the operation of the recommendation system.[^12] Once the experimental uplift methodology is in place, the recommendation system can constantly modify the recommendation list in light of subsequent data on the uplift.
+Predictions, then, are not bound by fixed parameters of the model.
 
-# 12. Similarity, substitution and complementarity
+[^12]: See [@Lo_2002] for an introduction to the construction of experimental set ups for uplift models.
+Market researchers and data miners seem to have started adopting a statistically grounded experimental predictive practice sometimes around 2000.
+See for instance [@Almquist_2001]  Response modelling, propensity modelling and uplift modelling all seek to identify associations between 'treatments' or interventions and the 'Responders', the people  affected by the treatment.
+Interestingly, these
+
+## 11. Infrastructure: realtime engine serves up top 5
+
+![Realtime infrastructure and its limits](~/lists/figure/hadoop.png)
+
+While much discussion of recent media has highlighted its realtime operation, Shreena's description of the infrastructure that implements the new model suggested a different temporal patterning.
+While raw data from Tesco Online feeds into the model every hour, it seems that recommendation list for each customer is only generated once a week.
+This makes sense.
+Customers only shop online every few days at the most, and in some cases, only every few weeks.
+To continually update the top 200 recommendations for every customer would be computationally expensive.
+
+Shreena did not say much else about the infrastructure but the 200 recommendations generated each week are certainly never seen by the customers.
+The realtime engine that actually generates results shown on the Tesco Online website only shows 5 after 'applying business rules.' Again, this might be seen as corroborating Hansen and Badiou's insistence on the necessarily limited grasp of any prediction system on the situation it encounters.
+Even the new model, situated amidst its experimental setup cannot fully encompass the surplus of sensibility it encounters.
+
+The brief mention of specific infrastructural elements such as `hadoop` is, however, significant.
+It points to the temporality of infrastructures themselves as a component of list practices.
+The possibility of adjusting the recommendations for every customer weekly derives from an infrastructure capable of collecting data, assimilating that data to predictive model able to generate recommendations that are relevant to individuals amongst 200,000 products.
+
+## 12. Similarity, substitution and complementarity
+
 - interesting contrasts here between the relations of things in a list
 - whereas the `arule` approach compares lists looking for similiarities and ways of completing lists, the more recent approaches look for micrological propensities in the list that relate to a particular time and place, amidst a continuously differentiable pathways between things ...
+- how do things associate with each other in complementarities, substitutions, etc.
 
-# 13. Complementarity drives conversion 
+## 13. Complementarity drives conversion 
 - the problem of action
 
-# Conclusion
-- lists have complex temporal structure
-- listing practices very closely entangled with problems of social order
-- lists undergoing a 'conversion' experience as they become more probabilistic
+## Conclusion
+- lists have complex temporal-material structure, since they mix prediction, infrastructural constraints, the individual characteristics of list-makers, statistical validations, database architectures and above all the surplus of sensibility associated with commodity production.
+- listing practices very closely entangled with problems of social order under predictive conditions or amidst predictive operations. 
+- lists undergoing a 'conversion' experience as they become more probabilistic, but their heterogeneous character as operations caught up in models, infrastructures, experiments, and the flow of commodities means that they can never form a closed set or dataset.
+Amidst prediction they become constitutively incomplete.
+
+## References

### 2016-10-11 lists_conversion_experience.rmd
starting basket similarity

@@
 Presentation presents another presentation I saw recently about online grocery shopping at Tesco, a large supermarket chain based in the UK.
-The lists in this setting are recommendations for further purchases under the title of 'Have you forgotten?'.
-When you shop for groceries online, the 'Have you forgotten' recommendations appear when at the checkout stage.
-While the title 'have you forgotten' implies that you will be shown items that could have been or are usually on a shopping list.
-The working assumption is a temporal one -- the list is incomplete either because you have forgotten to list something, or you didn't know what should be on your list, perhaps because you have never even thought of putting that thing on your list.
+The lists in this setting are recommendations for further grocery purchases under the title of 'Have you forgotten?'.
+When you shop for groceries online, the 'Have you forgotten' recommendations appear  at the checkout stage.
+While the title 'have you forgotten' implies that you will be shown items that could have been or are usually on a shopping list, the working assumption is a temporal one -- the list is incomplete either because you have forgotten to list something, or you didn't know what should be on your list, perhaps because you have never even thought of putting that thing on your list.
 
@@
 But a recommendation system finds it much harder to suggest items that you have not forgotten but may have never bought.
-If you do buy recommended items, then the recommender system has been effective.
+If you do buy recommended items, then the recommender system has been effective in shunting the situation differently.
 It has led to a **'conversion.'** From a supermarket perspective, the conversion should be high-value.
@@
 
-If we are interested in lists from various angles -- as asignifying semiotics as Maurizio Lazzarato calls it in his _Signs and Machines_ [@Lazzarato_2014] or as elements in 'a new order of spatio-temporal continuity for forms of economic, political and cultural life' as Celia Lury terms in her account of topological turn [@TBA], or in terms of the predictive conditions of contemporary experience [@Hansen_2015a] -- then contemporary list practices such as recommendations figure as sites of intersection between different forms.
+We could regard lists from various angles -- as asignifying semiotics as Maurizio Lazzarato calls it in his _Signs and Machines_ [@Lazzarato_2014] or as elements in 'a new order of spatio-temporal continuity for forms of economic, political and cultural life' as Celia Lury terms in her account of topological turn [@TBA].
+In this paper, I focus on lists and their propensities in terms of the predictive conditions of contemporary experience [@Hansen_2015a].
+List practices such as recommendations figure as sites of intersection between people, infrastructures, things, places and predictions.
 
@@
 The main narrative of the presentation concerned a shift from a set-based model for making lists to a predictive, probabilistic way of making lists in almost realtime and dealing with lists on a scale that is often difficult to grasp concretely.
-The change in list-making associated with learned models is increasingly widespread, and
+The change in list-making associated with learned models is increasingly widespread, and relies on a probabilistic treatment of people and things.
 
-I respond to the presentation in two ways -- firstly by recounting, contextualising and commenting on some of the main steps in making the 'Have your forgotten?' list  as a example of the sense-making done by data scientists and customer analytics teams in many places; and secondly, by trying to step into the list-making practices  as an occasion to intervene and participate in the ontological conversion practices they engage in.
+I respond to the presentation in two ways.
+First, by recounting, contextualising and commenting on some of the main steps in making the 'Have your forgotten?' list  as a example of the sense-making done by data scientists and customer analytics teams in many places.
+Many of these steps were mentioned only briefly in the presentation, and they are largely taken for granted as part of predictive analytic practice.
+Second, by trying to step into the list-making practices  as an occasion to intervene and participate in the ontological conversion practices they engage in.
 This is no mere methodological problem if, as Mark Hansen has recently suggested,
@@
 
-## 1. The context of the presentation - Consumption and shopping
+## 1. The context of the presentation: 'statistics and operations research in industry'
 
@@
 
-Shreena described her work at DunnHumby in terms of a shift in models from a 'rules-based list' to a 'relevancy model.' 
-This was a significant change in how the 'Have You Forgotten' list was constructed from one kind of model to another.
+Patel described her work at DunnHumby in terms of a shift in models from a 'rules-based list' to a 'relevancy model.' This was a significant change in how the 'Have You Forgotten' list was constructed from one kind of model to another.
 For me, the shift in models matters since it amounts to a much more probabilistic structuring of lists.
+
 I assume that the rules-based model derived from the early 1990s work on the `apriori` algorithm by Rakesh Agrawal and Ramakrishnan Srikanti at IBM Research Alameda [@Agrawal_1994].
@@
 [http://www.dssresources.com/newsletters/66.php]
-It is a key algorithm in 'market basket analysis' and attempts to re-structure shopping lists.
+The `apriori` algorithm is widely used in 'market basket analysis' and attempts to re-structure shopping lists in terms of associations between items that are frequently bought together.
 
@@
 The association rule-based model for the `Groceries` dataset suggests some of the characteristic ordering practices of data-mining approaches to lists. 
-The association between items is expressed in the form of _rules_. These rules, a term derived from decision support literature, define sets of associations between items. 
+The association between items is expressed in the form of _rules_.
+These rules, a term derived from decision support literature, define sets of associations between items.
 As a set theory-based practice, the association rules define likely sets of items or shopping baskets in terms of decreasing probability.
@@
 So any recommendation in the 'Have You Forgotten' list should probably put them in that order.
+
 Association rules and the `apriori` algorithm present social order in terms of sets of common associations.
-This means that that the algorithm can be understood in terms of the set-based multiples proposed by Alain Badiou.
+This means that that the algorithm can be understood in terms of the set-based multiples proposed by Alain Badiou:
 
@@
 Badiou in his account of multiples highlight the contrast between two different ways of grasping a multiple -- in terms of the 'numerical one' and the 'vital One' [@Badiou_2004, 72], between the 'closed' and the 'open' respectively.
+
 The `apriori` algorithm, might be seen as an attempt to deal with the excess of things in a Tesco supermarket, and the lack of interest or engagement that people have with them by identifying subsets whose membership is stronger by virtue of the other sets of associations that people bring with them.
@@
 
-Shreena observed that the 'Have you forgotten' lists matter to people, and we can see why in the light of this proliferation of things in the list.
+Patel observed that the 'Have you forgotten' lists matter to people, and we can see why in the light of this proliferation of things in the list.
 A `r total` list of items is quite hard to list let alone read.
-Shreena said that some people log on and go straight to the checkout, allowing the recommendation system to supply them with the suggestions that they then move to the shopping basket.
+Patel said that some people log on and go straight to the checkout, allowing the recommendation system to supply them with the suggestions that they then move to the shopping basket.
 This is the online equivalent of using an old shopping list and only adding new items when needed. 
@@
 But it also suggests a transformation in the way that lists exist in the world.
+They are less the product of a discrete agent than an effect of an environment teeming with possible combinations and complementaries, a surplus of possible purchases that the supermarket needs to continually maintain but also filter in order to make it possible for something to happen.
 
@@
 
-Shreena described the data they on grocery purchases. 'We have 52 weeks of purchase data for each customer.' 
+Patel described the data they on grocery purchases. 'We have 52 weeks of purchase data for each customer.' 
 We know that the construction of lists has been transformed by recommendation systems.
@@
 The problem that the proliferation of things on the shelves of supermarket or grocery warehouse confronts is partly combinatorial.
-With  `r total` products (actually Shreena mentions 200,000 products) can be combined in truly astronomical numbers of combinations.
+With  `r total` products (actually Patel mentions 200,000 products) can be combined in truly astronomical numbers of combinations.
 If a typical shopping list has 20 items, then there are `r format(choose(total, 20), 2)` possible combinations.
@@
 In some ways, Hansen's account resembles the Badiouean ontology of the multiple in which a closed set is surrounded and conditioned by an open set that is full of many more potentials.
-However, if we think of the  new model that Shreena describes as the delimited (or delimiting?) system within   the larger surplus of sensibility, how that help us understand what the 'Have you forgotten?' recommendations do?
+However, if we think of the  new model that Patel describes as the delimited (or delimiting?) system within   the larger surplus of sensibility, how that help us understand what the 'Have you forgotten?' recommendations do?
 
@@
 ## 6. Gain: use good features to build models
+
 - what counts as a good feature? See Domingos on this
@@
 
-It was surprising that Shreena spoke relatively briefly about the new models.
+It was surprising that Patel spoke relatively briefly about the new models.
 She showed a graph that compared the old rule-based model with some of the new alternatives -- logistic regression, random forests, and a few others.
 She dismissed most of the models quite quickly and focused only on one, the logistic regression model.
-Or more accurately, she just said that logistic regression did better than other, and there was little more to discusson on that front.
+Or more accurately, she just said that logistic regression did better than other, and there was little more to discuss on that front.
 The measure of doing better here is called precision, and it measures the proportion of the recommended products that are actually purchased.
@@
 
+The fact that so little was said about the models themselves suggests that they are now largely taken for granted in recommendation systems.
+The models she mentioned are all classifiers, predictive models that treat a set of variables in a datasets as a vector space in which good classification means finding a way of defining a decision boundary between points that belong to different classes of things.
+Here the classes of things are binary: `recommended` or `not recommended.` Some classifier models, including logistic regression, generate a probability rather than simply a classification.
+
 ## 7. Sparsity: output 20 recommendations from 200,000 products
@@
 Predictive models, we might say, encounter their own surplus of sensibility in the data as they search for the most accurate predictions.
-A  very brief comment from Shreena -- 'we have lots of zeros' -- points to the problem here.
+A  very brief comment from Patel -- 'we have lots of zeros' -- points to the problem here.
 Each of the millions of Tesco online grocery shoppers has bought some selection of the 200,000 products.
@@
 So the raw data -- the 52 weeks of purchases for each customer -- can be aggregated to increase the density of the data in ways that helps the model.
-Shreena mentions some of the 'features' used in the new model but somewhat strangely at this point, her presentation turns toward the difficulties of measuring the success of the model and the infrastructural limitations on making recommendations in ways that are more alive to what is happening right now.
+Patel mentions some of the 'features' used in the new model but somewhat strangely at this point, her presentation turns toward the difficulties of measuring the success of the model and the infrastructural limitations on making recommendations in ways that are more alive to what is happening right now.
 
@@
 The discussion of imperfection was brief - it incorporated the notion of recommendations weighted by price and the notion of precision (or the proportion of products actually purchased.).
-Shreena simply said that this metric was imperfect, but without elaborating why.
+Patel simply said that this metric was imperfect, but without elaborating why.
 Instead she moved to describe some of the ways in which the recommendation system worked in practice.
@@
 
-While much discussion of recent media has highlighted its realtime operation, Shreena's description of the infrastructure that implements the new model suggested a different temporal patterning.
+While much discussion of recent media has highlighted its realtime operation, Patel's description of the infrastructure that implements the new model suggested a different temporal patterning.
 While raw data from Tesco Online feeds into the model every hour, it seems that recommendation list for each customer is only generated once a week.
@@
 
-Shreena did not say much else about the infrastructure but the 200 recommendations generated each week are certainly never seen by the customers.
+Patel did not say much else about the infrastructure but the 200 recommendations generated each week are certainly never seen by the customers.
 The realtime engine that actually generates results shown on the Tesco Online website only shows 5 after 'applying business rules.' Again, this might be seen as corroborating Hansen and Badiou's insistence on the necessarily limited grasp of any prediction system on the situation it encounters.
@@
 
-## 12. Similarity, substitution and complementarity
+## 12. Similarity, substitution, complementarity and micrological propensities
 
-- interesting contrasts here between the relations of things in a list
-- whereas the `arule` approach compares lists looking for similiarities and ways of completing lists, the more recent approaches look for micrological propensities in the list that relate to a particular time and place, amidst a continuously differentiable pathways between things ...
-- how do things associate with each other in complementarities, substitutions, etc.
+![Basket similarity](figure/patel_basket_similarity.png)
+
+What does it matter that recommendations are relevant?
+Relevance is closely connected to propensities in the predictive condition.
+From DunnHumby's standpoint, the relevance of a recommendation relates directly to conversion.
+From the standpoint of the ontological conversion, and  if Hansen is right, the real or actual fabric of prediction resides in the propensities of things, what he calls the 'complex and diffuse calculus' that only subsequently and provisionally congeals as a conversion event.
+The 'micrological propensities' of things feed directly into an sense of the relevance of recommendations.
+
+How does the DunnHumby model deal with the micrological propensities of things?
+In her presentation, Patel  spent quite a lot of time addressing the problem of basket similarity in terms of substitutions and complementarities between things.
+She concludes by saying 'basket similarity is the future' and 'complementarities drive conversion.'
+
+Whereas the association rules focus on complementarities -- if you buy bread, you might want milk, if you buy beer, you might want diapers -- the substitutions look for micrological propensities that relate to gradations of preference, and variations in time and place.
+Micrological propensities in the list relate to a particular time and place, amidst a continuously differentiable pathways between things.
+
+How do things associate with each other in  substitutions and complementarities?
+Patel's comments on this are very brief.
+Like the mention of `hadoop`, they imply quite extensive but relatively well-established data practices.
+She speaks for instance of a 'design matrix of 14,000 columns' and the use of `L1 regularization` to drive coefficients down to zero.
+
+```{r biscuits, echo=FALSE, cache=TRUE}
+    library(dplyr)
+    library(ggplot2)
+    library(readr)
+    bisc = read_tsv('data/biscuits.csv')
+    bisc %>% select(price, name)  %>% order(price, decreasing=TRUE)    
+    bisc[order(bisc$price, decreasing=TRUE), c('name','price')]
+```
+We can imagine this in terms of biscuits.
 

### 2016-10-12 lists_conversion_experience.rmd
product similarity matrix

@@
 
-Presentation presents another presentation I saw recently about online grocery shopping at Tesco, a large supermarket chain based in the UK.
+I saw recently a presentation about online grocery shopping at Tesco, a large supermarket chain based in the UK.
 The lists in this setting are recommendations for further grocery purchases under the title of 'Have you forgotten?'.
-When you shop for groceries online, the 'Have you forgotten' recommendations appear  at the checkout stage.
+When you shop for groceries online, a list of 'Have you forgotten' recommendations appear  at the checkout stage.
 While the title 'have you forgotten' implies that you will be shown items that could have been or are usually on a shopping list, the working assumption is a temporal one -- the list is incomplete either because you have forgotten to list something, or you didn't know what should be on your list, perhaps because you have never even thought of putting that thing on your list.
 
-As we will see, the items on this list are recommendations. They are only a few of them: five. The question: which five items and in what order?
+The items on this list are recommendations. They are only a few of them: five. The question: which five items and in what order?
 
@@
 Shopping lists are quite prosaic -- a dozen lines scrawled on the back of a used envelope stuck on the refrigerator door  -- and sublime -- a million products ranked in some kind of order for tens or hundreds of millions of people.
-I regard this list of just five suggestions given to Tesco's online customers as symptomatic of a major contemporary form of ordering, the recommendation list, that we see in many  forms of news, media, entertainment, marketing, advertising.
+I regard the list of five suggestions show to Tesco's online customers at the checkout as symptomatic of a major contemporary form of ordering, the recommendation list, that we see in many  forms of news, media, entertainment, marketing, advertising.
 
 In this respect, the presentation was not really surprising or exceptional.
-Its ordinariness and prosaic character stands out as the most significant feature.
+The ordinariness and prosaic character of grocery shopping stands out as the most significant feature.
 The presentation itself is a working example for an argument about our experience of predictive ordering.
 In this paper, I re-present the presentation using a mixture of theoretical and practical resources.
-The theory is largely recent media theory, and the practical resources are that of coding and predictive models.
+The theory is largely recent media theory and its accounts of prediction in media.
+The practical resources are that of coding and predictive models.
 
@@
 On the other hand, very unexpected items are quite easy to remember and tend to even the score in the favour of the children.
-Balancing between the improbable and memorable, between a predictable list like 'apples, bananas, carrots' and an unusual list such as 'agave syrup, boccadilla,cilantro' is complicated.
-
+Balancing between the improbable and memorable, between a predictable list like 'apples, bananas, carrots' and an unusual list such as 'agave syrup, boccadilla, cilantro' is complicated.
 Something similar applies to recommendation systems.
@@
 
-But a recommendation system finds it much harder to suggest items that you have not forgotten but may have never bought.
+A recommendation system should suggest items that you have not forgotten but may have never bought.
 If you do buy recommended items, then the recommender system has been effective in shunting the situation differently.
@@
 We could regard lists from various angles -- as asignifying semiotics as Maurizio Lazzarato calls it in his _Signs and Machines_ [@Lazzarato_2014] or as elements in 'a new order of spatio-temporal continuity for forms of economic, political and cultural life' as Celia Lury terms in her account of topological turn [@TBA].
-In this paper, I focus on lists and their propensities in terms of the predictive conditions of contemporary experience [@Hansen_2015a].
+In this paper, I focus on lists and their propensities in terms of the predictive conditions of contemporary media experience [@Hansen_2015a].
 List practices such as recommendations figure as sites of intersection between people, infrastructures, things, places and predictions.
@@
 
-dir('data/')
 csvs = list.files('data', pattern = '*.csv',full.names=TRUE)
@@
 
-str(df)
-colnames(df)
 ggplot(df, aes(x=price,y=ContentsQuantity)) + geom_point() 
-ggplot(df, aes(x=price)) + geom_histogram(bins=200)
 fivenum(df$price)
@@
 How do things associate with each other in  substitutions and complementarities?
-Patel's comments on this are very brief.
-Like the mention of `hadoop`, they imply quite extensive but relatively well-established data practices.
+Patel's comments on this are very brief but suggest that the 'Have You Forgotten?' list becomes an increasingly mutable entity, open to many substitutions and re-writings.
+Like the mention of `hadoop`, what Patel said about substitution and complementarity implies extensive but relatively well-established data practices.
 She speaks for instance of a 'design matrix of 14,000 columns' and the use of `L1 regularization` to drive coefficients down to zero.
+This is machine learning talk.
+What is being modelled here?
+No longer whether to recommend a product to a customer or not, but whether something could be substituted for what is already in the recommendation list.
 
@@
 We can imagine this in terms of biscuits.
+A customer's recommendation list might already include `r bisc$name[10]`.
+The 'self-learning substitution model' might be able to substitute something in that list that is of higher value, either because it is bigger or more expensive.
+
+Similarly, to model the potential complementarities   between things entails a matrix of comparisons.
+`Nutella` complements `bananas,` but the converse is not necessarily true since `bananas` need not complement `nutella.`  If then the
 
@@
 ## Conclusion
+
+My investigation of lists has been centred around the transformation in their predictive underpinnings.
+Following Hansen, I have been suggesting that the predictive listing in the form of open-ended recommendation is symptomatic of changes in the distribution of media.
+The idea that media should no longer be considered as extensions or prostheses of human communication but the environmental situation in which different propensities play out is increasingly central to contemporary media theory.
+
+I diverge from Hansen in how I account for this change.
+Hansen refers to an ontological transformation in which predictive media or recommendation systems would be seen as the closed form of the open propensities of the world.
+I have not explored the Whitehead-derived philosophy of propensities in the world that he develops there, only because I am not sure that we need speculative theory to grapple with the material practices of prediction.
+Already in the DunnHumby model and the Tesco 'Have You Forgotten?' recommendations, many of the traits of predictive processes are at work.
+
 - lists have complex temporal-material structure, since they mix prediction, infrastructural constraints, the individual characteristics of list-makers, statistical validations, database architectures and above all the surplus of sensibility associated with commodity production.

### 2016-10-12 extract_images.sh

@@ -1,0 +1,1 @@
+grep '\(#\{1,2\}\)\|!\[\|> ' lists_conversion_experience.md > slides.md 

### 2016-10-13 extract_images.sh
slide work

@@ -1,1 +1,0 @@
-grep '\(#\{1,2\}\)\|!\[\|> ' lists_conversion_experience.md > slides.md 

### 2016-10-13 extract_slides.sh
extract slides

@@ -1,0 +1,2 @@
+#!/bin/bash
+grep '\(#\{1,2\}\)\|!\[\|> ' lists_conversion_experience.md > slides.md 

### 2016-10-16 lists_conversion_experience.rmd
slide work
playing with formats

--- 2016-10-11 lists_conversion_experience.rmd
@@
+---
+title: "An Example Using the Tufte Style"
+author: "John Smith"
+output:
+  tufte::tufte_handout: default
+tufte::tufte_html: default
+---
+
 # 'Have you forgotten?' Conversion experiences in list worlds
@@
     bisc = read_tsv('data/biscuits.csv')
-    bisc %>% select(price, name)  %>% order(price, decreasing=TRUE)    
     bisc[order(bisc$price, decreasing=TRUE), c('name','price')]

### 2016-10-16 lists_conversion_experience.rmd
messing up the formatting!

--- 2016-10-12 lists_conversion_experience.rmd
@@
+---
+title: "An Example Using the Tufte Style"
+author: "John Smith"
+output:
+  tufte::tufte_handout: default
+tufte::tufte_html: default
+---
+
 # 'Have you forgotten?' Conversion experiences in list worlds
+
+```{r setup, cache=FALSE, include=FALSE}
+library(knitr)
+output <- opts_knit$get("rmarkdown.pandoc.to")
+opts_chunk$set(warnings=FALSE)
+opts_chunk$set(messages=FALSE)
+```
 
@@
 
-![Realtime infrastructure and its limits](~/lists/figure/hadoop.png)
+![Realtime infrastructure and its limits](figure/hadoop.png)
 

### 2016-10-17 lists_conversion_experience.rmd
mainly playing with formatting

@@
 ---
-title: "An Example Using the Tufte Style"
-author: "John Smith"
+title: "'Have you forgotten?' Conversion experiences in list worlds"
+author: "Adrian Mackenzie"
 output:
@@
 output <- opts_knit$get("rmarkdown.pandoc.to")
-opts_chunk$set(warnings=FALSE)
-opts_chunk$set(messages=FALSE)
+opts_chunk$set(warning=FALSE)
+opts_chunk$set(message=FALSE)
+opts_chunk$set(error=TRUE)
 ```

### 2016-10-17 render_tufte.r

@@ -1,0 +1,6 @@
+library(tufte)
+library(rmarkdown)
+library(tint)
+render('lists_conversion_experience.rmd',output_format=tufte_handout())
+render('lists_conversion_experience.rmd',output_format=ioslides_presentation)
+render('lists_conversion_experience.rmd',output_format=tintPdf())

### 2016-10-18 lists_conversion_experience.rmd
adding a bit on propensities

--- 2016-10-16 lists_conversion_experience.rmd
@@
 ---
-title: "An Example Using the Tufte Style"
-author: "John Smith"
+title: "'Have you forgotten?' Conversion experiences in list worlds"
+author: "Adrian Mackenzie"
+date: "`r Sys.Date()`"
 output:
 tufte::tufte_handout: default
-tufte::tufte_html: default
+citation_package: biber
+latex_engine: xelatex
+bibliography: ~/ref_bibs/uni.bib
+link-citations: yes
 ---
@@
 library(knitr)
+library(tint)
+knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tint'))
+options(htmltools.dir.version = FALSE)
 output <- opts_knit$get("rmarkdown.pandoc.to")
-opts_chunk$set(warnings=FALSE)
-opts_chunk$set(messages=FALSE)
+opts_chunk$set(warning=FALSE)
+opts_chunk$set(message=FALSE)
+opts_chunk$set(error=TRUE)
 ```
@@
 
-Shopping lists are quite prosaic -- a dozen lines scrawled on the back of a used envelope stuck on the refrigerator door  -- and sublime -- a million products ranked in some kind of order for tens or hundreds of millions of people.
+Shopping lists such as Figure \ref{fig:shopping_list} are quite prosaic -- a dozen lines scrawled on the back of a used envelope stuck on the refrigerator door  -- and sublime -- a million products ranked in some kind of order for tens or hundreds of millions of people.
 I regard the list of five suggestions show to Tesco's online customers at the checkout as symptomatic of a major contemporary form of ordering, the recommendation list, that we see in many  forms of news, media, entertainment, marketing, advertising.
@@
 
-```{r simulate_basket, echo=FALSE, cache=TRUE}
+```{r simulate_basket, echo=FALSE, cache=TRUE, fig.margin=TRUE}
 library(dplyr)
@@
 
-```{r arules, echo=TRUE}
+```{r arules, echo=TRUE, fig.margin=TRUE}
 library(arules)
@@
 
-```{r tesco, cache=TRUE, }
+```{r tesco, cache=TRUE, fig.margin=TRUE}
     library(ggplot2)
@@
 
+## 13. Complementarity drives conversion
+
+![propensity_matrix](figure/propensity_conversions.png)
+
 Similarly, to model the potential complementarities   between things entails a matrix of comparisons.
-`Nutella` complements `bananas,` but the converse is not necessarily true since `bananas` need not complement `nutella.`  If then the
+`Nutella` complements `bananas,` but the converse is not necessarily true since `bananas` need not complement `nutella.`
 
-## 13. Complementarity drives conversion 
-- the problem of action
+How should we understand the effect of complementarity?
+Patel says that 'complementarity drives conversion.' That is, the matrix of similarities expressed as numerical associations between products best animates the conversion of recommendations into purchases.
+
+In the similarity matrix, shaped by the history of past recommendations and purchases made by this customer, lie the propensities of things that Hansen (following Popper) attributes to the world.
+Karl Popper speaks of the future as a 'crystallization of propensities' [@Popper_1990,18]
 

### 2016-10-18 tesco_fetch_product_images.sh

@@ -1,6 +1,6 @@
 #!/bin/bash
-
 cut -f11 $1|xargs wget -P $2 
 cd $2
-montage  -mode concatenate -tile 40x IDShot_90x90.jpg.* out.jpg
-eog out.jpg
+montage  -mode concatenate -tile 40x IDShot_90x90.jpg.* ../combined_images/$2.jpg
+cd ..
+eog combined_images/$2.jpg

### 2016-10-19 lists_conversion_experience.rmd
writing abstract

@@
 ---
-title: "'Have you forgotten?' Conversion experiences in list worlds"
+title: "'Have you forgotten?' Conversion experiences in lists"
 author: "Adrian Mackenzie"
@@
 
-# 'Have you forgotten?' Conversion experiences in list worlds
+# 'Have you forgotten?' Conversion experiences in lists
 
@@
 Shopping lists such as Figure \ref{fig:shopping_list} are quite prosaic -- a dozen lines scrawled on the back of a used envelope stuck on the refrigerator door  -- and sublime -- a million products ranked in some kind of order for tens or hundreds of millions of people.
-I regard the list of five suggestions show to Tesco's online customers at the checkout as symptomatic of a major contemporary form of ordering, the recommendation list, that we see in many  forms of news, media, entertainment, marketing, advertising.
+Such lists are highly repetitive, yet mix items of great diversity and display complicated temporal structures, spanning past and future, actual and potential.
+I regard the list of five suggestions show to Tesco's online customers at the checkout as exemplifying a major contemporary form of ordering, the recommendation list, that we see in many  forms of news, media, entertainment, marketing, advertising.
 
-In this respect, the presentation was not really surprising or exceptional.
-The ordinariness and prosaic character of grocery shopping stands out as the most significant feature.
-The presentation itself is a working example for an argument about our experience of predictive ordering.
+The presentation I saw was not really surprising or exceptional.
+It was a presentation by a data scientist to data scientists, describing an implementation of a predictive model as well as some of the problems and potentials of the predictive model.
+I see the presentation itself as providing materials for an argument about our experience of predictive ordering.
 In this paper, I re-present the presentation using a mixture of theoretical and practical resources.
 The theory is largely recent media theory and its accounts of prediction in media.
-The practical resources are that of coding and predictive models.
+The practical resources are that of coding and predictive models.[^12]
 
-The problem of incomplete lists reminds me of the game 'I went shopping and I bought ...
+[^12]: The problem of incomplete lists reminds me of the game 'I went shopping and I bought ...
 ' that adults sometimes play with children.
@@
 If you do buy recommended items, then the recommender system has been effective in shunting the situation differently.
-It has led to a **'conversion.'** From a supermarket perspective, the conversion should be high-value.
-There is little value in recommending a whole lot of small and cheap items such as rubber bands or toothpicks.
+It has led to a **'conversion.'** From a supermarket perspective, the conversion should be high in price.
+There is little value in successfully recommending a whole lot of small and cheap items such as rubber bands or toothpicks.
 So it might need to choose amongst different recommendations and weight high value items differently to low value items.
@@
 
-We could regard lists from various angles -- as asignifying semiotics as Maurizio Lazzarato calls it in his _Signs and Machines_ [@Lazzarato_2014] or as elements in 'a new order of spatio-temporal continuity for forms of economic, political and cultural life' as Celia Lury terms in her account of topological turn [@TBA].
-In this paper, I focus on lists and their propensities in terms of the predictive conditions of contemporary media experience [@Hansen_2015a].
-List practices such as recommendations figure as sites of intersection between people, infrastructures, things, places and predictions.
+**How do we approach lists as objects of research?** While lists such as recommendations intersect strongly with other meaning and sense-making practices in everyday life and popular culture, they are also somewhat refractory to discourse  and textual analysis methods.
+On the one hand, lists are highly associative.
+They can be interpreted or decoded semiotically, but they also exhibit great variations in textuality that multiply semiotic readings.
+On the other hand, as operational inscriptions, they can be treated ethnomethodologically, as the production of ('the order of social settings through ...
+shared sense making practices.')[https://en.wikipedia.org/wiki/Ethnomethodology]
 
-**How do we approach lists as objects of research?** While lists such as recommendations intersect strongly with other meaning and sense-making practices in everyday life and popular culture, they are also somewhere refractory to discourse analysis and textual analysis methods of humanities and social science research.
-For on the one hand, lists are highly associative.
-They can be interpreted or decoded semiotically, but they also exhibit great variations in textuality that multiply semiotic readings.
-For on the other hand, as operational entities, they can be treated ethnomethodologically, as the production of ('the order of social settings through ...
-shared sense making practices.')[https://en.wikipedia.org/wiki/Ethnomethodology]
+List practices such as online grocery recommendations figure as sites of intersection between people, infrastructures, things, places and predictions.
+We might, therefore,  approach lists from various novel angles -- as asignifying semiotics as Maurizio Lazzarato calls it in his _Signs and Machines_ [@Lazzarato_2014] or as elements in 'a new order of spatio-temporal continuity for forms of economic, political and cultural life' as Celia Lury terms in her account of topological turn [@Lury_2012].
+In this paper, I focus on lists and list-making in terms of the predictive conditions of contemporary media experience [@Hansen_2015a].
 
@@
 This effectively is an experiment in inhabiting two different worlds.
-This sets up a feedback look between the predictive model, the logistic regression, and the operation of the recommendation system.[^12] Once the experimental uplift methodology is in place, the recommendation system can constantly modify the recommendation list in light of subsequent data on the uplift.
+This sets up a feedback look between the predictive model, the logistic regression, and the operation of the recommendation system.[^122] Once the experimental uplift methodology is in place, the recommendation system can constantly modify the recommendation list in light of subsequent data on the uplift.
 Predictions, then, are not bound by fixed parameters of the model.
 
-[^12]: See [@Lo_2002] for an introduction to the construction of experimental set ups for uplift models.
+[^122]: See [@Lo_2002] for an introduction to the construction of experimental set ups for uplift models.
 Market researchers and data miners seem to have started adopting a statistically grounded experimental predictive practice sometimes around 2000.
@@
 
-![propensity_matrix](figure/propensity_conversions.png)
+![propensity_matrix](figure/propensity_conversions.jpg)
 

### 2016-10-20 lists_conversion_experience.rmd
fixing manuscript with graphics, etc

@@
 link-citations: yes
+header-includes:
+    - \usepackage{setspace}
+    - \doublespacing
 ---
@@
 ```
+
+## Abstract
+
+Amidst the predictive conditions of contemporary media-culture, the inevitable incompleteness of lists is both a problem and a potential. 
+This paper investigates practices of list-making associated with online grocery shopping at a  UK and global supermarket chain, Tesco. 
+It will analyse how a recommendation system suggests a small number of grocery items of personal relevance to each of the millions of online grocery shoppers. 
+The predictive modelling practices aim to up-sell and cross-sell some of the 250,000 items in the supermarket's inventory. 
+I find it interesting to approach the list-making practices  of the grocery recommendation system as a mundane case of the predictive intermediaries operating in many places. 
+The transformation or conversion of listing into predicting can be analysed in recommendation systems that extend existing lists with potential new items, thus transforming the temporal and agential fabric of the list.
+While all written lists mix past and future actions, actual and potential events, their predictive re-writing in recommendation systems adds a new operational dynamic.
+In this operational dynamic, the incompleteness of lists, their propensity to expand or change, becomes more important than their capacity to exhaustively enumerate things. 
+The conversion of lists from finite to in-definite in recommendation systems suggests we might need to re-conceptualise how we think about social order and structure in media cultures.
 
@@
 
-ggplot(df, aes(x=price,y=ContentsQuantity)) + geom_point() 
-fivenum(df$price)
+ggplot(df, aes(x=price,y=ContentsQuantity)) + geom_point() + theme(axis.text=element_text(size=10)) 
 df[df$price>=20,'name']
@@
 DunnHumby -- Edwina Dunn and Clive Humby -- are famous for having convinced the CEO of Tesco sometime in 1991 that a loyalty card program could change the supermarket chain's relationship to its customers.
-[@Humby_1984]
-Clive Humby's academic work is hard to track down, but an early paper of his given at the Conference of Your Operational Researchers in Nottingham in 1984 suggests the direction that he and later Tesco would take in constructing lists [@Humby_1984]
+[@Okeefe_1984]  Clive Humby's academic work is hard to track down, but an early paper of his given at the Conference of Your Operational Researchers in Nottingham in 1984 suggests the direction that he and later Tesco would take in constructing lists [@Okeefe_1984]
 
@@
 
-```{r tesco, cache=TRUE, fig.margin=TRUE}
+```{r tesco, cache=TRUE, fig.cap ='Tesco surplus',  fig.margin=TRUE}
     library(ggplot2)
@@
 
-    ggplot(tesco[tesco$actual>50,], aes(x=labels, y=actual, fill=labels)) +geom_bar(stat='identity') + coord_flip() + ggtitle('Tesco grocery items with more than 50 products') + theme(legend.position="none")
+    ggplot(tesco[tesco$actual>50,], aes(x=labels, y=actual, fill=labels)) +geom_bar(stat='identity') + coord_flip() + ggtitle('Tesco grocery items with more than 50 products') + theme(legend.position="none", axis.text=element_text(size=9))
 ```
@@
 
-```{r sparsity, cache=TRUE}
+```{r sparsity, echo=FALSE, cache=TRUE, fig.margin=TRUE, fig.cap ='The sparsity of data' }
+library(ggplot2)
+library(reshape2)
+p = as.logical(rbinom(n=2e5, size=1, prob=0.001))
+head(p, 30)
+m = (matrix(p, nrow=1000, byrow=TRUE))
 
-library(ggplot2)
-customers = seq(1000000,1)
-products = seq(200000,1)
-format(1000000 * 200000, 3)
-p = rbinom(2e5, n=1,prob=0.5)
-m = matrix(p, nrow=900, byrow=TRUE)
-df = as.data.frame(m)
-#TODO: plot here, could possibly use the images from tescoe
+melted <- melt(m)
+melted$value <- as.logical(melted$value)
+ggplot(melted, aes(x = Var2, y = Var1, fill = value)) + geom_tile() +
+    scale_fill_manual(values = c("grey", "black"))
 
@@
 
-```{r biscuits, echo=FALSE, cache=TRUE}
-    library(dplyr)
-    library(ggplot2)
+```{r biscuits, echo=FALSE, cache=TRUE, fig.margin=TRUE}
     library(readr)
+    library(knitr)
     bisc = read_tsv('data/biscuits.csv')
-    bisc[order(bisc$price, decreasing=TRUE), c('name','price')]
+    kable(bisc[order(bisc$price, decreasing=TRUE)[1:30], c('name','price')])
 ```
+
 We can imagine this in terms of biscuits.
@@
 
-![propensity_matrix](figure/propensity_conversions.jpg)
+![propensity_matrix](figure/propensity_conversions.png)
 
@@
 My investigation of lists has been centred around the transformation in their predictive underpinnings.
+How should we think about the list-making practices of these models, databases and computational infrastructures?
 Following Hansen, I have been suggesting that the predictive listing in the form of open-ended recommendation is symptomatic of changes in the distribution of media.
@@
 
-I diverge from Hansen in how I account for this change.
-Hansen refers to an ontological transformation in which predictive media or recommendation systems would be seen as the closed form of the open propensities of the world.
+Hansen refers to an ontological transformation in which predictive media or recommendation systems would be seen as the closed realisation of the actually open propensities of the world.
+By understanding the world in terms of propensities, he argues, we have a change to engage with contemporary predictive media.
+
 I have not explored the Whitehead-derived philosophy of propensities in the world that he develops there, only because I am not sure that we need speculative theory to grapple with the material practices of prediction.
@@
 
+I diverge from Hansen in how I account for this change. Rather than looking for a speculative theory that sets out a world in which prediction would be the closed form,
+
+- listing practices are very closely entangled with problems of social order under predictive conditions or amidst predictive operations. 
 - lists have complex temporal-material structure, since they mix prediction, infrastructural constraints, the individual characteristics of list-makers, statistical validations, database architectures and above all the surplus of sensibility associated with commodity production.
-- listing practices very closely entangled with problems of social order under predictive conditions or amidst predictive operations. 
 - lists undergoing a 'conversion' experience as they become more probabilistic, but their heterogeneous character as operations caught up in models, infrastructures, experiments, and the flow of commodities means that they can never form a closed set or dataset.

### 2016-10-21 abstract.rmd
fixing up formatting and adding new figures

--- 2016-10-20 lists_conversion_experience.rmd
@@
----
-title: "'Have you forgotten?' Conversion experiences in lists"
-author: "Adrian Mackenzie"
-date: "`r Sys.Date()`"
-output:
-tufte::tufte_handout: default
-citation_package: biber
-latex_engine: xelatex
-bibliography: ~/ref_bibs/uni.bib
-link-citations: yes
-header-includes:
-    - \usepackage{setspace}
-    - \doublespacing
----
-
-# 'Have you forgotten?' Conversion experiences in lists
-
-```{r setup, cache=FALSE, include=FALSE}
-library(knitr)
-library(tint)
-knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tint'))
-options(htmltools.dir.version = FALSE)
-output <- opts_knit$get("rmarkdown.pandoc.to")
-opts_chunk$set(warning=FALSE)
-opts_chunk$set(message=FALSE)
-opts_chunk$set(error=TRUE)
-```
-
-## Abstract
-
 Amidst the predictive conditions of contemporary media-culture, the inevitable incompleteness of lists is both a problem and a potential. 
@@
 The conversion of lists from finite to in-definite in recommendation systems suggests we might need to re-conceptualise how we think about social order and structure in media cultures.
-
-## Introduction: the presentation of a presentation
-
-I saw recently a presentation about online grocery shopping at Tesco, a large supermarket chain based in the UK.
-The lists in this setting are recommendations for further grocery purchases under the title of 'Have you forgotten?'.
-When you shop for groceries online, a list of 'Have you forgotten' recommendations appear  at the checkout stage.
-While the title 'have you forgotten' implies that you will be shown items that could have been or are usually on a shopping list, the working assumption is a temporal one -- the list is incomplete either because you have forgotten to list something, or you didn't know what should be on your list, perhaps because you have never even thought of putting that thing on your list.
-
-The items on this list are recommendations. They are only a few of them: five. The question: which five items and in what order?
-
-![shopping_list](figure/1_shopping.png )
-
-Shopping lists such as Figure \ref{fig:shopping_list} are quite prosaic -- a dozen lines scrawled on the back of a used envelope stuck on the refrigerator door  -- and sublime -- a million products ranked in some kind of order for tens or hundreds of millions of people.
-Such lists are highly repetitive, yet mix items of great diversity and display complicated temporal structures, spanning past and future, actual and potential.
-I regard the list of five suggestions show to Tesco's online customers at the checkout as exemplifying a major contemporary form of ordering, the recommendation list, that we see in many  forms of news, media, entertainment, marketing, advertising.
-
-The presentation I saw was not really surprising or exceptional.
-It was a presentation by a data scientist to data scientists, describing an implementation of a predictive model as well as some of the problems and potentials of the predictive model.
-I see the presentation itself as providing materials for an argument about our experience of predictive ordering.
-In this paper, I re-present the presentation using a mixture of theoretical and practical resources.
-The theory is largely recent media theory and its accounts of prediction in media.
-The practical resources are that of coding and predictive models.[^12]
-
-[^12]: The problem of incomplete lists reminds me of the game 'I went shopping and I bought ...
-' that adults sometimes play with children.
-The game is about remembering a well-ordered list of items - 'a' is apple, 'b' is for banana, 'c' is for carrot.
-The point of the game is to use the well-ordered set of labels provided by the sequence of letters in the alphabet to remember a random and perhaps very diverse list of items.
-The longer the game continues, the harder it is to remember since the shopping list grows longer.
-At the same time, each player repeats the list, and this  helps memorisation.
-One trick in playing this game is to choose items that are quite unexpected or increasingly random.
-Rather than `ham`, list `halva`; rather than `fish`, add `face cream.` This is especially useful for adults in beating children.
-The children have better memories for items they know, but adults know more unexpected items.
-On the other hand, very unexpected items are quite easy to remember and tend to even the score in the favour of the children.
-Balancing between the improbable and memorable, between a predictable list like 'apples, bananas, carrots' and an unusual list such as 'agave syrup, boccadilla, cilantro' is complicated.
-Something similar applies to recommendation systems.
-The system can remember very many things.
-It may know everything you bought in the last few years on that site, and it can easily see what you would normally buy and add those as suggestions.
-
-A recommendation system should suggest items that you have not forgotten but may have never bought.
-If you do buy recommended items, then the recommender system has been effective in shunting the situation differently.
-It has led to a **'conversion.'** From a supermarket perspective, the conversion should be high in price.
-There is little value in successfully recommending a whole lot of small and cheap items such as rubber bands or toothpicks.
-So it might need to choose amongst different recommendations and weight high value items differently to low value items.
-In either case, 'conversion'  is the valued event.
-
-**How do we approach lists as objects of research?** While lists such as recommendations intersect strongly with other meaning and sense-making practices in everyday life and popular culture, they are also somewhat refractory to discourse  and textual analysis methods.
-On the one hand, lists are highly associative.
-They can be interpreted or decoded semiotically, but they also exhibit great variations in textuality that multiply semiotic readings.
-On the other hand, as operational inscriptions, they can be treated ethnomethodologically, as the production of ('the order of social settings through ...
-shared sense making practices.')[https://en.wikipedia.org/wiki/Ethnomethodology]
-
-List practices such as online grocery recommendations figure as sites of intersection between people, infrastructures, things, places and predictions.
-We might, therefore,  approach lists from various novel angles -- as asignifying semiotics as Maurizio Lazzarato calls it in his _Signs and Machines_ [@Lazzarato_2014] or as elements in 'a new order of spatio-temporal continuity for forms of economic, political and cultural life' as Celia Lury terms in her account of topological turn [@Lury_2012].
-In this paper, I focus on lists and list-making in terms of the predictive conditions of contemporary media experience [@Hansen_2015a].
-
-The question for me here is **how** contemporary lists can still be understood as produced through 'shared sense-making practices.' All lists are forms of order, since even random ordering is operationally important, but what counts as order might historically transformable, in which case the order of social settings or the problem of social order might be deeply historically contingent.
-The problem of order lies at the core of lists making, and our relation to lists can be seen as a test of the way that order operates more generally.
-
-In this context, the fact that the lists and indeed supermarkets more generally are driven by database logics suggests that working with list data could well entail interacting with databases.
-
-To this end, I will be seeking to simulate the workings of the recommendation system by generating queries, and exploring results from  a very large supermarket chain database.
-
-```{r simulate_basket, echo=FALSE, cache=TRUE, fig.margin=TRUE}
-library(dplyr)
-library(ggplot2)
-library(readr)
-
-csvs = list.files('data', pattern = '*.csv',full.names=TRUE)
-df = bind_rows(sapply(csvs, read_tsv))
-
-ggplot(df, aes(x=price,y=ContentsQuantity)) + geom_point() + theme(axis.text=element_text(size=10)) 
-df[df$price>=20,'name']
-```
-
-Much of my paper, and my responses to the presentation I saw from a data scientist  will focus on how the ordering that is done in a recommendation lists relates to the problem of social order more generally.
-The main narrative of the presentation concerned a shift from a set-based model for making lists to a predictive, probabilistic way of making lists in almost realtime and dealing with lists on a scale that is often difficult to grasp concretely.
-The change in list-making associated with learned models is increasingly widespread, and relies on a probabilistic treatment of people and things.
-
-I respond to the presentation in two ways.
-First, by recounting, contextualising and commenting on some of the main steps in making the 'Have your forgotten?' list  as a example of the sense-making done by data scientists and customer analytics teams in many places.
-Many of these steps were mentioned only briefly in the presentation, and they are largely taken for granted as part of predictive analytic practice.
-Second, by trying to step into the list-making practices  as an occasion to intervene and participate in the ontological conversion practices they engage in.
-This is no mere methodological problem if, as Mark Hansen has recently suggested,
-
-> predictive analytics are discoveries of micrological propensities that are not directly correlated with human understanding and affectivity and that do not by themselves cohere into clearly identifiable events [@Hansen_2015a, 111-112]
-
-The question exercising much contemporary work on data, on media, platforms and finance pivots on the 'micrological propensities' that might be re-ordering lists.
-Hansen's argues that these propensities make probabilities into something real, some actually in things:
-
->Whatever explanatory and causal value predictive analytics of large datasets have is, I suggest, ultimately rooted in this ontological transformation whereby probabilities are understood to be expressions of the actual propensity of things [@Hansen_2015a, 120]
-
-As the quote shows, he presents this as an 'ontological transformation.' The practical and indeed experimental question for me here is whether such transformations can be detected and articulated in the prosaic setting of shopping lists and  shopping baskets.
-I will suggest that 'conversion' might be a better way of mapping these changes than 'transformation.'
-
-## 1. The context of the presentation: 'statistics and operations research in industry'
-
-I heard this presentation at an academic event with speakers from industry and government on statistics and operations research in industry.
-Speakers came from Shell, Disney, GCHQ, and various universities.
-Because I've done a lot of work on machine learning and data science more generally, I found the sessions both familiar and occasionally exciting.
-
-![patel_notes](figure/patel_overall.png)
-
-Shreena Patel works as a data scientist for DunnHumby, a well-known [ customer science company ](https://www.dunnhumby.com/)  focusing on online grocery shopping at Tesco.
-I found her presentation on shopping lists to be  practical and evocative since it was focused quite a lot on implementation of recommendations, and symptomatic in various ways.
-
-![Tesco ClubCard](figure/tesco-clubcard.jpg )
-
-Tesco is the largest supermarket chain in UK, famous for Clubcard starting in 1991; often cited as a leading early example of Big Data Analytics. 
-DunnHumby -- Edwina Dunn and Clive Humby -- are famous for having convinced the CEO of Tesco sometime in 1991 that a loyalty card program could change the supermarket chain's relationship to its customers.
-[@Okeefe_1984]  Clive Humby's academic work is hard to track down, but an early paper of his given at the Conference of Your Operational Researchers in Nottingham in 1984 suggests the direction that he and later Tesco would take in constructing lists [@Okeefe_1984]
-
-![Humby, 1984]( figure/2_humby_young_org_1994.png)
-
-While Tesco has been very successful in data-mining its customers, the online shopping situation is a bit different.
-It is no longer organised around campaigns involving special offers or redemption of points, but on continuous up-selling and cross-selling.
-
-- the problems of scale: Tesco is like Amazon or Netflix in this respect: too many people and too many products
-
->Tesco is the clear winner in the online grocery market, in fact it takes almost 50p of every £1 spent on food shopping on the internet.
-(https://econsultancy.com/blog/64841-how-seo-helps-tesco-to-dominate-the-online-grocery-market/)[https://econsultancy.com/blog/64841-how-seo-helps-tesco-to-dominate-the-online-grocery-market/]
-
-## 2. The web and the shopping basket: how it was reconstituted in the 1990s
-
-![People add to list](figure/online_lists_in_practice.png)
-
-- online commerce -- transforming shopping; 
-- personalization and recommendation as the killer apps for social media as a commercial process, but also in government etc.
-
-How would I characterise this shift? Use it to emphasize why attention to the algorithm is not enough -- it is the model that counts
-A little history of recommender systems -- uses Varian 1997 for a start. But has no real reference list. Not sure  
-Also can use Schafer -- back to Association Rule Mining from early 1990s
-
-When commentators first began talking about recommender systems in the 1990s, they highlighted the shift from demographic-driven market research techniques to personalized recommendations.
-For instance, writing in 1997, Paul Resnick and Hal Varian (at that time Dean of Information Sciences at UC Berkeley, but somehow now Chief Economist at Google), made much of this difference [@Resnick_1997].
-
-## 3. From the rules-based list to a 'relevancy model'
-
-Patel described her work at DunnHumby in terms of a shift in models from a 'rules-based list' to a 'relevancy model.' This was a significant change in how the 'Have You Forgotten' list was constructed from one kind of model to another.
-For me, the shift in models matters since it amounts to a much more probabilistic structuring of lists.
-
-I assume that the rules-based model derived from the early 1990s work on the `apriori` algorithm by Rakesh Agrawal and Ramakrishnan Srikanti at IBM Research Alameda [@Agrawal_1994].
-The `apriori` algorithm finds sets of items that commonly occur in transactions.
-These commonly occurring sets are expressed as association rules.
-For instance, using the `Groceries` dataset in the `R` package `arules,` we can readily see how Walmart might have discovered  -- but didn't -- the fabled association between diapers and beer.
-[http://www.dssresources.com/newsletters/66.php]
-The `apriori` algorithm is widely used in 'market basket analysis' and attempts to re-structure shopping lists in terms of associations between items that are frequently bought together.
-
-```{r arules, echo=TRUE, fig.margin=TRUE}
-library(arules)
-library(arulesViz)
-library(datasets)
-data(Groceries)
-itemFrequencyPlot(Groceries,topN=20,type="absolute")
-# Get the rules
-rules <- apriori(Groceries, parameter = list(supp = 0.001, conf = 0.8))
-
-# Show the top 5 rules, but only 2 digits
-options(digits=2)
-inspect(rules[1:5])
-summary(rules)
-
-rules<-sort(rules, by="confidence", decreasing=TRUE)
-inspect(rules[1:5])
-rules<-apriori(data=Groceries, parameter=list(supp=0.001,conf = 0.15,minlen=2), 
-           appearance = list(default="rhs",lhs="whole milk"),
-           control = list(verbose=F))
-rules<-sort(rules, decreasing=TRUE,by="confidence")
-inspect(rules[1:5])  
-items = Groceries@itemInfo
-```
-
-The association rule-based model for the `Groceries` dataset suggests some of the characteristic ordering practices of data-mining approaches to lists. 
-The association between items is expressed in the form of _rules_.
-These rules, a term derived from decision support literature, define sets of associations between items.
-As a set theory-based practice, the association rules define likely sets of items or shopping baskets in terms of decreasing probability.
-`r inspect(rules[1])` is a stronger association than `r inspect(rules[2])`.
-So any recommendation in the 'Have You Forgotten' list should probably put them in that order.
-
-Association rules and the `apriori` algorithm present social order in terms of sets of common associations.
-This means that that the algorithm can be understood in terms of the set-based multiples proposed by Alain Badiou:
-
-> Multiplicity will be called  ‘closed’ when grasped by the numerical one, and ‘open’ when grasped by the vital One.
-Every multiplicity is the joint effectuation of the closed and open, but its ‘veritable’ multiple-being lies on the side of the open.
-[@Badiou_2004, 72]
-
-Badiou in his account of multiples highlight the contrast between two different ways of grasping a multiple -- in terms of the 'numerical one' and the 'vital One' [@Badiou_2004, 72], between the 'closed' and the 'open' respectively.
-
-The `apriori` algorithm, might be seen as an attempt to deal with the excess of things in a Tesco supermarket, and the lack of interest or engagement that people have with them by identifying subsets whose membership is stronger by virtue of the other sets of associations that people bring with them.
-Association is stronger when it derives from an excess of different ways of being in sets.
-
-Badiou's argument is that being part of a set differs from belonging to a set.
-An open set arises from when we dwell within a set by 'taking the intersection of two elements.' What might this mean in terms of ordering?
-
-If we take the itemset from the `Groceries` dataset [@Hahsler_2006], and then run it through the Tesco online groceries site, each of the 169 items in the dataset yields dozens and sometimes thousands of results from Tesco online.
-
-```{r tesco, cache=TRUE, fig.cap ='Tesco surplus',  fig.margin=TRUE}
-    library(ggplot2)
-    tesco = read.csv('data/reference_data/tesco_groceries.csv')
-    total = sum(tesco$actual, na.rm=TRUE) 
-    tesco = na.omit(tesco)
-    tesco$actual = as.integer(tesco$actual)
-
-    ggplot(tesco[tesco$actual>50,], aes(x=labels, y=actual, fill=labels)) +geom_bar(stat='identity') + coord_flip() + ggtitle('Tesco grocery items with more than 50 products') + theme(legend.position="none", axis.text=element_text(size=9))
-```
-
-Each of the items in the `Groceries` dataset proliferates in the Tesco site, into a new list of branded variations.
-The 169 items of the `Groceries` dataset, itself from a local German or Austrian supermarket [@REF_TBA],  expand into roughly `r total` possible items.
-
-## 4. Hack? Customers go straight to the checkout
-
-Patel observed that the 'Have you forgotten' lists matter to people, and we can see why in the light of this proliferation of things in the list.
-A `r total` list of items is quite hard to list let alone read.
-Patel said that some people log on and go straight to the checkout, allowing the recommendation system to supply them with the suggestions that they then move to the shopping basket.
-This is the online equivalent of using an old shopping list and only adding new items when needed. 
-It is also a strategy to avoid having to view too many lists, since the only way Tesco can show people what they might buy is by showing them lists of things (e.g.
-chicken products), but these lists are often too long to work with.
-For instance, Tesco has more than 1000 chicken products on its site.
-People inhabit lists and work with them no matter how they are made; thus vindicating ethnomethodologists assertion that social order is the product of ongoing sense-making practices. 
-But it also suggests a transformation in the way that lists exist in the world.
-They are less the product of a discrete agent than an effect of an environment teeming with possible combinations and complementaries, a surplus of possible purchases that the supermarket needs to continually maintain but also filter in order to make it possible for something to happen.
-
-## 5. Data: 52 weeks of data for each customer: how is order constructed -- from history?
-
-Patel described the data they on grocery purchases. 'We have 52 weeks of purchase data for each customer.' 
-We know that the construction of lists has been transformed by recommendation systems.
-These transformations have been widely discussed in the context of big data, where questions of who has how much data are usually most prominent.
-In terms of the ways that lists order the world, the amount of data is not so amount as the way in which things in the world take on a list-like existence. 
-I want to dwell on this idea for a while. We have seen two different ways in which lists might be constructed in a recommender system. 
-The first uses history and existence of previous lists, including the customer's own list, to populate a new list. 
-The second uses many shopping lists to identify associations between things and then uses the strongest associations to recommend additions to a current list. 
-The 'relevancy model,' however, tries to do something different. 
-The problem that the proliferation of things on the shelves of supermarket or grocery warehouse confronts is partly combinatorial.
-With  `r total` products (actually Patel mentions 200,000 products) can be combined in truly astronomical numbers of combinations.
-If a typical shopping list has 20 items, then there are `r format(choose(total, 20), 2)` possible combinations.
-Some of this vast number of possible shopping lists is highly unlikely in terms of shopping lists and histories of shopping, but potentially highly interesting and value as recommendation.
-
-> This means, to put it slightly differently, that surrounding any delimited predictive system is a larger field of data -- what I elsewhere call a “surplus of sensibility” -- that, viewed speculatively, indexes the causal efficacy of the total situation within which this delimited system operates.
-(Effectively, the latter gains its reliability from closing off this larger surplus of sensibility, thereby transforming an always excessive propensity into a [provisionally] closed dataset.) Because it affords data that exceed whatever any given predictive system might include, the data of the world’s causal efficacy -- the data constituting its real potentiality -- always and in principle facilitates knowledge that cannot be restricted to any particular agenda [@Hansen_2015, 126-7]
-
-In some ways, Hansen's account resembles the Badiouean ontology of the multiple in which a closed set is surrounded and conditioned by an open set that is full of many more potentials.
-However, if we think of the  new model that Patel describes as the delimited (or delimiting?) system within   the larger surplus of sensibility, how that help us understand what the 'Have you forgotten?' recommendations do?
-
-- history of data accumulations;
-
-- shift from data mining to personalization? 
-- use some of the Varian, Linden, Resnick, Schaffer literature
-
-## 6. Gain: use good features to build models
-
-- what counts as a good feature? See Domingos on this
-- move from Dunnhumby's ClubCard focus to much more general world of predictive models (e.g. Facebook use a million models, etc.)
-- the model counts for surprisingly little -- at least, they all seem similar ...
-rather what counts is all that stuff on working with substitutions, similiarities and complementarities for each person.
-
-It was surprising that Patel spoke relatively briefly about the new models.
-She showed a graph that compared the old rule-based model with some of the new alternatives -- logistic regression, random forests, and a few others.
-She dismissed most of the models quite quickly and focused only on one, the logistic regression model.
-Or more accurately, she just said that logistic regression did better than other, and there was little more to discuss on that front.
-The measure of doing better here is called precision, and it measures the proportion of the recommended products that are actually purchased.
-
-Why did she say little about the model, which after all, produces the probabilistic predictions on which everything in the recommendation list depends and hence the conversion experience?
-
-![Models](figure/patel_recommendation1.png)
-
-The graph compares models in terms of their predictions and how those predictions came to pass.
-The two columns on the left show earlier models -- the global model, which is not personalized, the association rules model for individual customers, and then the new models (logistic regression, random forest, etc) for customers.
-An increase in precision means more sales.
-Her interest lay in the relevance of the recommendations, and in particular, in how 'most of the gain comes from thinking of good features.'
-
-The fact that so little was said about the models themselves suggests that they are now largely taken for granted in recommendation systems.
-The models she mentioned are all classifiers, predictive models that treat a set of variables in a datasets as a vector space in which good classification means finding a way of defining a decision boundary between points that belong to different classes of things.
-Here the classes of things are binary: `recommended` or `not recommended.` Some classifier models, including logistic regression, generate a probability rather than simply a classification.
-
-## 7. Sparsity: output 20 recommendations from 200,000 products
-
-What counts as a good feature in a world of predictive lists?
-Some things in the data that increases the precision of the model.
-The problem that Tesco faces, however, is that there are too many possible recommendations to be tested.
-They sell around 200,000 products and no customer wants to buy any more than a fraction of a per cent of those products.
-The recommendation list needs to very short.
-
-Predictive models, we might say, encounter their own surplus of sensibility in the data as they search for the most accurate predictions.
-A  very brief comment from Patel -- 'we have lots of zeros' -- points to the problem here.
-Each of the millions of Tesco online grocery shoppers has bought some selection of the 200,000 products.
-But the product-customer matrix, the basic vector-space in which all recommendation systems operate, remain very sparse and unpopulated.
-
-Say Tesco has 1 million online customers.
-A data matrix for just whether they ever bought a particular product or not will be `r format(1e6 * 2e5)` in size.
-Given that any one customer is like to only have bought 100 different products, the data matrix will be `r format(100-(100/2e5)*100, 2)`% empty.
-
-```{r sparsity, echo=FALSE, cache=TRUE, fig.margin=TRUE, fig.cap ='The sparsity of data' }
-library(ggplot2)
-library(reshape2)
-p = as.logical(rbinom(n=2e5, size=1, prob=0.001))
-head(p, 30)
-m = (matrix(p, nrow=1000, byrow=TRUE))
-
-melted <- melt(m)
-melted$value <- as.logical(melted$value)
-ggplot(melted, aes(x = Var2, y = Var1, fill = value)) + geom_tile() +
-    scale_fill_manual(values = c("grey", "black"))
-
-```
-
-Having too many things to list -- this must be happening in many settings -- videos, social networks, products, etc.
--- is not the real problem.
-The problem is that 'massively unbalanced data,' which means that different classes of products are not equally represented.
-Since so many people buy milk, the recommendation system might end up always recommending milk products.
-So the data needs to be 'corrected' by, in this case, random under-sampling or removal of some of the data for common purchases.
-Secondly, even if a recommendation system finds uncommon items that are good recommendations, they should neither be too cheap ('low spend'), nor, as we will see too similar to what the customer has already ordered.
-
-In Hansen's or Badiou's terms, this is one way in which the dataset is closed.
-While 200,000 products is still a closed dataset since it has a defined number of items, undersampling alters the frequency of occurrence of very common items so that the predictive models are not distracted by them.
-
-So the raw data -- the 52 weeks of purchases for each customer -- can be aggregated to increase the density of the data in ways that helps the model.
-Patel mentions some of the 'features' used in the new model but somewhat strangely at this point, her presentation turns toward the difficulties of measuring the success of the model and the infrastructural limitations on making recommendations in ways that are more alive to what is happening right now.
-
-## 8. Value: Balancing conversion vs spend
-The value of an ordering
-
-## 9. Metric: precision-at-K
-
-The discussion of imperfection was brief - it incorporated the notion of recommendations weighted by price and the notion of precision (or the proportion of products actually purchased.).
-Patel simply said that this metric was imperfect, but without elaborating why.
-Instead she moved to describe some of the ways in which the recommendation system worked in practice.
-
-## 10. Experimental settings: A/B allocation and uplift
-
-![A/B testing](figure/a_b_tests.png)
-
-First of all, the new model was deployed experimentally in a random A/B controlled trial.
-This means that all customers are allocated to one of four categories as shown in the figure.
-For an identical shopping list, or even for the same person, it is possible to receive different recommendations.
-
-This randomising of the application of the model, drawing on protocols for randomised clinical trials first developed in the 1960s, is widely used in social media platforms and hence in all recommender systems.
-It is difficult to gauge how much.
-Facebook reports that it has more than 1 million models operation in its infrastructure.
-[TBA: ref for FBflow]   We could also discuss the integration of experiments into the operational infrastructures at Google.[REF tba]
-
-Random allocation of customers to the four categories (Test A, Test B, Control A, Control B) adds another layer of probability to the recommendation system.
-This is done in the name of statistical validation of the effects or the uplift of the model.
-It also suggests that the effects of the model cannot be known in advance -- they can only be observed experimentally.
-Does this experimentality of the list point more to the propensities that Hansen says saturate experience?
-
-It is perhaps significant that this random allocation of customers to control and test groups occurs without any relation to the particular profile or propensities of the customer.
-It seeks to statistically validate the effects -- the *uplift* -- of the model on  conversion events by allowing the effects of the model on what people do to be measured.
-The lift or uplift refers to a change in events associated with the same groups of people.
-This effectively is an experiment in inhabiting two different worlds.
-This sets up a feedback look between the predictive model, the logistic regression, and the operation of the recommendation system.[^122] Once the experimental uplift methodology is in place, the recommendation system can constantly modify the recommendation list in light of subsequent data on the uplift.
-Predictions, then, are not bound by fixed parameters of the model.
-
-[^122]: See [@Lo_2002] for an introduction to the construction of experimental set ups for uplift models.
-Market researchers and data miners seem to have started adopting a statistically grounded experimental predictive practice sometimes around 2000.
-See for instance [@Almquist_2001]  Response modelling, propensity modelling and uplift modelling all seek to identify associations between 'treatments' or interventions and the 'Responders', the people  affected by the treatment.
-Interestingly, these
-
-## 11. Infrastructure: realtime engine serves up top 5
-
-![Realtime infrastructure and its limits](figure/hadoop.png)
-
-While much discussion of recent media has highlighted its realtime operation, Patel's description of the infrastructure that implements the new model suggested a different temporal patterning.
-While raw data from Tesco Online feeds into the model every hour, it seems that recommendation list for each customer is only generated once a week.
-This makes sense.
-Customers only shop online every few days at the most, and in some cases, only every few weeks.
-To continually update the top 200 recommendations for every customer would be computationally expensive.
-
-Patel did not say much else about the infrastructure but the 200 recommendations generated each week are certainly never seen by the customers.
-The realtime engine that actually generates results shown on the Tesco Online website only shows 5 after 'applying business rules.' Again, this might be seen as corroborating Hansen and Badiou's insistence on the necessarily limited grasp of any prediction system on the situation it encounters.
-Even the new model, situated amidst its experimental setup cannot fully encompass the surplus of sensibility it encounters.
-
-The brief mention of specific infrastructural elements such as `hadoop` is, however, significant.
-It points to the temporality of infrastructures themselves as a component of list practices.
-The possibility of adjusting the recommendations for every customer weekly derives from an infrastructure capable of collecting data, assimilating that data to predictive model able to generate recommendations that are relevant to individuals amongst 200,000 products.
-
-## 12. Similarity, substitution, complementarity and micrological propensities
-
-![Basket similarity](figure/patel_basket_similarity.png)
-
-What does it matter that recommendations are relevant?
-Relevance is closely connected to propensities in the predictive condition.
-From DunnHumby's standpoint, the relevance of a recommendation relates directly to conversion.
-From the standpoint of the ontological conversion, and  if Hansen is right, the real or actual fabric of prediction resides in the propensities of things, what he calls the 'complex and diffuse calculus' that only subsequently and provisionally congeals as a conversion event.
-The 'micrological propensities' of things feed directly into an sense of the relevance of recommendations.
-
-How does the DunnHumby model deal with the micrological propensities of things?
-In her presentation, Patel  spent quite a lot of time addressing the problem of basket similarity in terms of substitutions and complementarities between things.
-She concludes by saying 'basket similarity is the future' and 'complementarities drive conversion.'
-
-Whereas the association rules focus on complementarities -- if you buy bread, you might want milk, if you buy beer, you might want diapers -- the substitutions look for micrological propensities that relate to gradations of preference, and variations in time and place.
-Micrological propensities in the list relate to a particular time and place, amidst a continuously differentiable pathways between things.
-
-How do things associate with each other in  substitutions and complementarities?
-Patel's comments on this are very brief but suggest that the 'Have You Forgotten?' list becomes an increasingly mutable entity, open to many substitutions and re-writings.
-Like the mention of `hadoop`, what Patel said about substitution and complementarity implies extensive but relatively well-established data practices.
-She speaks for instance of a 'design matrix of 14,000 columns' and the use of `L1 regularization` to drive coefficients down to zero.
-This is machine learning talk.
-What is being modelled here?
-No longer whether to recommend a product to a customer or not, but whether something could be substituted for what is already in the recommendation list.
-
-```{r biscuits, echo=FALSE, cache=TRUE, fig.margin=TRUE}
-    library(readr)
-    library(knitr)
-    bisc = read_tsv('data/biscuits.csv')
-    kable(bisc[order(bisc$price, decreasing=TRUE)[1:30], c('name','price')])
-```
-
-We can imagine this in terms of biscuits.
-A customer's recommendation list might already include `r bisc$name[10]`.
-The 'self-learning substitution model' might be able to substitute something in that list that is of higher value, either because it is bigger or more expensive.
-
-## 13. Complementarity drives conversion
-
-![propensity_matrix](figure/propensity_conversions.png)
-
-Similarly, to model the potential complementarities   between things entails a matrix of comparisons.
-`Nutella` complements `bananas,` but the converse is not necessarily true since `bananas` need not complement `nutella.`
-
-How should we understand the effect of complementarity?
-Patel says that 'complementarity drives conversion.' That is, the matrix of similarities expressed as numerical associations between products best animates the conversion of recommendations into purchases.
-
-In the similarity matrix, shaped by the history of past recommendations and purchases made by this customer, lie the propensities of things that Hansen (following Popper) attributes to the world.
-Karl Popper speaks of the future as a 'crystallization of propensities' [@Popper_1990,18]
-
-## Conclusion
-
-My investigation of lists has been centred around the transformation in their predictive underpinnings.
-How should we think about the list-making practices of these models, databases and computational infrastructures?
-Following Hansen, I have been suggesting that the predictive listing in the form of open-ended recommendation is symptomatic of changes in the distribution of media.
-The idea that media should no longer be considered as extensions or prostheses of human communication but the environmental situation in which different propensities play out is increasingly central to contemporary media theory.
-
-Hansen refers to an ontological transformation in which predictive media or recommendation systems would be seen as the closed realisation of the actually open propensities of the world.
-By understanding the world in terms of propensities, he argues, we have a change to engage with contemporary predictive media.
-
-I have not explored the Whitehead-derived philosophy of propensities in the world that he develops there, only because I am not sure that we need speculative theory to grapple with the material practices of prediction.
-Already in the DunnHumby model and the Tesco 'Have You Forgotten?' recommendations, many of the traits of predictive processes are at work.
-
-I diverge from Hansen in how I account for this change. Rather than looking for a speculative theory that sets out a world in which prediction would be the closed form,
-
-- listing practices are very closely entangled with problems of social order under predictive conditions or amidst predictive operations. 
-- lists have complex temporal-material structure, since they mix prediction, infrastructural constraints, the individual characteristics of list-makers, statistical validations, database architectures and above all the surplus of sensibility associated with commodity production.
-- lists undergoing a 'conversion' experience as they become more probabilistic, but their heterogeneous character as operations caught up in models, infrastructures, experiments, and the flow of commodities means that they can never form a closed set or dataset.
-Amidst prediction they become constitutively incomplete.
-
-## References

### 2016-10-21 lists_conversion_experience.rmd
fixing up formatting and adding new figures

@@
 opts_chunk$set(error=TRUE)
+opts_chunk$set(cache=TRUE)
 ```
@@
 This paper investigates practices of list-making associated with online grocery shopping at a  UK and global supermarket chain, Tesco. 
-It will analyse how a recommendation system suggests a small number of grocery items of personal relevance to each of the millions of online grocery shoppers. 
+It will analyse how a recommender system suggests a small number of grocery items of personal relevance to each of the millions of online grocery shoppers. 
 The predictive modelling practices aim to up-sell and cross-sell some of the 250,000 items in the supermarket's inventory. 
-I find it interesting to approach the list-making practices  of the grocery recommendation system as a mundane case of the predictive intermediaries operating in many places. 
-The transformation or conversion of listing into predicting can be analysed in recommendation systems that extend existing lists with potential new items, thus transforming the temporal and agential fabric of the list.
-While all written lists mix past and future actions, actual and potential events, their predictive re-writing in recommendation systems adds a new operational dynamic.
+I find it interesting to approach the list-making practices  of the grocery recommender system as a mundane case of the predictive intermediaries operating in many places. 
+The transformation or conversion of listing into predicting can be analysed in recommender systems that extend existing lists with potential new items, thus transforming the temporal and agential fabric of the list.
+While all written lists mix past and future actions, actual and potential events, their predictive re-writing in recommender systems adds a new operational dynamic.
 In this operational dynamic, the incompleteness of lists, their propensity to expand or change, becomes more important than their capacity to exhaustively enumerate things. 
-The conversion of lists from finite to in-definite in recommendation systems suggests we might need to re-conceptualise how we think about social order and structure in media cultures.
+The conversion of lists from finite to in-definite in recommender systems suggests we might need to re-conceptualise how we think about social order and structure in media cultures.
 
@@
 While the title 'have you forgotten' implies that you will be shown items that could have been or are usually on a shopping list, the working assumption is a temporal one -- the list is incomplete either because you have forgotten to list something, or you didn't know what should be on your list, perhaps because you have never even thought of putting that thing on your list.
+- making some changes, including a slide of early recommendation systems from 1990s such as described in ~/archive/data_intensive/machine_learning/schafer_recommenders_ecommerce_2001.pdf
+-
 
@@
 Balancing between the improbable and memorable, between a predictable list like 'apples, bananas, carrots' and an unusual list such as 'agave syrup, boccadilla, cilantro' is complicated.
-Something similar applies to recommendation systems.
+Something similar applies to recommender systems.
 The system can remember very many things.
@@
 
-To this end, I will be seeking to simulate the workings of the recommendation system by generating queries, and exploring results from  a very large supermarket chain database.
+To this end, I will be seeking to simulate the workings of the recommender system by generating queries, and exploring results from  a very large supermarket chain database.
 
-```{r simulate_basket, echo=FALSE, cache=TRUE, fig.margin=TRUE}
+```{r simulate_basket, echo=FALSE, cache=TRUE, fig.margin=TRUE, fig.cap='Prices and quantities of things in the world'}
 library(dplyr)
@@
 library(readr)
+library(knitr)
 
@@
 ggplot(df, aes(x=price,y=ContentsQuantity)) + geom_point() + theme(axis.text=element_text(size=10)) 
-df[df$price>=20,'name']
+kable(df[df$price>=20,'name'])
 ```
@@
 How would I characterise this shift? Use it to emphasize why attention to the algorithm is not enough -- it is the model that counts
-A little history of recommender systems -- uses Varian 1997 for a start. But has no real reference list. Not sure  
+A little history of recommender systems -- uses Varian 1997 for a start. But has no real reference list. 
 Also can use Schafer -- back to Association Rule Mining from early 1990s
@@
 For instance, writing in 1997, Paul Resnick and Hal Varian (at that time Dean of Information Sciences at UC Berkeley, but somehow now Chief Economist at Google), made much of this difference [@Resnick_1997].
+
+![Early e-commerce recommender system architecture](figure/schafer_recommendation_2001.png)
+
+![The advent of recommender systems](figure/resnick_varian_page1_197.png)
 
@@
 
-```{r arules, echo=TRUE, fig.margin=TRUE}
+```{r arules, cache=TRUE, echo=TRUE, fig.margin=TRUE}
+library(knitr)
 library(arules)
@@
 itemFrequencyPlot(Groceries,topN=20,type="absolute")
-# Get the rules
 rules <- apriori(Groceries, parameter = list(supp = 0.001, conf = 0.8))
 
-# Show the top 5 rules, but only 2 digits
 options(digits=2)
-inspect(rules[1:5])
-summary(rules)
-
 rules<-sort(rules, by="confidence", decreasing=TRUE)
-inspect(rules[1:5])
 rules<-apriori(data=Groceries, parameter=list(supp=0.001,conf = 0.15,minlen=2), 
@@
 rules<-sort(rules, decreasing=TRUE,by="confidence")
-inspect(rules[1:5])  
+kable(inspect(rules[1:5]))
 items = Groceries@itemInfo
@@
 
-The association rule-based model for the `Groceries` dataset suggests some of the characteristic ordering practices of data-mining approaches to lists. 
+The association rule-based model for the `Groceries` dataset suggests some of the characteristic ordering practices of 1990s-style data-mining approaches to lists. 
 The association between items is expressed in the form of _rules_.
@@
 
-Association rules and the `apriori` algorithm present social order in terms of sets of common associations.
+Association rules and the `apriori` algorithm present social order in terms of sets of common associations that are already weighted by frequency and hence probability.
 This means that that the algorithm can be understood in terms of the set-based multiples proposed by Alain Badiou:
@@
 A `r total` list of items is quite hard to list let alone read.
-Patel said that some people log on and go straight to the checkout, allowing the recommendation system to supply them with the suggestions that they then move to the shopping basket.
+Patel said that some people log on and go straight to the checkout, allowing the recommender system to supply them with the suggestions that they then move to the shopping basket.
 This is the online equivalent of using an old shopping list and only adding new items when needed. 
@@
 Patel described the data they on grocery purchases. 'We have 52 weeks of purchase data for each customer.' 
-We know that the construction of lists has been transformed by recommendation systems.
+We know that the construction of lists has been transformed by recommender systems.
 These transformations have been widely discussed in the context of big data, where questions of who has how much data are usually most prominent.
@@
 
-The fact that so little was said about the models themselves suggests that they are now largely taken for granted in recommendation systems.
+The fact that so little was said about the models themselves suggests that they are now largely taken for granted in recommender systems.
 The models she mentioned are all classifiers, predictive models that treat a set of variables in a datasets as a vector space in which good classification means finding a way of defining a decision boundary between points that belong to different classes of things.
@@
 Each of the millions of Tesco online grocery shoppers has bought some selection of the 200,000 products.
-But the product-customer matrix, the basic vector-space in which all recommendation systems operate, remain very sparse and unpopulated.
+But the product-customer matrix, the basic vector-space in which all recommender systems operate, remain very sparse and unpopulated.
 
@@
 Patel simply said that this metric was imperfect, but without elaborating why.
-Instead she moved to describe some of the ways in which the recommendation system worked in practice.
+Instead she moved to describe some of the ways in which the recommender system worked in practice.
 
@@
 
-Random allocation of customers to the four categories (Test A, Test B, Control A, Control B) adds another layer of probability to the recommendation system.
+Random allocation of customers to the four categories (Test A, Test B, Control A, Control B) adds another layer of probability to the recommender system.
 This is done in the name of statistical validation of the effects or the uplift of the model.
@@
 This effectively is an experiment in inhabiting two different worlds.
-This sets up a feedback look between the predictive model, the logistic regression, and the operation of the recommendation system.[^122] Once the experimental uplift methodology is in place, the recommendation system can constantly modify the recommendation list in light of subsequent data on the uplift.
+This sets up a feedback look between the predictive model, the logistic regression, and the operation of the recommendation system.[^122] Once the experimental uplift methodology is in place, the recommender system can constantly modify the recommendation list in light of subsequent data on the uplift.
 Predictions, then, are not bound by fixed parameters of the model.
@@
 
-Hansen refers to an ontological transformation in which predictive media or recommendation systems would be seen as the closed realisation of the actually open propensities of the world.
+Hansen refers to an ontological transformation in which predictive media or recommender systems would be seen as the closed realisation of the actually open propensities of the world.
 By understanding the world in terms of propensities, he argues, we have a change to engage with contemporary predictive media.

### 2016-10-24 lists_conversion_experience.rmd
revising manuscript

@@
 
-**How do we approach lists as objects of research?** While lists such as recommendations intersect strongly with other meaning and sense-making practices in everyday life and popular culture, they are also somewhat refractory to discourse  and textual analysis methods.
+## How do we approach lists as objects of research?
+
+While lists such as recommendations intersect strongly with other meaning and sense-making practices in everyday life and popular culture, they are also somewhat refractory to discourse  and textual analysis methods.
 On the one hand, lists are highly associative.
@@
 
-To this end, I will be seeking to simulate the workings of the recommender system by generating queries, and exploring results from  a very large supermarket chain database.
+This is no mere methodological problem if, as Mark Hansen has recently suggested,
+
+> predictive analytics are discoveries of micrological propensities that are not directly correlated with human understanding and affectivity and that do not by themselves cohere into clearly identifiable events [@Hansen_2015a, 111-112]
+
+The question exercising much contemporary work on data, on media, platforms and finance pivots on the 'micrological propensities' that might be re-ordering lists.
+Hansen's argues that these propensities make probabilities into something real, some actually in things:
+
+>Whatever explanatory and causal value predictive analytics of large datasets have is, I suggest, ultimately rooted in this ontological transformation whereby probabilities are understood to be expressions of the actual propensity of things [@Hansen_2015a, 120]
+
+As the quote shows, he presents this as an 'ontological transformation.' The practical and indeed experimental question for me here is whether such transformations can be detected and articulated in the prosaic setting of shopping lists and  shopping baskets.
+I will suggest that 'conversion' might be a better way of mapping these changes than 'transformation.'
+
+## 1. The context of the presentation: 'statistics and operations research in industry'
+
+I heard this presentation at an academic event with speakers from industry and government on statistics and operations research in industry.
+Speakers came from Shell, Disney, GCHQ, and various universities.
+Because I've done a lot of work on machine learning and data science more generally, I found the sessions both familiar and occasionally exciting.
+
+![patel_notes](figure/patel_overall.png)
+
+Shreena Patel works as a data scientist for DunnHumby, a well-known [ customer science company ](https://www.dunnhumby.com/)  focusing on online grocery shopping at Tesco.
+I found her presentation on shopping lists to be  practical and evocative since it was focused quite a lot on implementation of recommendations, and symptomatic in various ways.
+It was filled with graphs, numbers, and tables concerning different aspects of the 'have you forgotten?' list.
+Understandably, this was not presented in a language of ontological transformation.
+
+To this end, I will be seeking to reconstruct some of the workings of the recommender system by generating queries, and exploring results from  a very large supermarket chain database.
 
@@
 Second, by trying to step into the list-making practices  as an occasion to intervene and participate in the ontological conversion practices they engage in.
-This is no mere methodological problem if, as Mark Hansen has recently suggested,
 
-> predictive analytics are discoveries of micrological propensities that are not directly correlated with human understanding and affectivity and that do not by themselves cohere into clearly identifiable events [@Hansen_2015a, 111-112]
+## 2. The web and the shopping basket: how it was reconstituted in the 1990s
 
-The question exercising much contemporary work on data, on media, platforms and finance pivots on the 'micrological propensities' that might be re-ordering lists.
-Hansen's argues that these propensities make probabilities into something real, some actually in things:
-
->Whatever explanatory and causal value predictive analytics of large datasets have is, I suggest, ultimately rooted in this ontological transformation whereby probabilities are understood to be expressions of the actual propensity of things [@Hansen_2015a, 120]
-
-As the quote shows, he presents this as an 'ontological transformation.' The practical and indeed experimental question for me here is whether such transformations can be detected and articulated in the prosaic setting of shopping lists and  shopping baskets.
-I will suggest that 'conversion' might be a better way of mapping these changes than 'transformation.'
-
-## 1. The context of the presentation: 'statistics and operations research in industry'
-
-I heard this presentation at an academic event with speakers from industry and government on statistics and operations research in industry.
-Speakers came from Shell, Disney, GCHQ, and various universities.
-Because I've done a lot of work on machine learning and data science more generally, I found the sessions both familiar and occasionally exciting.
-
-![patel_notes](figure/patel_overall.png)
-
-Shreena Patel works as a data scientist for DunnHumby, a well-known [ customer science company ](https://www.dunnhumby.com/)  focusing on online grocery shopping at Tesco.
-I found her presentation on shopping lists to be  practical and evocative since it was focused quite a lot on implementation of recommendations, and symptomatic in various ways.
-
-![Tesco ClubCard](figure/tesco-clubcard.jpg )
+!['Every little bit helps':Tesco ClubCard](figure/tesco-clubcard.jpg )
 
@@
 
-While Tesco has been very successful in data-mining its customers, the online shopping situation is a bit different.
-It is no longer organised around campaigns involving special offers or redemption of points, but on continuous up-selling and cross-selling.
-
-- the problems of scale: Tesco is like Amazon or Netflix in this respect: too many people and too many products
+While Tesco has been very successful in data-mining its customers,  and perhaps became the UK's biggest supermarket through data-mining in the 1990s, the online shopping situation is a bit different.
+It is no longer organised around campaigns involving special offers or redemption of points, but on continuous up-selling and cross-selling to customers who might be anywhere, but especially are not in the supermarket itself.
+.
 
 >Tesco is the clear winner in the online grocery market, in fact it takes almost 50p of every £1 spent on food shopping on the internet.
-(https://econsultancy.com/blog/64841-how-seo-helps-tesco-to-dominate-the-online-grocery-market/)[https://econsultancy.com/blog/64841-how-seo-helps-tesco-to-dominate-the-online-grocery-market/]
-
-## 2. The web and the shopping basket: how it was reconstituted in the 1990s
+[ https://econsultancy.com/blog/64841-how-seo-helps-tesco-to-dominate-the-online-grocery-market/ ]( https://econsultancy.com/blog/64841-how-seo-helps-tesco-to-dominate-the-online-grocery-market/ )
 
@@
 
-- online commerce -- transforming shopping; 
-- personalization and recommendation as the killer apps for social media as a commercial process, but also in government etc.
+How should characterise this shift from ClubCard membership data mining to online grocery markets?
+Patel described her work at DunnHumby in terms of a shift in models from a 'rules-based list' to a 'relevancy model.' This was a significant change in how the 'Have You Forgotten' list was constructed from one kind of model to another.
+For me, this shift is a part of broader re-organisation of prediction in which existing predictive practices are being re-distributed and intensified in particular ways.
+The shift in models matters since it amounts to a much more probabilistic structuring of lists.
 
-How would I characterise this shift? Use it to emphasize why attention to the algorithm is not enough -- it is the model that counts
-A little history of recommender systems -- uses Varian 1997 for a start. But has no real reference list. 
-Also can use Schafer -- back to Association Rule Mining from early 1990s
+When commentators first began talking about recommender systems in the 1990s, they highlighted the shift from demographic-driven market research or data mining techniques to personalized recommendations.
+For instance, writing in 1997 in a special section of the _Communications of the ACM_ on recommender systems, Paul Resnick and Hal Varian (at that time Dean of Information Sciences at UC Berkeley, but somehow now Chief Economist at Google), made much of this difference [@Resnick_1997].
+They emphasise the need for a term to distinguish the emerging practices from data mining.
+[^44]
 
-When commentators first began talking about recommender systems in the 1990s, they highlighted the shift from demographic-driven market research techniques to personalized recommendations.
-For instance, writing in 1997, Paul Resnick and Hal Varian (at that time Dean of Information Sciences at UC Berkeley, but somehow now Chief Economist at Google), made much of this difference [@Resnick_1997].
+[^44]: Also can use Schafer -- back to Association Rule Mining from early 1990s
 
@@
 
+> In everyday life, we rely on recommendations from other people.
+...
+Recommender systems augment this natural social process.
+In a typical recommender  system, people provide recommendations as inputs, which the system then aggregates and directs to appropriate recipients.
+In some cases the primary transformation is in the aggregation; in others the system’s value lies in its ability to make good matches between the recommenders and those seeking recommendations.[@Resnick_1997,56]
+
+Writing near the advent of early web-based e-Commerce, Resnick and Varian (now Chief Economist at Google) imagined recommender systems as automations of the 'collaborative filtering' that people do when they write and read reviews of products and services (for instance, on the travel website, TripAdvisor; [@REF_Scott_TBA]
+
 ## 3. From the rules-based list to a 'relevancy model'
 
-Patel described her work at DunnHumby in terms of a shift in models from a 'rules-based list' to a 'relevancy model.' This was a significant change in how the 'Have You Forgotten' list was constructed from one kind of model to another.
-For me, the shift in models matters since it amounts to a much more probabilistic structuring of lists.
-
+We can see something of this shift from collaborative filtering in the shift that Patel describes as moving from 'a rules-based' model to personal relevance.
 I assume that the rules-based model derived from the early 1990s work on the `apriori` algorithm by Rakesh Agrawal and Ramakrishnan Srikanti at IBM Research Alameda [@Agrawal_1994].
 The `apriori` algorithm finds sets of items that commonly occur in transactions.
+It was listed as one of the top ten data mining algorithms in a survey conducted amongst data miners [@TBA_2007].
 These commonly occurring sets are expressed as association rules.
@@
 [http://www.dssresources.com/newsletters/66.php]
+
 The `apriori` algorithm is widely used in 'market basket analysis' and attempts to re-structure shopping lists in terms of associations between items that are frequently bought together.
@@
 data(Groceries)
-itemFrequencyPlot(Groceries,topN=20,type="absolute")
+itemFrequencyPlot(Groceries,topN=10,col=rainbow(10), ntype="absolute",  cex.axis = 0.55)
 rules <- apriori(Groceries, parameter = list(supp = 0.001, conf = 0.8))
@@
 
-Badiou in his account of multiples highlight the contrast between two different ways of grasping a multiple -- in terms of the 'numerical one' and the 'vital One' [@Badiou_2004, 72], between the 'closed' and the 'open' respectively.
+Badiou in his account of multiples highlights the contrast between two different ways of grasping a multiple -- in terms of the 'numerical one' and the 'vital One' [@Badiou_2004, 72], between the 'closed' and the 'open' respectively.
+Note that Badiou does not explicitly oppose the closed and open.
+While Hansen tends to see predictive analytics as the closed-formed of the micrological propensities of things,  Badiou refers to a 'joint effectuation.'
 
-The `apriori` algorithm, might be seen as an attempt to deal with the excess of things in a Tesco supermarket, and the lack of interest or engagement that people have with them by identifying subsets whose membership is stronger by virtue of the other sets of associations that people bring with them.
+What would be Badiou's take on a grocery list as augmented by a recommender system?
+The `apriori` algorithm, might be seen as an attempt to grasp the excess of things in a Tesco supermarket by identifying subsets whose membership is stronger by virtue of the other sets of associations that people bring with them.
 Association is stronger when it derives from an excess of different ways of being in sets.
@@
 Badiou's argument is that being part of a set differs from belonging to a set.
-An open set arises from when we dwell within a set by 'taking the intersection of two elements.' What might this mean in terms of ordering?
+An open set arises from when we dwell within a set by 'taking the intersection of two elements.' What might this mean in terms of ordering carried out with specific grocery items in a supermarket?
 
-If we take the itemset from the `Groceries` dataset [@Hahsler_2006], and then run it through the Tesco online groceries site, each of the 169 items in the dataset yields dozens and sometimes thousands of results from Tesco online.
+If we take a typical grocery dataset used in data-mining modelling (e.g.
+from the `Groceries` dataset [@Hahsler_2006a] acquired from 'local German supermarket'), and then paste it into the 'shopping list' box on on the Tesco grocery website, each of the 169 items in the dataset yields dozens and sometimes thousands of results from Tesco online.
 
@@
 Each of the items in the `Groceries` dataset proliferates in the Tesco site, into a new list of branded variations.
-The 169 items of the `Groceries` dataset, itself from a local German or Austrian supermarket [@REF_TBA],  expand into roughly `r total` possible items.
+The 169 items of Hahsler `Groceries` dataset  expand into roughly `r total` possible items.
+This is, I would suggest, a form of open.
+The multiplicity grasped at the `Grocery` dataset becomes more open in this setting, where 'open' complements any possible closure with new intersections.
+Rather than seeing predictive analytics as a closed form of an open world, I am suggesting that we might begin to locate it in the joint effectuations or intersections within models.
 
@@
 
+Patel said that some people log on and go straight to the checkout, allowing the recommender system to supply them with the suggestions that they then move to the shopping basket.
 Patel observed that the 'Have you forgotten' lists matter to people, and we can see why in the light of this proliferation of things in the list.
 A `r total` list of items is quite hard to list let alone read.
-Patel said that some people log on and go straight to the checkout, allowing the recommender system to supply them with the suggestions that they then move to the shopping basket.
 This is the online equivalent of using an old shopping list and only adding new items when needed.
@@
 chicken products), but these lists are often too long to work with.
-For instance, Tesco has more than 1000 chicken products on its site.
+For instance, if Tesco has more than 1000 chicken products on its site, how can a customer even look through the list of chicken products?
+
 People inhabit lists and work with them no matter how they are made; thus vindicating ethnomethodologists assertion that social order is the product of ongoing sense-making practices.
-But it also suggests a transformation in the way that lists exist in the world.
-They are less the product of a discrete agent than an effect of an environment teeming with possible combinations and complementaries, a surplus of possible purchases that the supermarket needs to continually maintain but also filter in order to make it possible for something to happen.
+[^45] But it also suggests a transformation in the way that lists exist in the world.
+They are less the inscriptions of a discrete agent than an effect of an environment teeming with possible combinations and complementaries, a surplus of possible purchases that the supermarket needs to continually maintain but also filter in order to make it possible for something to happen.
 
-## 5. Data: 52 weeks of data for each customer: how is order constructed -- from history?
+[^45]: What have ethnomethodologists and others actually said about lists?
 
-Patel described the data they on grocery purchases. 'We have 52 weeks of purchase data for each customer.' 
+## 5. 'We have 52 weeks of data for each customer': how is order constructed -- from history?
+
 We know that the construction of lists has been transformed by recommender systems.
-These transformations have been widely discussed in the context of big data, where questions of who has how much data are usually most prominent.
-In terms of the ways that lists order the world, the amount of data is not so amount as the way in which things in the world take on a list-like existence. 
-I want to dwell on this idea for a while. We have seen two different ways in which lists might be constructed in a recommender system. 
+Patel described the data that Tesco keeps on grocery purchases: 'We have 52 weeks of purchase data for each customer.' 
+These transformations have been widely discussed in the context of big data, where questions of who has how much data are usually most prominent (see [@Kitchin_2014]).
+In terms of the ways that lists order the world, the amount of data is not perhaps not so important as the way in which list-making is re-distributed in the world. 
+I want to dwell on this idea for a while.
+
+HERE
+
+We have seen two different ways in which lists might be constructed in a recommender system. 
 The first uses history and existence of previous lists, including the customer's own list, to populate a new list. 

### 2016-10-26 lists_conversion_experience.rmd
log updates

@@
 ---
-title: "'Have you forgotten?' Conversion experiences in lists"
+title: "'Have you forgotten?' Conversion experiences in shopping lists"
 author: "Adrian Mackenzie"
@@
 
-# 'Have you forgotten?' Conversion experiences in lists
+# 'Have you forgotten?' Conversion experiences in shopping lists
 
@@
 ggplot(df, aes(x=price,y=ContentsQuantity)) + geom_point() + theme(axis.text=element_text(size=10)) 
-kable(df[df$price>=20,'name'])
+kable(head(df[df$price>=30,c('name','price')]))
 ```
@@
 
-![The advent of recommender systems](figure/resnick_varian_page1_197.png)
+![The advent of recommender systems](figure/resnick_varian_1997_p1.png)
 
@@
 
-```{r arules, cache=TRUE, echo=TRUE, fig.margin=TRUE}
+```{r arules, cache=TRUE, echo=FALSE, fig.margin=TRUE}
 library(knitr)
@@
 
-```{r tesco, cache=TRUE, fig.cap ='Tesco surplus',  fig.margin=TRUE}
+```{r tesco, echo=FALSE, cache=TRUE, fig.cap ='Tesco surplus',  fig.margin=TRUE}
     library(ggplot2)
@@
 People inhabit lists and work with them no matter how they are made; thus vindicating ethnomethodologists assertion that social order is the product of ongoing sense-making practices.
-[^45] But it also suggests a transformation in the way that lists exist in the world.
+[^45] But the predicted list also suggests a transformation in the way that lists exist in the world.
 They are less the inscriptions of a discrete agent than an effect of an environment teeming with possible combinations and complementaries, a surplus of possible purchases that the supermarket needs to continually maintain but also filter in order to make it possible for something to happen.
@@
 
-HERE
+We have seen one ways in which lists might be constructed in a recommender system.[^56]
 
-We have seen two different ways in which lists might be constructed in a recommender system. 
-The first uses history and existence of previous lists, including the customer's own list, to populate a new list. 
-The second uses many shopping lists to identify associations between things and then uses the strongest associations to recommend additions to a current list. 
+[^56]: The first uses history and existence of previous lists, including the customer's own list, to populate a new list.
+
+It uses many shopping lists to identify associations between things and then uses the strongest associations to recommend additions to a current list. 
 The 'relevancy model,' however, tries to do something different. 
@@
 
-In some ways, Hansen's account resembles the Badiouean ontology of the multiple in which a closed set is surrounded and conditioned by an open set that is full of many more potentials.
+Hansen's account opposes the closed prediction to the open propensities of the data-world.
+A closed set is surrounded and conditioned by an open set that is full of many more potentials.
 However, if we think of the  new model that Patel describes as the delimited (or delimiting?) system within   the larger surplus of sensibility, how that help us understand what the 'Have you forgotten?' recommendations do?
@@
 Or more accurately, she just said that logistic regression did better than other, and there was little more to discuss on that front.
-The measure of doing better here is called precision, and it measures the proportion of the recommended products that are actually purchased.
 
@@
 The graph compares models in terms of their predictions and how those predictions came to pass.
-The two columns on the left show earlier models -- the global model, which is not personalized, the association rules model for individual customers, and then the new models (logistic regression, random forest, etc) for customers.
+The two columns on the left show earlier models -- the global model, which is not personalized, the association rules model for individual customers, and then the new relevance models (logistic regression, random forest, etc) for customers.
 An increase in precision means more sales.
@@
 
-The fact that so little was said about the models themselves suggests that they are now largely taken for granted in recommender systems.
-The models she mentioned are all classifiers, predictive models that treat a set of variables in a datasets as a vector space in which good classification means finding a way of defining a decision boundary between points that belong to different classes of things.
+The fact that so little was said about the models themselves suggests that they are now largely taken for granted in recommender systems.[^57] The models Patel mentioned are all classifiers, predictive models that treat a set of variables in a datasets as a vector space in which good classification means finding a way of defining a decision boundary between points that belong to different classes of things.
 Here the classes of things are binary: `recommended` or `not recommended.` Some classifier models, including logistic regression, generate a probability rather than simply a classification.
+
+[^57]: The measure of doing better here is called precision, and it measures the proportion of the recommended products that are actually purchased.
+
+So, the model generates probabilities for each product for each customer.
+The 'good features' strengthen associations between people and things.
+Each thing has a propensity in relation to each customer.
 
@@
 A data matrix for just whether they ever bought a particular product or not will be `r format(1e6 * 2e5)` in size.
-Given that any one customer is like to only have bought 100 different products, the data matrix will be `r format(100-(100/2e5)*100, 2)`% empty.
+Given that any one customer is like to only have bought 100 different products, the probability matrix will be `r format(100-(100/2e5)*100, 2)`% empty.
 
@@
 p = as.logical(rbinom(n=2e5, size=1, prob=0.001))
-head(p, 30)
 m = (matrix(p, nrow=1000, byrow=TRUE))
@@
 
-So the raw data -- the 52 weeks of purchases for each customer -- can be aggregated to increase the density of the data in ways that helps the model.
+So the raw data -- the 52 weeks of purchases for each customer -- can be rewoven with the database of Tesco products to increase the density of the data in ways that helps the model.
 Patel mentions some of the 'features' used in the new model but somewhat strangely at this point, her presentation turns toward the difficulties of measuring the success of the model and the infrastructural limitations on making recommendations in ways that are more alive to what is happening right now.

### 2016-10-27 lists_conversion_experience.rmd
log update

@@
 latex_engine: xelatex
-bibliography: ~/ref_bibs/uni.bib
+bibliography: /home/mackenza/ref_bibs/uni.bib
 link-citations: yes
@@
 While the title 'have you forgotten' implies that you will be shown items that could have been or are usually on a shopping list, the working assumption is a temporal one -- the list is incomplete either because you have forgotten to list something, or you didn't know what should be on your list, perhaps because you have never even thought of putting that thing on your list.
-- making some changes, including a slide of early recommendation systems from 1990s such as described in ~/archive/data_intensive/machine_learning/schafer_recommenders_ecommerce_2001.pdf
--
 
-The items on this list are recommendations. They are only a few of them: five. The question: which five items and in what order?
+```{r random_basket, echo=FAlSE, cache=TRUE, fig.cap='Random basket'}
+library(dplyr)
+library(ggplot2)
+library(readr)
+library(knitr)
+csvs = list.files('data', pattern = '*.csv',full.names=TRUE)
+df = bind_rows(sapply(csvs, read_tsv))
+kable(sample_n(df, size=5)$name, col.names='item')
+```
+
+The items on this list could be recommendations. They are only a few of them: five. The question: which five items and in what order?
 
@@
 
-ggplot(df, aes(x=price,y=ContentsQuantity)) + geom_point() + theme(axis.text=element_text(size=10)) 
-kable(head(df[df$price>=30,c('name','price')]))
+g = ggplot(df, aes(x=price,y=ContentsQuantity)) + geom_point(alpha=0.3, size=0.9 ) + theme(axis.text=element_text(size=10))  + xlab('Price (£)')
+g + scale_x_log10() + geom_smooth()
+```
+
+```{r  simlate_basket_2, echo=FALSE, cache=TRUE, fig.margin=TRUE}
+s = sample_n(df[df$price>=30,], size=20)
+kable(s[, c('name','price')])
 ```
@@
 DunnHumby -- Edwina Dunn and Clive Humby -- are famous for having convinced the CEO of Tesco sometime in 1991 that a loyalty card program could change the supermarket chain's relationship to its customers.
-[@Okeefe_1984]  Clive Humby's academic work is hard to track down, but an early paper of his given at the Conference of Your Operational Researchers in Nottingham in 1984 suggests the direction that he and later Tesco would take in constructing lists [@Okeefe_1984]
+[@OKeefe_1984]  Clive Humby's academic work is hard to track down, but an early paper of his given at the Conference of Your Operational Researchers in Nottingham in 1984 suggests the direction that he and later Tesco would take in constructing lists [@OKeefe_1984]
 
@@
 
-![Early e-commerce recommender system architecture](figure/schafer_recommendation_2001.png)
-
-![The advent of recommender systems](figure/resnick_varian_1997_p1.png)
-
 > In everyday life, we rely on recommendations from other people.
@@
 
-## 3. From the rules-based list to a 'relevancy model'
+![Early e-commerce recommender system architecture](figure/schafer_recommendation_2001.png)
+
+![The advent of recommender systems](figure/resnick_varian_1997_p1.png)
+
+## 3. The rules-based list and the `apriori` algorithm
 
@@
 itemFrequencyPlot(Groceries,topN=10,col=rainbow(10), ntype="absolute",  cex.axis = 0.55)
-rules <- apriori(Groceries, parameter = list(supp = 0.001, conf = 0.8))
-
 options(digits=2)
-rules<-sort(rules, by="confidence", decreasing=TRUE)
-rules<-apriori(data=Groceries, parameter=list(supp=0.001,conf = 0.15,minlen=2), 
-           appearance = list(default="rhs",lhs="whole milk"),
-           control = list(verbose=F))
+rules<-apriori(data=Groceries, parameter=list(supp=0.001,conf = 0.15,minlen=2), appearance = list(default="rhs",lhs="whole milk"), control = list(verbose=F))
 rules<-sort(rules, decreasing=TRUE,by="confidence")
-kable(inspect(rules[1:5]))
+inspect(rules[1:5]) 
 items = Groceries@itemInfo
@@
 ## 4. Hack? Customers go straight to the checkout
+
+![Straight to the checkout?](figure/patel_straight_to_checkout.png)
 
@@
 
-![Models](figure/patel_recommendation1.png)
+![Models](figure/patel_compares_models.png)
 
@@
 
-```{r sparsity, echo=FALSE, cache=TRUE, fig.margin=TRUE, fig.cap ='The sparsity of data' }
+```{r sparsity, echo=FALSE, cache=TRUE, fig.margin=FALSE, fig.cap ='The sparsity of predictions' }
 library(ggplot2)
@@
 
-- listing practices are very closely entangled with problems of social order under predictive conditions or amidst predictive operations. 
-- lists have complex temporal-material structure, since they mix prediction, infrastructural constraints, the individual characteristics of list-makers, statistical validations, database architectures and above all the surplus of sensibility associated with commodity production.
+- the **constitutive incompleteness** and openness of lists are very closely entangled with problems of social order under predictive conditions or amidst predictive operations. 
+- shopping lists have complex temporal-material structure, since they mix prediction, infrastructural constraints, the individual characteristics of list-makers, statistical validations, database architectures and above all the **surplus sensibility** associated with  circulation of commodities.
+- shopping lists undergoing a **conversion experience** as they become more probabilistic, but the idea that probabilities are the closed form of propensities does not account for the distribution of probabilities in relations between things, between people, between people and things, and between things and people.
 - lists undergoing a 'conversion' experience as they become more probabilistic, but their heterogeneous character as operations caught up in models, infrastructures, experiments, and the flow of commodities means that they can never form a closed set or dataset.

### 2016-10-27 lists_slides.rmd
log update

--- 2016-10-27 lists_conversion_experience.rmd
@@
 title: "'Have you forgotten?' Conversion experiences in shopping lists"
-author: "Adrian Mackenzie"
+author: "Adrian Mackenzie, Lancaster University"
 date: "`r Sys.Date()`"
@@
     - \usepackage{setspace}
-    - \doublespacing
 ---
-
-# 'Have you forgotten?' Conversion experiences in shopping lists
 
@@
 
-## Abstract
+## How do we approach lists as objects of research?
 
-Amidst the predictive conditions of contemporary media-culture, the inevitable incompleteness of lists is both a problem and a potential. 
-This paper investigates practices of list-making associated with online grocery shopping at a  UK and global supermarket chain, Tesco. 
-It will analyse how a recommender system suggests a small number of grocery items of personal relevance to each of the millions of online grocery shoppers. 
-The predictive modelling practices aim to up-sell and cross-sell some of the 250,000 items in the supermarket's inventory. 
-I find it interesting to approach the list-making practices  of the grocery recommender system as a mundane case of the predictive intermediaries operating in many places. 
-The transformation or conversion of listing into predicting can be analysed in recommender systems that extend existing lists with potential new items, thus transforming the temporal and agential fabric of the list.
-While all written lists mix past and future actions, actual and potential events, their predictive re-writing in recommender systems adds a new operational dynamic.
-In this operational dynamic, the incompleteness of lists, their propensity to expand or change, becomes more important than their capacity to exhaustively enumerate things. 
-The conversion of lists from finite to in-definite in recommender systems suggests we might need to re-conceptualise how we think about social order and structure in media cultures.
+![shopping_list](figure/1_shopping.png )
 
-## Introduction: the presentation of a presentation
+-----------------------------
 
-I saw recently a presentation about online grocery shopping at Tesco, a large supermarket chain based in the UK.
-The lists in this setting are recommendations for further grocery purchases under the title of 'Have you forgotten?'.
-When you shop for groceries online, a list of 'Have you forgotten' recommendations appear  at the checkout stage.
-While the title 'have you forgotten' implies that you will be shown items that could have been or are usually on a shopping list, the working assumption is a temporal one -- the list is incomplete either because you have forgotten to list something, or you didn't know what should be on your list, perhaps because you have never even thought of putting that thing on your list.
-
-```{r random_basket, echo=FAlSE, cache=TRUE, fig.cap='Random basket'}
+```{r random_basket, echo=FALSE, cache=TRUE, fig.cap='Random basket'}
 library(dplyr)
@@
 df = bind_rows(sapply(csvs, read_tsv))
-kable(sample_n(df, size=5)$name, col.names='item')
+kable(sample_n(df, size=5)$name, col.names='item', caption='Have you forgotten?')
 ```
 
-The items on this list could be recommendations. They are only a few of them: five. The question: which five items and in what order?
-
-![shopping_list](figure/1_shopping.png )
-
-Shopping lists such as Figure \ref{fig:shopping_list} are quite prosaic -- a dozen lines scrawled on the back of a used envelope stuck on the refrigerator door  -- and sublime -- a million products ranked in some kind of order for tens or hundreds of millions of people.
-Such lists are highly repetitive, yet mix items of great diversity and display complicated temporal structures, spanning past and future, actual and potential.
-I regard the list of five suggestions show to Tesco's online customers at the checkout as exemplifying a major contemporary form of ordering, the recommendation list, that we see in many  forms of news, media, entertainment, marketing, advertising.
-
-The presentation I saw was not really surprising or exceptional.
-It was a presentation by a data scientist to data scientists, describing an implementation of a predictive model as well as some of the problems and potentials of the predictive model.
-I see the presentation itself as providing materials for an argument about our experience of predictive ordering.
-In this paper, I re-present the presentation using a mixture of theoretical and practical resources.
-The theory is largely recent media theory and its accounts of prediction in media.
-The practical resources are that of coding and predictive models.[^12]
-
-[^12]: The problem of incomplete lists reminds me of the game 'I went shopping and I bought ...
-' that adults sometimes play with children.
-The game is about remembering a well-ordered list of items - 'a' is apple, 'b' is for banana, 'c' is for carrot.
-The point of the game is to use the well-ordered set of labels provided by the sequence of letters in the alphabet to remember a random and perhaps very diverse list of items.
-The longer the game continues, the harder it is to remember since the shopping list grows longer.
-At the same time, each player repeats the list, and this  helps memorisation.
-One trick in playing this game is to choose items that are quite unexpected or increasingly random.
-Rather than `ham`, list `halva`; rather than `fish`, add `face cream.` This is especially useful for adults in beating children.
-The children have better memories for items they know, but adults know more unexpected items.
-On the other hand, very unexpected items are quite easy to remember and tend to even the score in the favour of the children.
-Balancing between the improbable and memorable, between a predictable list like 'apples, bananas, carrots' and an unusual list such as 'agave syrup, boccadilla, cilantro' is complicated.
-Something similar applies to recommender systems.
-The system can remember very many things.
-It may know everything you bought in the last few years on that site, and it can easily see what you would normally buy and add those as suggestions.
-
-A recommendation system should suggest items that you have not forgotten but may have never bought.
-If you do buy recommended items, then the recommender system has been effective in shunting the situation differently.
-It has led to a **'conversion.'** From a supermarket perspective, the conversion should be high in price.
-There is little value in successfully recommending a whole lot of small and cheap items such as rubber bands or toothpicks.
-So it might need to choose amongst different recommendations and weight high value items differently to low value items.
-In either case, 'conversion'  is the valued event.
-
-## How do we approach lists as objects of research?
-
-While lists such as recommendations intersect strongly with other meaning and sense-making practices in everyday life and popular culture, they are also somewhat refractory to discourse  and textual analysis methods.
-On the one hand, lists are highly associative.
-They can be interpreted or decoded semiotically, but they also exhibit great variations in textuality that multiply semiotic readings.
-On the other hand, as operational inscriptions, they can be treated ethnomethodologically, as the production of ('the order of social settings through ...
-shared sense making practices.')[https://en.wikipedia.org/wiki/Ethnomethodology]
-
-List practices such as online grocery recommendations figure as sites of intersection between people, infrastructures, things, places and predictions.
-We might, therefore,  approach lists from various novel angles -- as asignifying semiotics as Maurizio Lazzarato calls it in his _Signs and Machines_ [@Lazzarato_2014] or as elements in 'a new order of spatio-temporal continuity for forms of economic, political and cultural life' as Celia Lury terms in her account of topological turn [@Lury_2012].
-In this paper, I focus on lists and list-making in terms of the predictive conditions of contemporary media experience [@Hansen_2015a].
-
-The question for me here is **how** contemporary lists can still be understood as produced through 'shared sense-making practices.' All lists are forms of order, since even random ordering is operationally important, but what counts as order might historically transformable, in which case the order of social settings or the problem of social order might be deeply historically contingent.
-The problem of order lies at the core of lists making, and our relation to lists can be seen as a test of the way that order operates more generally.
-
-In this context, the fact that the lists and indeed supermarkets more generally are driven by database logics suggests that working with list data could well entail interacting with databases.
-
-This is no mere methodological problem if, as Mark Hansen has recently suggested,
+---------------------------------------
 
@@
 
-The question exercising much contemporary work on data, on media, platforms and finance pivots on the 'micrological propensities' that might be re-ordering lists.
-Hansen's argues that these propensities make probabilities into something real, some actually in things:
-
 >Whatever explanatory and causal value predictive analytics of large datasets have is, I suggest, ultimately rooted in this ontological transformation whereby probabilities are understood to be expressions of the actual propensity of things [@Hansen_2015a, 120]
-
-As the quote shows, he presents this as an 'ontological transformation.' The practical and indeed experimental question for me here is whether such transformations can be detected and articulated in the prosaic setting of shopping lists and  shopping baskets.
-I will suggest that 'conversion' might be a better way of mapping these changes than 'transformation.'
 
@@
 
-I heard this presentation at an academic event with speakers from industry and government on statistics and operations research in industry.
-Speakers came from Shell, Disney, GCHQ, and various universities.
-Because I've done a lot of work on machine learning and data science more generally, I found the sessions both familiar and occasionally exciting.
-
 ![patel_notes](figure/patel_overall.png)
 
-Shreena Patel works as a data scientist for DunnHumby, a well-known [ customer science company ](https://www.dunnhumby.com/)  focusing on online grocery shopping at Tesco.
-I found her presentation on shopping lists to be  practical and evocative since it was focused quite a lot on implementation of recommendations, and symptomatic in various ways.
-It was filled with graphs, numbers, and tables concerning different aspects of the 'have you forgotten?' list.
-Understandably, this was not presented in a language of ontological transformation.
+-----------------------------------------------
 
-To this end, I will be seeking to reconstruct some of the workings of the recommender system by generating queries, and exploring results from  a very large supermarket chain database.
-
-```{r simulate_basket, echo=FALSE, cache=TRUE, fig.margin=TRUE, fig.cap='Prices and quantities of things in the world'}
+```{r simulate_basket, echo=FALSE, cache=TRUE, fig.margin=FALSE, fig.cap='Prices and quantities of things in Tesco'}
 library(dplyr)
@@
 
-```{r  simlate_basket_2, echo=FALSE, cache=TRUE, fig.margin=TRUE}
+--------------------------------------------------
+
+```{r  simlate_basket_2, echo=FALSE, cache=TRUE, fig.margin=FALSE}
 s = sample_n(df[df$price>=30,], size=20)
@@
 ```
-
-Much of my paper, and my responses to the presentation I saw from a data scientist  will focus on how the ordering that is done in a recommendation lists relates to the problem of social order more generally.
-The main narrative of the presentation concerned a shift from a set-based model for making lists to a predictive, probabilistic way of making lists in almost realtime and dealing with lists on a scale that is often difficult to grasp concretely.
-The change in list-making associated with learned models is increasingly widespread, and relies on a probabilistic treatment of people and things.
-
-I respond to the presentation in two ways.
-First, by recounting, contextualising and commenting on some of the main steps in making the 'Have your forgotten?' list  as a example of the sense-making done by data scientists and customer analytics teams in many places.
-Many of these steps were mentioned only briefly in the presentation, and they are largely taken for granted as part of predictive analytic practice.
-Second, by trying to step into the list-making practices  as an occasion to intervene and participate in the ontological conversion practices they engage in.
 
@@
 
-Tesco is the largest supermarket chain in UK, famous for Clubcard starting in 1991; often cited as a leading early example of Big Data Analytics. 
-DunnHumby -- Edwina Dunn and Clive Humby -- are famous for having convinced the CEO of Tesco sometime in 1991 that a loyalty card program could change the supermarket chain's relationship to its customers.
-[@OKeefe_1984]  Clive Humby's academic work is hard to track down, but an early paper of his given at the Conference of Your Operational Researchers in Nottingham in 1984 suggests the direction that he and later Tesco would take in constructing lists [@OKeefe_1984]
+------------------------------------
 
@@
 
-While Tesco has been very successful in data-mining its customers,  and perhaps became the UK's biggest supermarket through data-mining in the 1990s, the online shopping situation is a bit different.
-It is no longer organised around campaigns involving special offers or redemption of points, but on continuous up-selling and cross-selling to customers who might be anywhere, but especially are not in the supermarket itself.
-.
+---------------------------
 
@@
 
+-------------------------------
+
 ![People add to list](figure/online_lists_in_practice.png)
 
-How should characterise this shift from ClubCard membership data mining to online grocery markets?
-Patel described her work at DunnHumby in terms of a shift in models from a 'rules-based list' to a 'relevancy model.' This was a significant change in how the 'Have You Forgotten' list was constructed from one kind of model to another.
-For me, this shift is a part of broader re-organisation of prediction in which existing predictive practices are being re-distributed and intensified in particular ways.
-The shift in models matters since it amounts to a much more probabilistic structuring of lists.
+----------------------------------------
 
-When commentators first began talking about recommender systems in the 1990s, they highlighted the shift from demographic-driven market research or data mining techniques to personalized recommendations.
-For instance, writing in 1997 in a special section of the _Communications of the ACM_ on recommender systems, Paul Resnick and Hal Varian (at that time Dean of Information Sciences at UC Berkeley, but somehow now Chief Economist at Google), made much of this difference [@Resnick_1997].
-They emphasise the need for a term to distinguish the emerging practices from data mining.
-[^44]
+![The advent of recommender systems](figure/resnick_varian_1997_p1.png)
 
-[^44]: Also can use Schafer -- back to Association Rule Mining from early 1990s
+----------------------------------------
 
@@
 
-Writing near the advent of early web-based e-Commerce, Resnick and Varian (now Chief Economist at Google) imagined recommender systems as automations of the 'collaborative filtering' that people do when they write and read reviews of products and services (for instance, on the travel website, TripAdvisor; [@REF_Scott_TBA]
+----------------------------------------
 
@@
 
-![The advent of recommender systems](figure/resnick_varian_1997_p1.png)
-
 ## 3. The rules-based list and the `apriori` algorithm
 
-We can see something of this shift from collaborative filtering in the shift that Patel describes as moving from 'a rules-based' model to personal relevance.
-I assume that the rules-based model derived from the early 1990s work on the `apriori` algorithm by Rakesh Agrawal and Ramakrishnan Srikanti at IBM Research Alameda [@Agrawal_1994].
-The `apriori` algorithm finds sets of items that commonly occur in transactions.
-It was listed as one of the top ten data mining algorithms in a survey conducted amongst data miners [@TBA_2007].
-These commonly occurring sets are expressed as association rules.
-For instance, using the `Groceries` dataset in the `R` package `arules,` we can readily see how Walmart might have discovered  -- but didn't -- the fabled association between diapers and beer.
-[http://www.dssresources.com/newsletters/66.php]
-
-The `apriori` algorithm is widely used in 'market basket analysis' and attempts to re-structure shopping lists in terms of associations between items that are frequently bought together.
-
-```{r arules, cache=TRUE, echo=FALSE, fig.margin=TRUE}
+```{r arules, cache=TRUE, echo=FALSE, fig.margin=FALSE}
 library(knitr)
@@
 itemFrequencyPlot(Groceries,topN=10,col=rainbow(10), ntype="absolute",  cex.axis = 0.55)
+```
+
+--------------
+
+```{r arules_2, cache=TRUE, echo=FALSE}
 options(digits=2)
@@
 inspect(rules[1:5]) 
-items = Groceries@itemInfo
 ```
 
-The association rule-based model for the `Groceries` dataset suggests some of the characteristic ordering practices of 1990s-style data-mining approaches to lists. 
-The association between items is expressed in the form of _rules_.
-These rules, a term derived from decision support literature, define sets of associations between items.
-As a set theory-based practice, the association rules define likely sets of items or shopping baskets in terms of decreasing probability.
-`r inspect(rules[1])` is a stronger association than `r inspect(rules[2])`.
-So any recommendation in the 'Have You Forgotten' list should probably put them in that order.
-
-Association rules and the `apriori` algorithm present social order in terms of sets of common associations that are already weighted by frequency and hence probability.
-This means that that the algorithm can be understood in terms of the set-based multiples proposed by Alain Badiou:
+----------------------------------------
 
@@
 
-Badiou in his account of multiples highlights the contrast between two different ways of grasping a multiple -- in terms of the 'numerical one' and the 'vital One' [@Badiou_2004, 72], between the 'closed' and the 'open' respectively.
-Note that Badiou does not explicitly oppose the closed and open.
-While Hansen tends to see predictive analytics as the closed-formed of the micrological propensities of things,  Badiou refers to a 'joint effectuation.'
+> A concept of the open is substantially established once we possess a multiple such that the we dwell within it by taking the intersection of two elements, or the union of many elements as we wish (even an infinity of elements).
+In the other words, the intersection of two opens is an open, and any union whatsoever of opens remains open.
+[@Badiou_2004, 73]
 
-What would be Badiou's take on a grocery list as augmented by a recommender system?
-The `apriori` algorithm, might be seen as an attempt to grasp the excess of things in a Tesco supermarket by identifying subsets whose membership is stronger by virtue of the other sets of associations that people bring with them.
-Association is stronger when it derives from an excess of different ways of being in sets.
+----------------------------------------
 
-Badiou's argument is that being part of a set differs from belonging to a set.
-An open set arises from when we dwell within a set by 'taking the intersection of two elements.' What might this mean in terms of ordering carried out with specific grocery items in a supermarket?
-
-If we take a typical grocery dataset used in data-mining modelling (e.g.
-from the `Groceries` dataset [@Hahsler_2006a] acquired from 'local German supermarket'), and then paste it into the 'shopping list' box on on the Tesco grocery website, each of the 169 items in the dataset yields dozens and sometimes thousands of results from Tesco online.
-
-```{r tesco, echo=FALSE, cache=TRUE, fig.cap ='Tesco surplus',  fig.margin=TRUE}
+```{r tesco, echo=FALSE, cache=TRUE, fig.cap ='Tesco surplus',  fig.margin=FALSE}
     library(ggplot2)
@@
 
-Each of the items in the `Groceries` dataset proliferates in the Tesco site, into a new list of branded variations.
-The 169 items of Hahsler `Groceries` dataset  expand into roughly `r total` possible items.
-This is, I would suggest, a form of open.
-The multiplicity grasped at the `Grocery` dataset becomes more open in this setting, where 'open' complements any possible closure with new intersections.
-Rather than seeing predictive analytics as a closed form of an open world, I am suggesting that we might begin to locate it in the joint effectuations or intersections within models.
-
 ## 4. Hack? Customers go straight to the checkout
@@
 
-Patel said that some people log on and go straight to the checkout, allowing the recommender system to supply them with the suggestions that they then move to the shopping basket.
-Patel observed that the 'Have you forgotten' lists matter to people, and we can see why in the light of this proliferation of things in the list.
-A `r total` list of items is quite hard to list let alone read.
-This is the online equivalent of using an old shopping list and only adding new items when needed.
-It is also a strategy to avoid having to view too many lists, since the only way Tesco can show people what they might buy is by showing them lists of things (e.g.
-chicken products), but these lists are often too long to work with.
-For instance, if Tesco has more than 1000 chicken products on its site, how can a customer even look through the list of chicken products?
-
-People inhabit lists and work with them no matter how they are made; thus vindicating ethnomethodologists assertion that social order is the product of ongoing sense-making practices.
-[^45] But the predicted list also suggests a transformation in the way that lists exist in the world.
-They are less the inscriptions of a discrete agent than an effect of an environment teeming with possible combinations and complementaries, a surplus of possible purchases that the supermarket needs to continually maintain but also filter in order to make it possible for something to happen.
-
-[^45]: What have ethnomethodologists and others actually said about lists?
-
 ## 5. 'We have 52 weeks of data for each customer': how is order constructed -- from history?
-
-We know that the construction of lists has been transformed by recommender systems.
-Patel described the data that Tesco keeps on grocery purchases: 'We have 52 weeks of purchase data for each customer.' 
-These transformations have been widely discussed in the context of big data, where questions of who has how much data are usually most prominent (see [@Kitchin_2014]).
-In terms of the ways that lists order the world, the amount of data is not perhaps not so important as the way in which list-making is re-distributed in the world. 
-I want to dwell on this idea for a while.
-
-We have seen one ways in which lists might be constructed in a recommender system.[^56]
-
-[^56]: The first uses history and existence of previous lists, including the customer's own list, to populate a new list.
-
-It uses many shopping lists to identify associations between things and then uses the strongest associations to recommend additions to a current list. 
-The 'relevancy model,' however, tries to do something different. 
-The problem that the proliferation of things on the shelves of supermarket or grocery warehouse confronts is partly combinatorial.
-With  `r total` products (actually Patel mentions 200,000 products) can be combined in truly astronomical numbers of combinations.
-If a typical shopping list has 20 items, then there are `r format(choose(total, 20), 2)` possible combinations.
-Some of this vast number of possible shopping lists is highly unlikely in terms of shopping lists and histories of shopping, but potentially highly interesting and value as recommendation.
 
@@
 
-Hansen's account opposes the closed prediction to the open propensities of the data-world.
-A closed set is surrounded and conditioned by an open set that is full of many more potentials.
-However, if we think of the  new model that Patel describes as the delimited (or delimiting?) system within   the larger surplus of sensibility, how that help us understand what the 'Have you forgotten?' recommendations do?
-
-- history of data accumulations;
-
-- shift from data mining to personalization? 
-- use some of the Varian, Linden, Resnick, Schaffer literature
-
 ## 6. Gain: use good features to build models
-
-- what counts as a good feature? See Domingos on this
-- move from Dunnhumby's ClubCard focus to much more general world of predictive models (e.g. Facebook use a million models, etc.)
-- the model counts for surprisingly little -- at least, they all seem similar ...
-rather what counts is all that stuff on working with substitutions, similiarities and complementarities for each person.
-
-It was surprising that Patel spoke relatively briefly about the new models.
-She showed a graph that compared the old rule-based model with some of the new alternatives -- logistic regression, random forests, and a few others.
-She dismissed most of the models quite quickly and focused only on one, the logistic regression model.
-Or more accurately, she just said that logistic regression did better than other, and there was little more to discuss on that front.
-
-Why did she say little about the model, which after all, produces the probabilistic predictions on which everything in the recommendation list depends and hence the conversion experience?
 
@@
 
-The graph compares models in terms of their predictions and how those predictions came to pass.
-The two columns on the left show earlier models -- the global model, which is not personalized, the association rules model for individual customers, and then the new relevance models (logistic regression, random forest, etc) for customers.
-An increase in precision means more sales.
-Her interest lay in the relevance of the recommendations, and in particular, in how 'most of the gain comes from thinking of good features.'
-
-The fact that so little was said about the models themselves suggests that they are now largely taken for granted in recommender systems.[^57] The models Patel mentioned are all classifiers, predictive models that treat a set of variables in a datasets as a vector space in which good classification means finding a way of defining a decision boundary between points that belong to different classes of things.
-Here the classes of things are binary: `recommended` or `not recommended.` Some classifier models, including logistic regression, generate a probability rather than simply a classification.
-
-[^57]: The measure of doing better here is called precision, and it measures the proportion of the recommended products that are actually purchased.
-
-So, the model generates probabilities for each product for each customer.
-The 'good features' strengthen associations between people and things.
-Each thing has a propensity in relation to each customer.
-
 ## 7. Sparsity: output 20 recommendations from 200,000 products
-
-What counts as a good feature in a world of predictive lists?
-Some things in the data that increases the precision of the model.
-The problem that Tesco faces, however, is that there are too many possible recommendations to be tested.
-They sell around 200,000 products and no customer wants to buy any more than a fraction of a per cent of those products.
-The recommendation list needs to very short.
-
-Predictive models, we might say, encounter their own surplus of sensibility in the data as they search for the most accurate predictions.
-A  very brief comment from Patel -- 'we have lots of zeros' -- points to the problem here.
-Each of the millions of Tesco online grocery shoppers has bought some selection of the 200,000 products.
-But the product-customer matrix, the basic vector-space in which all recommender systems operate, remain very sparse and unpopulated.
-
-Say Tesco has 1 million online customers.
-A data matrix for just whether they ever bought a particular product or not will be `r format(1e6 * 2e5)` in size.
-Given that any one customer is like to only have bought 100 different products, the probability matrix will be `r format(100-(100/2e5)*100, 2)`% empty.
 
@@
 
-Having too many things to list -- this must be happening in many settings -- videos, social networks, products, etc.
--- is not the real problem.
-The problem is that 'massively unbalanced data,' which means that different classes of products are not equally represented.
-Since so many people buy milk, the recommendation system might end up always recommending milk products.
-So the data needs to be 'corrected' by, in this case, random under-sampling or removal of some of the data for common purchases.
-Secondly, even if a recommendation system finds uncommon items that are good recommendations, they should neither be too cheap ('low spend'), nor, as we will see too similar to what the customer has already ordered.
-
-In Hansen's or Badiou's terms, this is one way in which the dataset is closed.
-While 200,000 products is still a closed dataset since it has a defined number of items, undersampling alters the frequency of occurrence of very common items so that the predictive models are not distracted by them.
-
-So the raw data -- the 52 weeks of purchases for each customer -- can be rewoven with the database of Tesco products to increase the density of the data in ways that helps the model.
-Patel mentions some of the 'features' used in the new model but somewhat strangely at this point, her presentation turns toward the difficulties of measuring the success of the model and the infrastructural limitations on making recommendations in ways that are more alive to what is happening right now.
-
-## 8. Value: Balancing conversion vs spend
-The value of an ordering
-
-## 9. Metric: precision-at-K
-
-The discussion of imperfection was brief - it incorporated the notion of recommendations weighted by price and the notion of precision (or the proportion of products actually purchased.).
-Patel simply said that this metric was imperfect, but without elaborating why.
-Instead she moved to describe some of the ways in which the recommender system worked in practice.
-
 ## 10. Experimental settings: A/B allocation and uplift
@@
 ![A/B testing](figure/a_b_tests.png)
-
-First of all, the new model was deployed experimentally in a random A/B controlled trial.
-This means that all customers are allocated to one of four categories as shown in the figure.
-For an identical shopping list, or even for the same person, it is possible to receive different recommendations.
-
-This randomising of the application of the model, drawing on protocols for randomised clinical trials first developed in the 1960s, is widely used in social media platforms and hence in all recommender systems.
-It is difficult to gauge how much.
-Facebook reports that it has more than 1 million models operation in its infrastructure.
-[TBA: ref for FBflow]   We could also discuss the integration of experiments into the operational infrastructures at Google.[REF tba]
-
-Random allocation of customers to the four categories (Test A, Test B, Control A, Control B) adds another layer of probability to the recommender system.
-This is done in the name of statistical validation of the effects or the uplift of the model.
-It also suggests that the effects of the model cannot be known in advance -- they can only be observed experimentally.
-Does this experimentality of the list point more to the propensities that Hansen says saturate experience?
-
-It is perhaps significant that this random allocation of customers to control and test groups occurs without any relation to the particular profile or propensities of the customer.
-It seeks to statistically validate the effects -- the *uplift* -- of the model on  conversion events by allowing the effects of the model on what people do to be measured.
-The lift or uplift refers to a change in events associated with the same groups of people.
-This effectively is an experiment in inhabiting two different worlds.
-This sets up a feedback look between the predictive model, the logistic regression, and the operation of the recommendation system.[^122] Once the experimental uplift methodology is in place, the recommender system can constantly modify the recommendation list in light of subsequent data on the uplift.
-Predictions, then, are not bound by fixed parameters of the model.
-
-[^122]: See [@Lo_2002] for an introduction to the construction of experimental set ups for uplift models.
-Market researchers and data miners seem to have started adopting a statistically grounded experimental predictive practice sometimes around 2000.
-See for instance [@Almquist_2001]  Response modelling, propensity modelling and uplift modelling all seek to identify associations between 'treatments' or interventions and the 'Responders', the people  affected by the treatment.
-Interestingly, these
 
@@
 
-While much discussion of recent media has highlighted its realtime operation, Patel's description of the infrastructure that implements the new model suggested a different temporal patterning.
-While raw data from Tesco Online feeds into the model every hour, it seems that recommendation list for each customer is only generated once a week.
-This makes sense.
-Customers only shop online every few days at the most, and in some cases, only every few weeks.
-To continually update the top 200 recommendations for every customer would be computationally expensive.
-
-Patel did not say much else about the infrastructure but the 200 recommendations generated each week are certainly never seen by the customers.
-The realtime engine that actually generates results shown on the Tesco Online website only shows 5 after 'applying business rules.' Again, this might be seen as corroborating Hansen and Badiou's insistence on the necessarily limited grasp of any prediction system on the situation it encounters.
-Even the new model, situated amidst its experimental setup cannot fully encompass the surplus of sensibility it encounters.
-
-The brief mention of specific infrastructural elements such as `hadoop` is, however, significant.
-It points to the temporality of infrastructures themselves as a component of list practices.
-The possibility of adjusting the recommendations for every customer weekly derives from an infrastructure capable of collecting data, assimilating that data to predictive model able to generate recommendations that are relevant to individuals amongst 200,000 products.
-
 ## 12. Similarity, substitution, complementarity and micrological propensities
@@
 
-What does it matter that recommendations are relevant?
-Relevance is closely connected to propensities in the predictive condition.
-From DunnHumby's standpoint, the relevance of a recommendation relates directly to conversion.
-From the standpoint of the ontological conversion, and  if Hansen is right, the real or actual fabric of prediction resides in the propensities of things, what he calls the 'complex and diffuse calculus' that only subsequently and provisionally congeals as a conversion event.
-The 'micrological propensities' of things feed directly into an sense of the relevance of recommendations.
+-------------------------------------------------
 
-How does the DunnHumby model deal with the micrological propensities of things?
-In her presentation, Patel  spent quite a lot of time addressing the problem of basket similarity in terms of substitutions and complementarities between things.
-She concludes by saying 'basket similarity is the future' and 'complementarities drive conversion.'
-
-Whereas the association rules focus on complementarities -- if you buy bread, you might want milk, if you buy beer, you might want diapers -- the substitutions look for micrological propensities that relate to gradations of preference, and variations in time and place.
-Micrological propensities in the list relate to a particular time and place, amidst a continuously differentiable pathways between things.
-
-How do things associate with each other in  substitutions and complementarities?
-Patel's comments on this are very brief but suggest that the 'Have You Forgotten?' list becomes an increasingly mutable entity, open to many substitutions and re-writings.
-Like the mention of `hadoop`, what Patel said about substitution and complementarity implies extensive but relatively well-established data practices.
-She speaks for instance of a 'design matrix of 14,000 columns' and the use of `L1 regularization` to drive coefficients down to zero.
-This is machine learning talk.
-What is being modelled here?
-No longer whether to recommend a product to a customer or not, but whether something could be substituted for what is already in the recommendation list.
-
-```{r biscuits, echo=FALSE, cache=TRUE, fig.margin=TRUE}
+```{r biscuits, echo=FALSE, cache=TRUE, fig.margin=FALSE}
     library(readr)
@@
 
-We can imagine this in terms of biscuits.
-A customer's recommendation list might already include `r bisc$name[10]`.
-The 'self-learning substitution model' might be able to substitute something in that list that is of higher value, either because it is bigger or more expensive.
-
 ## 13. Complementarity drives conversion
@@
 
-Similarly, to model the potential complementarities   between things entails a matrix of comparisons.
-`Nutella` complements `bananas,` but the converse is not necessarily true since `bananas` need not complement `nutella.`
-
-How should we understand the effect of complementarity?
-Patel says that 'complementarity drives conversion.' That is, the matrix of similarities expressed as numerical associations between products best animates the conversion of recommendations into purchases.
-
-In the similarity matrix, shaped by the history of past recommendations and purchases made by this customer, lie the propensities of things that Hansen (following Popper) attributes to the world.
-Karl Popper speaks of the future as a 'crystallization of propensities' [@Popper_1990,18]
-
 ## Conclusion
 
-My investigation of lists has been centred around the transformation in their predictive underpinnings.
-How should we think about the list-making practices of these models, databases and computational infrastructures?
-Following Hansen, I have been suggesting that the predictive listing in the form of open-ended recommendation is symptomatic of changes in the distribution of media.
-The idea that media should no longer be considered as extensions or prostheses of human communication but the environmental situation in which different propensities play out is increasingly central to contemporary media theory.
-
-Hansen refers to an ontological transformation in which predictive media or recommender systems would be seen as the closed realisation of the actually open propensities of the world.
-By understanding the world in terms of propensities, he argues, we have a change to engage with contemporary predictive media.
-
-I have not explored the Whitehead-derived philosophy of propensities in the world that he develops there, only because I am not sure that we need speculative theory to grapple with the material practices of prediction.
-Already in the DunnHumby model and the Tesco 'Have You Forgotten?' recommendations, many of the traits of predictive processes are at work.
-
-I diverge from Hansen in how I account for this change. Rather than looking for a speculative theory that sets out a world in which prediction would be the closed form,
-
-- the **constitutive incompleteness** and openness of lists are very closely entangled with problems of social order under predictive conditions or amidst predictive operations. 
+- the **constitutive incompleteness** and openness of lists are very closely entangled with problems of social order under predictive conditions
 - shopping lists have complex temporal-material structure, since they mix prediction, infrastructural constraints, the individual characteristics of list-makers, statistical validations, database architectures and above all the **surplus sensibility** associated with  circulation of commodities.
 - shopping lists undergoing a **conversion experience** as they become more probabilistic, but the idea that probabilities are the closed form of propensities does not account for the distribution of probabilities in relations between things, between people, between people and things, and between things and people.
-- lists undergoing a 'conversion' experience as they become more probabilistic, but their heterogeneous character as operations caught up in models, infrastructures, experiments, and the flow of commodities means that they can never form a closed set or dataset.
-Amidst prediction they become constitutively incomplete.
 
-## References
+--------------------------------------------

### 2017-02-01 lists_conversion_experience.rmd
planning edit

@@
 
-```{r random_basket, echo=FAlSE, cache=TRUE, fig.cap='Random basket'}
+```{r random-basket, echo=FALSE, cache=TRUE, fig.cap='Random basket'}
 library(dplyr)
@@
 options(digits=2)
-rules<-apriori(data=Groceries, parameter=list(supp=0.001,conf = 0.15,minlen=2), appearance = list(default="rhs",lhs="whole milk"), control = list(verbose=F))
-rules<-sort(rules, decreasing=TRUE,by="confidence")
-inspect(rules[1:5]) 
-items = Groceries@itemInfo
+rules<- apriori(data=Groceries, parameter=list(supp=0.001,conf = 0.15,minlen=2), appearance = list(default="rhs",lhs="whole milk"), control = list(verbose=F))
+rules_sorted <- sort(rules, decreasing=TRUE,by="confidence")
+rule1 <- inspect(rules_sorted[1])
+rule2 <- inspect(rules_sorted[2])
+kable(inspect(rules[1:5]))
+items <- Groceries@itemInfo
 ```
@@
 The association between items is expressed in the form of _rules_.
-These rules, a term derived from decision support literature, define sets of associations between items.
+These rules, a term derived from older decision support literature, define sets of associations between items.
 As a set theory-based practice, the association rules define likely sets of items or shopping baskets in terms of decreasing probability.
-`r inspect(rules[1])` is a stronger association than `r inspect(rules[2])`.
+Table \@ref(tab:arules) shows that the first rule has a stronger association than the second.
+Milk and vegetables is more likely an association than milk and buns.
 So any recommendation in the 'Have You Forgotten' list should probably put them in that order.

### 2017-02-01 render_tufte.r

@@ -1,6 +1,6 @@
 library(tufte)
 library(rmarkdown)
 library(tint)
-render('lists_conversion_experience.rmd',output_format=tufte_handout())
-render('lists_conversion_experience.rmd',output_format=ioslides_presentation)
+#render('lists_conversion_experience.rmd',output_format=tufte_handout())
+#render('lists_conversion_experience.rmd',output_format=ioslides_presentation)
 render('lists_conversion_experience.rmd',output_format=tintPdf())

### 2017-02-07 lists_conversion_experience.rmd
some local changes?

--- 2016-10-27 lists_conversion_experience.rmd
@@
 output:
-tufte::tufte_handout: default
+tint::tintPdf: default
 citation_package: biber
@@
 ---
-
-# 'Have you forgotten?' Conversion experiences in shopping lists
 
@@
 
-## Abstract
+# Abstract
 
@@
 
-## Introduction: the presentation of a presentation
+# Introduction: the presentation of a presentation
 
@@
 
-```{r random_basket, echo=FAlSE, cache=TRUE, fig.cap='Random basket'}
+```{r random_basket, echo=FALSE, cache=TRUE, fig.cap='Random basket', fig.margin=TRUE}
 library(dplyr)
-library(ggplot2)
 library(readr)
@@
 df = bind_rows(sapply(csvs, read_tsv))
-kable(sample_n(df, size=5)$name, col.names='item')
+kable(sample_n(df, size=5)$name, col.names='Items')
 ```
@@
 
-## How do we approach lists as objects of research?
+# How do we approach lists as objects of research?
 
@@
 
-## 1. The context of the presentation: 'statistics and operations research in industry'
+# 1. The context of the presentation: 'statistics and operations research in industry'
 
@@
 
-## 2. The web and the shopping basket: how it was reconstituted in the 1990s
+# 2. The web and the shopping basket: how it was reconstituted in the 1990s
 
@@
 
-## 3. The rules-based list and the `apriori` algorithm
+# 3. The rules-based list and the `apriori` algorithm
 
@@
 inspect(rules[1:5]) 
+rule2 = inspect(rules[2])
+rule1 = inspect(rules[1])
 items = Groceries@itemInfo
@@
 As a set theory-based practice, the association rules define likely sets of items or shopping baskets in terms of decreasing probability.
-`r inspect(rules[1])` is a stronger association than `r inspect(rules[2])`.
+`r rule1` is a stronger association than `r rule2`.
 So any recommendation in the 'Have You Forgotten' list should probably put them in that order.
@@
 
-## 4. Hack? Customers go straight to the checkout
+# 4. Hack? Customers go straight to the checkout
 
@@
 
-## 5. 'We have 52 weeks of data for each customer': how is order constructed -- from history?
+# 5. 'We have 52 weeks of data for each customer': how is order constructed -- from history?
 
@@
 
-## 6. Gain: use good features to build models
+# 6. Gain: use good features to build models
 
@@
 
-## 7. Sparsity: output 20 recommendations from 200,000 products
+# 7. Sparsity: output 20 recommendations from 200,000 products
 
@@
 
-## 8. Value: Balancing conversion vs spend
-The value of an ordering
-
-## 9. Metric: precision-at-K
+# 9. Metric: precision-at-K
 
@@
 
-## 10. Experimental settings: A/B allocation and uplift
+# 10. Experimental settings: A/B allocation and uplift
 
@@
 
-## 11. Infrastructure: realtime engine serves up top 5
+# 11. Infrastructure: realtime engine serves up top 5
 
@@
 
-## 12. Similarity, substitution, complementarity and micrological propensities
+# 12. Similarity, substitution, complementarity and micrological propensities
 
@@
 
-## 13. Complementarity drives conversion
+# 13. Complementarity drives conversion
 
@@
 
-## Conclusion
+# Conclusion
 
@@
 
-## References
+# References

### 2017-02-07 render_tufte.r

@@ -2,5 +2,5 @@ library(tufte)
 library(rmarkdown)
 library(tint)
 render('lists_conversion_experience.rmd',output_format=tufte_handout())
-render('lists_conversion_experience.rmd',output_format=ioslides_presentation)
-render('lists_conversion_experience.rmd',output_format=tintPdf())
+#render('lists_conversion_experience.rmd',output_format=ioslides_presentation)
+#render('lists_conversion_experience.rmd',output_format=tintPdf())

### 2017-02-09 lists_conversion_experience.rmd
done 10.5 pages

--- 2017-02-01 lists_conversion_experience.rmd
@@
 
-Amidst the predictive conditions of contemporary media-culture, the inevitable incompleteness of lists is both a problem and a potential. 
 This paper investigates practices of list-making associated with online grocery shopping at a  UK and global supermarket chain, Tesco. 
-It will analyse how a recommender system suggests a small number of grocery items of personal relevance to each of the millions of online grocery shoppers. 
-The predictive modelling practices aim to up-sell and cross-sell some of the 250,000 items in the supermarket's inventory. 
-I find it interesting to approach the list-making practices  of the grocery recommender system as a mundane case of the predictive intermediaries operating in many places. 
-The transformation or conversion of listing into predicting can be analysed in recommender systems that extend existing lists with potential new items, thus transforming the temporal and agential fabric of the list.
+It reconstructs an account of how a recommender system suggests a small number of grocery items of personal relevance to each of the millions of online grocery shoppers. 
+The recommender system  aim to up-sell and cross-sell some of the 250,000 items in the supermarket's inventory. 
+I suggest that we can understand devices such as recommender systems as transformations or conversions of the temporal and agential fabric of the important social ordering device, the list.
 While all written lists mix past and future actions, actual and potential events, their predictive re-writing in recommender systems adds a new operational dynamic.
-In this operational dynamic, the incompleteness of lists, their propensity to expand or change, becomes more important than their capacity to exhaustively enumerate things. 
+The incompleteness of lists, their propensity to expand or change, becomes more important than their capacity to  enumerate or rank things. 
 The conversion of lists from finite to in-definite in recommender systems suggests we might need to re-conceptualise how we think about social order and structure in media cultures.
 
-## Introduction: the presentation of a presentation
+##
 
-I saw recently a presentation about online grocery shopping at Tesco, a large supermarket chain based in the UK.
-The lists in this setting are recommendations for further grocery purchases under the title of 'Have you forgotten?'.
-When you shop for groceries online, a list of 'Have you forgotten' recommendations appear  at the checkout stage.
-While the title 'have you forgotten' implies that you will be shown items that could have been or are usually on a shopping list, the working assumption is a temporal one -- the list is incomplete either because you have forgotten to list something, or you didn't know what should be on your list, perhaps because you have never even thought of putting that thing on your list.
+Online grocery shopping at Tesco, a large supermarket chain based in the UK, includes recommendations for further grocery purchases under the title of 'Have you forgotten?'.
+When customers shop for groceries online, a list of five recommendations appear  at the checkout stage.
+These recommendations are the product of a recommender system, an important category of operational device in big data (see for instance [@Striphas_2015] for analysis of Netflix recommendation; [@Morris_2015]  or [@Seaver_2015] for an account of music recommendations]).
+The question 'have you forgotten?' is followed by some  grocery items that could have been or are usually on a shopping list.
+The title of the suggestions is a bit misleading: the recommender system, as we will see, is not concerned with forgetting, with the many slips and oversights associated with shopping lists,  but rather with items that you didn't know what should be on your list, perhaps because you have never thought of putting them there.
+
+The shelves of large contemporary supermarket are the endpoint of global logistic supply chains, in all their logistical, value-transforming and brand-mediated hypercomplexity (see [@Nielsen_TBA; @Tsing_2009] on logistics and power are the endpoint of global logistic supply chains, in all their logistical, value-transforming and brand-mediated hypercomplexity (see [@Nielsen_TBA; @Tsing_2009] on logistics and power)) Groceries display a geography of agriculture, industry, transport, communication and financialisation driven by flows of labour and capital.
+Encounters between this hyper-complex geography and people, even in the confines of a supermarket, are no simple matter, either for shoppers or for supermarket operators such as Tesco.
+While there are many facets to shopping [@Miller_2013], the shopping list, whether written on the back of an envelope, or saved as a list in an online grocery shopping system, is one intersectional ordering device.
+In a shopping trajectory (trawling along aisles packed with thousands of products, scrolling down screens or searching for particular brands amidst search results), lists inscribe some kind of order that filters or reduces the excessive propensities, claims, dazzle and distraction of commodities.
+
+Hand-written shopping lists undoubtedly have ongoing importance and display interesting mixed ordering practices (see the list of handwritten shopping lists at [Grocery List](http://www.grocerylist.org)), online shopping lists take shape at the intersection of web and internet infrastructures, supply chain logistics, individualised practices and *habitus,* and increasingly, the predictive operations of  recommender systems.
+Just as the aisles and shelves of a supermarket present a commercially ordered geography of the world,  online shopping recommender systems generate lists that seek to align people to commodities they otherwise might have little relation to.
 
@@
 The items on this list could be recommendations. They are only a few of them: five. The question: which five items and in what order?
+```{r shopping-hand}
+include_graphics('figure/1_shopping.png')
+```
 
-![shopping_list](figure/1_shopping.png )
-
-Shopping lists such as Figure \ref{fig:shopping_list} are quite prosaic -- a dozen lines scrawled on the back of a used envelope stuck on the refrigerator door  -- and sublime -- a million products ranked in some kind of order for tens or hundreds of millions of people.
-Such lists are highly repetitive, yet mix items of great diversity and display complicated temporal structures, spanning past and future, actual and potential.
-I regard the list of five suggestions show to Tesco's online customers at the checkout as exemplifying a major contemporary form of ordering, the recommendation list, that we see in many  forms of news, media, entertainment, marketing, advertising.
+In terms of content, there are many similarities between shopping lists such as Figure \@ref(fig:shopping-hand)  -- a dozen lines scrawled on the back of a used envelope stuck on the refrigerator door  --  and the five ranked items shown to a customer on an online shopping site (as shown in Table \@ref(tab:random-basket)).
+While they are similar in content,  the list of five suggestions shown to Tesco's online customers at the checkout  exemplify a transformation in forms of ordering, a transformation that may affect the ontological texture of the world in the same way that commodification does.
+In its re-writing of the list, the recommender system effects a conversion experience in which individualised desire, consumption practices, value-chains, infrastructural time-space dynamics, the lives of others and the propensities of things all come into play.
+ing a major contemporary form of ordering, the recommendation list, that we see in many  forms of news, media, entertainment, marketing, advertising.
 
@@
 
-A recommendation system should suggest items that you have not forgotten but may have never bought.
-If you do buy recommended items, then the recommender system has been effective in shunting the situation differently.
-It has led to a **'conversion.'** From a supermarket perspective, the conversion should be high in price.
-There is little value in successfully recommending a whole lot of small and cheap items such as rubber bands or toothpicks.
-So it might need to choose amongst different recommendations and weight high value items differently to low value items.
-In either case, 'conversion'  is the valued event.
+Despite being titled 'Have You Forgotten?' Tesco's recommender system does not suggest forgotten items.
+It focuses on 'conversion events.' If a shopper does buy  recommended items, 'conversion' has occurred.
+From a supermarket perspective, high-value conversion are desirable.
+There is little value in successfully recommending small or cheap items such as rubber bands or toothpicks.
+From a data science perspective, conversion events, when they occur, confirm the power of predictive analytics to change the world in some way.
+From a sociological perspective, constructing and configuring recommender system in the name of conversion events, the recommender system effects an epistemic-operational transformation central to big data as an operational formation that responds to an ongoing control crisis [@Beniger_1986].
+This central transformation is that the idea that data affords apprehension, visibility and management of associative propensities in the world.
+Recommender systems, then, are concerned with the conversion of social order.
 
-## How do we approach lists as objects of research?
+## Lists in and as data
 
-While lists such as recommendations intersect strongly with other meaning and sense-making practices in everyday life and popular culture, they are also somewhat refractory to discourse  and textual analysis methods.
+Lists have long interested historians, anthropologists, sociologists, literary and media theorists.
+If, as the social anthropologist Jack Goody argued, lists are historically primary as forms of writing [@Goody_1986], and if list-making and its later variants (e.g.
+tables) precede discursive and narrative writing practices, then we might expect lists to function as powerful social ordering devices.
+As literary scholars suggest (see [@Mainberger_2003]), although lists have often been de-valued as literary forms, list-making often appears, even in literary form, whenever writing seeks to address, name, group or evoke totality, profusion, excess or abundance.
+From both  anthropological and literary perspectives, shopping lists have excellent reasons to exist: they are deeply rooted in urban social formations historically and culturally.
+Lists such as online grocery recommendations weave together people, infrastructures, things, places and predictions.[^1.456]
+
+[^1.456]: While lists intersect strongly with other meaning and sense-making practices in everyday life and popular culture, they can be refractory to discourse  and textual analysis methods built around models of language or speech, with its rules of syntax and grammar.
 On the one hand, lists are highly associative.
-They can be interpreted or decoded semiotically, but they also exhibit great variations in textuality that multiply semiotic readings.
-On the other hand, as operational inscriptions, they can be treated ethnomethodologically, as the production of ('the order of social settings through ...
-shared sense making practices.')[https://en.wikipedia.org/wiki/Ethnomethodology]
+They can be interpreted or decoded semiotically, although they exhibit  variations in textuality -- how they are written and read -- that thwart semiotic readings.
+On the other hand, as operational inscriptions, they can be treated ethnomethodologically, as the production of social order in a given setting.
 
-List practices such as online grocery recommendations figure as sites of intersection between people, infrastructures, things, places and predictions.
-We might, therefore,  approach lists from various novel angles -- as asignifying semiotics as Maurizio Lazzarato calls it in his _Signs and Machines_ [@Lazzarato_2014] or as elements in 'a new order of spatio-temporal continuity for forms of economic, political and cultural life' as Celia Lury terms in her account of topological turn [@Lury_2012].
+We might, in the light of their mutability, permeability and embedding in social order,   approach lists from various theoretical angles -- as asignifying semiotics as Maurizio Lazzarato calls it in his _Signs and Machines_ [@Lazzarato_2014] or as elements in 'a new order of spatio-temporal continuity for forms of economic, political and cultural life' as Celia Lury puts it in her account of topological turn [@Lury_2012].
 In this paper, I focus on lists and list-making in terms of the predictive conditions of contemporary media experience [@Hansen_2015a].
 
-The question for me here is **how** contemporary lists can still be understood as produced through 'shared sense-making practices.' All lists are forms of order, since even random ordering is operationally important, but what counts as order might historically transformable, in which case the order of social settings or the problem of social order might be deeply historically contingent.
-The problem of order lies at the core of lists making, and our relation to lists can be seen as a test of the way that order operates more generally.
+All lists are forms of order.
+Mathematicians distinguish lists from sets on the basis of order.
+Formally, both lists and sets are collections, but a list also contains a mapping of its elements to an ordering, usually the natural or counting numbers.
+Sometimes this order is explicit.
+Numbers are written down one side of the list.
+Often the order is implicit to the spatial arrangement of the list.
+Even random order can be  operationally important (for instance, in gambling machine or a psychology experiment, randomly order lists might be important).
+Mappings between list order and social order are manifestly highly contingent.
+Listing practices in a given social setting, however, always index its engagement with necessary unstable dynamics of social order.
+From the perspective of social order, shopping lists, as they are co-constructed by a supermarket chain's recommender system, might be important intersectional zones for the conversion of the average everydayness of lists through data-analytic propensities.
 
-In this context, the fact that the lists and indeed supermarkets more generally are driven by database logics suggests that working with list data could well entail interacting with databases.
-
-This is no mere methodological problem if, as Mark Hansen has recently suggested,
+Mark Hansen has recently suggested,
 
@@
 
-As the quote shows, he presents this as an 'ontological transformation.' The practical and indeed experimental question for me here is whether such transformations can be detected and articulated in the prosaic setting of shopping lists and  shopping baskets.
-I will suggest that 'conversion' might be a better way of mapping these changes than 'transformation.'
+Hansen predicates prediction as an 'ontological transformation' transforming the relation between probabilities and things.
+The practical and indeed empirical  question is whether such transformations in probability that bring 'the propensities of things'  can be detected and articulated in the prosaic setting of shopping lists and  shopping baskets.
+'Conversion' might be a preferable term for these changes.
+It highlights a range of re-orientations in subjects, experience, things, numbers and structures  entailed more than the subjective 'understood-to-be-expressed' formulation that Hansen emphasises.
 
-## 1. The context of the presentation: 'statistics and operations research in industry'
+## DunnHumby's work for Tesco
 
-I heard this presentation at an academic event with speakers from industry and government on statistics and operations research in industry.
-Speakers came from Shell, Disney, GCHQ, and various universities.
-Because I've done a lot of work on machine learning and data science more generally, I found the sessions both familiar and occasionally exciting.
+In one of the many industry-meets-academia events occurring in increasingly data science-oriented higher education institutions in the UK, speakers from industry, government and commerce described their work.[^11.1] Shreena Patel, a PhD graduate in statistics and operations research,  works as a data scientist for DunnHumby, a well-known [ customer science company ](https://www.dunnhumby.com/)  focusing on online grocery shopping at Tesco.
+Speaking to an audience of statisticians and operations researchers, Patel   focused on the models underlying  shopping list recommendations.
+Her presentation was filled with graphs, numbers, and tables concerning different aspects of the 'Have you forgotten?' recommender system.
+Patel's presentation was not sociological.
+In order to explore the operational environment she described, and its predictive operations, I will reconstruct key features of her presentation using sample data drawn from the Tesco API ([Application Programmer Interface](https://devportal.tescolabs.com/docs/services/56c73b1bf205fd0ed81dbe7a/operations/56d5b2bee2813e0a0053c47c), [@Tesco.com_2017])  A sample of the most highly priced items in Tesco stores is shown in Table  \@ref(tab:expensive-groceries) and a range of data science, statistics and graphics libraries implemented in the widely-used statistical programming language `R` [@RDevelopmentCoreTeam_2010].[^11.2]
 
-![patel_notes](figure/patel_overall.png)
+[^11.1]: More than 100  Data Science Institutes have been set up in North America, Europe and UK since around 2012.
+The traffic between higher education and data analytics in industry is intense and flows in several forms: people funding, research findings, software and technical devices (code) and training.
 
-Shreena Patel works as a data scientist for DunnHumby, a well-known [ customer science company ](https://www.dunnhumby.com/)  focusing on online grocery shopping at Tesco.
-I found her presentation on shopping lists to be  practical and evocative since it was focused quite a lot on implementation of recommendations, and symptomatic in various ways.
-It was filled with graphs, numbers, and tables concerning different aspects of the 'have you forgotten?' list.
-Understandably, this was not presented in a language of ontological transformation.
+[^11.2]:
 
-To this end, I will be seeking to reconstruct some of the workings of the recommender system by generating queries, and exploring results from  a very large supermarket chain database.
-
-```{r simulate_basket, echo=FALSE, cache=TRUE, fig.margin=TRUE, fig.cap='Prices and quantities of things in the world'}
+```{r simulate-inventory, echo=FALSE, cache=TRUE, fig.margin=TRUE, fig.cap='Prices and quantities of things in the world'}
 library(dplyr)
@@
 g = ggplot(df, aes(x=price,y=ContentsQuantity)) + geom_point(alpha=0.3, size=0.9 ) + theme(axis.text=element_text(size=10))  + xlab('Price (£)')
-g + scale_x_log10() + geom_smooth()
+g + ylab('Number of products') + scale_x_log10() + geom_smooth()
 ```
 
-```{r  simlate_basket_2, echo=FALSE, cache=TRUE, fig.margin=TRUE}
-s = sample_n(df[df$price>=30,], size=20)
-kable(s[, c('name','price')])
+Figure \@ref(fig:simulate-inventory) plots something of the density of commodities in the Tesco grocery database.
+By virtue of their sheer abundance, the plot cannot easily separate the many items that Tesco sells, especially in the price range £1-£10.
+
+```{r  expensive-groceries, echo=FALSE, cache=TRUE, fig.margin=TRUE}
+s = sample_n(df[df$price>=30,], size=10)
+kable(s[order(s$price), c('name','price')])
 ```
 
-Much of my paper, and my responses to the presentation I saw from a data scientist  will focus on how the ordering that is done in a recommendation lists relates to the problem of social order more generally.
-The main narrative of the presentation concerned a shift from a set-based model for making lists to a predictive, probabilistic way of making lists in almost realtime and dealing with lists on a scale that is often difficult to grasp concretely.
-The change in list-making associated with learned models is increasingly widespread, and relies on a probabilistic treatment of people and things.
+I respond to Patel's presentation in two ways.
+First, by recounting, contextualising and commenting on some of the main steps in making the 'Have your forgotten?' list  as a example of the predictive sense-making done by data scientists and customer analytics teams working with transactional data in many commercial settings.
+Patel mentioned many of these steps only fleetingly in the presentation, for they are largely taken for granted as part of predictive analytic practice.
+Second, by trying to step into the list-making practices  as an occasion to apprehend in the wider conversions they signal.
+As a sociological response to a data scientist presenting a model to an audience of statisticians and operations researchers, my interest lies less in specific technical innovations and more on how the ordering work done by recommender systems relates to problems of social order more generally.
+The method of reconstruction borrows from the philosopher John Dewey a broad interest in how practical and abstract elements of experience are interwoven [@Dewey_1957].
+More concretely, it draws heavily on the vision of 'interface methods' proposed by Noortje Marres [@Marres_2015] in order to keep in view both the highly technical orderings of lists in recommender system and the changes in social orderings they might imply or impel.
 
-I respond to the presentation in two ways.
-First, by recounting, contextualising and commenting on some of the main steps in making the 'Have your forgotten?' list  as a example of the sense-making done by data scientists and customer analytics teams in many places.
-Many of these steps were mentioned only briefly in the presentation, and they are largely taken for granted as part of predictive analytic practice.
-Second, by trying to step into the list-making practices  as an occasion to intervene and participate in the ontological conversion practices they engage in.
+## How the shopping list was reconstituted in the 1990s
 
-## 2. The web and the shopping basket: how it was reconstituted in the 1990s
-
-!['Every little bit helps':Tesco ClubCard](figure/tesco-clubcard.jpg )
-
-Tesco is the largest supermarket chain in UK, famous for Clubcard starting in 1991; often cited as a leading early example of Big Data Analytics. 
-DunnHumby -- Edwina Dunn and Clive Humby -- are famous for having convinced the CEO of Tesco sometime in 1991 that a loyalty card program could change the supermarket chain's relationship to its customers.
-[@OKeefe_1984]  Clive Humby's academic work is hard to track down, but an early paper of his given at the Conference of Your Operational Researchers in Nottingham in 1984 suggests the direction that he and later Tesco would take in constructing lists [@OKeefe_1984]
+The main narrative of Patel's presentation concerned a shift from a well-established demographically-targeted data mining model for making recommendations to a predictive, probabilistic, personally relevant  re-writing of the shopping list in almost-realtime.
+The changes Patel described are increasingly widespread.
+While they are configured in Tesco-specific ways by DunnHumby (and this reflects a longer history), they are also broadly symptomatic.
+Tesco is the largest supermarket chain in UK, famous for its customer loyalty and targeted marketing programme known as 'Tesco  Clubcard,' which started in 1991.
+DunnHumby -- founded by operations researchers Edwina Dunn and Clive Humby -- is famous for having convinced the CEO of Tesco sometime in 1991 that a loyalty card program could change the supermarket chain's relationship to its customers.
+Clive Humby's academic publications  are hard to track down, but an early paper given at the _Conference of Young Operational Researchers_ in Nottingham in 1984 suggests the direction that he, DunnHumby and later Tesco would take in constructing lists [@OKeefe_1984].
+The abstract for Humby's presentation pre-figures an ongoing trajectory for data mining techniques aimed at eliciting detailed information on individual customer preferences.
 
@@
 
-While Tesco has been very successful in data-mining its customers,  and perhaps became the UK's biggest supermarket through data-mining in the 1990s, the online shopping situation is a bit different.
-It is no longer organised around campaigns involving special offers or redemption of points, but on continuous up-selling and cross-selling to customers who might be anywhere, but especially are not in the supermarket itself.
-.
-
->Tesco is the clear winner in the online grocery market, in fact it takes almost 50p of every £1 spent on food shopping on the internet.
-[ https://econsultancy.com/blog/64841-how-seo-helps-tesco-to-dominate-the-online-grocery-market/ ]( https://econsultancy.com/blog/64841-how-seo-helps-tesco-to-dominate-the-online-grocery-market/ )
+While Tesco  succeeded in data-mining its customers,  and perhaps became the UK's biggest supermarket with the help of data-mining in the 1990s, the online shopping environment  in 2016 is a bit different.
+It is no longer organised around campaigns involving special offers or redemption of points for demographically segmented customer (DunnHumby made heavy use of UK Census data).
+It can no longer rely on placement of goods in carefully chosen locations in stores.
+It needs or at least might want to continuously up-sell and cross-sell to customers who might hardly ever be in the supermarket itself.
+If 'Tesco is the clear winner in the online grocery market, in fact it takes almost 50p of every £1 spent on food shopping on the internet' [How SEO helps Tesco to dominate the online grocery market](https://econsultancy.com/blog/64841-how-seo-helps-tesco-to-dominate-the-online-grocery-market)[@Silverwood-Cope_2014], then Tesco itself has undergone some kind of conversion?
 
@@
 
-How should characterise this shift from ClubCard membership data mining to online grocery markets?
-Patel described her work at DunnHumby in terms of a shift in models from a 'rules-based list' to a 'relevancy model.' This was a significant change in how the 'Have You Forgotten' list was constructed from one kind of model to another.
-For me, this shift is a part of broader re-organisation of prediction in which existing predictive practices are being re-distributed and intensified in particular ways.
-The shift in models matters since it amounts to a much more probabilistic structuring of lists.
+How could we characterise the shift from ClubCard data mining to online grocery markets?
+Patel described her work at DunnHumby in terms of a shift in models from a 'rules-based list' to a 'relevancy model.' The relevancy model changed how the 'Have You Forgotten' list was constructed.
+This shift is a typical of broader re-organisation of prediction in which existing predictive practices are being re-distributed and intensified in particular ways.
+The shift in models pivots on a more probabilistic structuring of lists.
+The ostensible banality of the 'Have you forgotten?' recommendations betrays micrological signs of a topologically complex predictive infrastructure that subducts people and things in significant ways.
+
+## Reconfiguring recommendations
+
+HERE
 

### 2017-02-10 lists_conversion_experience.rmd
got to p16 of edits

@@
 
-HERE
-
-When commentators first began talking about recommender systems in the 1990s, they highlighted the shift from demographic-driven market research or data mining techniques to personalized recommendations.
-For instance, writing in 1997 in a special section of the _Communications of the ACM_ on recommender systems, Paul Resnick and Hal Varian (at that time Dean of Information Sciences at UC Berkeley, but somehow now Chief Economist at Google), made much of this difference [@Resnick_1997].
-They emphasise the need for a term to distinguish the emerging practices from data mining.
-[^44]
-
-[^44]: Also can use Schafer -- back to Association Rule Mining from early 1990s
+Academic researchers first began writing about recommender systems in the mid-1990s.
+They highlighted a potential shift from demographic-driven market research or data mining techniques to personalized recommendations.
+For instance, writing in 1997 in a special section of the _Communications of the ACM_ on recommender systems, Paul Resnick and Hal Varian (at that time Dean of Information Sciences at UC Berkeley, but currently Chief Economist at Google), made much of this difference [@Resnick_1997].
+Resnick and Varian  emphasise the need to distinguish the emerging practices from data mining:
 
@@
 
-Writing near the advent of early web-based e-Commerce, Resnick and Varian (now Chief Economist at Google) imagined recommender systems as automations of the 'collaborative filtering' that people do when they write and read reviews of products and services (for instance, on the travel website, TripAdvisor; [@REF_Scott_TBA]
+Writing just after the advent of web-based e-Commerce, Resnick and Varian  imagined recommender systems augmenting the 'collaborative filtering' that people do when they write and read reviews of products and services (for instance, on the travel website, TripAdvisor; [@Scott_2012]).
+In 1997, two decades ago, Resnick and Varian thought that 'people provide recommendations' and recommender systems would aggregate and rank for recipients.
+A slightly later review, [@Schafer_2001], diagrams an augmented 'natural social process' with a range of elements, technologies, inputs and outputs, with varying degrees of personalization.
 
@@
 
-![The advent of recommender systems](figure/resnick_varian_1997_p1.png)
+The shift from or conversion of  collaborative filtering appeared in Patel's presentations.
+Patel describes the move from 'a rules-based system'  to  a 'personal relevance model'.
+Each term in that phrase -- personal, relevance and model -- has some weight in the conversion work.
+We can assume that What Patel refers to as  the rule-based system refers  to the `apriori` algorithm, developed  by Rakesh Agrawal and Ramakrishnan Srikanti at IBM Research Alameda in the early 1990s [@Agrawal_1994].
+A now-classic approach to 'market basket analysis,' it was listed as one of the top ten data mining algorithms in a survey conducted amongst data miners [@Wu_2008] and usually attracts a chapter in data-mining and some machine learning textbooks (e.g.
+[@Hastie_2009]).
 
-## 3. The rules-based list and the `apriori` algorithm
-
-We can see something of this shift from collaborative filtering in the shift that Patel describes as moving from 'a rules-based' model to personal relevance.
-I assume that the rules-based model derived from the early 1990s work on the `apriori` algorithm by Rakesh Agrawal and Ramakrishnan Srikanti at IBM Research Alameda [@Agrawal_1994].
-The `apriori` algorithm finds sets of items that commonly occur in transactions.
-It was listed as one of the top ten data mining algorithms in a survey conducted amongst data miners [@TBA_2007].
-These commonly occurring sets are expressed as association rules.
-For instance, using the `Groceries` dataset in the `R` package `arules,` we can readily see how Walmart might have discovered  -- but didn't -- the fabled association between diapers and beer.
-[http://www.dssresources.com/newsletters/66.php]
-
-The `apriori` algorithm is widely used in 'market basket analysis' and attempts to re-structure shopping lists in terms of associations between items that are frequently bought together.
-
-```{r arules, cache=TRUE, echo=FALSE, fig.margin=TRUE}
+```{r arules, cache=TRUE, messages=FALSE, echo=FALSE, fig.margin=TRUE}
 library(knitr)
@@
 rules_sorted <- sort(rules, decreasing=TRUE,by="confidence")
-rule1 <- inspect(rules_sorted[1])
-rule2 <- inspect(rules_sorted[2])
-kable(inspect(rules[1:5]))
-items <- Groceries@itemInfo
+kable(inspect(rules_sorted[1:5]))
 ```
 
-The association rule-based model for the `Groceries` dataset suggests some of the characteristic ordering practices of 1990s-style data-mining approaches to lists.
-The association between items is expressed in the form of _rules_.
+The `apriori` algorithm finds sets of items that commonly occur together in transactions.
+These commonly occurring sets are expressed as 'association rules.' For instance, using the `Groceries` dataset in the `R` package `arules,` we can readily see how Walmart *might* have discovered   the fabled association between diapers and beer.[^149.1] Figure \@ref(fig:arules) shows how often common items appear.
+`Whole milk` appears most frequently.
+The `apriori` algorithm attempts to re-structure shopping lists using associations between items frequently bought together.
+The association rule-based model for the `Groceries` dataset illustrates some characteristic ordering practices of 1990s-style data-mining approaches to lists (the `Groceries` dataset  was acquired from a 'local German supermarket' [@Hahsler_2006a]).
+The association between items is expressed in the form of _rules_ not a model.
 These rules, a term derived from older decision support literature, define sets of associations between items.
-As a set theory-based practice, the association rules define likely sets of items or shopping baskets in terms of decreasing probability.
+As the product of a  counting practice, the association rules define sets of items or shopping baskets in terms of decreasing strength.
 Table \@ref(tab:arules) shows that the first rule has a stronger association than the second.
 Milk and vegetables is more likely an association than milk and buns.
-So any recommendation in the 'Have You Forgotten' list should probably put them in that order.
+A rules-based recommender system should, therefore, might put them in that order in the 'Have You Forgotten' list.
 
-Association rules and the `apriori` algorithm present social order in terms of sets of common associations that are already weighted by frequency and hence probability.
-This means that that the algorithm can be understood in terms of the set-based multiples proposed by Alain Badiou:
+[^149.1]: One of the founding narratives of big data customer analytics, the beer-diaper association was not, according to [@Power_2002] found through data mining at all.
 
-> Multiplicity will be called  ‘closed’ when grasped by the numerical one, and ‘open’ when grasped by the vital One.
-Every multiplicity is the joint effectuation of the closed and open, but its ‘veritable’ multiple-being lies on the side of the open.
-[@Badiou_2004, 72]
+While association rules and the `apriori` algorithm present social order in terms of sets of common associations weighted by frequency and hence implicitly by probability, an `apriori`-based recommender system contains no statistical model.
+It has no grasp on personal relevance since the associations it defines are only between things.
+Put in terms of the big data conversion, the rules-based system is not predictive.
+The `apriori` algorithm might be seen as an attempt to grasp the excess of things in a Tesco supermarket by identifying subsets whose closure derives from the associations between things that people construct (through buying habits, by virtue of a shopping list, etc.).
+In association rules, personhood or experience is irrelevant.
+The openness of these rules -- the `apriori` algorithm can product any number of rules of diminishing strength -- points towards the propensity of things, but cannot cannot individuate that propensity.
 
-Badiou in his account of multiples highlights the contrast between two different ways of grasping a multiple -- in terms of the 'numerical one' and the 'vital One' [@Badiou_2004, 72], between the 'closed' and the 'open' respectively.
-Note that Badiou does not explicitly oppose the closed and open.
-While Hansen tends to see predictive analytics as the closed-formed of the micrological propensities of things,  Badiou refers to a 'joint effectuation.'
-
-What would be Badiou's take on a grocery list as augmented by a recommender system?
-The `apriori` algorithm, might be seen as an attempt to grasp the excess of things in a Tesco supermarket by identifying subsets whose membership is stronger by virtue of the other sets of associations that people bring with them.
-Association is stronger when it derives from an excess of different ways of being in sets.
-
-Badiou's argument is that being part of a set differs from belonging to a set.
-An open set arises from when we dwell within a set by 'taking the intersection of two elements.' What might this mean in terms of ordering carried out with specific grocery items in a supermarket?
-
-If we take a typical grocery dataset used in data-mining modelling (e.g.
-from the `Groceries` dataset [@Hahsler_2006a] acquired from 'local German supermarket'), and then paste it into the 'shopping list' box on on the Tesco grocery website, each of the 169 items in the dataset yields dozens and sometimes thousands of results from Tesco online.
+Although data mining for association rules finds relations things, it struggles with the propensity of commodities to multiply.
+A simple illustration of the combinatorial problem faced by recommender system can be developed using the `Grocery` dataset.
+If we take all the items in the `Grocery` dataset and paste them into the 'shopping list' box on the Tesco grocery website, each of the 169 items in the dataset yields dozens and sometimes thousands of products from Tesco online.
 
@@
 
-Each of the items in the `Groceries` dataset proliferates in the Tesco site, into a new list of branded variations.
-The 169 items of Hahsler `Groceries` dataset  expand into roughly `r total` possible items.
-This is, I would suggest, a form of open.
-The multiplicity grasped at the `Grocery` dataset becomes more open in this setting, where 'open' complements any possible closure with new intersections.
+Items in the `Groceries` dataset proliferate into a Tesco's list of branded variations.
+The 169 items of the `Groceries` dataset  expand into roughly `r total` Tesco items.
+This is, I would suggest, a form of propensity in Hansen's terms, albeit one that occurs in or very close to the recommender system rather than  'in the world.'  The multiple listed  in the `Grocery` dataset becomes more open in this setting through new possible intersections.
+HERE , where 'open' complements any possible closure with new intersections.
 Rather than seeing predictive analytics as a closed form of an open world, I am suggesting that we might begin to locate it in the joint effectuations or intersections within models.

### 2017-02-13 lists_conversion_experience.rmd
adding up to p 22 revs

@@
 
-Items in the `Groceries` dataset proliferate into a Tesco's list of branded variations.
+Items in the`Groceries` dataset proliferate into a Tesco's list of branded variations.
 The 169 items of the `Groceries` dataset  expand into roughly `r total` Tesco items.
-This is, I would suggest, a form of propensity in Hansen's terms, albeit one that occurs in or very close to the recommender system rather than  'in the world.'  The multiple listed  in the `Grocery` dataset becomes more open in this setting through new possible intersections.
-HERE , where 'open' complements any possible closure with new intersections.
-Rather than seeing predictive analytics as a closed form of an open world, I am suggesting that we might begin to locate it in the joint effectuations or intersections within models.
+This is, I would suggest, a form of propensity in Hansen's terms, albeit one that occurs in, or very close to, the recommender system rather than  'in the world.' The multiple listed  in the `Grocery` dataset becomes more open in this setting through new impersonal intersections that must be bounded and ordered.
+The problem that the proliferation of things on the shelves of supermarket or grocery warehouse confronts is partly combinatorial.
+`r total` products (actually Patel mentioned 200,000 products) can be combined ways.
+If a typical shopping list has 20 items, then there are `r format(choose(total, 20), 2)` possible lists.
+Some of this vast number of possible shopping lists are highly unlikely in terms of typical purchasing patterns, but potentially highly interesting and value as recommendation.
 
-## 4. Hack? Customers go straight to the checkout
+Amidst this proliferation of possibles multiple, the recommender system  starts to become a site of ordering -- in both senses of that term -- of things and people together in shopping lists.
+They are less the inscriptions of a discrete agent than an effect of an environment teeming with possible combinations and complementarities, a surplus of associations that the market needs to both maintain and limit.
+For instance, given that Tesco lists more than 1000 chicken products on its site, how can a customer practically discriminate between chicken products?
+Patel mentioned that some people log on and go straight to the checkout, allowing the recommender system to supply them with the suggestions that they then move to the shopping basket.
+A `r total` list of items is quite hard to list let alone read.
+Going straight to the recommendation list is the predictive equivalent of using an old shopping list and only adding new items when needed.
 
-![Straight to the checkout?](figure/patel_straight_to_checkout.png)
+## Ordering relevance through personhood
 
-Patel said that some people log on and go straight to the checkout, allowing the recommender system to supply them with the suggestions that they then move to the shopping basket.
-Patel observed that the 'Have you forgotten' lists matter to people, and we can see why in the light of this proliferation of things in the list.
-A `r total` list of items is quite hard to list let alone read.
-This is the online equivalent of using an old shopping list and only adding new items when needed.
-It is also a strategy to avoid having to view too many lists, since the only way Tesco can show people what they might buy is by showing them lists of things (e.g.
-chicken products), but these lists are often too long to work with.
-For instance, if Tesco has more than 1000 chicken products on its site, how can a customer even look through the list of chicken products?
+Combinatorial surplus, however, does not exhaust what might be found in the data.
+The relevancy model that DunnHumby has implemented for Tesco (as for so many other online businesses) seeks to balance the tensions between proliferating associations and relative closure.
+It responds to the problem of a world where the propensities of things and the propensities of people mingle with each other, yet threaten to disorder the list, the site of their mutual ordering.[^181.1] Big data discourse in its promissory mode attributes potency to personalization: 'most important, using big data we hope to identify specific individuals rather than groups' [@Mayer-Schönberger_2013].
+The ontological transformation that media theorists such as Hansen suggests actually play out somewhat more conventionally: the associative propensities of things in the shopping list (as constructed by a recommender system) come together with forms of personhood associated with buying.
+A model of personhood based on purchase data delimits the propensities of things.
 
-People inhabit lists and work with them no matter how they are made; thus vindicating ethnomethodologists assertion that social order is the product of ongoing sense-making practices.
-[^45] But the predicted list also suggests a transformation in the way that lists exist in the world.
-They are less the inscriptions of a discrete agent than an effect of an environment teeming with possible combinations and complementaries, a surplus of possible purchases that the supermarket needs to continually maintain but also filter in order to make it possible for something to happen.
+[^181.1]: Rather than seeing predictive analytics as a closed form of an open world, I am suggesting that we might begin to locate it in the joint effectuations or intersections within models.
 
-[^45]: What have ethnomethodologists and others actually said about lists?
-
-## 5. 'We have 52 weeks of data for each customer': how is order constructed -- from history?
-
-We know that the construction of lists has been transformed by recommender systems.
-Patel described the data that Tesco keeps on grocery purchases: 'We have 52 weeks of purchase data for each customer.' 
-These transformations have been widely discussed in the context of big data, where questions of who has how much data are usually most prominent (see [@Kitchin_2014]).
-In terms of the ways that lists order the world, the amount of data is not perhaps not so important as the way in which list-making is re-distributed in the world. 
-I want to dwell on this idea for a while.
-
-We have seen one ways in which lists might be constructed in a recommender system.[^56]
-
-[^56]: The first uses history and existence of previous lists, including the customer's own list, to populate a new list.
+Patel described the data that Tesco keeps on grocery purchases: 'We have 52 weeks of purchase data for each customer.' These transformations have been widely discussed in the context of big data, where questions of who has how much data are usually most prominent (see [@Kitchin_2014]).
+In terms of the ways that lists order the world, the amount of data is not perhaps not so important as the way in which list-making is re-ordered via a model of personhood.
 
@@
 The 'relevancy model,' however, tries to do something different.
-The problem that the proliferation of things on the shelves of supermarket or grocery warehouse confronts is partly combinatorial.
-With  `r total` products (actually Patel mentions 200,000 products) can be combined in truly astronomical numbers of combinations.
-If a typical shopping list has 20 items, then there are `r format(choose(total, 20), 2)` possible combinations.
-Some of this vast number of possible shopping lists is highly unlikely in terms of shopping lists and histories of shopping, but potentially highly interesting and value as recommendation.
 
-> This means, to put it slightly differently, that surrounding any delimited predictive system is a larger field of data -- what I elsewhere call a “surplus of sensibility” -- that, viewed speculatively, indexes the causal efficacy of the total situation within which this delimited system operates.
-(Effectively, the latter gains its reliability from closing off this larger surplus of sensibility, thereby transforming an always excessive propensity into a [provisionally] closed dataset.) Because it affords data that exceed whatever any given predictive system might include, the data of the world’s causal efficacy -- the data constituting its real potentiality -- always and in principle facilitates knowledge that cannot be restricted to any particular agenda [@Hansen_2015, 126-7]
+Hansen attributes 'real potential' to data. In terms that somewhat echo big data epistemics, data has an associative propensity, according to Hansen:
 
-Hansen's account opposes the closed prediction to the open propensities of the data-world.
-A closed set is surrounded and conditioned by an open set that is full of many more potentials.
-However, if we think of the  new model that Patel describes as the delimited (or delimiting?) system within   the larger surplus of sensibility, how that help us understand what the 'Have you forgotten?' recommendations do?
+> [S]urrounding any delimited predictive system is a larger field of data -- what I elsewhere call a “surplus of sensibility” -- that, viewed speculatively, indexes the causal efficacy of the total situation within which this delimited system operates.
+(Effectively, the latter gains its reliability from closing off this larger surplus of sensibility, thereby transforming an always excessive propensity into a [provisionally] closed dataset.) Because it affords data that exceed whatever any given predictive system might include, the data of the world’s causal efficacy -- the data constituting its real potentiality -- always and in principle facilitates knowledge that cannot be restricted to any particular agenda [@Hansen_2015, 126-7].
 
-- history of data accumulations;
+The underlying contrast between an open world and a 'provisionally closed dataset' is both enabling and limiting.
+Hansen's account opposes closed prediction to the open propensities of a data-world.
+A closed set is surrounded and conditioned by an open set teeming with potentials.
+If we think of the  new model that Patel describes as the delimited (or delimiting?) system within   the larger surplus of sensibility, how would that help us understand what the 'Have you forgotten?' recommendations do?
+It enables us to understand -- as we will see next -- what the Tesco relevancy model borrows from the world.
+It limits, on the other hand, our understanding of how predictive analytics, or any other ordering device, themselves remain open, require configuration, and only exist amidst technical ensembles whose operation remains obscure.
 
-- shift from data mining to personalization? 
-- use some of the Varian, Linden, Resnick, Schaffer literature
+```{r models}
+include_graphics('figure/patel_compares_models.png')
+```
 
-## 6. Gain: use good features to build models
+Patel introduced the relevancy model with a graph familiar to machine learning and statistical modelling.
+(A sketch of her graph appears in Figure \@ref(fig:models).)  The graph plots the _precision_ -- the proportion of the recommended products that customers actually purchase -- for several different models.
+In media theoretical terms, the graph indexes the 'causal efficacy' of a recommender system, its capacity to include and transform propensities or 'real potentialities' into operational knowledge.
+The graph compares the old rule-based recommender systems with some of the relevancy-model alternatives -- logistic regression, random forests, and a few others -- in terms of their predictions and how those predictions turned out.
+Patel dismissed most of the models quite quickly and focused only on one, the logistic regression model, which did better than other alternatives.
+Why did she say little about the model, which after all, produces the probabilistic predictions on which everything in the recommendation list depends and hence the conversion experience?
+Her interest lay in the precision of the recommendations, and, crucially for our purposes, in how 'most of the gain comes from thinking of good features.'
 
-- what counts as a good feature? See Domingos on this
-- move from Dunnhumby's ClubCard focus to much more general world of predictive models (e.g. Facebook use a million models, etc.)
-- the model counts for surprisingly little -- at least, they all seem similar ...
-rather what counts is all that stuff on working with substitutions, similiarities and complementarities for each person.
-
-It was surprising that Patel spoke relatively briefly about the new models.
-She showed a graph that compared the old rule-based model with some of the new alternatives -- logistic regression, random forests, and a few others.
-She dismissed most of the models quite quickly and focused only on one, the logistic regression model.
-Or more accurately, she just said that logistic regression did better than other, and there was little more to discuss on that front.
-
-Why did she say little about the model, which after all, produces the probabilistic predictions on which everything in the recommendation list depends and hence the conversion experience?
-
-![Models](figure/patel_compares_models.png)
-
-The graph compares models in terms of their predictions and how those predictions came to pass.
-The two columns on the left show earlier models -- the global model, which is not personalized, the association rules model for individual customers, and then the new relevance models (logistic regression, random forest, etc) for customers.
-An increase in precision means more sales.
-Her interest lay in the relevance of the recommendations, and in particular, in how 'most of the gain comes from thinking of good features.'
-
-The fact that so little was said about the models themselves suggests that they are now largely taken for granted in recommender systems.[^57] The models Patel mentioned are all classifiers, predictive models that treat a set of variables in a datasets as a vector space in which good classification means finding a way of defining a decision boundary between points that belong to different classes of things.
-Here the classes of things are binary: `recommended` or `not recommended.` Some classifier models, including logistic regression, generate a probability rather than simply a classification.
-
-[^57]: The measure of doing better here is called precision, and it measures the proportion of the recommended products that are actually purchased.
-
-So, the model generates probabilities for each product for each customer.
-The 'good features' strengthen associations between people and things.
-Each thing has a propensity in relation to each customer.
+The models Patel so briefly mentioned are all classifiers, predictive models that spatially order  variables in a dataset as a high dimensional vector space.
+In the vector space, classification means drawing a decision boundary between points that henceforth belong to different classes of things [@Mackenzie_2015].
+Here the classes of things are binary: `recommended` or `not recommended,` and for many of the models shown in the graph, membership of a class is expressed as a probability.
+The logistic regression model generates probabilities for each product for each customer.
+Patel did not articulate the implication of the contrast between the left and right hand side of the graphic.
+Although the statistical framing appears explicitly (e.g.
+in measuring the precision of  recommender systems), the underlying shift to a statistical model is barely mentioned.
+In becoming predictive, the 'Have You Forgotten?' recommendations have included a version of the social in the form of a  model that seeks to identify relations between variables.
+The relevance of recommendations pivots on the apprehension of social order by a model.
 
@@
 
-What counts as a good feature in a world of predictive lists?
-Some things in the data that increases the precision of the model.
-The problem that Tesco faces, however, is that there are too many possible recommendations to be tested.
-They sell around 200,000 products and no customer wants to buy any more than a fraction of a per cent of those products.
-The recommendation list needs to very short.
+Note that neither the infrastructure that marshals the Tesco customer data  - Patel will mention `Hadoop` -- nor the predictive analytic device -- the logistic regression model -- loom very large in the conversion of the recommender system.
+'Most of the gain comes,' Patel observes, 'from thinking of good features.' While this is almost a truism in big data practice (see, for instance, [@Domingos_2012]), what counts as a good feature in a shopping list?
 
-Predictive models, we might say, encounter their own surplus of sensibility in the data as they search for the most accurate predictions.
-A  very brief comment from Patel -- 'we have lots of zeros' -- points to the problem here.
-Each of the millions of Tesco online grocery shoppers has bought some selection of the 200,000 products.
+A  very brief comment from Patel -- 'we have lots of zeros' -- points to the locus of this thinking: the vast, nearly empty matrix of customer-product associations.
+Given that most things in Tesco remain relatively inert in relation to each other, a matrix that records association between individual people and different products is bound to be mostly empty.
+Say Tesco has 1 million online customers.
+Each of online  shopper has bought some selection of the 200,000 products.
 But the product-customer matrix, the basic vector-space in which all recommender systems operate, remain very sparse and unpopulated.
+The customer-product data matrix will be `r format(1e6 * 2e5)` in size.
+Given that any one customer is likely to only have bought 100 different products, the probability matrix will be close to `r format(100-(100/2e5)*100, 2)`% empty.
+The data is, as Patel put it, 'massively unbalanced', and such imbalance would heavily bias recommendation towards commonly and only impersonally relevant suggestions, suggestions that would likely not produce the desired conversion experience.
+Rather than a surplus of sensibility, predictive models, we might say, encounter their own dearth of sensibility in the data.
 
-Say Tesco has 1 million online customers.
-A data matrix for just whether they ever bought a particular product or not will be `r format(1e6 * 2e5)` in size.
-Given that any one customer is like to only have bought 100 different products, the probability matrix will be `r format(100-(100/2e5)*100, 2)`% empty.
-
-```{r sparsity, echo=FALSE, cache=TRUE, fig.margin=FALSE, fig.cap ='The sparsity of predictions' }
+```{r sparsity, echo=FALSE, cache=TRUE, fig.margin=FALSE, fig.cap ='Sparsity of the customer-product association matrix' }
 library(ggplot2)
@@
 ggplot(melted, aes(x = Var2, y = Var1, fill = value)) + geom_tile() +
-    scale_fill_manual(values = c("grey", "black"))
+    scale_fill_manual(values = c("grey", "black")) + xlab('Customer id') + ylab('Product id')
 
@@
 
-Having too many things to list -- this must be happening in many settings -- videos, social networks, products, etc.
--- is not the real problem.
-The problem is that 'massively unbalanced data,' which means that different classes of products are not equally represented.
 Since so many people buy milk, the recommendation system might end up always recommending milk products.
-So the data needs to be 'corrected' by, in this case, random under-sampling or removal of some of the data for common purchases.
+So the data needs to be 'corrected' by, as Patel reports, removing -- under-sampling -- some of the data for common purchases.
 Secondly, even if a recommendation system finds uncommon items that are good recommendations, they should neither be too cheap ('low spend'), nor, as we will see too similar to what the customer has already ordered.
+
+HERE
 

### 2017-02-14 lists_conversion_experience.rmd
doing edits

@@
 
-HERE
+Amidst this imbalance, distortion and deficit of sensibility given in the data, 'thinking of good feature' -- ways of traversing the data -- becomes central.
+For our purposes, the exact details of these features matter less than the way in which propensities become probabilities.
+The dataset, as we will see, will not be practically treated as 'closed,' as Hansen puts it.
+Although it undergoes acts of enclosure (for instance, in the transformation of the customer-product matrix into vector space), these acts are not conclusive.
+'Thinking of good features' potentially re-opens the dataset and sometimes completely transforms it.
+This thinking, which I suggest occurs mostly in quite obscure yet central zones of data practice, is highly mixed.
+It combines experimental techniques, infrastructural dynamics, seasonal and daily temporalities, and above all, an ongoing engagement with probabilities and their fungibilities.
 
-In Hansen's or Badiou's terms, this is one way in which the dataset is closed.
-While 200,000 products is still a closed dataset since it has a defined number of items, undersampling alters the frequency of occurrence of very common items so that the predictive models are not distracted by them.
+##  Experimental settings: A/B allocation and uplift
 
-So the raw data -- the 52 weeks of purchases for each customer -- can be rewoven with the database of Tesco products to increase the density of the data in ways that helps the model.
-Patel mentions some of the 'features' used in the new model but somewhat strangely at this point, her presentation turns toward the difficulties of measuring the success of the model and the infrastructural limitations on making recommendations in ways that are more alive to what is happening right now.
+|Test A|Test B|
+|---|---|
+|Control A|Control B|
 
-## 8. Value: Balancing conversion vs spend
-The value of an ordering
+The relevancy model was deployed experimentally in a random A/B controlled trial on the Tesco website.
+All customers were  allocated to one of four categories as shown in the table.
+In the  A/B trial,  an identical shopping list, or even the same person, might receive different recommendations.
+The randomised application of the model, drawing on protocols for randomised clinical trials first developed in the 1960s, is widely used in social media platforms and hence in all recommender systems.[^238.1] Random allocation of customers to the four categories (Test A, Test B, Control A, Control B) adds another layer of probability to the recommender system in the name of statistical validation of the effects or the 'uplift' of the model.[^122] Ironically,  the effects of a predictive model cannot be known in advance.
+They can only be observed experimentally.
 
-## 9. Metric: precision-at-K
-
-The discussion of imperfection was brief - it incorporated the notion of recommendations weighted by price and the notion of precision (or the proportion of products actually purchased.).
-Patel simply said that this metric was imperfect, but without elaborating why.
-Instead she moved to describe some of the ways in which the recommender system worked in practice.
-
-## 10. Experimental settings: A/B allocation and uplift
-
-![A/B testing](figure/a_b_tests.png)
-
-First of all, the new model was deployed experimentally in a random A/B controlled trial.
-This means that all customers are allocated to one of four categories as shown in the figure.
-For an identical shopping list, or even for the same person, it is possible to receive different recommendations.
-
-This randomising of the application of the model, drawing on protocols for randomised clinical trials first developed in the 1960s, is widely used in social media platforms and hence in all recommender systems.
-It is difficult to gauge how much.
-Facebook reports that it has more than 1 million models operation in its infrastructure.
-[TBA: ref for FBflow]   We could also discuss the integration of experiments into the operational infrastructures at Google.[REF tba]
-
-Random allocation of customers to the four categories (Test A, Test B, Control A, Control B) adds another layer of probability to the recommender system.
-This is done in the name of statistical validation of the effects or the uplift of the model.
-It also suggests that the effects of the model cannot be known in advance -- they can only be observed experimentally.
-Does this experimentality of the list point more to the propensities that Hansen says saturate experience?
+[^238.1]: It is difficult to gauge how much. Facebook reports that it has more than 1 million models operation in its infrastructure. TBA[]
 
@@
 It seeks to statistically validate the effects -- the *uplift* -- of the model on  conversion events by allowing the effects of the model on what people do to be measured.
-The lift or uplift refers to a change in events associated with the same groups of people.
-This effectively is an experiment in inhabiting two different worlds.
-This sets up a feedback look between the predictive model, the logistic regression, and the operation of the recommendation system.[^122] Once the experimental uplift methodology is in place, the recommender system can constantly modify the recommendation list in light of subsequent data on the uplift.
+The lift or uplift refers to conversion events associated with the same groups of people.
+Effectively is an experiment in inhabiting two different worlds, the randomised control trial  sets up a feedback look between the predictive model, the logistic regression, and the world.
+Once the experimental uplift methodology is in place, the recommender system can constantly modify the recommendation list in light of subsequent data on the uplift.
 Predictions, then, are not bound by fixed parameters of the model.
@@
 Market researchers and data miners seem to have started adopting a statistically grounded experimental predictive practice sometimes around 2000.
-See for instance [@Almquist_2001]  Response modelling, propensity modelling and uplift modelling all seek to identify associations between 'treatments' or interventions and the 'Responders', the people  affected by the treatment.
-Interestingly, these
+See for instance [@Almquist_2001]  Response modelling, 'propensity' modelling and uplift modelling all seek to identify associations between 'treatments' or interventions and the 'Responders', the people  affected by the treatment.
 
-## 11. Infrastructure: realtime engine serves up top 5
+## Infrastructure: realtime engine serves up top 5
 
@@
 
-While much discussion of recent media has highlighted its realtime operation, Patel's description of the infrastructure that implements the new model suggested a different temporal patterning.
-While raw data from Tesco Online feeds into the model every hour, it seems that recommendation list for each customer is only generated once a week.
-This makes sense.
+Much discussion of recent media has highlighted its real-time operation (see [@Kittler_1993; @Virilio_1997; @Hansen_2015; @Bollmer_2016]).
+Patel's description of the relevancy model infrastructure suggested a different temporal patterning.
+While raw data from Tesco Online feeds into the model every hour, a recommendation list for each customer is only generated once a week.
 Customers only shop online every few days at the most, and in some cases, only every few weeks.
@@
 
-Patel did not say much else about the infrastructure but the 200 recommendations generated each week are certainly never seen by the customers.
-The realtime engine that actually generates results shown on the Tesco Online website only shows 5 after 'applying business rules.' Again, this might be seen as corroborating Hansen and Badiou's insistence on the necessarily limited grasp of any prediction system on the situation it encounters.
-Even the new model, situated amidst its experimental setup cannot fully encompass the surplus of sensibility it encounters.
+A customer does not see their 200 weekly recommendations.
+The recommender system generates recommendations, but after 'applying business rules, only five are  shown on the Tesco Online website.
+Patel briefly mentioned  specific infrastructural elements such as `hadoop` [@ApacheSoftwareFoundation_2009].
+The logistics of running a predictive model affect listing practices.
+The possibility of adjusting the recommendations for every customer weekly derives from an infrastructure capable of collecting data, assimilating that data to a predictive model able to generate approximately 200 million relevant recommendations, where 'relevance' depends on a matrix of probabilities of associations between people and things that shifts in time.
+Like the global supply chain logistics optimised for flow of commodities [@Tsing_2009; @Nielson_2012], `hadoop` and its legion of variants (`mahout`, `spark`, `hive`, `pig`, `yarn`) operationalise flows of data at an infrastructural rather than analytic scale.
+In many respects, the quickly glossed-over infrastructural deployment of DunnHumby's relevancy model is the primary conversion event: the logistic regression model at the heart of the recommender system is no longer an analytical device but an operational one.
 
-The brief mention of specific infrastructural elements such as `hadoop` is, however, significant.
-It points to the temporality of infrastructures themselves as a component of list practices.
-The possibility of adjusting the recommendations for every customer weekly derives from an infrastructure capable of collecting data, assimilating that data to predictive model able to generate recommendations that are relevant to individuals amongst 200,000 products.
-
-## 12. Similarity, substitution, complementarity and micrological propensities
+## Similarity, substitution, complementarity and micrological propensities
 
@@
 
-What does it matter that recommendations are relevant?
+The propensities of things become probabilities in predictive analytics only to the extent that probabilities themselves become propensities.
+Recommendations are relevant in a recommender system or any other predictive model to the precise extent of their probability.
 Relevance is closely connected to propensities in the predictive condition.
-From DunnHumby's standpoint, the relevance of a recommendation relates directly to conversion.
-From the standpoint of the ontological conversion, and  if Hansen is right, the real or actual fabric of prediction resides in the propensities of things, what he calls the 'complex and diffuse calculus' that only subsequently and provisionally congeals as a conversion event.
-The 'micrological propensities' of things feed directly into an sense of the relevance of recommendations.
+From DunnHumby's standpoint, the relevance of a recommendation relates directly to conversion events.
+Conversion events happen in the world.
+They actualise a propensity that hitherto remained either latent or potential.
+From the standpoint of the ontological conversion between probability and propensity,  prediction stems from the propensities of things, what Hansen describes as the 'complex and diffuse calculus' that only subsequently and provisionally congeals as a conversion event.
+The 'micrological propensities' of things support the relevance of recommendations.
 
 How does the DunnHumby model deal with the micrological propensities of things?
-In her presentation, Patel  spent quite a lot of time addressing the problem of basket similarity in terms of substitutions and complementarities between things.
-She concludes by saying 'basket similarity is the future' and 'complementarities drive conversion.'
+In her presentation, Patel  spent quite a lot of time addressing the problem of 'basket similarity,' the problem that the recommender system might recommend, for good reason, something similar to items already in a customer's basket.
+The recommender system needs to know, or the statistical model needs to know how things are related to each other so that it does not waste recommendations on overly similar items.
+A customer might be willing to substitute a similar item for something they have already chosen, but they are more likely to accept a recommendation for an item that complements items in their list.
 
-Whereas the association rules focus on complementarities -- if you buy bread, you might want milk, if you buy beer, you might want diapers -- the substitutions look for micrological propensities that relate to gradations of preference, and variations in time and place.
+Whereas the association rule-based recommender system focused on thing-thing complementarities -- if you buy bread, you might want milk, if you buy beer, you might want diapers -- the substitutions look for micrological propensities that relate to gradations of preference, and variations in time and place.
 Micrological propensities in the list relate to a particular time and place, amidst a continuously differentiable pathways between things.
 
-How do things associate with each other in  substitutions and complementarities?
-Patel's comments on this are very brief but suggest that the 'Have You Forgotten?' list becomes an increasingly mutable entity, open to many substitutions and re-writings.
-Like the mention of `hadoop`, what Patel said about substitution and complementarity implies extensive but relatively well-established data practices.
-She speaks for instance of a 'design matrix of 14,000 columns' and the use of `L1 regularization` to drive coefficients down to zero.
-This is machine learning talk.
+How could the recommender system sense the possible substitutions and complementaries between things for a given customer, especially since a person's sensibilities and susceptibilities concerning groceries might be shaped by wider social groups, orderings and circumstances?
+Patel's comments on this were very brief but suggest that the 'Have You Forgotten?' list becomes an increasingly mutable entity, open to many substitutions and intensive re-writings.
+Like the mention of `hadoop`, what Patel said about substitution and complementarity draws on extensive but relatively well-established data practices.
+She mentioned,  for instance,  a 'design matrix of 14,000 columns' and the use of '`L1 regularization` to drive coefficients down to zero' in a new feature of the model called 'self-learning substitutes,' a model that examines the recommendation list and predicts substitutes and complementarities for 'all products' in the list.
+The 'self learning substitutes' feature of the relevancy model draws on another data matrix, the product similarity matrix, itself generated from all previous recommendations that have led to conversion events (that is, 'a user clicked as the response').
 What is being modelled here?
-No longer whether to recommend a product to a customer or not, but whether something could be substituted for what is already in the recommendation list.
-
+No longer whether to recommend a product to a customer or not, but whether something could be substitute or complement  what is already in the recommendation list.
+HERE
 ```{r biscuits, echo=FALSE, cache=TRUE, fig.margin=TRUE}

### 2017-02-14 lists_conversion_experience.rmd
doing edits

@@
 In its re-writing of the list, the recommender system effects a conversion experience in which individualised desire, consumption practices, value-chains, infrastructural time-space dynamics, the lives of others and the propensities of things all come into play.
-ing a major contemporary form of ordering, the recommendation list, that we see in many  forms of news, media, entertainment, marketing, advertising.
+A major contemporary form of ordering, the recommendation list, that we see in many  forms of news, media, entertainment, marketing, advertising, can be seen as what the philosopher of science Karl Popper called  a 'crystallisation of propensities' [@Popper_1990, 18].
 
@@
 The 'self learning substitutes' feature of the relevancy model draws on another data matrix, the product similarity matrix, itself generated from all previous recommendations that have led to conversion events (that is, 'a user clicked as the response').
-What is being modelled here?
+A customer's recommendation list might already include `Tesco Ginger Nuts 300g.`   The 'self-learning substitution model' might be able to substitute something in that list that is of higher value, either because it is bigger or more expensive: `Mcvitie's Ginger Moments 400g.` What is being modelled here?
 No longer whether to recommend a product to a customer or not, but whether something could be substitute or complement  what is already in the recommendation list.
-HERE
-```{r biscuits, echo=FALSE, cache=TRUE, fig.margin=TRUE}
-    library(readr)
-    library(knitr)
-    bisc = read_tsv('data/biscuits.csv')
-    kable(bisc[order(bisc$price, decreasing=TRUE)[1:30], c('name','price')])
-```
 
-We can imagine this in terms of biscuits.
-A customer's recommendation list might already include `r bisc$name[10]`.
-The 'self-learning substitution model' might be able to substitute something in that list that is of higher value, either because it is bigger or more expensive.
-
-## 13. Complementarity drives conversion
-
-![propensity_matrix](figure/propensity_conversions.png)
-
-Similarly, to model the potential complementarities   between things entails a matrix of comparisons.
-`Nutella` complements `bananas,` but the converse is not necessarily true since `bananas` need not complement `nutella.`
-
-How should we understand the effect of complementarity?
-Patel says that 'complementarity drives conversion.' That is, the matrix of similarities expressed as numerical associations between products best animates the conversion of recommendations into purchases.
-
-In the similarity matrix, shaped by the history of past recommendations and purchases made by this customer, lie the propensities of things that Hansen (following Popper) attributes to the world.
-Karl Popper speaks of the future as a 'crystallization of propensities' [@Popper_1990,18]
+Similarly, modelling complementarities between things for a particular person entails a matrix of comparisons.
+`Nutella` complements `bananas,` but the converse is not necessarily true since `bananas` need not complement `nutella.` How should we understand the effect of complementarity?
+From the standpoint of selling more things, Patel emphasises that 'complementarity drives conversion.' Bringing the matrix of similarities expressed as numerical associations between products into the model improves the conversion of recommendations into purchases.
+From the standpoint of a social ordering that identifies people and things as propensities, complementarities suggest another possible outcome.
+What does it mean for our sense of social order if the relations between things shift in accordance with our experience of the world?
+At base, there is no ontological transformation here, but an inclusion of some part of the world into the model.
+In the similarity matrix, shaped by the history of past recommendations and purchases made by this customer, the propensities of things drive down the surplus of sensibility associated with what people want.
 
@@
 
-My investigation of lists has been centred around the transformation in their predictive underpinnings.
-How should we think about the list-making practices of these models, databases and computational infrastructures?
-Following Hansen, I have been suggesting that the predictive listing in the form of open-ended recommendation is symptomatic of changes in the distribution of media.
+In their intersection with mundane practices of ordering, recommender systems offer some guidance to the ways that predictive practices affect existing  forms of social order in the world.
+We have seen several types of conversion events: ongoing transformation of models and infrastructures, the experimentalisation  of recommendations, and the broader diffusion of a calculus of probabilities into a fibrous weave of contingencies and conditionalities, warping over time.
+HERE Following Hansen, I have been suggesting that the predictive listing in the form of open-ended recommendation is symptomatic of changes in social order.
 The idea that media should no longer be considered as extensions or prostheses of human communication but the environmental situation in which different propensities play out is increasingly central to contemporary media theory.

### 2017-02-15 lists_conversion_experience.rmd
writing conclusion

@@
 
-This paper investigates practices of list-making associated with online grocery shopping at a  UK and global supermarket chain, Tesco. 
+This paper investigates practices of list-making associated with online grocery shopping at a UK-global supermarket chain, Tesco. 
 It reconstructs an account of how a recommender system suggests a small number of grocery items of personal relevance to each of the millions of online grocery shoppers. 
@@
 I suggest that we can understand devices such as recommender systems as transformations or conversions of the temporal and agential fabric of the important social ordering device, the list.
-While all written lists mix past and future actions, actual and potential events, their predictive re-writing in recommender systems adds a new operational dynamic.
+While all written lists mix past and future actions, actual and potential events, their re-writing by recommender systems adds a new operational dynamic.
 The incompleteness of lists, their propensity to expand or change, becomes more important than their capacity to  enumerate or rank things. 
@@
 
-The shelves of large contemporary supermarket are the endpoint of global logistic supply chains, in all their logistical, value-transforming and brand-mediated hypercomplexity (see [@Nielsen_TBA; @Tsing_2009] on logistics and power are the endpoint of global logistic supply chains, in all their logistical, value-transforming and brand-mediated hypercomplexity (see [@Nielsen_TBA; @Tsing_2009] on logistics and power)) Groceries display a geography of agriculture, industry, transport, communication and financialisation driven by flows of labour and capital.
+The shelves of large contemporary supermarket are the endpoint of global logistic supply chains, in all their logistical, value-transforming and brand-mediated hypercomplexity (see [@Nielson_2012; @Tsing_2008] on logistics and power are the endpoint of global logistic supply chains, in all their logistical, value-transforming and brand-mediated hypercomplexity).
+Groceries imply a geography of agriculture, industry, transport, communication and financialisation driven by flows of labour and capital.
 Encounters between this hyper-complex geography and people, even in the confines of a supermarket, are no simple matter, either for shoppers or for supermarket operators such as Tesco.
-While there are many facets to shopping [@Miller_2013], the shopping list, whether written on the back of an envelope, or saved as a list in an online grocery shopping system, is one intersectional ordering device.
-In a shopping trajectory (trawling along aisles packed with thousands of products, scrolling down screens or searching for particular brands amidst search results), lists inscribe some kind of order that filters or reduces the excessive propensities, claims, dazzle and distraction of commodities.
+While there are many facets to shopping [@Miller_2013], the shopping list, whether written on the back of an envelope, or saved as a list in an online grocery shopping system, is one intersectional ordering device that configures a universe of possible references, and a teeming multitude of propensities, into a possible order (in both senses of that term: an ordering and an imperative).
+As people shop, either by  trawling along aisles packed with thousands of products, or scrolling down screens or searching for particular brands amidst search results, lists inscribe some kind of order that filters or reduces the excessive propensities, claims, dazzle and distraction of commodities.
 
-Hand-written shopping lists undoubtedly have ongoing importance and display interesting mixed ordering practices (see the list of handwritten shopping lists at [Grocery List](http://www.grocerylist.org)), online shopping lists take shape at the intersection of web and internet infrastructures, supply chain logistics, individualised practices and *habitus,* and increasingly, the predictive operations of  recommender systems.
-Just as the aisles and shelves of a supermarket present a commercially ordered geography of the world,  online shopping recommender systems generate lists that seek to align people to commodities they otherwise might have little relation to.
+While hand-written shopping lists undoubtedly have ongoing importance and display interestingly mixed ordering practices (see the list of handwritten shopping lists at [Grocery List](http://www.grocerylist.org)), online shopping lists take shape at the intersection of web and internet infrastructures, supply chain logistics, individualised practices and *habitus,* and increasingly, the predictive operations of  recommender systems.
+Just as the aisles and shelves of a supermarket present a commercially-ordered geography of the world,  online shopping recommender systems generate lists that seek to align people to commodities they otherwise might have little relation to.
+
+[HERE]
 
@@
 
-The items on this list could be recommendations. They are only a few of them: five. The question: which five items and in what order?
 ```{r shopping-hand}
@@
 
+The items on this list could be recommendations.
+They are only a few of them: five.
+The question is: which five items and in what order?
 In terms of content, there are many similarities between shopping lists such as Figure \@ref(fig:shopping-hand)  -- a dozen lines scrawled on the back of a used envelope stuck on the refrigerator door  --  and the five ranked items shown to a customer on an online shopping site (as shown in Table \@ref(tab:random-basket)).
@@
 As literary scholars suggest (see [@Mainberger_2003]), although lists have often been de-valued as literary forms, list-making often appears, even in literary form, whenever writing seeks to address, name, group or evoke totality, profusion, excess or abundance.
+More recent sociological work on lists (see [@deGoede_2016] for an oveview) confirms the potency of lists as ordering devices.
 From both  anthropological and literary perspectives, shopping lists have excellent reasons to exist: they are deeply rooted in urban social formations historically and culturally.
@@
 
+HERE 
 All lists are forms of order.
@@
 
-Mark Hansen has recently suggested,
+The media theorist Mark Hansen has recently suggested that
 
@@
 
-The question exercising much contemporary work on data, on media, platforms and finance pivots on the 'micrological propensities' that might be re-ordering lists.
-Hansen's argues that these propensities make probabilities into something real, some actually in things:
+This assertion resonates with many accounts of big data.
+The expectation exercising much contemporary work on data, on media, platforms and finance pivots on 'micrological propensities' that might only become evident given enough data.
+Hansen argues that an ontological transformation links the micrological propensities of things in the world to prediction:
 
@@
 
-Hansen predicates prediction as an 'ontological transformation' transforming the relation between probabilities and things.
-The practical and indeed empirical  question is whether such transformations in probability that bring 'the propensities of things'  can be detected and articulated in the prosaic setting of shopping lists and  shopping baskets.
-'Conversion' might be a preferable term for these changes.
-It highlights a range of re-orientations in subjects, experience, things, numbers and structures  entailed more than the subjective 'understood-to-be-expressed' formulation that Hansen emphasises.
+Hansen predicates prediction as an 'ontological transformation' thata ffects probabilities.
+It links probabilities as calculations to propensities or the mutable agencies of things.
+The practical and indeed empirical  question is whether such transformations in probability that bring 'the propensities of things'  can be detected and articulated in the prosaic setting of shopping lists and  recommendations.
+I will argue that 'conversion' is a preferable term for these changes.
+It highlights  re-orientations in subjects, experience, things, numbers and structures that ranges more widely than Hansen's subjective 'understood-to-be-expressed' formulation.
 
-## DunnHumby's work for Tesco
+## Conversion events: DunnHumby's work for Tesco
 
-In one of the many industry-meets-academia events occurring in increasingly data science-oriented higher education institutions in the UK, speakers from industry, government and commerce described their work.[^11.1] Shreena Patel, a PhD graduate in statistics and operations research,  works as a data scientist for DunnHumby, a well-known [ customer science company ](https://www.dunnhumby.com/)  focusing on online grocery shopping at Tesco.
+In one of the many industry-meets-academia events occurring in increasingly data science-oriented higher education institutions in the UK, speakers from industry, government and commerce described their work with predictive modles.[^11.1] Shreena Patel, a PhD graduate in statistics and operations research,  works as a data scientist for DunnHumby, a well-known [ customer science company ](https://www.dunnhumby.com/).
+Her work focuses on online grocery shopping at the supermarket chain Tesco.
 Speaking to an audience of statisticians and operations researchers, Patel   focused on the models underlying  shopping list recommendations.
@@
 
-## 7. Sparsity: output 20 recommendations from 200,000 products
+## Sparsity: output 20 recommendations from 200,000 products
 
@@
 In their intersection with mundane practices of ordering, recommender systems offer some guidance to the ways that predictive practices affect existing  forms of social order in the world.
+Rather than directly positing an ontological transformation, I have explored some of the practices of conversion associated with a shift from a data-mining rule-based system to a predictive recommender system.
+The DunnHumby model and the Tesco 'Have You Forgotten?' recommendations present many of the traits of the conversion processes associated with big data.
 We have seen several types of conversion events: ongoing transformation of models and infrastructures, the experimentalisation  of recommendations, and the broader diffusion of a calculus of probabilities into a fibrous weave of contingencies and conditionalities, warping over time.
-HERE Following Hansen, I have been suggesting that the predictive listing in the form of open-ended recommendation is symptomatic of changes in social order.
-The idea that media should no longer be considered as extensions or prostheses of human communication but the environmental situation in which different propensities play out is increasingly central to contemporary media theory.
 
-Hansen refers to an ontological transformation in which predictive media or recommender systems would be seen as the closed realisation of the actually open propensities of the world.
-By understanding the world in terms of propensities, he argues, we have a change to engage with contemporary predictive media.
+Following Hansen, I have been suggesting that the predictive listing in the form of open-ended recommendation is symptomatic of changes in social order. 
+Predictive media or recommender systems would be seen as the closed realisation of the  open propensities of the world.
+By understanding the world in terms of propensities, he argues, we have a chance  to engage with contemporary predictive processes in ways that both acknowledge their relentless and exhausting pressure to consume more and retain some chance of inhabiting the world differently.
 
-I have not explored the Whitehead-derived philosophy of propensities in the world that he develops there, only because I am not sure that we need speculative theory to grapple with the material practices of prediction.
-Already in the DunnHumby model and the Tesco 'Have You Forgotten?' recommendations, many of the traits of predictive processes are at work.
+I diverge from Hansen in accounting for this change.
+Rather than looking for a speculative theory that sets out a world in which prediction would be the closed form of an ontologically open world ('the propensities of things'), we might understand the predictive conversion as constitutively incomplete.
+Shopping lists  exemplify this incompleteness and ineradicable openness of prediction.
+The constitutive incompleteness and openness of lists are very closely entangled with problems of social order amidst predictive operations.
+Lists undergoing a 'conversion' experience as they become more probabilistic, but their heterogeneous character as operations caught up in models, infrastructures, experiments, and the flow of commodities means that they can never form a closed set or dataset.
+The problem of 'constructing good features,' a problem that any predictive model must address, persists as the site of experiment with structures that lack closure.
+A constitutive openness to new associations, new combinations, new groupings or actors is integral to the conversion event.
 
-I diverge from Hansen in how I account for this change. Rather than looking for a speculative theory that sets out a world in which prediction would be the closed form,
+In their account of the Netflix recommender system, Ted Hallinan and Peter Striphas make the following argument:
 
-- the **constitutive incompleteness** and openness of lists are very closely entangled with problems of social order under predictive conditions or amidst predictive operations. 
-- shopping lists have complex temporal-material structure, since they mix prediction, infrastructural constraints, the individual characteristics of list-makers, statistical validations, database architectures and above all the **surplus sensibility** associated with  circulation of commodities.
-- shopping lists undergoing a **conversion experience** as they become more probabilistic, but the idea that probabilities are the closed form of propensities does not account for the distribution of probabilities in relations between things, between people, between people and things, and between things and people.
-- lists undergoing a 'conversion' experience as they become more probabilistic, but their heterogeneous character as operations caught up in models, infrastructures, experiments, and the flow of commodities means that they can never form a closed set or dataset.
-Amidst prediction they become constitutively incomplete.
+> These systems, and more importantly their algorithms, play a critical role in deciding which articles (or parts thereof) gain admission to the cultural realm, and in what form.
+Their doing so thus points in the direction of another universe of reference—a court of algorithmic appeal in which objects, ideas, and practices are heard, cross-examined, and judged independently, in part, of human beings [@Hallinan_2014, 13]
+
+The 'universe of reference' includes objects, ideas and practices mingle.
+Something similar happens in the Dunnhumby relevance model.
+In the customer-product association matrix, in the product-product similarity matrix, in the 14,000 variables included in the logistic regression model, a 'universe of reference' takes shape.
+The shopping list, as a result, take on a complex temporal-material structure that mixes prediction, infrastructural constraints, the individual characteristics of list-makers, statistical validations, database architectures and above all the **surplus sensibility** associated with  circulation of commodities.
+Shopping lists undergoing a **conversion experience** as they become more probabilistic, but the idea that probabilities are the closed form of propensities does not account for the distribution of probabilities in relations between things, between people, between people and things, and between things and people.
 

### 2017-02-16 lists_conversion_experience.rmd
adding popper and checking

@@
 opts_chunk$set(message=FALSE)
-opts_chunk$set(error=TRUE)
+opts_chunk$set(echo=FALSE)
+opts_chunk$set(error=FALSE)
 opts_chunk$set(cache=TRUE)
@@
 
-The shelves of large contemporary supermarket are the endpoint of global logistic supply chains, in all their logistical, value-transforming and brand-mediated hypercomplexity (see [@Nielson_2012; @Tsing_2008] on logistics and power are the endpoint of global logistic supply chains, in all their logistical, value-transforming and brand-mediated hypercomplexity).
+The shelves of large contemporary supermarket are the endpoint of global logistic supply chains, in all their logistical, value-transforming and brand-mediated hypercomplexity (see [@Neilson_2012; @Tsing_2009] on logistics and power are the endpoint of global logistic supply chains, in all their logistical, value-transforming and brand-mediated hypercomplexity).
 Groceries imply a geography of agriculture, industry, transport, communication and financialisation driven by flows of labour and capital.
@@
 Just as the aisles and shelves of a supermarket present a commercially-ordered geography of the world,  online shopping recommender systems generate lists that seek to align people to commodities they otherwise might have little relation to.
-
-[HERE]
 
@@
 
-The presentation I saw was not really surprising or exceptional.
+The Tesco presentation I saw was not really surprising or exceptional.
 It was a presentation by a data scientist to data scientists, describing an implementation of a predictive model as well as some of the problems and potentials of the predictive model.
-I see the presentation itself as providing materials for an argument about our experience of predictive ordering.
+The presentation provides some materials for an exploration of the form, techniques, and to certain extent, experience of predictive ordering.
 In this paper, I re-present the presentation using a mixture of theoretical and practical resources.
-The theory is largely recent media theory and its accounts of prediction in media.
-The practical resources are that of coding and predictive models.[^12]
+The theory is largely recent media theory and its accounts of prediction.
+The practical resources are that of coding and predictive models, and online grocery data.[^12]
 
@@
 From a data science perspective, conversion events, when they occur, confirm the power of predictive analytics to change the world in some way.
-From a sociological perspective, constructing and configuring recommender system in the name of conversion events, the recommender system effects an epistemic-operational transformation central to big data as an operational formation that responds to an ongoing control crisis [@Beniger_1986].
-This central transformation is that the idea that data affords apprehension, visibility and management of associative propensities in the world.
-Recommender systems, then, are concerned with the conversion of social order.
+From a sociological perspective, matters are less obvious.
+Constructing and configuring recommender system in the name of conversion events, the recommender system effects an epistemic-operational transformation central to big data.
+We might regard the recommender system as an operational formation that responds to an ongoing control crisis associated with information [@Beniger_1986].
+We might also explore these conversion events in terms of how predictive analytics happens in the world, how the tendency of things to realize certain events -- what the philosopher science Karl Popper terms the propensities of things [@Popper_1990, 11] -- happens in and through data practices.
+Data in concert with calculation, models, infrastructures, and experiments, affords apprehension, visibility and management of associative propensities in the world.
+Recommender systems, then, are vitally concerned not just with conversion events, but also with the what counts as order in the world.
 
-## Lists in and as data
+## Lists and order
 
@@
 
-HERE 
 All lists are forms of order.
@@
 
-![People add to list](figure/online_lists_in_practice.png)
-
 How could we characterise the shift from ClubCard data mining to online grocery markets?
@@
 A slightly later review, [@Schafer_2001], diagrams an augmented 'natural social process' with a range of elements, technologies, inputs and outputs, with varying degrees of personalization.
-
-![Early e-commerce recommender system architecture](figure/schafer_recommendation_2001.png)
 
@@
 
-    ggplot(tesco[tesco$actual>50,], aes(x=labels, y=actual, fill=labels)) +geom_bar(stat='identity') + coord_flip() + ggtitle('Tesco grocery items with more than 50 products') + theme(legend.position="none", axis.text=element_text(size=9))
+    ggplot(tesco[tesco$actual>50,], aes(x=labels, y=actual, fill=labels)) +geom_bar(stat='identity') + coord_flip() + ggtitle('Tesco grocery items with more than 50 products') + theme(legend.position="none", axis.text=element_text(size=7))
 ```
 
-Items in the`Groceries` dataset proliferate into a Tesco's list of branded variations.
+Items in the `Groceries` dataset proliferate into a Tesco's list of branded variations.
 The 169 items of the `Groceries` dataset  expand into roughly `r total` Tesco items.
@@
 > [S]urrounding any delimited predictive system is a larger field of data -- what I elsewhere call a “surplus of sensibility” -- that, viewed speculatively, indexes the causal efficacy of the total situation within which this delimited system operates.
-(Effectively, the latter gains its reliability from closing off this larger surplus of sensibility, thereby transforming an always excessive propensity into a [provisionally] closed dataset.) Because it affords data that exceed whatever any given predictive system might include, the data of the world’s causal efficacy -- the data constituting its real potentiality -- always and in principle facilitates knowledge that cannot be restricted to any particular agenda [@Hansen_2015, 126-7].
+(Effectively, the latter gains its reliability from closing off this larger surplus of sensibility, thereby transforming an always excessive propensity into a [provisionally] closed dataset.) Because it affords data that exceed whatever any given predictive system might include, the data of the world’s causal efficacy -- the data constituting its real potentiality -- always and in principle facilitates knowledge that cannot be restricted to any particular agenda [@Hansen_2015a, 126-7].
 
@@
 library(reshape2)
-p = as.logical(rbinom(n=2e5, size=1, prob=0.001))
+p = as.logical(rnorm(n=2e5, size=1, prob=0.01))
 m = (matrix(p, nrow=1000, byrow=TRUE))
-
 melted <- melt(m)
@@
 ggplot(melted, aes(x = Var2, y = Var1, fill = value)) + geom_tile() +
-    scale_fill_manual(values = c("grey", "black")) + xlab('Customer id') + ylab('Product id')
-
+    scale_fill_manual(values = c("grey", "red")) + xlab('Customer id') + ylab('Product id') + theme(legend.position='none')
 ```
@@
 
-![Realtime infrastructure and its limits](figure/hadoop.png)
-
 Much discussion of recent media has highlighted its real-time operation (see [@Kittler_1993; @Virilio_1997; @Hansen_2015; @Bollmer_2016]).
@@
 ## Similarity, substitution, complementarity and micrological propensities
-
-![Basket similarity](figure/patel_basket_similarity.png)
 
@@
 The DunnHumby model and the Tesco 'Have You Forgotten?' recommendations present many of the traits of the conversion processes associated with big data.
-We have seen several types of conversion events: ongoing transformation of models and infrastructures, the experimentalisation  of recommendations, and the broader diffusion of a calculus of probabilities into a fibrous weave of contingencies and conditionalities, warping over time.
+We have seen several types of conversion events: ongoing transformation of models and infrastructures, the experimentalisation  of recommendations, and the broader diffusion of a calculus of probabilities into a fibrous weave of propensities and conditionalities, warping over time.
+The predictive assemblages respond to problems of logistical ensembles more generally.
 
-Following Hansen, I have been suggesting that the predictive listing in the form of open-ended recommendation is symptomatic of changes in social order. 
-Predictive media or recommender systems would be seen as the closed realisation of the  open propensities of the world.
-By understanding the world in terms of propensities, he argues, we have a chance  to engage with contemporary predictive processes in ways that both acknowledge their relentless and exhausting pressure to consume more and retain some chance of inhabiting the world differently.
+Following Hansen, I have been suggesting that the predictive listing in the form of open-ended recommendation changes social order.
+It alters the way in which repetition or reiteration, exemplified in the list, participating in real situations, and their possibilities.
+By following Popper in understanding the world in terms of propensities, we have a chance  to engage with contemporary predictive processes in ways that both acknowledge their relentless and exhausting pressure to consume more and retain some chance of inhabiting the world differently.
 
-I diverge from Hansen in accounting for this change.
-Rather than looking for a speculative theory that sets out a world in which prediction would be the closed form of an ontologically open world ('the propensities of things'), we might understand the predictive conversion as constitutively incomplete.
-Shopping lists  exemplify this incompleteness and ineradicable openness of prediction.
+I diverge from Hansen in making sense of predictive practice.
+Rather than understanding prediction as the closed form of an ontologically open world ('the propensities of things'), the DunnHumby relevance model, in all its statistical, infrastructural, organisational and banal complexity, suggest that predictive conversion remains constitutively incomplete.
+Predictive practice abounds in propensities.
+Ironically, shopping lists  exemplify this incompleteness and ineradicable openness of prediction practice.
 The constitutive incompleteness and openness of lists are very closely entangled with problems of social order amidst predictive operations.

### 2017-02-16 render_tufte.r

@@ -1,6 +1,6 @@
-library(tufte)
 library(rmarkdown)
 library(tint)
+library(tufte)
 #render('lists_conversion_experience.rmd',output_format=tufte_handout())
 #render('lists_conversion_experience.rmd',output_format=ioslides_presentation)
 render('lists_conversion_experience.rmd',output_format=tintPdf())

### 2017-02-17 lists_conversion_experience.rmd
tidying up; adding bit of Miller

@@
 bibliography: /home/mackenza/ref_bibs/uni.bib
+bibliostyle: 'harvard'
 link-citations: yes
@@
 
-```{r setup, cache=FALSE, include=FALSE}
+```{r setup2, cache=FALSE, include=FALSE}
 library(knitr)
@@
 It reconstructs an account of how a recommender system suggests a small number of grocery items of personal relevance to each of the millions of online grocery shoppers. 
-The recommender system  aim to up-sell and cross-sell some of the 250,000 items in the supermarket's inventory. 
 I suggest that we can understand devices such as recommender systems as transformations or conversions of the temporal and agential fabric of the important social ordering device, the list.
@@
 These recommendations are the product of a recommender system, an important category of operational device in big data (see for instance [@Striphas_2015] for analysis of Netflix recommendation; [@Morris_2015]  or [@Seaver_2015] for an account of music recommendations]).
-The question 'have you forgotten?' is followed by some  grocery items that could have been or are usually on a shopping list.
-The title of the suggestions is a bit misleading: the recommender system, as we will see, is not concerned with forgetting, with the many slips and oversights associated with shopping lists,  but rather with items that you didn't know what should be on your list, perhaps because you have never thought of putting them there.
+The question 'have you forgotten?' is followed by a list of some  grocery items that could have been or are usually on a shopping list.
+The title of the suggestions is a bit misleading: the recommender system, as we will see, is not concerned with forgetting, with the many slips and oversights associated with shopping lists,  but rather with items that you didn't know should be on your list, perhaps because you have never thought of putting them there.
 
@@
 Groceries imply a geography of agriculture, industry, transport, communication and financialisation driven by flows of labour and capital.
-Encounters between this hyper-complex geography and people, even in the confines of a supermarket, are no simple matter, either for shoppers or for supermarket operators such as Tesco.
-While there are many facets to shopping [@Miller_2013], the shopping list, whether written on the back of an envelope, or saved as a list in an online grocery shopping system, is one intersectional ordering device that configures a universe of possible references, and a teeming multitude of propensities, into a possible order (in both senses of that term: an ordering and an imperative).
+Encounters between this hyper-complex geography and people, even in the familiar confines of a supermarket, are no simple matter, either for shoppers or for supermarket operators such as Tesco.
+While there are many facets to shopping (see [@Miller_2012] for an account of shopping as a practice that addresses discrepancies between normative and actual social order, and as a sacrificial ritual that instates social order), the shopping list, whether written on the back of an envelope, or saved as a list in an online grocery shopping system, is one intersectional ordering device that configures a universe of possible references, and a teeming multitude of propensities, into a possible order (in both senses of that term: an ordering and an imperative).
 As people shop, either by  trawling along aisles packed with thousands of products, or scrolling down screens or searching for particular brands amidst search results, lists inscribe some kind of order that filters or reduces the excessive propensities, claims, dazzle and distraction of commodities.
 
-While hand-written shopping lists undoubtedly have ongoing importance and display interestingly mixed ordering practices (see the list of handwritten shopping lists at [Grocery List](http://www.grocerylist.org)), online shopping lists take shape at the intersection of web and internet infrastructures, supply chain logistics, individualised practices and *habitus,* and increasingly, the predictive operations of  recommender systems.
-Just as the aisles and shelves of a supermarket present a commercially-ordered geography of the world,  online shopping recommender systems generate lists that seek to align people to commodities they otherwise might have little relation to.
+While hand-written shopping lists undoubtedly have ongoing importance and display interestingly mixed ordering practices (see the list of handwritten shopping lists at [Grocery List](http://www.grocerylist.org)), online shopping lists are written and re-written at the intersection of web and internet infrastructures, supply chain logistics, individualised practices and *habitus,* and increasingly, the predictive operations of  recommender systems.
+Just as the aisles and shelves of a supermarket present a commercially-ordered geography of needs and desires,  online shopping recommender systems generate lists that seek to align people to commodities they otherwise might have little relation to.
 
@@
 We might regard the recommender system as an operational formation that responds to an ongoing control crisis associated with information [@Beniger_1986].
-We might also explore these conversion events in terms of how predictive analytics happens in the world, how the tendency of things to realize certain events -- what the philosopher science Karl Popper terms the propensities of things [@Popper_1990, 11] -- happens in and through data practices.
+We might also explore these conversion events in terms of how predictive analytics happens in the world, how the tendency of things to realize certain events -- what Popper terms the 'propensities of things' [@Popper_1990, 11] -- happens in and through data practices.
 Data in concert with calculation, models, infrastructures, and experiments, affords apprehension, visibility and management of associative propensities in the world.
@@
 
-## Lists and order
+## Lists and orders
 
@@
 As literary scholars suggest (see [@Mainberger_2003]), although lists have often been de-valued as literary forms, list-making often appears, even in literary form, whenever writing seeks to address, name, group or evoke totality, profusion, excess or abundance.
-More recent sociological work on lists (see [@deGoede_2016] for an oveview) confirms the potency of lists as ordering devices.
+More recent sociological work on lists (see [@deGoede_2016] for an overview) confirms the potency of lists as ordering devices.
 From both  anthropological and literary perspectives, shopping lists have excellent reasons to exist: they are deeply rooted in urban social formations historically and culturally.
@@
 
-Hansen predicates prediction as an 'ontological transformation' thata ffects probabilities.
+Hansen predicates prediction as an 'ontological transformation' that affects probabilities.
 It links probabilities as calculations to propensities or the mutable agencies of things.
@@
 
-In one of the many industry-meets-academia events occurring in increasingly data science-oriented higher education institutions in the UK, speakers from industry, government and commerce described their work with predictive modles.[^11.1] Shreena Patel, a PhD graduate in statistics and operations research,  works as a data scientist for DunnHumby, a well-known [ customer science company ](https://www.dunnhumby.com/).
+In one of the many industry-meets-academia events occurring in increasingly data science-oriented higher education institutions in the UK, speakers from industry, government and commerce described their work with predictive models.[^11.1] Shreena Patel, a PhD graduate in statistics and operations research,  works as a data scientist for DunnHumby, a well-known [ customer science company ](https://www.dunnhumby.com/).
 Her work focuses on online grocery shopping at the supermarket chain Tesco.
@@
 Patel's presentation was not sociological.
-In order to explore the operational environment she described, and its predictive operations, I will reconstruct key features of her presentation using sample data drawn from the Tesco API ([Application Programmer Interface](https://devportal.tescolabs.com/docs/services/56c73b1bf205fd0ed81dbe7a/operations/56d5b2bee2813e0a0053c47c), [@Tesco.com_2017])  A sample of the most highly priced items in Tesco stores is shown in Table  \@ref(tab:expensive-groceries) and a range of data science, statistics and graphics libraries implemented in the widely-used statistical programming language `R` [@RDevelopmentCoreTeam_2010].[^11.2]
+In order to explore the operational environment she described, and its predictive operations, I will reconstruct key features of her presentation using sample data drawn from the Tesco API ([Application Programmer Interface](https://devportal.tescolabs.com/docs/services/56c73b1bf205fd0ed81dbe7a/operations/56d5b2bee2813e0a0053c47c), [@Tesco.com_2017])   and a range of data science, statistics and graphics libraries implemented in the widely-used statistical programming language `R` [@RDevelopmentCoreTeam_2010].
 
@@
 The traffic between higher education and data analytics in industry is intense and flows in several forms: people funding, research findings, software and technical devices (code) and training.
-
-[^11.2]:
 
@@
 By virtue of their sheer abundance, the plot cannot easily separate the many items that Tesco sells, especially in the price range £1-£10.
+A sample of the most highly priced items in Tesco stores is shown in Table  \@ref(tab:expensive-groceries)
 
@@
 
-I respond to Patel's presentation in two ways.
+Patel's presentation presents two opportunities.
 First, by recounting, contextualising and commenting on some of the main steps in making the 'Have your forgotten?' list  as a example of the predictive sense-making done by data scientists and customer analytics teams working with transactional data in many commercial settings.
 Patel mentioned many of these steps only fleetingly in the presentation, for they are largely taken for granted as part of predictive analytic practice.
-Second, by trying to step into the list-making practices  as an occasion to apprehend in the wider conversions they signal.
+What stands out from the presentation is the normality of the recommender system: it is part of a long-standing and ongoing transformation or conversion of grocey shopping.
+Second, Patel's explanation of the renovation of a recommender system offered its audience a chance to step into the list-making practices situated at the centre of a mundane situation, grocery shopping.
 As a sociological response to a data scientist presenting a model to an audience of statisticians and operations researchers, my interest lies less in specific technical innovations and more on how the ordering work done by recommender systems relates to problems of social order more generally.
@@
 The main narrative of Patel's presentation concerned a shift from a well-established demographically-targeted data mining model for making recommendations to a predictive, probabilistic, personally relevant  re-writing of the shopping list in almost-realtime.
-The changes Patel described are increasingly widespread.
-While they are configured in Tesco-specific ways by DunnHumby (and this reflects a longer history), they are also broadly symptomatic.
-Tesco is the largest supermarket chain in UK, famous for its customer loyalty and targeted marketing programme known as 'Tesco  Clubcard,' which started in 1991.
-DunnHumby -- founded by operations researchers Edwina Dunn and Clive Humby -- is famous for having convinced the CEO of Tesco sometime in 1991 that a loyalty card program could change the supermarket chain's relationship to its customers.
+The changes Patel described are increasingly widespread and common.
+While they are configured in Tesco-specific ways by DunnHumby (and this reflects a longer history), they are also broadly typical.
+Tesco is the largest supermarket chain in UK, famous for a customer loyalty and targeted marketing programme known as 'Tesco  Clubcard,' which started in 1991.
+DunnHumby -- founded by operations researchers Edwina Dunn and Clive Humby -- is said to have convinced the CEO of Tesco sometime in 1991 that a loyalty card program could change the supermarket chain's relationship to its customers.
 Clive Humby's academic publications  are hard to track down, but an early paper given at the _Conference of Young Operational Researchers_ in Nottingham in 1984 suggests the direction that he, DunnHumby and later Tesco would take in constructing lists [@OKeefe_1984].
@@
 
-```{r arules, cache=TRUE, messages=FALSE, echo=FALSE, fig.margin=TRUE}
+```{r arules-ex, cache=TRUE, message=FALSE, echo=FALSE, fig.margin=TRUE}
 library(knitr)
@@
 options(digits=2)
-rules<- apriori(data=Groceries, parameter=list(supp=0.001,conf = 0.15,minlen=2), appearance = list(default="rhs",lhs="whole milk"), control = list(verbose=F))
+rules<- apriori(data=Groceries, parameter=list(supp=0.001,conf = 0.15,minlen=2), appearance = list(default="rhs",lhs="whole milk"), control = list(verbose=FALSE))
 rules_sorted <- sort(rules, decreasing=TRUE,by="confidence")
@@
 The `apriori` algorithm finds sets of items that commonly occur together in transactions.
-These commonly occurring sets are expressed as 'association rules.' For instance, using the `Groceries` dataset in the `R` package `arules,` we can readily see how Walmart *might* have discovered   the fabled association between diapers and beer.[^149.1] Figure \@ref(fig:arules) shows how often common items appear.
+These commonly occurring sets are expressed as 'association rules.' For instance, using the `Groceries` dataset in the `R` package `arules,` we can readily see how Walmart *might* have discovered   the fabled association between diapers and beer.[^149.1] Figure \@ref(fig:arules-ex) shows how often common items appear.
 `Whole milk` appears most frequently.
@@
 As the product of a  counting practice, the association rules define sets of items or shopping baskets in terms of decreasing strength.
-Table \@ref(tab:arules) shows that the first rule has a stronger association than the second.
+Table \@ref(tab:arules-ex) shows that the first rule has a stronger association than the second.
 Milk and vegetables is more likely an association than milk and buns.
@@
 
-```{r tesco, echo=FALSE, cache=TRUE, fig.cap ='Tesco surplus',  fig.margin=TRUE}
+```{r tesco-ex, echo=FALSE, cache=TRUE, fig.cap ='Tesco surplus',  fig.margin=TRUE}
     library(ggplot2)
@@
     tesco$actual = as.integer(tesco$actual)
-
     ggplot(tesco[tesco$actual>50,], aes(x=labels, y=actual, fill=labels)) +geom_bar(stat='identity') + coord_flip() + ggtitle('Tesco grocery items with more than 50 products') + theme(legend.position="none", axis.text=element_text(size=7))
@@
 Items in the `Groceries` dataset proliferate into a Tesco's list of branded variations.
-The 169 items of the `Groceries` dataset  expand into roughly `r total` Tesco items.
+The 169 items of the `Groceries` dataset  expand into roughly `r total` Tesco items (see Figure \@ref(fig:tesco-ex)).
 This is, I would suggest, a form of propensity in Hansen's terms, albeit one that occurs in, or very close to, the recommender system rather than  'in the world.' The multiple listed  in the `Grocery` dataset becomes more open in this setting through new impersonal intersections that must be bounded and ordered.
@@
 
-```{r sparsity, echo=FALSE, cache=TRUE, fig.margin=FALSE, fig.cap ='Sparsity of the customer-product association matrix' }
+```{r sparsity-ex, echo=FALSE, cache=TRUE, fig.margin=FALSE, fig.cap ='Sparsity of the customer-product association matrix' }
 library(ggplot2)
 library(reshape2)
-p = as.logical(rnorm(n=2e5, size=1, prob=0.01))
+p = as.logical(rbinom(n=2e5, size=20,0.001 ))
 m = (matrix(p, nrow=1000, byrow=TRUE))
@@
 
-## Infrastructure: realtime engine serves up top 5
+## Infrastructure: the realtime engine serves
 
@@
 The DunnHumby model and the Tesco 'Have You Forgotten?' recommendations present many of the traits of the conversion processes associated with big data.
+Rather than discontinuous changes, the conversion of the recommender system via a relevance model suggests an ongoing problematization of propensities.
 We have seen several types of conversion events: ongoing transformation of models and infrastructures, the experimentalisation  of recommendations, and the broader diffusion of a calculus of probabilities into a fibrous weave of propensities and conditionalities, warping over time.
@@
 
-Following Hansen, I have been suggesting that the predictive listing in the form of open-ended recommendation changes social order.
-It alters the way in which repetition or reiteration, exemplified in the list, participating in real situations, and their possibilities.
+I have been suggesting that predictive listing in the form of open-ended recommendation changes social order.
+It alters the way in which repetition or reiteration, exemplified in the shopping list, can also participate in change and emergent order.
 By following Popper in understanding the world in terms of propensities, we have a chance  to engage with contemporary predictive processes in ways that both acknowledge their relentless and exhausting pressure to consume more and retain some chance of inhabiting the world differently.
 
-I diverge from Hansen in making sense of predictive practice.
-Rather than understanding prediction as the closed form of an ontologically open world ('the propensities of things'), the DunnHumby relevance model, in all its statistical, infrastructural, organisational and banal complexity, suggest that predictive conversion remains constitutively incomplete.
+I diverge from some understandings of predictive practice.
+Rather than understanding prediction as the closed form of an ontologically open world ('the propensities of things'), the DunnHumby relevance model, in all its statistical, infrastructural, organisational and banal complexity, suggests that predictive conversion remains constitutively incomplete.
 Predictive practice abounds in propensities.
@@
 The 'universe of reference' includes objects, ideas and practices mingle.
-Something similar happens in the Dunnhumby relevance model.
+Something similar happens to shopping in the Dunnhumby relevance model.
 In the customer-product association matrix, in the product-product similarity matrix, in the 14,000 variables included in the logistic regression model, a 'universe of reference' takes shape.
-The shopping list, as a result, take on a complex temporal-material structure that mixes prediction, infrastructural constraints, the individual characteristics of list-makers, statistical validations, database architectures and above all the **surplus sensibility** associated with  circulation of commodities.
+The shopping list, the commonplace inscription of normative and actual social order, takes on a complex temporal-material structure that mixes prediction, infrastructural constraints, the individual characteristics of list-makers, statistical validations, database architectures and above all the **surplus sensibility** associated with  circulation of commodities.
 Shopping lists undergoing a **conversion experience** as they become more probabilistic, but the idea that probabilities are the closed form of propensities does not account for the distribution of probabilities in relations between things, between people, between people and things, and between things and people.

### 2017-03-16 lists_conversion_experience.rmd
log update

@@
 tufte::tufte_handout: default
+fontsize: 12pt
 citation_package: biber

### 2017-03-16 lists_conversion_experience.rmd
20 edits added in

--- 2017-02-17 lists_conversion_experience.rmd
@@
 ---
-title: "'Have you forgotten?' Conversion experiences in shopping lists"
+title: "Probabilities and propensities: conversion experiences in online grocery shopping"
 author: "Adrian Mackenzie"
@@
 
-This paper investigates practices of list-making associated with online grocery shopping at a UK-global supermarket chain, Tesco. 
-It reconstructs an account of how a recommender system suggests a small number of grocery items of personal relevance to each of the millions of online grocery shoppers. 
-I suggest that we can understand devices such as recommender systems as transformations or conversions of the temporal and agential fabric of the important social ordering device, the list.
-While all written lists mix past and future actions, actual and potential events, their re-writing by recommender systems adds a new operational dynamic.
-The incompleteness of lists, their propensity to expand or change, becomes more important than their capacity to  enumerate or rank things. 
-The conversion of lists from finite to in-definite in recommender systems suggests we might need to re-conceptualise how we think about social order and structure in media cultures.
+We can understand devices such as recommender systems as attempts to transform or convert the temporal and agential fabric of social ordering practices. 
+This paper investigates practices of ordering in the setting of shopping lists and  online grocery orders. 
+Although all shopping lists mix past and future actions, actual and potential events, and normative and actual values, their re-writing by recommender systems adds a new operational dynamic.
+It reconstructs an account of how a recommender system suggests a small number of grocery items of personal relevance to each of the millions of online grocery shoppers in  a major UK supermarket chain.
+The paper explores how the constitutive incompleteness of shopping lists, their propensity to expand or change, becomes more important than their capacity to  enumerate or rank things. 
+It argues that the conversion of lists from finite ordering to indeterminate propensity in recommender systems requires new ways of conceptualising social order.
 
-##
+# Probabilities and propensities: conversion experiences in online grocery shopping
 
 Online grocery shopping at Tesco, a large supermarket chain based in the UK, includes recommendations for further grocery purchases under the title of 'Have you forgotten?'.
-When customers shop for groceries online, a list of five recommendations appear at the checkout stage.
-These recommendations are the product of a recommender system, an important category of operational device in big data (see for instance [@Striphas_2015] for analysis of Netflix recommendation; [@Morris_2015]  or [@Seaver_2015] for an account of music recommendations]).
+When Tesco customers shop for groceries online, a list of five recommendations appear at the checkout stage.
+The recommendations are the product of a recommender system, an important category of operational device in big data (see for instance [@Striphas_2015] for analysis of Netflix recommendation; [@Morris_2015]  or [@Seaver_2015] for an account of music recommendations]).
 The question 'have you forgotten?' is followed by a list of some  grocery items that could have been or are usually on a shopping list.
-The title of the suggestions is a bit misleading: the recommender system, as we will see, is not concerned with forgetting, with the many slips and oversights associated with shopping lists,  but rather with items that you didn't know should be on your list, perhaps because you have never thought of putting them there.
+The title of the suggestions is a bit misleading: the recommender system, as we will see, is not concerned with forgetting, with the many slips and oversights associated to which shopping is prone,  but rather with items that customers had not selected perhaps because they had never thought of buying them in the first place.
 
-The shelves of large contemporary supermarket are the endpoint of global logistic supply chains, in all their logistical, value-transforming and brand-mediated hypercomplexity (see [@Neilson_2012; @Tsing_2009] on logistics and power are the endpoint of global logistic supply chains, in all their logistical, value-transforming and brand-mediated hypercomplexity).
-Groceries imply a geography of agriculture, industry, transport, communication and financialisation driven by flows of labour and capital.
-Encounters between this hyper-complex geography and people, even in the familiar confines of a supermarket, are no simple matter, either for shoppers or for supermarket operators such as Tesco.
-While there are many facets to shopping (see [@Miller_2012] for an account of shopping as a practice that addresses discrepancies between normative and actual social order, and as a sacrificial ritual that instates social order), the shopping list, whether written on the back of an envelope, or saved as a list in an online grocery shopping system, is one intersectional ordering device that configures a universe of possible references, and a teeming multitude of propensities, into a possible order (in both senses of that term: an ordering and an imperative).
-As people shop, either by  trawling along aisles packed with thousands of products, or scrolling down screens or searching for particular brands amidst search results, lists inscribe some kind of order that filters or reduces the excessive propensities, claims, dazzle and distraction of commodities.
+The shelves of large contemporary supermarket are the endpoint of global logistic supply chains, in all their logistical, value-transforming and brand-mediated hypercomplexity (see [@Neilson_2012; @Tsing_2009]).
+Groceries imply a planetary geography of agriculture, industry, transport, communication and financialisation animated by flows of labour and capital.
+Encounters between this hyper-complex commodity-geography and people, even in the familiar confines of a supermarket, are no simple matter, either for shoppers or for supermarket operators such as Tesco.
+For shoppers, as Daniel Miller has argued [@Miller_2012],  shopping takes place as a negotiation of discrepancies between normative and actual social order (for instance, between ideals of health and commitments to thrift).
+It can also be seen  as a sacrificial ritual that re-instates social order in various ways.
 
-While hand-written shopping lists undoubtedly have ongoing importance and display interestingly mixed ordering practices (see the list of handwritten shopping lists at [Grocery List](http://www.grocerylist.org)), online shopping lists are written and re-written at the intersection of web and internet infrastructures, supply chain logistics, individualised practices and *habitus,* and increasingly, the predictive operations of  recommender systems.
-Just as the aisles and shelves of a supermarket present a commercially-ordered geography of needs and desires,  online shopping recommender systems generate lists that seek to align people to commodities they otherwise might have little relation to.
+The shopping list, whether written on the back of an envelope, or saved as a list in an online grocery shopping system, lies at the intersection of these flows, orderings, values and negotiations.
+It is an intersectional ordering device that encapsulates a universe of possible references, and a teeming multitude of propensities with an actual local order (in both senses of that term: an ordering and an imperative).
+As people shop, either by  trawling along aisles packed with thousands of products, or scrolling down screens or searching for particular brands amidst search results, lists inscribe some kind of order that filters or reduces the excessive propensities, claims, dazzle and distraction of commodities to the practicalities of domestic economy.
+
+Hand-written shopping lists undoubtedly have ongoing importance and display interestingly mixed ordering practices (see the montage of handwritten shopping lists at [Grocery List](http://www.grocerylist.org)).
+But online shopping lists take shape at the intersection of web and internet infrastructures, supply chain logistics, individualised practices and *habitus,* and increasingly, the predictive operations of  recommender systems.
+Just as the aisles and shelves of a supermarket present a densely-woven semiotics of objects competing for visual attention by offering distinctions of taste, thrift, expedience, novelty, indulgence, health and so forth, online shopping recommender systems generate lists that seek to align people to commodities they otherwise might have little relation to.
 
@@
 
-The items on this list could be recommendations.
+The items on Tesco's 'Have You Forgotten?' list are recommendations.
 They are only a few of them: five.
@@
 In terms of content, there are many similarities between shopping lists such as Figure \@ref(fig:shopping-hand)  -- a dozen lines scrawled on the back of a used envelope stuck on the refrigerator door  --  and the five ranked items shown to a customer on an online shopping site (as shown in Table \@ref(tab:random-basket)).
-While they are similar in content,  the list of five suggestions shown to Tesco's online customers at the checkout  exemplify a transformation in forms of ordering, a transformation that may affect the ontological texture of the world in the same way that commodification does.
-In its re-writing of the list, the recommender system effects a conversion experience in which individualised desire, consumption practices, value-chains, infrastructural time-space dynamics, the lives of others and the propensities of things all come into play.
-A major contemporary form of ordering, the recommendation list, that we see in many  forms of news, media, entertainment, marketing, advertising, can be seen as what the philosopher of science Karl Popper called  a 'crystallisation of propensities' [@Popper_1990, 18].
+Even if they are similar in content,  the list of five suggestions shown to Tesco's online customers at the checkout  exemplifies a transformation in forms of ordering.
+In its re-writing of the list, the recommender system reassembles individualised desire, consumption practices, value-chains, infrastructural time-space dynamics, the lives of others and the propensities of things.
+A major contemporary form of ordering, the recommender system can be found in many  forms of news, media, entertainment, marketing, and advertising.
 
-The Tesco presentation I saw was not really surprising or exceptional.
-It was a presentation by a data scientist to data scientists, describing an implementation of a predictive model as well as some of the problems and potentials of the predictive model.
-The presentation provides some materials for an exploration of the form, techniques, and to certain extent, experience of predictive ordering.
-In this paper, I re-present the presentation using a mixture of theoretical and practical resources.
-The theory is largely recent media theory and its accounts of prediction.
-The practical resources are that of coding and predictive models, and online grocery data.[^12]
-
-[^12]: The problem of incomplete lists reminds me of the game 'I went shopping and I bought ...
-' that adults sometimes play with children.
-The game is about remembering a well-ordered list of items - 'a' is apple, 'b' is for banana, 'c' is for carrot.
-The point of the game is to use the well-ordered set of labels provided by the sequence of letters in the alphabet to remember a random and perhaps very diverse list of items.
-The longer the game continues, the harder it is to remember since the shopping list grows longer.
-At the same time, each player repeats the list, and this  helps memorisation.
-One trick in playing this game is to choose items that are quite unexpected or increasingly random.
-Rather than `ham`, list `halva`; rather than `fish`, add `face cream.` This is especially useful for adults in beating children.
-The children have better memories for items they know, but adults know more unexpected items.
-On the other hand, very unexpected items are quite easy to remember and tend to even the score in the favour of the children.
-Balancing between the improbable and memorable, between a predictable list like 'apples, bananas, carrots' and an unusual list such as 'agave syrup, boccadilla, cilantro' is complicated.
-Something similar applies to recommender systems.
-The system can remember very many things.
-It may know everything you bought in the last few years on that site, and it can easily see what you would normally buy and add those as suggestions.
-
-Despite being titled 'Have You Forgotten?' Tesco's recommender system does not suggest forgotten items.
-It focuses on 'conversion events.' If a shopper does buy  recommended items, 'conversion' has occurred.
+Despite being titled 'Have You Forgotten?' Tesco's recommender system  focuses on 'conversion events.' If a shopper does buy a recommended item, 'conversion' has occurred.
 From a supermarket perspective, high-value conversion are desirable.
 There is little value in successfully recommending small or cheap items such as rubber bands or toothpicks.
-From a data science perspective, conversion events, when they occur, confirm the power of predictive analytics to change the world in some way.
-From a sociological perspective, matters are less obvious.
-Constructing and configuring recommender system in the name of conversion events, the recommender system effects an epistemic-operational transformation central to big data.
-We might regard the recommender system as an operational formation that responds to an ongoing control crisis associated with information [@Beniger_1986].
-We might also explore these conversion events in terms of how predictive analytics happens in the world, how the tendency of things to realize certain events -- what Popper terms the 'propensities of things' [@Popper_1990, 11] -- happens in and through data practices.
+Conversion events, when they occur, confirm the power of predictive analytics to change the world in some way.
+From a sociological perspective, the significance of conversion events is far less obvious.
+In the name of conversion events, recommender systems stage an epistemic-operational transformation central to big data.
+In conversion events,  predictive analytics happens in the world.
+The tendency of things to realize certain events -- what Popper terms the 'propensities of things' [@Popper_1990, 11] -- happens  through data practices.
 Data in concert with calculation, models, infrastructures, and experiments, affords apprehension, visibility and management of associative propensities in the world.
-Recommender systems, then, are vitally concerned not just with conversion events, but also with the what counts as order in the world.
 
@@
 
-Lists have long interested historians, anthropologists, sociologists, literary and media theorists.
+In grocery shopping, conversion events derive from and affect lists.
+Whether  they write out shopping lists or not, online grocery shoppers interact with commodities through lists.
+Lists have long interested historians, anthropologists, sociologists, literary and media theorists as everyday ordering devices.
 If, as the social anthropologist Jack Goody argued, lists are historically primary as forms of writing [@Goody_1986], and if list-making and its later variants (e.g.
 tables) precede discursive and narrative writing practices, then we might expect lists to function as powerful social ordering devices.
-As literary scholars suggest (see [@Mainberger_2003]), although lists have often been de-valued as literary forms, list-making often appears, even in literary form, whenever writing seeks to address, name, group or evoke totality, profusion, excess or abundance.
-More recent sociological work on lists (see [@deGoede_2016] for an overview) confirms the potency of lists as ordering devices.
-From both  anthropological and literary perspectives, shopping lists have excellent reasons to exist: they are deeply rooted in urban social formations historically and culturally.
-Lists such as online grocery recommendations weave together people, infrastructures, things, places and predictions.[^1.456]
+More recent sociological work on lists (see [@deGoede_2016] for an overview) explore the social and political potency of lists as ordering devices.
+As literary scholars suggest (see [@Mainberger_2003]), even if lists have often been de-valued as literary forms, list-making commonly appears in literary form whenever writing seeks to address, name, group or evoke totality, profusion, excess or abundance.
+From both  anthropological and literary perspectives, shopping lists have excellent reasons to exist: they are deeply rooted in organisational life and infrastructures.
+Lists weave together people, infrastructures, things, and places.[^1.456]
 
@@
 On the other hand, as operational inscriptions, they can be treated ethnomethodologically, as the production of social order in a given setting.
+We might, in the light of their mutability, permeability and embedding in social order,  approach lists from various theoretical angles -- as asignifying semiotics as Maurizio Lazzarato calls it in his _Signs and Machines_ [@Lazzarato_2014] or as elements in 'a new order of spatio-temporal continuity for forms of economic, political and cultural life' as Celia Lury puts it in her account of topological turn [@Lury_2012].
 
-We might, in the light of their mutability, permeability and embedding in social order,  approach lists from various theoretical angles -- as asignifying semiotics as Maurizio Lazzarato calls it in his _Signs and Machines_ [@Lazzarato_2014] or as elements in 'a new order of spatio-temporal continuity for forms of economic, political and cultural life' as Celia Lury puts it in her account of topological turn [@Lury_2012].
-In this paper, I focus on lists and list-making in terms of the predictive conditions of contemporary media experience [@Hansen_2015a].
-
+In this paper, I focus on lists and list-making in terms of the predictive social orderings.
 All lists are forms of order.
@@
 Often the order is implicit to the spatial arrangement of the list.
-Even random order can be  operationally important (for instance, in gambling machine or a psychology experiment, randomly order lists might be important).
 Mappings between list order and social order are manifestly highly contingent.
-Listing practices in a given social setting, however, always index its engagement with necessary unstable dynamics of social order.
-From the perspective of social order, shopping lists, as they are co-constructed by a supermarket chain's recommender system, might be important intersectional zones for the conversion of the average everydayness of lists through data-analytic propensities.
+A flight checklist differs greatly from a personal to-do list, even if they share many formal properties.
+Listing practices in a given social setting, therefore, always index an engagement between the formal properties of the list and the necessarily more fluid dynamics of social order.
+From the perspective of social order, shopping lists, as they are co-constructed by a supermarket chain's recommender system, might be important intersectional zones for the conversion of average everydayness through data-analytic propensities.
 
-The media theorist Mark Hansen has recently suggested that
+How should we view the predictive re-writing of lists? The media theorist Mark Hansen has recently suggested that
 
@@
 
-This assertion resonates with many accounts of big data.
-The expectation exercising much contemporary work on data, on media, platforms and finance pivots on 'micrological propensities' that might only become evident given enough data.
-Hansen argues that an ontological transformation links the micrological propensities of things in the world to prediction:
+Hansen's account resonates with wider views of big data.
+The expectation energising much predictive practice  pivots on 'micrological propensities' elicited from data.
+Hansen  links his Whitehead-inspired ontology to propensities of things in the world via probabilities:
 
@@
 
-Hansen predicates prediction as an 'ontological transformation' that affects probabilities.
-It links probabilities as calculations to propensities or the mutable agencies of things.
-The practical and indeed empirical  question is whether such transformations in probability that bring 'the propensities of things'  can be detected and articulated in the prosaic setting of shopping lists and  recommendations.
-I will argue that 'conversion' is a preferable term for these changes.
-It highlights  re-orientations in subjects, experience, things, numbers and structures that ranges more widely than Hansen's subjective 'understood-to-be-expressed' formulation.
+Hansen predicates prediction as an 'ontological transformation' that deploys probabilities in an evermore closely woven and encompassing expression of the animated, eventful, propensities of things.
+Predictive analytics links probabilities as calculations to propensities or the mutable agencies of things.
+The practical and indeed empirical  question is whether such transformations in probability that bring 'the propensities of things'  can be detected and articulated in the prosaic setting of shopping lists and recommender systems.[^20.1]
+
+[^20.1]: I will argue that 'conversion' is a preferable term for these changes.
+It highlights  re-orientations in subjects, experience, things, numbers and structures that ranges more widely than Hansen's 'ontological transformation' formulation.
 

### 2017-03-17 lists_conversion_experience.rmd
next 20 edits

@@
 But online shopping lists take shape at the intersection of web and internet infrastructures, supply chain logistics, individualised practices and *habitus,* and increasingly, the predictive operations of  recommender systems.
-Just as the aisles and shelves of a supermarket present a densely-woven semiotics of objects competing for visual attention by offering distinctions of taste, thrift, expedience, novelty, indulgence, health and so forth, online shopping recommender systems generate lists that seek to align people to commodities they otherwise might have little relation to.
+Just as the aisles and shelves of a supermarket present a densely-woven semiotics of objects competing for visual attention by offering distinctions of taste, thrift, expedience, novelty, indulgence, health and increasingly, online shopping recommender systems generate lists that seek to align people to commodities they otherwise might have little relation to (see [@Turow_2015] for an overview).
 
@@
 Predictive analytics links probabilities as calculations to propensities or the mutable agencies of things.
-The practical and indeed empirical  question is whether such transformations in probability that bring 'the propensities of things'  can be detected and articulated in the prosaic setting of shopping lists and recommender systems.[^20.1]
+The practical and indeed empirical  question is whether such transformations or 'conversion' in relations between probabilities and 'the propensities of things'  can be detected and articulated in the prosaic setting of shopping lists and recommender systems.[^20.1]
 
@@
 
-In one of the many industry-meets-academia events occurring in increasingly data science-oriented higher education institutions in the UK, speakers from industry, government and commerce described their work with predictive models.[^11.1] Shreena Patel, a PhD graduate in statistics and operations research,  works as a data scientist for DunnHumby, a well-known [ customer science company ](https://www.dunnhumby.com/).
-Her work focuses on online grocery shopping at the supermarket chain Tesco.
-Speaking to an audience of statisticians and operations researchers, Patel   focused on the models underlying  shopping list recommendations.
-Her presentation was filled with graphs, numbers, and tables concerning different aspects of the 'Have you forgotten?' recommender system.
-Patel's presentation was not sociological.
-In order to explore the operational environment she described, and its predictive operations, I will reconstruct key features of her presentation using sample data drawn from the Tesco API ([Application Programmer Interface](https://devportal.tescolabs.com/docs/services/56c73b1bf205fd0ed81dbe7a/operations/56d5b2bee2813e0a0053c47c), [@Tesco.com_2017])   and a range of data science, statistics and graphics libraries implemented in the widely-used statistical programming language `R` [@RDevelopmentCoreTeam_2010].
+At one of the many industry-meets-academia events occurring in increasingly data science-oriented higher education institutions in the UK, speakers from industry, government and commerce described their work with predictive models.[^11.1] Shreena Patel, a PhD graduate in statistics and operations research,  works as a data scientist for DunnHumby, a well-known [ customer science company ](https://www.dunnhumby.com/) [@dunnhumby_2017].
+Her work at DunnHumby focuses on online grocery shopping at the supermarket chain Tesco.
+Speaking to an audience of statisticians and operations researchers, Patel   focused on the development and operation of predictive models underlying  shopping list recommendations.
+Her presentation was filled with graphs, numbers, and tables concerning ongoing development of the 'Have you forgotten?' recommender system.
 
@@
 
-Figure \@ref(fig:simulate-inventory) plots something of the density of commodities in the Tesco grocery database.
+Like many recommender systems, Patel's models address a large inventory of commodities.
+Figure \@ref(fig:simulate-inventory) plots something of the density of commodity prices in the Tesco grocery inventory.
 By virtue of their sheer abundance, the plot cannot easily separate the many items that Tesco sells, especially in the price range £1-£10.
-A sample of the most highly priced items in Tesco stores is shown in Table  \@ref(tab:expensive-groceries)
+A random sample of the more highly priced items in Tesco stores is shown in Table  \@ref(tab:expensive-groceries)
 
@@
 s = sample_n(df[df$price>=30,], size=10)
-kable(s[order(s$price), c('name','price')])
+kable(s[order(s$price), c('Label,'Price (£)')])
 ```
 
-Patel's presentation presents two opportunities.
-First, by recounting, contextualising and commenting on some of the main steps in making the 'Have your forgotten?' list  as a example of the predictive sense-making done by data scientists and customer analytics teams working with transactional data in many commercial settings.
+Against the background of the sheer number of commodities and their distribution of prices,  Patel's presentation presents two opportunities.
+First, by recounting, contextualising and commenting on the main steps in making the 'Have your forgotten?' list, we might follow some of the predictive sense-making done by data scientists and customer analytics teams working with transactional data in a typical commercial setting.
 Patel mentioned many of these steps only fleetingly in the presentation, for they are largely taken for granted as part of predictive analytic practice.
-What stands out from the presentation is the normality of the recommender system: it is part of a long-standing and ongoing transformation or conversion of grocey shopping.
-Second, Patel's explanation of the renovation of a recommender system offered its audience a chance to step into the list-making practices situated at the centre of a mundane situation, grocery shopping.
-As a sociological response to a data scientist presenting a model to an audience of statisticians and operations researchers, my interest lies less in specific technical innovations and more on how the ordering work done by recommender systems relates to problems of social order more generally.
-The method of reconstruction borrows from the philosopher John Dewey a broad interest in how practical and abstract elements of experience are interwoven [@Dewey_1957].
-More concretely, it draws heavily on the vision of 'interface methods' proposed by Noortje Marres [@Marres_2015] in order to keep in view both the highly technical orderings of lists in recommender system and the changes in social orderings they might imply or impel.
+Second, Patel focused on the renovation and updating of long-standing data-mining practice via a much more explicitly 'big data' and 'machine learning' oriented implementation.
+What stands out from the presentation is the normality of the recommender system: it is part of a long-standing and ongoing transformation or conversion of grocery shopping.
+From a sociological standpoint, the interest lies less in specific technical innovations and more on how the ordering work done by recommender systems relates to problems of social order more generally.
+
+Since Patel was not able to release her presentation slides to a wider audience, I rely on data-driven simulations and statistical programming as ways of reconstructing elements of the recommender system.
+In order to explore the operational environment she described, and its predictive operations, I model key features of her presentation using sample grocery data drawn from the Tesco API ([Application Programmer Interface](https://devportal.tescolabs.com/docs/services/56c73b1bf205fd0ed81dbe7a/operations/56d5b2bee2813e0a0053c47c), [@Tesco.com_2017])   and a range of data science, statistics and graphics libraries implemented in the widely-used statistical programming language `R` [@RDevelopmentCoreTeam_2010].
+At a methodological level, it draws heavily on the vision of 'interface methods' proposed by Noortje Marres [@Marres_2015] in order to keep in view both the highly technical orderings of lists in recommender system and the changes in social orderings they might imply or impel.
 
@@
 
-The main narrative of Patel's presentation concerned a shift from a well-established demographically-targeted data mining model for making recommendations to a predictive, probabilistic, personally relevant  re-writing of the shopping list in almost-realtime.
+The main narrative of Patel's presentation concerned a shift from a well-established loyalty-card based data-mining model to a predictive, probabilistic, personally relevant re-writing of the shopping list in almost-realtime.
 The changes Patel described are increasingly widespread and common.
 While they are configured in Tesco-specific ways by DunnHumby (and this reflects a longer history), they are also broadly typical.
-Tesco is the largest supermarket chain in UK, famous for a customer loyalty and targeted marketing programme known as 'Tesco  Clubcard,' which started in 1991.
+Tesco is the largest supermarket chain in UK, famous amongst retailers for its  customer loyalty and targeted marketing programme known as 'Tesco  Clubcard,' which started in 1991.
 DunnHumby -- founded by operations researchers Edwina Dunn and Clive Humby -- is said to have convinced the CEO of Tesco sometime in 1991 that a loyalty card program could change the supermarket chain's relationship to its customers.
@@
 
-While Tesco  succeeded in data-mining its customers,  and perhaps became the UK's biggest supermarket with the help of data-mining in the 1990s, the online shopping environment  in 2016 is a bit different.
-It is no longer organised around campaigns involving special offers or redemption of points for demographically segmented customer (DunnHumby made heavy use of UK Census data).
+While Tesco  succeeded in data-mining its customers using demographic segmentation,  and perhaps became the UK's biggest supermarket with the help of data-mining in the 1990s, the shopping environment  in 2017 is a bit different [@Turow_2017].
+It is no longer organised around campaigns involving special offers or redemption of points for demographically segmented loyal customers (DunnHumby made heavy use of UK Census data).
 It can no longer rely on placement of goods in carefully chosen locations in stores.
-It needs or at least might want to continuously up-sell and cross-sell to customers who might hardly ever be in the supermarket itself.
-If 'Tesco is the clear winner in the online grocery market, in fact it takes almost 50p of every £1 spent on food shopping on the internet' [How SEO helps Tesco to dominate the online grocery market](https://econsultancy.com/blog/64841-how-seo-helps-tesco-to-dominate-the-online-grocery-market)[@Silverwood-Cope_2014], then Tesco itself has undergone some kind of conversion?
+It needs or at least might want to continuously up-sell and cross-sell to customers who might hardly ever visits the supermarket itself.
 
 How could we characterise the shift from ClubCard data mining to online grocery markets?
-Patel described her work at DunnHumby in terms of a shift in models from a 'rules-based list' to a 'relevancy model.' The relevancy model changed how the 'Have You Forgotten' list was constructed.
-This shift is a typical of broader re-organisation of prediction in which existing predictive practices are being re-distributed and intensified in particular ways.
-The shift in models pivots on a more probabilistic structuring of lists.
-The ostensible banality of the 'Have you forgotten?' recommendations betrays micrological signs of a topologically complex predictive infrastructure that subducts people and things in significant ways.
+If 'Tesco is the clear winner in the online grocery market, in fact it takes almost 50p of every £1 spent on food shopping on the internet' [How SEO helps Tesco to dominate the online grocery market](https://econsultancy.com/blog/64841-how-seo-helps-tesco-to-dominate-the-online-grocery-market)[@Silverwood-Cope_2014], then Tesco itself has undergone some kind of conversion?
+Patel described her work at DunnHumby as shifting the recommender system from a 'rules-based list' to a 'relevancy model.' The relevancy model changed how the 'Have You Forgotten' list was constructed.
+This shift is a typical of broader re-organisation of prediction in which existing data analysis practices are being re-distributed and intensified in particular ways.
+The shift in models results in a more probabilistic structuring of lists.
+The ostensible banality of the 'Have you forgotten?' recommendations betrays micrological signs of a topologically complex predictive infrastructure that constructs alignments between people and things in significant ways.
 
@@
 Academic researchers first began writing about recommender systems in the mid-1990s.
-They highlighted a potential shift from demographic-driven market research or data mining techniques to personalized recommendations.
-For instance, writing in 1997 in a special section of the _Communications of the ACM_ on recommender systems, Paul Resnick and Hal Varian (at that time Dean of Information Sciences at UC Berkeley, but currently Chief Economist at Google), made much of this difference [@Resnick_1997].
+From the outset they highlighted a potential shift from demographic-driven market research or data mining techniques to personalized recommendations.
+For instance, writing in 1997 in a special section of the _Communications of the ACM_ on recommender systems, Paul Resnick and Hal Varian (at that time Dean of Information Sciences at UC Berkeley, but currently Chief Economist at Google), made much of this personalisation [@Resnick_1997].
 Resnick and Varian  emphasise the need to distinguish the emerging practices from data mining:
@@
 Writing just after the advent of web-based e-Commerce, Resnick and Varian  imagined recommender systems augmenting the 'collaborative filtering' that people do when they write and read reviews of products and services (for instance, on the travel website, TripAdvisor; [@Scott_2012]).
-In 1997, two decades ago, Resnick and Varian thought that 'people provide recommendations' and recommender systems would aggregate and rank for recipients.
+In 1997,  Resnick and Varian thought that 'people provide recommendations' and recommender systems would aggregate and rank for recipients.
 A slightly later review, [@Schafer_2001], diagrams an augmented 'natural social process' with a range of elements, technologies, inputs and outputs, with varying degrees of personalization.
 
-The shift from or conversion of  collaborative filtering appeared in Patel's presentations.
+The conversion of  collaborative filtering into predictive personalisation appeared in Patel's presentations.
 Patel describes the move from 'a rules-based system'  to  a 'personal relevance model'.
-Each term in that phrase -- personal, relevance and model -- has some weight in the conversion work.
-We can assume that What Patel refers to as  the rule-based system refers  to the `apriori` algorithm, developed  by Rakesh Agrawal and Ramakrishnan Srikanti at IBM Research Alameda in the early 1990s [@Agrawal_1994].
-A now-classic approach to 'market basket analysis,' it was listed as one of the top ten data mining algorithms in a survey conducted amongst data miners [@Wu_2008] and usually attracts a chapter in data-mining and some machine learning textbooks (e.g.
+Each term in that phrase -- personal, relevance and model -- has some weight in the conversion event.
+We can assume that what Patel describes as  the 'rule-based system' refers  to the `apriori` algorithm, developed  by Rakesh Agrawal and Ramakrishnan Srikanti at IBM Research Alameda in the early 1990s [@Agrawal_1994].
+A now-classic approach to 'market basket analysis,' it was listed as one of the top ten data mining algorithms in a survey conducted amongst data miners [@Wu_2008] and usually attracts a chapter in data-mining and machine learning textbooks (e.g.
 [@Hastie_2009]).
@@
 `Whole milk` appears most frequently.
-The `apriori` algorithm attempts to re-structure shopping lists using associations between items frequently bought together.
 The association rule-based model for the `Groceries` dataset illustrates some characteristic ordering practices of 1990s-style data-mining approaches to lists (the `Groceries` dataset  was acquired from a 'local German supermarket' [@Hahsler_2006a]).
@@
 These rules, a term derived from older decision support literature, define sets of associations between items.
-As the product of a  counting practice, the association rules define sets of items or shopping baskets in terms of decreasing strength.
 Table \@ref(tab:arules-ex) shows that the first rule has a stronger association than the second.
@@
 
-While association rules and the `apriori` algorithm present social order in terms of sets of common associations weighted by frequency and hence implicitly by probability, an `apriori`-based recommender system contains no statistical model.
-It has no grasp on personal relevance since the associations it defines are only between things.
-Put in terms of the big data conversion, the rules-based system is not predictive.
+Although association rules and the `apriori` algorithm present social order in terms of sets of frequency-weighted associations, an `apriori`-based recommender system contains no statistical model.
+It also no grasp on personal relevance since the associations it defines are only between things.
+Put in terms of the big data conversion, the rules-based system is not predictive, or personal.
 The `apriori` algorithm might be seen as an attempt to grasp the excess of things in a Tesco supermarket by identifying subsets whose closure derives from the associations between things that people construct (through buying habits, by virtue of a shopping list, etc.).
 In association rules, personhood or experience is irrelevant.
-The openness of these rules -- the `apriori` algorithm can product any number of rules of diminishing strength -- points towards the propensity of things, but cannot cannot individuate that propensity.
+The openness of these rules -- the `apriori` algorithm can product any number of rules of diminishing strength -- points towards the propensity of things to find themselves together in shopping baskets, but cannot cannot individuate that propensity beyond demographic quantiles and categories.
 
-Although data mining for association rules finds relations things, it struggles with the propensity of commodities to multiply.
+Even if `apriori`  elicits associations between things, it struggles with the propensity of commodities to multiply.
 A simple illustration of the combinatorial problem faced by recommender system can be developed using the `Grocery` dataset.
-If we take all the items in the `Grocery` dataset and paste them into the 'shopping list' box on the Tesco grocery website, each of the 169 items in the dataset yields dozens and sometimes thousands of products from Tesco online.
+If we take all the items in the `Grocery` dataset and paste them into the 'shopping list' box on the Tesco grocery website (or as I side, run them as searches on the TescoLab Product API), each of the 169 items in the dataset yields dozens and sometimes thousands of products from Tesco online.
 
@@
 The 169 items of the `Groceries` dataset  expand into roughly `r total` Tesco items (see Figure \@ref(fig:tesco-ex)).
-This is, I would suggest, a form of propensity in Hansen's terms, albeit one that occurs in, or very close to, the recommender system rather than  'in the world.' The multiple listed  in the `Grocery` dataset becomes more open in this setting through new impersonal intersections that must be bounded and ordered.
-The problem that the proliferation of things on the shelves of supermarket or grocery warehouse confronts is partly combinatorial.
-`r total` products (actually Patel mentioned 200,000 products) can be combined ways.
+There is, I would suggest, a logistic propensity in things that recommender systems encounter in the ways of global supply chains.
+The multiple listed  in the `Grocery` dataset becomes more open in this setting through new impersonal intersections that must be bounded and ordered.
+The proliferation of things on the shelves of supermarket or grocery warehouse confronts produces a combinatorial problem for data miners.
+`r total` products (actually Patel mentioned 200,000 products) can be combined in a vast number of ways.
 If a typical shopping list has 20 items, then there are `r format(choose(total, 20), 2)` possible lists.
-Some of this vast number of possible shopping lists are highly unlikely in terms of typical purchasing patterns, but potentially highly interesting and value as recommendation.
+Some of this vast number of possible shopping lists are highly unlikely in terms of typical purchasing patterns, but potentially valuable as recommendations.
 

### 2017-03-20 lists_conversion_experience.rmd
next 20 edits

@@
 A major contemporary form of ordering, the recommender system can be found in many  forms of news, media, entertainment, marketing, and advertising.
+Tesco's recommender systems differs greatly from others associated with Amazon, Netflix or Ebay.
+Grocery shopping is a high-frequency, multi-purchase shopping practice and the potential for recommendations proliferates accordingly.
 
@@
 
-Amidst this proliferation of possibles multiple, the recommender system  starts to become a site of ordering -- in both senses of that term -- of things and people together in shopping lists.
-They are less the inscriptions of a discrete agent than an effect of an environment teeming with possible combinations and complementarities, a surplus of associations that the market needs to both maintain and limit.
-For instance, given that Tesco lists more than 1000 chicken products on its site, how can a customer practically discriminate between chicken products?
-Patel mentioned that some people log on and go straight to the checkout, allowing the recommender system to supply them with the suggestions that they then move to the shopping basket.
-A `r total` list of items is quite hard to list let alone read.
-Going straight to the recommendation list is the predictive equivalent of using an old shopping list and only adding new items when needed.
-
-## Ordering relevance through personhood
+## Ordering relevance through personhood or thinghood?
 
 Combinatorial surplus, however, does not exhaust what might be found in the data.
-The relevancy model that DunnHumby has implemented for Tesco (as for so many other online businesses) seeks to balance the tensions between proliferating associations and relative closure.
-It responds to the problem of a world where the propensities of things and the propensities of people mingle with each other, yet threaten to disorder the list, the site of their mutual ordering.[^181.1] Big data discourse in its promissory mode attributes potency to personalization: 'most important, using big data we hope to identify specific individuals rather than groups' [@Mayer-Schönberger_2013].
-The ontological transformation that media theorists such as Hansen suggests actually play out somewhat more conventionally: the associative propensities of things in the shopping list (as constructed by a recommender system) come together with forms of personhood associated with buying.
+The relevancy model that DunnHumby implemented for Tesco (as for so many other  businesses) seeks to balance the tensions between proliferating associations and relative closure.
+It responds to the problem of finding an ordering in which the propensities of things as commodities and the propensities of people mingle with each other, without overflowing  the site of their  ordering encounter, the list.[^181.1] Big data discourse in its promissory mode  attributes potency to personalization: 'most important, using big data we hope to identify specific individuals rather than groups' [@Mayer-Schönberger_2013].
+As Turow and co-authors conclude in their account of transformation of retail space: 'through it all, knowingly and not, and away from the spotlights of fierce social debate, retailers are encouraging daily routines that accept data-driven personalization as a centrifugal public force' [@Turow_2015, 476].
+
+Mark Hansen offers a different perspective on what might happen in these ordering encounters. Hansen attributes 'real potential' to data. Hansen attributes an associative propensity to data:
+
+> [S]urrounding any delimited predictive system is a larger field of data -- what I elsewhere call a “surplus of sensibility” -- that, viewed speculatively, indexes the causal efficacy of the total situation within which this delimited system operates.
+(Effectively, the latter gains its reliability from closing off this larger surplus of sensibility, thereby transforming an always excessive propensity into a [provisionally] closed dataset.) Because it affords data that exceed whatever any given predictive system might include, the data of the world’s causal efficacy -- the data constituting its real potentiality -- always and in principle facilitates knowledge that cannot be restricted to any particular agenda [@Hansen_2015a, 126-7].
+
+Many affirmative and critical analyses of big data and predictive analytics highlight the divisive social orderings implemented through personalization.
+For good reason, they attend to changes in personhood.
+Hansen, by contrast, seeks to conceive of the potential of data to support or provoke different relations to the world.
+Even if the underlying contrast between an open world and a 'provisionally closed dataset' limits the account in ways that I will explore, the ontological transformation  Hansen describes  involves people, commodity-things, and infrastructures.
+The recommender system generates common codings for  the associative propensities of things, the logistical value-chains of supermarket commerce and  the forms of personhood associated with shopping.
 A model of personhood based on purchase data delimits the propensities of things.
@@
 
-Patel described the data that Tesco keeps on grocery purchases: 'We have 52 weeks of purchase data for each customer.' These transformations have been widely discussed in the context of big data, where questions of who has how much data are usually most prominent (see [@Kitchin_2014]).
-In terms of the ways that lists order the world, the amount of data is not perhaps not so important as the way in which list-making is re-ordered via a model of personhood.
-
-It uses many shopping lists to identify associations between things and then uses the strongest associations to recommend additions to a current list.
-The 'relevancy model,' however, tries to do something different.
-
-Hansen attributes 'real potential' to data. In terms that somewhat echo big data epistemics, data has an associative propensity, according to Hansen:
-
-> [S]urrounding any delimited predictive system is a larger field of data -- what I elsewhere call a “surplus of sensibility” -- that, viewed speculatively, indexes the causal efficacy of the total situation within which this delimited system operates.
-(Effectively, the latter gains its reliability from closing off this larger surplus of sensibility, thereby transforming an always excessive propensity into a [provisionally] closed dataset.) Because it affords data that exceed whatever any given predictive system might include, the data of the world’s causal efficacy -- the data constituting its real potentiality -- always and in principle facilitates knowledge that cannot be restricted to any particular agenda [@Hansen_2015a, 126-7].
-
-The underlying contrast between an open world and a 'provisionally closed dataset' is both enabling and limiting.
-Hansen's account opposes closed prediction to the open propensities of a data-world.
-A closed set is surrounded and conditioned by an open set teeming with potentials.
-If we think of the  new model that Patel describes as the delimited (or delimiting?) system within   the larger surplus of sensibility, how would that help us understand what the 'Have you forgotten?' recommendations do?
+If we think of the  new model that Patel describes as the delimited (or delimiting?) system within   the larger surplus of sensibility, how would the idea of a predictive system as a provisional enclosure of the world's causal efficacy, its 'surplus of sensibility' (a terms that carries some connotations of the critical political economy of surplus value) help us understand what the 'Have you forgotten?' recommendations do?
 It enables us to understand -- as we will see next -- what the Tesco relevancy model borrows from the world.
-It limits, on the other hand, our understanding of how predictive analytics, or any other ordering device, themselves remain open, require configuration, and only exist amidst technical ensembles whose operation remains obscure.
+By contrast, despite its ontological ambitions, it limits our understanding of how predictive analytics  themselves exist amidst technical ensembles whose operation remains obscure yet central to contemporary worlds.
 
@@
 
-Patel introduced the relevancy model with a graph familiar to machine learning and statistical modelling.
+Patel described the data that Tesco keeps on grocery purchases: 'We have 52 weeks of purchase data for each customer.' These transformations have been widely discussed in the context of big data, where questions of who has how much data are usually most prominent (see [@Kitchin_2014]).
+In terms of the ways that lists order the world, the amount of data is not perhaps not so important as the way in which list-making is re-ordered via a model of personhood.
+Patel introduced the new relevancy model with a graph familiar to machine learners and statistical modellers.
 (A sketch of her graph appears in Figure \@ref(fig:models).)  The graph plots the _precision_ -- the proportion of the recommended products that customers actually purchase -- for several different models.
-In media theoretical terms, the graph indexes the 'causal efficacy' of a recommender system, its capacity to include and transform propensities or 'real potentialities' into operational knowledge.
+In Hansen's media theoretical terms, the graph indexes the 'causal efficacy' of the recommender system, its capacity to include and transform propensities or 'real potentialities' into operational knowledge.
 The graph compares the old rule-based recommender systems with some of the relevancy-model alternatives -- logistic regression, random forests, and a few others -- in terms of their predictions and how those predictions turned out.
@@
 The models Patel so briefly mentioned are all classifiers, predictive models that spatially order  variables in a dataset as a high dimensional vector space.
-In the vector space, classification means drawing a decision boundary between points that henceforth belong to different classes of things [@Mackenzie_2015].
-Here the classes of things are binary: `recommended` or `not recommended,` and for many of the models shown in the graph, membership of a class is expressed as a probability.
+Operating in the vector space, models classify people and events.
+Much of the operational formation of the classifier models received scant reference in Patel's presentation.
+It is likely that in DunnHumby's models the classifying model at the heart of the recommender system constructs a decision boundary in vector space between points that belong to different classes of things [@Mackenzie_2015].
+The classes of things are binary: `recommended` or `not recommended,` and for many of the models shown in the graph, membership of a class is expressed as a probability.
 The logistic regression model generates probabilities for each product for each customer.
@@
 
-Note that neither the infrastructure that marshals the Tesco customer data  - Patel will mention `Hadoop` -- nor the predictive analytic device -- the logistic regression model -- loom very large in the conversion of the recommender system.
+Note that neither the data analytic infrastructure that processes  Tesco customer data  - Patel will mention `Hadoop` -- nor the predictive analytic device -- the logistic regression model -- loom very large in the conversion of the recommender system.
 'Most of the gain comes,' Patel observes, 'from thinking of good features.' While this is almost a truism in big data practice (see, for instance, [@Domingos_2012]), what counts as a good feature in a shopping list?
 
-A  very brief comment from Patel -- 'we have lots of zeros' -- points to the locus of this thinking: the vast, nearly empty matrix of customer-product associations.
-Given that most things in Tesco remain relatively inert in relation to each other, a matrix that records association between individual people and different products is bound to be mostly empty.
+Rather than a surplus of sensibility, predictive models often encounter a dearth of sensibility in the data.
+A  very brief comment from Patel -- 'we have lots of zeros' -- points to the difficult terrain of  'good feature' thinking: the vast, nearly empty matrix of customer-product associations.
+Given that most things in Tesco remain relatively unknown to each other and to customers.
+A matrix that records associations between individual people and different products is bound to be mostly empty.
 Say Tesco has 1 million online customers.
-Each of online  shopper has bought some selection of the 200,000 products.
-But the product-customer matrix, the basic vector-space in which all recommender systems operate, remain very sparse and unpopulated.
+Each online  shopper has bought some selection of the 200,000 products.
 The customer-product data matrix will be `r format(1e6 * 2e5)` in size.
-Given that any one customer is likely to only have bought 100 different products, the probability matrix will be close to `r format(100-(100/2e5)*100, 2)`% empty.
-The data is, as Patel put it, 'massively unbalanced', and such imbalance would heavily bias recommendation towards commonly and only impersonally relevant suggestions, suggestions that would likely not produce the desired conversion experience.
-Rather than a surplus of sensibility, predictive models, we might say, encounter their own dearth of sensibility in the data.
+The product-customer matrix, the basic vector-space in which all recommender systems operate, remains very sparse and unpopulated.
+Given that any one customer is likely to only have bought 100 or so different products, the probability matrix will be close to `r format(100-(100/2e5)*100, 2)`% empty.
+The data is, as Patel put it, 'massively unbalanced' (any dataset where the items of interest are much rarer than some normal values is said to be unbalanced) and such imbalance would heavily bias the recommender system towards common and only impersonally relevant suggestions, suggestions that might  not produce the desired conversion experience.
 
@@
 
-Amidst this imbalance, distortion and deficit of sensibility given in the data, 'thinking of good feature' -- ways of traversing the data -- becomes central.
-For our purposes, the exact details of these features matter less than the way in which propensities become probabilities.
+Given this imbalance, distortion and deficit of sensibility in the data, 'thinking of good features' -- ways of traversing the data -- becomes central.
+For our purposes, the exact details of these features matter less than attending to the way in which propensities become probabilities.
 The dataset, as we will see, will not be practically treated as 'closed,' as Hansen puts it.

### 2017-03-21 lists_conversion_experience.rmd
finished 92 edits

@@
 
-The shopping list, whether written on the back of an envelope, or saved as a list in an online grocery shopping system, lies at the intersection of these flows, orderings, values and negotiations.
+The shopping list, whether written on the back of an envelope, or saved as a list in an online grocery shopping system, lies at the intersection of these logistical flows, infrastructural orderings, cultural values and  lively negotiations.
 It is an intersectional ordering device that encapsulates a universe of possible references, and a teeming multitude of propensities with an actual local order (in both senses of that term: an ordering and an imperative).
@@
 But online shopping lists take shape at the intersection of web and internet infrastructures, supply chain logistics, individualised practices and *habitus,* and increasingly, the predictive operations of  recommender systems.
-Just as the aisles and shelves of a supermarket present a densely-woven semiotics of objects competing for visual attention by offering distinctions of taste, thrift, expedience, novelty, indulgence, health and increasingly, online shopping recommender systems generate lists that seek to align people to commodities they otherwise might have little relation to (see [@Turow_2015] for an overview).
+Just as the aisles and shelves of a supermarket present a densely-woven semiotics of objects competing for visual attention by offering distinctions of taste, thrift, expedience, novelty, indulgence, health and increasingly, online shopping recommender systems generate lists that seek to align people to commodities that they otherwise might have little relation to (see [@Turow_2015] for an overview).
 
@@
 
-While Tesco  succeeded in data-mining its customers using demographic segmentation,  and perhaps became the UK's biggest supermarket with the help of data-mining in the 1990s, the shopping environment  in 2017 is a bit different [@Turow_2017].
+While Tesco  succeeded in data-mining its customers using demographic segmentation,  and perhaps became the UK's biggest supermarket with the help of data-mining in the 1990s, the shopping environment  in 2017 is markedly different [@Turow_2017].
 It is no longer organised around campaigns involving special offers or redemption of points for demographically segmented loyal customers (DunnHumby made heavy use of UK Census data).
@@
 Given this imbalance, distortion and deficit of sensibility in the data, 'thinking of good features' -- ways of traversing the data -- becomes central.
-For our purposes, the exact details of these features matter less than attending to the way in which propensities become probabilities.
+For our purposes, the exact details of these features matter less than attending to the way in which, contrary to Hansen's account of analytic closure,  probabilities become propensities in data practice.
 The dataset, as we will see, will not be practically treated as 'closed,' as Hansen puts it.
@@
 'Thinking of good features' potentially re-opens the dataset and sometimes completely transforms it.
-This thinking, which I suggest occurs mostly in quite obscure yet central zones of data practice, is highly mixed.
+This transmogrification, which I suggest occurs mostly in quite obscure yet central zones of data practice, is highly mixed.
 It combines experimental techniques, infrastructural dynamics, seasonal and daily temporalities, and above all, an ongoing engagement with probabilities and their fungibilities.
@@
 
-The relevancy model was deployed experimentally in a random A/B controlled trial on the Tesco website.
+For instance, the predictions of the recommender system themselves are the subject of probabilistic treatment.
+Patel described the experimental deployment of  the relevancy model experimentally in a random A/B controlled trial on the Tesco website.
 All customers were  allocated to one of four categories as shown in the table.
 In the  A/B trial,  an identical shopping list, or even the same person, might receive different recommendations.
-The randomised application of the model, drawing on protocols for randomised clinical trials first developed in the 1960s, is widely used in social media platforms and hence in all recommender systems.[^238.1] Random allocation of customers to the four categories (Test A, Test B, Control A, Control B) adds another layer of probability to the recommender system in the name of statistical validation of the effects or the 'uplift' of the model.[^122] Ironically,  the effects of a predictive model cannot be known in advance.
+The randomised application of the model, drawing on protocols for randomised clinical trials first developed in the 1960s, is widely used in social media platforms and hence in the implementation and observation of the effects of recommender systems.[^238.1] Random allocation of customers to the four categories (Test A, Test B, Control A, Control B) adds a layer of probability to the recommender system in the name of statistical validation of the effects or the 'uplift' of the model.[^122] Ironically,  the effects of a predictive model cannot be known in advance.
 They can only be observed experimentally.
 
-[^238.1]: It is difficult to gauge how much. Facebook reports that it has more than 1 million models operation in its infrastructure. TBA[]
+[^238.1]: It is difficult to gauge how much. Facebook reports that it has more than 1 million models operating in its infrastructure. [@Dunn_2016]
 
@@
 It seeks to statistically validate the effects -- the *uplift* -- of the model on  conversion events by allowing the effects of the model on what people do to be measured.
-The lift or uplift refers to conversion events associated with the same groups of people.
-Effectively is an experiment in inhabiting two different worlds, the randomised control trial  sets up a feedback look between the predictive model, the logistic regression, and the world.
-Once the experimental uplift methodology is in place, the recommender system can constantly modify the recommendation list in light of subsequent data on the uplift.
-Predictions, then, are not bound by fixed parameters of the model.
+The uplift refers to conversion events associated with the same groups of people.
+Effectively an experiment in inhabiting two different worlds, the randomised control trial  sets up a feedback look between the predictive model (the logistic regression), and the world.
 
@@
 
+A second layer of probability affects the operation of the recommender system.
 Much discussion of recent media has highlighted its real-time operation (see [@Kittler_1993; @Virilio_1997; @Hansen_2015; @Bollmer_2016]).
-Patel's description of the relevancy model infrastructure suggested a different temporal patterning.
+Patel's description of the relevancy model infrastructure sketched a different temporal patterning, suggesting immediacy in relation to shopping, but scheduling and periodicity in relation to the infrastructural operations of the predictive model.
 While raw data from Tesco Online feeds into the model every hour, a recommendation list for each customer is only generated once a week.
@@
 The possibility of adjusting the recommendations for every customer weekly derives from an infrastructure capable of collecting data, assimilating that data to a predictive model able to generate approximately 200 million relevant recommendations, where 'relevance' depends on a matrix of probabilities of associations between people and things that shifts in time.
-Like the global supply chain logistics optimised for flow of commodities [@Tsing_2009; @Nielson_2012], `hadoop` and its legion of variants (`mahout`, `spark`, `hive`, `pig`, `yarn`) operationalise flows of data at an infrastructural rather than analytic scale.
+Like the global supply chain logistics optimised for flow of commodities [@Tsing_2009; @Neilson_2012], `hadoop` and its legion of variants (`mahout`, `spark`, `hive`, `pig`, `yarn`) operationalise flows of data at an infrastructural rather than analytic scale.
 In many respects, the quickly glossed-over infrastructural deployment of DunnHumby's relevancy model is the primary conversion event: the logistic regression model at the heart of the recommender system is no longer an analytical device but an operational one.
@@
 
-The propensities of things become probabilities in predictive analytics only to the extent that probabilities themselves become propensities.
-Recommendations are relevant in a recommender system or any other predictive model to the precise extent of their probability.
+There is more to probabilities in prediction than a simply enclosure and quantification of the surplus of sensibility, the overwhelming variety, differences and noise of the world.
+I am suggesting, contrary to Hansen, that the propensities of things are enclosed as  probabilities in predictive analytics only to the extent that probabilities themselves become propensities.
+Recommendations are relevant in a recommender system or any other operational predictive model to the precise extent of their probability.
 Relevance is closely connected to propensities in the predictive condition.
@@
 Conversion events happen in the world.
-They actualise a propensity that hitherto remained either latent or potential.
-From the standpoint of the ontological conversion between probability and propensity,  prediction stems from the propensities of things, what Hansen describes as the 'complex and diffuse calculus' that only subsequently and provisionally congeals as a conversion event.
+When a recommender system discovers a common coding of human practice, and logistical configurations, with all their value differentiations, they actualise a propensity that hitherto remained potential.
+From the standpoint of the ontological exchangeability between probability and propensity,  prediction stems from the propensities of things, what Hansen describes as the 'complex and diffuse calculus' that only provisionally congeals as a conversion event.
 The 'micrological propensities' of things support the relevance of recommendations.
 
-How does the DunnHumby model deal with the micrological propensities of things?
-In her presentation, Patel  spent quite a lot of time addressing the problem of 'basket similarity,' the problem that the recommender system might recommend, for good reason, something similar to items already in a customer's basket.
-The recommender system needs to know, or the statistical model needs to know how things are related to each other so that it does not waste recommendations on overly similar items.
+How does the DunnHumby model deal with the exchange and oscillation between propensities and probabilities?
+In her presentation, Patel  spent much time addressing the problem of 'basket similarity,' the problem that the recommender system might recommend, for good reason, something similar to items already in a customer's basket.
+The recommender system or statistical model needs to know how things are related to each other so that it does not waste recommendations on overly similar items.
 A customer might be willing to substitute a similar item for something they have already chosen, but they are more likely to accept a recommendation for an item that complements items in their list.
 
-Whereas the association rule-based recommender system focused on thing-thing complementarities -- if you buy bread, you might want milk, if you buy beer, you might want diapers -- the substitutions look for micrological propensities that relate to gradations of preference, and variations in time and place.
+Whereas the association rule-based recommender system focused on thing-thing complementarities -- if you buy bread, you might want milk, if you buy beer, you might want diapers -- the personal relevance substitutions look for micrological propensities that relate to gradations of preference, and variations in time and place.
 Micrological propensities in the list relate to a particular time and place, amidst a continuously differentiable pathways between things.
@@
 Patel's comments on this were very brief but suggest that the 'Have You Forgotten?' list becomes an increasingly mutable entity, open to many substitutions and intensive re-writings.
-Like the mention of `hadoop`, what Patel said about substitution and complementarity draws on extensive but relatively well-established data practices.
+
+Like the mention of `hadoop`, what Patel said about substitution and complementarity draws on extensive but relatively well-established data practices in machine learning.
 She mentioned,  for instance,  a 'design matrix of 14,000 columns' and the use of '`L1 regularization` to drive coefficients down to zero' in a new feature of the model called 'self-learning substitutes,' a model that examines the recommendation list and predicts substitutes and complementarities for 'all products' in the list.
 The 'self learning substitutes' feature of the relevancy model draws on another data matrix, the product similarity matrix, itself generated from all previous recommendations that have led to conversion events (that is, 'a user clicked as the response').
-A customer's recommendation list might already include `Tesco Ginger Nuts 300g.`   The 'self-learning substitution model' might be able to substitute something in that list that is of higher value, either because it is bigger or more expensive: `Mcvitie's Ginger Moments 400g.` What is being modelled here?
-No longer whether to recommend a product to a customer or not, but whether something could be substitute or complement  what is already in the recommendation list.
+A customer's recommendation list might already include `Tesco Ginger Nuts 300g.`   The 'self-learning substitution model' seeks to replace list items with similar but higher value items: `Mcvitie's Ginger Moments 400g.` What is being modelled here?
+No longer whether to recommend a product to a customer or not, but micrological conversions.
 
 Similarly, modelling complementarities between things for a particular person entails a matrix of comparisons.
-`Nutella` complements `bananas,` but the converse is not necessarily true since `bananas` need not complement `nutella.` How should we understand the effect of complementarity?
-From the standpoint of selling more things, Patel emphasises that 'complementarity drives conversion.' Bringing the matrix of similarities expressed as numerical associations between products into the model improves the conversion of recommendations into purchases.
-From the standpoint of a social ordering that identifies people and things as propensities, complementarities suggest another possible outcome.
-What does it mean for our sense of social order if the relations between things shift in accordance with our experience of the world?
-At base, there is no ontological transformation here, but an inclusion of some part of the world into the model.
+`Nutella` complements `bananas,` but the converse is not necessarily true since `bananas` need not complement `nutella,` especially if I have nut allergies.
+From the standpoint of customer data science, Patel emphasised the operational significance of predicting complementarities:  'complementarity drives conversion.' Bringing the matrix of similarities expressed as numerical associations between products into the model improves the conversion of recommendations into purchases.
+From the standpoint of a sociological account of social ordering that associate people and things through propensities, complementarities pose a different problem: what does it mean for social order if the relations between things shift in accordance with our experience of the world as rendered in the intricately varying weights of an L1-regularized logistic regression model?
 In the similarity matrix, shaped by the history of past recommendations and purchases made by this customer, the propensities of things drive down the surplus of sensibility associated with what people want.
@@
 
-In their intersection with mundane practices of ordering, recommender systems offer some guidance to the ways that predictive practices affect existing forms of social order in the world.
-Rather than directly positing an ontological transformation, I have explored some of the practices of conversion associated with a shift from a data-mining rule-based system to a predictive recommender system.
+In their mundane intersectional practices of ordering, recommender systems show how  predictive practices affect existing forms of social order in the world.
+Rather than directly positing an ontological transformation (as Hansen does), or pointing to the erosion of social order (as Turow does), I have explored some of the conversion events associated with a shift from a data-mining rule-based system to a predictive recommender system.
 The DunnHumby model and the Tesco 'Have You Forgotten?' recommendations present many of the traits of the conversion processes associated with big data.
@@
 We have seen several types of conversion events: ongoing transformation of models and infrastructures, the experimentalisation  of recommendations, and the broader diffusion of a calculus of probabilities into a fibrous weave of propensities and conditionalities, warping over time.
-The predictive assemblages respond to problems of logistical ensembles more generally.
+The predictive assemblages connect logistical ensembles to the negotiation, compromises, rituals and local orderings of everyday life.
 
 I have been suggesting that predictive listing in the form of open-ended recommendation changes social order.
-It alters the way in which repetition or reiteration, exemplified in the shopping list, can also participate in change and emergent order.
-By following Popper in understanding the world in terms of propensities, we have a chance  to engage with contemporary predictive processes in ways that both acknowledge their relentless and exhausting pressure to consume more and retain some chance of inhabiting the world differently.
+It establishes a common coding between machinic and ritual repetition (as exemplified in the shopping list).
+It also suggests some ways in which repetition or reiteration  can also participate in changing and emergent order.
+By understanding the world in terms of propensities, we have a chance  to engage with contemporary predictive processes in ways that both acknowledge their relentless and exhausting pressure to consume more and retain some chance of inhabiting the world differently.
 
-I diverge from some understandings of predictive practice.
-Rather than understanding prediction as the closed form of an ontologically open world ('the propensities of things'), the DunnHumby relevance model, in all its statistical, infrastructural, organisational and banal complexity, suggests that predictive conversion remains constitutively incomplete.
-Predictive practice abounds in propensities.
-Ironically, shopping lists  exemplify this incompleteness and ineradicable openness of prediction practice.
-The constitutive incompleteness and openness of lists are very closely entangled with problems of social order amidst predictive operations.
+Rather than understanding prediction as the closed form of an ontologically open world ('the propensities of things'), the DunnHumby relevance model, in all its statistical, infrastructural, organisational and banal complexity, suggests that the predictive conversion remains constitutively incomplete.
+Shopping lists in all their banal yet primary orderings are fertile terrain for prediction because they are incomplete, somewhat open and mutable forms.
+The constitutive incompleteness and openness of lists are very closely entangled with ongoing processes of ordering and othering, and at the same time, can accommodate predictive operations.
 Lists undergoing a 'conversion' experience as they become more probabilistic, but their heterogeneous character as operations caught up in models, infrastructures, experiments, and the flow of commodities means that they can never form a closed set or dataset.
-The problem of 'constructing good features,' a problem that any predictive model must address, persists as the site of experiment with structures that lack closure.
-A constitutive openness to new associations, new combinations, new groupings or actors is integral to the conversion event.
+The problem of 'constructing good features,' a problem that any predictive model must address, persists as the site of experiments in partly open structures.
 
+A constitutive openness to new associations, new combinations, new groupings or actors is integral to the conversion events associated with contemporary data practice.
 In their account of the Netflix recommender system, Ted Hallinan and Peter Striphas make the following argument:
@@
 
-The 'universe of reference' includes objects, ideas and practices mingle.
+Despite the somewhat troubling reference to 'another universe,' as if Netflix or DunnHumby's recommender systems occupy a different reality, Hallinan and Striphas usefully and aptly highlight how such systems  mingle  objects, ideas and practices.
+
 Something similar happens to shopping in the Dunnhumby relevance model.
 In the customer-product association matrix, in the product-product similarity matrix, in the 14,000 variables included in the logistic regression model, a 'universe of reference' takes shape.
-The shopping list, the commonplace inscription of normative and actual social order, takes on a complex temporal-material structure that mixes prediction, infrastructural constraints, the individual characteristics of list-makers, statistical validations, database architectures and above all the **surplus sensibility** associated with  circulation of commodities.
-Shopping lists undergoing a **conversion experience** as they become more probabilistic, but the idea that probabilities are the closed form of propensities does not account for the distribution of probabilities in relations between things, between people, between people and things, and between things and people.
+The shopping list, the commonplace inscription of normative and actual social order, takes on a complex temporal-material structure that mixes prediction, infrastructural constraints, the individual characteristics of list-makers, statistical validations, database architectures and above all the surplus sensibility associated with  circulation and consumption of commodities.
+Shopping lists undergoing a conversion experience as they become more probabilistic, but the idea that probabilities are the closed form of propensities does not account for the distribution of probabilities in relations between things, between people, between people and things, and between things and people.
 

### 2017-03-22 lists_conversion_experience.rmd
fixing script bugs

@@
 
-How should we view the predictive re-writing of lists? The media theorist Mark Hansen has recently suggested that
-
-> predictive analytics are discoveries of micrological propensities that are not directly correlated with human understanding and affectivity and that do not by themselves cohere into clearly identifiable events [@Hansen_2015a, 111-112]
-
-Hansen's account resonates with wider views of big data.
+How should we view the predictive re-writing of lists?
+The media theorist Mark Hansen has recently suggested that 'predictive analytics are discoveries of micrological propensities that are not directly correlated with human understanding and affectivity and that do not by themselves cohere into clearly identifiable events.' [@Hansen_2015a, 111-112] Hansen's account resonates with wider views of big data.
 The expectation energising much predictive practice  pivots on 'micrological propensities' elicited from data.
-Hansen  links his Whitehead-inspired ontology to propensities of things in the world via probabilities:
+Unlike many other accounts of data, Hansen  links data to propensities of things in the world via probabilities:
 
@@
 
-Hansen predicates prediction as an 'ontological transformation' that deploys probabilities in an evermore closely woven and encompassing expression of the animated, eventful, propensities of things.
-Predictive analytics links probabilities as calculations to propensities or the mutable agencies of things.
+In his Whitehead-inspired account, Hansen theorises prediction as an 'ontological transformation' that deploys probabilities in an evermore closely woven and encompassing expression of animated, eventful, propensities of things.
+Predictive analytics links probabilities as calculations to propensities or the mutable associative agencies of things.
 The practical and indeed empirical  question is whether such transformations or 'conversion' in relations between probabilities and 'the propensities of things'  can be detected and articulated in the prosaic setting of shopping lists and recommender systems.[^20.1]
@@
 
-```{r simulate-inventory, echo=FALSE, cache=TRUE, fig.margin=TRUE, fig.cap='Prices and quantities of things in the world'}
+```{r simulate-inventory, echo=FALSE, cache=TRUE, fig.margin=TRUE, fig.cap='Prices and quantities of things in the Tesco universe of reference'}
 library(dplyr)
@@
 
+```{r  expensive-groceries, echo=FALSE, cache=TRUE, fig.margin=TRUE}
+s = sample_n(df[df$price>=30,], size=10)
+s$Price = s$price
+s$Label = s$name
+kable(s[order(s$price), c('Label','Price')], col.names=c('Item', 'Price (£)'))
+```
+
 Like many recommender systems, Patel's models address a large inventory of commodities.
@@
 A random sample of the more highly priced items in Tesco stores is shown in Table  \@ref(tab:expensive-groceries)
-
-```{r  expensive-groceries, echo=FALSE, cache=TRUE, fig.margin=TRUE}
-s = sample_n(df[df$price>=30,], size=10)
-kable(s[order(s$price), c('Label,'Price (£)')])
-```
 

### 2017-03-22 render_tufte.r

@@ -1,6 +1,8 @@
 library(rmarkdown)
+library(bookdown)
 library(tint)
 library(tufte)
 #render('lists_conversion_experience.rmd',output_format=tufte_handout())
 #render('lists_conversion_experience.rmd',output_format=ioslides_presentation)
-render('lists_conversion_experience.rmd',output_format=tintPdf())
+#render('lists_conversion_experience.rmd',output_format=tintPdf())
+render('lists_conversion_experience.rmd',output_format=word_document2())

### 2017-03-22 lists_conversion_experience.rmd
fixing references

@@
 The relevancy model that DunnHumby implemented for Tesco (as for so many other  businesses) seeks to balance the tensions between proliferating associations and relative closure.
-It responds to the problem of finding an ordering in which the propensities of things as commodities and the propensities of people mingle with each other, without overflowing  the site of their  ordering encounter, the list.[^181.1] Big data discourse in its promissory mode  attributes potency to personalization: 'most important, using big data we hope to identify specific individuals rather than groups' [@Mayer-Schönberger_2013].
+It responds to the problem of finding an ordering in which the propensities of things as commodities and the propensities of people mingle with each other, without overflowing  the site of their  ordering encounter, the list.[^181.1] Big data discourse in its promissory mode  attributes potency to personalization: 'most important, using big data we hope to identify specific individuals rather than groups' [@Mayer-Schonberger_2013].
 As Turow and co-authors conclude in their account of transformation of retail space: 'through it all, knowingly and not, and away from the spotlights of fierce social debate, retailers are encouraging daily routines that accept data-driven personalization as a centrifugal public force' [@Turow_2015, 476].
@@
 Much of the operational formation of the classifier models received scant reference in Patel's presentation.
-It is likely that in DunnHumby's models the classifying model at the heart of the recommender system constructs a decision boundary in vector space between points that belong to different classes of things [@Mackenzie_2015].
+It is likely that in DunnHumby's models the classifying model at the heart of the recommender system constructs a decision boundary in vector space between points that belong to different classes of things [@Mackenzie_2015e].
 The classes of things are binary: `recommended` or `not recommended,` and for many of the models shown in the graph, membership of a class is expressed as a probability.

### 2017-03-22 lists_conversion_experience.rmd

@@
 bibliography: /home/mackenza/ref_bibs/uni.bib
-bibliostyle: 'harvard'
+csl: "sage-harvard.csl"
 link-citations: yes
@@
 df = bind_rows(sapply(csvs, read_tsv))
-kable(sample_n(df, size=5)$name, col.names='item')
+kable(sample_n(df, size=5)$name, col.names='item', caption='A random list of Tesco grocery items')
 ```
 
-```{r shopping-hand}
+```{r shopping-hand, fig.cap='A hand-written shopping list'}
 include_graphics('figure/1_shopping.png')
@@
 
-```{r  expensive-groceries, echo=FALSE, cache=TRUE, fig.margin=TRUE}
+```{r expensive-groceries, echo=FALSE, cache=TRUE}
 s = sample_n(df[df$price>=30,], size=10)
@@
 s$Label = s$name
-kable(s[order(s$price), c('Label','Price')], col.names=c('Item', 'Price (£)'))
+kable(s[order(s$price), c('Label','Price')], col.names=c('Item', 'Price (£)'), caption='Random sample of higher cost items from Tesco')
 ```
@@
 
-```{r arules-ex, cache=TRUE, message=FALSE, echo=FALSE, fig.margin=TRUE}
+```{r arules-ex, cache=TRUE, fig.cap='Frequency of association betwene items in the `Grocery` dataset', message=FALSE, echo=FALSE, fig.margin=TRUE}
 library(knitr)
@@
 itemFrequencyPlot(Groceries,topN=10,col=rainbow(10), ntype="absolute",  cex.axis = 0.55)
+```
+
+```{r arules-ex2}
 options(digits=2)
@@
 rules_sorted <- sort(rules, decreasing=TRUE,by="confidence")
-kable(inspect(rules_sorted[1:5]))
+kable(inspect(rules_sorted[1:5]), caption='The first five association rules for the `Groceries` dataset')
 ```
@@
 These rules, a term derived from older decision support literature, define sets of associations between items.
-Table \@ref(tab:arules-ex) shows that the first rule has a stronger association than the second.
+Table \@ref(tab:arules-ex2) shows that the first rule has a stronger association than the second.
 Milk and vegetables is more likely an association than milk and buns.
@@
 
-```{r models}
+```{r models, fig.cap='Precision measurements for different machine learning models'}
 include_graphics('figure/patel_compares_models.png')

### 2017-03-25 lists_conversion_experience.rmd
fixing humby reference to image
trying gh pages for paper

@@
 
-## Lists and orders
+# Lists and orders
 
@@
 
-## Conversion events: DunnHumby's work for Tesco
+# Conversion events: DunnHumby's work for Tesco
 
@@
 
-## How the shopping list was reconstituted in the 1990s
+# How the shopping list was reconstituted in the 1990s
 
@@
 
-![Humby, 1984]( figure/2_humby_young_org_1994.png)
+```{r humby-1984, fig.cap = 'Abstract from a Clive Humby presentation (Humby,1984)'}
+include_graphics('figure/2_humby_young_org_1994.png')
+```
 
@@
 
-## Reconfiguring recommendations
+# Reconfiguring recommendations
 
@@
 
-## Ordering relevance through personhood or thinghood?
+# Ordering relevance through personhood or thinghood?
 
@@
 
-## Sparsity: output 20 recommendations from 200,000 products
+# Sparsity: output 20 recommendations from 200,000 products
 
@@
 
-## Experimental settings: A/B allocation and uplift
+# Experimental settings: A/B allocation and uplift
 
@@
 
-## Infrastructure: the realtime engine serves
+# Infrastructure: the realtime engine serves
 
@@
 
-## Similarity, substitution, complementarity and micrological propensities
+# Similarity, substitution, complementarity and micrological propensities
 

### 2017-03-25 render_tufte.r

@@ -5,4 +5,5 @@ library(tufte)
 #render('lists_conversion_experience.rmd',output_format=tufte_handout())
 #render('lists_conversion_experience.rmd',output_format=ioslides_presentation)
 #render('lists_conversion_experience.rmd',output_format=tintPdf())
-render('lists_conversion_experience.rmd',output_format=word_document2())
+#render('lists_conversion_experience.rmd',output_format=word_document2())
+render('lists_conversion_experience.rmd',output_format=html_document2(toc=TRUE, toc_depth = 1, toc_float=TRUE, theme='journal'))

### 2017-03-25 render_tufte.r
trying tufte html

@@ -6,4 +6,5 @@ library(tufte)
 #render('lists_conversion_experience.rmd',output_format=ioslides_presentation)
 #render('lists_conversion_experience.rmd',output_format=tintPdf())
 #render('lists_conversion_experience.rmd',output_format=word_document2())
-render('lists_conversion_experience.rmd',output_format=html_document2(toc=TRUE, toc_depth = 1, toc_float=TRUE, theme='journal'))
+#render('lists_conversion_experience.rmd',output_file='index.html', output_format=html_document2(toc=TRUE, toc_depth = 1, toc_float=TRUE, theme='journal'))
+render('lists_conversion_experience.rmd',output_file='index.html', output_format=tufte_html2(toc=TRUE, toc_depth = 1))

### 2017-03-25 render_tufte.r
reverting back to html2

@@ -6,5 +6,5 @@ library(tufte)
 #render('lists_conversion_experience.rmd',output_format=ioslides_presentation)
 #render('lists_conversion_experience.rmd',output_format=tintPdf())
 #render('lists_conversion_experience.rmd',output_format=word_document2())
-#render('lists_conversion_experience.rmd',output_file='index.html', output_format=html_document2(toc=TRUE, toc_depth = 1, toc_float=TRUE, theme='journal'))
-render('lists_conversion_experience.rmd',output_file='index.html', output_format=tufte_html2(toc=TRUE, toc_depth = 1))
+render('lists_conversion_experience.rmd',output_file='index.html', output_format=html_document2(toc=TRUE, toc_depth = 1, toc_float=FALSE, theme='journal'))
+#render('lists_conversion_experience.rmd',output_file='index.html', output_format=tufte_html2(toc=TRUE, toc_depth = 1))

### 2017-03-25 render_tufte.r
tweaking rendering

@@ -5,5 +5,5 @@ library(tufte)
 #render('lists_conversion_experience.rmd',output_format=tufte_handout())
 #render('lists_conversion_experience.rmd',output_format=ioslides_presentation)
 #render('lists_conversion_experience.rmd',output_format=tintPdf())
-#render('lists_conversion_experience.rmd',output_format=word_document2())
-render('lists_conversion_experience.rmd',output_format=html_document2(toc=TRUE, toc_depth = 1, toc_float=TRUE, theme='journal'))
+render('lists_conversion_experience.rmd',output_format=word_document2())
+

### 2017-03-28 lists_conversion_experience.rmd
preparing submission branch

@@
 title: "Probabilities and propensities: conversion experiences in online grocery shopping"
-author: "Adrian Mackenzie"
-date: "`r Sys.Date()`"
-output:
-tufte::tufte_handout: default
+author: Adrian Mackenzie^[University of Somewhere] 
+date: "`r format(Sys.time(), '%B %Y')`"
 fontsize: 12pt
@@
     - \doublespacing
+abstract: 
+    "Recommender systems attempt to transform or convert the temporal and agential fabric of social ordering by substituting probabilistic calculation for social practices.
+This paper investigates practices of ordering in the setting of shopping lists and  online grocery orders.
+It reconstructs an account of how a recommender system suggests a small number of grocery items of personal relevance to each of the millions of online grocery shoppers in  a major UK supermarket chain.
+Although all shopping lists mix past and future actions, actual and potential events, and normative and actual values, their re-writing by recommender systems adds a new operational dynamic.The paper explores how in the context of recommender systems the constitutive incompleteness of shopping lists, their propensity to expand or change, becomes more important than their capacity to inscribe social order.
+This suggests that the conversion of lists from finite ordering to indeterminate propensity in recommender systems requires new ways of conceptualising social order."
+keywords: prediction, probability, order, recommender system, shopping
 ---
@@
 ```
-
-# Abstract
-
-We can understand devices such as recommender systems as attempts to transform or convert the temporal and agential fabric of social ordering practices. 
-This paper investigates practices of ordering in the setting of shopping lists and  online grocery orders. 
-Although all shopping lists mix past and future actions, actual and potential events, and normative and actual values, their re-writing by recommender systems adds a new operational dynamic.
-It reconstructs an account of how a recommender system suggests a small number of grocery items of personal relevance to each of the millions of online grocery shoppers in  a major UK supermarket chain.
-The paper explores how the constitutive incompleteness of shopping lists, their propensity to expand or change, becomes more important than their capacity to  enumerate or rank things. 
-It argues that the conversion of lists from finite ordering to indeterminate propensity in recommender systems requires new ways of conceptualising social order.
 

### 2017-03-28 title.rmd
preparing submission branch

--- 2017-03-28 lists_conversion_experience.rmd
@@
 title: "Probabilities and propensities: conversion experiences in online grocery shopping"
-author: Adrian Mackenzie^[University of Somewhere] 
+author: Adrian Mackenzie, Sociology, Lancaster University 
 date: "`r format(Sys.time(), '%B %Y')`"
+output:
+    word_document: 
+     reference_docx: reference_docx.docx  
 fontsize: 12pt
-citation_package: biber
-latex_engine: xelatex
-bibliography: /home/mackenza/ref_bibs/uni.bib
-csl: "sage-harvard.csl"
-link-citations: yes
-header-includes:
-    - \usepackage{setspace}
-    - \doublespacing
 abstract: 
@@
 ---
-
-```{r setup2, cache=FALSE, include=FALSE}
-library(knitr)
-library(tint)
-knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tint'))
-options(htmltools.dir.version = FALSE)
-output <- opts_knit$get("rmarkdown.pandoc.to")
-opts_chunk$set(warning=FALSE)
-opts_chunk$set(message=FALSE)
-opts_chunk$set(echo=FALSE)
-opts_chunk$set(error=FALSE)
-opts_chunk$set(cache=TRUE)
-```
-
-# Probabilities and propensities: conversion experiences in online grocery shopping
-
-Online grocery shopping at Tesco, a large supermarket chain based in the UK, includes recommendations for further grocery purchases under the title of 'Have you forgotten?'.
-When Tesco customers shop for groceries online, a list of five recommendations appear at the checkout stage.
-The recommendations are the product of a recommender system, an important category of operational device in big data (see for instance [@Striphas_2015] for analysis of Netflix recommendation; [@Morris_2015]  or [@Seaver_2015] for an account of music recommendations]).
-The question 'have you forgotten?' is followed by a list of some  grocery items that could have been or are usually on a shopping list.
-The title of the suggestions is a bit misleading: the recommender system, as we will see, is not concerned with forgetting, with the many slips and oversights associated to which shopping is prone,  but rather with items that customers had not selected perhaps because they had never thought of buying them in the first place.
-
-The shelves of large contemporary supermarket are the endpoint of global logistic supply chains, in all their logistical, value-transforming and brand-mediated hypercomplexity (see [@Neilson_2012; @Tsing_2009]).
-Groceries imply a planetary geography of agriculture, industry, transport, communication and financialisation animated by flows of labour and capital.
-Encounters between this hyper-complex commodity-geography and people, even in the familiar confines of a supermarket, are no simple matter, either for shoppers or for supermarket operators such as Tesco.
-For shoppers, as Daniel Miller has argued [@Miller_2012],  shopping takes place as a negotiation of discrepancies between normative and actual social order (for instance, between ideals of health and commitments to thrift).
-It can also be seen  as a sacrificial ritual that re-instates social order in various ways.
-
-The shopping list, whether written on the back of an envelope, or saved as a list in an online grocery shopping system, lies at the intersection of these logistical flows, infrastructural orderings, cultural values and  lively negotiations.
-It is an intersectional ordering device that encapsulates a universe of possible references, and a teeming multitude of propensities with an actual local order (in both senses of that term: an ordering and an imperative).
-As people shop, either by  trawling along aisles packed with thousands of products, or scrolling down screens or searching for particular brands amidst search results, lists inscribe some kind of order that filters or reduces the excessive propensities, claims, dazzle and distraction of commodities to the practicalities of domestic economy.
-
-Hand-written shopping lists undoubtedly have ongoing importance and display interestingly mixed ordering practices (see the montage of handwritten shopping lists at [Grocery List](http://www.grocerylist.org)).
-But online shopping lists take shape at the intersection of web and internet infrastructures, supply chain logistics, individualised practices and *habitus,* and increasingly, the predictive operations of  recommender systems.
-Just as the aisles and shelves of a supermarket present a densely-woven semiotics of objects competing for visual attention by offering distinctions of taste, thrift, expedience, novelty, indulgence, health and increasingly, online shopping recommender systems generate lists that seek to align people to commodities that they otherwise might have little relation to (see [@Turow_2015] for an overview).
-
-```{r random-basket, echo=FALSE, cache=TRUE, fig.cap='Random basket'}
-library(dplyr)
-library(ggplot2)
-library(readr)
-library(knitr)
-csvs = list.files('data', pattern = '*.csv',full.names=TRUE)
-df = bind_rows(sapply(csvs, read_tsv))
-kable(sample_n(df, size=5)$name, col.names='item', caption='A random list of Tesco grocery items')
-```
-
-```{r shopping-hand, fig.cap='A hand-written shopping list'}
-include_graphics('figure/1_shopping.png')
-```
-
-The items on Tesco's 'Have You Forgotten?' list are recommendations.
-They are only a few of them: five.
-The question is: which five items and in what order?
-In terms of content, there are many similarities between shopping lists such as Figure \@ref(fig:shopping-hand)  -- a dozen lines scrawled on the back of a used envelope stuck on the refrigerator door  --  and the five ranked items shown to a customer on an online shopping site (as shown in Table \@ref(tab:random-basket)).
-Even if they are similar in content,  the list of five suggestions shown to Tesco's online customers at the checkout  exemplifies a transformation in forms of ordering.
-In its re-writing of the list, the recommender system reassembles individualised desire, consumption practices, value-chains, infrastructural time-space dynamics, the lives of others and the propensities of things.
-A major contemporary form of ordering, the recommender system can be found in many  forms of news, media, entertainment, marketing, and advertising.
-Tesco's recommender systems differs greatly from others associated with Amazon, Netflix or Ebay.
-Grocery shopping is a high-frequency, multi-purchase shopping practice and the potential for recommendations proliferates accordingly.
-
-Despite being titled 'Have You Forgotten?' Tesco's recommender system  focuses on 'conversion events.' If a shopper does buy a recommended item, 'conversion' has occurred.
-From a supermarket perspective, high-value conversion are desirable.
-There is little value in successfully recommending small or cheap items such as rubber bands or toothpicks.
-Conversion events, when they occur, confirm the power of predictive analytics to change the world in some way.
-From a sociological perspective, the significance of conversion events is far less obvious.
-In the name of conversion events, recommender systems stage an epistemic-operational transformation central to big data.
-In conversion events,  predictive analytics happens in the world.
-The tendency of things to realize certain events -- what Popper terms the 'propensities of things' [@Popper_1990, 11] -- happens  through data practices.
-Data in concert with calculation, models, infrastructures, and experiments, affords apprehension, visibility and management of associative propensities in the world.
-
-# Lists and orders
-
-In grocery shopping, conversion events derive from and affect lists.
-Whether  they write out shopping lists or not, online grocery shoppers interact with commodities through lists.
-Lists have long interested historians, anthropologists, sociologists, literary and media theorists as everyday ordering devices.
-If, as the social anthropologist Jack Goody argued, lists are historically primary as forms of writing [@Goody_1986], and if list-making and its later variants (e.g.
-tables) precede discursive and narrative writing practices, then we might expect lists to function as powerful social ordering devices.
-More recent sociological work on lists (see [@deGoede_2016] for an overview) explore the social and political potency of lists as ordering devices.
-As literary scholars suggest (see [@Mainberger_2003]), even if lists have often been de-valued as literary forms, list-making commonly appears in literary form whenever writing seeks to address, name, group or evoke totality, profusion, excess or abundance.
-From both  anthropological and literary perspectives, shopping lists have excellent reasons to exist: they are deeply rooted in organisational life and infrastructures.
-Lists weave together people, infrastructures, things, and places.[^1.456]
-
-[^1.456]: While lists intersect strongly with other meaning and sense-making practices in everyday life and popular culture, they can be refractory to discourse  and textual analysis methods built around models of language or speech, with its rules of syntax and grammar.
-On the one hand, lists are highly associative.
-They can be interpreted or decoded semiotically, although they exhibit  variations in textuality -- how they are written and read -- that thwart semiotic readings.
-On the other hand, as operational inscriptions, they can be treated ethnomethodologically, as the production of social order in a given setting.
-We might, in the light of their mutability, permeability and embedding in social order,  approach lists from various theoretical angles -- as asignifying semiotics as Maurizio Lazzarato calls it in his _Signs and Machines_ [@Lazzarato_2014] or as elements in 'a new order of spatio-temporal continuity for forms of economic, political and cultural life' as Celia Lury puts it in her account of topological turn [@Lury_2012].
-
-In this paper, I focus on lists and list-making in terms of the predictive social orderings.
-All lists are forms of order.
-Mathematicians distinguish lists from sets on the basis of order.
-Formally, both lists and sets are collections, but a list also contains a mapping of its elements to an ordering, usually the natural or counting numbers.
-Sometimes this order is explicit.
-Numbers are written down one side of the list.
-Often the order is implicit to the spatial arrangement of the list.
-Mappings between list order and social order are manifestly highly contingent.
-A flight checklist differs greatly from a personal to-do list, even if they share many formal properties.
-Listing practices in a given social setting, therefore, always index an engagement between the formal properties of the list and the necessarily more fluid dynamics of social order.
-From the perspective of social order, shopping lists, as they are co-constructed by a supermarket chain's recommender system, might be important intersectional zones for the conversion of average everydayness through data-analytic propensities.
-
-How should we view the predictive re-writing of lists?
-The media theorist Mark Hansen has recently suggested that 'predictive analytics are discoveries of micrological propensities that are not directly correlated with human understanding and affectivity and that do not by themselves cohere into clearly identifiable events.' [@Hansen_2015a, 111-112] Hansen's account resonates with wider views of big data.
-The expectation energising much predictive practice  pivots on 'micrological propensities' elicited from data.
-Unlike many other accounts of data, Hansen  links data to propensities of things in the world via probabilities:
-
->Whatever explanatory and causal value predictive analytics of large datasets have is, I suggest, ultimately rooted in this ontological transformation whereby probabilities are understood to be expressions of the actual propensity of things [@Hansen_2015a, 120]
-
-In his Whitehead-inspired account, Hansen theorises prediction as an 'ontological transformation' that deploys probabilities in an evermore closely woven and encompassing expression of animated, eventful, propensities of things.
-Predictive analytics links probabilities as calculations to propensities or the mutable associative agencies of things.
-The practical and indeed empirical  question is whether such transformations or 'conversion' in relations between probabilities and 'the propensities of things'  can be detected and articulated in the prosaic setting of shopping lists and recommender systems.[^20.1]
-
-[^20.1]: I will argue that 'conversion' is a preferable term for these changes.
-It highlights  re-orientations in subjects, experience, things, numbers and structures that ranges more widely than Hansen's 'ontological transformation' formulation.
-
-# Conversion events: DunnHumby's work for Tesco
-
-At one of the many industry-meets-academia events occurring in increasingly data science-oriented higher education institutions in the UK, speakers from industry, government and commerce described their work with predictive models.[^11.1] Shreena Patel, a PhD graduate in statistics and operations research,  works as a data scientist for DunnHumby, a well-known [ customer science company ](https://www.dunnhumby.com/) [@dunnhumby_2017].
-Her work at DunnHumby focuses on online grocery shopping at the supermarket chain Tesco.
-Speaking to an audience of statisticians and operations researchers, Patel focused on the development and operation of predictive models underlying  shopping list recommendations.
-Her presentation was filled with graphs, numbers, and tables concerning ongoing development of the 'Have you forgotten?' recommender system.
-
-[^11.1]: More than 100  Data Science Institutes have been set up in North America, Europe and UK since around 2012.
-The traffic between higher education and data analytics in industry is intense and flows in several forms: people funding, research findings, software and technical devices (code) and training.
-
-```{r simulate-inventory, echo=FALSE, cache=TRUE, fig.margin=TRUE, fig.cap='Prices and quantities of things in the Tesco universe of reference'}
-library(dplyr)
-library(ggplot2)
-library(readr)
-library(knitr)
-
-csvs = list.files('data', pattern = '*.csv',full.names=TRUE)
-df = bind_rows(sapply(csvs, read_tsv))
-
-g = ggplot(df, aes(x=price,y=ContentsQuantity)) + geom_point(alpha=0.3, size=0.9 ) + theme(axis.text=element_text(size=10))  + xlab('Price (£)')
-g + ylab('Number of products') + scale_x_log10() + geom_smooth()
-```
-
-```{r expensive-groceries, echo=FALSE, cache=TRUE}
-s = sample_n(df[df$price>=30,], size=10)
-s$Price = s$price
-s$Label = s$name
-kable(s[order(s$price), c('Label','Price')], col.names=c('Item', 'Price (£)'), caption='Random sample of higher cost items from Tesco')
-```
-
-Like many recommender systems, Patel's models address a large inventory of commodities.
-Figure \@ref(fig:simulate-inventory) plots something of the density of commodity prices in the Tesco grocery inventory.
-By virtue of their sheer abundance, the plot cannot easily separate the many items that Tesco sells, especially in the price range £1-£10.
-A random sample of the more highly priced items in Tesco stores is shown in Table  \@ref(tab:expensive-groceries)
-
-Against the background of the sheer number of commodities and their distribution of prices,  Patel's presentation presents two opportunities.
-First, by recounting, contextualising and commenting on the main steps in making the 'Have your forgotten?' list, we might follow some of the predictive sense-making done by data scientists and customer analytics teams working with transactional data in a typical commercial setting.
-Patel mentioned many of these steps only fleetingly in the presentation, for they are largely taken for granted as part of predictive analytic practice.
-Second, Patel focused on the renovation and updating of long-standing data-mining practice via a much more explicitly 'big data' and 'machine learning' oriented implementation.
-What stands out from the presentation is the normality of the recommender system: it is part of a long-standing and ongoing transformation or conversion of grocery shopping.
-From a sociological standpoint, the interest lies less in specific technical innovations and more on how the ordering work done by recommender systems relates to problems of social order more generally.
-
-Since Patel was not able to release her presentation slides to a wider audience, I rely on data-driven simulations and statistical programming as ways of reconstructing elements of the recommender system.
-In order to explore the operational environment she described, and its predictive operations, I model key features of her presentation using sample grocery data drawn from the Tesco API ([Application Programmer Interface](https://devportal.tescolabs.com/docs/services/56c73b1bf205fd0ed81dbe7a/operations/56d5b2bee2813e0a0053c47c), [@Tesco.com_2017])   and a range of data science, statistics and graphics libraries implemented in the widely-used statistical programming language `R` [@RDevelopmentCoreTeam_2010].
-At a methodological level, it draws heavily on the vision of 'interface methods' proposed by Noortje Marres [@Marres_2015] in order to keep in view both the highly technical orderings of lists in recommender system and the changes in social orderings they might imply or impel.
-
-# How the shopping list was reconstituted in the 1990s
-
-The main narrative of Patel's presentation concerned a shift from a well-established loyalty-card based data-mining model to a predictive, probabilistic, personally relevant re-writing of the shopping list in almost-realtime.
-The changes Patel described are increasingly widespread and common.
-While they are configured in Tesco-specific ways by DunnHumby (and this reflects a longer history), they are also broadly typical.
-Tesco is the largest supermarket chain in UK, famous amongst retailers for its  customer loyalty and targeted marketing programme known as 'Tesco  Clubcard,' which started in 1991.
-DunnHumby -- founded by operations researchers Edwina Dunn and Clive Humby -- is said to have convinced the CEO of Tesco sometime in 1991 that a loyalty card program could change the supermarket chain's relationship to its customers.
-Clive Humby's academic publications  are hard to track down, but an early paper given at the _Conference of Young Operational Researchers_ in Nottingham in 1984 suggests the direction that he, DunnHumby and later Tesco would take in constructing lists [@OKeefe_1984].
-The abstract for Humby's presentation pre-figures an ongoing trajectory for data mining techniques aimed at eliciting detailed information on individual customer preferences.
-
-```{r humby-1984, fig.cap = 'Abstract from a Clive Humby presentation (Humby,1984)'}
-include_graphics('figure/2_humby_young_org_1994.png')
-```
-
-While Tesco  succeeded in data-mining its customers using demographic segmentation,  and perhaps became the UK's biggest supermarket with the help of data-mining in the 1990s, the shopping environment  in 2017 is markedly different [@Turow_2017].
-It is no longer organised around campaigns involving special offers or redemption of points for demographically segmented loyal customers (DunnHumby made heavy use of UK Census data).
-It can no longer rely on placement of goods in carefully chosen locations in stores.
-It needs or at least might want to continuously up-sell and cross-sell to customers who might hardly ever visits the supermarket itself.
-
-How could we characterise the shift from ClubCard data mining to online grocery markets?
-If 'Tesco is the clear winner in the online grocery market, in fact it takes almost 50p of every £1 spent on food shopping on the internet' [How SEO helps Tesco to dominate the online grocery market](https://econsultancy.com/blog/64841-how-seo-helps-tesco-to-dominate-the-online-grocery-market)[@Silverwood-Cope_2014], then Tesco itself has undergone some kind of conversion?
-Patel described her work at DunnHumby as shifting the recommender system from a 'rules-based list' to a 'relevancy model.' The relevancy model changed how the 'Have You Forgotten' list was constructed.
-This shift is a typical of broader re-organisation of prediction in which existing data analysis practices are being re-distributed and intensified in particular ways.
-The shift in models results in a more probabilistic structuring of lists.
-The ostensible banality of the 'Have you forgotten?' recommendations betrays micrological signs of a topologically complex predictive infrastructure that constructs alignments between people and things in significant ways.
-
-# Reconfiguring recommendations
-
-Academic researchers first began writing about recommender systems in the mid-1990s.
-From the outset they highlighted a potential shift from demographic-driven market research or data mining techniques to personalized recommendations.
-For instance, writing in 1997 in a special section of the _Communications of the ACM_ on recommender systems, Paul Resnick and Hal Varian (at that time Dean of Information Sciences at UC Berkeley, but currently Chief Economist at Google), made much of this personalisation [@Resnick_1997].
-Resnick and Varian  emphasise the need to distinguish the emerging practices from data mining:
-
-> In everyday life, we rely on recommendations from other people.
-...
-Recommender systems augment this natural social process.
-In a typical recommender  system, people provide recommendations as inputs, which the system then aggregates and directs to appropriate recipients.
-In some cases the primary transformation is in the aggregation; in others the system’s value lies in its ability to make good matches between the recommenders and those seeking recommendations.[@Resnick_1997,56]
-
-Writing just after the advent of web-based e-Commerce, Resnick and Varian  imagined recommender systems augmenting the 'collaborative filtering' that people do when they write and read reviews of products and services (for instance, on the travel website, TripAdvisor; [@Scott_2012]).
-In 1997,  Resnick and Varian thought that 'people provide recommendations' and recommender systems would aggregate and rank for recipients.
-A slightly later review, [@Schafer_2001], diagrams an augmented 'natural social process' with a range of elements, technologies, inputs and outputs, with varying degrees of personalization.
-
-The conversion of  collaborative filtering into predictive personalisation appeared in Patel's presentations.
-Patel describes the move from 'a rules-based system'  to  a 'personal relevance model'.
-Each term in that phrase -- personal, relevance and model -- has some weight in the conversion event.
-We can assume that what Patel describes as  the 'rule-based system' refers  to the `apriori` algorithm, developed  by Rakesh Agrawal and Ramakrishnan Srikanti at IBM Research Alameda in the early 1990s [@Agrawal_1994].
-A now-classic approach to 'market basket analysis,' it was listed as one of the top ten data mining algorithms in a survey conducted amongst data miners [@Wu_2008] and usually attracts a chapter in data-mining and machine learning textbooks (e.g.
-[@Hastie_2009]).
-
-```{r arules-ex, cache=TRUE, fig.cap='Frequency of association betwene items in the `Grocery` dataset', message=FALSE, echo=FALSE, fig.margin=TRUE}
-library(knitr)
-library(arules)
-library(arulesViz)
-library(datasets)
-data(Groceries)
-itemFrequencyPlot(Groceries,topN=10,col=rainbow(10), ntype="absolute",  cex.axis = 0.55)
-```
-
-```{r arules-ex2}
-options(digits=2)
-rules<- apriori(data=Groceries, parameter=list(supp=0.001,conf = 0.15,minlen=2), appearance = list(default="rhs",lhs="whole milk"), control = list(verbose=FALSE))
-rules_sorted <- sort(rules, decreasing=TRUE,by="confidence")
-kable(inspect(rules_sorted[1:5]), caption='The first five association rules for the `Groceries` dataset')
-```
-
-The `apriori` algorithm finds sets of items that commonly occur together in transactions.
-These commonly occurring sets are expressed as 'association rules.' For instance, using the `Groceries` dataset in the `R` package `arules,` we can readily see how Walmart *might* have discovered   the fabled association between diapers and beer.[^149.1] Figure \@ref(fig:arules-ex) shows how often common items appear.
-`Whole milk` appears most frequently.
-The association rule-based model for the `Groceries` dataset illustrates some characteristic ordering practices of 1990s-style data-mining approaches to lists (the `Groceries` dataset  was acquired from a 'local German supermarket' [@Hahsler_2006a]).
-The association between items is expressed in the form of _rules_ not a model.
-These rules, a term derived from older decision support literature, define sets of associations between items.
-Table \@ref(tab:arules-ex2) shows that the first rule has a stronger association than the second.
-Milk and vegetables is more likely an association than milk and buns.
-A rules-based recommender system should, therefore, might put them in that order in the 'Have You Forgotten' list.
-
-[^149.1]: One of the founding narratives of big data customer analytics, the beer-diaper association was not, according to [@Power_2002] found through data mining at all.
-
-Although association rules and the `apriori` algorithm present social order in terms of sets of frequency-weighted associations, an `apriori`-based recommender system contains no statistical model.
-It also no grasp on personal relevance since the associations it defines are only between things.
-Put in terms of the big data conversion, the rules-based system is not predictive, or personal.
-The `apriori` algorithm might be seen as an attempt to grasp the excess of things in a Tesco supermarket by identifying subsets whose closure derives from the associations between things that people construct (through buying habits, by virtue of a shopping list, etc.).
-In association rules, personhood or experience is irrelevant.
-The openness of these rules -- the `apriori` algorithm can product any number of rules of diminishing strength -- points towards the propensity of things to find themselves together in shopping baskets, but cannot cannot individuate that propensity beyond demographic quantiles and categories.
-
-Even if `apriori`  elicits associations between things, it struggles with the propensity of commodities to multiply.
-A simple illustration of the combinatorial problem faced by recommender system can be developed using the `Grocery` dataset.
-If we take all the items in the `Grocery` dataset and paste them into the 'shopping list' box on the Tesco grocery website (or as I side, run them as searches on the TescoLab Product API), each of the 169 items in the dataset yields dozens and sometimes thousands of products from Tesco online.
-
-```{r tesco-ex, echo=FALSE, cache=TRUE, fig.cap ='Tesco surplus',  fig.margin=TRUE}
-    library(ggplot2)
-    tesco = read.csv('data/reference_data/tesco_groceries.csv')
-    total = sum(tesco$actual, na.rm=TRUE) 
-    tesco = na.omit(tesco)
-    tesco$actual = as.integer(tesco$actual)
-    ggplot(tesco[tesco$actual>50,], aes(x=labels, y=actual, fill=labels)) +geom_bar(stat='identity') + coord_flip() + ggtitle('Tesco grocery items with more than 50 products') + theme(legend.position="none", axis.text=element_text(size=7))
-```
-
-Items in the `Groceries` dataset proliferate into a Tesco's list of branded variations.
-The 169 items of the `Groceries` dataset  expand into roughly `r total` Tesco items (see Figure \@ref(fig:tesco-ex)).
-There is, I would suggest, a logistic propensity in things that recommender systems encounter in the ways of global supply chains.
-The multiple listed  in the `Grocery` dataset becomes more open in this setting through new impersonal intersections that must be bounded and ordered.
-The proliferation of things on the shelves of supermarket or grocery warehouse confronts produces a combinatorial problem for data miners.
-`r total` products (actually Patel mentioned 200,000 products) can be combined in a vast number of ways.
-If a typical shopping list has 20 items, then there are `r format(choose(total, 20), 2)` possible lists.
-Some of this vast number of possible shopping lists are highly unlikely in terms of typical purchasing patterns, but potentially valuable as recommendations.
-
-# Ordering relevance through personhood or thinghood?
-
-Combinatorial surplus, however, does not exhaust what might be found in the data.
-The relevancy model that DunnHumby implemented for Tesco (as for so many other  businesses) seeks to balance the tensions between proliferating associations and relative closure.
-It responds to the problem of finding an ordering in which the propensities of things as commodities and the propensities of people mingle with each other, without overflowing  the site of their  ordering encounter, the list.[^181.1] Big data discourse in its promissory mode  attributes potency to personalization: 'most important, using big data we hope to identify specific individuals rather than groups' [@Mayer-Schonberger_2013].
-As Turow and co-authors conclude in their account of transformation of retail space: 'through it all, knowingly and not, and away from the spotlights of fierce social debate, retailers are encouraging daily routines that accept data-driven personalization as a centrifugal public force' [@Turow_2015, 476].
-
-Mark Hansen offers a different perspective on what might happen in these ordering encounters. Hansen attributes 'real potential' to data. Hansen attributes an associative propensity to data:
-
-> [S]urrounding any delimited predictive system is a larger field of data -- what I elsewhere call a “surplus of sensibility” -- that, viewed speculatively, indexes the causal efficacy of the total situation within which this delimited system operates.
-(Effectively, the latter gains its reliability from closing off this larger surplus of sensibility, thereby transforming an always excessive propensity into a [provisionally] closed dataset.) Because it affords data that exceed whatever any given predictive system might include, the data of the world’s causal efficacy -- the data constituting its real potentiality -- always and in principle facilitates knowledge that cannot be restricted to any particular agenda [@Hansen_2015a, 126-7].
-
-Many affirmative and critical analyses of big data and predictive analytics highlight the divisive social orderings implemented through personalization.
-For good reason, they attend to changes in personhood.
-Hansen, by contrast, seeks to conceive of the potential of data to support or provoke different relations to the world.
-Even if the underlying contrast between an open world and a 'provisionally closed dataset' limits the account in ways that I will explore, the ontological transformation  Hansen describes  involves people, commodity-things, and infrastructures.
-The recommender system generates common codings for  the associative propensities of things, the logistical value-chains of supermarket commerce and  the forms of personhood associated with shopping.
-A model of personhood based on purchase data delimits the propensities of things.
-
-[^181.1]: Rather than seeing predictive analytics as a closed form of an open world, I am suggesting that we might begin to locate it in the joint effectuations or intersections within models.
-
-If we think of the  new model that Patel describes as the delimited (or delimiting?) system within   the larger surplus of sensibility, how would the idea of a predictive system as a provisional enclosure of the world's causal efficacy, its 'surplus of sensibility' (a terms that carries some connotations of the critical political economy of surplus value) help us understand what the 'Have you forgotten?' recommendations do?
-It enables us to understand -- as we will see next -- what the Tesco relevancy model borrows from the world.
-By contrast, despite its ontological ambitions, it limits our understanding of how predictive analytics  themselves exist amidst technical ensembles whose operation remains obscure yet central to contemporary worlds.
-
-```{r models, fig.cap='Precision measurements for different machine learning models'}
-include_graphics('figure/patel_compares_models.png')
-```
-
-Patel described the data that Tesco keeps on grocery purchases: 'We have 52 weeks of purchase data for each customer.' These transformations have been widely discussed in the context of big data, where questions of who has how much data are usually most prominent (see [@Kitchin_2014]).
-In terms of the ways that lists order the world, the amount of data is not perhaps not so important as the way in which list-making is re-ordered via a model of personhood.
-Patel introduced the new relevancy model with a graph familiar to machine learners and statistical modellers.
-(A sketch of her graph appears in Figure \@ref(fig:models).)  The graph plots the _precision_ -- the proportion of the recommended products that customers actually purchase -- for several different models.
-In Hansen's media theoretical terms, the graph indexes the 'causal efficacy' of the recommender system, its capacity to include and transform propensities or 'real potentialities' into operational knowledge.
-The graph compares the old rule-based recommender systems with some of the relevancy-model alternatives -- logistic regression, random forests, and a few others -- in terms of their predictions and how those predictions turned out.
-Patel dismissed most of the models quite quickly and focused only on one, the logistic regression model, which did better than other alternatives.
-Why did she say little about the model, which after all, produces the probabilistic predictions on which everything in the recommendation list depends and hence the conversion experience?
-Her interest lay in the precision of the recommendations, and, crucially for our purposes, in how 'most of the gain comes from thinking of good features.'
-
-The models Patel so briefly mentioned are all classifiers, predictive models that spatially order  variables in a dataset as a high dimensional vector space.
-Operating in the vector space, models classify people and events.
-Much of the operational formation of the classifier models received scant reference in Patel's presentation.
-It is likely that in DunnHumby's models the classifying model at the heart of the recommender system constructs a decision boundary in vector space between points that belong to different classes of things [@Mackenzie_2015e].
-The classes of things are binary: `recommended` or `not recommended,` and for many of the models shown in the graph, membership of a class is expressed as a probability.
-The logistic regression model generates probabilities for each product for each customer.
-Patel did not articulate the implication of the contrast between the left and right hand side of the graphic.
-Although the statistical framing appears explicitly (e.g.
-in measuring the precision of  recommender systems), the underlying shift to a statistical model is barely mentioned.
-In becoming predictive, the 'Have You Forgotten?' recommendations have included a version of the social in the form of a  model that seeks to identify relations between variables.
-The relevance of recommendations pivots on the apprehension of social order by a model.
-
-# Sparsity: output 20 recommendations from 200,000 products
-
-Note that neither the data analytic infrastructure that processes  Tesco customer data  - Patel will mention `Hadoop` -- nor the predictive analytic device -- the logistic regression model -- loom very large in the conversion of the recommender system.
-'Most of the gain comes,' Patel observes, 'from thinking of good features.' While this is almost a truism in big data practice (see, for instance, [@Domingos_2012]), what counts as a good feature in a shopping list?
-
-Rather than a surplus of sensibility, predictive models often encounter a dearth of sensibility in the data.
-A  very brief comment from Patel -- 'we have lots of zeros' -- points to the difficult terrain of  'good feature' thinking: the vast, nearly empty matrix of customer-product associations.
-Given that most things in Tesco remain relatively unknown to each other and to customers.
-A matrix that records associations between individual people and different products is bound to be mostly empty.
-Say Tesco has 1 million online customers.
-Each online  shopper has bought some selection of the 200,000 products.
-The customer-product data matrix will be `r format(1e6 * 2e5)` in size.
-The product-customer matrix, the basic vector-space in which all recommender systems operate, remains very sparse and unpopulated.
-Given that any one customer is likely to only have bought 100 or so different products, the probability matrix will be close to `r format(100-(100/2e5)*100, 2)`% empty.
-The data is, as Patel put it, 'massively unbalanced' (any dataset where the items of interest are much rarer than some normal values is said to be unbalanced) and such imbalance would heavily bias the recommender system towards common and only impersonally relevant suggestions, suggestions that might  not produce the desired conversion experience.
-
-```{r sparsity-ex, echo=FALSE, cache=TRUE, fig.margin=FALSE, fig.cap ='Sparsity of the customer-product association matrix' }
-library(ggplot2)
-library(reshape2)
-p = as.logical(rbinom(n=2e5, size=20,0.001 ))
-m = (matrix(p, nrow=1000, byrow=TRUE))
-melted <- melt(m)
-melted$value <- as.logical(melted$value)
-ggplot(melted, aes(x = Var2, y = Var1, fill = value)) + geom_tile() +
-    scale_fill_manual(values = c("grey", "red")) + xlab('Customer id') + ylab('Product id') + theme(legend.position='none')
-```
-
-Since so many people buy milk, the recommendation system might end up always recommending milk products.
-So the data needs to be 'corrected' by, as Patel reports, removing -- under-sampling -- some of the data for common purchases.
-Secondly, even if a recommendation system finds uncommon items that are good recommendations, they should neither be too cheap ('low spend'), nor, as we will see too similar to what the customer has already ordered.
-
-Given this imbalance, distortion and deficit of sensibility in the data, 'thinking of good features' -- ways of traversing the data -- becomes central.
-For our purposes, the exact details of these features matter less than attending to the way in which, contrary to Hansen's account of analytic closure,  probabilities become propensities in data practice.
-The dataset, as we will see, will not be practically treated as 'closed,' as Hansen puts it.
-Although it undergoes acts of enclosure (for instance, in the transformation of the customer-product matrix into vector space), these acts are not conclusive.
-'Thinking of good features' potentially re-opens the dataset and sometimes completely transforms it.
-This transmogrification, which I suggest occurs mostly in quite obscure yet central zones of data practice, is highly mixed.
-It combines experimental techniques, infrastructural dynamics, seasonal and daily temporalities, and above all, an ongoing engagement with probabilities and their fungibilities.
-
-# Experimental settings: A/B allocation and uplift
-
-|Test A|Test B|
-|---|---|
-|Control A|Control B|
-
-For instance, the predictions of the recommender system themselves are the subject of probabilistic treatment.
-Patel described the experimental deployment of  the relevancy model experimentally in a random A/B controlled trial on the Tesco website.
-All customers were  allocated to one of four categories as shown in the table.
-In the  A/B trial,  an identical shopping list, or even the same person, might receive different recommendations.
-The randomised application of the model, drawing on protocols for randomised clinical trials first developed in the 1960s, is widely used in social media platforms and hence in the implementation and observation of the effects of recommender systems.[^238.1] Random allocation of customers to the four categories (Test A, Test B, Control A, Control B) adds a layer of probability to the recommender system in the name of statistical validation of the effects or the 'uplift' of the model.[^122] Ironically,  the effects of a predictive model cannot be known in advance.
-They can only be observed experimentally.
-
-[^238.1]: It is difficult to gauge how much. Facebook reports that it has more than 1 million models operating in its infrastructure. [@Dunn_2016]
-
-It is perhaps significant that this random allocation of customers to control and test groups occurs without any relation to the particular profile or propensities of the customer.
-It seeks to statistically validate the effects -- the *uplift* -- of the model on  conversion events by allowing the effects of the model on what people do to be measured.
-The uplift refers to conversion events associated with the same groups of people.
-Effectively an experiment in inhabiting two different worlds, the randomised control trial  sets up a feedback look between the predictive model (the logistic regression), and the world.
-
-[^122]: See [@Lo_2002] for an introduction to the construction of experimental set ups for uplift models.
-Market researchers and data miners seem to have started adopting a statistically grounded experimental predictive practice sometimes around 2000.
-See for instance [@Almquist_2001]  Response modelling, 'propensity' modelling and uplift modelling all seek to identify associations between 'treatments' or interventions and the 'Responders', the people  affected by the treatment.
-
-# Infrastructure: the realtime engine serves
-
-A second layer of probability affects the operation of the recommender system.
-Much discussion of recent media has highlighted its real-time operation (see [@Kittler_1993; @Virilio_1997; @Hansen_2015; @Bollmer_2016]).
-Patel's description of the relevancy model infrastructure sketched a different temporal patterning, suggesting immediacy in relation to shopping, but scheduling and periodicity in relation to the infrastructural operations of the predictive model.
-While raw data from Tesco Online feeds into the model every hour, a recommendation list for each customer is only generated once a week.
-Customers only shop online every few days at the most, and in some cases, only every few weeks.
-To continually update the top 200 recommendations for every customer would be computationally expensive.
-
-A customer does not see their 200 weekly recommendations.
-The recommender system generates recommendations, but after 'applying business rules, only five are  shown on the Tesco Online website.
-Patel briefly mentioned  specific infrastructural elements such as `hadoop` [@ApacheSoftwareFoundation_2009].
-The logistics of running a predictive model affect listing practices.
-The possibility of adjusting the recommendations for every customer weekly derives from an infrastructure capable of collecting data, assimilating that data to a predictive model able to generate approximately 200 million relevant recommendations, where 'relevance' depends on a matrix of probabilities of associations between people and things that shifts in time.
-Like the global supply chain logistics optimised for flow of commodities [@Tsing_2009; @Neilson_2012], `hadoop` and its legion of variants (`mahout`, `spark`, `hive`, `pig`, `yarn`) operationalise flows of data at an infrastructural rather than analytic scale.
-In many respects, the quickly glossed-over infrastructural deployment of DunnHumby's relevancy model is the primary conversion event: the logistic regression model at the heart of the recommender system is no longer an analytical device but an operational one.
-
-# Similarity, substitution, complementarity and micrological propensities
-
-There is more to probabilities in prediction than a simply enclosure and quantification of the surplus of sensibility, the overwhelming variety, differences and noise of the world.
-I am suggesting, contrary to Hansen, that the propensities of things are enclosed as  probabilities in predictive analytics only to the extent that probabilities themselves become propensities.
-Recommendations are relevant in a recommender system or any other operational predictive model to the precise extent of their probability.
-Relevance is closely connected to propensities in the predictive condition.
-From DunnHumby's standpoint, the relevance of a recommendation relates directly to conversion events.
-Conversion events happen in the world.
-When a recommender system discovers a common coding of human practice, and logistical configurations, with all their value differentiations, they actualise a propensity that hitherto remained potential.
-From the standpoint of the ontological exchangeability between probability and propensity,  prediction stems from the propensities of things, what Hansen describes as the 'complex and diffuse calculus' that only provisionally congeals as a conversion event.
-The 'micrological propensities' of things support the relevance of recommendations.
-
-How does the DunnHumby model deal with the exchange and oscillation between propensities and probabilities?
-In her presentation, Patel  spent much time addressing the problem of 'basket similarity,' the problem that the recommender system might recommend, for good reason, something similar to items already in a customer's basket.
-The recommender system or statistical model needs to know how things are related to each other so that it does not waste recommendations on overly similar items.
-A customer might be willing to substitute a similar item for something they have already chosen, but they are more likely to accept a recommendation for an item that complements items in their list.
-
-Whereas the association rule-based recommender system focused on thing-thing complementarities -- if you buy bread, you might want milk, if you buy beer, you might want diapers -- the personal relevance substitutions look for micrological propensities that relate to gradations of preference, and variations in time and place.
-Micrological propensities in the list relate to a particular time and place, amidst a continuously differentiable pathways between things.
-
-How could the recommender system sense the possible substitutions and complementaries between things for a given customer, especially since a person's sensibilities and susceptibilities concerning groceries might be shaped by wider social groups, orderings and circumstances?
-Patel's comments on this were very brief but suggest that the 'Have You Forgotten?' list becomes an increasingly mutable entity, open to many substitutions and intensive re-writings.
-
-Like the mention of `hadoop`, what Patel said about substitution and complementarity draws on extensive but relatively well-established data practices in machine learning.
-She mentioned,  for instance,  a 'design matrix of 14,000 columns' and the use of '`L1 regularization` to drive coefficients down to zero' in a new feature of the model called 'self-learning substitutes,' a model that examines the recommendation list and predicts substitutes and complementarities for 'all products' in the list.
-The 'self learning substitutes' feature of the relevancy model draws on another data matrix, the product similarity matrix, itself generated from all previous recommendations that have led to conversion events (that is, 'a user clicked as the response').
-A customer's recommendation list might already include `Tesco Ginger Nuts 300g.`   The 'self-learning substitution model' seeks to replace list items with similar but higher value items: `Mcvitie's Ginger Moments 400g.` What is being modelled here?
-No longer whether to recommend a product to a customer or not, but micrological conversions.
-
-Similarly, modelling complementarities between things for a particular person entails a matrix of comparisons.
-`Nutella` complements `bananas,` but the converse is not necessarily true since `bananas` need not complement `nutella,` especially if I have nut allergies.
-From the standpoint of customer data science, Patel emphasised the operational significance of predicting complementarities:  'complementarity drives conversion.' Bringing the matrix of similarities expressed as numerical associations between products into the model improves the conversion of recommendations into purchases.
-From the standpoint of a sociological account of social ordering that associate people and things through propensities, complementarities pose a different problem: what does it mean for social order if the relations between things shift in accordance with our experience of the world as rendered in the intricately varying weights of an L1-regularized logistic regression model?
-In the similarity matrix, shaped by the history of past recommendations and purchases made by this customer, the propensities of things drive down the surplus of sensibility associated with what people want.
-
-# Conclusion
-
-In their mundane intersectional practices of ordering, recommender systems show how  predictive practices affect existing forms of social order in the world.
-Rather than directly positing an ontological transformation (as Hansen does), or pointing to the erosion of social order (as Turow does), I have explored some of the conversion events associated with a shift from a data-mining rule-based system to a predictive recommender system.
-The DunnHumby model and the Tesco 'Have You Forgotten?' recommendations present many of the traits of the conversion processes associated with big data.
-Rather than discontinuous changes, the conversion of the recommender system via a relevance model suggests an ongoing problematization of propensities.
-We have seen several types of conversion events: ongoing transformation of models and infrastructures, the experimentalisation  of recommendations, and the broader diffusion of a calculus of probabilities into a fibrous weave of propensities and conditionalities, warping over time.
-The predictive assemblages connect logistical ensembles to the negotiation, compromises, rituals and local orderings of everyday life.
-
-I have been suggesting that predictive listing in the form of open-ended recommendation changes social order.
-It establishes a common coding between machinic and ritual repetition (as exemplified in the shopping list).
-It also suggests some ways in which repetition or reiteration  can also participate in changing and emergent order.
-By understanding the world in terms of propensities, we have a chance  to engage with contemporary predictive processes in ways that both acknowledge their relentless and exhausting pressure to consume more and retain some chance of inhabiting the world differently.
-
-Rather than understanding prediction as the closed form of an ontologically open world ('the propensities of things'), the DunnHumby relevance model, in all its statistical, infrastructural, organisational and banal complexity, suggests that the predictive conversion remains constitutively incomplete.
-Shopping lists in all their banal yet primary orderings are fertile terrain for prediction because they are incomplete, somewhat open and mutable forms.
-The constitutive incompleteness and openness of lists are very closely entangled with ongoing processes of ordering and othering, and at the same time, can accommodate predictive operations.
-Lists undergoing a 'conversion' experience as they become more probabilistic, but their heterogeneous character as operations caught up in models, infrastructures, experiments, and the flow of commodities means that they can never form a closed set or dataset.
-The problem of 'constructing good features,' a problem that any predictive model must address, persists as the site of experiments in partly open structures.
-
-A constitutive openness to new associations, new combinations, new groupings or actors is integral to the conversion events associated with contemporary data practice.
-In their account of the Netflix recommender system, Ted Hallinan and Peter Striphas make the following argument:
-
-> These systems, and more importantly their algorithms, play a critical role in deciding which articles (or parts thereof) gain admission to the cultural realm, and in what form.
-Their doing so thus points in the direction of another universe of reference—a court of algorithmic appeal in which objects, ideas, and practices are heard, cross-examined, and judged independently, in part, of human beings [@Hallinan_2014, 13]
-
-Despite the somewhat troubling reference to 'another universe,' as if Netflix or DunnHumby's recommender systems occupy a different reality, Hallinan and Striphas usefully and aptly highlight how such systems  mingle  objects, ideas and practices.
-
-Something similar happens to shopping in the Dunnhumby relevance model.
-In the customer-product association matrix, in the product-product similarity matrix, in the 14,000 variables included in the logistic regression model, a 'universe of reference' takes shape.
-The shopping list, the commonplace inscription of normative and actual social order, takes on a complex temporal-material structure that mixes prediction, infrastructural constraints, the individual characteristics of list-makers, statistical validations, database architectures and above all the surplus sensibility associated with  circulation and consumption of commodities.
-Shopping lists undergoing a conversion experience as they become more probabilistic, but the idea that probabilities are the closed form of propensities does not account for the distribution of probabilities in relations between things, between people, between people and things, and between things and people.
-
-# References

### 2017-03-28 lists_conversion_experience.rmd
this commit has the manuscript formatted as a word doc

@@
 This suggests that the conversion of lists from finite ordering to indeterminate propensity in recommender systems requires new ways of conceptualising social order."
-keywords: prediction, probability, order, recommender system, shopping
+keywords: prediction, probability, order, lists, recommender system, shopping
 ---

### 2017-03-28 lists_conversion_experience.rmd
fixing author affiliatino problem

@@
 title: "Probabilities and propensities: conversion experiences in online grocery shopping"
-author: Adrian Mackenzie^[University of Somewhere] 
+author: Adrian Mackenzie, Sociology, Lancaster University
 date: "`r format(Sys.time(), '%B %Y')`"

### 2017-07-12 reply_to_reviews.md
planning revisions

Thanks very much to reviewers for their work on what must have seemed like a frustratingly elusive and adumbratory piece of work. The reviews have been a great help in re-writing the piece. The reviews were quite long, so I won't summarise all the changes I've made in response. There were many. My summary of the  major issues identified in the reviews  is below.

    1. Perceives **lack of cohesion and clarity in the argument;** re-frame opening to be less about commodities and supply chain, and more about problems of association, correlation, categorisation, etc.; establish a consistent role for the list; elaborate the methodology of reconstruction; clearly say what kind of transformation occurs in the transformation, and in the propensities; and what is the 'so what' of these clarifications and conceptualisation? 
    2. Asks for more substantial empirical engagement, reorganization of the piece and **clarification of overall argument and purpose,** and specific section purposes; several dozen detailed points and calls for correction; 
    3. Asks for clarification about the contribution of the piece; description of the case is vague in detail; look at Bucher and Barocas for state of the art; deal with disconnect between abstract and actual paper, especially around social order; **main argument is incomprehensible.**

### 2017-07-12 log.rmd
planning revisions

# Lists siegen october 2016

## possible titles
Lists that always change and lists that hardly ever change
What is not a list?
Lists as prediction
List heterogeneity: change and stability
Predictive listing
Lists and prediction
Lists as predictive setting
List in a predictive configuration
Streams as lists: the problem of lists in time
Streams, pipelines and trees: living with unending lists
**Have you forgotten? List experiences in our predictive condition** ...

## simple examples of lists
- `ls` -- what is the history of this? See Stallman and MacKenzie; also cat, etc. 
- indexes -- Abbott on this
- indexing of elements in data structures and in mathematical forms
- list on wikipedia
- working memory and its addressability
- job listings -- the Kaggle contest on this - but it is only in relation to testing 
- program listings and the problem of branches and the flow of control
- image as a list -- AlphaGo
- text as a list -- RNN
- discussion amongst programmers of why they hardly ever use 'lists'
- python list comprehension
- intel i7 and samsung branch prediction
- lists of ingredients in Chef Watson
- blockchain as an unchangeable list
- streams as indefinite lists
- Godel numbers as lists
- 'have you forgotten' list at Tesco

## Theory
- Fuller quotes
- Whitehead on sequences and series
- Stuart on sequences
- Badiou on sequences
- lists article in _Devices_ book
- hansen on prediction

## to-do
- look at information retrieval textbooks -- Manning, etc
- go through Whitehead mathematics notes on sequences and series: DONE
- go through _Elements of Statistical Learning_ for stuff on lists and ranking: do they deal with it?

## Tue Aug 23 08:41:32 BST 2016

- 'list' of ideas for this paper
- lists as sequences, and how sequences converge: this is only known through summation (Whitehead) and convergence to a limit; which links lists to neighbourhoods and ideas of convergence; via sequence, can also get into vector space
- listings -- program code -- and lists of instructions executed by programs  -- and what differences listings make; why doesn't anyone really refer to program listings anymore
- the list as predictive site -- how are lists becoming predictive: this could include many ranked lists (e.g. search results); -- but also the very architecture of the CPU itself as a predictive process.  -- the branch prediction idea and the way it allows pipelining of processes
- ranking and lists: the obvious connection is ranking, but what if I don't want to do ranking? 
- the substrate of lists and their space -- a topological account?
- focus on the operations of listing, including indexing, ordering, ranking, rewriting, and retrieving
- does the work on working memory, or on recursive processes help in exploring what is going on with lists?
- could keep things simple and just focus on 'listings' dealt with in machine learning - how is a ranking actually done in  a model and why would that matter. The answer is probabilistic?  
- lists and the problems of scale

## Wed Aug 24 09:09:36 BST 2016

- what if lists are fundamentally changing in nature because they are becoming probabilistic, vectorised and optimised.   The key question would be how to demonstrate that. 
- large lists, indefinite lists, infinite lists ... how do they matter?  Because the series they produce becomes a kind of limit
- second main idea: massive expansion of lists -- everything is a list (e.g. images, text, etc)

## Fri Aug 26 11:14:10 BST 2016

 - finalising title; had idea about lists that change and those that don't; lists that adapt to individual circumstance and those that seek maximum stabilisation; are there enough tensions between the agent and structures, between change and stability; between control/order and event/invention to explore here? Tensions between predictive or probabilised lists (that adapt fully to context) and those that seek to minimise change (blockchain, working memory, sequence forms,etc.)

## Mon Sep  5 13:27:05 BST 2016

- started looking at Godel numbers and Godel sequences, and from there got to the lists entry on wikipedia. Can think of stream as an infinite list, and this seems to be a useful way of thinking about lists in their contemporary form -- they are open ended; they contain time-orderings, but these orderings are not fixed. Could look at stream processing as an extension of lists; and then run an argument about scale. Do we oppose scale? Do we scale? What about scaleability?

## Mon Sep 12 15:19:55 BST 2016

- just listened to Shreena Patel from DunnHumby speaking about 'Have you forgotten?' in the shopping basket at Tesco. Really interesting description of the pipeline for the recommendations that are shown at the end. Could base the whole talk around this rather than the high-falutin talk I've sent in. Shopping lists are the prosaic example I've been looking for. 
- see pages  43-47, **August 2016** notebook; now scanned as figure/single_patel.pdf; also single pages as figure/patel_page3_SKM_C36816101109420.pdf and as figure/SKM_C36816100609150.pdf and figure/SKM_C36816100609170.pdf and figure/SKM_C36816100609171.pdf.

## Mon Sep 26 08:34:00 BST 2016

- went through Shreena Patel DunnHumby presentation and turned it into around 10 points that could be the basis of the presentation. 
- would be good to have her presentation. 
- emailed her for it
- put the 10 or so main points into presentation order and started to write some notes around these
- could also point to (Recommender Systems Conference)[https://recsys.acm.org/recsys16/] as an example -- spent some time looking through these -- interesting papers on Youtube deeplearning, which I put in archive (~/archive/data_intensive/machine_learning/covington_youtube_deeplearning_2016.pdf). See there how matrix factorization has been been the main approach for recommender systems. 
- would like to simulate some of the things that Patel talked about. 
- got paper on Trilling, Kinsey Report -- started to think about all the accumulation of data in recommender systems. How long has this been happening? Tesco Clubcard seen as an origin, but customer/business analytics must be much older. How older? DONE: WoS search on research literature - DONE
- did a bit of this -- first mention of the term is highly cited 1997 ACM article by Varian & Resnick; describes some existing systems; point to the term 'collaborative filtering' as first used in 1991; paper is in archive ~/archive/data_intensive/machine_learning/varian_resnick_recommenders_1997.pdf
- 1997 special issue of ACM on recsys? Not sure. Went through time in WoS a bit and found 2001 article on recommender systems -- Schafer et al ~/archive/data_intensive/machine_learning/schafer_recommenders_ecommerce_2001.pdf -- which is a bit better. Has a lot of taxonomies and late 1990s references 
- DONE: go through the main points and think about where I can put my own stuff in; DONE 
- DONE: look up ethnomethodology on lists ...
- DONE: DONE go through book manuscript looking for points of intersection - could be on population; the table; optimisation/function; probabilisation, dimensionality; enunciative function;

## Tue Sep 27 09:12:58 BST 2016

- started looking for any literature on Amazon recommendation engine -- found their ML service; much discussion of how important recommendations have been ... downloaded Linden's amazon paper; quite useful for all its material on the problems of large scale ... 
- tried to find Humby's work -- only one paper in Journal of Market Research, 1999, 'New developments in demographic targeting--the implications of 1991.'i about the ACORN system; this 1989 paper not n the archives of Market Research Society -- too early.  Found abstract for a paper given at 'young OR conference' in 1994. 
- Took screenshot of Humby 1984 and saved in figures folder -- it will be a slide ... 
- following leads on collaborative filtering -- how it differs from market basket analysis; how item-based and user-based CF techniques differ; tutorials in R on these
- also thinking about ordinals and well-ordered sets again: recommendation lists are well-ordered sets; in that sense they are something like ordinal numbers; that is, ways of describing well-ordered sets. 'Ordinal numbers are thus the "labels" needed to arrange collections of objects in order.' writes Wikipedia. 
- did search on 'lists' in my notes and made file lists.md. Has useful stuff from Lury, Lazzarato, Fuller, d 
- looking at this blog (Salem Marafi)[http://www.salemmarafi.com/code/market-basket-analysis-with-r/], installed arulesViz and played a bit with Grocery dataset.  
- wrote an intro to list, highlighting problems of how to study them as forms of order, and how to participate in them
- NOTDONE: follow up on the Hobbes problem of social order: are new forms of operational changing that social problem. No longer contract, economic relations, etc. ... 
- feel better about getting some code for arules in the paper -- this is something that I've wanted to do anyway
- brought in all of Salem Marafi's code just to play with it -- will need more commentary to make sense of it. Am I saying that the shift in models from arules to CF is important? In what ways?

## Wed 28 Sep 09:33:36 BST 2016

- will look at the most interesting bits of dunnhumby presentation first, and then see if the other bits can be fitted in time ... 
- wrote a bit about arules and apriori, trying to link it to set theory; see ~/archive/data_intensive/machine_learning/agrawal_apriori_94apriori.pdf
- DONE: find old shopping list and scan it DONE
- added a bit of badiou, but I'm not convinced.
- got distracted by non-human book (Grusin) and spent time downloading Hansen's article on prediction. And then Bogost on carpentry. 
- added a bit more on the badiou front -- the open and excess.
- DONE: image of a ClubCard -- DONE
- spent time pasting itemset from `Groceries` dataset into Tesco's shopping list. Data is in data/grocery_items.csv. Can paste the whole list into the shopping site, or use the url (http://www.tesco.com/groceries/Product/Search/Default.aspx?notepad=yogurt&newSort=true)[http://www.tesco.com/groceries/Product/Search/Default.aspx?notepad=yogurt&newSort=true] substituting any term you want into it to get the results.  Would need to scrape a bit to get them all. Can get the number and then step through using http://www.tesco.com/groceries/product/search/default.aspx?notepad=chicken&viewMode=grid&N=0&Nao=40 
- tracking down the `Groceries` dataset -- found reference in reference.pdf for `arules`. Added this to archive ~/archives/data_intensive/machine_learning/ 
- what do the brands do?

## Thu Sep 29 09:31:46 BST 2016

- Played this game at the tea table:   'I went shopping and bought '  -- this game is an example of a well-ordered set, since it labels every item with a letter of the alphabet, and those letters are well-ordered.  
- the game is about remembering what is in the list, and the alphabet helps remind you.
- could use this game at the opening.  The baseline account of lists as a form of number. 
- wrote a bit about this at the beginning -- it feels a bit better than other stuff because it gets into sets
- added in Hansen on prediction and how probability becomes actual ... Also put quotes in new notes file. His argument that prediction is a closed probabilistic determination of propensities that always exceed it as a field of data. The goal is to locate those excesses and perceptualise them somehow.  
- DONE bring the quote about technical interface to the data of sensibility into the conclusion
- read some Hansen at lunch and marked useful bits on paper copy -- DONE:  DONE put those into notes file and then use them to structure the argument?
- started to bring some of the probability idea into 'I went shopping and I bought ...'

## Fri Sep 30 10:16:16 BST 2016

- pfaffing with rmarkdown, setting up shortcuts, etc. will be a help I think 
- trying to get to core of the issue- what is probabilistic and potentiating in the sense that Hansen speaks about ...  
- wrote python script to scrape all the product counts from tesco in the groceries list; took 1 hour; to get a total figure that is accurate I suppose
- saved results as csv data/tesco_groceries.csv and a did of them 
- eventually got somewhere with the Tesco data. Plotted it (> 50); 
- started writing about the number of combinations of these items they constitute a surplus of sensibility.  
- wrote some points for the conclusion

## Wed Oct  5 07:48:28 BST 2016
- have written about 2500 words, but need to get to the point: how does a subject of the list take shape in the data
- have finished reading the Hansen which is at ~/archive/Hansen feel like I want to get away from any opposition between the actual and the statistical, and say that the feeling of agency comes from the list.
- like the argument that subject is a feeling of the list ... and Hansen's argument in general about 'our predictive condition' 
- 'condition' reminds of Lyotard - downloaded and added to archive here ~/archive/Lyotard_Postmodern_Condition_1981.pdf great line there about continuous differentiable functions won't be the basis of prediction ... this quote in notes file ~/notes/lyotard_postmodern.md.  
- DONE: get quote on conversion and **relevance** and why it matters from Hansen
- also had an idea of thinking about an event that happens in recommender system -- does it ever happen that something is recommended that matters to me ... 
- TESCO api -- signed up for this but can't get key for API for some reason -- can't see it there, but its probably a distraction ... 
- DONE: DONE scan Shreena's graph of the models and use as image. 
- went back to notes on 200000 products and looked at undersampling -- randomly drop some of the products to get a more balanced dataset; also see that much of the presentation was really about the topic of basket similarity and how to deal with it since it affects the **relevance** of the the recommendations. This is another tie-in to the Hansen argument. 
- thought too that the conversion experience Tesco wants -- no low value conversion -- is precisely what they are suffering from in competition with the German supermarkets that do no online delivery.

## Thu  6 Oct 09:27:40 BST 2016

- scanned notes from notebook as figures. They are figure/single_patel.pdf as well as in single files 
- how to deal with surplus -- is that the issue? Could talk about the undersampling of the 200000 products.  
- went back to the TESCO api and actually managed to register properly
- wrote a small python script to get items -- actually more interested in the data that the api gives back. Its quite extensive in its attempts to qualify each item.
- got a bit into the 'unbalanced data' issue and started to write about that.

## Fri Oct  7 10:30:53 BST 2016
- writing scripts to get tesco products -- now have that working and can make montage of whole collections of products. tesco_get_groceries.py
- not sure why I did that -- can see no use for it except to a sense of the surplus of sensibility by showing montages. But what is the point of that?
- ran scripts enough times to get around 20k products. Might be overlap between categories, but this is about 10% of Tesco's product range in groceries .. 
- had idea of trying to simulate strange interests by working with the tesco data and then building models that could predict those.
- could also use the images to build classifiers; or use all the product metadata to construct new features and classifications
- had idea of trying to simulate strange interests by working with the tesco data and then building models that could predict those.
- could also use the images to build classifiers; or use all the product metadata to construct new features and classifications.
- added a bit of stuff on the experimental allocation of customers
- also started adding fragments of my notes as images; I'm hoping that I can write a script to extract the images and the knitr blocks into a presentation file and then present that.
- wrote script simulate_baskets.r to pull all the results into a single dataframe for tesco items and then use that to run queries against. 
- DONE: get some ideas on conversion from James' ~/archive/james_texts/james_varieties.pdf
-  MAYBE: some more work on simulating the baskets and perhaps the customers using the GBCS data
- started to write a script to simulate baskets in terms of range of items, total cost, number of items,, etac 
- DONE: write r chunks in outline to extract high cost conversion items DONE

## Mon Oct 10 09:27:43 BST 2016

- reread hansen a bit to get a sense of the actual argument: the main thing seems to be that it is not the event itself that matters but the organisation and accessing of the propensities in the world that come to light. 4  
- also the hansen point about re-thinking media as distributed in the world
- tried to track down the Vernes reference -- 1982, critique de raison aleatoire -- would need to ill that.
- wrote some on the uplift model -- turns out that has only recently come into marking (around 1999) and is under active development in market research. I thought it would be standard practice ...  downloaded a paper on it by Victor Lo 2002 ~/archive/data_intensive/markets/lo_lift_model_1982_10.1.1.99.7064.pdf
- uplift models are one of the ways that markets link machine learning into marketing ... 
- DONE: DONE write script to extract all headings, quotes, images and knitr chunks to a file and then convert to ioslides, etc. DONE
- DONE: get some references for realtime in media: Hansen doesn't speak about this much -- might be able to adapt some of his temporality talk. ... 
- tried t write the infrastructure section -- not much to say about hadoop, except that without it, no lists ... is all of this a product of infrastructural transformation?

## Tue 11 Oct 09:27:24 BST 2016

- DONE: use montage to create image of all out.jpg as a slide
- had to rescan patel page 3 because it was cropped figure/patel_page3_SKM_C36816101109420.pdf
- went through paper adding a bit on classifiers and tangentially approaching the topic of basket similarity. 
- finding it hard to write about that bit of the presentation  - some small cues -- 'modelling substitutes'. Does that mean that they build models for reach product? Or for each of the recommendations in the 200 in the list? That is, a second layer of modelling that connects the recommendations to what is in the basket? Shreena talks about it in that way ...

## Wed 12 Oct 11:41:17 BST 2016
- hard to get started today -- 11.41 and just starting. Spent quite a lot of time on poLCA and this doesn't have relation to the topic.  
- will try to write about the substitute and similarity modelling
- looking up Popper on propensity and downloaded to ~/archive/popper_propensities_1990.pdf; also printed it
- read it a bit and it might help in the discussion of the weighting of the list
- had the idea in the pool that I don't need to go much further except to say something about how media theory is grappling with the growth of prediction; premediation, etc. This also is an attempt to grapple with that. DONE: put into intro and conclusion  DONE
- added extract_images.sh to create slides; DONE: try it with Rmarkdown reveal, etc. DONE

## Thu 13 Oct 15:58:47 BST 2016

- didn't get much time today, but did a little on slides -- putting in chunk options, which didn't seem to work, and trying to get slides working. Some of them come out, but not all.

## Sun Oct 16 20:43:09 BST 2016

- playing with tint formatting, but didn't really get it working.
- managed to mess up the repo a bit by forgetting to pull down

## Mon 17 Oct 12:55:25 BST 2016
- still playing around with tufte, tint etc. Got it working reasonably well after an hour. Good example document is here: https://raw.githubusercontent.com/rstudio/tufte/master/inst/rmarkdown/templates/tufte_html/skeleton/skeleton.Rmd
- also found Morris paper on recommender systems that I should read as point of comparison:  DONE read and extract key points from ~/archive/data_intensive/morris_recommender_European Journal of Cultural Studies-2015-Morris-446-63.pdf and added some notes here ~/notes/morris_recommender_2015.md

- key thing to get it working is:
>library('tint'), render('test_tufte.rmd',output_format=tintPdf())

or
>render('test_tufte.rmd',output_format=ioslides_presentation())

- made a small script for this and added it as render_tufte.r
- modified it to run for lists_conversion_experience.rmd

## Tue Oct 18 10:55:59 BST 2016

- still need to write something on the similarity matrix; how does it fit into the narrative? 
- playing around with figures in the margins instead of working on the similarity matrix
- wrote a little on the similarity matrix, but it still doesn't go far. 
- building an image of similarity matrix using all images; getting more images for snacks, sauces, etc. Idea is that I will be able to say something about propensities and hence conversion(?) by having these images in vast montage. 
- finished making the image, but hard to get it showing in the document ... maybe just too big

## Wed 19 Oct 09:24:20 BST 2016

- figure/propensity_conversions.png works ok at work laptop. Not sure why not showing at home. Some problem with eog. 
- trying to write abstract to send, but not sure I have one. 
- wrote abstract and sent it

## Thu 20 Oct 11:10:53 BST 2016

- changing font size on graphics2
- added a bit to the conclusion and then printed to get something to work with
- doing graphic for matrix of zeros
- and trying to read everything

## Fri 21 Oct 11:07:10 BST 2016

- laid out all pages on the whiteboard and looked at the flow of the argument
- making some changes, including a slide of early recommendation systems from 1990s such as described in ~/archive/data_intensive/machine_learning/schafer_recommenders_ecommerce_2001.pdf
- feeling very distracted, but also adding ~/archive/data_intensive/machine_learning/varian_resnick_recommenders_1997.pdf
- also did some re-formatting of tables and stuff to cut down on the output. 
- Feels like the bigger ideas around transformation in probability are still being sorted out here. Do I think that the probability practices associated with the personal relevance model can be understood as propensities of things in the world? Or am I arguing for more intermediate positioning? That is, we don't to do an infrastructural transformation in order to account for the conversion experiences. 
- Maybe I can play around a bit more with conversion experiences in theory.

## Mon 24 Oct 15:51:50 BST 2016

- trying to fix up the flow a bit. Adding quote from Varian and Resnick. 
- fleshing out some sections of the discussion by adding more commentary -- got to page

## Wed Oct 26 09:43:43 BST 2016

- did some revisions on paper yesterday, but didn't manage to enter them all.
- realised during revisions that predictions are in the product-customer matrix, not after that. Important for my argument that   
- had to print out with revision halfway through ... 
- working on train and flight. Re-wrote conclusion in notebook blue august 2016, 66-67. Also has some comments DONE: DONE bring these comments and ideas into paper

## Thu Oct 27 08:27:50 BST 2016
- how to do ioslides -- take the main text; cut out everything except headings, quotes, images and r chunks; use --------- to make slide breaks ... 
- added new patel screenshots to figure;
- fixing plots to be a bit more meaningful
- DONE: check whether Miller's work on shopping would be helpful here ... a good point of comparison. 
- wrote lots of notes on the 2nd paper copy: DONE: bring them into the lists_conversion_experience.rmd. Also check the 1st printout as it has some unprocessed notes and changes. 
- DONE: look at SI of E&P D on lists; also Stark and Esposito; also  maybe Eco; 
- MAYBE: would be good to  add a bit of the formalism of Agrawal to flesh out the rules
- DONE: add Popper point about conditional probability
- TODO: flesh out infrastructure section by reference to the ways in which the list is constructed around some infrastructural underpinnings  
- this fits with Carolins point that many big data practices -- not all -- end up producing a ranked list. Could also use Jack Goody on lists and the seduction to order

## Fri Oct 28 08:07:00 BST 2016

- questions from presentation and further discussion thoughts:
    - doesn't the idea of the open need qualification? Yes, good idea -- would be good to develop some different ways of thinking about how to qualify the open;
    - What are the politics of doing this with the Tesco shopping list? Not sure how to answer this -- certainly doesn't have the politico-epistemically loaded value of elite examples ....
    - How does this relate to the future? Is it about open future? Maybe it is -- what if these systems do shape what happens more and more? If we live in the anticipatory mode? 
    - Where is trust in this setup? How do I trust over time? Again, couldn't answer this - I haven't look at how people relate to it. But wider issues of trust -- people going to supermarkets where the data science is not so obviously shaping everything; where the branding is not so extensively shaped by the supply chain logistics, etc. 
    - Valuation practices -- question from Martin -- about how these fit. Reminded me that I could add Marx on this ... on the forms of relevance embedded in the commodity form.
    - Also looked up precision and recall again -- TODO: both of them are connected by the notion of relevance; look at Whitehead on this -- where does he talk about relevance? The flipside of this would be an irrelevant list, but possibly there would be a relevance that felt different to the ones described by precision and recall. 
    - What about lists of images on the Tesco -- lists of images are shown ... 
    - also could bring in grocerylists.org for many shopping lists
    - think of the list as a sub-structured by a much deeper set of plans and orientations. But also as the surplus -- between the two, what happens?
- notes on the workshop august blue notebook pp 66-81 add more ...

## Thu 24 Nov 16:24:08 GMT 2016

- doing prep for socl940 session: 2 things came up. Striphas and Hallinan's Netflix prize article is on recommendation systems. DONE: include their comments on culture there; also suggests need to do a better lit search DONE

## Wed 01 Feb 2017 10:24:50 GMT
- new branch 'bdas'; starting work on journal submission; have manuscript from siegen
- DONE: add reference to morris
- DONE: read manuscript again looking for ensemble focus
- DONE: read the notes above
- DONE: read the notes in blue notebook
- trying to calculate how many hours I need to revise the manuscript -- see blue_aug2016 notebook 152-153
- starting reading the siegen manuscript and writing a few notes on it. 
- reframe the list as important topic for bdas -- use peters argument for media theory notes/durham-peters-cloud.md - data comes first, and in a list; 
- spent 25 minutes fixing knit elements of siegen version in order to get a clean manuscript. 
- TODO: add the key point that big data is a response to the problem of logistical ensembles; this might be the linchpin for the whole paper. And definitely would suit it to BDAS
- printed clean copy
- NEXT: edit 6 pages

## Thu 02 Feb 2017 09:27:56 GMT
- according to my calculations - blue_aug2016 152, editing and submission should take about 27 hours of work. I can do 2 hours today. Implies submission on 17 Feb
- DONE: look at recent articles in bdas and collect any relevant ones; have been through 100 or so -- best candidates -- Kitchin,  Beer,  added some to lists subcollection in data_forms collection in zotero
- TODO: look again at Amoore in Econ&Society, add Peters on List; could add List as evil media;
- Going through notes from the siegen event -- reminded me of
    - engagement with probability and propensity; a key site of debate
    - transformation of interestedness of order
    - alignments between our interests and interestedness of the list
    - mixed list types in the same list?
    - look at specific quantifications within lists -- often linked
    - affective becomings in list, especially via gaps
    - really rich set of list ideas in the workshop
- spent quite a lot of time on bdas site -- around 134 articles there - I tried to download all of the biblio into zotero, but that wasn't working. I'd like to read of them.- spent quite a lot of time on bdas site -- around 134 articles there - I tried to download all of the biblio into zotero, but that wasn't working. I'd like to read of them. 
- oh, no, it was working, just in a different place. so spent 20 minutes getting rid of duplicates
- starting to work on the clean paper copy and want to address
    1. probability and propensities in the world -- the Hansen argument
    2. recommender system as construction of propensities
    3. but also as response to logistical worlds and complexity of transformations of value -- 'conversion experiences'
    4. the conversion experience associated with 'big data' is? conversion of rules to associations; conversion of propensity to actuality; conversion of belief in big data into the operations of an ensemble (logistics, media platform, operational control). 
- managed to do 5 pages today, but then had all the pfaff described above as well. 22.5 hours work to go

## Fri 03 Feb 2017 13:08:17 GMT
- working on next 5-6 pages of manuscript; did to the end of page 12; so that's on schedule  2.15 hours
- TODO: look up quote on reconstruction from Dewey; can't find it electronically; need to find book at home;

## Mon 06 Feb 2017 11:20:42 GMT
- limited time today, but thinking that 'propensity' should go in the title? Would I have to change a lot for that? Or do I need to speak much more about conversion? Or combine them: from propensity to actuality: conversion experiences in shopping lists ; the propensity discourse in big data is quite important since it underpins the promise. The promise of big data is a propensity promise. So to analyse the doing of propensity is to analyse the conversion experience ...
- TODO: add the practice of propensity idea into the abstract, intro and conclusion framings. 
- did 8 pages of revisions today, leaving about 9 for tomorrow.

## Tue 07 Feb 2017 08:58:03 GMT
- hoping to do about 9 pages today, but we'll see. Some harder conceptual stuff around propensity to work through. Have read Hansen a bit again, and he is not so bad on open/closed.
- pfaffed a bit with branches and commits between master/siegen version and newer bdas version. Hard to see sometimes in git how things are distributed. I used:
    git diff bdas  master -- lists_conversion_experience.rmd
    to look at differences between the branches. 
- did around 7 pages -now finishing page 26, quite close to conclusion given the presence of a table. But I've hit a slight snag -- the design matrix, the similiarity matrix for products, has to be included in the overall process. How

## Wed 08 Feb 2017 09:32:30 GMT
-working on the conclusion on paper
-TODO: I was looking at DunnHumby yesterday. Could make more of what they claim about themselves. At least in a footnote. Also state an empirical limitation in the project -- commercial sensitivity -- and how I deal with it; could link this back to the forensic book on shelves at home
-TODO also thinking about diapers and beer again. Could I make that into the headline story? An anchor point that I set out to reconstruct? But hasn't that been done? Should look up diaper on bdas.
- finished paper revisions at 10.10. Still need do abstract; and many todos above on references, of various kinds. Especially Morris, Peters, etc.  
- starting file revisions. 
- DONE: get Miller - Theory of Shopping, TJRC (M); actually I got _Consumption and its Consequences_, 2012; seems just as good
- added a backup file; did almost 5 pages of revs in the file.

## Thu 09 Feb 2017 10:07:25 GMT
-limited time today due to various home stuff, but trying to get 5 pages of edits into text.  
- did 2 hours of edits -- gradually adding more references, etc

## Fri 10 Feb 2017 13:33:59 GMT
- read a bit of dewey last night --  
    > Concrete things have _ways_ of acting, as many ways of acting as they points of interaction with other things. One thing is callous, unresponsive, inert in the presence of some other things; it is alert, eager and on the aggressive with respect to other things; in a third case, it is receptive, docile. 153

    >Organization is no more merely nominal or mental in any art, including the art of inquiry, than it is in a department store or railway system. 154

    >[a classification] ... is rather a repertory of weapons for attack upon the future and the unknown. 154

- these might be useful -- classification, organization, the future -- but only within the broader framing of a non-transcendental ontology. 
- riding to work --  thought that I should emphasize the shift from rules to models, from associations to prediction; this is the major shift, and its use of probability as the connective tissue, as the abstract-concrete conditioning is really important to recognise. It cuts across the closed-open distinction that Hansen assumes; that said, the obscure of this conversion, the fact that it cannot be fully brought to light in reconstruction cannot be eliminated by phenomenological reduction; forensics, etc. It operates within the model itself, with the 'need for good features.'  
- TODO: footnote how DunnHumby itself markets to customers. There is a regress, perhaps not infinite, but multi-stepped that runs from customer analytics; selling customer analytics; the images of analytics; the selling of images of analytics; etc. 
- did manage to get 5 more pages

## Mon 13 Feb 2017 09:21:53 GMT
- still roughly on schedule I think; reached edit 24 of 44
- read Will Davies on happiness industry on the weekend -- he has a lot on the problems of measuring people, including measuring faces, reactions, etc. + the rise of marketing and behavioural prediction. This is the side of things that I don't really talk about. That designing more effective recommendation really does not address an underlying malaise in which not speaking, not articulating, not contesting or engaging with the question of what you want, or what might make you happy, becomes the problem in itself. The better the recommender, the worse for the recommendee. 
- reached top of page 22, edit 34 done;  which is just 5 pages. But they were hard work. Still not sure if they are doing enough. The main thing would be the identification of person and things, and the list as the site of their encounter.  How does this differ from advertising?

## Tue 14 Feb 2017 09:52:01 GMT
- late start due to school/run ; been thinking about modelisation, and how/why that matters. It is both limiting -- since its all still driven by customer behaviour -- and interesting because it identifies people and things. If probability is propensity of things, as Popper puts it, that is, all probability is conditional, and big data is a massive expansion of that conditionality, then what about conversion? There is no conversion, only an incremental indexing and linking up? 
- heavy weather in these revisions -- not sure this part of the paper is working well enough. 
-made it to edit 42 of about 46 or page 28 of 31

## Wed 15 Feb 2017 09:36:21 GMT
- hoping to finish the paper revisions today.
- work up last night thinking that I should include Hallinan and Striphas' argument about re-making culture; am I arguing the same thing? That social order is being re-done? Yes, but with a difference -- the attention to conversion events suggests that this is still being worked out; this is a social ordering process; the entanglement of human and non-human, the intersections of infrastructure, calculation, etc.  
- DONE: quote from hallinan about objects, etc
- for conclusion, need to write the different senses of conversion; and perhaps add something about relation between inversion and conversion. Is there one? Method of infrastructural inversion; no real relation because one is a methodological maxim and the latter is a term describing something that takes place.  
- writing a conclusion today -- nothing else really
- have started putting some extra stuff - hallinan, more hansen, etc near the introduction. Feels like it needs another run through to fix many loose ends, etc. 
- started to deGoede from e&P ~/archive/ensembles/deGoede_SI_lists_2016.pdf -- could add a bit more. 
- started to go through TODO lists -- still more to follow up there.

## Thu 16 Feb 2017 09:36:46 GMT
- TODO will start by adding deGoede today - have made some ~/notes/deGoede_lists_2016.md and should include some of those in main text 
- have made some notes on deGoede -- not sure they open up any new insights. Should read Staheli's piece. But they are useful support. TODO: include in conclusion
- looking at todos -- Popper is important; started reading that again and making ~/notes/popper_propensities.md
- have added some Popper into intro and conclusion (too implicit?). Feels like I need to go through again and check the argument: which is now implicitly: conversion comes from propensities; .
- TODO: check the overall argument by going through introduction, abstract and conclusion
- DONE: check section headings map the overall argument

## Fri 17 Feb 2017 13:24:29 GMT
- got Miller on Consumption; peanut-butter theory of shopping may be useful in framing what is going on with the shopping list; puts some notes in ~/notes/miller_consumptions_2012.md; two connections to social order -- the discrepancies between normative and actual; the sacrificial constitution of social order; 
- DONE: emphasize in conclusion that the 'conversion' in Tesco is an example of ongoing modification; of a propensity that existed as a possibility for quite a long time. So the conversion event is just as much about Tesco itself as about what happens to people. Always this double face of the event. 
- quite a lot of pfaffing around with images and other stuff; like labels etcs 
- TODO: commentary on the different items show in the items table and what they might be for an individual shopper;

## Thu 16 Mar 2017 10:39:43 GMT
- have done another edit on the paper, smoothing out the argument more.  
- TODO: go through the bdas bibliography to see what should be included from there. started doing this and found dourish and garnett;  
- starting to put paper edits in. How many in 30 minutes?
- roughly 80 edits to do. 
- did 20 edits in about 90 minutes of work. With 80 edits, that means about 3 more sessions of work. Should be able to submit next week.

## Fri 17 Mar 2017 14:22:25 GMT
- aiming for 20 edits today: DONE 
- finally got some turow -- ~/archive/ensembles/Turow_data_mining_2015.pdf bit is pretty relevant -- focused on shopping ... but how would I relate to it? Don't disagree with it, but doesn't offer much of a way forward. Maybe could compare to Hansen?

## Mon 20 Mar 2017 12:04:19 GMT
- have got only 2 hours today maximum - working on 20 edits I hope 
- not sure what to do about Turow -- read a bit and added notes in ~/notes/turow_2015.md Could use some of this in a footnote, or to distinguish what I'm doing in the article ... 
- do a bdas search on personalization/personalisation

## Tue 21 Mar 2017 08:57:50 GMT
- had the idea while running that the problem of the Jt argument about personalization is that (1) it sees personalization in purely personal terms; rather than as something that has a history and forms of practice (2) it sees the fabric of social order as not already woven by the power differentials associated with ownership and infrastructure. I could develop an argument around this, but not sure it would fit with this article. It seems too that Cathy O'Neill's stuff would be saying something similar.     
- the key argument -- propensities are enclosed as probabilities only to the extent that probabilities become propensities -- needs more highlighting. Get it in the abstract? TODO: put in the conclusion
- finished entering 92 edits
- current title: **Probabilities and propensities: conversion experiences in online grocery shopping**

## Wed 22 Mar 2017 09:15:08 GMT
- trying to get figure numbers working and made a new branch to do this. 
- problems running arules -- updating packages to deal with that. 
- found the root of all problems with figure numbering -- the lack of captions. Without a caption, knitr doesn't find the label. You must have captions on everything -- tables and figure -- for the cross-references to work.

## Fri 24 Mar 2017 12:16:48 GMT
- been meaning to submit -- looking at guidelines https://us.sagepub.com/en-us/nam/journal/big-data-society#ManuscriptPrep

## Tue 28 Mar 2017 11:28:42 BST
- rewrote abstract a bit, but played a lot with yaml, etc in trying to get title page to work. Ended up constructing one using title.rmd and then manual edits

## Fri 28 Apr 2017 11:33:20 BST
- accidentally came across Thrift 1999, lists come live and put into zotero ensembles folder

## Mon 10 Jul 2017 15:50:59 BST
- need to update this log with stuff about reports;
- TODO; notes from Dewey about reconstruction are here ~/notes/dewey_reconstruction_1957.md; and the full text is in ~/archive/dewey_reconstruction.txt;
- Dewey:  grepped through the full text; the basic idea is that reconstruction is the way that experience can be intelligently re-configured; it draws on the  medium it inhabits, but changes/transforms elements of that medium, in the same way that any living thing does. The challenge of reconstruction is to deal with problem of the disconnect between power-knowledge and social life.  
- if I were to rewrite the lists paper as a philosophical reconstruction, how would it differ from an empirical investigation of recommender systems.
- dewey's argument is that a new conception of experience is needed, and a new relation between reason and experience (83). Psychology + biology: transformative nature of experience -- it changes the world; 'experience becomes an affair primarily of doing'  - this in the notes file ...

## Wed 12 Jul 2017 14:29:19 BST
- historical precedent -- Tesco shopping 1984 http://www.bbc.co.uk/news/magazine-24091393 [@Winterman_2013]  - TODO -- add in 
- reading through the reports, and summarizing here what they are asking
    1. The friendly one  - lack of cohesion and clarity in the argument; re-frame opening to be less about commodities and supply chain, and more about problems of association, correlation, categorisation, etc.; establish a consistent role for the list; elaborate the methodology of reconstruction; clearly say what kind of transformation occurs in the transformation, and in the propensities; and what is the 'so what' of these clarifications and conceptualisation? 
    2. Asks for more substantial empirical engagement, reorganization of the piece and clarification of overall argument and purpose, and specific section purposes; many detailed points and calls for clarification; 
    3. Asks for clarification about the contribution of the piece; description of the case is vague in detail; look at Bucher and Barocas for state of the art; deal with disconnect between abstract and actual paper, especially around social order; main argument is incomprehensible.  
- started writing reply_to_reviews.md
- there are 36 double-spaced pages of content (39 including references). About 6 of these are figures/tables/images. So at 6 pages/day revisions, it would take about 5 days. Starting tomorrow 13 July, that would be 13, 14, 17, 18, 19 July working on the pages.  Entering the edits would take another 5 days (20, 21, 24, 25, 26). That assumes not major obstacles. So submission could be before end of July. 
- I'm not I can satisfy R2 in their desires for empirical, or sociological rigour. I'm drawn much more to the idea of reconstructing an encounter with a system, where reconstruction is a pragmatist/Dewey-oriented attempt to change our relation to experience. This includes a form of technical practice working with data and models, very much in the way of the actual recommender system.

### 2017-07-13 log.rmd

@@
 
+## Wed 05 Jul 2017 11:03:59 BST
+- received reports a week or two ago: all in report.md
+- printed them out to read; nothing too major, but they didn't get the paper. Revision will need to cut away whatever obstructs their view, I guess. 
+- programme the work on this: I've estimated 15 hours on it. Might work;
+- DONE: read reports properly
+- added reports as report.md.gpg
+
 ## Mon 10 Jul 2017 15:50:59 BST

### 2017-07-13 log.rmd
writing new intro on revision

@@
 This includes a form of technical practice working with data and models, very much in the way of the actual recommender system.
+
+## Thu 13 Jul 2017 15:29:02 BST
+- starting revisions by going through ref2; 
+- philosophico-methodological preramble:
+
+> This paper is a philosophical reconstruction [@Dewey_2017] of an online grocery ordering system.
+The reconstruction is empirical in certain respects.
+It departs from an ethnographic moment: being a member of the audience at an industry/academic conference where a data scientist was describing the system.
+It is experimental in reconstructing, using widely available code resources resources such as API (Application Programmer Interfaces) and software libraries for machine learning, some important elements of the system in question.
+It is archaeological in the sense developed by Michel Foucault in that it seeks to identify and investigate the practices of knowing, of operating, ordering and enunciating associated with the system as it takes shape over time.
+Like some recent work in science and technology studies and anthropology [@Marres_2017; @Rabinow_2004], the paper itself provisionally enacts a mode of relating to the problem it addresses.
+At times, this mode of empirical philosophical engagement with its objects may well seem overly concerned with the technical minutiae of the grocery ordering system and its recommendation, and lacking in empirical sociological rigour.
+With some forbearance on the part of readers, however, the main philosophical argument of the paper, should resonate sociologically.
+It concerns social order, or the consistency and regularity of relating and acting amid a world in where probabilities and calculations based on probabilities play a more pervasive role.

### 2017-07-14 log.rmd
writing main argument for it

@@
 - starting revisions by going through ref2; 
-- philosophico-methodological preramble:
+- philosophico-methodological preramble: TODO: put this as opening paragraph
 
@@
 The reconstruction is empirical in certain respects.
-It departs from an ethnographic moment: being a member of the audience at an industry/academic conference where a data scientist was describing the system.
+It departs from an ethnographic moment: being a member of the audience at an industry/academic conference where a data scientist was describing how recommendations were created for online grocery orders.
+It draws on anthropological and sociological research concerning shopping and lists.
 It is experimental in reconstructing, using widely available code resources resources such as API (Application Programmer Interfaces) and software libraries for machine learning, some important elements of the system in question.
 It is archaeological in the sense developed by Michel Foucault in that it seeks to identify and investigate the practices of knowing, of operating, ordering and enunciating associated with the system as it takes shape over time.
-Like some recent work in science and technology studies and anthropology [@Marres_2017; @Rabinow_2004], the paper itself provisionally enacts a mode of relating to the problem it addresses.
-At times, this mode of empirical philosophical engagement with its objects may well seem overly concerned with the technical minutiae of the grocery ordering system and its recommendation, and lacking in empirical sociological rigour.
+Not least, draws on philosophy of science and media theory to frame its engagement with data, algorithms and predictions.
+Like some recent work in science and technology studies and anthropology [@Marres_2017; @Rabinow_2004], through this combination of approaches the paper itself provisionally allows its own reconstruction to be shaped by  the object it describes.
+At times, this mode of empirical philosophical engagement with its objects may well seem overly concerned with the technical minutiae of the grocery ordering system and its recommendations, and lacking in empirical sociological rigour.
 With some forbearance on the part of readers, however, the main philosophical argument of the paper, should resonate sociologically.
 It concerns social order, or the consistency and regularity of relating and acting amid a world in where probabilities and calculations based on probabilities play a more pervasive role.
+
+## Fri 14 Jul 2017 11:30:07 BST
+- read popper on propensities again, updated notes in ~/notes/popper_propensities.md, and also marked up the paper copy more;  
+- TODO: for methods add sandvig on audit [@Sandvig_2015a]
+
+> The main argument of the paper concerns the operational role of probability in Big Data situations.
+While calculations of probability have  long-standing importance in many settings (insurance, medicine, experimental and field sciences, operations research, risk analysis, engineering design, public health,  economic modelling, etc.), probabilities have taken on a much thicker situational texture.
+Operational probabilities differ from epistemic probabilities in their position in the world, as part of infrastructures, transactions, and event-structures.
+Taking their complex siting into account, and accepting their _physical_ reality, we might come to understand the world they belong to as what the philosopher of science Karl Popper understood as a 'world of propensities, as an unfolding process of realizing possibilities and of unfolding new possibilities' [@Popper_1990,19].
+
+> The broad implication of the argument concerns how we think about social order or regularities: proponents and critics of Big Data present the advent of large volumes of data and analytical techniques ( machine learning in particular) as transforming organisations, institutions, economic processes and social life.
+I suggest too that a transformation or conversion event is occurring, but one which might open up the possibility of experiencing the world differently, and conceiving of social ordering processes somewhat afresh.
+Even if the argument is based on an account of an eminently social practice such as shopping, the analysis of how a world of propensities continuously converting into realizations (or not) applies across the board, from the most physical to the most ideal or abstract settings.
+The banality of the example here -- grocery shopping -- allows some first hand engagement with the messiness, the entanglements, and potentials of this conversion event in ways that other interesting settings -- particle physics for instance -- do not.

### 2017-07-17 lists_conversion_experience.rmd
starting reconstruction

--- 2017-03-28 lists_conversion_experience.rmd
@@
 It can no longer rely on placement of goods in carefully chosen locations in stores.
-It needs or at least might want to continuously up-sell and cross-sell to customers who might hardly ever visits the supermarket itself.
+It needs or at least might want to continuously up-sell and cross-sell to customers who might hardly ever visit the supermarket itself.
 
-How could we characterise the shift from ClubCard data mining to online grocery markets?
-If 'Tesco is the clear winner in the online grocery market, in fact it takes almost 50p of every £1 spent on food shopping on the internet' [How SEO helps Tesco to dominate the online grocery market](https://econsultancy.com/blog/64841-how-seo-helps-tesco-to-dominate-the-online-grocery-market)[@Silverwood-Cope_2014], then Tesco itself has undergone some kind of conversion?
+How could we characterise the shift from _ClubCard_ data mining to online grocery markets?
+If 'Tesco is the clear winner in the online grocery market, in fact it takes almost 50p of every £1 spent on food shopping on the internet' [@Silverwood-Cope_2014](https://econsultancy.com/blog/64841-how-seo-helps-tesco-to-dominate-the-online-grocery-market), then Tesco itself has undergone some kind of conversion?
 Patel described her work at DunnHumby as shifting the recommender system from a 'rules-based list' to a 'relevancy model.' The relevancy model changed how the 'Have You Forgotten' list was constructed.

### 2017-07-17 log.rmd
starting reconstruction

--- 2017-07-13 log.rmd
@@
 This includes a form of technical practice working with data and models, very much in the way of the actual recommender system.
+
+## Thu 13 Jul 2017 15:29:02 BST
+- starting revisions by going through ref2; 
+- philosophico-methodological preramble: TODO: put this as opening paragraph
+
+> This paper is a philosophical reconstruction [@Dewey_2017] of an online grocery ordering system.
+The reconstruction is empirical in certain respects.
+It departs from an ethnographic moment: being a member of the audience at an industry/academic conference where a data scientist was describing how recommendations were created for online grocery orders.
+Given this highly mediated and limited encounter with a recommender system and one of its designers, the paper constructs a multi-layered account.
+It draws on anthropological and sociological research concerning shopping and list-making to situate the recommender system amongst the social ordering practices of shopping.
+It conducts several code-based experiments in order to reconstruct, using widely available code resources resources such as API (Application Programmer Interfaces) and software libraries for machine learning, some important elements of the system in question.
+It makes use of archaeological approaches, in the sense developed by Michel Foucault [@Foucault_1972] in order to identify and map the functional statements and practices underpinning the knowledge and order generated by such systems over time.
+Not least, I draw on the philosophy of science and media theory to frame an engagement with data, algorithms and recommendations around probabilities and their enactments.
+Like some recent work in science and technology studies,  anthropology and media studies [@Marres_2017; @Rabinow_2004; @Bogost_2012], through this combination of approaches,  the paper itself provisionally allows its own reconstruction to be shaped by an experiential encounter with the object it describes.
+At times, this mode of empirical philosophical engagement with its objects may well seem overly concerned with the technical minutiae of the grocery ordering system, and lacking in  sociological rigour or empirical evidence.
+With some forbearance on the part of readers, however, the main philosophical argument of the paper should resonate sociologically.
+It concerns social order, or the consistency and regularity of relating and acting amid a world in where probabilities and calculations based on probabilities play a more pervasive role.
+
+## Fri 14 Jul 2017 11:30:07 BST
+- read popper on propensities again, updated notes in ~/notes/popper_propensities.md, and also marked up the paper copy more;  
+- TODO: for methods add sandvig on audit [@Sandvig_2015a]
+
+> The main argument of the paper concerns how  probability operates in Big Data situations.
+While calculations of probability have  long-standing importance in many settings (insurance, medicine, experimental and field sciences, operations research, risk analysis, engineering design, public health,  economic modelling, etc.), probabilities have taken on a much thicker situational texture.
+Operational probabilities differ from epistemic probabilities in their position in the world, as part of infrastructures, transactions, and event-structures.
+Taking their complex siting into account, and accepting their _physical_ reality, we might come to understand the world they belong to as what the philosopher of science Karl Popper understood as a 'world of propensities, as an unfolding process of realizing possibilities and of unfolding new possibilities' [@Popper_1990,19].
+Given this understanding of operational probability, our understanding of what is enacted in contemporary data analytics, especially in association with platforms [@Gillespie_2010] might shift.
+
+> The broad implication of the argument concerns how we think about social order or regularities: proponents and critics of Big Data present the advent of large volumes of data and analytical techniques (machine learning in particular) as transforming organisations, institutions, economic processes and social life.
+I suggest too that a transformation or conversion event is occurring, but one which might open up the possibility of experiencing the world differently, and conceiving of social ordering processes somewhat afresh.
+Even if the argument is centred on an eminently social practice such as shopping, the analysis of how a world of propensities continuously and unevenly converts into realizations applies across the board, from the most physical to the most ideal or abstract settings.
+The average everydayness of the example here -- grocery shopping -- allows some first hand engagement with the messiness, the entanglements, and potentials of this conversion event in ways that other interesting settings -- particle physics for instance -- do not.
+
+## Mon 17 Jul 2017 11:46:32 BST
+- still not sure how radically to restructure this. 
+- the simplest order I can think of
+    - A. Introduction - includes a,b,c from above
+    - B. Patel section + opening description
+    - C. Shopping and lists -- Miller, and list section
+    - D. Archaeology of recommendations -- from 1984 -2007
+    - E. Experimenting with probabilistic conversions
+    - F. Conclusion
+- will see if this structure works with a rough cut and paste
+- re-ordering has gone ok to the end of D. But section E is really tricky -- this is the conversion event; do I include the machine learning stuff as part of D; or use it to be begin E?
+    - E: shift from apriori to logistic regression; then think of good features; experimentalising/trials; infrastructural operationalisation; but then what about substitution; complementarity?
+All of this section is meant to be about probabilisation?
+And the propensities idea;
+- TODO: write short paragraph describing my overall social theory of lists (approx 200 words)
+- new title: Probabilities in the data analytic conversion of social order
+- TODO: play with the structuring of Section E. 
+-

### 2017-07-17 reconstruction.rmd
starting reconstruction

--- 2017-03-28 lists_conversion_experience.rmd
@@
----
-title: "Probabilities and propensities: conversion experiences in online grocery shopping"
-author: Adrian Mackenzie^[University of Somewhere] 
-date: "`r format(Sys.time(), '%B %Y')`"
-fontsize: 12pt
-citation_package: biber
-latex_engine: xelatex
-bibliography: /home/mackenza/ref_bibs/uni.bib
-csl: "sage-harvard.csl"
-link-citations: yes
-header-includes:
-    - \usepackage{setspace}
-    - \doublespacing
-abstract: 
-    "Recommender systems attempt to transform or convert the temporal and agential fabric of social ordering by substituting probabilistic calculation for social practices.
-This paper investigates practices of ordering in the setting of shopping lists and  online grocery orders.
-It reconstructs an account of how a recommender system suggests a small number of grocery items of personal relevance to each of the millions of online grocery shoppers in  a major UK supermarket chain.
-Although all shopping lists mix past and future actions, actual and potential events, and normative and actual values, their re-writing by recommender systems adds a new operational dynamic.The paper explores how in the context of recommender systems the constitutive incompleteness of shopping lists, their propensity to expand or change, becomes more important than their capacity to inscribe social order.
-This suggests that the conversion of lists from finite ordering to indeterminate propensity in recommender systems requires new ways of conceptualising social order."
-keywords: prediction, probability, order, lists, recommender system, shopping
----
+# Probabilities in the data analytic conversion of social order
 
-```{r setup2, cache=FALSE, include=FALSE}
-library(knitr)
-library(tint)
-knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tint'))
-options(htmltools.dir.version = FALSE)
-output <- opts_knit$get("rmarkdown.pandoc.to")
-opts_chunk$set(warning=FALSE)
-opts_chunk$set(message=FALSE)
-opts_chunk$set(echo=FALSE)
-opts_chunk$set(error=FALSE)
-opts_chunk$set(cache=TRUE)
-```
+# A. Introduction  includes a,b,c from above
 
-# Probabilities and propensities: conversion experiences in online grocery shopping
+This paper is a philosophical reconstruction [@Dewey_2017] of an online grocery ordering system.
+The reconstruction is empirical in certain respects.
+It departs from an ethnographic moment: being a member of the audience at an industry/academic conference where a data scientist was describing how recommendations were created for online grocery orders.
+Given this highly mediated and limited encounter with a recommender system and one of its designers, the paper constructs a multi-layered account.
+It draws on anthropological and sociological research concerning shopping and list-making to situate the recommender system amongst the social ordering practices of shopping.
+It conducts several code-based experiments in order to reconstruct, using widely available code resources resources such as API (Application Programmer Interfaces) and software libraries for machine learning, some important elements of the system in question.
+It makes use of archaeological approaches, in the sense developed by Michel Foucault [@Foucault_1972] in order to identify and map the functional statements and practices underpinning the knowledge and order generated by such systems over time.
+Not least, I draw on the philosophy of science and media theory to frame an engagement with data, algorithms and recommendations around probabilities and their enactments.
+Like some recent work in science and technology studies,  anthropology and media studies [@Marres_2017; @Rabinow_2004; @Bogost_2012], through this combination of approaches,  the paper itself provisionally allows its own reconstruction to be shaped by an experiential encounter with the object it describes.
+At times, this mode of empirical philosophical engagement with its objects may well seem overly concerned with the technical minutiae of the grocery ordering system, and lacking in  sociological rigour or empirical evidence.
+With some forbearance on the part of readers, however, the main philosophical argument of the paper should resonate sociologically.
+It concerns social order, or the consistency and regularity of relating and acting amid a world in where probabilities and calculations based on probabilities play a more pervasive role.
 
-Online grocery shopping at Tesco, a large supermarket chain based in the UK, includes recommendations for further grocery purchases under the title of 'Have you forgotten?'.
+The main argument of the paper concerns how  probability operates in Big Data situations.
+While calculations of probability have  long-standing importance in many settings (insurance, medicine, experimental and field sciences, operations research, risk analysis, engineering design, public health,  economic modelling, etc.), probabilities have taken on a much thicker situational texture.
+Operational probabilities differ from epistemic probabilities in their position in the world, as part of infrastructures, transactions, and event-structures.
+Taking their complex siting into account, and accepting their _physical_ reality, we might come to understand the world they belong to as what the philosopher of science Karl Popper understood as a 'world of propensities, as an unfolding process of realizing possibilities and of unfolding new possibilities' [@Popper_1990,19].
+Given this understanding of operational probability, our understanding of what is enacted in contemporary data analytics, especially in association with platforms [@Gillespie_2010] might shift.
+
+The broad implication of the argument concerns how we think about social order or regularities: proponents and critics of Big Data present the advent of large volumes of data and analytical techniques (machine learning in particular) as transforming organisations, institutions, economic processes and social life.
+I suggest too that a transformation or conversion event is occurring, but one which might open up the possibility of experiencing the world differently, and conceiving of social ordering processes somewhat afresh.
+Even if the argument is centred on an eminently social practice such as shopping, the analysis of how a world of propensities continuously and unevenly converts into realizations applies across the board, from the most physical to the most ideal or abstract settings.
+The average everydayness of the example here -- grocery shopping -- allows some first hand engagement with the messiness, the entanglements, and potentials of this conversion event in ways that other interesting settings -- particle physics for instance -- do not.
+
+# B. Patel section + opening description
+
+At one of the many industry-meets-academia events occurring in increasingly data science-oriented higher education institutions in the UK, speakers from industry, government and commerce described their work with predictive models.[^11.1] Shreena Patel, a PhD graduate in statistics and operations research,  works as a data scientist for DunnHumby, a well-known [ customer science company ](https://www.dunnhumby.com/) [@dunnhumby_2017].
+Her work at DunnHumby focuses on online grocery shopping at the supermarket chain Tesco.
+Speaking to an audience of statisticians and operations researchers, Patel focused on the development and operation of predictive models underlying  shopping list recommendations.
+Her presentation was filled with graphs, numbers, and tables concerning ongoing development of the 'Have you forgotten?' recommender system.
+
+[^11.1]: More than 100  Data Science Institutes have been set up in North America, Europe and UK since around 2012.
+The traffic between higher education and data analytics in industry is intense and flows in several forms: people funding, research findings, software and technical devices (code) and training.
+
+Against the background of the sheer number of commodities and their distribution of prices,  Patel's presentation presents two opportunities.
+First, by recounting, contextualising and commenting on the main steps in making the 'Have your forgotten?' list, we might follow some of the predictive sense-making done by data scientists and customer analytics teams working with transactional data in a typical commercial setting.
+Patel mentioned many of these steps only fleetingly in the presentation, for they are largely taken for granted as part of predictive analytic practice.
+Second, Patel focused on the renovation and updating of long-standing data-mining practice via a much more explicitly 'big data' and 'machine learning' oriented implementation.
+Her presentation concerns what we might term, for reasons that will gradually emerge, a 'conversion event' in which a long-standing recommender system was replaced by a new, explicitly 'big data'-style system concerned with 'personally relevant' recommendations.
+What stands out from the presentation is the normality of the recommender system: it is part of a long-standing and ongoing transformation or conversion of grocery shopping.
+From a sociological standpoint, the interest lies less in specific technical innovations and more on how the ordering work done by recommender systems relates to problems of social order more generally.
+
+Online grocery shopping at [Tesco](https://www.tesco.com/groceries/?icid=dchp_groceriesshopgroceries), the largest supermarket chain in the UK, includes recommendations for further grocery purchases under the title of 'Have you forgotten?'.
 When Tesco customers shop for groceries online, a list of five recommendations appear at the checkout stage.
-The recommendations are the product of a recommender system, an important category of operational device in big data (see for instance [@Striphas_2015] for analysis of Netflix recommendation; [@Morris_2015]  or [@Seaver_2015] for an account of music recommendations]).
+The recommendations are the product of a recommender system, an important category of operational device in big data (see for instance [@Striphas_2015] for analysis of Netflix recommendation; [@Morris_2015]  or [@Seaver_2015] for an account of music recommendations).
 The question 'have you forgotten?' is followed by a list of some  grocery items that could have been or are usually on a shopping list.
@@
 Encounters between this hyper-complex commodity-geography and people, even in the familiar confines of a supermarket, are no simple matter, either for shoppers or for supermarket operators such as Tesco.
-For shoppers, as Daniel Miller has argued [@Miller_2012],  shopping takes place as a negotiation of discrepancies between normative and actual social order (for instance, between ideals of health and commitments to thrift).
-It can also be seen  as a sacrificial ritual that re-instates social order in various ways.
 
-The shopping list, whether written on the back of an envelope, or saved as a list in an online grocery shopping system, lies at the intersection of these logistical flows, infrastructural orderings, cultural values and  lively negotiations.
-It is an intersectional ordering device that encapsulates a universe of possible references, and a teeming multitude of propensities with an actual local order (in both senses of that term: an ordering and an imperative).
+# C. Shopping  lists and  a theory of social ordering
+
+The anthropologist  Daniel Miller has argued [@Miller_2012] that  shopping takes place as a negotiation of discrepancies between normative and actual social order (for instance, between ideals of health and commitments to thrift).
+What Miller suggests here is that ...
+[TBA]  It can also be seen  as a sacrificial ritual that re-instates social order in various ways.
+
+From an ethnomethodological standpoint, shopping lists provide important clues to how local social order is constituted, maintained, and repaired.
+Social ordering is done in shopping, and the shopping list as an inscriptive device stabilises certain forms of ordering at the expense of other possible orderings.
+
+The shopping list, whether written on the back of an envelope, or saved as a list in an online grocery shopping system, lies at the intersection of logistical flows, infrastructural orderings, and  lively negotiations around actual and normative social orders.
+Shopping lists, therefore  are intersectional ordering devices that encapsulates a universe of possible references, and a teeming multitude of propensities with an actual local order (in both senses of that term: an ordering and an imperative).
 As people shop, either by  trawling along aisles packed with thousands of products, or scrolling down screens or searching for particular brands amidst search results, lists inscribe some kind of order that filters or reduces the excessive propensities, claims, dazzle and distraction of commodities to the practicalities of domestic economy.
@@
 
-```{r random-basket, echo=FALSE, cache=TRUE, fig.cap='Random basket'}
-library(dplyr)
-library(ggplot2)
-library(readr)
-library(knitr)
-csvs = list.files('data', pattern = '*.csv',full.names=TRUE)
-df = bind_rows(sapply(csvs, read_tsv))
-kable(sample_n(df, size=5)$name, col.names='item', caption='A random list of Tesco grocery items')
-```
-
-```{r shopping-hand, fig.cap='A hand-written shopping list'}
-include_graphics('figure/1_shopping.png')
-```
-
-The items on Tesco's 'Have You Forgotten?' list are recommendations.
-They are only a few of them: five.
-The question is: which five items and in what order?
-In terms of content, there are many similarities between shopping lists such as Figure \@ref(fig:shopping-hand)  -- a dozen lines scrawled on the back of a used envelope stuck on the refrigerator door  --  and the five ranked items shown to a customer on an online shopping site (as shown in Table \@ref(tab:random-basket)).
-Even if they are similar in content,  the list of five suggestions shown to Tesco's online customers at the checkout  exemplifies a transformation in forms of ordering.
-In its re-writing of the list, the recommender system reassembles individualised desire, consumption practices, value-chains, infrastructural time-space dynamics, the lives of others and the propensities of things.
-A major contemporary form of ordering, the recommender system can be found in many  forms of news, media, entertainment, marketing, and advertising.
-Tesco's recommender systems differs greatly from others associated with Amazon, Netflix or Ebay.
-Grocery shopping is a high-frequency, multi-purchase shopping practice and the potential for recommendations proliferates accordingly.
-
-Despite being titled 'Have You Forgotten?' Tesco's recommender system  focuses on 'conversion events.' If a shopper does buy a recommended item, 'conversion' has occurred.
-From a supermarket perspective, high-value conversion are desirable.
-There is little value in successfully recommending small or cheap items such as rubber bands or toothpicks.
-Conversion events, when they occur, confirm the power of predictive analytics to change the world in some way.
-From a sociological perspective, the significance of conversion events is far less obvious.
-In the name of conversion events, recommender systems stage an epistemic-operational transformation central to big data.
-In conversion events,  predictive analytics happens in the world.
-The tendency of things to realize certain events -- what Popper terms the 'propensities of things' [@Popper_1990, 11] -- happens  through data practices.
-Data in concert with calculation, models, infrastructures, and experiments, affords apprehension, visibility and management of associative propensities in the world.
-
-# Lists and orders
-
-In grocery shopping, conversion events derive from and affect lists.
-Whether  they write out shopping lists or not, online grocery shoppers interact with commodities through lists.
-Lists have long interested historians, anthropologists, sociologists, literary and media theorists as everyday ordering devices.
-If, as the social anthropologist Jack Goody argued, lists are historically primary as forms of writing [@Goody_1986], and if list-making and its later variants (e.g.
+If, as the social anthropologist Jack Goody argues, lists are historically primary as forms of writing in urbanizing cultures [@Goody_1986], and if list-making and its later variants (e.g.
 tables) precede discursive and narrative writing practices, then we might expect lists to function as powerful social ordering devices.
@@
 
-In this paper, I focus on lists and list-making in terms of the predictive social orderings.
-All lists are forms of order.
-Mathematicians distinguish lists from sets on the basis of order.
+Finally, and more formally, we might also consider some of the formal properties of lists as orderings.
+All lists imply order, but what kind of order?
+Mathematicians distinguish lists from sets on the basis of a distinction focused on order.
 Formally, both lists and sets are collections, but a list also contains a mapping of its elements to an ordering, usually the natural or counting numbers.
@@
 Often the order is implicit to the spatial arrangement of the list.
+This formal idea of the list as a mapping is helpful since it suggests relational possibilities.
+Mappings occasion relations between sets, collections, group or multiplicities.
 Mappings between list order and social order are manifestly highly contingent.
@@
 
-How should we view the predictive re-writing of lists?
-The media theorist Mark Hansen has recently suggested that 'predictive analytics are discoveries of micrological propensities that are not directly correlated with human understanding and affectivity and that do not by themselves cohere into clearly identifiable events.' [@Hansen_2015a, 111-112] Hansen's account resonates with wider views of big data.
-The expectation energising much predictive practice  pivots on 'micrological propensities' elicited from data.
-Unlike many other accounts of data, Hansen  links data to propensities of things in the world via probabilities:
+TBA - short paragraph summarising the social theory of list ordering
 
->Whatever explanatory and causal value predictive analytics of large datasets have is, I suggest, ultimately rooted in this ontological transformation whereby probabilities are understood to be expressions of the actual propensity of things [@Hansen_2015a, 120]
-
-In his Whitehead-inspired account, Hansen theorises prediction as an 'ontological transformation' that deploys probabilities in an evermore closely woven and encompassing expression of animated, eventful, propensities of things.
-Predictive analytics links probabilities as calculations to propensities or the mutable associative agencies of things.
-The practical and indeed empirical  question is whether such transformations or 'conversion' in relations between probabilities and 'the propensities of things'  can be detected and articulated in the prosaic setting of shopping lists and recommender systems.[^20.1]
-
-[^20.1]: I will argue that 'conversion' is a preferable term for these changes.
-It highlights  re-orientations in subjects, experience, things, numbers and structures that ranges more widely than Hansen's 'ontological transformation' formulation.
-
-# Conversion events: DunnHumby's work for Tesco
-
-At one of the many industry-meets-academia events occurring in increasingly data science-oriented higher education institutions in the UK, speakers from industry, government and commerce described their work with predictive models.[^11.1] Shreena Patel, a PhD graduate in statistics and operations research,  works as a data scientist for DunnHumby, a well-known [ customer science company ](https://www.dunnhumby.com/) [@dunnhumby_2017].
-Her work at DunnHumby focuses on online grocery shopping at the supermarket chain Tesco.
-Speaking to an audience of statisticians and operations researchers, Patel focused on the development and operation of predictive models underlying  shopping list recommendations.
-Her presentation was filled with graphs, numbers, and tables concerning ongoing development of the 'Have you forgotten?' recommender system.
-
-[^11.1]: More than 100  Data Science Institutes have been set up in North America, Europe and UK since around 2012.
-The traffic between higher education and data analytics in industry is intense and flows in several forms: people funding, research findings, software and technical devices (code) and training.
-
-```{r simulate-inventory, echo=FALSE, cache=TRUE, fig.margin=TRUE, fig.cap='Prices and quantities of things in the Tesco universe of reference'}
-library(dplyr)
-library(ggplot2)
-library(readr)
-library(knitr)
-
-csvs = list.files('data', pattern = '*.csv',full.names=TRUE)
-df = bind_rows(sapply(csvs, read_tsv))
-
-g = ggplot(df, aes(x=price,y=ContentsQuantity)) + geom_point(alpha=0.3, size=0.9 ) + theme(axis.text=element_text(size=10))  + xlab('Price (£)')
-g + ylab('Number of products') + scale_x_log10() + geom_smooth()
-```
-
-```{r expensive-groceries, echo=FALSE, cache=TRUE}
-s = sample_n(df[df$price>=30,], size=10)
-s$Price = s$price
-s$Label = s$name
-kable(s[order(s$price), c('Label','Price')], col.names=c('Item', 'Price (£)'), caption='Random sample of higher cost items from Tesco')
-```
-
-Like many recommender systems, Patel's models address a large inventory of commodities.
-Figure \@ref(fig:simulate-inventory) plots something of the density of commodity prices in the Tesco grocery inventory.
-By virtue of their sheer abundance, the plot cannot easily separate the many items that Tesco sells, especially in the price range £1-£10.
-A random sample of the more highly priced items in Tesco stores is shown in Table  \@ref(tab:expensive-groceries)
-
-Against the background of the sheer number of commodities and their distribution of prices,  Patel's presentation presents two opportunities.
-First, by recounting, contextualising and commenting on the main steps in making the 'Have your forgotten?' list, we might follow some of the predictive sense-making done by data scientists and customer analytics teams working with transactional data in a typical commercial setting.
-Patel mentioned many of these steps only fleetingly in the presentation, for they are largely taken for granted as part of predictive analytic practice.
-Second, Patel focused on the renovation and updating of long-standing data-mining practice via a much more explicitly 'big data' and 'machine learning' oriented implementation.
-What stands out from the presentation is the normality of the recommender system: it is part of a long-standing and ongoing transformation or conversion of grocery shopping.
-From a sociological standpoint, the interest lies less in specific technical innovations and more on how the ordering work done by recommender systems relates to problems of social order more generally.
-
-Since Patel was not able to release her presentation slides to a wider audience, I rely on data-driven simulations and statistical programming as ways of reconstructing elements of the recommender system.
-In order to explore the operational environment she described, and its predictive operations, I model key features of her presentation using sample grocery data drawn from the Tesco API ([Application Programmer Interface](https://devportal.tescolabs.com/docs/services/56c73b1bf205fd0ed81dbe7a/operations/56d5b2bee2813e0a0053c47c), [@Tesco.com_2017])   and a range of data science, statistics and graphics libraries implemented in the widely-used statistical programming language `R` [@RDevelopmentCoreTeam_2010].
-At a methodological level, it draws heavily on the vision of 'interface methods' proposed by Noortje Marres [@Marres_2015] in order to keep in view both the highly technical orderings of lists in recommender system and the changes in social orderings they might imply or impel.
-
-# How the shopping list was reconstituted in the 1990s
+# D. Archaeology of recommendations  from 1984  to 2007
 
@@
 The shift in models results in a more probabilistic structuring of lists.
-The ostensible banality of the 'Have you forgotten?' recommendations betrays micrological signs of a topologically complex predictive infrastructure that constructs alignments between people and things in significant ways.
+The ostensible banality of the 'Have you forgotten?' recommendations betrays micrological signs of a topologically complex predictive infrastructure that intervenes in list-based social ordering to reconfigure alignments between people and things.
 
-# Reconfiguring recommendations
-
+This is not a drastic or abrupt event.
 Academic researchers first began writing about recommender systems in the mid-1990s.
@@
 
-```{r arules-ex, cache=TRUE, fig.cap='Frequency of association betwene items in the `Grocery` dataset', message=FALSE, echo=FALSE, fig.margin=TRUE}
-library(knitr)
-library(arules)
-library(arulesViz)
-library(datasets)
-data(Groceries)
-itemFrequencyPlot(Groceries,topN=10,col=rainbow(10), ntype="absolute",  cex.axis = 0.55)
-```
+# E. Experimenting with probabilistic conversions
 
-```{r arules-ex2}
-options(digits=2)
-rules<- apriori(data=Groceries, parameter=list(supp=0.001,conf = 0.15,minlen=2), appearance = list(default="rhs",lhs="whole milk"), control = list(verbose=FALSE))
-rules_sorted <- sort(rules, decreasing=TRUE,by="confidence")
-kable(inspect(rules_sorted[1:5]), caption='The first five association rules for the `Groceries` dataset')
-```
-
-The `apriori` algorithm finds sets of items that commonly occur together in transactions.
-These commonly occurring sets are expressed as 'association rules.' For instance, using the `Groceries` dataset in the `R` package `arules,` we can readily see how Walmart *might* have discovered   the fabled association between diapers and beer.[^149.1] Figure \@ref(fig:arules-ex) shows how often common items appear.
-`Whole milk` appears most frequently.
-The association rule-based model for the `Groceries` dataset illustrates some characteristic ordering practices of 1990s-style data-mining approaches to lists (the `Groceries` dataset  was acquired from a 'local German supermarket' [@Hahsler_2006a]).
-The association between items is expressed in the form of _rules_ not a model.
-These rules, a term derived from older decision support literature, define sets of associations between items.
-Table \@ref(tab:arules-ex2) shows that the first rule has a stronger association than the second.
-Milk and vegetables is more likely an association than milk and buns.
-A rules-based recommender system should, therefore, might put them in that order in the 'Have You Forgotten' list.
-
-[^149.1]: One of the founding narratives of big data customer analytics, the beer-diaper association was not, according to [@Power_2002] found through data mining at all.
-
-Although association rules and the `apriori` algorithm present social order in terms of sets of frequency-weighted associations, an `apriori`-based recommender system contains no statistical model.
-It also no grasp on personal relevance since the associations it defines are only between things.
-Put in terms of the big data conversion, the rules-based system is not predictive, or personal.
-The `apriori` algorithm might be seen as an attempt to grasp the excess of things in a Tesco supermarket by identifying subsets whose closure derives from the associations between things that people construct (through buying habits, by virtue of a shopping list, etc.).
-In association rules, personhood or experience is irrelevant.
-The openness of these rules -- the `apriori` algorithm can product any number of rules of diminishing strength -- points towards the propensity of things to find themselves together in shopping baskets, but cannot cannot individuate that propensity beyond demographic quantiles and categories.
-
-Even if `apriori`  elicits associations between things, it struggles with the propensity of commodities to multiply.
-A simple illustration of the combinatorial problem faced by recommender system can be developed using the `Grocery` dataset.
-If we take all the items in the `Grocery` dataset and paste them into the 'shopping list' box on the Tesco grocery website (or as I side, run them as searches on the TescoLab Product API), each of the 169 items in the dataset yields dozens and sometimes thousands of products from Tesco online.
-
-```{r tesco-ex, echo=FALSE, cache=TRUE, fig.cap ='Tesco surplus',  fig.margin=TRUE}
-    library(ggplot2)
-    tesco = read.csv('data/reference_data/tesco_groceries.csv')
-    total = sum(tesco$actual, na.rm=TRUE) 
-    tesco = na.omit(tesco)
-    tesco$actual = as.integer(tesco$actual)
-    ggplot(tesco[tesco$actual>50,], aes(x=labels, y=actual, fill=labels)) +geom_bar(stat='identity') + coord_flip() + ggtitle('Tesco grocery items with more than 50 products') + theme(legend.position="none", axis.text=element_text(size=7))
-```
-
-Items in the `Groceries` dataset proliferate into a Tesco's list of branded variations.
-The 169 items of the `Groceries` dataset  expand into roughly `r total` Tesco items (see Figure \@ref(fig:tesco-ex)).
-There is, I would suggest, a logistic propensity in things that recommender systems encounter in the ways of global supply chains.
-The multiple listed  in the `Grocery` dataset becomes more open in this setting through new impersonal intersections that must be bounded and ordered.
-The proliferation of things on the shelves of supermarket or grocery warehouse confronts produces a combinatorial problem for data miners.
-`r total` products (actually Patel mentioned 200,000 products) can be combined in a vast number of ways.
-If a typical shopping list has 20 items, then there are `r format(choose(total, 20), 2)` possible lists.
-Some of this vast number of possible shopping lists are highly unlikely in terms of typical purchasing patterns, but potentially valuable as recommendations.
-
-# Ordering relevance through personhood or thinghood?
-
-Combinatorial surplus, however, does not exhaust what might be found in the data.
-The relevancy model that DunnHumby implemented for Tesco (as for so many other  businesses) seeks to balance the tensions between proliferating associations and relative closure.
-It responds to the problem of finding an ordering in which the propensities of things as commodities and the propensities of people mingle with each other, without overflowing  the site of their  ordering encounter, the list.[^181.1] Big data discourse in its promissory mode  attributes potency to personalization: 'most important, using big data we hope to identify specific individuals rather than groups' [@Mayer-Schonberger_2013].
-As Turow and co-authors conclude in their account of transformation of retail space: 'through it all, knowingly and not, and away from the spotlights of fierce social debate, retailers are encouraging daily routines that accept data-driven personalization as a centrifugal public force' [@Turow_2015, 476].
-
-Mark Hansen offers a different perspective on what might happen in these ordering encounters. Hansen attributes 'real potential' to data. Hansen attributes an associative propensity to data:
-
-> [S]urrounding any delimited predictive system is a larger field of data -- what I elsewhere call a “surplus of sensibility” -- that, viewed speculatively, indexes the causal efficacy of the total situation within which this delimited system operates.
-(Effectively, the latter gains its reliability from closing off this larger surplus of sensibility, thereby transforming an always excessive propensity into a [provisionally] closed dataset.) Because it affords data that exceed whatever any given predictive system might include, the data of the world’s causal efficacy -- the data constituting its real potentiality -- always and in principle facilitates knowledge that cannot be restricted to any particular agenda [@Hansen_2015a, 126-7].
-
-Many affirmative and critical analyses of big data and predictive analytics highlight the divisive social orderings implemented through personalization.
-For good reason, they attend to changes in personhood.
-Hansen, by contrast, seeks to conceive of the potential of data to support or provoke different relations to the world.
-Even if the underlying contrast between an open world and a 'provisionally closed dataset' limits the account in ways that I will explore, the ontological transformation  Hansen describes  involves people, commodity-things, and infrastructures.
-The recommender system generates common codings for  the associative propensities of things, the logistical value-chains of supermarket commerce and  the forms of personhood associated with shopping.
-A model of personhood based on purchase data delimits the propensities of things.
-
-[^181.1]: Rather than seeing predictive analytics as a closed form of an open world, I am suggesting that we might begin to locate it in the joint effectuations or intersections within models.
-
-If we think of the  new model that Patel describes as the delimited (or delimiting?) system within   the larger surplus of sensibility, how would the idea of a predictive system as a provisional enclosure of the world's causal efficacy, its 'surplus of sensibility' (a terms that carries some connotations of the critical political economy of surplus value) help us understand what the 'Have you forgotten?' recommendations do?
-It enables us to understand -- as we will see next -- what the Tesco relevancy model borrows from the world.
-By contrast, despite its ontological ambitions, it limits our understanding of how predictive analytics  themselves exist amidst technical ensembles whose operation remains obscure yet central to contemporary worlds.
-
-```{r models, fig.cap='Precision measurements for different machine learning models'}
-include_graphics('figure/patel_compares_models.png')
-```
-
-Patel described the data that Tesco keeps on grocery purchases: 'We have 52 weeks of purchase data for each customer.' These transformations have been widely discussed in the context of big data, where questions of who has how much data are usually most prominent (see [@Kitchin_2014]).
-In terms of the ways that lists order the world, the amount of data is not perhaps not so important as the way in which list-making is re-ordered via a model of personhood.
-Patel introduced the new relevancy model with a graph familiar to machine learners and statistical modellers.
-(A sketch of her graph appears in Figure \@ref(fig:models).)  The graph plots the _precision_ -- the proportion of the recommended products that customers actually purchase -- for several different models.
-In Hansen's media theoretical terms, the graph indexes the 'causal efficacy' of the recommender system, its capacity to include and transform propensities or 'real potentialities' into operational knowledge.
-The graph compares the old rule-based recommender systems with some of the relevancy-model alternatives -- logistic regression, random forests, and a few others -- in terms of their predictions and how those predictions turned out.
-Patel dismissed most of the models quite quickly and focused only on one, the logistic regression model, which did better than other alternatives.
-Why did she say little about the model, which after all, produces the probabilistic predictions on which everything in the recommendation list depends and hence the conversion experience?
-Her interest lay in the precision of the recommendations, and, crucially for our purposes, in how 'most of the gain comes from thinking of good features.'
-
-The models Patel so briefly mentioned are all classifiers, predictive models that spatially order  variables in a dataset as a high dimensional vector space.
-Operating in the vector space, models classify people and events.
-Much of the operational formation of the classifier models received scant reference in Patel's presentation.
-It is likely that in DunnHumby's models the classifying model at the heart of the recommender system constructs a decision boundary in vector space between points that belong to different classes of things [@Mackenzie_2015e].
-The classes of things are binary: `recommended` or `not recommended,` and for many of the models shown in the graph, membership of a class is expressed as a probability.
-The logistic regression model generates probabilities for each product for each customer.
-Patel did not articulate the implication of the contrast between the left and right hand side of the graphic.
-Although the statistical framing appears explicitly (e.g.
-in measuring the precision of  recommender systems), the underlying shift to a statistical model is barely mentioned.
-In becoming predictive, the 'Have You Forgotten?' recommendations have included a version of the social in the form of a  model that seeks to identify relations between variables.
-The relevance of recommendations pivots on the apprehension of social order by a model.
-
-# Sparsity: output 20 recommendations from 200,000 products
-
-Note that neither the data analytic infrastructure that processes  Tesco customer data  - Patel will mention `Hadoop` -- nor the predictive analytic device -- the logistic regression model -- loom very large in the conversion of the recommender system.
-'Most of the gain comes,' Patel observes, 'from thinking of good features.' While this is almost a truism in big data practice (see, for instance, [@Domingos_2012]), what counts as a good feature in a shopping list?
-
-Rather than a surplus of sensibility, predictive models often encounter a dearth of sensibility in the data.
-A  very brief comment from Patel -- 'we have lots of zeros' -- points to the difficult terrain of  'good feature' thinking: the vast, nearly empty matrix of customer-product associations.
-Given that most things in Tesco remain relatively unknown to each other and to customers.
-A matrix that records associations between individual people and different products is bound to be mostly empty.
-Say Tesco has 1 million online customers.
-Each online  shopper has bought some selection of the 200,000 products.
-The customer-product data matrix will be `r format(1e6 * 2e5)` in size.
-The product-customer matrix, the basic vector-space in which all recommender systems operate, remains very sparse and unpopulated.
-Given that any one customer is likely to only have bought 100 or so different products, the probability matrix will be close to `r format(100-(100/2e5)*100, 2)`% empty.
-The data is, as Patel put it, 'massively unbalanced' (any dataset where the items of interest are much rarer than some normal values is said to be unbalanced) and such imbalance would heavily bias the recommender system towards common and only impersonally relevant suggestions, suggestions that might  not produce the desired conversion experience.
-
-```{r sparsity-ex, echo=FALSE, cache=TRUE, fig.margin=FALSE, fig.cap ='Sparsity of the customer-product association matrix' }
-library(ggplot2)
-library(reshape2)
-p = as.logical(rbinom(n=2e5, size=20,0.001 ))
-m = (matrix(p, nrow=1000, byrow=TRUE))
-melted <- melt(m)
-melted$value <- as.logical(melted$value)
-ggplot(melted, aes(x = Var2, y = Var1, fill = value)) + geom_tile() +
-    scale_fill_manual(values = c("grey", "red")) + xlab('Customer id') + ylab('Product id') + theme(legend.position='none')
-```
-
-Since so many people buy milk, the recommendation system might end up always recommending milk products.
-So the data needs to be 'corrected' by, as Patel reports, removing -- under-sampling -- some of the data for common purchases.
-Secondly, even if a recommendation system finds uncommon items that are good recommendations, they should neither be too cheap ('low spend'), nor, as we will see too similar to what the customer has already ordered.
-
-Given this imbalance, distortion and deficit of sensibility in the data, 'thinking of good features' -- ways of traversing the data -- becomes central.
-For our purposes, the exact details of these features matter less than attending to the way in which, contrary to Hansen's account of analytic closure,  probabilities become propensities in data practice.
-The dataset, as we will see, will not be practically treated as 'closed,' as Hansen puts it.
-Although it undergoes acts of enclosure (for instance, in the transformation of the customer-product matrix into vector space), these acts are not conclusive.
-'Thinking of good features' potentially re-opens the dataset and sometimes completely transforms it.
-This transmogrification, which I suggest occurs mostly in quite obscure yet central zones of data practice, is highly mixed.
-It combines experimental techniques, infrastructural dynamics, seasonal and daily temporalities, and above all, an ongoing engagement with probabilities and their fungibilities.
-
-# Experimental settings: A/B allocation and uplift
-
-|Test A|Test B|
-|---|---|
-|Control A|Control B|
-
-For instance, the predictions of the recommender system themselves are the subject of probabilistic treatment.
-Patel described the experimental deployment of  the relevancy model experimentally in a random A/B controlled trial on the Tesco website.
-All customers were  allocated to one of four categories as shown in the table.
-In the  A/B trial,  an identical shopping list, or even the same person, might receive different recommendations.
-The randomised application of the model, drawing on protocols for randomised clinical trials first developed in the 1960s, is widely used in social media platforms and hence in the implementation and observation of the effects of recommender systems.[^238.1] Random allocation of customers to the four categories (Test A, Test B, Control A, Control B) adds a layer of probability to the recommender system in the name of statistical validation of the effects or the 'uplift' of the model.[^122] Ironically,  the effects of a predictive model cannot be known in advance.
-They can only be observed experimentally.
-
-[^238.1]: It is difficult to gauge how much. Facebook reports that it has more than 1 million models operating in its infrastructure. [@Dunn_2016]
-
-It is perhaps significant that this random allocation of customers to control and test groups occurs without any relation to the particular profile or propensities of the customer.
-It seeks to statistically validate the effects -- the *uplift* -- of the model on  conversion events by allowing the effects of the model on what people do to be measured.
-The uplift refers to conversion events associated with the same groups of people.
-Effectively an experiment in inhabiting two different worlds, the randomised control trial  sets up a feedback look between the predictive model (the logistic regression), and the world.
-
-[^122]: See [@Lo_2002] for an introduction to the construction of experimental set ups for uplift models.
-Market researchers and data miners seem to have started adopting a statistically grounded experimental predictive practice sometimes around 2000.
-See for instance [@Almquist_2001]  Response modelling, 'propensity' modelling and uplift modelling all seek to identify associations between 'treatments' or interventions and the 'Responders', the people  affected by the treatment.
-
-# Infrastructure: the realtime engine serves
-
-A second layer of probability affects the operation of the recommender system.
-Much discussion of recent media has highlighted its real-time operation (see [@Kittler_1993; @Virilio_1997; @Hansen_2015; @Bollmer_2016]).
-Patel's description of the relevancy model infrastructure sketched a different temporal patterning, suggesting immediacy in relation to shopping, but scheduling and periodicity in relation to the infrastructural operations of the predictive model.
-While raw data from Tesco Online feeds into the model every hour, a recommendation list for each customer is only generated once a week.
-Customers only shop online every few days at the most, and in some cases, only every few weeks.
-To continually update the top 200 recommendations for every customer would be computationally expensive.
-
-A customer does not see their 200 weekly recommendations.
-The recommender system generates recommendations, but after 'applying business rules, only five are  shown on the Tesco Online website.
-Patel briefly mentioned  specific infrastructural elements such as `hadoop` [@ApacheSoftwareFoundation_2009].
-The logistics of running a predictive model affect listing practices.
-The possibility of adjusting the recommendations for every customer weekly derives from an infrastructure capable of collecting data, assimilating that data to a predictive model able to generate approximately 200 million relevant recommendations, where 'relevance' depends on a matrix of probabilities of associations between people and things that shifts in time.
-Like the global supply chain logistics optimised for flow of commodities [@Tsing_2009; @Neilson_2012], `hadoop` and its legion of variants (`mahout`, `spark`, `hive`, `pig`, `yarn`) operationalise flows of data at an infrastructural rather than analytic scale.
-In many respects, the quickly glossed-over infrastructural deployment of DunnHumby's relevancy model is the primary conversion event: the logistic regression model at the heart of the recommender system is no longer an analytical device but an operational one.
-
-# Similarity, substitution, complementarity and micrological propensities
-
-There is more to probabilities in prediction than a simply enclosure and quantification of the surplus of sensibility, the overwhelming variety, differences and noise of the world.
-I am suggesting, contrary to Hansen, that the propensities of things are enclosed as  probabilities in predictive analytics only to the extent that probabilities themselves become propensities.
-Recommendations are relevant in a recommender system or any other operational predictive model to the precise extent of their probability.
-Relevance is closely connected to propensities in the predictive condition.
-From DunnHumby's standpoint, the relevance of a recommendation relates directly to conversion events.
-Conversion events happen in the world.
-When a recommender system discovers a common coding of human practice, and logistical configurations, with all their value differentiations, they actualise a propensity that hitherto remained potential.
-From the standpoint of the ontological exchangeability between probability and propensity,  prediction stems from the propensities of things, what Hansen describes as the 'complex and diffuse calculus' that only provisionally congeals as a conversion event.
-The 'micrological propensities' of things support the relevance of recommendations.
-
-How does the DunnHumby model deal with the exchange and oscillation between propensities and probabilities?
-In her presentation, Patel  spent much time addressing the problem of 'basket similarity,' the problem that the recommender system might recommend, for good reason, something similar to items already in a customer's basket.
-The recommender system or statistical model needs to know how things are related to each other so that it does not waste recommendations on overly similar items.
-A customer might be willing to substitute a similar item for something they have already chosen, but they are more likely to accept a recommendation for an item that complements items in their list.
-
-Whereas the association rule-based recommender system focused on thing-thing complementarities -- if you buy bread, you might want milk, if you buy beer, you might want diapers -- the personal relevance substitutions look for micrological propensities that relate to gradations of preference, and variations in time and place.
-Micrological propensities in the list relate to a particular time and place, amidst a continuously differentiable pathways between things.
-
-How could the recommender system sense the possible substitutions and complementaries between things for a given customer, especially since a person's sensibilities and susceptibilities concerning groceries might be shaped by wider social groups, orderings and circumstances?
-Patel's comments on this were very brief but suggest that the 'Have You Forgotten?' list becomes an increasingly mutable entity, open to many substitutions and intensive re-writings.
-
-Like the mention of `hadoop`, what Patel said about substitution and complementarity draws on extensive but relatively well-established data practices in machine learning.
-She mentioned,  for instance,  a 'design matrix of 14,000 columns' and the use of '`L1 regularization` to drive coefficients down to zero' in a new feature of the model called 'self-learning substitutes,' a model that examines the recommendation list and predicts substitutes and complementarities for 'all products' in the list.
-The 'self learning substitutes' feature of the relevancy model draws on another data matrix, the product similarity matrix, itself generated from all previous recommendations that have led to conversion events (that is, 'a user clicked as the response').
-A customer's recommendation list might already include `Tesco Ginger Nuts 300g.`   The 'self-learning substitution model' seeks to replace list items with similar but higher value items: `Mcvitie's Ginger Moments 400g.` What is being modelled here?
-No longer whether to recommend a product to a customer or not, but micrological conversions.
-
-Similarly, modelling complementarities between things for a particular person entails a matrix of comparisons.
-`Nutella` complements `bananas,` but the converse is not necessarily true since `bananas` need not complement `nutella,` especially if I have nut allergies.
-From the standpoint of customer data science, Patel emphasised the operational significance of predicting complementarities:  'complementarity drives conversion.' Bringing the matrix of similarities expressed as numerical associations between products into the model improves the conversion of recommendations into purchases.
-From the standpoint of a sociological account of social ordering that associate people and things through propensities, complementarities pose a different problem: what does it mean for social order if the relations between things shift in accordance with our experience of the world as rendered in the intricately varying weights of an L1-regularized logistic regression model?
-In the similarity matrix, shaped by the history of past recommendations and purchases made by this customer, the propensities of things drive down the surplus of sensibility associated with what people want.
-
-# Conclusion
-
-In their mundane intersectional practices of ordering, recommender systems show how  predictive practices affect existing forms of social order in the world.
-Rather than directly positing an ontological transformation (as Hansen does), or pointing to the erosion of social order (as Turow does), I have explored some of the conversion events associated with a shift from a data-mining rule-based system to a predictive recommender system.
-The DunnHumby model and the Tesco 'Have You Forgotten?' recommendations present many of the traits of the conversion processes associated with big data.
-Rather than discontinuous changes, the conversion of the recommender system via a relevance model suggests an ongoing problematization of propensities.
-We have seen several types of conversion events: ongoing transformation of models and infrastructures, the experimentalisation  of recommendations, and the broader diffusion of a calculus of probabilities into a fibrous weave of propensities and conditionalities, warping over time.
-The predictive assemblages connect logistical ensembles to the negotiation, compromises, rituals and local orderings of everyday life.
-
-I have been suggesting that predictive listing in the form of open-ended recommendation changes social order.
-It establishes a common coding between machinic and ritual repetition (as exemplified in the shopping list).
-It also suggests some ways in which repetition or reiteration  can also participate in changing and emergent order.
-By understanding the world in terms of propensities, we have a chance  to engage with contemporary predictive processes in ways that both acknowledge their relentless and exhausting pressure to consume more and retain some chance of inhabiting the world differently.
-
-Rather than understanding prediction as the closed form of an ontologically open world ('the propensities of things'), the DunnHumby relevance model, in all its statistical, infrastructural, organisational and banal complexity, suggests that the predictive conversion remains constitutively incomplete.
-Shopping lists in all their banal yet primary orderings are fertile terrain for prediction because they are incomplete, somewhat open and mutable forms.
-The constitutive incompleteness and openness of lists are very closely entangled with ongoing processes of ordering and othering, and at the same time, can accommodate predictive operations.
-Lists undergoing a 'conversion' experience as they become more probabilistic, but their heterogeneous character as operations caught up in models, infrastructures, experiments, and the flow of commodities means that they can never form a closed set or dataset.
-The problem of 'constructing good features,' a problem that any predictive model must address, persists as the site of experiments in partly open structures.
-
-A constitutive openness to new associations, new combinations, new groupings or actors is integral to the conversion events associated with contemporary data practice.
-In their account of the Netflix recommender system, Ted Hallinan and Peter Striphas make the following argument:
-
-> These systems, and more importantly their algorithms, play a critical role in deciding which articles (or parts thereof) gain admission to the cultural realm, and in what form.
-Their doing so thus points in the direction of another universe of reference—a court of algorithmic appeal in which objects, ideas, and practices are heard, cross-examined, and judged independently, in part, of human beings [@Hallinan_2014, 13]
-
-Despite the somewhat troubling reference to 'another universe,' as if Netflix or DunnHumby's recommender systems occupy a different reality, Hallinan and Striphas usefully and aptly highlight how such systems  mingle  objects, ideas and practices.
-
-Something similar happens to shopping in the Dunnhumby relevance model.
-In the customer-product association matrix, in the product-product similarity matrix, in the 14,000 variables included in the logistic regression model, a 'universe of reference' takes shape.
-The shopping list, the commonplace inscription of normative and actual social order, takes on a complex temporal-material structure that mixes prediction, infrastructural constraints, the individual characteristics of list-makers, statistical validations, database architectures and above all the surplus sensibility associated with  circulation and consumption of commodities.
-Shopping lists undergoing a conversion experience as they become more probabilistic, but the idea that probabilities are the closed form of propensities does not account for the distribution of probabilities in relations between things, between people, between people and things, and between things and people.
-
-# References
+# F. Conclusion

### 2017-07-18 log.rmd
restructuring last section

@@
 - TODO: play with the structuring of Section E.
--
+
+## Tue 18 Jul 2017 07:18:46 BST
+- adding Marcus on prototyping method into intro; 
+- starting on section E: it will be rough
+- how do I actually use the social theory of lists - what happens to the mapping? to the discrepancies between normative and actual, and to the siting of the list?
+- TODO: write about this in the conclusion
+- remapped last 12 pages a bit, taking into account the Popper argument and critiquing the Hansen a bit more;  see page 21 of July printout for summary; still feels like need more experiments or prototypes; particularly for the model itself;
+- TODO: rewrite section on logistic regression in terms of the theory of the list -- as a mapping;
+- restructuring of Section E: the idea is to combine the Hansen reading of Popper; the aspects of the relevance model (logistic regression,  much data, a/b experiments, good features, substitution/complementarity), and the social theory of the list
+- popper quotes in order:
+
+> the introduction of propensities amounts to generalizing and extending the idea of forces again  14
+
+> they should be regarded as _inherent in a situation_ 14
+
+> we need _a calculus of relative or conditional probabilities_ as opposed _a calculus of absolute probabilities_ 16
+
+> in many kinds of events this is not the case, and the propensities cannot be measured because the relevant situation changes and cannot be repeated 17
+
+> in our real changing world, the situation and, with it, the possibilities, and thus the propensities, change all the time.
+They may certainly change if we, or any other organisms,  _prefer_ one possibility to another; or if we _discover_ a possibility where we have not seen one before 17
+
+> experiments. ... by creating, at will, artificial conditions that either exclude, or reduce to zero, all the interfering and disturbing propensities 23
+
+> What may happen in the future .. is, to some extent, open. There are many possibilities trying to realize themselves, but few of them have a very high propensity, given the initial conditions. 22
+
+- an experiment for the conditional probability argument using logistic regression;
+    - simulated 500k customers in terms of Taking Part data
+    - simulate 52 weeks of purchase history using Tesco API data
+    - build a separate model for each customer; 
+    - see how that model can recommend substitutions or complementarities; these are 'new possibilities'   
+- TODO: move the personalization discussion higher up, perhaps in the introduction
+- rough reconstruction is ~4.2k, but haven't brought in the experiments much

### 2017-07-18 reconstruction.rmd
restructuring last section

@@
 
-This paper is a philosophical reconstruction [@Dewey_2017] of an online grocery ordering system.
+This paper is a philosophical reconstruction [@Dewey_2017] of an online grocery ordering system undergoing an upgrade.
 The reconstruction is empirical in certain respects.
 It departs from an ethnographic moment: being a member of the audience at an industry/academic conference where a data scientist was describing how recommendations were created for online grocery orders.
-Given this highly mediated and limited encounter with a recommender system and one of its designers, the paper constructs a multi-layered account.
-It draws on anthropological and sociological research concerning shopping and list-making to situate the recommender system amongst the social ordering practices of shopping.
-It conducts several code-based experiments in order to reconstruct, using widely available code resources resources such as API (Application Programmer Interfaces) and software libraries for machine learning, some important elements of the system in question.
-It makes use of archaeological approaches, in the sense developed by Michel Foucault [@Foucault_1972] in order to identify and map the functional statements and practices underpinning the knowledge and order generated by such systems over time.
-Not least, I draw on the philosophy of science and media theory to frame an engagement with data, algorithms and recommendations around probabilities and their enactments.
-Like some recent work in science and technology studies,  anthropology and media studies [@Marres_2017; @Rabinow_2004; @Bogost_2012], through this combination of approaches,  the paper itself provisionally allows its own reconstruction to be shaped by an experiential encounter with the object it describes.
+Starting from this highly mediated and fragmentary encounter with a recommender system and one of its designers, the paper explores the problem the system addresses.
+It draws on anthropological and sociological research concerning shopping and list-making to situate the recommender system amongst the social ordering practices of shopping and shopping lists.
+It makes use of archaeological approaches, in the sense developed by Michel Foucault [@Foucault_1972] in order to identify and map important functional statements and practices configuring the knowledge and order generated by such systems over time.
+It conducts several code-based experiments in order to reconstruct, using widely available code resources resources such as API (Application Programmer Interfaces) and software libraries for machine learning, some prototypical elements of the system in question.
+In doing this, it draws  on the philosophy of science and media theory to frame an engagement with data, algorithms and recommendations around probabilities and their worldly enactments.
+
+Like some recent work in science and technology studies,  anthropology and media studies [@Marres_2017; @Marcus_2014; @Bogost_2012], through this combination of approaches,  the paper allows its own reconstruction to be provisionally shaped by an experiential encounter with the object it describes.
 At times, this mode of empirical philosophical engagement with its objects may well seem overly concerned with the technical minutiae of the grocery ordering system, and lacking in  sociological rigour or empirical evidence.
@@
 While calculations of probability have  long-standing importance in many settings (insurance, medicine, experimental and field sciences, operations research, risk analysis, engineering design, public health,  economic modelling, etc.), probabilities have taken on a much thicker situational texture.
+In recent research and opinion on  data practice, personalization has been seen as the linchpin of predictive modelling.
+The reconstruction of the recommender system suggests a different way of seeing these transformations.
 Operational probabilities differ from epistemic probabilities in their position in the world, as part of infrastructures, transactions, and event-structures.
-Taking their complex siting into account, and accepting their _physical_ reality, we might come to understand the world they belong to as what the philosopher of science Karl Popper understood as a 'world of propensities, as an unfolding process of realizing possibilities and of unfolding new possibilities' [@Popper_1990,19].
-Given this understanding of operational probability, our understanding of what is enacted in contemporary data analytics, especially in association with platforms [@Gillespie_2010] might shift.
+Taking their complex siting into account, and accepting their _physical_ reality, we might come to understand the world operative probabilities belong to as what the philosopher of science Karl Popper understood as a 'world of propensities, as an unfolding process of realizing possibilities and of unfolding new possibilities' [@Popper_1990,19].
+Given this understanding of operational probability, our understanding of what is enacted in contemporary data analytics, especially in association with platforms [@Gillespie_2010] where the operationality of probabilities are heightened,  might shift.
 
-The broad implication of the argument concerns how we think about social order or regularities: proponents and critics of Big Data present the advent of large volumes of data and analytical techniques (machine learning in particular) as transforming organisations, institutions, economic processes and social life.
-I suggest too that a transformation or conversion event is occurring, but one which might open up the possibility of experiencing the world differently, and conceiving of social ordering processes somewhat afresh.
-Even if the argument is centred on an eminently social practice such as shopping, the analysis of how a world of propensities continuously and unevenly converts into realizations applies across the board, from the most physical to the most ideal or abstract settings.
+The broad implication of the argument concerns how we think about social order or regularities amidst change or transformation.
+Proponents and critics of Big Data present the advent of large volumes of data and analytical techniques (machine learning in particular) as transforming for better or worse organisations, institutions, economic processes and social life.
+Personalization has been a pivotal element, they argue, of this transformation.
+I suggest too that a transformation or conversion event is occurring, but one which might also, even in the midst of much ongoing personalization,  open up the possibility of experiencing the world differently, and conceiving of social ordering processes somewhat afresh.
+Even if the argument is centred on shopping, the analysis of how a world of propensities continuously and unevenly converts into realizations applies across the board, from the most physical to the most ideal or abstract settings.
 The average everydayness of the example here -- grocery shopping -- allows some first hand engagement with the messiness, the entanglements, and potentials of this conversion event in ways that other interesting settings -- particle physics for instance -- do not.
@@
 
-The shelves of large contemporary supermarket are the endpoint of global logistic supply chains, in all their logistical, value-transforming and brand-mediated hypercomplexity (see [@Neilson_2012; @Tsing_2009]).
+[TBA: THIS PARA could be moved down] The shelves of large contemporary supermarket are the endpoint of global logistic supply chains, in all their logistical, value-transforming and brand-mediated hypercomplexity (see [@Neilson_2012; @Tsing_2009]).
 Groceries imply a planetary geography of agriculture, industry, transport, communication and financialisation animated by flows of labour and capital.
@@
 
-The conversion of  collaborative filtering into predictive personalisation appeared in Patel's presentations.
-Patel describes the move from 'a rules-based system'  to  a 'personal relevance model'.
-Each term in that phrase -- personal, relevance and model -- has some weight in the conversion event.
-We can assume that what Patel describes as  the 'rule-based system' refers  to the `apriori` algorithm, developed  by Rakesh Agrawal and Ramakrishnan Srikanti at IBM Research Alameda in the early 1990s [@Agrawal_1994].
-A now-classic approach to 'market basket analysis,' it was listed as one of the top ten data mining algorithms in a survey conducted amongst data miners [@Wu_2008] and usually attracts a chapter in data-mining and machine learning textbooks (e.g.
-[@Hastie_2009]).
+[BREAK HERE]
 
@@
 
+## The realization of propensities
+
+>the introduction of propensities amounts to generalizing and extending the idea of forces again  14
+
+>they should be regarded as _inherent in a situation_ 14
+
+The conversion of  collaborative filtering into predictive personalisation appeared throughout  Patel's presentations.
+Patel described a move or 'conversion'  from 'a rules-based system'  to  a 'personal relevance model' as the basis of the recommender system.
+Each term in that phrase -- personal, relevance and model -- has some weight in the conversion event.
+Again, dealing with the sometimes telegraphic style of machine learning presentations means making some assumptions.
+It is reasonable to assume that what Patel describes as  the 'rule-based system' refers  to the extremely well-known `apriori` algorithm, developed  by computer scientists Rakesh Agrawal and Ramakrishnan Srikanti working at IBM Research Alameda in the early 1990s [@Agrawal_1994].
+A now-classic approach to 'market basket analysis,' it was listed as one of the top ten data mining algorithms in a survey conducted amongst data miners [@Wu_2008] and usually attracts a chapter in data-mining and machine learning textbooks (e.g.
+[@Hastie_2009]).
+The interest of `apriori` for our purposes is that it relies on massive datasets, but does not personalize its recommendations.
+
+The media theorist Mark Hansen has recently suggested that 'predictive analytics are discoveries of micrological propensities that are not directly correlated with human understanding and affectivity and that do not by themselves cohere into clearly identifiable events' [@Hansen_2015a, 111-112].
+Hansen's account differs from many widely shared views of big data.
+In the face of the so much personalization (and Patel's presentation exemplifies this), Hansen suggests a somewhat processes of impersonalization.
+If  much predictive practice attempts to elicit 'micrological propensities' from data,   Hansen, drawing on the work of the philosopher of science Karl Popper,  links data to things via probabilities:
+
+>Whatever explanatory and causal value predictive analytics of large datasets have is, I suggest, ultimately rooted in this ontological transformation whereby probabilities are understood to be expressions of the actual propensity of things [@Hansen_2015a, 120]
+
+Hansen theorises big data or predictive analytics as an 'ontological transformation' that deploys probabilities in an evermore closely woven and encompassing expression of animated, eventful, propensities of things.
+Predictive analytics links probabilities as calculations to propensities or the mutable associative agencies of things.
+The practical and indeed empirical  question is whether such transformations or 'conversion' in relations between probabilities and 'the propensities of things'  can be detected and articulated in the prosaic setting of shopping lists and recommender systems.[^20.1] The series of prototyping reconstructions I am about to discuss will explore the tenability of this rather radical alternative to regular narratives of personalization.
+
+[^20.1]: I will argue that 'conversion' is a preferable term for these changes.
+It highlights  re-orientations in subjects, experience, things, numbers and structures that ranges more widely than Hansen's 'ontological transformation' formulation.
+Not much hinges on the choice of terms, but 'conversion' happens to be the term used by Patel.
+
+## Apriori as conditional probabilities
+
+> we need _a calculus of relative or conditional probabilities_ as opposed _a calculus of absolute probabilities_ 16
+
+```{r arules-ex, cache=TRUE, fig.cap='Frequency of association betwene items in the `Grocery` dataset', message=FALSE, echo=FALSE, fig.margin=TRUE}
+library(knitr)
+library(arules)
+library(arulesViz)
+library(datasets)
+data(Groceries)
+itemFrequencyPlot(Groceries,topN=10,col=rainbow(10), ntype="absolute",  cex.axis = 0.55)
+```
+
+```{r arules-ex2}
+options(digits=2)
+rules<- apriori(data=Groceries, parameter=list(supp=0.001,conf = 0.15,minlen=2), appearance = list(default="rhs",lhs="whole milk"), control = list(verbose=FALSE))
+rules_sorted <- sort(rules, decreasing=TRUE,by="confidence")
+kable(inspect(rules_sorted[1:5]), caption='The first five association rules for the `Groceries` dataset')
+```
+
+The `apriori` algorithm finds sets of items that commonly occur together in transactions.
+These commonly occurring sets are expressed as 'association rules.' For instance, using the `Groceries` dataset in the `R` package `arules,` we can readily see how Walmart *might* have discovered   the fabled association between diapers and beer.[^149.1] Figure \@ref(fig:arules-ex) shows how often common items appear.
+`Whole milk` appears most frequently.
+The association rule-based model for the `Groceries` dataset illustrates some characteristic ordering practices of 1990s-style data-mining approaches to lists (the `Groceries` dataset  was acquired from a 'local German supermarket' [@Hahsler_2006a]).
+The association between items is expressed in the form of _rules_ not a model.
+These rules, a term derived from older decision support literature, define sets of associations between items.
+Table \@ref(tab:arules-ex2) shows that the first rule has a stronger association than the second.
+Milk and vegetables is more likely an association than milk and buns.
+A rules-based recommender system should, therefore, might put them in that order in the 'Have You Forgotten' list.
+
+[^149.1]: One of the founding narratives of big data customer analytics, the beer-diaper association was not, according to [@Power_2002] found through data mining at all.
+
+Although association rules and the `apriori` algorithm present social order in terms of sets of frequency-weighted associations, an `apriori`-based recommender system contains no statistical model.
+It also no grasp on personal relevance since the associations it defines are only between things.
+Put in terms of the big data conversion, the rules-based system is not predictive, or personal.
+The `apriori` algorithm might be seen as an attempt to grasp the excess of things in a Tesco supermarket by identifying subsets whose closure derives from the associations between things that people construct (through buying habits, by virtue of a shopping list, etc.).
+In association rules, personhood or experience is irrelevant.
+The openness of these rules -- the `apriori` algorithm can product any number of rules of diminishing strength -- points towards the propensity of things to find themselves together in shopping baskets, but cannot cannot individuate that propensity beyond demographic quantiles and categories.
+
+## `Apriori` meets the API
+
+Even if `apriori`  elicits associations between things, it struggles with the propensity of commodities to multiply.
+A simple illustration of the combinatorial problem faced by recommender system can be developed using the `Grocery` dataset.
+If we take all the items in the `Grocery` dataset and paste them into the 'shopping list' box on the Tesco grocery website (or as I side, run them as searches on the TescoLab Product API), each of the 169 items in the dataset yields dozens and sometimes thousands of products from Tesco online.
+
+```{r tesco-ex, echo=FALSE, cache=TRUE, fig.cap ='Tesco surplus',  fig.margin=TRUE}
+    library(ggplot2)
+    tesco = read.csv('data/reference_data/tesco_groceries.csv')
+    total = sum(tesco$actual, na.rm=TRUE) 
+    tesco = na.omit(tesco)
+    tesco$actual = as.integer(tesco$actual)
+    ggplot(tesco[tesco$actual>50,], aes(x=labels, y=actual, fill=labels)) +geom_bar(stat='identity') + coord_flip() + ggtitle('Tesco grocery items with more than 50 products') + theme(legend.position="none", axis.text=element_text(size=7))
+```
+
+Items in the `Groceries` dataset proliferate into a Tesco's list of branded variations.
+The 169 items of the `Groceries` dataset  expand into roughly `r total` Tesco items (see Figure \@ref(fig:tesco-ex)).
+There is, I would suggest, a logistic propensity in things that recommender systems encounter in the ways of global supply chains.
+The multiple listed  in the `Grocery` dataset becomes more open in this setting through new impersonal intersections that must be bounded and ordered.
+The proliferation of things on the shelves of supermarket or grocery warehouse confronts produces a combinatorial problem for data miners.
+`r total` products (actually Patel mentioned 200,000 products) can be combined in a vast number of ways.
+If a typical shopping list has 20 items, then there are `r format(choose(total, 20), 2)` possible lists.
+Some of this vast number of possible shopping lists are highly unlikely in terms of typical purchasing patterns, but potentially valuable as recommendations.
+
+##  The list in its mapping to probabilities
+
+Conditional probability; the models
+[TBA - section on the role of probability; and logr]
+
+## Repeating sufficiently often to measure
+
+52 weeks of data
+> in many kinds of events this is not the case, and the propensities cannot be measured because the relevant situation changes and cannot be repeated 17
+
+##  Platform Experiments that reduce interfering propensities
+
+> experiments. ... by creating, at will, artificial conditions that either exclude, or reduce to zero, all the interfering and disturbing propensities 23
+
+## The openness of the data -- new possibilities
+
+> What may happen in the future .. is, to some extent, open. There are many possibilities trying to realize themselves, but few of them have a very high propensity, given the initial conditions. 22
+
 # F. Conclusion
+
+In their mundane intersectional practices of ordering, recommender systems show how  predictive practices affect existing forms of social order in the world.
+Rather than directly positing an ontological transformation (as Hansen does), or pointing to the erosion of social order (as Turow does), I have explored some of the conversion events associated with a shift from a data-mining rule-based system to a predictive recommender system.
+The DunnHumby model and the Tesco 'Have You Forgotten?' recommendations present many of the traits of the conversion processes associated with big data.
+Rather than discontinuous changes, the conversion of the recommender system via a relevance model suggests an ongoing problematization of propensities.
+We have seen several types of conversion events: ongoing transformation of models and infrastructures, the experimentalisation  of recommendations, and the broader diffusion of a calculus of probabilities into a fibrous weave of propensities and conditionalities, warping over time.
+The predictive assemblages connect logistical ensembles to the negotiation, compromises, rituals and local orderings of everyday life.
+
+I have been suggesting that predictive listing in the form of open-ended recommendation changes social order.
+It establishes a common coding between machinic and ritual repetition (as exemplified in the shopping list).
+It also suggests some ways in which repetition or reiteration  can also participate in changing and emergent order.
+By understanding the world in terms of propensities, we have a chance  to engage with contemporary predictive processes in ways that both acknowledge their relentless and exhausting pressure to consume more and retain some chance of inhabiting the world differently.
+
+Rather than understanding prediction as the closed form of an ontologically open world ('the propensities of things'), the DunnHumby relevance model, in all its statistical, infrastructural, organisational and banal complexity, suggests that the predictive conversion remains constitutively incomplete.
+Shopping lists in all their banal yet primary orderings are fertile terrain for prediction because they are incomplete, somewhat open and mutable forms.
+The constitutive incompleteness and openness of lists are very closely entangled with ongoing processes of ordering and othering, and at the same time, can accommodate predictive operations.
+Lists undergoing a 'conversion' experience as they become more probabilistic, but their heterogeneous character as operations caught up in models, infrastructures, experiments, and the flow of commodities means that they can never form a closed set or dataset.
+The problem of 'constructing good features,' a problem that any predictive model must address, persists as the site of experiments in partly open structures.
+
+A constitutive openness to new associations, new combinations, new groupings or actors is integral to the conversion events associated with contemporary data practice.
+In their account of the Netflix recommender system, Ted Hallinan and Peter Striphas make the following argument:
+
+> These systems, and more importantly their algorithms, play a critical role in deciding which articles (or parts thereof) gain admission to the cultural realm, and in what form.
+Their doing so thus points in the direction of another universe of reference—a court of algorithmic appeal in which objects, ideas, and practices are heard, cross-examined, and judged independently, in part, of human beings [@Hallinan_2014, 13]
+
+Despite the somewhat troubling reference to 'another universe,' as if Netflix or DunnHumby's recommender systems occupy a different reality, Hallinan and Striphas usefully and aptly highlight how such systems  mingle  objects, ideas and practices.
+
+Something similar happens to shopping in the Dunnhumby relevance model.
+In the customer-product association matrix, in the product-product similarity matrix, in the 14,000 variables included in the logistic regression model, a 'universe of reference' takes shape.
+The shopping list, the commonplace inscription of normative and actual social order, takes on a complex temporal-material structure that mixes prediction, infrastructural constraints, the individual characteristics of list-makers, statistical validations, database architectures and above all the surplus sensibility associated with  circulation and consumption of commodities.
+Shopping lists undergoing a conversion experience as they become more probabilistic, but the idea that probabilities are the closed form of propensities does not account for the distribution of probabilities in relations between things, between people, between people and things, and between things and people.

### 2017-07-19 log.rmd
rewriting conclusion

@@
 - rough reconstruction is ~4.2k, but haven't brought in the experiments much
+
+## Wed 19 Jul 2017 06:46:17 BST
+- try writing the conclusion to give shape to various shifts in emphasis;  
+- made more notes in Dewey on reconstruction; see dewey_reconstruction_1957.md
+- TODO: put more Dewey in intro and conclusion
+
+- order of points in the conclusion:
+    - the specific situation is grocery list  and online shopping recommendations, not shopping in general, but the conversion or transformation of a  recommender system. 
+    - a small theory of shopping and shopping lists: shopping as a constant negotiation between normative and actual orders; list as an intersectional ordering device -- it creates a set of associations; it maps them to an order; but that order can be changed and can be re-associated
+    - these systems largely understood in terms of personalization; undoubtedly occurring. But  something else underpins  personalization and allows it to occur
+    - probabilization allows personalization and exceeds it: associations seen in the recommender system include person-thing, thing-thing, person-in-time; these associations overflow closure; the argument here has focused on describing how this affects attempts to construct predictive models
+    - suggests an alternative theory of social order in which association is not any single organization but propensities, or tendencies towards realization; 
+    - conversion events are what is going on in big data more generally: how do we think about them without being caught in inherited patterns of thought. 
+    - reconstruction and what it does: focuses on specific situations in order to get at the trouble, the problem, the conflict that attracts our interest to them; specifically addresses a broader inherited problem in relation between science and society
+
+- ok, that might work, but still a lot happening; there is no single clear point here at the moment; will have to simplify this more to make it work as article; the point that matters most to me is the personalization one; but I'm not sure exactly why ...
+I know philosophically -- I'm interested in probability -- but not so sure why in terms of the argument of the paper;   using probability and
+- now trying to draft the conclusion: wrote 500 words;

### 2017-07-19 reconstruction.rmd
rewriting conclusion

@@
 
+The 'minute and extensive scrutinizing' [@Dewey_1957] of a recommender system undergoing a conversion from an older data-mining approach to a current 'personal relevance model' approach addresses a specific situation: online grocery shopping (not shopping in general) and the ordering of lists.
+
+The theory of shopping lists constructed at the beginning of the paper was a scaffolding component.
+It draws on anthropology, sociology and mathematics to suggest that we see shopping lists as the site of a complex negotiation between different social orders.
+The list itself as an inscriptive device functions as an ordering device, a device that accommodates an immense variety of specific habits, dispositions, values and associations.
+The mapping component of the list, the order that arranges the set of items according to different interests and aims, is the operational focus of recommender systems.
+
+The personalization of recommendations has been a defining characteristic of many big data conversion events.
+Rightly, the effects of all-consuming personalization have been the target of critics and critical research.
+The archaeology of the Tesco recommender system, including its early experiments with demographic data-mining, its adoption of a rule-based 'market basket' analysis system, and its current implementation of a 'personal relevance model' certainly follows that trajectory.
+But the conversion event is not reducible to personalization.
+
+Probabilization precedes personalization.
+When we observe what is practically done to construct a personal relevance model and apply it to a shopping list, we see, conditional probabilities, predictive models, the importance of data accumulated over time,  platform experiences, the rhythms of infrastructural maintenance, and open-ended construction of new features in data, all of which negotiate shifting and complicated associations between people and things, people and people, and things with things.
+If social order is made of associations, is to be social is to associate, then the forms of association taking shape in this and other similar conversion events might be understood as understand as operationalizing probability in matrices of propensity.
+As Popper argues, probabilities are being realized in the same way that fields of force were realized in modern physics.
+Probabilization does not take place in physics laboratories but on platforms, combining databases, compute capacity, interfaces and global supply chain logistics.
+
+The fact that probabilisation largely takes the form of personalization tends to obscure the re-framing of social order taking place.
+This ordering  depends on  the mappings that inscriptive devices such as shopping lists  display.
+Only because shopping is already somewhat ordered by lists can a partial re-mapping, provisionally conducive to new events, take place.
+The conversion entailed in this re-mapping is not just about perosnalization.
+It is more ontological than that since it includes relations running multi-laterally between people and things in time.
+It is less ontological, however, in that it is not a sudden transformation, but an ongoing process of adjusting, scaffolding, intervening and configuring orders.
+
+Reconstructions of specific conversion events associated with big data can work in different ways.
+The combination of social theory, archaeology, philosophy and model/data prototypes used in this paper sketches one possible way of addressing the trouble, the disturbance or instability that seems to affect many discussions of big data.
+
+------------
+
 In their mundane intersectional practices of ordering, recommender systems show how  predictive practices affect existing forms of social order in the world.

### 2017-07-20 log.rmd
working on section E probability

@@
 - now trying to draft the conclusion: wrote 500 words;
+
+## Thu 20 Jul 2017 08:59:07 BST
+- today is Section E; I have about 7 hour left to finish this; so will need to get that section done today. 
+- working hard on the probability stuff - wrote a section introducting Popper; reframed arules discussion. 
+- TODO - short paragraph summarising the social theory of list ordering; make sure to add stuff on the ordering-remapping of order in the list;

### 2017-07-20 reconstruction.rmd
working on section E probability

@@
 
-TBA - short paragraph summarising the social theory of list ordering
+TBA - short paragraph summarising the social theory of list ordering; make sure to add stuff on the ordering-remapping of order in the list;
 
@@
 Writing just after the advent of web-based e-Commerce, Resnick and Varian  imagined recommender systems augmenting the 'collaborative filtering' that people do when they write and read reviews of products and services (for instance, on the travel website, TripAdvisor; [@Scott_2012]).
-In 1997,  Resnick and Varian thought that 'people provide recommendations' and recommender systems would aggregate and rank for recipients.
+In 1997,  Resnick and Varian expected that 'people [would] provide recommendations' and recommender systems would aggregate and rank recommendations for recipients.
 A slightly later review, [@Schafer_2001], diagrams an augmented 'natural social process' with a range of elements, technologies, inputs and outputs, with varying degrees of personalization.
-
-[BREAK HERE]
 
@@
 
-## The realization of propensities
+This section of the paper returns directly to Patel's presentation, and the character of the newly implemented  recommender system.
+In the discussion that follows, the description of components of the new recommender system -- predictive models and their parameters, infrastructural provisioning, product similarity matrices --  and the problems of  recommendation -- too many products, changes over time, unstable propensities of things in their associations with people -- together suggest that personalization, and the 'personal relevance model' have a troubled relation to social order.
+I develop an analysis centred on probabilities and processes of probabilisation -- rendering situations calculable as probabilities -- in exploring Patel's account of the recommender system.
+The motivation to focus on probabilities and probabilisation is explicitly philosophically reconstructive.
+It represents an attempt to reframe the operation of the recommender system in a way freed from lingering incompatibilities between calculation and social life [@Dewey_1957, 26] .
 
->the introduction of propensities amounts to generalizing and extending the idea of forces again  14
+In order to do this, we need a working notion of probability.
+I draw directly on the work of Karl Popper for this.
+In an essay written towards the end of his career, Popper presents an alternative account of probabilities as real processes, in the same way that contemporary sciences might see forces as real [@Popper_1990].
+Somewhat counter-intuitively, Popper does not identify probability with either degree of belief (likelihood) or frequency of events.
+Instead he suggests 'they should be regarded as _inherent in a situation_ ' [@Popper_1990, 14].
+On this account, probabilities are tendencies towards realization inherent in a situation.
+Probabilities have or indeed are _propensities_, tendencies to realize the event (11).
+While this somewhat abstract philosophical account of probability as propensity might seem to be anchored far away from the concerns of online grocery shopping, Patel's presentation of the personal relevance model and the ways in which it seeks to calculate probabilities of purchase.
+The series of prototyping reconstructions I am about to discuss will explore the tenability of this rather radical alternative to regular narratives of personalization.[^65.1]
 
->they should be regarded as _inherent in a situation_ 14
-
-The conversion of  collaborative filtering into predictive personalisation appeared throughout  Patel's presentations.
-Patel described a move or 'conversion'  from 'a rules-based system'  to  a 'personal relevance model' as the basis of the recommender system.
-Each term in that phrase -- personal, relevance and model -- has some weight in the conversion event.
-Again, dealing with the sometimes telegraphic style of machine learning presentations means making some assumptions.
-It is reasonable to assume that what Patel describes as  the 'rule-based system' refers  to the extremely well-known `apriori` algorithm, developed  by computer scientists Rakesh Agrawal and Ramakrishnan Srikanti working at IBM Research Alameda in the early 1990s [@Agrawal_1994].
-A now-classic approach to 'market basket analysis,' it was listed as one of the top ten data mining algorithms in a survey conducted amongst data miners [@Wu_2008] and usually attracts a chapter in data-mining and machine learning textbooks (e.g.
-[@Hastie_2009]).
-The interest of `apriori` for our purposes is that it relies on massive datasets, but does not personalize its recommendations.
-
-The media theorist Mark Hansen has recently suggested that 'predictive analytics are discoveries of micrological propensities that are not directly correlated with human understanding and affectivity and that do not by themselves cohere into clearly identifiable events' [@Hansen_2015a, 111-112].
+[^65.1]: The media theorist Mark Hansen has recently applied Popper's account to argue  that 'predictive analytics are discoveries of micrological propensities that are not directly correlated with human understanding and affectivity and that do not by themselves cohere into clearly identifiable events' [@Hansen_2015a, 111-112].
 Hansen's account differs from many widely shared views of big data.
-In the face of the so much personalization (and Patel's presentation exemplifies this), Hansen suggests a somewhat processes of impersonalization.
-If  much predictive practice attempts to elicit 'micrological propensities' from data,   Hansen, drawing on the work of the philosopher of science Karl Popper,  links data to things via probabilities:
-
->Whatever explanatory and causal value predictive analytics of large datasets have is, I suggest, ultimately rooted in this ontological transformation whereby probabilities are understood to be expressions of the actual propensity of things [@Hansen_2015a, 120]
-
+In the face of the so much personalization (and Patel's presentation exemplifies this), Hansen attributes a somewhat impersonalizing force to big data.
+If  much predictive practice attempts to elicit 'micrological propensities' from data, Hansen links data to things via probabilities: 'whatever explanatory and causal value predictive analytics of large datasets have is, I suggest, ultimately rooted in this ontological transformation whereby probabilities are understood to be expressions of the actual propensity of things' [@Hansen_2015a, 120].
 Hansen theorises big data or predictive analytics as an 'ontological transformation' that deploys probabilities in an evermore closely woven and encompassing expression of animated, eventful, propensities of things.
 Predictive analytics links probabilities as calculations to propensities or the mutable associative agencies of things.
-The practical and indeed empirical  question is whether such transformations or 'conversion' in relations between probabilities and 'the propensities of things'  can be detected and articulated in the prosaic setting of shopping lists and recommender systems.[^20.1] The series of prototyping reconstructions I am about to discuss will explore the tenability of this rather radical alternative to regular narratives of personalization.
-
-[^20.1]: I will argue that 'conversion' is a preferable term for these changes.
-It highlights  re-orientations in subjects, experience, things, numbers and structures that ranges more widely than Hansen's 'ontological transformation' formulation.
+The practical and indeed empirical  question is whether such transformations or 'conversion' in relations between probabilities and 'the propensities of things'  can be detected and articulated in the prosaic setting of shopping lists and recommender systems.
+'Conversion' is a preferable term for these changes.
 Not much hinges on the choice of terms, but 'conversion' happens to be the term used by Patel.
+It highlights  re-orientations in subjects, experience, things, numbers and infrastructures that ranges more widely and are more grounded than Hansen's 'ontological transformation' formulation.
 
@@
 
-> we need _a calculus of relative or conditional probabilities_ as opposed _a calculus of absolute probabilities_ 16
+Popper suggests that one of the reasons probability has been difficult to work with philosophically, even as it pervades scientific thought, is that a certain limited domain of probability practice focused on dice rolls, coin tosses, urns with balls and other seemingly random events has occupied centre stage in concepts of probability.
+Even if it has been enormously productive and transformative, it is insufficient because it privileges _absolute_ probability at the expense of _conditional_ probabilities.
+'We need,' Popper urges, '_a calculus of relative or conditional probabilities_ as opposed _a calculus of absolute probabilities_' [@Popper_1990, 16].
+Relative or conditional probabilities are propensities that depend on other events for their own realization.
+All events require other events, so all probabilities are conditional, even if probability calculations have typically preferred to determine  the weight of possibilities devoid of the inevitable physical conditioning of their realization.
+
+Viewed from the standpoint of probabilities, many predictive systems are highly crafted arrangements for the calculation of conditional probabilities.
+For instance, the first element of the new recommender system consisted in a change of the underpinning algorithms and model.
+Patel described a move from 'a rules-based system'  to  a 'personal relevance model' as the basis of the new recommender system.
+It is likely that what Patel describes as  the 'rules-based system' refers  to the extremely well-known `apriori` algorithm or association rules learning technique, developed  by computer scientists Rakesh Agrawal and Ramakrishnan Srikanti working at IBM Research Alameda in the early 1990s [@Agrawal_1994].
+A now-classic approach to 'market basket analysis,' it was listed as one of the top ten data mining algorithms in a survey conducted amongst data miners [@Wu_2008] and usually attracts a chapter in data-mining and machine learning textbooks (e.g.
+[@Hastie_2009]).
+The interest of `apriori` for our purposes is that it begins to address the problem of understanding large numbers of shopping transactions as a matter of conditional probability.
 
@@
 
+The notion of conditional probability at work in `aprior` is relatively simple, and assumes that the frequency of transactions provides the best guide to what shoppers are likely to do.
+The `apriori` algorithm finds sets of items that commonly occur together in transactions.
+In this sense, is still oriented by a notion of absolute probability, inflected by some elements of conditional probability.
+Commonly occurring sets are expressed as 'association rules.' For instance, using the `Groceries` dataset in the `R` package `arules,`  the `apriori` algorithm counts frequencies of purchase in the overall set of all items purchased in a supermarket (the `Groceries` dataset  was acquired from a 'local German supermarket' [@Hahsler_2006a]).
+Figure \@ref(fig:arules-ex) shows how often the most common items appear.
+`Whole milk` appears most frequently.
+
 ```{r arules-ex2}
@@
 
-The `apriori` algorithm finds sets of items that commonly occur together in transactions.
-These commonly occurring sets are expressed as 'association rules.' For instance, using the `Groceries` dataset in the `R` package `arules,` we can readily see how Walmart *might* have discovered   the fabled association between diapers and beer.[^149.1] Figure \@ref(fig:arules-ex) shows how often common items appear.
-`Whole milk` appears most frequently.
-The association rule-based model for the `Groceries` dataset illustrates some characteristic ordering practices of 1990s-style data-mining approaches to lists (the `Groceries` dataset  was acquired from a 'local German supermarket' [@Hahsler_2006a]).
-The association between items is expressed in the form of _rules_ not a model.
-These rules, a term derived from older decision support literature, define sets of associations between items.
-Table \@ref(tab:arules-ex2) shows that the first rule has a stronger association than the second.
-Milk and vegetables is more likely an association than milk and buns.
-A rules-based recommender system should, therefore, might put them in that order in the 'Have You Forgotten' list.
+The association rule-based model for the `Groceries` dataset illustrates some characteristic ordering practices of 1990s-style data-mining approaches to lists.
+The association between items is expressed in the form of _rules_ whose importance is can be ranked by how frequently items are found in the same purchase.
+Although the `apriori` model does not use the language of probability calculus, the 'rules,' a term derived from older decision support literature, rank associations between items.
+When we run the `apriori` algorithm on the `Groceries` dataset, the resulting set of rules (shown in Table \@ref(tab:arules-ex2)) indicates that milk and vegetables has a stronger association than milk and buns.
+The calculations of `support` (how frequently the association appears in the dataset), `confidence` (how often the rule applies in the dataset)  and   `lift` (a ratio between the `support` and the independent frequencies of the items) in the table attempt to measure the strength of these associations in different ways.
 
-[^149.1]: One of the founding narratives of big data customer analytics, the beer-diaper association was not, according to [@Power_2002] found through data mining at all.
+## 2500 sauces: `Apriori` meets the API
 
-Although association rules and the `apriori` algorithm present social order in terms of sets of frequency-weighted associations, an `apriori`-based recommender system contains no statistical model.
-It also no grasp on personal relevance since the associations it defines are only between things.
-Put in terms of the big data conversion, the rules-based system is not predictive, or personal.
-The `apriori` algorithm might be seen as an attempt to grasp the excess of things in a Tesco supermarket by identifying subsets whose closure derives from the associations between things that people construct (through buying habits, by virtue of a shopping list, etc.).
-In association rules, personhood or experience is irrelevant.
-The openness of these rules -- the `apriori` algorithm can product any number of rules of diminishing strength -- points towards the propensity of things to find themselves together in shopping baskets, but cannot cannot individuate that propensity beyond demographic quantiles and categories.
-
-## `Apriori` meets the API
-
-Even if `apriori`  elicits associations between things, it struggles with the propensity of commodities to multiply.
-A simple illustration of the combinatorial problem faced by recommender system can be developed using the `Grocery` dataset.
-If we take all the items in the `Grocery` dataset and paste them into the 'shopping list' box on the Tesco grocery website (or as I side, run them as searches on the TescoLab Product API), each of the 169 items in the dataset yields dozens and sometimes thousands of products from Tesco online.
+Even as `apriori`  elicits associations between things, it struggles with the propensity of commodities to multiply.
+A simple illustration of the combinatorial problem faced by recommender system can be developed by bringing the `Grocery` dataset together with the actual list of items that sells in its online grocery shopping site.
+If we take all the items in the `Grocery` dataset and paste them into the 'shopping list' box on the Tesco grocery website (or as I did, run them as searches on the TescoLab Product API [@Tesco_2016]), each of the 169 items in the `Grocery` dataset yields dozens and sometimes thousands of products from Tesco online.
 
@@
     tesco$actual = as.integer(tesco$actual)
-    ggplot(tesco[tesco$actual>50,], aes(x=labels, y=actual, fill=labels)) +geom_bar(stat='identity') + coord_flip() + ggtitle('Tesco grocery items with more than 50 products') + theme(legend.position="none", axis.text=element_text(size=7))
+    ggplot(tesco[tesco$actual>50,],
+        aes(x=labels, y=actual, fill=labels)) +
+        geom_bar(stat='identity') + coord_flip() +
+        ggtitle('Tesco grocery items with more than 50 products') +
+        theme(legend.position="none", axis.text=element_text(size=7))
+
 ```
 
-Items in the `Groceries` dataset proliferate into a Tesco's list of branded variations.
+Items in the `Groceries` dataset proliferate into a Tesco's list of branded products.
 The 169 items of the `Groceries` dataset  expand into roughly `r total` Tesco items (see Figure \@ref(fig:tesco-ex)).
-There is, I would suggest, a logistic propensity in things that recommender systems encounter in the ways of global supply chains.
-The multiple listed  in the `Grocery` dataset becomes more open in this setting through new impersonal intersections that must be bounded and ordered.
-The proliferation of things on the shelves of supermarket or grocery warehouse confronts produces a combinatorial problem for data miners.
+There is, I would suggest, a logistic proliferation that recommender systems encounter.
+The multiple listed  in the `Grocery` dataset becomes more open in this setting through new impersonal intersections that could find themselves in associations of various strengths.
+If we think just of the almost 2500 sauces or 1200 `rice` products supplied by the Tesco API,  then a tremendous number of associations between sauce and rice are possible.
+The propensities of actual sauce and rice products to find themselves together in an individual purchase vary greatly.
+The proliferation of things on the shelves of supermarket or grocery warehouse confronts produces a combinatorial problem for data-mining machine learning approaches such as association rules.
 `r total` products (actually Patel mentioned 200,000 products) can be combined in a vast number of ways.
 If a typical shopping list has 20 items, then there are `r format(choose(total, 20), 2)` possible lists.
-Some of this vast number of possible shopping lists are highly unlikely in terms of typical purchasing patterns, but potentially valuable as recommendations.
+Some of this vast number of possible shopping lists have propensities or tendencies to realization that are close to zero.
+But others with somewhat higher propensities might furnish the basis of interesting recommendations.
+
+More importantly, the limited conditioning of probability of the association rules approach suggests a compelling reason for personalization.
+Even if the association rules and the `apriori` algorithm   generate recommendable sets of frequency-weighted associations, an `apriori`-based recommender system contains no statistical model.
+Its probabilisation of shopping is incomplete since it only works on associations between things.
+The tendency of some milk to find itself in a shopping basket alongside bread attests to an important commonality of practice in certain parts of the world.
+But the conditional probabilities implicit in the model do not bring in much of the world and its teeming sieve of propensities moving towards realization.
+These associations are not trivial, but they are very open-ended.
+Put in terms of the big data conversion, the rules-based system lacks, in principle, a means of crystallizing a limited or enclosed set of possibilities.
 
@@
 
-Conditional probability; the models
-[TBA - section on the role of probability; and logr]
+Predictive models are central technical elements in many big data arrangements because they provide a way of constructing conditional probabilities that incorporate more of the situation in which any propensity -- to purchase a bottle of Vueve Cliquot champagne for instance -- is inherent.
+In the new recommender system described by Patel, the business goal is to extend the list of the items customers have selected for purchase with  a few well-chosen items.
+In order to do this, the list of items selected for purchase will be extended by recommendations that have, according to a predictive model, the most chance of 'conversion'; that is, actual purchase.
+The predictive model carries the burden of calculating the conditional probability of purchase given everything known about the situation at this moment in time.
+
+Patel introduced the new 'personal relevance model' with a data graphic familiar to machine learners and statistical modellers.
+(A sketch of her graph appears in Figure \@ref(fig:models).)  The graph plots the _precision_ -- the proportion of the recommended products that customers actually purchase -- for several different statistical models.
+The graph indexes the 'causal efficacy' of the recommender system, its capacity to include and transform propensities or 'real potentialities' into operational events or purchases.
+The graph compares the previous rule-based recommender systems with some of the relevancy-model alternatives -- logistic regression, random forests, gradient boosting and a few others -- in terms of their predictions and how those predictions turned out.
+Patel dismissed most of the models quite quickly and focused only on one, the logistic regression model, which did as well or slight other than other alternatives.
+
+The models Patel so briefly mentioned are all 'classifiers,' predictive models that 'classify' particular outcomes by calculating their probability.
+Much of the core architecture of the machine learning classifier models received scant reference in Patel's presentation.
+It is most likely that in DunnHumby's work, the classifying model at the heart of the recommender system constructs a decision boundary in vector space between points that belong to different classes of things [@Mackenzie_2015e].
+The classes of things are binary: `recommended` or `not recommended,` and for many of the models shown in the graph, membership of a thing such as grocery to the binary classes are expressed as a probability.
+The logistic regression model generates probabilities for each product for each customer.
+The crucial point here is that the shopping list written by the customer is extended, quasi-invisibly, in the recommender system to potentially any of the 250,000 items in the Tesco inventory.
+
+[HERE]
+
+Patel did not articulate the implication of the contrast between the left and right hand side of the graphic.
+Although the statistical framing appears explicitly (e.g.
+in measuring the precision of  recommender systems), the underlying shift to a statistical model is barely mentioned.
 
@@
 
-------------
-
-In their mundane intersectional practices of ordering, recommender systems show how  predictive practices affect existing forms of social order in the world.
-Rather than directly positing an ontological transformation (as Hansen does), or pointing to the erosion of social order (as Turow does), I have explored some of the conversion events associated with a shift from a data-mining rule-based system to a predictive recommender system.
-The DunnHumby model and the Tesco 'Have You Forgotten?' recommendations present many of the traits of the conversion processes associated with big data.
-Rather than discontinuous changes, the conversion of the recommender system via a relevance model suggests an ongoing problematization of propensities.
-We have seen several types of conversion events: ongoing transformation of models and infrastructures, the experimentalisation  of recommendations, and the broader diffusion of a calculus of probabilities into a fibrous weave of propensities and conditionalities, warping over time.
-The predictive assemblages connect logistical ensembles to the negotiation, compromises, rituals and local orderings of everyday life.
-
-I have been suggesting that predictive listing in the form of open-ended recommendation changes social order.
-It establishes a common coding between machinic and ritual repetition (as exemplified in the shopping list).
-It also suggests some ways in which repetition or reiteration  can also participate in changing and emergent order.
-By understanding the world in terms of propensities, we have a chance  to engage with contemporary predictive processes in ways that both acknowledge their relentless and exhausting pressure to consume more and retain some chance of inhabiting the world differently.
-
-Rather than understanding prediction as the closed form of an ontologically open world ('the propensities of things'), the DunnHumby relevance model, in all its statistical, infrastructural, organisational and banal complexity, suggests that the predictive conversion remains constitutively incomplete.
-Shopping lists in all their banal yet primary orderings are fertile terrain for prediction because they are incomplete, somewhat open and mutable forms.
-The constitutive incompleteness and openness of lists are very closely entangled with ongoing processes of ordering and othering, and at the same time, can accommodate predictive operations.
-Lists undergoing a 'conversion' experience as they become more probabilistic, but their heterogeneous character as operations caught up in models, infrastructures, experiments, and the flow of commodities means that they can never form a closed set or dataset.
-The problem of 'constructing good features,' a problem that any predictive model must address, persists as the site of experiments in partly open structures.
-
-A constitutive openness to new associations, new combinations, new groupings or actors is integral to the conversion events associated with contemporary data practice.
-In their account of the Netflix recommender system, Ted Hallinan and Peter Striphas make the following argument:
-
-> These systems, and more importantly their algorithms, play a critical role in deciding which articles (or parts thereof) gain admission to the cultural realm, and in what form.
-Their doing so thus points in the direction of another universe of reference—a court of algorithmic appeal in which objects, ideas, and practices are heard, cross-examined, and judged independently, in part, of human beings [@Hallinan_2014, 13]
-
-Despite the somewhat troubling reference to 'another universe,' as if Netflix or DunnHumby's recommender systems occupy a different reality, Hallinan and Striphas usefully and aptly highlight how such systems  mingle  objects, ideas and practices.
-
-Something similar happens to shopping in the Dunnhumby relevance model.
-In the customer-product association matrix, in the product-product similarity matrix, in the 14,000 variables included in the logistic regression model, a 'universe of reference' takes shape.
-The shopping list, the commonplace inscription of normative and actual social order, takes on a complex temporal-material structure that mixes prediction, infrastructural constraints, the individual characteristics of list-makers, statistical validations, database architectures and above all the surplus sensibility associated with  circulation and consumption of commodities.
-Shopping lists undergoing a conversion experience as they become more probabilistic, but the idea that probabilities are the closed form of propensities does not account for the distribution of probabilities in relations between things, between people, between people and things, and between things and people.
+# References

### 2017-07-21 log.rmd
got further into section E -- cutting down to model

@@
 - TODO - short paragraph summarising the social theory of list ordering; make sure to add stuff on the ordering-remapping of order in the list;
+
+## Fri 21 Jul 2017 09:22:28 BST
+- did so much work on this yesterday -- today I want it to go more easily; and I don't want to do so much work on it. 
+- link everything in the presentation to the logistic regression model; that's the linchpin -- a massive conditional probability statement
+        - TODO: add a bit about this model ... and think how Popper material helps here 
+- grocery shopping in principle cannot be about personalization because shopping itself isn't. Thats a fundamental problem.
+        - DONE - at least in part: update the theory of shopping to reflect this
+        - TODO: exemplify it in a number of ways
+- although I didn't do as much, paper felt better because I've managed to localize down on the logistic regression model; it helps me combat the personalization problem better. 
+- reached the customer-product matrix;

### 2017-07-21 reconstruction.rmd
got further into section E -- cutting down to model

@@
 
-[TBA: THIS PARA could be moved down] The shelves of large contemporary supermarket are the endpoint of global logistic supply chains, in all their logistical, value-transforming and brand-mediated hypercomplexity (see [@Neilson_2012; @Tsing_2009]).
-Groceries imply a planetary geography of agriculture, industry, transport, communication and financialisation animated by flows of labour and capital.
-Encounters between this hyper-complex commodity-geography and people, even in the familiar confines of a supermarket, are no simple matter, either for shoppers or for supermarket operators such as Tesco.
-
 # C. Shopping  lists and  a theory of social ordering
@@
 The anthropologist  Daniel Miller has argued [@Miller_2012] that  shopping takes place as a negotiation of discrepancies between normative and actual social order (for instance, between ideals of health and commitments to thrift).
-What Miller suggests here is that ...
-[TBA]  It can also be seen  as a sacrificial ritual that re-instates social order in various ways.
+What Miller suggests here is that shopping practices attempt to resolve difference between how people think they should live and how they actually live: 'we have to watch how shopping helps resolve these discrepancies between the normative and the actual, but we also need some ideas of where the normative comes from in the first place' [@Miller_2012, 72].
+Whatever theory we have of how normative social order comes about, the key argument is that shopping inhabits a space between normative/idea and actual order.
+Importantly, and this will be a key consideration for recommender systems, shopping is not particularly personal or individual.
+Grocery shopping in particular intersects forms of social order concerning family and other forms of social grouping.
 
@@
 
-The shopping list, whether written on the back of an envelope, or saved as a list in an online grocery shopping system, lies at the intersection of logistical flows, infrastructural orderings, and  lively negotiations around actual and normative social orders.
-Shopping lists, therefore  are intersectional ordering devices that encapsulates a universe of possible references, and a teeming multitude of propensities with an actual local order (in both senses of that term: an ordering and an imperative).
+The shopping list, whether written on the back of an envelope, or saved as a list in an online grocery shopping system, lies at the intersection of logistical flows, infrastructural orderings, and  lively negotiations around actual and normative social orders.[^30.1]  Shopping lists, therefore  are intersectional ordering devices that encapsulates a universe of possible references, and a teeming multitude of propensities with an actual local order (in both senses of that term: an ordering and an imperative).
 As people shop, either by  trawling along aisles packed with thousands of products, or scrolling down screens or searching for particular brands amidst search results, lists inscribe some kind of order that filters or reduces the excessive propensities, claims, dazzle and distraction of commodities to the practicalities of domestic economy.
+
+[^30.1]: The shelves of large contemporary supermarket are the endpoint of global logistic supply chains, in all their logistical, value-transforming and brand-mediated hypercomplexity (see [@Neilson_2012; @Tsing_2009]).
+Groceries imply a planetary geography of agriculture, industry, transport, communication and financialisation animated by flows of labour and capital.
+Encounters between this hyper-complex commodity-geography and people, even in the familiar confines of a supermarket, are no simple matter, either for shoppers or for supermarket operators such as Tesco.
 
@@
 The graph indexes the 'causal efficacy' of the recommender system, its capacity to include and transform propensities or 'real potentialities' into operational events or purchases.
-The graph compares the previous rule-based recommender systems with some of the relevancy-model alternatives -- logistic regression, random forests, gradient boosting and a few others -- in terms of their predictions and how those predictions turned out.
+Although she did not dwell on it, Patel used the graph to compare the previous rule-based recommender systems with some of the 'personal relevance model' alternatives -- logistic regression, random forests, gradient boosting and a few others -- in terms of their predictions and how those predictions turned out.
 Patel dismissed most of the models quite quickly and focused only on one, the logistic regression model, which did as well or slight other than other alternatives.
 
-The models Patel so briefly mentioned are all 'classifiers,' predictive models that 'classify' particular outcomes by calculating their probability.
+The models Patel so briefly mentioned are all 'classifiers,' predictive models that 'classify' particular events or outcomes by calculating their probability of membership of some class of event or outcome.
 Much of the core architecture of the machine learning classifier models received scant reference in Patel's presentation.
-It is most likely that in DunnHumby's work, the classifying model at the heart of the recommender system constructs a decision boundary in vector space between points that belong to different classes of things [@Mackenzie_2015e].
-The classes of things are binary: `recommended` or `not recommended,` and for many of the models shown in the graph, membership of a thing such as grocery to the binary classes are expressed as a probability.
-The logistic regression model generates probabilities for each product for each customer.
+It is most likely that in DunnHumby's work, the classifier at the heart of the recommender system calculates for a given customer the most likely products to be purchased.
+The classes of events are binary: `recommended` or `not recommended.` Products are allocated to one of these classes depending on their calculated probability.
+A probability greater than `0.5` typically would be `recommended.`  The logistic regression model generates probabilities of purchase for each product for each customer.
+
 The crucial point here is that the shopping list written by the customer is extended, quasi-invisibly, in the recommender system to potentially any of the 250,000 items in the Tesco inventory.
-
-[HERE]
-
-Patel did not articulate the implication of the contrast between the left and right hand side of the graphic.
-Although the statistical framing appears explicitly (e.g.
-in measuring the precision of  recommender systems), the underlying shift to a statistical model is barely mentioned.
+But the basis of this re-writing of the grocery list is an intricate conditional probability statement.
+Even if Patel assumed that the audience understood the working of the logistic regression model, the bulk of her presentation concerned the obstacles and problems that arise in trying to probabilise recommendations in ways that lead to the much-desired 'conversion events' or sales.
 
@@
 
-52 weeks of data
-> in many kinds of events this is not the case, and the propensities cannot be measured because the relevant situation changes and cannot be repeated 17
+I will not discuss in detail the different dimensions of probabilisation implicit in the architecture of the new model.
+It is worthwhile, however, highlighting some key elements that point directly to the troubled notion of 'personal relevance' and personalization more generally.
+These include the problem of repetition, attempts to create new views of the data, experiments on the platform, and the temporalities of calculation.
+
+One fundamental difficulty here is that the propensity of a customer to buy a recommended product at a particular point in time is changing.
+Popper's notion of probabilities as propensities as tendencies to realization inherent in a situation directly implies this.
+Measuring propensities and expressing these measurements as numbers between `0` and `1` becomes difficult because the situation is change and the propensities themselves change.
+As Popper observes, 'the propensities cannot be measured because the relevant situation changes and cannot be repeated' [@Popper_1990, 17].
+No doubt the contingencies associated with grocery shopping are legion, perhaps more so than other kinds of products (books, music, films),  and any attempt to measure propensities associated with particular commodities  will encounter many dynamic changes.
+
+One way in which recommender systems attempt to address the problem of the relational dynamics of propensities is by looking for repetitions.
+Patel mentioned that the new model uses  '52 weeks of data' for each customer.
+In including this data in the model, the assumption is that the probability that a specific customer will buy a specific product will be shaped by previous purchases.
+The history of previous purchases constitutes forms of repetition that imply a stabilisation of propensities (high or low; e.g.
+a vegan customer will never have purchased chicken products, so the measured propensities for any of the 1000 or so chicken products sold be Tesco will remain close to `0`).
+
+When the logistic regression model includes the 52 weeks of previous purchases, the conditional probability calculation undertaken by the recommender system ramifies tremendously in several respects.
+Each of Tesco's 200,000 products becomes a variable in the classifier.
+Practically, most of the these variables will not influence the calculation of the probabilities for recommendations very much.
+As Patel observed,  again assuming that this would be obvious to the audience of data scientists,  'we have lots of zeros'.
+-- points to the difficult terrain of  'good feature' thinking: the vast, nearly empty matrix of customer-product associations.
+Given that most things in Tesco remain relatively unknown to each other and to customers.
+A matrix that records associations between individual people and different products is bound to be mostly empty.
+Say Tesco has 1 million online customers.
+Each online  shopper has bought some selection of the 200,000 products.
+The customer-product data matrix will be `r format(1e6 * 2e5)` in size.
+The product-customer matrix, the basic vector-space in which all recommender systems operate, remains very sparse and unpopulated.
+Given that any one customer is likely to only have bought 100 or so different products, the probability matrix will be close to `r format(100-(100/2e5)*100, 2)`% empty.
+The data is, as Patel put it, 'massively unbalanced' (any dataset where the items of interest are much rarer than some normal values is said to be unbalanced) and such imbalance would heavily bias the recommender system towards common and only impersonally relevant suggestions, suggestions that might  not produce the desired conversion experience.
+
+HERE
+
+```{r sparsity-ex, echo=FALSE, cache=TRUE, fig.margin=FALSE, fig.cap ='Sparsity of the customer-product association matrix' }
+library(ggplot2)
+library(reshape2)
+p = as.logical(rbinom(n=2e5, size=20,0.001 ))
+m = (matrix(p, nrow=1000, byrow=TRUE))
+melted <- melt(m)
+melted$value <- as.logical(melted$value)
+ggplot(melted, aes(x = Var2, y = Var1, fill = value)) + geom_tile() +
+    scale_fill_manual(values = c("grey", "red")) + xlab('Customer id') + ylab('Product id') + theme(legend.position='none')
+```
+
+Since so many people buy milk, the recommendation system might end up always recommending milk products.
+So the data needs to be 'corrected' by, as Patel reports, removing -- under-sampling -- some of the data for common purchases.
+Secondly, even if a recommendation system finds uncommon items that are good recommendations, they should neither be too cheap ('low spend'), nor, as we will see too similar to what the customer has already ordered.
 

### 2017-07-26 reconstruction.rmd
starting a/b test

@@
 
+In his account of probabilities as physical propensities, Popper emphasises the specific configurations established in laboratory experiments.
+
 > experiments. ... by creating, at will, artificial conditions that either exclude, or reduce to zero, all the interfering and disturbing propensities 23
+
+|Test A|Test B|
+|------|------|
+|Control A|Control B|
+|------|------|
+
+For instance, the predictions of the recommender system themselves are the subject of probabilistic treatment.
+Patel described the experimental deployment of  the relevancy model experimentally in a random A/B controlled trial on the Tesco website.
+All customers were  allocated to one of four categories as shown in the table.
+In the  A/B trial,  an identical shopping list, or even the same person, might receive different recommendations.
+The randomised application of the model, drawing on protocols for randomised clinical trials first developed in the 1960s, is widely used in social media platforms and hence in the implementation and observation of the effects of recommender systems.[^238.1] Random allocation of customers to the four categories (Test A, Test B, Control A, Control B) adds a layer of probability to the recommender system in the name of statistical validation of the effects or the 'uplift' of the model.[^122] Ironically,  the effects of a predictive model cannot be known in advance.
+They can only be observed experimentally.
+
+[^238.1]: It is difficult to gauge how much. Facebook reports that it has more than 1 million models operating in its infrastructure. [@Dunn_2016]
+
+It is perhaps significant that this random allocation of customers to control and test groups occurs without any relation to the particular profile or propensities of the customer.
+It seeks to statistically validate the effects -- the *uplift* -- of the model on  conversion events by allowing the effects of the model on what people do to be measured.
+The uplift refers to conversion events associated with the same groups of people.
+Effectively an experiment in inhabiting two different worlds, the randomised control trial  sets up a feedback look between the predictive model (the logistic regression), and the world.
+
+[^122]: See [@Lo_2002] for an introduction to the construction of experimental set ups for uplift models.
+Market researchers and data miners seem to have started adopting a statistically grounded experimental predictive practice sometimes around 2000.
+See for instance [@Almquist_2001]  Response modelling, 'propensity' modelling and uplift modelling all seek to identify associations between 'treatments' or interventions and the 'Responders', the people  affected by the treatment.
 
@@
 
-The 'minute and extensive scrutinizing' [@Dewey_1957] of a recommender system undergoing a conversion from an older data-mining approach to a current 'personal relevance model' approach addresses a specific situation: online grocery shopping (not shopping in general) and the ordering of lists.
+The 'minute and extensive scrutinizing' [@Dewey_1957] of a recommender system undergoing a conversion from an older data-mining approach to a current 'personal relevance model' approach addresses a specific situation: online grocery shopping (not shopping in general) and, even more micro-sociologically, the ordering of lists.
 
 The theory of shopping lists constructed at the beginning of the paper was a scaffolding component.
-It draws on anthropology, sociology and mathematics to suggest that we see shopping lists as the site of a complex negotiation between different social orders.
+It draws on anthropology, sociology and mathematics to suggest that we see shopping lists as the site of a complex negotiation between different social orders, between norms and actualities.
 The list itself as an inscriptive device functions as an ordering device, a device that accommodates an immense variety of specific habits, dispositions, values and associations.
-The mapping component of the list, the order that arranges the set of items according to different interests and aims, is the operational focus of recommender systems.
+The mapping element of the list, the order that arranges the set of items according to different interests, values, and aims, is the operational focus of recommender systems.
 

### 2017-07-26 log.rmd
starting a/b test
trying to realize themselves

@@
 - reached the customer-product matrix;
+
+## Wed 26 Jul 2017 12:06:20 BST
+- at the heart of the paper is the contrast between the association rules and the personal relevance model. The differences between them aren't as great as might seem to be ...   
+- where I finished last time -- line 153 -- there is still a gap.
+I didn't get to say everything I wanted to about the logistic regression model;  I wanted to add some stuff about how the model includes all the different things
+- starting adding the probability section and adding in a bit of the Popper argument  
+- got to the point of trying to say what 'trying to realize themselves' might mean. I intend to connect this to both 'trying to construct good features.'

### 2017-07-26 reconstruction.rmd
trying to realize themselves

@@
 
-This paper is a philosophical reconstruction [@Dewey_2017] of an online grocery ordering system undergoing an upgrade.
+This paper philosophically reconstructs [@Dewey_2017] an online grocery ordering system undergoing an upgrade.
 The reconstruction is empirical in certain respects.
@@
 It makes use of archaeological approaches, in the sense developed by Michel Foucault [@Foucault_1972] in order to identify and map important functional statements and practices configuring the knowledge and order generated by such systems over time.
-It conducts several code-based experiments in order to reconstruct, using widely available code resources resources such as API (Application Programmer Interfaces) and software libraries for machine learning, some prototypical elements of the system in question.
-In doing this, it draws  on the philosophy of science and media theory to frame an engagement with data, algorithms and recommendations around probabilities and their worldly enactments.
+It conducts several small-scale code-based experiments in order to reconstruct, using widely available code resources resources such as API (Application Programmer Interfaces) and software libraries for machine learning, some prototypical elements of the system in question.
+In doing this, it draws  on the philosophy of science and media theory to frame an engagement with data, algorithms and recommendations in terms of probabilities and their worldly enactments.
 
 Like some recent work in science and technology studies,  anthropology and media studies [@Marres_2017; @Marcus_2014; @Bogost_2012], through this combination of approaches,  the paper allows its own reconstruction to be provisionally shaped by an experiential encounter with the object it describes.
-At times, this mode of empirical philosophical engagement with its objects may well seem overly concerned with the technical minutiae of the grocery ordering system, and lacking in  sociological rigour or empirical evidence.
+At times, this mode of empirical philosophical engagement with its objects may well seem overly concerned with the technical minutiae of the groceries and recommender systems, and lacking in  sociological rigour or empirical evidence.
 With some forbearance on the part of readers, however, the main philosophical argument of the paper should resonate sociologically.
@@
 
-HERE
-
 ```{r sparsity-ex, echo=FALSE, cache=TRUE, fig.margin=FALSE, fig.cap ='Sparsity of the customer-product association matrix' }
@@
 
+Many other slight adjustments, tweaks and local modifications would probably occur around the logistic regression model.
+The model's recommendations themselves have propensities that need to be adjusted to the scale of values of grocery retail.
+[FEELS LIKE A JUMP HERE:  see notes 26 July 2017]
+
 ## Platform Experiments that reduce interfering propensities
 
-In his account of probabilities as physical propensities, Popper emphasises the specific configurations established in laboratory experiments.
+In his account of probabilities as physical propensities, Popper emphasises the specific configurations established in laboratory experiments:
 
-> experiments. ... by creating, at will, artificial conditions that either exclude, or reduce to zero, all the interfering and disturbing propensities 23
+> experiments work ... by creating, at will, artificial conditions that either exclude, or reduce to zero, all the interfering and disturbing propensities [@Popper_1990,23].
+
+Like many accounts of laboratory practice [@Latour_1097; @Rheinberger, @Knorr-Cetina; @Lynch - epistopics ref], Popper understands experiments as stabilizing the conditions so that phenomena can be saved from their own variability.
+Experiments play an important role in conversion events, or in the process of bring new predictive models to bear on shopping practice.
+A  crucial consideration for a predictive model in a recommender system will be how to experimentalize its own predictions so that they are not interfered with or disturbed by other propensities.
+This is obviously an inordinately difficult challenge, since, as we have seen, grocery shopping takes place at the intersection of normative and actual social orders, individual and group belongings, and logistically global economic processes.
 
@@
 
-For instance, the predictions of the recommender system themselves are the subject of probabilistic treatment.
-Patel described the experimental deployment of  the relevancy model experimentally in a random A/B controlled trial on the Tesco website.
+Nonetheless, experimentalization, the practice of creating conditions that reduce disturbing propensities, is very much part of the convernsion between models.
+For instance, the predictions of the recommender system themselves are the subject of experimental.
+Patel described the deployment of  the relevancy model experimentally in a random A/B controlled trial on the Tesco website.
 All customers were  allocated to one of four categories as shown in the table.
-In the  A/B trial,  an identical shopping list, or even the same person, might receive different recommendations.
-The randomised application of the model, drawing on protocols for randomised clinical trials first developed in the 1960s, is widely used in social media platforms and hence in the implementation and observation of the effects of recommender systems.[^238.1] Random allocation of customers to the four categories (Test A, Test B, Control A, Control B) adds a layer of probability to the recommender system in the name of statistical validation of the effects or the 'uplift' of the model.[^122] Ironically,  the effects of a predictive model cannot be known in advance.
+In the  A/B trial,  an identical shopping list, or even the same person, might receive recommendations from different models.
+The randomised application of  competing predictive models, draws on protocols for randomised clinical trials first developed in the 1960's, and is widely used in social media platforms and hence in the implementation and observation of the effects of recommender systems.[^238.1] Random allocation of customers to the four categories (Test A, Test B, Control A, Control B) adds a layer of probability to the recommender system in the name of statistical validation of the effects or the 'uplift' of the model.[^122] Ironically,  the effects of a predictive model cannot be known in advance.
 They can only be observed experimentally.
@@
 
-> What may happen in the future .. is, to some extent, open. There are many possibilities trying to realize themselves, but few of them have a very high propensity, given the initial conditions. 22
+Whatever disturbing factors or interferences the experimental trials of different predictive models reduce, a broader and ineluctable stream of propensities remains in operation.
+Popper's account of probabilities as propensities presents openness as concomitant to realization:
+
+>What may happen in the future ..
+is, to some extent, open.
+There are many possibilities trying to realize themselves, but few of them have a very high propensity, given the initial conditions [@Popper_1990, 22].
+
+There are two ways of reading this kind of statement.
+One is to see Popper as simply stating one of the consequencies of his physicalist understanding of probabilities as tendencies inherent in a situation.
+Another is to focus on the pivotal phrase 'possibilities trying to realize themselves' by asking what 'trying to realize' might mean.
+I would suggest that Patel's description of DunnHumby's work on the new personal relevance model is just such a 'trying to realize' in important ways.
+
+HERE
 

### 2017-07-27 log.rmd
self-learning substitute

@@
 - got to the point of trying to say what 'trying to realize themselves' might mean. I intend to connect this to both 'trying to construct good features.'
+
+## Thu 27 Jul 2017 15:57:03 BST
+- late in the day to be starting this -- just wanted to get the one point in about building 'good feature.'  
+- brought in the self-learning substitutes discussion and the complementarities
+- TODO: decide whether the hadoop and the weekly model running, etc should go in

### 2017-07-27 reconstruction.rmd
self-learning substitute

@@
 
+Several times, Patel emphasized the importance of 'constructing good features' from the data, and much of her presentation concerned different attempts to construct 'good features.' A 'feature' in the context of machine learning refers to a variable included in a predictive model.
+A 'good feature' is one that somehow improves the accuracy, precision, specificity or any of the panoply of measures applied to machine learning models in practice.
+I find it interesting that the main efforts that DunnHumby made to  construct good features were not closely associated with personhood, but with the relations -- the propensities -- between things.
+
+Patel, for instance, described  the problem of 'basket similarity,' the problem that the recommender system might recommend something similar to items already in a customer's basket.
+A customer might be willing to substitute a similar item for something they have already chosen, but they are more likely to accept a recommendation for an item that complements rather than duplicates items in their list.
+
+How could the recommender system sense the possible substitutions and complementaries between things for a given customer, especially since a person's sensibilities and susceptibilities concerning groceries will be subtly shaped by wider social groups, orderings and circumstances?
+It could only do that by also modelling the relation between recommendations and items already ordered to construct a new feature from the data concerning how the substitutabilities and complementarities between product.
+Taking all the baskets of items purchased, rather in the way that association rules does, a new feature, what DunnyHumby termed 'self-learning substitutes' was derived.
+The 'self learning substitutes' feature of the relevancy model draws on another data matrix, the product similarity matrix, itself generated from all previous recommendations that have led to conversion events (that is, 'a user clicked as the response').
+A customer's recommendation list might already include `Tesco Ginger Nuts 300g.`   The 'self-learning substitution model' seeks to replace list items with similar but higher value items: `Mcvitie's Ginger Moments 400g.`
+
+This new feature is complex.
+Patel mentioned,  for instance, that it took the form of a 'design matrix of 14,000 columns.' This means that 14,000 new explanatory variables were added to the logistic regression model producing the recommendations.
+Each recommendation would be subtly re-weighted by this complex derived feature.
+The complexity of the model itself introduces new instabilities, new 'possibilities trying to realize themselves' and the DunnyHumby data scientists used  '`L1 regularization` to drive coefficients down to zero', or, put more simply, to exclude variables there were not contributing much to the predictions produced by the model.
+
+Even this added feature, a complex addition in many respects, does not exhaust DunnHumby's attempts to re-shape the recommender system.
+They were just beginning, Patel reported, to address complementarities between things.
+`Nutella` complements `bananas,` but the converse is not necessarily true since `bananas` need not complement `nutella,` especially if I have nut allergies.
+From the standpoint of customer data science, Patel emphasised the operational significance of predicting complementarities:  'complementarity drives conversion.' Bringing the matrix of similarities expressed as numerical associations between products into the model improves the conversion of recommendations into purchases.
+
 HERE

### 2017-07-28 log.rmd
adding file to be printed

--- 2017-07-26 log.rmd
@@
 - starting revisions by going through ref2; 
-- philosophico-methodological preramble: TODO: put this as opening paragraph
+- philosophico-methodological preramble: DONE: put this as opening paragraph
 
@@
 - new title: Probabilities in the data analytic conversion of social order
-- TODO: play with the structuring of Section E.
+- DONE: play with the structuring of Section E.
 
@@
     - see how that model can recommend substitutions or complementarities; these are 'new possibilities'   
-- TODO: move the personalization discussion higher up, perhaps in the introduction
+- DONE: move the personalization discussion higher up, perhaps in the introduction
 - rough reconstruction is ~4.2k, but haven't brought in the experiments much
@@
 - got to the point of trying to say what 'trying to realize themselves' might mean. I intend to connect this to both 'trying to construct good features.'
+
+## Thu 27 Jul 2017 15:57:03 BST
+- late in the day to be starting this -- just wanted to get the one point in about building 'good feature.'  
+- brought in the self-learning substitutes discussion and the complementarities
+- DONE: decide whether the hadoop and the weekly model running, etc should go in
+
+## Fri 28 Jul 2017 12:18:30 BST
+- adding the hadoop stuff as a footnote to para on local mods.  
+- moved the personalization stuff into the intro
+- starting to rewrite the abstract; finished a version of it. 
+- DONE: go through and fix subheadings
+- TODO: go through TODOs
+- TODO: print

### 2017-07-28 reconstruction.rmd
adding file to be printed

@@
-# Probabilities in the data analytic conversion of social order
+---
+title: "Personalization and probabilities: the expression of propensities in online grocery shopping"
+author: Adrian Mackenzie^[Sociology, Lancaster University]
+date: "`r format(Sys.time(), '%B %Y')`"
+fontsize: 12pt
+citation_package: biber
+latex_engine: xelatex
+bibliography: /home/mackenza/ref_bibs/uni.bib
+csl: "sage-harvard.csl"
+link-citations: yes
+header-includes:
+    - \usepackage{setspace}
+    - \doublespacing
+abstract:
+    "Many accounts of big data systems assume that they target individuals.
+Personalization, with all the risks of discrimination and bias it entails, has been the focus of inquiry.
+This paper suggests that personalization occurs against a more expansive background that in principle is not reducible to individual attributes and desires, but includes many different forms of relations in flux.
+It describes the 'personalization' of an online grocery shopping recommender system.
+It reconstructs an account of how a recommender system constructs an ordered list of a small number of grocery items of personal relevance for each of the millions of online grocery shoppers in  a major UK supermarket chain.
+Drawing on a theory of probability proposed by the philsopher of science Karl Popper, it suggests that personal relevance is inevitably bound up with relations to others, and the relations between things.
+Using a mixture of discourse analysis and code-based reconstruction of key elements of the recommender system, it suggests that personalization is one provisional stabilization of a much more open-ended weave of propensities associated with people and things in contemporary big data configurations.
+The paper explores how in the context of recommender systems the constitutive incompleteness of shopping lists, their propensity to expand or change, becomes more important than their capacity to be personalized."
+---
 
-# A. Introduction  includes a,b,c from above
+```{r setup2, cache=FALSE, include=FALSE}
+library(knitr)
+library(tint)
+knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tint'))
+options(htmltools.dir.version = FALSE)
+output <- opts_knit$get("rmarkdown.pandoc.to")
+opts_chunk$set(warning=FALSE)
+opts_chunk$set(message=FALSE)
+opts_chunk$set(echo=FALSE)
+opts_chunk$set(error=FALSE)
+opts_chunk$set(cache=TRUE)
+```
 
-This paper philosophically reconstructs [@Dewey_2017] an online grocery ordering system undergoing an upgrade.
+This paper philosophically reconstructs [@Dewey_2017] an online grocery ordering system undergoing an upgrade from a demographically-based recommendation system to a 'personal relevance' model.
 The reconstruction is empirical in certain respects.
@@
 Starting from this highly mediated and fragmentary encounter with a recommender system and one of its designers, the paper explores the problem the system addresses.
-It draws on anthropological and sociological research concerning shopping and list-making to situate the recommender system amongst the social ordering practices of shopping and shopping lists.
+It draws on anthropological and sociological research concerning shopping and list-making to situate the recommender system amongst the social ordering practices of shopping and lists.
 It makes use of archaeological approaches, in the sense developed by Michel Foucault [@Foucault_1972] in order to identify and map important functional statements and practices configuring the knowledge and order generated by such systems over time.
 It conducts several small-scale code-based experiments in order to reconstruct, using widely available code resources resources such as API (Application Programmer Interfaces) and software libraries for machine learning, some prototypical elements of the system in question.
-In doing this, it draws  on the philosophy of science and media theory to frame an engagement with data, algorithms and recommendations in terms of probabilities and their worldly enactments.
+Throughout, it draws  on the philosophy of science and media theory to frame an engagement with data, algorithms and recommendations in terms of probabilities and their worldly enactments.
 
@@
 The main argument of the paper concerns how  probability operates in Big Data situations.
-While calculations of probability have  long-standing importance in many settings (insurance, medicine, experimental and field sciences, operations research, risk analysis, engineering design, public health,  economic modelling, etc.), probabilities have taken on a much thicker situational texture.
+While calculations of probability have  long-standing importance in many settings (insurance, medicine, experimental and field sciences, operations research, risk analysis, engineering design, public health,  economic modelling, etc.), probabilities have taken on a much thicker situational texture on digital platforms associated with the web and social media.
 In recent research and opinion on data practice, personalization has been seen as the linchpin of predictive modelling.
-The reconstruction of the recommender system suggests a different way of seeing these transformations.
+Big data discourse in its promissory mode  attributes potency to personalization: 'most important, using big data we hope to identify specific individuals rather than groups' [@Mayer-Schonberger_2013].
+Similarly, Joseph Turow and co-authors conclude in their critical account of the transformation of retail space: 'through it all, knowingly and not, and away from the spotlights of fierce social debate, retailers are encouraging daily routines that accept data-driven personalization as a centrifugal public' [@Turow_2015, 476].
+Analysis of internet filter bubbles and the growth of predictive platforms in industry and government make similar points [@Pariser_2011; @ONeil_2016].
+
+The reconstruction of the recommender system suggests a different way of seeing predictive personalization by expanding the key underpinning notion of probability.
+Probabilities in practice -- in this case, in recommender system -- diverge from probability theory and indeed from statistics as a science.
 Operational probabilities differ from epistemic probabilities in their position in the world, as part of infrastructures, transactions, and event-structures.
-Taking their complex siting into account, and accepting their _physical_ reality, we might come to understand the world operative probabilities belong to as what the philosopher of science Karl Popper understood as a 'world of propensities, as an unfolding process of realizing possibilities and of unfolding new possibilities' [@Popper_1990,19].
-Given this understanding of operational probability, our understanding of what is enacted in contemporary data analytics, especially in association with platforms [@Gillespie_2010] where the operationality of probabilities are heightened,  might shift.
+Taking their complex siting into account, and accepting their _technical_ reality, we might come to understand operative probabilities along the lines of the philosopher of science Karl Popper understood as a 'world of propensities, as an unfolding process of realizing possibilities and of unfolding new possibilities' [@Popper_1990,19].
+Given this understanding of operational probability, our understanding of what is enacted in contemporary data analytics, especially in association with platforms [@Gillespie_2010] where the operationality of probabilities are heightened,  might shift focus away from personalization.
 
-The broad implication of the argument concerns how we think about social order or regularities amidst change or transformation.
+The broad implication of this expansion  concerns how we approach social order or regularities amidst change or transformation.
 Proponents and critics of Big Data present the advent of large volumes of data and analytical techniques (machine learning in particular) as transforming for better or worse organisations, institutions, economic processes and social life.
-Personalization has been a pivotal element, they argue, of this transformation.
+Personalization has been the pivotal element, they argue, of this transformation.
 I suggest too that a transformation or conversion event is occurring, but one which might also, even in the midst of much ongoing personalization,  open up the possibility of experiencing the world differently, and conceiving of social ordering processes somewhat afresh.
-Even if the argument is centred on shopping, the analysis of how a world of propensities continuously and unevenly converts into realizations applies across the board, from the most physical to the most ideal or abstract settings.
-The average everydayness of the example here -- grocery shopping -- allows some first hand engagement with the messiness, the entanglements, and potentials of this conversion event in ways that other interesting settings -- particle physics for instance -- do not.
+Even if the argument is centred on grocery shopping, the analysis of how a world of propensities continuously and unevenly converts into realizations applies across the board, from the most physical to the most ideal or abstract settings.
+The average everydayness of the example here -- grocery shopping -- allows some first hand engagement with the messiness, the entanglements, and potentials of this conversion event in ways that other interesting settings -- social media for instance -- do not.
 
-# B. Patel section + opening description
+# B. The conversion from demographic to personalized recommendation
 
@@
 
-# C. Shopping  lists and  a theory of social ordering
+# C. Shopping  and  a theory of list orderings
 
-The anthropologist  Daniel Miller has argued [@Miller_2012] that  shopping takes place as a negotiation of discrepancies between normative and actual social order (for instance, between ideals of health and commitments to thrift).
+The anthropologist  Daniel Miller has argued [@Miller_2012] that  shopping takes place as a negotiation of discrepancies between normative and actual social order (for instance, between the ideals of health emblematised by organic products and commitments to thrift embodied in generic products).
+More provocatively, Miller claims that 'shopping is largely a technology for the expression of love' (85).
 What Miller suggests here is that shopping practices attempt to resolve difference between how people think they should live and how they actually live: 'we have to watch how shopping helps resolve these discrepancies between the normative and the actual, but we also need some ideas of where the normative comes from in the first place' [@Miller_2012, 72].
@@
 Importantly, and this will be a key consideration for recommender systems, shopping is not particularly personal or individual.
-Grocery shopping in particular intersects forms of social order concerning family and other forms of social grouping.
+Grocery shopping in particular is woven through with forms of social order concerning family and other forms of social grouping, and their associated relations (love, etc.).
 
-From an ethnomethodological standpoint, shopping lists provide important clues to how local social order is constituted, maintained, and repaired.
-Social ordering is done in shopping, and the shopping list as an inscriptive device stabilises certain forms of ordering at the expense of other possible orderings.
-
+Shopping lists provide important clues to how local social order is constituted, maintained, and repaired in shopping.
+If social ordering is done in shopping,  then shopping list as an inscriptive device stabilises certain forms of ordering at the expense of other possible orderings.
 The shopping list, whether written on the back of an envelope, or saved as a list in an online grocery shopping system, lies at the intersection of logistical flows, infrastructural orderings, and  lively negotiations around actual and normative social orders.[^30.1]  Shopping lists, therefore  are intersectional ordering devices that encapsulates a universe of possible references, and a teeming multitude of propensities with an actual local order (in both senses of that term: an ordering and an imperative).
@@
 
-# D. Archaeology of recommendations  from 1984  to 2007
+# D. Archaeology of recommendations:  from 1984  to 2007
 
@@
 
-## Apriori as conditional probabilities
+## `Apriori` and conditional probabilities
 
@@
 Many other slight adjustments, tweaks and local modifications would probably occur around the logistic regression model.
-The model's recommendations themselves have propensities that need to be adjusted to the scale of values of grocery retail.
-[FEELS LIKE A JUMP HERE:  see notes 26 July 2017]
+The model's recommendations themselves have propensities that need to be adjusted to the scale of values of grocery retail, to the available infrastructures, and to the capacities of the online grocery system to merge recommendations into a grocery order in a timely fashion.
+While raw data from Tesco Online feeds into the model every hour, a recommendation list for each customer is only generated once a week.
+Customers only shop online every few days at the most, and in some cases, only every few weeks.
+To continually update the top 200 recommendations for every customer would be computationally expensive.
+Patel briefly mentioned  specific infrastructural elements such as `hadoop` [@ApacheSoftwareFoundation_2009].
+The logistics of running a predictive model affect listing practices.
+The possibility of adjusting the recommendations for every customer weekly derives from an infrastructure capable of collecting data, assimilating that data to a predictive model able to generate approximately 200 million relevant recommendations, where 'relevance' depends on a matrix of probabilities of associations between people and things that shifts in time.
+`Hadoop` and its legion of 'big data' variants (`mahout`, `spark`, `hive`, `pig`, `yarn`) operationalise flows of data at an infrastructural rather than analytic scale.
+In many respects, the quickly glossed-over infrastructural deployment of DunnHumby's relevancy model is the primary conversion event: the logistic regression model at the heart of the recommender system is no longer an analytical device but an operational one.
 
@@
 
-There are two ways of reading this kind of statement.
+There are two ways of reading this statement.
 One is to see Popper as simply stating one of the consequencies of his physicalist understanding of probabilities as tendencies inherent in a situation.
@@
 A 'good feature' is one that somehow improves the accuracy, precision, specificity or any of the panoply of measures applied to machine learning models in practice.
+The precise meaning of 'good', however, remains open to many different attempts, experiments and variations.
 I find it interesting that the main efforts that DunnHumby made to  construct good features were not closely associated with personhood, but with the relations -- the propensities -- between things.
 
-Patel, for instance, described  the problem of 'basket similarity,' the problem that the recommender system might recommend something similar to items already in a customer's basket.
+Patel, for instance, addressed the problem of 'basket similarity,' the problem that the recommender system might recommend something similar to items already in a customer's basket.
 A customer might be willing to substitute a similar item for something they have already chosen, but they are more likely to accept a recommendation for an item that complements rather than duplicates items in their list.
-
 How could the recommender system sense the possible substitutions and complementaries between things for a given customer, especially since a person's sensibilities and susceptibilities concerning groceries will be subtly shaped by wider social groups, orderings and circumstances?
-It could only do that by also modelling the relation between recommendations and items already ordered to construct a new feature from the data concerning how the substitutabilities and complementarities between product.
+It could only do that if had some sense of the relation between items already in the list and the field of possible recommendations.
+The 'personal relevance' model, therefore, sought to include the relation between recommendations and items already ordered.
+The data scientists constructed a new feature from the data concerning how the substitutabilities and complementarities between product.
 Taking all the baskets of items purchased, rather in the way that association rules does, a new feature, what DunnyHumby termed 'self-learning substitutes' was derived.
@@
 This new feature is complex.
-Patel mentioned,  for instance, that it took the form of a 'design matrix of 14,000 columns.' This means that 14,000 new explanatory variables were added to the logistic regression model producing the recommendations.
+Patel mentioned that it took the form of a 'design matrix of 14,000 columns.' This means that 14,000 new explanatory variables were added to the logistic regression model producing the recommendations.
 Each recommendation would be subtly re-weighted by this complex derived feature.
@@
 From the standpoint of customer data science, Patel emphasised the operational significance of predicting complementarities:  'complementarity drives conversion.' Bringing the matrix of similarities expressed as numerical associations between products into the model improves the conversion of recommendations into purchases.
-
-HERE
 
@@
 Only because shopping is already somewhat ordered by lists can a partial re-mapping, provisionally conducive to new events, take place.
-The conversion entailed in this re-mapping is not just about perosnalization.
+The conversion entailed in this re-mapping is not just about personalization.
 It is more ontological than that since it includes relations running multi-laterally between people and things in time.

### 2017-07-28 render_tufte.r

@@ -4,6 +4,6 @@ library(tint)
 library(tufte)
 #render('lists_conversion_experience.rmd',output_format=tufte_handout())
 #render('lists_conversion_experience.rmd',output_format=ioslides_presentation)
-#render('lists_conversion_experience.rmd',output_format=tintPdf())
-render('lists_conversion_experience.rmd',output_format=word_document2())
+render('reconstruction.rmd',output_format=tintPdf())
+render('reconstruction.rmd',output_format=word_document2())
 

### 2017-07-31 log.rmd
lu print

@@
 - TODO: go through TODOs
-- TODO: print
+- DONE: print
+
+## Mon 31 Jul 2017 11:47:42 BST
+- printing as docx.
+There are around 35 pages on manuscript.
+Plan for revision: 7 pages/day for one week; (7 hours)  go through referee reports again alongside the manuscript to put them in (2 hours); redo diagrams and references (3 hours), submission (1 hour).
+Total of about 12 hours.
+Might be to do about 6-7 hours this week.

### 2017-07-31 reconstruction.rmd
updating slides

@@
 Similarly, Joseph Turow and co-authors conclude in their critical account of the transformation of retail space: 'through it all, knowingly and not, and away from the spotlights of fierce social debate, retailers are encouraging daily routines that accept data-driven personalization as a centrifugal public' [@Turow_2015, 476].
-Analysis of internet filter bubbles and the growth of predictive platforms in industry and government make similar points [@Pariser_2011; @ONeil_2016].
+Analysis of internet filter bubbles and the growth of predictive platforms in industry and government make similar arguments [@Pariser_2011; @ONeil_2016a].
 

### 2017-08-01 reconstruction.rmd
entered 13/51 edits

@@
 ---
-title: "Personalization and probabilities: the expression of propensities in online grocery shopping"
+title: "Personalization and probabilities: impersonal propensities in online grocery shopping"
 author: Adrian Mackenzie^[Sociology, Lancaster University]
@@
 
-This paper philosophically reconstructs [@Dewey_2017] an online grocery ordering system undergoing an upgrade from a demographically-based recommendation system to a 'personal relevance' model.
+This paper philosophically reconstructs [@Dewey_1957] an online grocery ordering system undergoing an upgrade from a demographically-based recommendation system to a 'personal relevance' model.
 The reconstruction is empirical in certain respects.
 It departs from an ethnographic moment: being a member of the audience at an industry/academic conference where a data scientist was describing how recommendations were created for online grocery orders.
-Starting from this highly mediated and fragmentary encounter with a recommender system and one of its designers, the paper explores the problem the system addresses.
+Starting from this highly mediated and fragmentary encounter with a recommender system and one of its designers, the paper explores the problem of personally relevant recommendations.
 It draws on anthropological and sociological research concerning shopping and list-making to situate the recommender system amongst the social ordering practices of shopping and lists.
@@
 
-Like some recent work in science and technology studies,  anthropology and media studies [@Marres_2017; @Marcus_2014; @Bogost_2012], through this combination of approaches,  the paper allows its own reconstruction to be provisionally shaped by an experiential encounter with the object it describes.
-At times, this mode of empirical philosophical engagement with its objects may well seem overly concerned with the technical minutiae of the groceries and recommender systems, and lacking in  sociological rigour or empirical evidence.
-With some forbearance on the part of readers, however, the main philosophical argument of the paper should resonate sociologically.
-It concerns social order, or the consistency and regularity of relating and acting amid a world in where probabilities and calculations based on probabilities play a more pervasive role.
-
-The main argument of the paper concerns how  probability operates in Big Data situations.
-While calculations of probability have  long-standing importance in many settings (insurance, medicine, experimental and field sciences, operations research, risk analysis, engineering design, public health,  economic modelling, etc.), probabilities have taken on a much thicker situational texture on digital platforms associated with the web and social media.
+The main argument of the paper concerns how  personalization occurs in Big Data situations.
 In recent research and opinion on data practice, personalization has been seen as the linchpin of predictive modelling.
 Big data discourse in its promissory mode  attributes potency to personalization: 'most important, using big data we hope to identify specific individuals rather than groups' [@Mayer-Schonberger_2013].
-Similarly, Joseph Turow and co-authors conclude in their critical account of the transformation of retail space: 'through it all, knowingly and not, and away from the spotlights of fierce social debate, retailers are encouraging daily routines that accept data-driven personalization as a centrifugal public' [@Turow_2015, 476].
+Conversely, Joseph Turow and co-authors conclude in their critical account of the transformation of retail space: 'through it all, knowingly and not, and away from the spotlights of fierce social debate, retailers are encouraging daily routines that accept data-driven personalization as a centrifugal public' [@Turow_2015, 476].
 Analysis of internet filter bubbles and the growth of predictive platforms in industry and government make similar arguments [@Pariser_2011; @ONeil_2016a].
@@
 The reconstruction of the recommender system suggests a different way of seeing predictive personalization by expanding the key underpinning notion of probability.
-Probabilities in practice -- in this case, in recommender system -- diverge from probability theory and indeed from statistics as a science.
+While calculations of probability have  long-standing importance in many settings (insurance, medicine, experimental and field sciences, operations research, risk analysis, engineering design, public health,  economic modelling, etc.), probabilities have taken on a much more operational texture on digital platforms associated with the web and social media.
+Probabilities in practice -- in this case, in recommender systems -- diverge from probability theory and indeed from statistics as a science.
 Operational probabilities differ from epistemic probabilities in their position in the world, as part of infrastructures, transactions, and event-structures.
@@
 The average everydayness of the example here -- grocery shopping -- allows some first hand engagement with the messiness, the entanglements, and potentials of this conversion event in ways that other interesting settings -- social media for instance -- do not.
+
+Like some recent work in science and technology studies,  anthropology and media studies [@Marres_2017; @Marcus_2014; @Bogost_2012],  the paper allows its own reconstruction to be provisionally shaped by an experiential encounter with the object it describes.
+At times, this mode of empirical philosophical engagement with its objects may well seem overly concerned with the technical minutiae of the groceries and recommender systems, and lacking in  sociological rigour or empirical evidence.
+With some forbearance on the part of readers, however, the main philosophical argument of the paper should resonate sociologically.
+It concerns social order, or the consistency and regularity of relating and acting amid a world in where probabilities and calculations based on probabilities play a more pervasive role.
 
@@
 Patel mentioned many of these steps only fleetingly in the presentation, for they are largely taken for granted as part of predictive analytic practice.
-Second, Patel focused on the renovation and updating of long-standing data-mining practice via a much more explicitly 'big data' and 'machine learning' oriented implementation.
-Her presentation concerns what we might term, for reasons that will gradually emerge, a 'conversion event' in which a long-standing recommender system was replaced by a new, explicitly 'big data'-style system concerned with 'personally relevant' recommendations.
+Second, Patel focused on the renovation and updating of long-standing data-mining practice via a much more explicitly 'big data' and 'machine learning'- oriented implementation.
+Her presentation concerns what we might term, for reasons that will gradually emerge, a 'conversion event' in which a long-standing recommender system was replaced by a 'big data'-style system concerned with 'personally relevant' recommendations.
 What stands out from the presentation is the normality of the recommender system: it is part of a long-standing and ongoing transformation or conversion of grocery shopping.
-From a sociological standpoint, the interest lies less in specific technical innovations and more on how the ordering work done by recommender systems relates to problems of social order more generally.
+From a sociological standpoint, the interest lies less in specific technical innovations and more on how the ordering work done by recommender systems relates to the role of prediction in social order more generally.
+
+# C. Shopping  and  a theory of list orderings
 
@@
 When Tesco customers shop for groceries online, a list of five recommendations appear at the checkout stage.
-The recommendations are the product of a recommender system, an important category of operational device in big data (see for instance [@Striphas_2015] for analysis of Netflix recommendation; [@Morris_2015]  or [@Seaver_2015] for an account of music recommendations).
+The recommendations are the product of a recommender system, an important category of operational device in big data (see for instance [@Hallinan_2014] for analysis of Netflix recommendation; [@Morris_2015]  or [@Seaver_2015] for an account of music recommendations).
 The question 'have you forgotten?' is followed by a list of some  grocery items that could have been or are usually on a shopping list.
 The title of the suggestions is a bit misleading: the recommender system, as we will see, is not concerned with forgetting, with the many slips and oversights associated to which shopping is prone,  but rather with items that customers had not selected perhaps because they had never thought of buying them in the first place.
-
-# C. Shopping  and  a theory of list orderings
 
@@
 What Miller suggests here is that shopping practices attempt to resolve difference between how people think they should live and how they actually live: 'we have to watch how shopping helps resolve these discrepancies between the normative and the actual, but we also need some ideas of where the normative comes from in the first place' [@Miller_2012, 72].
-Whatever theory we have of how normative social order comes about, the key argument is that shopping inhabits a space between normative/idea and actual order.
+Whatever theory we have of how normative social order arises, the key argument is that shopping inhabits a space between normative/idea and actual order.
 Importantly, and this will be a key consideration for recommender systems, shopping is not particularly personal or individual.
-Grocery shopping in particular is woven through with forms of social order concerning family and other forms of social grouping, and their associated relations (love, etc.).
+Grocery shopping in particular is saturated  by fluxing forms of social order concerning family and other forms of social grouping, and their associated relations (love, etc.).
 

### 2017-08-02 log.rmd
updating slides
18/32 pages revised
27/31 pages

@@
 There are around 35 pages on manuscript.
-Plan for revision: 7 pages/day for one week; (7 hours)  go through referee reports again alongside the manuscript to put them in (2 hours); redo diagrams and references (3 hours), submission (1 hour).
-Total of about 12 hours.
+Plan for revision: 7 pages/day for one week; (5 hours); enter edits 5 hours;  go through referee reports again alongside the manuscript to put them in (2 hours); redo diagrams and references (3 hours), submission (1 hour).
+Total of about 16 hours.
 Might be to do about 6-7 hours this week.
+- did 8 pages of revisions; and 20 edits
+
+## Tue 01 Aug 2017 15:53:17 BST
+- did another 10 pages of revisions -- up to 51 edits
+
+## Wed 02 Aug 2017 10:49:46 BST
+- reached page 27/31 in edits  on paper

### 2017-08-31 log.rmd
entered 89/89 edits

@@
 - today is Section E; I have about 7 hour left to finish this; so will need to get that section done today.
-- working hard on the probability stuff - wrote a section introducting Popper; reframed arules discussion.
+- working hard on the probability stuff - wrote a section introducing Popper; reframed arules discussion.
 - TODO - short paragraph summarising the social theory of list ordering; make sure to add stuff on the ordering-remapping of order in the list;
@@
 - reached page 27/31 in edits  on paper
+
+## Thu 31 Aug 2017 16:34:43 BST
+- long gap here -- due to important other activities such as summer holidays;  
+- found the 1989 Humby article on demographics -- argues for the need to add lifestage to that of neighbourhood. 
+- TODO: add reference to [@Humby_1989]

### 2017-08-31 reconstruction.rmd
entered 89/89 edits

@@
 
-This paper philosophically reconstructs [@Dewey_1957] an online grocery ordering system undergoing an upgrade from a demographically-based recommendation system to a 'personal relevance' model.
+This paper reconstructs [@Dewey_1957] an online grocery ordering system undergoing an upgrade from a demographically-based recommendation system to a 'personal relevance' model.
 The reconstruction is empirical in certain respects.
@@
 There are two ways of reading this statement.
-One is to see Popper as simply stating one of the consequencies of his physicalist understanding of probabilities as tendencies inherent in a situation.
-Another is to focus on the pivotal phrase 'possibilities trying to realize themselves' by asking what 'trying to realize' might mean.
-I would suggest that Patel's description of DunnHumby's work on the new personal relevance model is just such a 'trying to realize' in important ways.
+One is to see Popper as stating a triusm,  the obvious consequence of his physicalist understanding of probabilities as tendencies inherent in a situation.
+Another is to focus on the pivotal phrase 'possibilities trying to realize themselves' by asking what 'trying to realize' might mean in practice.
+I would suggest that DunnHumby's work on the new personal relevance model is just such a 'trying to realize' in important ways.
 
@@
 A 'good feature' is one that somehow improves the accuracy, precision, specificity or any of the panoply of measures applied to machine learning models in practice.
+HERE
 The precise meaning of 'good', however, remains open to many different attempts, experiments and variations.
 I find it interesting that the main efforts that DunnHumby made to  construct good features were not closely associated with personhood, but with the relations -- the propensities -- between things.
-
 Patel, for instance, addressed the problem of 'basket similarity,' the problem that the recommender system might recommend something similar to items already in a customer's basket.
-A customer might be willing to substitute a similar item for something they have already chosen, but they are more likely to accept a recommendation for an item that complements rather than duplicates items in their list.
+A customer might be willing to substitute a similar item for something they have already chosen, but they are more likely to accept a recommendation that complements rather than duplicates items in their list.
 How could the recommender system sense the possible substitutions and complementaries between things for a given customer, especially since a person's sensibilities and susceptibilities concerning groceries will be subtly shaped by wider social groups, orderings and circumstances?

### 2017-08-31 reconstruction.rmd
abstract tweak

@@
 abstract:
-    "Many accounts of big data systems assume that they target individuals.
+    "Many accounts of big data platforms assume that they target individuals.
 Personalization, with all the risks of discrimination and bias it entails, has been the focus of inquiry.
@@
 Drawing on a theory of probability proposed by the philsopher of science Karl Popper, it suggests that personal relevance is inevitably bound up with relations to others, and the relations between things.
-Using a mixture of discourse analysis and code-based reconstruction of key elements of the recommender system, it suggests that personalization is one provisional stabilization of a much more open-ended weave of propensities associated with people and things in contemporary big data configurations.
-The paper explores how in the context of recommender systems the constitutive incompleteness of shopping lists, their propensity to expand or change, becomes more important than their capacity to be personalized."
+Using a mixture of discourse analysis and code-based reconstruction of key elements of the recommender system, it suggests that personalization is one facet of a much more open-ended weave of propensities associated with people and things in contemporary big data configurations.
+The paper explores how in the context of recommender systems the constitutive incompleteness of shopping lists, their propensity to expand or change, might be more important than their capacity to be personalized."
 ---

### 2017-09-07 reconstruction.rmd
29/31 pages edited

@@
 DunnHumby -- founded by operations researchers Edwina Dunn and Clive Humby -- is said to have convinced the CEO of Tesco sometime in 1991 that a loyalty card program could change the supermarket chain's relationship to its customers.
-Clive Humby's academic publications  are hard to track down, but an early paper given at the _Conference of Young Operational Researchers_ in Nottingham in 1984 suggests the direction that he, DunnHumby and later Tesco would take in constructing lists [@OKeefe_1984].
-The abstract for Humby's presentation pre-figures an ongoing trajectory for data mining techniques aimed at eliciting detailed information on individual customer preferences.
+Clive Humby's academic publications are hard to track down (see [@Humby_1989] where Humby argues for the need to add lifestage data to neighbourhood data in marketing research), but an early paper given at the _Conference of Young Operational Researchers_ in Nottingham in 1984 suggests the direction that he, DunnHumby and later Tesco would take in constructing lists [@OKeefe_1984].
+The abstract for Humby's presentation pre-figures an ongoing trajectory for data mining techniques aimed at eliciting detailed information on individual customer references.
 
@@
 
-[^122]: See [@Lo_2002] for an introduction to the construction of experimental set ups for uplift models.
-Market researchers and data miners seem to have started adopting a statistically grounded experimental predictive practice sometimes around 2000.
+[^122]: See [@Lo_2002] for an introduction to the construction of experimental set ups for uplift models in marketing research.
+Market researchers and data miners seem to have started adopting a statistically-grounded experimental predictive practice around 2000.
 See for instance [@Almquist_2001]  Response modelling, 'propensity' modelling and uplift modelling all seek to identify associations between 'treatments' or interventions and the 'Responders', the people  affected by the treatment.
 
-## The openness of the data -- new possibilities
+## The openness of the data: new features
 
@@
 
-There are two ways of reading this statement.
-One is to see Popper as stating a triusm,  the obvious consequence of his physicalist understanding of probabilities as tendencies inherent in a situation.
-Another is to focus on the pivotal phrase 'possibilities trying to realize themselves' by asking what 'trying to realize' might mean in practice.
-I would suggest that DunnHumby's work on the new personal relevance model is just such a 'trying to realize' in important ways.
+There are different ways of reading this statement.
+We can read Popper as stating a truism,  the obvious consequence of his physicalist understanding of probabilities as tendencies inherent in a situation: anything can happen in the future.
+An alternative, perhaps more interesting reading  focuses on the pivotal phrase: 'possibilities trying to realize themselves.' What does 'trying to realize' might mean in practice?
+I would suggest that DunnHumby's work on the new personal relevance model, and much big data practice in general,  contains just such a 'trying to realize' in practice.
 
-Several times, Patel emphasized the importance of 'constructing good features' from the data, and much of her presentation concerned different attempts to construct 'good features.' A 'feature' in the context of machine learning refers to a variable included in a predictive model.
-A 'good feature' is one that somehow improves the accuracy, precision, specificity or any of the panoply of measures applied to machine learning models in practice.
-HERE
-The precise meaning of 'good', however, remains open to many different attempts, experiments and variations.
-I find it interesting that the main efforts that DunnHumby made to  construct good features were not closely associated with personhood, but with the relations -- the propensities -- between things.
-Patel, for instance, addressed the problem of 'basket similarity,' the problem that the recommender system might recommend something similar to items already in a customer's basket.
+Several times in her presentation, Patel emphasized the importance of 'constructing good features' from the data, and much of her presentation concerned DunnHumby's efforts to construct 'good features.' A 'feature' in the context of machine learning refers to a variable included in a predictive model.
+A 'good feature' contributes to the accuracy, precision, specificity or any of the panoply of measures applied to machine learning models in practice.
+The construction of 'good features', however, remains open to many different possibilities, some of which are more practically feasible than others.
+
+The main efforts that DunnHumby made to  construct good features were not closely associated with personhood, but with the relations -- the propensities -- between things.
+Patel, for instance, addressed the problem of 'basket similarity.' The recommender system should not recommend items that are too similar to groceries already in a customer's basket.
 A customer might be willing to substitute a similar item for something they have already chosen, but they are more likely to accept a recommendation that complements rather than duplicates items in their list.
-How could the recommender system sense the possible substitutions and complementaries between things for a given customer, especially since a person's sensibilities and susceptibilities concerning groceries will be subtly shaped by wider social groups, orderings and circumstances?
-It could only do that if had some sense of the relation between items already in the list and the field of possible recommendations.
-The 'personal relevance' model, therefore, sought to include the relation between recommendations and items already ordered.
-The data scientists constructed a new feature from the data concerning how the substitutabilities and complementarities between product.
-Taking all the baskets of items purchased, rather in the way that association rules does, a new feature, what DunnyHumby termed 'self-learning substitutes' was derived.
-The 'self learning substitutes' feature of the relevancy model draws on another data matrix, the product similarity matrix, itself generated from all previous recommendations that have led to conversion events (that is, 'a user clicked as the response').
+How could the recommender system sense the possible similiarities and complementaries between things in general or for a given customer, especially since a person's sensibilities and susceptibilities concerning groceries will be subtly shaped by wider social groups, orderings and circumstances?
+
+The predictive model could only do that if it had some sense of the relation between items already in the list and the field of possible recommendations.
+The 'personal relevance' model, therefore, sought to include the similarities and complementarities between items already ordered and recommendations.
+The DunnyHumby data scientists constructed a new feature from the data measuring substitutability and complementarity between products.
+Taking all the baskets of items purchased, a new feature, what DunnyHumby termed 'self-learning substitutes' was derived.
+Added to the relevancy model, the 'self learning substitutes' feature adds another data matrix, the product similarity matrix, itself generated from all previous recommendations that have led to conversion events (that is, 'a user clicked as the response').
 A customer's recommendation list might already include `Tesco Ginger Nuts 300g.`   The 'self-learning substitution model' seeks to replace list items with similar but higher value items: `Mcvitie's Ginger Moments 400g.`
+The self-learning substitutes feature is complex.
+Patel mentioned that it took the form of a 'design matrix of 14,000 columns.' 14,000 new explanatory variables were added to the logistic regression model producing the recommendations.
+Each recommendation would be subtly re-weighted by this complex derived feature.[^215]
 
-This new feature is complex.
-Patel mentioned that it took the form of a 'design matrix of 14,000 columns.' This means that 14,000 new explanatory variables were added to the logistic regression model producing the recommendations.
-Each recommendation would be subtly re-weighted by this complex derived feature.
-The complexity of the model itself introduces new instabilities, new 'possibilities trying to realize themselves' and the DunnyHumby data scientists used  '`L1 regularization` to drive coefficients down to zero', or, put more simply, to exclude variables there were not contributing much to the predictions produced by the model.
-
-Even this added feature, a complex addition in many respects, does not exhaust DunnHumby's attempts to re-shape the recommender system.
-They were just beginning, Patel reported, to address complementarities between things.
+[^215]: The proliferation of variables or features in the model itself introduces new instabilities, new 'possibilities trying to realize themselves' and the DunnyHumby data scientists used  '`L1 regularization` to drive coefficients down to zero', or, put more simply, to exclude variables there were not contributing much to the predictions produced by the model.
+Even this added feature does not exhaust DunnHumby's attempts to re-shape the recommender system.
+They were just beginning, Patel reported, to model complementarities between things.
 `Nutella` complements `bananas,` but the converse is not necessarily true since `bananas` need not complement `nutella,` especially if I have nut allergies.
-From the standpoint of customer data science, Patel emphasised the operational significance of predicting complementarities:  'complementarity drives conversion.' Bringing the matrix of similarities expressed as numerical associations between products into the model improves the conversion of recommendations into purchases.
+From the standpoint of customer data science, Patel emphasised the operational significance of predicting complementarities:  'complementarity drives conversion.'
 

### 2017-09-08 log.rmd
29/31 pages edited
31/31 pages entered

@@
 - found the 1989 Humby article on demographics -- argues for the need to add lifestage to that of neighbourhood. 
-- TODO: add reference to [@Humby_1989]
+- DONE: add reference to [@Humby_1989]
+- reached conclusion in adding edits. Not sure if the section on Popper and possibiities trying to realize themselves is fleshed out enough.
+- TODO: check if the last section does enough with propensities and the new features.
+
+## Fri 08 Sep 2017 09:24:22 BST
+- re-writing the conclusion today - did this on paper and now in file; it still needs more work
+- but in order to do that, I've had to go back to beginning and go through the paper looking at the argument again.< 
+- TODO: jump between D. and E. needs fixing. 
+- TODO: section on list theory is still awkward -- Miller is fine, but what then?

### 2017-09-08 reconstruction.rmd
31/31 pages entered

@@
 Personalization, with all the risks of discrimination and bias it entails, has been the focus of inquiry.
-This paper suggests that personalization occurs against a more expansive background that in principle is not reducible to individual attributes and desires, but includes many different forms of relations in flux.
+This paper suggests that personalization occurs against a more expansive background of probabilization that in principle is not reducible to individual attributes and desires, but includes many different forms of relations in flux.
 It describes the 'personalization' of an online grocery shopping recommender system.
-It reconstructs an account of how a recommender system constructs an ordered list of a small number of grocery items of personal relevance for each of the millions of online grocery shoppers in  a major UK supermarket chain.
-Drawing on a theory of probability proposed by the philsopher of science Karl Popper, it suggests that personal relevance is inevitably bound up with relations to others, and the relations between things.
+It reconstructs an account of how a recommender system arranges an ordered list of a small number of grocery items of personal relevance for each of the millions of online grocery shoppers in  a major UK supermarket chain.
+Drawing on a theory of probability proposed by the philsopher of science Karl Popper, it suggests that personal relevance is inevitably caught up in relations to others, and the relations between things.
 Using a mixture of discourse analysis and code-based reconstruction of key elements of the recommender system, it suggests that personalization is one facet of a much more open-ended weave of propensities associated with people and things in contemporary big data configurations.
@@
 
+## Introduction
+
 This paper reconstructs [@Dewey_1957] an online grocery ordering system undergoing an upgrade from a demographically-based recommendation system to a 'personal relevance' model.
 The reconstruction is empirical in certain respects.
-It departs from an ethnographic moment: being a member of the audience at an industry/academic conference where a data scientist was describing how recommendations were created for online grocery orders.
+It departs from an ethnographic moment: being a member of the audience at an industry/academic conference where a data scientist was describing how personalized recommendations were created for online grocery orders.
 Starting from this highly mediated and fragmentary encounter with a recommender system and one of its designers, the paper explores the problem of personally relevant recommendations.
@@
 
-The main argument of the paper concerns how  personalization occurs in Big Data situations.
+The main argument of the paper concerns how  personalization occurs in Big Data 'conversion events' (a term that refers to the common before-after narrative forms that proponents, developers and advocates of these systems adopt).
 In recent research and opinion on data practice, personalization has been seen as the linchpin of predictive modelling.
@@
 Her presentation concerns what we might term, for reasons that will gradually emerge, a 'conversion event' in which a long-standing recommender system was replaced by a 'big data'-style system concerned with 'personally relevant' recommendations.
-What stands out from the presentation is the normality of the recommender system: it is part of a long-standing and ongoing transformation or conversion of grocery shopping.
+What stands out from the presentation is the normality of the recommender system: it is part of a long-standing and ongoing transformation of grocery shopping.
 From a sociological standpoint, the interest lies less in specific technical innovations and more on how the ordering work done by recommender systems relates to the role of prediction in social order more generally.
@@
 The question 'have you forgotten?' is followed by a list of some  grocery items that could have been or are usually on a shopping list.
-The title of the suggestions is a bit misleading: the recommender system, as we will see, is not concerned with forgetting, with the many slips and oversights associated to which shopping is prone,  but rather with items that customers had not selected perhaps because they had never thought of buying them in the first place.
+The title of the suggestions is a bit misleading.
+The recommender system, as we will see, is not concerned with forgetting, with the many slips and oversights associated to which shopping is prone,  but rather with items that customers had not selected perhaps because they had never thought of buying them in the first place.
 
-The anthropologist  Daniel Miller has argued [@Miller_2012] that  shopping takes place as a negotiation of discrepancies between normative and actual social order (for instance, between the ideals of health emblematised by organic products and commitments to thrift embodied in generic products).
+The anthropologist Daniel Miller has argued [@Miller_2012] that  shopping takes place as a negotiation of discrepancies between normative and actual social order (for instance, between the ideals of health emblematised by organic products and commitments to thrift embodied in lower cost generic products).
+What Miller suggests here is that shopping practices attempt to resolve difference between how people think they should live and how they actually live: 'we have to watch how shopping helps resolve these discrepancies between the normative and the actual, but we also need some ideas of where the normative comes from in the first place' [@Miller_2012, 72].
 More provocatively, Miller claims that 'shopping is largely a technology for the expression of love' (85).
-What Miller suggests here is that shopping practices attempt to resolve difference between how people think they should live and how they actually live: 'we have to watch how shopping helps resolve these discrepancies between the normative and the actual, but we also need some ideas of where the normative comes from in the first place' [@Miller_2012, 72].
-Whatever theory we have of how normative social order arises, the key argument is that shopping inhabits a space between normative/idea and actual order.
+It is, to use the language of actor-network theory, a practice of translation [@Callon_1991].
+Whatever theory we have of how normative social order arises, the key implication is that shopping inhabits a mundane space between normative/idea and actual order.
 Importantly, and this will be a key consideration for recommender systems, shopping is not particularly personal or individual.
@@
 
+If social ordering is done in shopping, then what do shopping lists do?
 Shopping lists provide important clues to how local social order is constituted, maintained, and repaired in shopping.
-If social ordering is done in shopping,  then shopping list as an inscriptive device stabilises certain forms of ordering at the expense of other possible orderings.
-The shopping list, whether written on the back of an envelope, or saved as a list in an online grocery shopping system, lies at the intersection of logistical flows, infrastructural orderings, and  lively negotiations around actual and normative social orders.[^30.1]  Shopping lists, therefore  are intersectional ordering devices that encapsulates a universe of possible references, and a teeming multitude of propensities with an actual local order (in both senses of that term: an ordering and an imperative).
+As an inscriptive device, a list stabilises certain quite complex forms of social order at the expense of other possible orderings.
 As people shop, either by  trawling along aisles packed with thousands of products, or scrolling down screens or searching for particular brands amidst search results, lists inscribe some kind of order that filters or reduces the excessive propensities, claims, dazzle and distraction of commodities to the practicalities of domestic economy.
+The shopping list, whether written on the back of an envelope, saved as a list in an online grocery shopping system, or remembered, lies at the intersection of logistical flows, infrastructural orderings, and  lively negotiations around actual and normative social orders.[^30.1]  Shopping lists  are intersectional ordering devices that encapsulate a universe of possible references, and a teeming multitude of propensities with an actual local order.
 
@@
 But online shopping lists take shape at the intersection of web and internet infrastructures, supply chain logistics, individualised practices and *habitus,* and increasingly, the predictive operations of  recommender systems.
-Just as the aisles and shelves of a supermarket present a densely-woven semiotics of objects competing for visual attention by offering distinctions of taste, thrift, expedience, novelty, indulgence, health and increasingly, online shopping recommender systems generate lists that seek to align people to commodities that they otherwise might have little relation to (see [@Turow_2015] for an overview).
+Just as the aisles and shelves of a supermarket present a densely-woven semiotics of objects competing for visual attention by offering distinctions of taste, thrift, expedience, novelty, indulgence, health and increasingly, online shopping recommender systems generate lists that seek to align people to commodities that they otherwise might have little relation to (see [@Turow_2015] for an overview of the development of these systems).
 
@@
 
-The main narrative of Patel's presentation concerned a shift from a well-established loyalty-card based data-mining model to a predictive, probabilistic, personally relevant re-writing of the shopping list in almost-realtime.
-The changes Patel described are increasingly widespread and common.
-While they are configured in Tesco-specific ways by DunnHumby (and this reflects a longer history), they are also broadly typical.
+The main narrative of Patel's conversion narrative concerned the shift from a well-established loyalty-card based data-mining model to a predictive, probabilistic, personally relevant construction of the shopping list in almost-realtime.
+While the changes Patel described are configured in Tesco-specific ways by DunnHumby, they are also broadly typical.
 Tesco is the largest supermarket chain in UK, famous amongst retailers for its  customer loyalty and targeted marketing programme known as 'Tesco Clubcard,' which started in 1991.
 DunnHumby -- founded by operations researchers Edwina Dunn and Clive Humby -- is said to have convinced the CEO of Tesco sometime in 1991 that a loyalty card program could change the supermarket chain's relationship to its customers.
-Clive Humby's academic publications are hard to track down (see [@Humby_1989] where Humby argues for the need to add lifestage data to neighbourhood data in marketing research), but an early paper given at the _Conference of Young Operational Researchers_ in Nottingham in 1984 suggests the direction that he, DunnHumby and later Tesco would take in constructing lists [@OKeefe_1984].
+Clive Humby's academic publications are hard to track down (see [@Humby_1989] where Humby argues for the need to add lifestage data to neighbourhood data in marketing research), but an early paper given at the _Conference of Young Operational Researchers_ in Nottingham in 1984 suggests the direction that he, the company Dunn and Humby formed (DunnHumby) and later Tesco would take in constructing lists [@OKeefe_1984].
 The abstract for Humby's presentation pre-figures an ongoing trajectory for data mining techniques aimed at eliciting detailed information on individual customer references.
@@
 It can no longer rely on placement of goods in carefully chosen locations in stores.
-It needs or at least might want to continuously up-sell and cross-sell to customers who might hardly ever visits the supermarket itself.
+It needs or at least might want to continuously up-sell and cross-sell to customers who might hardly ever visit the supermarket itself.
 
 How could we characterise the shift from ClubCard data mining to online grocery markets?
-If 'Tesco is the clear winner in the online grocery market, in fact it takes almost 50p of every £1 spent on food shopping on the internet' [How SEO helps Tesco to dominate the online grocery market](https://econsultancy.com/blog/64841-how-seo-helps-tesco-to-dominate-the-online-grocery-market)[@Silverwood-Cope_2014], then Tesco itself has undergone some kind of conversion?
-Patel described her work at DunnHumby as shifting the recommender system from a 'rules-based list' to a 'relevancy model.' The relevancy model changed how the 'Have You Forgotten' list was constructed.
+If 'Tesco is the clear winner in the online grocery market, in fact it takes almost 50p of every £1 spent on food shopping on the internet' [How SEO helps Tesco to dominate the online grocery market](https://econsultancy.com/blog/64841-how-seo-helps-tesco-to-dominate-the-online-grocery-market)[@Silverwood-Cope_2014], then has Tesco itself undergone some kind of conversion?
+Patel described her work at DunnHumby as converting the recommender system from a 'rules-based list' to a 'relevancy model.' The relevancy model changed how the 'Have You Forgotten' list was constructed.
 This shift is a typical of broader re-organisation of prediction in which existing data analysis practices are being re-distributed and intensified in particular ways.
@@
 This is not a drastic or abrupt event.
-Academic researchers first began writing about recommender systems in the mid-1990s.
+Academic researchers first began writing about personalized recommender systems in the mid-1990s.
 From the outset they highlighted a potential shift from demographic-driven market research or data mining techniques to personalized recommendations.
-For instance, writing in 1997 in a special section of the _Communications of the ACM_ on recommender systems, Paul Resnick and Hal Varian (at that time Dean of Information Sciences at UC Berkeley, but currently Chief Economist at Google), made much of this personalisation [@Resnick_1997].
+For instance, writing in 1997 in a special section of the _Communications of the ACM_ on recommender systems, Paul Resnick and Hal Varian (at that time Dean of Information Sciences at UC Berkeley, but currently Chief Economist at Google), made much of this personalization [@Resnick_1997].
 Resnick and Varian  emphasise the need to distinguish the emerging practices from data mining:
 
-> In everyday life, we rely on recommendations from other people.
+>In everyday life, we rely on recommendations from other people.
 ...
@@
 
-Writing just after the advent of web-based e-Commerce, Resnick and Varian  imagined recommender systems augmenting the 'collaborative filtering' that people do when they write and read reviews of products and services (for instance, on the travel website, TripAdvisor; [@Scott_2012]).
+Writing just after the advent of web-based e-Commerce, Resnick and Varian  imagined recommender systems drawing on  the 'collaborative filtering' that people do when they write and read reviews of products and services (for instance, on the travel website, TripAdvisor; [@Scott_2012]).
 In 1997,  Resnick and Varian expected that 'people [would] provide recommendations' and recommender systems would aggregate and rank recommendations for recipients.
@@
 
-The 'minute and extensive scrutinizing' [@Dewey_1957] of a recommender system undergoing a conversion from an older data-mining approach to a current 'personal relevance model' approach addresses a specific situation: online grocery shopping (not shopping in general) and, even more micro-sociologically, the ordering of lists.
+The personalization of recommendations has been a defining characteristic of many big data conversion events.
+The exploitative and discriminatory effects of all-consuming personalization have been the target of critics and critical research on data economies.
+The archaeology of the Tesco recommender system, including its early experiments with demographic data-mining, its adoption of a rule-based 'market basket' analysis system, and its current implementation of a 'personal relevance model' follows the trajectory of personalization.
 
-The theory of shopping lists constructed at the beginning of the paper was a scaffolding component.
-It draws on anthropology, sociology and mathematics to suggest that we see shopping lists as the site of a complex negotiation between different social orders, between norms and actualities.
-The list itself as an inscriptive device functions as an ordering device, a device that accommodates an immense variety of specific habits, dispositions, values and associations.
-The mapping element of the list, the order that arranges the set of items according to different interests, values, and aims, is the operational focus of recommender systems.
-
-The personalization of recommendations has been a defining characteristic of many big data conversion events.
-Rightly, the effects of all-consuming personalization have been the target of critics and critical research.
-The archaeology of the Tesco recommender system, including its early experiments with demographic data-mining, its adoption of a rule-based 'market basket' analysis system, and its current implementation of a 'personal relevance model' certainly follows that trajectory.
-But the conversion event is not reducible to personalization.
-
-Probabilization precedes personalization.
-When we observe what is practically done to construct a personal relevance model and apply it to a shopping list, we see, conditional probabilities, predictive models, the importance of data accumulated over time,  platform experiences, the rhythms of infrastructural maintenance, and open-ended construction of new features in data, all of which negotiate shifting and complicated associations between people and things, people and people, and things with things.
-If social order is made of associations, is to be social is to associate, then the forms of association taking shape in this and other similar conversion events might be understood as understand as operationalizing probability in matrices of propensity.
-As Popper argues, probabilities are being realized in the same way that fields of force were realized in modern physics.
-Probabilization does not take place in physics laboratories but on platforms, combining databases, compute capacity, interfaces and global supply chain logistics.
-
-The fact that probabilisation largely takes the form of personalization tends to obscure the re-framing of social order taking place.
-This ordering  depends on  the mappings that inscriptive devices such as shopping lists  display.
+The fact that the conversion event largely takes the form of personalization,  an individualizing personalization that is so patently ill-fit to the socially complex negotiations of grocery shopping, tends to obscure the wider re-framing of social order taking place.
+on  the mappings that inscriptive devices such as shopping lists  display.
 Only because shopping is already somewhat ordered by lists can a partial re-mapping, provisionally conducive to new events, take place.
@@
 
+But the conversion event associated with big data is not reducible to personalization or the individual.
+Probabilization precedes and exceeds personalization.
+When we observe what is practically done to construct a personal relevance model and apply it to a shopping list, we see conditional probabilities, predictive models, the importance of data accumulated over time,  platform experiences, the rhythms of infrastructural maintenance, and open-ended construction of new features in data, all of which shift and complicate the conditional associations between people and things, people and people, and things with things.
+If social order is made of associations, if to be social is to associate [@Latour_TBA], then the associations taking shape in this and other similar big data conversion events might be understood as operationalizing probability in matrices of propensity.
+If, as Popper argues, probabilities are being realized in the same way that 'fields' were realized in modern physics, then we should see the compute  platforms, databases, software libraries, web interfaces and global supply chain logistics as part of this conversion event.
+
 Reconstructions of specific conversion events associated with big data can work in different ways.

### 2017-09-25 reply_to_reviews.md
starting final revisions on main text

@@
-Thanks very much to reviewers for their work on what must have seemed like a frustratingly elusive and adumbratory piece of work.
+Thanks very much to reviewers for their work on what must have seemed like a frustrating and elusive manuscript.
 The reviews have been a great help in re-writing the piece.
@@
 There were many.
-My summary of the  major issues identified in the reviews  is below.
+The main focus in the re-writing of the manuscript was the overall argument.
+
+Much of the confusion about the main argument stemmed from its multiple threads.
+It was engaging in several different debates at once -- about recommender systems and their exploitative effects on  consumption practices, about recommender systems as examples of how big data practices take root in specific sectors, and recommender systems as symptom of some kind of more ontological transformation in sociotechnical collectives.
+In the re-write, I've disentangled these threads in order to clarify the argument: grocery recommender systems are a mundane example of changes associated with big data; personalization is a powerful ordering technique in big data practice; but personalization cannot account for either the range of practices associated with big data or the broader associated transformations.
+The 'conversion' of the Tesco grocery recommender system to a personal relevancy model helps make this argument: grocery shopping is a complex social negotiation, intimately entwined with social groups (family, household, etc).
+Data-mining to target grocery shoppers pre-dates by several decades recent growth in predictive analytics.
+A personalized recommender system, based even on post-demographic data, will struggle to effectively personalize grocery recommendations.
+
+The manuscript has been rewritten from almost the ground up.
+It would be hard to list all the changes.
+The basic idea of the rewrite was remove any thread of discussion that didn't relate to the main argument.
+A lot of the discussion of list practice was distracting so that has been reduced to a minimum.
+I wasn't able to provide the more substantial empirical engagement requested by Reviewer 2.
+That would require a new research project.
+The case study cannot make any claims to ethnographic thickness, so what the paper does to deepen its argumentative engagement around personalization in order to develop a more genealogical account based  on  research literature, widely used software and some examples of data from the Tesco API.
 

### 2017-09-25 log.rmd
lu
starting final revisions on main text

@@
 - TODO: section on list theory is still awkward -- Miller is fine, but what then?
+
+## Tue 12 Sep 2017 14:29:19 CEST
+- DONE: get Boltanski, 'finding ones ways in social space', 1983, has stuff on items and people; cf Distinction; added to ~/archive/boltanski_navigating_1983.pdf
+
+## Mon 25 Sep 2017 09:37:00 BST
+- started writing the revision letter and have detailed the main changes there: reply_to_reviews.md 
+- redid the introduction again, take out lots of adjectives
+- TODO: rewrite abstract and conclusion
+- TODO: maybe should take out a lot of the list discussion. Is it too distracting?

### 2017-09-25 reconstruction.rmd
starting final revisions on main text

@@
 abstract:
-    "Many accounts of big data platforms assume that they target individuals.
-Personalization, with all the risks of discrimination and bias it entails, has been the focus of inquiry.
-This paper suggests that personalization occurs against a more expansive background of probabilization that in principle is not reducible to individual attributes and desires, but includes many different forms of relations in flux.
-It describes the 'personalization' of an online grocery shopping recommender system.
-It reconstructs an account of how a recommender system arranges an ordered list of a small number of grocery items of personal relevance for each of the millions of online grocery shoppers in  a major UK supermarket chain.
-Drawing on a theory of probability proposed by the philsopher of science Karl Popper, it suggests that personal relevance is inevitably caught up in relations to others, and the relations between things.
-Using a mixture of discourse analysis and code-based reconstruction of key elements of the recommender system, it suggests that personalization is one facet of a much more open-ended weave of propensities associated with people and things in contemporary big data configurations.
-The paper explores how in the context of recommender systems the constitutive incompleteness of shopping lists, their propensity to expand or change, might be more important than their capacity to be personalized."
+    "Accounts of big data practices often assume that they target individuals.
+Personalization, with all the risks of discrimination and bias it entails, has been the critical focus in accounts of consumption, government, social media and health.
+This paper argues that personalization occurs against a more expansive change in probabilization that,  in principle, is not reducible to individual attributes and actions.
+It describes the 'personalization' of an online grocery shopping recommender system to list a small number of grocery items of personal relevance for each of the millions of online grocery shoppers at a  a major UK supermarket chain.
+Drawing on a theory of probability proposed by the philsopher of science Karl Popper and anthropological work on shopping, it suggests that personal relevance is caught up in relations to others and things.
+Using a mixture of discourse analysis and code-based reconstruction of key elements of the recommender system, it suggests that personalization is one facet of an open-ended weave of propensities associated with people and things in contemporary big data configurations.
+The paper explores how, in the context of recommender systems, the constitutive incompleteness of shopping lists, their propensity to expand or change, might be more important than their capacity to be personalized."
 ---
@@
 
-This paper reconstructs [@Dewey_1957] an online grocery ordering system undergoing an upgrade from a demographically-based recommendation system to a 'personal relevance' model.
-The reconstruction is empirical in certain respects.
+Much big data practice converts an existing set of practices by adding a model or predictive elements based on the collection and analysis of data.
+This paper recounts the conversion of  an online grocery ordering system  from an older demographically-based recommendation system to a 'personal relevance' model.
+The re-counting is empirical in certain respects.
 It departs from an ethnographic moment: being a member of the audience at an industry/academic conference where a data scientist was describing how personalized recommendations were created for online grocery orders.
-Starting from this highly mediated and fragmentary encounter with a recommender system and one of its designers, the paper explores the problem of personally relevant recommendations.
-It draws on anthropological and sociological research concerning shopping and list-making to situate the recommender system amongst the social ordering practices of shopping and lists.
-It makes use of archaeological approaches, in the sense developed by Michel Foucault [@Foucault_1972] in order to identify and map important functional statements and practices configuring the knowledge and order generated by such systems over time.
+Starting from this highly mediated and fragmentary encounter with a recommender system designer, the paper  draws on anthropological and sociological research concerning shopping and list-making to situate the recommender system amongst the social ordering practices of shopping and lists.
+It makes use of archaeological approaches, in the sense developed by Michel Foucault [@Foucault_1972] to identify and map functional statements and practices configuring the knowledge and order generated by such systems over time.
 It conducts several small-scale code-based experiments in order to reconstruct, using widely available code resources resources such as API (Application Programmer Interfaces) and software libraries for machine learning, some prototypical elements of the system in question.
-Throughout, it draws  on the philosophy of science and media theory to frame an engagement with data, algorithms and recommendations in terms of probabilities and their worldly enactments.
+It draws  on the philosophy of science and media theory to frame an engagement with grocery recommendations in terms of probabilities and their worldly enactments.
 
-The main argument of the paper concerns how  personalization occurs in Big Data 'conversion events' (a term that refers to the common before-after narrative forms that proponents, developers and advocates of these systems adopt).
+The main argument of the paper concerns how  personalization occurs in Big Data 'conversion events.' The term 'conversion event'  refers to the common before-after narrative forms that proponents, developers and advocates of these systems adopt.
 In recent research and opinion on data practice, personalization has been seen as the linchpin of predictive modelling.
@@
 Conversely, Joseph Turow and co-authors conclude in their critical account of the transformation of retail space: 'through it all, knowingly and not, and away from the spotlights of fierce social debate, retailers are encouraging daily routines that accept data-driven personalization as a centrifugal public' [@Turow_2015, 476].
-Analysis of internet filter bubbles and the growth of predictive platforms in industry and government make similar arguments [@Pariser_2011; @ONeil_2016a].
+Analysis of internet filter bubbles and the growth of predictive platforms in industry and government  [@Pariser_2011; @ONeil_2016a] also centre on personalization.
+In this paper, I agree that a transformation or conversion event is occurring, but one which also, even in the midst of much ongoing personalization,  opens the possibility of conceiving of social ordering processes afresh.
+The average everydayness and familiarity of the example here -- grocery shopping -- allows some first hand engagement with the messiness, the entanglements, and potentials of this conversion event in ways that other interesting settings -- social media for instance -- do not.
 
-The reconstruction of the recommender system suggests a different way of seeing predictive personalization by expanding the key underpinning notion of probability.
-While calculations of probability have  long-standing importance in many settings (insurance, medicine, experimental and field sciences, operations research, risk analysis, engineering design, public health,  economic modelling, etc.), probabilities have taken on a much more operational texture on digital platforms associated with the web and social media.
-Probabilities in practice -- in this case, in recommender systems -- diverge from probability theory and indeed from statistics as a science.
-Operational probabilities differ from epistemic probabilities in their position in the world, as part of infrastructures, transactions, and event-structures.
-Taking their complex siting into account, and accepting their _technical_ reality, we might come to understand operative probabilities along the lines of the philosopher of science Karl Popper understood as a 'world of propensities, as an unfolding process of realizing possibilities and of unfolding new possibilities' [@Popper_1990,19].
-Given this understanding of operational probability, our understanding of what is enacted in contemporary data analytics, especially in association with platforms [@Gillespie_2010] where the operationality of probabilities are heightened,  might shift focus away from personalization.
-
-The broad implication of this expansion  concerns how we approach social order or regularities amidst change or transformation.
-Proponents and critics of Big Data present the advent of large volumes of data and analytical techniques (machine learning in particular) as transforming for better or worse organisations, institutions, economic processes and social life.
-Personalization has been the pivotal element, they argue, of this transformation.
-I suggest too that a transformation or conversion event is occurring, but one which might also, even in the midst of much ongoing personalization,  open up the possibility of experiencing the world differently, and conceiving of social ordering processes somewhat afresh.
-Even if the argument is centred on grocery shopping, the analysis of how a world of propensities continuously and unevenly converts into realizations applies across the board, from the most physical to the most ideal or abstract settings.
-The average everydayness of the example here -- grocery shopping -- allows some first hand engagement with the messiness, the entanglements, and potentials of this conversion event in ways that other interesting settings -- social media for instance -- do not.
+'Data-driven personalization' relies on predictions, themselves dependent on data practices ranging across acquisition, storage, transformation, exchange, integration, modelling and experimentation.
+The reconstruction of the recommender system and its  personalized predictions I undertake here focuses on the notion of probability and its role in modelling and experimentation.
+Although probabilistic calculations of events have  long-standing importance in many settings (insurance, medicine, experimental and field sciences, operations research, risk analysis, engineering design, public health,  economic modelling, etc.), probabilities have taken on an operational function in digital platforms associated with the web and social media.
+They have become part of infrastructures, transactions, and event-structures.
+Taking their their _operation_ reality into account, we might understand probabilities in quasi-physicalist terms  defined by the philosopher of science Karl Popper as a 'world of propensities, as an unfolding process of realizing possibilities and of unfolding new possibilities' [@Popper_1990,19].
+This operational-physicalist or sociotechnical notion of probability shifts our understanding of data analytics, especially in association with platforms [@Gillespie_2010], away from personalization.
 
 Like some recent work in science and technology studies,  anthropology and media studies [@Marres_2017; @Marcus_2014; @Bogost_2012],  the paper allows its own reconstruction to be provisionally shaped by an experiential encounter with the object it describes.
-At times, this mode of empirical philosophical engagement with its objects may well seem overly concerned with the technical minutiae of the groceries and recommender systems, and lacking in  sociological rigour or empirical evidence.
-With some forbearance on the part of readers, however, the main philosophical argument of the paper should resonate sociologically.
-It concerns social order, or the consistency and regularity of relating and acting amid a world in where probabilities and calculations based on probabilities play a more pervasive role.
+At times, this mode of empirical philosophy may seem fixated on the technical minutiae of groceries and recommender systems, and lacking in  sociological rigour or empirical evidence.
+With some forbearance on the part of readers, however, the argument of the paper should resonate sociologically.
+It concerns social order, or the consistency and regularity of relating and acting amid a world in where probabilities and calculations based on probabilities play a pervasive role.
 
@@
 
-At one of the many industry-meets-academia events occurring in increasingly data science-oriented higher education institutions in the UK, speakers from industry, government and commerce described their work with predictive models.[^11.1] Shreena Patel, a PhD graduate in statistics and operations research,  works as a data scientist for DunnHumby, a well-known [ customer science company ](https://www.dunnhumby.com/) [@dunnhumby_2017].
+At one of the many industry-meets-academia events occurring in data science-oriented higher education institutions in the UK, speakers from industry, government and commerce described their work with predictive models.[^11.1] Shreena Patel, a PhD graduate in statistics and operations research,  works as a data scientist for DunnHumby, a well-known [ customer science company ](https://www.dunnhumby.com/) [@dunnhumby_2017].
 Her work at DunnHumby focuses on online grocery shopping at the supermarket chain Tesco.

### 2017-09-26 log.rmd
entered 22/34 page edits

--- 2017-08-31 log.rmd
@@
 - found the 1989 Humby article on demographics -- argues for the need to add lifestage to that of neighbourhood. 
-- TODO: add reference to [@Humby_1989]
+- DONE: add reference to [@Humby_1989]
+- reached conclusion in adding edits. Not sure if the section on Popper and possibiities trying to realize themselves is fleshed out enough.
+- TODO: check if the last section does enough with propensities and the new features.
+
+## Fri 08 Sep 2017 09:24:22 BST
+- re-writing the conclusion today - did this on paper and now in file; it still needs more work
+- but in order to do that, I've had to go back to beginning and go through the paper looking at the argument again.< 
+- DONE: jump between D. and E. needs fixing. 
+- DONE: section on list theory is still awkward -- Miller is fine, but what then?
+
+## Tue 12 Sep 2017 14:29:19 CEST
+- DONE: get Boltanski, 'finding ones ways in social space', 1983, has stuff on items and people; cf Distinction; added to ~/archive/boltanski_navigating_1983.pdf
+
+## Mon 25 Sep 2017 09:37:00 BST
+- started writing the revision letter and have detailed the main changes there: reply_to_reviews.md 
+- redid the introduction again, take out lots of adjectives
+- TODO: rewrite abstract and conclusion
+- DONE: maybe should take out a lot of the list discussion. Is it too distracting?
+
+## Tue 26 Sep 2017 09:13:38 BST
+- realised I haven't actually entered the paper edits -- so going through the paper copy entering them. There are 34 pages -- do 15 today?  
+- going through these edits, cutting out the list discussion as much as possible. In particular took out the more formal part of the list discussion -- not sure how it fits at all. 
+- reached li 117/222, p18/34 today. 
+- TODO: check list section again - does it flow?
+- did some more -- p 22/34, li 153/222

### 2017-09-26 reconstruction.rmd
entered 22/34 page edits

@@
 
-# C. Shopping and a theory of list orderings
+# C. Shopping list orderings
 
-Online grocery shopping at [Tesco](https://www.tesco.com/groceries/?icid=dchp_groceriesshopgroceries), the largest supermarket chain in the UK, includes recommendations for further grocery purchases under the title of 'Have you forgotten?'.
+Online grocery shopping at [Tesco](https://www.tesco.com/groceries/?icid=dchp_groceriesshopgroceries) includes recommendations for further grocery purchases under the title of 'Have you forgotten?'.
 When Tesco customers shop for groceries online, a list of five recommendations appear at the checkout stage.
@@
 
-The anthropologist Daniel Miller has argued [@Miller_2012] that  shopping takes place as a negotiation of discrepancies between normative and actual social order (for instance, between the ideals of health emblematised by organic products and commitments to thrift embodied in lower cost generic products).
-What Miller suggests here is that shopping practices attempt to resolve difference between how people think they should live and how they actually live: 'we have to watch how shopping helps resolve these discrepancies between the normative and the actual, but we also need some ideas of where the normative comes from in the first place' [@Miller_2012, 72].
+The anthropologist Daniel Miller has argued [@Miller_2012] that  shopping negotiates discrepancies between normative and actual social order (for instance, between the ideals of health emblematised by organic products and commitments to thrift embodied in lower cost generic products).
+Miller suggests that shopping practices attempt to resolve difference between how people think they should live and how they actually live: 'we have to watch how shopping helps resolve these discrepancies between the normative and the actual, but we also need some ideas of where the normative comes from in the first place' [@Miller_2012, 72].
 More provocatively, Miller claims that 'shopping is largely a technology for the expression of love' (85).
 It is, to use the language of actor-network theory, a practice of translation [@Callon_1991].
-Whatever theory we have of how normative social order arises, the key implication is that shopping inhabits a mundane space between normative/idea and actual order.
-Importantly, and this will be a key consideration for recommender systems, shopping is not particularly personal or individual.
+Whatever theory we have of how normative social order arises, the key implication is that shopping inhabits a mundane  but highly variable space between normative/idea and actual order.
+Importantly, and this will be a key consideration for recommender systems, shopping is not necessarily personal or individual.
 Grocery shopping in particular is saturated  by fluxing forms of social order concerning family and other forms of social grouping, and their associated relations (love, etc.).
 
-If social ordering is done in shopping, then what do shopping lists do?
-Shopping lists provide important clues to how local social order is constituted, maintained, and repaired in shopping.
-As an inscriptive device, a list stabilises certain quite complex forms of social order at the expense of other possible orderings.
-As people shop, either by  trawling along aisles packed with thousands of products, or scrolling down screens or searching for particular brands amidst search results, lists inscribe some kind of order that filters or reduces the excessive propensities, claims, dazzle and distraction of commodities to the practicalities of domestic economy.
-The shopping list, whether written on the back of an envelope, saved as a list in an online grocery shopping system, or remembered, lies at the intersection of logistical flows, infrastructural orderings, and  lively negotiations around actual and normative social orders.[^30.1]  Shopping lists  are intersectional ordering devices that encapsulate a universe of possible references, and a teeming multitude of propensities with an actual local order.
+Shopping lists provisionally stabilise the complex social ordering of grocery shopping.
+They provide important clues as to how local social order is constituted, maintained, and repaired in shopping.
+As people shop, either by  trawling along aisles packed with thousands of products, or scrolling down screens or searching for particular brands amidst search results, lists filter or reduce the excessive abundance, claims on attention, dazzle and distraction of commodities to the practicalities of domestic economy.
+The shopping list, whether written on the back of an envelope, saved as a list in an online grocery shopping system, or remembered, lies at the intersection of logistical flows, infrastructural orderings, and  lively negotiations around actual and normative social orders.[^30.1]  Shopping lists  are intersectional ordering devices that encapsulate a universe of possible references, and a teeming multitude of propensities with an actual local order.[^1.456]
 
@@
 
-Hand-written shopping lists undoubtedly have ongoing importance and display interestingly mixed ordering practices (see the montage of handwritten shopping lists at [Grocery List](http://www.grocerylist.org)).
-But online shopping lists take shape at the intersection of web and internet infrastructures, supply chain logistics, individualised practices and *habitus,* and increasingly, the predictive operations of  recommender systems.
-Just as the aisles and shelves of a supermarket present a densely-woven semiotics of objects competing for visual attention by offering distinctions of taste, thrift, expedience, novelty, indulgence, health and increasingly, online shopping recommender systems generate lists that seek to align people to commodities that they otherwise might have little relation to (see [@Turow_2015] for an overview of the development of these systems).
+Hand-written shopping lists have ongoing practical importance and mixed ordering practices (see the montage of handwritten shopping lists at [Grocery List](http://www.grocerylist.org)).
+Online shopping lists, by contrast,  reside at the intersection of web and internet infrastructures, supply chain logistics, individualised practices and *habitus,* and increasingly, the predictive operations of  recommender systems.
+Whereas the aisles and shelves of a supermarket present a densely-woven mesh of objects competing for visual attention by offering distinctions of taste, thrift, expedience, novelty, indulgence, health, online shopping recommender systems generate lists that seek to align people to products that they otherwise might have little relation to (see [@Turow_2015] for an overview of the development of these systems).
 
-If, as the social anthropologist Jack Goody argues, lists are historically primary as forms of writing in urbanizing cultures [@Goody_1986], and if list-making and its later variants (e.g.
+[^1.456]: If, as the social anthropologist Jack Goody argues, lists are historically primary as forms of writing in urbanizing cultures [@Goody_1986], and if list-making and its later variants (e.g.
 tables) precede discursive and narrative writing practices, then we might expect lists to function as powerful social ordering devices.
@@
 From both  anthropological and literary perspectives, shopping lists have excellent reasons to exist: they are deeply rooted in organisational life and infrastructures.
-Lists weave together people, infrastructures, things, and places.[^1.456]
-
-[^1.456]: While lists intersect strongly with other meaning and sense-making practices in everyday life and popular culture, they can be refractory to discourse  and textual analysis methods built around models of language or speech, with its rules of syntax and grammar.
+Lists weave together people, infrastructures, things, and places.While lists intersect strongly with other meaning and sense-making practices in everyday life and popular culture, they can be refractory to discourse  and textual analysis methods built around models of language or speech, with its rules of syntax and grammar.
 On the one hand, lists are highly associative.
@@
 
-Finally, and more formally, we might also consider some of the formal properties of lists as orderings.
-All lists imply order, but what kind of order?
-Mathematicians distinguish lists from sets on the basis of a distinction focused on order.
-Formally, both lists and sets are collections, but a list also contains a mapping of its elements to an ordering, usually the natural or counting numbers.
-Sometimes this order is explicit.
-Numbers are written down one side of the list.
-Often the order is implicit to the spatial arrangement of the list.
-This formal idea of the list as a mapping is helpful since it suggests relational possibilities.
-Mappings occasion relations between sets, collections, group or multiplicities.
-Mappings between list order and social order are manifestly highly contingent.
-A flight checklist differs greatly from a personal to-do list, even if they share many formal properties.
-Listing practices in a given social setting, therefore, always index an engagement between the formal properties of the list and the necessarily more fluid dynamics of social order.
-From the perspective of social order, shopping lists, as they are co-constructed by a supermarket chain's recommender system, might be important intersectional zones for the conversion of average everydayness through data-analytic propensities.
-
-TBA - short paragraph summarising the social theory of list ordering; make sure to add stuff on the ordering-remapping of order in the list;
-
 # D. Archaeology of recommendations:  from 1984  to 2007
 
-The main narrative of Patel's conversion narrative concerned the shift from a well-established loyalty-card based data-mining model to a predictive, probabilistic, personally relevant construction of the shopping list in almost-realtime.
+The main narrative of Patel's conversion narrative concerned the shift from a well-established loyalty-card based data-mining model developed in the 1990s to a predictive, probabilistic, 'personal relevance' model that would  append to the shopping list in almost-realtime.
 While the changes Patel described are configured in Tesco-specific ways by DunnHumby, they are also broadly typical.
-Tesco is the largest supermarket chain in UK, famous amongst retailers for its  customer loyalty and targeted marketing programme known as 'Tesco Clubcard,' which started in 1991.
+Tesco is the largest supermarket chain in UK.
+Its  customer loyalty and targeted marketing programme known as 'Tesco Clubcard' started in 1991.
 DunnHumby -- founded by operations researchers Edwina Dunn and Clive Humby -- is said to have convinced the CEO of Tesco sometime in 1991 that a loyalty card program could change the supermarket chain's relationship to its customers.
-Clive Humby's academic publications are hard to track down (see [@Humby_1989] where Humby argues for the need to add lifestage data to neighbourhood data in marketing research), but an early paper given at the _Conference of Young Operational Researchers_ in Nottingham in 1984 suggests the direction that he, the company Dunn and Humby formed (DunnHumby) and later Tesco would take in constructing lists [@OKeefe_1984].
+Clive Humby's academic publications are hard to track down.
+An early paper given at the _Conference of Young Operational Researchers_ in Nottingham in 1984 suggests the direction that he, the company Dunn and Humby formed (DunnHumby) and later Tesco would take in constructing lists [@OKeefe_1984].
 The abstract for Humby's presentation pre-figures an ongoing trajectory for data mining techniques aimed at eliciting detailed information on individual customer references.
@@
 
+In  [@Humby_1989], Humby argues for the need to add lifestage data to neighbourhood data in marketing research.
 While Tesco  succeeded in data-mining its customers using demographic segmentation,  and perhaps became the UK's biggest supermarket with the help of data-mining in the 1990s, the shopping environment  in 2017 is markedly different [@Turow_2017].
@@
 It can no longer rely on placement of goods in carefully chosen locations in stores.
-It needs or at least might want to continuously up-sell and cross-sell to customers who might hardly ever visit the supermarket itself.
+It needs, or at least might want to, up-sell and cross-sell to customers who only sometimes visit the supermarket itself.
 
 How could we characterise the shift from ClubCard data mining to online grocery markets?
-If 'Tesco is the clear winner in the online grocery market, in fact it takes almost 50p of every £1 spent on food shopping on the internet' [How SEO helps Tesco to dominate the online grocery market](https://econsultancy.com/blog/64841-how-seo-helps-tesco-to-dominate-the-online-grocery-market)[@Silverwood-Cope_2014], then has Tesco itself undergone some kind of conversion?
-Patel described her work at DunnHumby as converting the recommender system from a 'rules-based list' to a 'relevancy model.' The relevancy model changed how the 'Have You Forgotten' list was constructed.
-This shift is a typical of broader re-organisation of prediction in which existing data analysis practices are being re-distributed and intensified in particular ways.
+If 'Tesco is the clear winner in the online grocery market, in fact it takes almost 50p of every £1 spent on food shopping on the internet' [@Silverwood-Cope_2014], then has Tesco itself undergone some kind of conversion?
+Patel described her work at DunnHumby as converting the recommender system from a 'rules-based list' to a 'relevancy model.' The relevancy model affected the construction of the 'Have You Forgotten' list.
+This very localised intervention is typical of broader re-organisation of prediction.
 The shift in models results in a more probabilistic structuring of lists.
-The ostensible banality of the 'Have you forgotten?' recommendations betrays micrological signs of a topologically complex predictive infrastructure that intervenes in list-based social ordering to reconfigure alignments between people and things.
+As is so often the case in big data conversion narratives, the triviality of the 'Have you forgotten?' recommendations provides only peripheral signs of the complex predictive infrastructure underpinning them.
 
-This is not a drastic or abrupt event.
 Academic researchers first began writing about personalized recommender systems in the mid-1990s.
@@
 For instance, writing in 1997 in a special section of the _Communications of the ACM_ on recommender systems, Paul Resnick and Hal Varian (at that time Dean of Information Sciences at UC Berkeley, but currently Chief Economist at Google), made much of this personalization [@Resnick_1997].
-Resnick and Varian  emphasise the need to distinguish the emerging practices from data mining:
+Resnick and Varian  emphasised the need to distinguish the emerging practices from data mining:
 
@@
 
-# E. Experimenting with probabilistic conversions
+# E. Probabilistic conversion events
 
-This section of the paper returns directly to Patel's presentation, and the character of the newly implemented  recommender system.
-In the discussion that follows, the description of components of the new recommender system -- predictive models and their parameters, infrastructural provisioning, product similarity matrices --  and the problems of  recommendation -- too many products, changes over time, unstable propensities of things in their associations with people -- together suggest that personalization, and the 'personal relevance model' have a troubled relation to social order.
-I develop an analysis centred on probabilities and processes of probabilisation -- rendering situations calculable as probabilities -- in exploring Patel's account of the recommender system.
-The motivation to focus on probabilities and probabilisation is explicitly philosophically reconstructive.
-It represents an attempt to reframe the operation of the recommender system in a way freed from lingering incompatibilities between calculation and social life [@Dewey_1957, 26] .
+Given personalized recommendations stretch back two decades, what in  the newly implemented  recommender system changes?
+The components of the new recommender system -- predictive models and their parameters, infrastructural provisioning to run the models, product similarity matrices --  address the challenges of personally-relevant recommendations -- too many products, changes over time, unstable propensities of things in their associations with people.
+The 'personal relevance model' and personalization more generally have a troubled relation to a social order because it is, as grocery shopping attests, intrinsically social.
+The model will need to render grocery shopping calculable in a way that somehow includes the associated social ordering and negotiation.
+The model, like many big data practices, predicates probabilities as a way to render the situation calculable.
 
-In order to do this, we need a working notion of probability.
-I draw directly on the work of Karl Popper for this.
-In an essay written towards the end of his career, Popper presents an alternative account of probabilities as real processes, in the same way that contemporary sciences might see forces as real [@Popper_1990].
-Somewhat counter-intuitively, Popper does not identify probability with either degree of belief (likelihood) or frequency of events.
+From the standpoint of probability, recommendations are conditional probabilities or probabilities whose calculation takes into the account the occurrence of other events.
+But what is a probability today?
+In exploring Patel's account of the recommender system, my focus probabilities seeks to re-frame the operation of the recommender system in a way freed from lingering incompatibilities between calculation and social life [@Dewey_1957, 26].
+In order to re-frame probability, I draw directly on the work of Karl Popper.
+In an essay written towards the end of his career, Popper presents a non-standard account of probabilities as real processes.
+He argues that probabilities have a physical reality equivalent to forces and fields [@Popper_1990].
+Against standard interpretations,  Popper does not identify probabilities with either degree of belief (likelihood) or frequency of events.
 Instead he suggests 'they should be regarded as _inherent in a situation_ ' [@Popper_1990, 14].
-On this account, probabilities are tendencies towards realization inherent in a situation.
+In his account, probabilities are tendencies towards realization inherent in a situation.
 Probabilities have or indeed are _propensities_, tendencies to realize the event (11).
-While this somewhat abstract philosophical account of probability as propensity might seem to be anchored far away from the concerns of online grocery shopping, Patel's presentation of the personal relevance model and the ways in which it seeks to calculate probabilities of purchase.
-The series of prototyping reconstructions I am about to discuss will explore the tenability of this rather radical alternative to regular narratives of personalization.[^65.1]
+While Popper's concept  of probability as propensity might seem remote from the concerns of online grocery shopping, it applies quite well to Patel's presentation of the personal relevance model and the ways in which it seeks to calculate probabilities of purchase.
+The reconstructions of the model I am about to discuss will explore the implications of this alternative account of probability for regular narratives of personalization.[^65.1]
 
@@
 
-Popper suggests that one of the reasons probability has been difficult to work with philosophically, even as it pervades scientific thought, is that a certain limited domain of probability practice focused on dice rolls, coin tosses, urns with balls and other seemingly random events has occupied centre stage in concepts of probability.
-Even if it has been enormously productive and transformative, it is insufficient because it privileges _absolute_ probability at the expense of _conditional_ probabilities.
-'We need,' Popper urges, '_a calculus of relative or conditional probabilities_ as opposed _a calculus of absolute probabilities_' [@Popper_1990, 16].
+Probability has been difficult to work with philosophically according to Popper because  dice rolls, coin tosses, urns with balls and other seemingly random events have occupied centre stage.
+Even dice-rolling and coin-tossing has been enormously productive and transformative in scientific thought and practice, it privileges _absolute_ probability at the expense of _conditional_ probabilities.
+'We need,' Popper urges, '_a calculus of relative or conditional probabilities_ as opposed to _a calculus of absolute probabilities_' [@Popper_1990, 16].
 Relative or conditional probabilities are propensities that depend on other events for their own realization.
-All events require other events, so all probabilities are conditional, even if probability calculations have typically preferred to determine  the weight of possibilities devoid of the inevitable physical conditioning of their realization.
+All events require other events, so all probabilities are conditional, even if probability calculations typically abstract or ignore the inevitable physical conditioning of their realization.
 
-Viewed from the standpoint of probabilities, many predictive systems are highly crafted arrangements for the calculation of conditional probabilities.
+Viewed from the standpoint of probabilities, predictive systems are highly crafted arrangements for the calculation of conditional probabilities.
 For instance, the first element of the new recommender system consisted in a change of the underpinning algorithms and model.
-Patel described a move from 'a rules-based system'  to  a 'personal relevance model' as the basis of the new recommender system.
-It is likely that what Patel describes as  the 'rules-based system' refers  to the extremely well-known `apriori` algorithm or association rules learning technique, developed  by computer scientists Rakesh Agrawal and Ramakrishnan Srikanti working at IBM Research Alameda in the early 1990s [@Agrawal_1994].
+Patel described  move away from 'a rules-based system'.
+It is likely that what Patel describes as  the 'rules-based system' refers  to the extremely well-known association rules or `apriori` algorithm  learning technique, developed  by computer scientists Rakesh Agrawal and Ramakrishnan Srikanti working at IBM Research Alameda in the early 1990s [@Agrawal_1994].
 A now-classic approach to 'market basket analysis,' it was listed as one of the top ten data mining algorithms in a survey conducted amongst data miners [@Wu_2008] and usually attracts a chapter in data-mining and machine learning textbooks (e.g.
-[@Hastie_2009]).
+[@Hastie_2008]).
 The interest of `apriori` for our purposes is that it begins to address the problem of understanding large numbers of shopping transactions as a matter of conditional probability.
@@
 
-The notion of conditional probability at work in `aprior` is relatively simple, and assumes that the frequency of transactions provides the best guide to what shoppers are likely to do.
+The notion of conditional probability at work in `aprior` is relatively simple, and assumes that the frequency of transactions provides the best guide to what shoppers are likely to buy.
 The `apriori` algorithm finds sets of items that commonly occur together in transactions.
 In this sense, is still oriented by a notion of absolute probability, inflected by some elements of conditional probability.
-Commonly occurring sets are expressed as 'association rules.' For instance, using the `Groceries` dataset in the `R` package `arules,`  the `apriori` algorithm counts frequencies of purchase in the overall set of all items purchased in a supermarket (the `Groceries` dataset  was acquired from a 'local German supermarket' [@Hahsler_2006a]).
+Commonly occurring sets are expressed as 'association rules.' For instance, applied to a dataset of generic, unbranded grocery purchases, the `apriori` algorithm counts frequencies of purchase in the overall set of all items purchased in a supermarket (the `Groceries` dataset  was acquired from a 'local German supermarket' [@Hahsler_2006a]).
 Figure \@ref(fig:arules-ex) shows how often the most common items appear.
@@
 Although the `apriori` model does not use the language of probability calculus, the 'rules,' a term derived from older decision support literature, rank associations between items.
-When we run the `apriori` algorithm on the `Groceries` dataset, the resulting set of rules (shown in Table \@ref(tab:arules-ex2)) indicates that milk and vegetables has a stronger association than milk and buns.
+The resulting set of rules (shown in Table \@ref(tab:arules-ex2)) indicates that milk and vegetables has a stronger association than milk and buns.
 The calculations of `support` (how frequently the association appears in the dataset), `confidence` (how often the rule applies in the dataset)  and   `lift` (a ratio between the `support` and the independent frequencies of the items) in the table attempt to measure the strength of these associations in different ways.
@@
 
-Even as `apriori`  elicits associations between things, it struggles with the propensity of commodities to multiply.
-A simple illustration of the combinatorial problem faced by recommender system can be developed by bringing the `Grocery` dataset together with the actual list of items that sells in its online grocery shopping site.
-If we take all the items in the `Grocery` dataset and paste them into the 'shopping list' box on the Tesco grocery website (or as I did, run them as searches on the TescoLab Product API [@Tesco_2016]), each of the 169 items in the `Grocery` dataset yields dozens and sometimes thousands of products from Tesco online.
+Even as `apriori`  expresses associations between things as relative probabilities, it struggles with the propensity of commodities to multiply, especially in supermarkets and grocery shopping.
+A simple illustration of the combinatorial problem faced by recommender system can be drawn by bringing the `Grocery` dataset together with the actual list of items that Tesco sells online.
+If we take all the items in the `Grocery` dataset and paste them into the 'shopping list' box on the Tesco grocery website (or as I did, run them as searches on the TescoLab Product API [@Tesco_2016]), each of the 169 generic items in the `Grocery` dataset yields dozens and sometimes thousands of products in the Tesco inventory.
 
@@
 The 169 items of the `Groceries` dataset  expand into roughly `r total` Tesco items (see Figure \@ref(fig:tesco-ex)).
-There is, I would suggest, a logistic proliferation that recommender systems encounter.
-The multiple listed  in the `Grocery` dataset becomes more open in this setting through new impersonal intersections that could find themselves in associations of various strengths.
-If we think just of the almost 2500 sauces or 1200 `rice` products supplied by the Tesco API,  then a tremendous number of associations between sauce and rice are possible.
-The propensities of actual sauce and rice products to find themselves together in an individual purchase vary greatly.
-The proliferation of things on the shelves of supermarket or grocery warehouse confronts produces a combinatorial problem for data-mining machine learning approaches such as association rules.
-`r total` products (actually Patel mentioned 200,000 products) can be combined in a vast number of ways.
+Recommender systems confront, I would suggest, a logistical proliferation of commodities.
+The association rules derived  from  the `Grocery` dataset becomes more open to identifying sets of items that have only low association with each other.
+Given almost 2500 `sauces` and 1200 `rice` products listed by the Tesco API,  a tremendous number of associations between sauce and rice are possible.
+The propensities of any given sauce and rice product to find themselves together in a shopping basket greatly.
+The proliferation of things on the shelves of supermarket or grocery warehouse produces a combinatorial problem for data-mining machine learning approaches such as association rules.
+`r total` products (actually Patel mentioned 200,000 products) can be combined in a many ways.
 If a typical shopping list has 20 items, then there are `r format(choose(total, 20), 2)` possible lists.
-Some of this vast number of possible shopping lists have propensities or tendencies to realization that are close to zero.
-But others with somewhat higher propensities might furnish the basis of interesting recommendations.
+Most possible shopping lists have propensities or tendencies to realization expressed as probabilities close to zero.
+Others with somewhat higher propensities might furnish the basis of interesting recommendations.
 
-More importantly, the limited conditioning of probability of the association rules approach suggests a compelling reason for personalization.
-Even if the association rules and the `apriori` algorithm   generate recommendable sets of frequency-weighted associations, an `apriori`-based recommender system contains no statistical model.
+More importantly, association rules limit the conditioning of probabilities to counting frequency of co-occurrence of items in shopping baskets.
+Even if the association rules provide recommendable sets of frequency-weighted associations, an `apriori`-based recommender system has no way of individualising its recommendations.
 Its probabilisation of shopping is incomplete since it only works on associations between things.
-The tendency of some milk to find itself in a shopping basket alongside bread attests to an important commonality of practice in certain parts of the world.
-But the conditional probabilities implicit in the model do not bring in much of the world and its teeming sieve of propensities moving towards realization.
+The tendency of some milk to find itself in a shopping basket alongside bread attests to an ordinary propensity in certain parts of the world.
+The conditional probabilities implicit in the association rules  do not, however, include much of the world.
 These associations are not trivial, but they are very open-ended.
-Put in terms of the big data conversion, the rules-based system lacks, in principle, a means of crystallizing a limited or enclosed set of possibilities.
+Put in terms of Popper's account of probability as propensity, the rules-based system has, in principle, limited means of crystallizing a limited or enclosed set of possibilities.[^135.1]
+
+[^135.1]: The fact that DunnHumby and Tesco use demographic data derived from government census and other sources to differentiate association rules does not significantly overcome this limitation.
 
@@
 
-Predictive models are central technical elements in many big data arrangements because they provide a way of constructing conditional probabilities that incorporate more of the situation in which any propensity -- to purchase a bottle of Vueve Cliquot champagne for instance -- is inherent.
-In the new recommender system described by Patel, the business goal is to extend the list of the items customers have selected for purchase with  a few well-chosen items.
-In order to do this, the list of items selected for purchase will be extended by recommendations that have, according to a predictive model, the most chance of 'conversion'; that is, actual purchase.
-The predictive model carries the burden of calculating the conditional probability of purchase given everything known about the situation at this moment in time.
+Predictive models have become central technical elements in many big data conversion events because they offer a way of operationally calculating probabilities that incorporate more of the propensities -- to purchase a magnum of French champagne for instance -- inherent in situations.
+In the new recommender system described by Patel, the business goal is to extend the list of the items customers have selected for purchase with  a few recommended items.
+In order to do this, the list of items selected for purchase will be extended by recommendations that have, according to a predictive model, the most chance of 'conversion' or actual purchase.
+Most recommender system designers assume that modelling 'personal relevance' is the best way to do this.
+The predictive model carries the burden of calculating the conditional probability of purchase given everything known about the person at a given moment in time.
+
+```{r models, fig.cap='Precision measurements for different machine learning models'}
+include_graphics('figure/patel_compares_models.png')
+```
 
 Patel introduced the new 'personal relevance model' with a data graphic familiar to machine learners and statistical modellers.
-(A sketch of her graph appears in Figure \@ref(fig:models).)  The graph plots the _precision_ -- the proportion of the recommended products that customers actually purchase -- for several different statistical models.
+(A sketch of her graph appears in Figure \@ref(fig:models).)   Patel assumed that the audience understood the working of the logistic regression model, and the bulk of her presentation concerned the obstacles and problems that arise in trying to personalize recommendations in ways that lead to the much-desired 'conversion events' or sales.
+The graph plots the _precision_ -- the proportion of the recommended products that customers actually purchase -- for several different statistical models.
 The graph indexes the 'causal efficacy' of the recommender system, its capacity to include and transform propensities or 'real potentialities' into operational events or purchases.
-Although she did not dwell on it, Patel used the graph to compare the previous rule-based recommender systems with some of the 'personal relevance model' alternatives -- logistic regression, random forests, gradient boosting and a few others -- in terms of their predictions and how those predictions turned out.
+Patel used the graph to compare the previous rule-based recommender systems with some of the 'personal relevance model' alternatives -- logistic regression, random forests, gradient boosting and a few others -- in terms of their predictions and how those predictions turned out.
 Patel dismissed most of the models quite quickly and focused only on one, the logistic regression model, which did as well or slight other than other alternatives.
 
-The models Patel so briefly mentioned are all 'classifiers,' predictive models that 'classify' particular events or outcomes by calculating their probability of membership of some class of event or outcome.
-Much of the core architecture of the machine learning classifier models received scant reference in Patel's presentation.
-It is most likely that in DunnHumby's work, the classifier at the heart of the recommender system calculates for a given customer the most likely products to be purchased.
-The classes of events are binary: `recommended` or `not recommended.` Products are allocated to one of these classes depending on their calculated probability.
+The models Patel mentioned are all 'classifiers,' predictive models that 'classify' particular outcomes by calculating their probability of membership of some class of outcome.
+Much of the core architecture of the machine learning classifiers received glancing reference in Patel's presentation.
+In DunnHumby's work, the classifier at the heart of the recommender system calculates for a given customer the most likely products to be purchased.
+The predicted classes are binary: `recommended` or `not recommended.` Products are allocated to one of these classes depending on their calculated probability.
 A probability greater than `0.5` typically would be `recommended.`  The logistic regression model generates probabilities of purchase for each product for each customer.
-
-The crucial point here is that the shopping list written by the customer is extended, quasi-invisibly, in the recommender system to potentially any of the 250,000 items in the Tesco inventory.
-But the basis of this re-writing of the grocery list is an intricate conditional probability statement.
-Even if Patel assumed that the audience understood the working of the logistic regression model, the bulk of her presentation concerned the obstacles and problems that arise in trying to probabilise recommendations in ways that lead to the much-desired 'conversion events' or sales.
+The fundamental  shift  here is that the classifier model extends the reach of  the recommendations generated for a customer to not only  the 250,000 items in the Tesco inventory but to any item, relationship, similarity or event that can be constructed as a variable in the classifier model.
+In this sense, the recommendations shown to a customer in the 'Have You Forgotten' list derive from an intricate conditional probability statement, woven from, as we will soon see, an open-ended field of relations.
 

### 2017-09-27 log.rmd
entered 29/32 page edits

@@
 - did some more -- p 22/34, li 153/222
+
+## Wed 27 Sep 2017 09:21:08 BST
+- plan for next 3 days:
+    - today: enter last 8 pages of refs & update ref_bibs - actually, didn't get anywhere near doing this;
+    - Thurs: rewrite abstract & conclusion; read the whole thing through
+    - Fri: submission
+- TODO: all change 'relevancy' to relevance; all change inverted commans to none for this paper
+- TODO: l.201 need to rethink the point of self-learning substitutes -- does it not fit with personalization?
+Or if it does, it might overflow that ...
+in its groupiness ..
+Found this sentence in abstract 'At the point of delivery, a customer may be offered a substitute for an unavailable item in their order.
+‘Self-Learning Substitutes’ is a technique for identifying suitable and affordable product substitutions in this situation, based on historical rejection rates.
+'

### 2017-09-27 reconstruction.rmd
entered 29/32 page edits

@@
 
-## Repeating sufficiently often to measure
+## Repeating sufficiently often to matter
 
-I will not discuss in detail the different dimensions of probabilisation implicit in the architecture of the new model.
-It is worthwhile, however, highlighting some key elements that point directly to the troubled notion of 'personal relevance' and personalization more generally.
-These include the problem of repetition, attempts to create new views of the data, experiments on the platform, and the temporalities of calculation.
+I will not discuss all the dimensions of probabilisation implicit in the architecture of the new model but highlight those elements that trouble the notion of 'personal relevance' and personalization more generally.
+These include the problem of repetition and temporality of calculation,  experiments on the platform,  and attempts to create new views of the data.
 
-One fundamental difficulty here is that the propensity of a customer to buy a recommended product at a particular point in time is changing.
-Popper's notion of probabilities as propensities as tendencies to realization inherent in a situation directly implies this.
-Measuring propensities and expressing these measurements as numbers between `0` and `1` becomes difficult because the situation is change and the propensities themselves change.
+One fundamental difficulty is that the propensity of a customer to buy a recommended product changes.
+Popper's notion of probabilities as tendencies to realization inherent in a situation implies this.
 As Popper observes, 'the propensities cannot be measured because the relevant situation changes and cannot be repeated' [@Popper_1990, 17].
-No doubt the contingencies associated with grocery shopping are legion, perhaps more so than other kinds of products (books, music, films),  and any attempt to measure propensities associated with particular commodities  will encounter many dynamic changes.
+Measuring propensities and expressing these measurements as numbers between `0` and `1` becomes difficult because as a situation changes,  the propensities themselves change.
+No doubt the contingencies associated with grocery shopping are legion, perhaps more so than other kinds of products (books, music, films),  and any attempt to measure propensities associated with particular commodities  will encounter many changes.
 
-One way in which recommender systems attempt to address the problem of the relational dynamics of propensities is by looking for repetitions.
+One way in which recommender systems attempt to address the problem of the relational dynamics of propensities is by assuming that purchases will be repeated.
 Patel mentioned that the new model uses  '52 weeks of data' for each customer.
-In including this data in the model, the assumption is that the probability that a specific customer will buy a specific product will be shaped by previous purchases.
+In including this data in the model, the assumption is that the probability that a specific customer will buy a specific product will be increased by previous purchases of that product.
 The history of previous purchases constitutes forms of repetition that imply a stabilisation of propensities (high or low; e.g.
 a vegan customer will never have purchased chicken products, so the measured propensities for any of the 1000 or so chicken products sold be Tesco will remain close to `0`).
+But the changing situation of the customer only figures here in the accumulated year of purchase data.
 
 When the logistic regression model includes the 52 weeks of previous purchases, the conditional probability calculation undertaken by the recommender system ramifies tremendously in several respects.
-Each of Tesco's 200,000 products becomes a variable in the classifier.
-Practically, most of the these variables will not influence the calculation of the probabilities for recommendations very much.
+Potentially, each of Tesco's 200,000 products becomes a variable in the classifier.
+We can imagine this as an arithmetic sum extending along a series of 200,000 terms.
+Practically, most of the these variables will only slightly influence the sum of the probabilities for recommendations since most customers will have purchased only a fraction of the inventory.
 As Patel observed,  again assuming that this would be obvious to the audience of data scientists,  'we have lots of zeros'.
--- points to the difficult terrain of  'good feature' thinking: the vast, nearly empty matrix of customer-product associations.
-Given that most things in Tesco remain relatively unknown to each other and to customers.
-A matrix that records associations between individual people and different products is bound to be mostly empty.
+A matrix that records associations between individual people and products is bound to be mostly empty.
 Say Tesco has 1 million online customers.
@@
 The product-customer matrix, the basic vector-space in which all recommender systems operate, remains very sparse and unpopulated.
-Given that any one customer is likely to only have bought 100 or so different products, the probability matrix will be close to `r format(100-(100/2e5)*100, 2)`% empty.
-The data is, as Patel put it, 'massively unbalanced' (any dataset where the items of interest are much rarer than some normal values is said to be unbalanced) and such imbalance would heavily bias the recommender system towards common and only impersonally relevant suggestions, suggestions that might  not produce the desired conversion experience.
+Given that any one customer is likely to only have bought 100 or so different products, the  matrix contain  `r format(100-(100/2e5)*100, 2)`% zero probabilities.
 
-```{r sparsity-ex, echo=FALSE, cache=TRUE, fig.margin=FALSE, fig.cap ='Sparsity of the customer-product association matrix' }
-library(ggplot2)
-library(reshape2)
-p = as.logical(rbinom(n=2e5, size=20,0.001 ))
-m = (matrix(p, nrow=1000, byrow=TRUE))
-melted <- melt(m)
-melted$value <- as.logical(melted$value)
-ggplot(melted, aes(x = Var2, y = Var1, fill = value)) + geom_tile() +
-    scale_fill_manual(values = c("grey", "red")) + xlab('Customer id') + ylab('Product id') + theme(legend.position='none')
-```
+Any dataset where the items of interest are much rarer than other values is said to be 'unbalanced.' The purchase history data is, as Patel put it, 'massively unbalanced,' and imbalance heavily biases the model towards common and somewhat impersonal suggestions, suggestions that might  not produce the desired conversion experience for either the individual customer or DunnHumby's renovation of Tesco's recommender system.
+Since so many people buy milk, the recommendation system might end up always recommending milk.
+So the data needs to be 'corrected' by, as Patel reports, removing -- 'under-sampling' -- some of the data for common purchases.
+'Having all the data,' one of the anchoring claims of big data conversion narratives, also creates the need to delete some data.
 
-Since so many people buy milk, the recommendation system might end up always recommending milk products.
-So the data needs to be 'corrected' by, as Patel reports, removing -- under-sampling -- some of the data for common purchases.
-Secondly, even if a recommendation system finds uncommon items that are good recommendations, they should neither be too cheap ('low spend'), nor, as we will see too similar to what the customer has already ordered.
+The problem of repetition, the fact that situations change and therefore propensities change too, is known to designers of recommender systems.
+But they do not have the ideal conditions envisaged in probability theory (unbiased coins, flipped any number of times).
+The model's own operation as a data-intensive calculation need to be adjusted to the scale of values of grocery retail, to the available computational, database and network infrastructures, and to the capacities of the online grocery system to inject recommendations into the flow of grocery orders in a timely fashion.
+While raw data from Tesco Online transactions  feeds into the model's dataset  every hour, a recommendation list for each customer is only generated once a week.
+Customers shop online every few days at the most, and in some cases, only every few weeks.
+To update the top 200 recommendations for several millions customers  demands much computation.
+Patel briefly mentioned  specific infrastructural elements such as `hadoop` [@ApacheSoftwareFoundation_2009].
 
-Many other slight adjustments, tweaks and local modifications would probably occur around the logistic regression model.
-The model's recommendations themselves have propensities that need to be adjusted to the scale of values of grocery retail, to the available infrastructures, and to the capacities of the online grocery system to merge recommendations into a grocery order in a timely fashion.
-While raw data from Tesco Online feeds into the model every hour, a recommendation list for each customer is only generated once a week.
-Customers only shop online every few days at the most, and in some cases, only every few weeks.
-To continually update the top 200 recommendations for every customer would be computationally expensive.
-Patel briefly mentioned  specific infrastructural elements such as `hadoop` [@ApacheSoftwareFoundation_2009].
-The logistics of running a predictive model affect listing practices.
-The possibility of adjusting the recommendations for every customer weekly derives from an infrastructure capable of collecting data, assimilating that data to a predictive model able to generate approximately 200 million relevant recommendations, where 'relevance' depends on a matrix of probabilities of associations between people and things that shifts in time.
-`Hadoop` and its legion of 'big data' variants (`mahout`, `spark`, `hive`, `pig`, `yarn`) operationalise flows of data at an infrastructural rather than analytic scale.
-In many respects, the quickly glossed-over infrastructural deployment of DunnHumby's relevancy model is the primary conversion event: the logistic regression model at the heart of the recommender system is no longer an analytical device but an operational one.
+The possibility of adjusting the recommendations for each customer every week depends on an infrastructure capable of collecting data, and assimilating that data in a predictive model.
+'Personal relevance' depends on a matrix of probabilities of associations between people and things that shifts in time.
+`Hadoop` and its legion of 'big data' variants (`mahout`, `spark`, `hive`, `pig`, `yarn`) operationalise repetition at an infrastructural rather than analytic scale.
+Patel's quick gloss of the infrastructural deployment of DunnHumby's relevancy model is  primary to the conversion event: the logistic regression model at the heart of the recommender system is no longer an analytical device but an operational one because it seeks to revise recommendations as situations change.
+[TBA: this paragraph is promising but doesn't deliver yet]
 
-## Platform Experiments that reduce interfering propensities
+## Platform experiments reduce interfering propensities
 
-In his account of probabilities as physical propensities, Popper emphasises the specific configurations established in laboratory experiments:
+In his account of probabilities as physical propensities, Popper emphasises why laboratory experiments are important to probability:
 
@@
 
-Like many accounts of laboratory practice [@Latour_1097; @Rheinberger, @Knorr-Cetina; @Lynch - epistopics ref], Popper understands experiments as stabilizing the conditions so that phenomena can be saved from their own variability.
+Like the accounts of laboratory practice provided by science studies [@Latour_1097; @Rheinberger, @Knorr-Cetina; @Lynch_1994], Popper understands experiments as stabilizing the conditions so that probabilities become less variable.
 Experiments play an important role in conversion events, or in the process of bring new predictive models to bear on shopping practice.
-A  crucial consideration for a predictive model in a recommender system will be how to experimentalize its own predictions so that they are not interfered with or disturbed by other propensities.
-This is obviously an inordinately difficult challenge, since, as we have seen, grocery shopping takes place at the intersection of normative and actual social orders, individual and group belongings, and logistically global economic processes.
+A crucial consideration for  a recommender system designer will be how to experimentalize the systems recommendations so that they are not disturbed by other propensities.
+This is an inordinately difficult challenge, since, as we have seen, grocery shopping takes place at the intersection of normative and actual social orders, individual and group belongings, and logistically global economic processes.
 
@@
 |Control A|Control B|
-|------|------|
+---------------------
 
-Nonetheless, experimentalization, the practice of creating conditions that reduce disturbing propensities, is very much part of the convernsion between models.
-For instance, the predictions of the recommender system themselves are the subject of experimental.
-Patel described the deployment of  the relevancy model experimentally in a random A/B controlled trial on the Tesco website.
-All customers were  allocated to one of four categories as shown in the table.
-In the  A/B trial,  an identical shopping list, or even the same person, might receive recommendations from different models.
+Experimentalization, the practice of creating conditions that reduce disturbing propensities, runs deep in the conversion event.
+The predictions of the recommender system themselves are the subject of experiment.
+Patel described the deployment of  the personal relevance model in a random A/B controlled trial on the Tesco website.
+All customers were  allocated to one of four categories as shown in  table.
+In the  A/B testing,  customers  receive recommendations from different models (the old recommender systems vs.
+the new one; a logistic regression model vs.
+a random forest model, etc.).
 The randomised application of  competing predictive models, draws on protocols for randomised clinical trials first developed in the 1960's, and is widely used in social media platforms and hence in the implementation and observation of the effects of recommender systems.[^238.1] Random allocation of customers to the four categories (Test A, Test B, Control A, Control B) adds a layer of probability to the recommender system in the name of statistical validation of the effects or the 'uplift' of the model.[^122] Ironically,  the effects of a predictive model cannot be known in advance.
@@
 
-[^238.1]: It is difficult to gauge how much. Facebook reports that it has more than 1 million models operating in its infrastructure. [@Dunn_2016]
+[^238.1]: It is difficult to gauge how much.
+Facebook reports that it has more than 1 million models operating in its infrastructure [@Dunn_2016].[@Christian_2012] provides information on the extent of A/B testing in web and social media.
 
-It is perhaps significant that this random allocation of customers to control and test groups occurs without any relation to the particular profile or propensities of the customer.
-It seeks to statistically validate the effects -- the *uplift* -- of the model on  conversion events by allowing the effects of the model on what people do to be measured.
+Random allocation of customers to control and test groups occurs without taking individual propensities into account.
+A/B testing  seeks to statistically validate effects -- the *uplift* -- of the model on  conversion rates by directly measuring the effects of the model on what people do.
 The uplift refers to conversion events associated with the same groups of people.
-Effectively an experiment in inhabiting two different worlds, the randomised control trial  sets up a feedback look between the predictive model (the logistic regression), and the world.
+Effectively an experiment in creating micrologically different worlds, the randomised control trial  sets up a control mechanism that connects  the predictive model (the logistic regression), and the conversion event more broadly.
+With this experimentalisation, the conversion event narrative lacks grounding.[@TBA: not happy with this ending]
 
@@
 
-Whatever disturbing factors or interferences the experimental trials of different predictive models reduce, a broader and ineluctable stream of propensities remains in operation.
-Popper's account of probabilities as propensities presents openness as concomitant to realization:
+Whatever disturbing factors or interferences the experimental trials of different predictive models reduce, another stream of propensities runs away from the goal of personal relevance.
+Popper suggest that the realization of tendencies is open-ended:
 
->What may happen in the future ..
+>What may happen in the future ...
 is, to some extent, open.
@@
 There are different ways of reading this statement.
-We can read Popper as stating a truism,  the obvious consequence of his physicalist understanding of probabilities as tendencies inherent in a situation: anything can happen in the future.
-An alternative, perhaps more interesting reading  focuses on the pivotal phrase: 'possibilities trying to realize themselves.' What does 'trying to realize' might mean in practice?
-I would suggest that DunnHumby's work on the new personal relevance model, and much big data practice in general,  contains just such a 'trying to realize' in practice.
+We can read Popper as stating a truism,  an obvious consequence of his physicalist understanding of probabilities as tendencies inherent in a situation: anything can happen in the future.
+An alternative, perhaps more interesting reading  focuses on the pivotal phrase: 'possibilities trying to realize themselves.' What might 'trying to realize' mean in practice?
+DunnHumby's work on the personal relevance model, and much big data practice in general, presents just such a 'trying to realize' in practice.
 
-Several times in her presentation, Patel emphasized the importance of 'constructing good features' from the data, and much of her presentation concerned DunnHumby's efforts to construct 'good features.' A 'feature' in the context of machine learning refers to a variable included in a predictive model.
-A 'good feature' contributes to the accuracy, precision, specificity or any of the panoply of measures applied to machine learning models in practice.
-The construction of 'good features', however, remains open to many different possibilities, some of which are more practically feasible than others.
+Several times in her presentation, Patel emphasized the importance of 'good features' in the data, and much of her presentation concerned DunnHumby's efforts to construct 'good features.' A 'feature' in the context of machine learning and predictve modelling refers to a variable included in a predictive model [@Domingos_2012].
+A 'good feature' contributes to the accuracy, precision, specificity or any of the other measures of prediction applied to machine learning models in practice.
+The construction of 'good features', however, remains open to many different possibilities, some of which are more practically feasible than others, and some of which are more aligned with personal relevance than others.
 
-The main efforts that DunnHumby made to  construct good features were not closely associated with personhood, but with the relations -- the propensities -- between things.
-Patel, for instance, addressed the problem of 'basket similarity.' The recommender system should not recommend items that are too similar to groceries already in a customer's basket.
-A customer might be willing to substitute a similar item for something they have already chosen, but they are more likely to accept a recommendation that complements rather than duplicates items in their list.
-How could the recommender system sense the possible similiarities and complementaries between things in general or for a given customer, especially since a person's sensibilities and susceptibilities concerning groceries will be subtly shaped by wider social groups, orderings and circumstances?
+The main efforts that DunnHumby made to  construct good features were not closely focused on individuals  but sought to include  relations between things in the predictive model.
+Patel, for instance, described the problem of 'basket similarity.' The recommender system should not recommend items that are too similar to groceries already in a customer's basket.
+A customer might be willing to substitute a similar item for something they have already chosen, but they are more likely to accept a recommendation that complements already selected items.
+How could the recommender system avoid similiarities and  prefer complementaries between things for a given customer, especially since a person's sensibilities and susceptibilities concerning similarity and complementarity are shaped by social groups, orderings and circumstances?
+A predictive model could only do that if it had some sense of the relation between items already in the list and the products elsewhere.
 
-The predictive model could only do that if it had some sense of the relation between items already in the list and the field of possible recommendations.
-The 'personal relevance' model, therefore, sought to include the similarities and complementarities between items already ordered and recommendations.
-The DunnyHumby data scientists constructed a new feature from the data measuring substitutability and complementarity between products.
-Taking all the baskets of items purchased, a new feature, what DunnyHumby termed 'self-learning substitutes' was derived.
-Added to the relevancy model, the 'self learning substitutes' feature adds another data matrix, the product similarity matrix, itself generated from all previous recommendations that have led to conversion events (that is, 'a user clicked as the response').
-A customer's recommendation list might already include `Tesco Ginger Nuts 300g.`   The 'self-learning substitution model' seeks to replace list items with similar but higher value items: `Mcvitie's Ginger Moments 400g.`
+The 'personal relevance' model sought to include the similarities and complementarities between items.
+The DunnyHumby data scientists constructed a new feature from the previous purchase data measuring substitutability and complementarity between products.
+Taking all the baskets of items purchased on the basis of recommendations, they derived a  new feature, what DunnyHumby termed 'self-learning substitutes.' Added to the relevancy model, the 'self learning substitutes' feature adds another data matrix, the product similarity matrix, itself generated from all previous recommendations that have led to conversion events (that is, 'a user clicked as the response').
 The self-learning substitutes feature is complex.
-Patel mentioned that it took the form of a 'design matrix of 14,000 columns.' 14,000 new explanatory variables were added to the logistic regression model producing the recommendations.
-Each recommendation would be subtly re-weighted by this complex derived feature.[^215]
+Patel mentioned that it took the form of a 'design matrix of 14,000 columns.' 14,000 new explanatory variables were added to the logistic regression model.
+Recommendation would be subtly re-weighted by this complex derived feature.[^215]
 
@@
 
-The personalization of recommendations has been a defining characteristic of many big data conversion events.
-The exploitative and discriminatory effects of all-consuming personalization have been the target of critics and critical research on data economies.
+The personalization of recommendations has been a distinctive characteristic in many big data conversion narratives and practice.
+The exploitative and discriminatory effects of personalization have rightly been the target of critics and critical research on data economies.
 The archaeology of the Tesco recommender system, including its early experiments with demographic data-mining, its adoption of a rule-based 'market basket' analysis system, and its current implementation of a 'personal relevance model' follows the trajectory of personalization.

### 2017-09-28 log.rmd
51/221 lines in reread

@@
 - redid the introduction again, take out lots of adjectives
-- TODO: rewrite abstract and conclusion
+- DONE: rewrite abstract and conclusion
 - DONE: maybe should take out a lot of the list discussion. Is it too distracting?
@@
 '
+
+## Thu 28 Sep 2017 09:24:50 BST
+- that self-learning substitutes thing is a bit different to what I expected 
+- went through the conclusion and the introduction, and indeed started to go through the whole thing. 
+- at l.55/222 in going through again this morning.

### 2017-09-28 reconstruction.rmd
51/221 lines in reread

@@
 Personalization, with all the risks of discrimination and bias it entails, has been the critical focus in accounts of consumption, government, social media and health.
-This paper argues that personalization occurs against a more expansive change in probabilization that,  in principle, is not reducible to individual attributes and actions.
-It describes the 'personalization' of an online grocery shopping recommender system to list a small number of grocery items of personal relevance for each of the millions of online grocery shoppers at a  a major UK supermarket chain.
-Drawing on a theory of probability proposed by the philsopher of science Karl Popper and anthropological work on shopping, it suggests that personal relevance is caught up in relations to others and things.
+This paper argues that personalization is part of a more expansive change in probabilization that, in principle, is not reducible to individual attributes and actions.
+It describes the 'personalization' of an online grocery shopping recommender system to list a small number of grocery items of personal relevance for each of the millions of online grocery shoppers at a major UK supermarket chain.
+Drawing on a theory of probability proposed by the philosopher of science Karl Popper and anthropological work on shopping, it suggests that the attempt to generate  personalized predictions changes relations to others and things.
 Using a mixture of discourse analysis and code-based reconstruction of key elements of the recommender system, it suggests that personalization is one facet of an open-ended weave of propensities associated with people and things in contemporary big data configurations.
@@
 
-## Introduction
+# A.  Introduction
 
@@
 This paper recounts the conversion of  an online grocery ordering system  from an older demographically-based recommendation system to a 'personal relevance' model.
-The re-counting is empirical in certain respects.
+The account is empirical in several respects.
 It departs from an ethnographic moment: being a member of the audience at an industry/academic conference where a data scientist was describing how personalized recommendations were created for online grocery orders.
-Starting from this highly mediated and fragmentary encounter with a recommender system designer, the paper  draws on anthropological and sociological research concerning shopping and list-making to situate the recommender system amongst the social ordering practices of shopping and lists.
+The paper  draws on anthropological and sociological research concerning shopping and list-making to situate the recommender system amongst the social ordering practices of shopping and lists.
 It makes use of archaeological approaches, in the sense developed by Michel Foucault [@Foucault_1972] to identify and map functional statements and practices configuring the knowledge and order generated by such systems over time.
 It conducts several small-scale code-based experiments in order to reconstruct, using widely available code resources resources such as API (Application Programmer Interfaces) and software libraries for machine learning, some prototypical elements of the system in question.
-It draws  on the philosophy of science and media theory to frame an engagement with grocery recommendations in terms of probabilities and their worldly enactments.
+Finally,  it  draws  on the philosophy of science to frame an engagement with grocery recommendations in terms of probabilities and their worldly enactments.
 
-The main argument of the paper concerns how  personalization occurs in Big Data 'conversion events.' The term 'conversion event'  refers to the common before-after narrative forms that proponents, developers and advocates of these systems adopt.
+The main argument of the paper concerns how  personalization occurs in big data 'conversion events.' The term 'conversion event' has a dual sense.
+In recommender system design and web ecommerce, it refers occasions when a customer, user or site visitor clicks or purchases.
+I also use it to   refer to the common before-after narrative forms that proponents, developers and advocates of recommender systems adopt in describing, promoting, explaining or otherwise talking about big data.
 In recent research and opinion on data practice, personalization has been seen as the linchpin of predictive modelling.
@@
 'Data-driven personalization' relies on predictions, themselves dependent on data practices ranging across acquisition, storage, transformation, exchange, integration, modelling and experimentation.
-The reconstruction of the recommender system and its  personalized predictions I undertake here focuses on the notion of probability and its role in modelling and experimentation.
+The reconstruction of a recommender system and its  personalized predictions I undertake here focuses on the notion of probability and its role in modelling and experimentation.
 Although probabilistic calculations of events have  long-standing importance in many settings (insurance, medicine, experimental and field sciences, operations research, risk analysis, engineering design, public health,  economic modelling, etc.), probabilities have taken on an operational function in digital platforms associated with the web and social media.
 They have become part of infrastructures, transactions, and event-structures.
-Taking their their _operation_ reality into account, we might understand probabilities in quasi-physicalist terms  defined by the philosopher of science Karl Popper as a 'world of propensities, as an unfolding process of realizing possibilities and of unfolding new possibilities' [@Popper_1990,19].
+In order this _operational_ reality into account, we might understand probabilities in quasi-physicalist terms  defined by the philosopher of science Karl Popper as a 'world of propensities, as an unfolding process of realizing possibilities and of unfolding new possibilities' [@Popper_1990,19].
 This operational-physicalist or sociotechnical notion of probability shifts our understanding of data analytics, especially in association with platforms [@Gillespie_2010], away from personalization.
@@
 Like some recent work in science and technology studies,  anthropology and media studies [@Marres_2017; @Marcus_2014; @Bogost_2012], the paper allows its own reconstruction to be provisionally shaped by an experiential encounter with the object it describes.
-At times, this mode of empirical philosophy may seem fixated on the technical minutiae of groceries and recommender systems, and lacking in  sociological rigour or empirical evidence.
+At times, this mode of empirical philosophy may seem fixated on the technical minutiae of algorithms and recommender systems.
 With some forbearance on the part of readers, however, the argument of the paper should resonate sociologically.
-It concerns social order, or the consistency and regularity of relating and acting amid a world in where probabilities and calculations based on probabilities play a pervasive role.
+It concerns the relationality of social life, or the consistency and regularity of relating and acting amid a world in where probabilities and calculations based on probabilities play a pervasive role.
 
@@
 
-At one of the many industry-meets-academia events occurring in data science-oriented higher education institutions in the UK, speakers from industry, government and commerce described their work with predictive models.[^11.1] Shreena Patel, a PhD graduate in statistics and operations research,  works as a data scientist for DunnHumby, a well-known [ customer science company ](https://www.dunnhumby.com/) [@dunnhumby_2017].
+At one of the many industry-meets-academia events occurring in data science-oriented higher education institutions in the UK, speakers from industry, government and commerce described their work with predictive models.[^11.1] Shreena Patel, a PhD graduate in statistics and operations research,  works as a data scientist for DunnHumby, a well-known [customer science company ](https://www.dunnhumby.com/) [@dunnhumby_2017].
 Her work at DunnHumby focuses on online grocery shopping at the supermarket chain Tesco.
@@
 Second, Patel focused on the renovation and updating of long-standing data-mining practice via a much more explicitly 'big data' and 'machine learning'- oriented implementation.
-Her presentation concerns what we might term, for reasons that will gradually emerge, a 'conversion event' in which a long-standing recommender system was replaced by a 'big data'-style system concerned with 'personally relevant' recommendations.
+Her presentation concerns what  a typical big data conversion event  in which a long-standing recommender system was replaced by a 'big data'-style system delivering 'personally relevant' recommendations.
 What stands out from the presentation is the normality of the recommender system: it is part of a long-standing and ongoing transformation of grocery shopping.
@@
 The recommendations are the product of a recommender system, an important category of operational device in big data (see for instance [@Hallinan_2014] for analysis of Netflix recommendation; [@Morris_2015]  or [@Seaver_2015] for an account of music recommendations).
-The question 'have you forgotten?' is followed by a list of some  grocery items that could have been or are usually on a shopping list.
+The question 'Have you forgotten?' is followed by a list of some  grocery items that could have been or are usually on a shopping list.
 The title of the suggestions is a bit misleading.
-The recommender system, as we will see, is not concerned with forgetting, with the many slips and oversights associated to which shopping is prone,  but rather with items that customers had not selected perhaps because they had never thought of buying them in the first place.
+The recommender system, as we will see, is not concerned with forgetting, with the many slips and oversights associated to which shopping is prone,  but rather with substituting and adding items that customers had not selected perhaps because they had never thought of buying them in the first place.
 
+Grocery shopping is difficult to personalize.
 The anthropologist Daniel Miller has argued [@Miller_2012] that  shopping negotiates discrepancies between normative and actual social order (for instance, between the ideals of health emblematised by organic products and commitments to thrift embodied in lower cost generic products).
@@
 More provocatively, Miller claims that 'shopping is largely a technology for the expression of love' (85).
-It is, to use the language of actor-network theory, a practice of translation [@Callon_1991].
-Whatever theory we have of how normative social order arises, the key implication is that shopping inhabits a mundane  but highly variable space between normative/idea and actual order.
-Importantly, and this will be a key consideration for recommender systems, shopping is not necessarily personal or individual.
-Grocery shopping in particular is saturated  by fluxing forms of social order concerning family and other forms of social grouping, and their associated relations (love, etc.).
+Whatever theory we have of how normative social order arises, Miller's argument  implies that shopping inhabits a mundane  but highly variable space between normative/idea and actual order.
+Importantly, and this will be a key consideration for the Tesco recommender systems, grocery shopping is not necessarily personal or individual.
+It  is saturated  by fluxing forms of social order concerning family and other forms of social grouping, and their associated relations (love, etc.).
 
@@
 
-## `Apriori` and conditional probabilities
+## `Apriori` conditional probabilities
 
@@
 
-## The list in its mapping to probabilities
+## The list as relational field
 
@@
 
-The personalization of recommendations has been a distinctive characteristic in many big data conversion narratives and practice.
-The exploitative and discriminatory effects of personalization have rightly been the target of critics and critical research on data economies.
-The archaeology of the Tesco recommender system, including its early experiments with demographic data-mining, its adoption of a rule-based 'market basket' analysis system, and its current implementation of a 'personal relevance model' follows the trajectory of personalization.
+The personalization of recommendations has been a distinctive feature of big data conversion narratives and practice.
+The Tesco recommender system, starting from its early experiments with demographic data-mining, its later adoption of a rule-based 'market basket' analysis system, and its recent implementation of a machine learning 'personal relevance model' follows the trajectory of personalization.
+Its conversion event largely takes the form of personalization.
+Only because grocery shopping is  personally pre-ordered by lists can a predictive recommendation, provisionally conducive to new conversion events, gain traction.
+Individualizing personalization, however,  is ill-fit to the socially complex negotiations of grocery shopping, and tends to obscure other implications.
+Conversion events, both in the sense of the purchases made on the basis of recommendation and the narratives of changes in modelling practice related by data scientists, do not easily map to personalization.
+They include relations running between people and things in time.
+They include an ongoing process of adjusting, scaffolding, intervening and configuring orders.
 
-The fact that the conversion event largely takes the form of personalization,  an individualizing personalization that is so patently ill-fit to the socially complex negotiations of grocery shopping, tends to obscure the wider re-framing of social order taking place.
-on  the mappings that inscriptive devices such as shopping lists  display.
-Only because shopping is already somewhat ordered by lists can a partial re-mapping, provisionally conducive to new events, take place.
-The conversion entailed in this re-mapping is not just about personalization.
-It is more ontological than that since it includes relations running multi-laterally between people and things in time.
-It is less ontological, however, in that it is not a sudden transformation, but an ongoing process of adjusting, scaffolding, intervening and configuring orders.
-
-But the conversion event associated with big data is not reducible to personalization or the individual.
-Probabilization precedes and exceeds personalization.
-When we observe what is practically done to construct a personal relevance model and apply it to a shopping list, we see conditional probabilities, predictive models, the importance of data accumulated over time,  platform experiences, the rhythms of infrastructural maintenance, and open-ended construction of new features in data, all of which shift and complicate the conditional associations between people and things, people and people, and things with things.
-If social order is made of associations, if to be social is to associate [@Latour_TBA], then the associations taking shape in this and other similar big data conversion events might be understood as operationalizing probability in matrices of propensity.
+I suggest that we might understand what precedes and exceeds personalization as probabiliization.
 If, as Popper argues, probabilities are being realized in the same way that 'fields' were realized in modern physics, then we should see the compute  platforms, databases, software libraries, web interfaces and global supply chain logistics as part of this conversion event.
-
-Reconstructions of specific conversion events associated with big data can work in different ways.
-The combination of social theory, archaeology, philosophy and model/data prototypes used in this paper sketches one possible way of addressing the trouble, the disturbance or instability that seems to affect many discussions of big data.
+When we observe what is practically done to construct a personal relevance model and apply it to a shopping list, we see conditional probabilities assembled in predictive models, the rhythms of  platform and infrastructural configurations in gathering and processing data,  and the open-ended construction of features in data.
+In various ways, they shift and complicate the  associations between people and things, people and people, and things with things.
+If social order is made of associations, if to be social is to associate, then big data conversion events operationalize association in matrices of propensity.
 

### 2017-09-29 log.rmd
redoing TBAs and intro

@@
 - DONE: look at recent articles in bdas and collect any relevant ones; have been through 100 or so -- best candidates -- Kitchin,  Beer,  added some to lists subcollection in data_forms collection in zotero
-- TODO: look again at Amoore in Econ&Society, add Peters on List; could add List as evil media;
+- DONE: look again at Amoore in Econ&Society, add Peters on List; could add List as evil media;
 - Going through notes from the siegen event -- reminded me of
@@
 Also state an empirical limitation in the project -- commercial sensitivity -- and how I deal with it; could link this back to the forensic book on shelves at home
--TODO also thinking about diapers and beer again. Could I make that into the headline story? An anchor point that I set out to reconstruct? But hasn't that been done? Should look up diaper on bdas.
+- DONE also thinking about diapers and beer again. Could I make that into the headline story? An anchor point that I set out to reconstruct? But hasn't that been done? Should look up diaper on bdas.
 - finished paper revisions at 10.10. Still need do abstract; and many todos above on references, of various kinds. Especially Morris, Peters, etc.  
@@
 It operates within the model itself, with the 'need for good features.'
-- TODO: footnote how DunnHumby itself markets to customers.
+- DONE: footnote how DunnHumby itself markets to customers.
 There is a regress, perhaps not infinite, but multi-stepped that runs from customer analytics; selling customer analytics; the images of analytics; the selling of images of analytics; etc.
@@
 - have added some Popper into intro and conclusion (too implicit?). Feels like I need to go through again and check the argument: which is now implicitly: conversion comes from propensities; .
-- TODO: check the overall argument by going through introduction, abstract and conclusion
+- DONE: check the overall argument by going through introduction, abstract and conclusion
 - DONE: check section headings map the overall argument
@@
 - need to update this log with stuff about reports;
-- TODO; notes from Dewey about reconstruction are here ~/notes/dewey_reconstruction_1957.md; and the full text is in ~/archive/dewey_reconstruction.txt;
+- DONE; notes from Dewey about reconstruction are here ~/notes/dewey_reconstruction_1957.md; and the full text is in ~/archive/dewey_reconstruction.txt;
 - Dewey:  grepped through the full text; the basic idea is that reconstruction is the way that experience can be intelligently re-configured; it draws on the  medium it inhabits, but changes/transforms elements of that medium, in the same way that any living thing does.
@@
 ## Wed 12 Jul 2017 14:29:19 BST
-- historical precedent -- Tesco shopping 1984 http://www.bbc.co.uk/news/magazine-24091393 [@Winterman_2013]  - TODO -- add in
+- historical precedent -- Tesco shopping 1984 http://www.bbc.co.uk/news/magazine-24091393 [@Winterman_2013]  - DONE -- add in
 - reading through the reports, and summarizing here what they are asking
@@
 - how do I actually use the social theory of lists - what happens to the mapping? to the discrepancies between normative and actual, and to the siting of the list?
-- TODO: write about this in the conclusion
+- DONE: write about this in the conclusion
 - remapped last 12 pages a bit, taking into account the Popper argument and critiquing the Hansen a bit more;  see page 21 of July printout for summary; still feels like need more experiments or prototypes; particularly for the model itself;
@@
         - DONE - at least in part: update the theory of shopping to reflect this
-        - TODO: exemplify it in a number of ways
+        - DONE: exemplify it in a number of ways
 - although I didn't do as much, paper felt better because I've managed to localize down on the logistic regression model; it helps me combat the personalization problem better.
@@
 - DONE: go through and fix subheadings
-- TODO: go through TODOs
+- DONE: go through TODOs
 - DONE: print
@@
 - reached conclusion in adding edits. Not sure if the section on Popper and possibiities trying to realize themselves is fleshed out enough.
-- TODO: check if the last section does enough with propensities and the new features.
+- DONE: check if the last section does enough with propensities and the new features.
 
@@
 - reached li 117/222, p18/34 today. 
-- TODO: check list section again - does it flow?
+- DONE: check list section again - does it flow?
 - did some more -- p 22/34, li 153/222
@@
     - Fri: submission
-- TODO: all change 'relevancy' to relevance; all change inverted commans to none for this paper
+- DONE: all change 'relevancy' to relevance; all change inverted commans to none for this paper
 - TODO: l.201 need to rethink the point of self-learning substitutes -- does it not fit with personalization?
@@
 - at l.55/222 in going through again this morning. 
+- going through TODOs;
+- quick read [@Amoore_2015] and found this quote to differ from:
+	>They carve out and convert radical heterogeneity into flat difference of degree, such that it appears as though everything is calculable, everything about the uncertain future is nonetheless decidable.361
+- Also this:
+	>In current debates on protecting people from the worst vicissitudes of data mining and algorithmic decision, the emphasis is placed overwhelmingly on the restriction of processes to specified people of interest and specific queries defined in advance.
+As we have argued, this critical framing entirely misunderstands the work of the analytics.
+The stories of the mass or bulk data are precisely the means by which the queries are generated – the analytics require the mass data in order to decide on what or who is interesting, and this can only ever be retroactive.
+361
+
+## Fri 29 Sep 2017 09:42:48 BST
+- just doing an hour on this today. Mainly on the TBAs - DONE
+- checking the list section; DONE
+- doing a bit more around experiments -- adding note on random variables.
+- went through intro again and re-ordered. How many times have I done that?
+- new title:  "Personalization and probabilities: conversion events in online grocery shopping"
+- ready to resubmit?

### 2017-09-29 reconstruction.rmd
redoing TBAs and intro

@@
 ---
-title: "Personalization and probabilities: impersonal propensities in online grocery shopping"
+title: "Personalization and probabilities: conversion events in online grocery shopping"
 author: Adrian Mackenzie^[Sociology, Lancaster University]
@@
 Using a mixture of discourse analysis and code-based reconstruction of key elements of the recommender system, it suggests that personalization is one facet of an open-ended weave of propensities associated with people and things in contemporary big data configurations.
-The paper explores how, in the context of recommender systems, the constitutive incompleteness of shopping lists, their propensity to expand or change, might be more important than their capacity to be personalized."
+The paper explores how, in the context of recommender systems, the constitutive incompleteness of social life, its propensity to expand or change, might be just as important as its capacity to be personalized."
 ---
@@
 
-Much big data practice converts an existing set of practices by adding a model or predictive elements based on the collection and analysis of data.
-This paper recounts the conversion of  an online grocery ordering system  from an older demographically-based recommendation system to a 'personal relevance' model.
+Big data systems promise to change existing practices by adding a model or predictive elements based on the collection and analysis of data.
+This paper recounts the conversion of  an online grocery ordering system  from an older demographically-based recommendation system to a 'personal relevance' predictive model.
+The main argument of the paper concerns how  personalization occurs in big data 'conversion events.' The term 'conversion event' has a dual sense.
+In recommender system design and web ecommerce, it refers to those occasions when a customer, user or site visitor clicks a specific hyperlink or purchases something.
+I also use it to refer to the common before-after narrative forms that proponents, developers and advocates of recommender systems adopt in describing, promoting, explaining or otherwise talking about big data, analytics or predictive modelling practices.
+
+On recent research and opinion on recommendations, personalization has been seen as the goal of predictive modelling.
+Big data discourse in its promissory mode attributes potency to personalization: 'most important, using big data we hope to identify specific individuals rather than groups' [@Mayer-Schonberger_2013].
+Conversely, current debates about the problems of big data emphasize the need to protect people from personalization.
+Joseph Turow and co-authors conclude in their critical account of the transformation of retail space by business analytics: 'through it all, knowingly and not, and away from the spotlights of fierce social debate, retailers are encouraging daily routines that accept data-driven personalization as a centrifugal public' [@Turow_2015, 476].
+Analysis of internet filter bubbles and the growth of predictive platforms in industry and government  [@Pariser_2011; @ONeil_2016a] also centre on personalization.
+
+In this paper, I agree that a transformation or conversion event is occurring, but one which also, even in the midst of much ongoing personalization,  opens the possibility of conceiving of social ordering processes afresh.
+Like Louise Amoore and Viola Piotukh call for a new epistemology of population  [@Amoore_2015, 360], I will argue that contemporary recommender systems afford us the opportunity to begin to re-conceptualise long-standing assumptions about social order and structure.
+The average everydayness and familiarity of the example here -- grocery shopping -- allows some first hand engagement with the messiness, the entanglements, and potentials of this conversion event in ways that other interesting settings -- social media for instance -- do not.
+
+'Data-driven personalization' relies on predictions, themselves dependent on data practices ranging across acquisition, storage, transformation, exchange, integration, modelling and experimentation.
+The reconstruction of a recommender system and its  personalized predictions I undertake here focuses  on the notion of probability and its role in modelling and experimentation.
+Even if probabilistic calculations of events have  long-standing importance in many settings (insurance, medicine, experimental and field sciences, operations research, risk analysis, engineering design, public health,  economic modelling, etc.), probabilities have taken on a new operational function in digital platforms associated with the web and social media.
+They have become part of infrastructures, transactions, and event-structures.
+In order to take this _operational_ reality into account, we might understand probabilities in quasi-physicalist terms  defined by the philosopher of science Karl Popper as a 'world of propensities, as an unfolding process of realizing possibilities and of unfolding new possibilities' [@Popper_1990,19].
+This operational-physicalist or sociotechnical notion of probability shifts our understanding of data analytics, especially in association with platforms [@Gillespie_2010], away from personalization.
+
 The account is empirical in several respects.
@@
 It conducts several small-scale code-based experiments in order to reconstruct, using widely available code resources resources such as API (Application Programmer Interfaces) and software libraries for machine learning, some prototypical elements of the system in question.
-Finally,  it  draws  on the philosophy of science to frame an engagement with grocery recommendations in terms of probabilities and their worldly enactments.
-
-The main argument of the paper concerns how  personalization occurs in big data 'conversion events.' The term 'conversion event' has a dual sense.
-In recommender system design and web ecommerce, it refers occasions when a customer, user or site visitor clicks or purchases.
-I also use it to   refer to the common before-after narrative forms that proponents, developers and advocates of recommender systems adopt in describing, promoting, explaining or otherwise talking about big data.
-In recent research and opinion on data practice, personalization has been seen as the linchpin of predictive modelling.
-Big data discourse in its promissory mode  attributes potency to personalization: 'most important, using big data we hope to identify specific individuals rather than groups' [@Mayer-Schonberger_2013].
-Conversely, Joseph Turow and co-authors conclude in their critical account of the transformation of retail space: 'through it all, knowingly and not, and away from the spotlights of fierce social debate, retailers are encouraging daily routines that accept data-driven personalization as a centrifugal public' [@Turow_2015, 476].
-Analysis of internet filter bubbles and the growth of predictive platforms in industry and government  [@Pariser_2011; @ONeil_2016a] also centre on personalization.
-In this paper, I agree that a transformation or conversion event is occurring, but one which also, even in the midst of much ongoing personalization,  opens the possibility of conceiving of social ordering processes afresh.
-The average everydayness and familiarity of the example here -- grocery shopping -- allows some first hand engagement with the messiness, the entanglements, and potentials of this conversion event in ways that other interesting settings -- social media for instance -- do not.
-
-'Data-driven personalization' relies on predictions, themselves dependent on data practices ranging across acquisition, storage, transformation, exchange, integration, modelling and experimentation.
-The reconstruction of a recommender system and its  personalized predictions I undertake here focuses on the notion of probability and its role in modelling and experimentation.
-Although probabilistic calculations of events have  long-standing importance in many settings (insurance, medicine, experimental and field sciences, operations research, risk analysis, engineering design, public health,  economic modelling, etc.), probabilities have taken on an operational function in digital platforms associated with the web and social media.
-They have become part of infrastructures, transactions, and event-structures.
-In order this _operational_ reality into account, we might understand probabilities in quasi-physicalist terms  defined by the philosopher of science Karl Popper as a 'world of propensities, as an unfolding process of realizing possibilities and of unfolding new possibilities' [@Popper_1990,19].
-This operational-physicalist or sociotechnical notion of probability shifts our understanding of data analytics, especially in association with platforms [@Gillespie_2010], away from personalization.
-
-Like some recent work in science and technology studies,  anthropology and media studies [@Marres_2017; @Marcus_2014; @Bogost_2012], the paper allows its own reconstruction to be provisionally shaped by an experiential encounter with the object it describes.
+Like some recent work in science and technology studies,  anthropology and media studies [@Marres_2017; @Marcus_2014; @Bogost_2012], the paper is  shaped by a practical encounter with the 'object' -- code, data, web platforms --  it describes.
 At times, this mode of empirical philosophy may seem fixated on the technical minutiae of algorithms and recommender systems.
@@
 Speaking to an audience of statisticians and operations researchers, Patel focused on the development and operation of predictive models underlying  shopping list recommendations.
-Her presentation was filled with graphs, numbers, and tables concerning ongoing development of the 'Have you forgotten?' recommender system.
+The presentation was filled with graphs, numbers, and tables concerning ongoing development of the 'Have you forgotten?' recommender system.
 
@@
 The recommendations are the product of a recommender system, an important category of operational device in big data (see for instance [@Hallinan_2014] for analysis of Netflix recommendation; [@Morris_2015]  or [@Seaver_2015] for an account of music recommendations).
-The question 'Have you forgotten?' is followed by a list of some  grocery items that could have been or are usually on a shopping list.
+The question 'Have you forgotten?' is followed by a list of some  grocery items that could have been or are usually on a shopping list, or that substitute for unavailable items.
 The title of the suggestions is a bit misleading.
-The recommender system, as we will see, is not concerned with forgetting, with the many slips and oversights associated to which shopping is prone,  but rather with substituting and adding items that customers had not selected perhaps because they had never thought of buying them in the first place.
+The recommender system, as we will see, is not concerned with forgetting or the many slips and oversights associated to which shopping is prone.
+The 'Have you forgotten' list substitutes and adds items that customers had not selected perhaps because they had never thought of buying them in the first place.
 
-Grocery shopping is difficult to personalize.
+Grocery shopping is difficult to personalize for several reasons.
+It has a complex relation to social order and structures.
 The anthropologist Daniel Miller has argued [@Miller_2012] that  shopping negotiates discrepancies between normative and actual social order (for instance, between the ideals of health emblematised by organic products and commitments to thrift embodied in lower cost generic products).
@@
 More provocatively, Miller claims that 'shopping is largely a technology for the expression of love' (85).
-Whatever theory we have of how normative social order arises, Miller's argument  implies that shopping inhabits a mundane  but highly variable space between normative/idea and actual order.
+Whatever theory we have of how normative social order arises or love technologies work, Miller's argument  implies that shopping inhabits a mundane  but highly variable space between normative/idea and actual order.
 Importantly, and this will be a key consideration for the Tesco recommender systems, grocery shopping is not necessarily personal or individual.
@@
 
-Shopping lists provisionally stabilise the complex social ordering of grocery shopping.
+Shopping lists are a form of writing that provisionally stabilise the complex social ordering of grocery shopping.
 They provide important clues as to how local social order is constituted, maintained, and repaired in shopping.
-As people shop, either by  trawling along aisles packed with thousands of products, or scrolling down screens or searching for particular brands amidst search results, lists filter or reduce the excessive abundance, claims on attention, dazzle and distraction of commodities to the practicalities of domestic economy.
-The shopping list, whether written on the back of an envelope, saved as a list in an online grocery shopping system, or remembered, lies at the intersection of logistical flows, infrastructural orderings, and  lively negotiations around actual and normative social orders.[^30.1]  Shopping lists  are intersectional ordering devices that encapsulate a universe of possible references, and a teeming multitude of propensities with an actual local order.[^1.456]
+As people shop, either by  trawling along aisles packed with thousands of products, or scrolling down screens or searching for particular brands amidst search results, lists filter or reduce the excessive abundance, claims on attention, dazzle and distraction of commodities on display (screens or shelves) to the exigencies and compromises of domestic economy.
+The shopping list, whether written on the back of an envelope, saved as a list in an online grocery shopping system, or simply remembered as 'I must remember to buy X,' lies at the intersection of logistical flows, infrastructural orderings, and  lively negotiations around actual and normative social orders.[^30.1]  Shopping lists are intersectional ordering devices that inscribe a universe of possible references or a multitude of propensities into a local order.[^1.456]
 
@@
 
-Hand-written shopping lists have ongoing practical importance and mixed ordering practices (see the montage of handwritten shopping lists at [Grocery List](http://www.grocerylist.org)).
-Online shopping lists, by contrast,  reside at the intersection of web and internet infrastructures, supply chain logistics, individualised practices and *habitus,* and increasingly, the predictive operations of  recommender systems.
-Whereas the aisles and shelves of a supermarket present a densely-woven mesh of objects competing for visual attention by offering distinctions of taste, thrift, expedience, novelty, indulgence, health, online shopping recommender systems generate lists that seek to align people to products that they otherwise might have little relation to (see [@Turow_2015] for an overview of the development of these systems).
+Hand-written shopping lists have ongoing practical importance as an ordering practice (see the montage of handwritten shopping lists at [Grocery List](http://www.grocerylist.org)).
+Online shopping lists, however,  shift the site of inscription onto a platform.
+They reside at the intersection of web and internet infrastructures, supply chain logistics, individualised practices and *habitus,* and increasingly, the predictive operations of  recommender systems.
+Whereas the aisles and shelves of a supermarket present a densely-woven mesh of objects competing for visual attention by offering distinctions of taste, thrift, expedience, novelty, indulgence, health, online shopping recommender systems generate lists that seek to interest people in products that they otherwise might have little relation to (see [@Turow_2015] for an overview of the development of these systems).
 
@@
 tables) precede discursive and narrative writing practices, then we might expect lists to function as powerful social ordering devices.
-More recent sociological work on lists (see [@deGoede_2016] for an overview) explore the social and political potency of lists as ordering devices.
-As literary scholars suggest (see [@Mainberger_2003]), even if lists have often been de-valued as literary forms, list-making commonly appears in literary form whenever writing seeks to address, name, group or evoke totality, profusion, excess or abundance.
+Recent sociological work on lists (see [@deGoede_2016] for an overview) explore the social and political potency of lists as ordering devices.
+As literary scholars suggest (see [@Mainberger_2003]),  list-making commonly appears in literary form whenever writing seeks to address, name, group or evoke totality, profusion, excess or abundance.
 From both  anthropological and literary perspectives, shopping lists have excellent reasons to exist: they are deeply rooted in organisational life and infrastructures.
-Lists weave together people, infrastructures, things, and places.While lists intersect strongly with other meaning and sense-making practices in everyday life and popular culture, they can be refractory to discourse  and textual analysis methods built around models of language or speech, with its rules of syntax and grammar.
+Lists weave together people, infrastructures, things, and places.
+While lists intersect with other meaning and sense-making practices in everyday life and popular culture, they can be refractory to discourse  and textual analysis methods built around models of language or speech, with its rules of syntax and grammar.
 On the one hand, lists are highly associative.
@@
 On the other hand, as operational inscriptions, they can be treated ethnomethodologically, as the production of social order in a given setting.
-We might, in the light of their mutability, permeability and embedding in social order,  approach lists from various theoretical angles -- as asignifying semiotics as Maurizio Lazzarato calls it in his _Signs and Machines_ [@Lazzarato_2014] or as elements in 'a new order of spatio-temporal continuity for forms of economic, political and cultural life' as Celia Lury puts it in her account of topological turn [@Lury_2012].
 
@@
 The main narrative of Patel's conversion narrative concerned the shift from a well-established loyalty-card based data-mining model developed in the 1990s to a predictive, probabilistic, 'personal relevance' model that would  append to the shopping list in almost-realtime.
-While the changes Patel described are configured in Tesco-specific ways by DunnHumby, they are also broadly typical.
-Tesco is the largest supermarket chain in UK.
-Its  customer loyalty and targeted marketing programme known as 'Tesco Clubcard' started in 1991.
-DunnHumby -- founded by operations researchers Edwina Dunn and Clive Humby -- is said to have convinced the CEO of Tesco sometime in 1991 that a loyalty card program could change the supermarket chain's relationship to its customers.
-Clive Humby's academic publications are hard to track down.
+While the changes Patel described are configured in Tesco-specific ways by DunnHumby, they are also broadly typical of a big data conversion event.
+Analytics service providers such as DunnHumby attempt to convert their customers by stories of conversion.[^65.1] Patel's presentation was part of this effort.
+Tesco is the largest supermarket chain in UK and a notable success story for DunnHumby.
+Tesco's  customer loyalty and targeted marketing programme known as 'Tesco Clubcard' started in 1991.
+DunnHumby -- founded by operations researchers Edwina Dunn and Clive Humby -- is said to have convinced the CEO of Tesco sometime in 1991 that a loyalty card program could change the supermarket chain's relationship to its customers.[^65.2]  Clive Humby's academic publications are hard to track down.
 An early paper given at the _Conference of Young Operational Researchers_ in Nottingham in 1984 suggests the direction that he, the company Dunn and Humby formed (DunnHumby) and later Tesco would take in constructing lists [@OKeefe_1984].
 The abstract for Humby's presentation pre-figures an ongoing trajectory for data mining techniques aimed at eliciting detailed information on individual customer references.
+
+[^65.1] The web pages of the big data analytic companies providing products to various sectors of industry, government and media are stocked with images, graphics and statements centred on this conversion experience.
+[@Amoore_2015] analyses some of these.
+
+[^65.2]: Tesco changed its relations to customers in 1984 when they launched a very online grocery shopping system [@Winterman_2013].
 
@@
 If 'Tesco is the clear winner in the online grocery market, in fact it takes almost 50p of every £1 spent on food shopping on the internet' [@Silverwood-Cope_2014], then has Tesco itself undergone some kind of conversion?
-Patel described her work at DunnHumby as converting the recommender system from a 'rules-based list' to a 'relevancy model.' The relevancy model affected the construction of the 'Have You Forgotten' list.
+Patel described her work at DunnHumby as converting the recommender system from a 'rules-based list' to a 'relevance model.' The relevance model affected the construction of the 'Have You Forgotten' list.
 This very localised intervention is typical of broader re-organisation of prediction.
@@
 When the logistic regression model includes the 52 weeks of previous purchases, the conditional probability calculation undertaken by the recommender system ramifies tremendously in several respects.
-Potentially, each of Tesco's 200,000 products becomes a variable in the classifier.
-We can imagine this as an arithmetic sum extending along a series of 200,000 terms.
-Practically, most of the these variables will only slightly influence the sum of the probabilities for recommendations since most customers will have purchased only a fraction of the inventory.
+In principle, each of Tesco's 200,000 products becomes a variable in the classifier.
+We can imagine the predictive model as repeatedly adding together weighted values of these  200,000 terms, and adjusting the weighting of each term to produce a sum, the probability of a particular item being purchased.
+In the final analysis, most of the variables will only slightly influence the sum of the probabilities for recommendations since most customers will have purchased only a fraction of the inventory.
 As Patel observed,  again assuming that this would be obvious to the audience of data scientists,  'we have lots of zeros'.
 A matrix that records associations between individual people and products is bound to be mostly empty.
-Say Tesco has 1 million online customers.
+(Say Tesco has 1 million online customers.
 Each online  shopper has bought some selection of the 200,000 products.
@@
 The product-customer matrix, the basic vector-space in which all recommender systems operate, remains very sparse and unpopulated.
-Given that any one customer is likely to only have bought 100 or so different products, the  matrix contain  `r format(100-(100/2e5)*100, 2)`% zero probabilities.
+Given that any one customer is likely to only have bought 100 or so different products, the  matrix contain  `r format(100-(100/2e5)*100, 2)`% zero probabilities.)
 
@@
 
+'Personal relevance' depends on a matrix of probabilities of associations between people and things that shifts in time.
 The possibility of adjusting the recommendations for each customer every week depends on an infrastructure capable of collecting data, and assimilating that data in a predictive model.
-'Personal relevance' depends on a matrix of probabilities of associations between people and things that shifts in time.
-`Hadoop` and its legion of 'big data' variants (`mahout`, `spark`, `hive`, `pig`, `yarn`) operationalise repetition at an infrastructural rather than analytic scale.
-Patel's quick gloss of the infrastructural deployment of DunnHumby's relevancy model is  primary to the conversion event: the logistic regression model at the heart of the recommender system is no longer an analytical device but an operational one because it seeks to revise recommendations as situations change.
-[TBA: this paragraph is promising but doesn't deliver yet]
+Even `Hadoop` and the legion of 'big data' variants (`mahout`, `spark`, `hive`, `pig`, `yarn`) that the new Tesco model relies on operationalise repetition at an infrastructural rather than situational scale.
+Patel's quick gloss of the weekly updates of DunnHumby's relevance model  is  primary to the conversion event: the logistic regression model at the heart of the recommender system is no longer an analytical device standing outside the platform but an operational one.
+It seeks to revise recommendations as situations change but is still governed by the infrastructural rhythms of weekly batch jobs.
 
@@
 
-Like the accounts of laboratory practice provided by science studies [@Latour_1097; @Rheinberger, @Knorr-Cetina; @Lynch_1994], Popper understands experiments as stabilizing the conditions so that probabilities become less variable.
-Experiments play an important role in conversion events, or in the process of bring new predictive models to bear on shopping practice.
+Like the accounts of laboratory practice provided by science studies [@Latour_1986;  @Lynch_1994], Popper understands experiments as stabilizing the conditions so that propensities become less variable.[^178.1] Experiments play an important role in conversion events, or in the process of bring new predictive models to bear on shopping practice.
 A crucial consideration for  a recommender system designer will be how to experimentalize the systems recommendations so that they are not disturbed by other propensities.
 This is an inordinately difficult challenge, since, as we have seen, grocery shopping takes place at the intersection of normative and actual social orders, individual and group belongings, and logistically global economic processes.
+
+[^178.1]: At the same time, variables in contemporary science and engineering have become more variable as random processes take on greater importance.
+The notion of a 'random variable,' for instance, incorporates a probability distribution into the variable itself.
+See [@MackenzieAdrian_2016a]
 
@@
 Experimentalization, the practice of creating conditions that reduce disturbing propensities, runs deep in the conversion event.
-The predictions of the recommender system themselves are the subject of experiment.
-Patel described the deployment of  the personal relevance model in a random A/B controlled trial on the Tesco website.
+The predictions of the recommender systems themselves are the object of the experiment.
+Patel described tests of  the personal relevance model in a random A/B controlled trial on the Tesco website.
 All customers were  allocated to one of four categories as shown in  table.
@@
 Random allocation of customers to control and test groups occurs without taking individual propensities into account.
-A/B testing  seeks to statistically validate effects -- the *uplift* -- of the model on  conversion rates by directly measuring the effects of the model on what people do.
+A/B testing  seeks to statistically validate predictions -- the *uplift* -- of different models on conversion rates by directly measuring differing  effects on what people do.
 The uplift refers to conversion events associated with the same groups of people.
-Effectively an experiment in creating micrologically different worlds, the randomised control trial  sets up a control mechanism that connects  the predictive model (the logistic regression), and the conversion event more broadly.
-With this experimentalisation, the conversion event narrative lacks grounding.[@TBA: not happy with this ending]
+Effectively an experiment in creating micrologically different worlds, the randomised control trial  calibrates  the predictive model (the logistic regression) by testing it under somewhat controlled conditions.
+Without this experimentalisation of the online shopping platform, the conversion event narrative lacks grounding.
 
@@
 The DunnyHumby data scientists constructed a new feature from the previous purchase data measuring substitutability and complementarity between products.
-Taking all the baskets of items purchased on the basis of recommendations, they derived a  new feature, what DunnyHumby termed 'self-learning substitutes.' Added to the relevancy model, the 'self learning substitutes' feature adds another data matrix, the product similarity matrix, itself generated from all previous recommendations that have led to conversion events (that is, 'a user clicked as the response').
+Taking all the baskets of items purchased on the basis of recommendations, they derived a  new feature, what DunnyHumby termed 'self-learning substitutes.' Added to the relevance model, the 'self learning substitutes' feature adds another data matrix, the product similarity matrix, itself generated from all previous recommendations that have led to conversion events (that is, 'a user clicked as the response').
 The self-learning substitutes feature is complex.
@@
 
-I suggest that we might understand what precedes and exceeds personalization as probabiliization.
-If, as Popper argues, probabilities are being realized in the same way that 'fields' were realized in modern physics, then we should see the compute  platforms, databases, software libraries, web interfaces and global supply chain logistics as part of this conversion event.
-When we observe what is practically done to construct a personal relevance model and apply it to a shopping list, we see conditional probabilities assembled in predictive models, the rhythms of  platform and infrastructural configurations in gathering and processing data,  and the open-ended construction of features in data.
+I suggest that we might understand what precedes and exceeds personalization as probabilization.
+If, as Popper argues, a propensity-based account of probabilities 'amounts to generalizing and extending the idea of forces again' [@Popper_1992,14], then we should see the compute  platforms, databases, software libraries, web interfaces and global supply chain logistics as part of this conversion event, as a tendency in the process of realization.
+When we track what is practically done to construct a personal relevance model and apply it to a shopping list, we see conditional probabilities assembled in predictive models, the rhythms of  platform and infrastructural configurations in gathering and processing data,  and the open-ended construction of features in data.
 In various ways, they shift and complicate the  associations between people and things, people and people, and things with things.
-If social order is made of associations, if to be social is to associate, then big data conversion events operationalize association in matrices of propensity.
+The framing of this conversion as a 'flat difference of degree, such that it appears as though everything is calculable' [@Amoore_2015, 361] risks misunderstanding the different propensities tending to realization.
+If social order is made of propensities to associate, if to be social is a propensity to associate, then big data conversion events operationalize association in matrices of propensity.
 

### 2017-10-02 log.rmd
fixing up lost manuscript bits and references

@@
 - ready to resubmit?
+
+## Mon 02 Oct 2017 12:14:12 BST
+- facing resubmission build. Bound to be missing references. I have them all somewhere.  
+- using render_tufte.r
+- drats - problem with the work I did on Friday -- some kind of crash (battery ran out?) has lost work I did later on Friday morning. Need to redo that now. The TBAs are still there for instance. 
+- have got most of the references in; need to get Christian_2012 in there. 
+- TODO: go abstract and intro again;
+    - go through conclusion again;
+    - redo TBAs
+    - check whether the extra Amoore point about homogeneous calculative space is there

### 2017-10-02 reconstruction.rmd
fixing up lost manuscript bits and references

--- 2017-09-28 reconstruction.rmd
@@
 
-Much big data practice converts an existing set of practices by adding a model or predictive elements based on the collection and analysis of data.
+Big data practice often converts an existing state of affairs by adding a model or predictive elements based on the collection and analysis of data.
 This paper recounts the conversion of  an online grocery ordering system  from an older demographically-based recommendation system to a 'personal relevance' model.
+The main argument of the paper concerns how  personalization occurs in big data 'conversion events.' The term 'conversion event' has a dual sense.
+In recommender system design and web ecommerce, it refers to those occasions when a customer, user or site visitor clicks a specific hyperlink or purchases something.
+I also use it to refer to the common before-after narrative forms that proponents, developers and advocates of recommender systems adopt in describing, promoting, explaining or otherwise talking about big data, analytics or predictive modelling practices.
+In recent research and opinion on recommendations, personalization has been seen as the goal of predictive modelling.
+Big data discourse in its promissory mode attributes potency to personalization: 'most important, using big data we hope to identify specific individuals rather than groups' [@Mayer-Schonberger_2013].
+Conversely, current debates about the problems of big data emphasize the need to protect people from personalization.
+Joseph Turow and co-authors conclude in their critical account of the transformation of retail space by business analytics: 'through it all, knowingly and not, and away from the spotlights of fierce social debate, retailers are encouraging daily routines that accept data-driven personalization as a centrifugal public' [@Turow_2015, 476].
+Analysis of internet filter bubbles and the growth of predictive platforms in industry and government  [@Pariser_2011; @ONeil_2016a] also centre on personalization.
+In this paper, I agree that a transformation or conversion event is occurring, but one which also, even in the midst of much ongoing personalization,  opens the possibility of conceiving of social ordering processes afresh.
+Like Louise Amoore and Viola Piotukh's call for a new epistemology of population  [@Amoore_2015, 360], I will argue that contemporary recommender systems afford us the opportunity to re-conceptualise long-standing assumptions about social order and structure.
+The average everydayness and familiarity of the example here -- grocery shopping -- allows some first hand engagement with the messiness, the entanglements, and potentials of this conversion event in ways that other interesting settings -- social media for instance -- do not.
+
+'Data-driven personalization' relies on predictions, themselves dependent on data practices ranging across acquisition, storage, transformation, exchange, integration, modelling and experimentation.
+The reconstruction of a recommender system and its  personalized predictions that I undertake here centres on the notion of probability and its role in modelling and experimentation.
+Although probabilistic calculations of events have  long-standing importance in many settings (insurance, medicine, experimental and field sciences, operations research, risk analysis, engineering design, public health,  economic modelling, etc.), probabilities have taken on an operational function in digital platforms associated with the web and social media.
+They have become part of infrastructures, transactions, and event-structures.
+In order to take this _operational_ reality into account, we might understand probabilities in quasi-physicalist terms  defined by the philosopher of science Karl Popper as a 'world of propensities, as an unfolding process of realizing possibilities and of unfolding new possibilities' [@Popper_1990,19].
+This operational-physicalist or sociotechnical notion of probability shifts our understanding of data analytics, especially in association with platforms [@Gillespie_2010], away from personalization.
+Finally,  it  draws  on the philosophy of science to frame an engagement with grocery recommendations in terms of probabilities and their worldly enactments.
+
 The account is empirical in several respects.
@@
 It conducts several small-scale code-based experiments in order to reconstruct, using widely available code resources resources such as API (Application Programmer Interfaces) and software libraries for machine learning, some prototypical elements of the system in question.
-Finally,  it  draws  on the philosophy of science to frame an engagement with grocery recommendations in terms of probabilities and their worldly enactments.
-
-The main argument of the paper concerns how  personalization occurs in big data 'conversion events.' The term 'conversion event' has a dual sense.
-In recommender system design and web ecommerce, it refers occasions when a customer, user or site visitor clicks or purchases.
-I also use it to   refer to the common before-after narrative forms that proponents, developers and advocates of recommender systems adopt in describing, promoting, explaining or otherwise talking about big data.
-In recent research and opinion on data practice, personalization has been seen as the linchpin of predictive modelling.
-Big data discourse in its promissory mode  attributes potency to personalization: 'most important, using big data we hope to identify specific individuals rather than groups' [@Mayer-Schonberger_2013].
-Conversely, Joseph Turow and co-authors conclude in their critical account of the transformation of retail space: 'through it all, knowingly and not, and away from the spotlights of fierce social debate, retailers are encouraging daily routines that accept data-driven personalization as a centrifugal public' [@Turow_2015, 476].
-Analysis of internet filter bubbles and the growth of predictive platforms in industry and government  [@Pariser_2011; @ONeil_2016a] also centre on personalization.
-In this paper, I agree that a transformation or conversion event is occurring, but one which also, even in the midst of much ongoing personalization,  opens the possibility of conceiving of social ordering processes afresh.
-The average everydayness and familiarity of the example here -- grocery shopping -- allows some first hand engagement with the messiness, the entanglements, and potentials of this conversion event in ways that other interesting settings -- social media for instance -- do not.
-
-'Data-driven personalization' relies on predictions, themselves dependent on data practices ranging across acquisition, storage, transformation, exchange, integration, modelling and experimentation.
-The reconstruction of a recommender system and its  personalized predictions I undertake here focuses on the notion of probability and its role in modelling and experimentation.
-Although probabilistic calculations of events have  long-standing importance in many settings (insurance, medicine, experimental and field sciences, operations research, risk analysis, engineering design, public health,  economic modelling, etc.), probabilities have taken on an operational function in digital platforms associated with the web and social media.
-They have become part of infrastructures, transactions, and event-structures.
-In order this _operational_ reality into account, we might understand probabilities in quasi-physicalist terms  defined by the philosopher of science Karl Popper as a 'world of propensities, as an unfolding process of realizing possibilities and of unfolding new possibilities' [@Popper_1990,19].
-This operational-physicalist or sociotechnical notion of probability shifts our understanding of data analytics, especially in association with platforms [@Gillespie_2010], away from personalization.
-
-Like some recent work in science and technology studies,  anthropology and media studies [@Marres_2017; @Marcus_2014; @Bogost_2012], the paper allows its own reconstruction to be provisionally shaped by an experiential encounter with the object it describes.
-At times, this mode of empirical philosophy may seem fixated on the technical minutiae of algorithms and recommender systems.
+Like some recent work in science and technology studies,  anthropology and media studies [@Marres_2017; @Marcus_2014; @Bogost_2012], the paper is  shaped by practical encounter with the 'object' -- code, data, web platforms --  it describes.
+At times, this mode of empirical philosophy may seem overly fixated on the technical minutiae of algorithms and recommender systems and inattentive to lived experience or practices.
 With some forbearance on the part of readers, however, the argument of the paper should resonate sociologically.
-It concerns the relationality of social life, or the consistency and regularity of relating and acting amid a world in where probabilities and calculations based on probabilities play a pervasive role.
+It concerns the relationality of social life, or the consistency and regularity of relating and acting amid a world in where probabilities and calculations based on probabilities are widespread.
 
@@
 Second, Patel focused on the renovation and updating of long-standing data-mining practice via a much more explicitly 'big data' and 'machine learning'- oriented implementation.
-Her presentation concerns what  a typical big data conversion event  in which a long-standing recommender system was replaced by a 'big data'-style system delivering 'personally relevant' recommendations.
+Her presentation concerns a typical big data conversion event  in which a long-standing recommender system was replaced by a 'big data'-style system delivering 'personally relevant' recommendations.
 What stands out from the presentation is the normality of the recommender system: it is part of a long-standing and ongoing transformation of grocery shopping.
@@
 
-The main narrative of Patel's conversion narrative concerned the shift from a well-established loyalty-card based data-mining model developed in the 1990s to a predictive, probabilistic, 'personal relevance' model that would  append to the shopping list in almost-realtime.
-While the changes Patel described are configured in Tesco-specific ways by DunnHumby, they are also broadly typical.
-Tesco is the largest supermarket chain in UK.
-Its  customer loyalty and targeted marketing programme known as 'Tesco Clubcard' started in 1991.
-DunnHumby -- founded by operations researchers Edwina Dunn and Clive Humby -- is said to have convinced the CEO of Tesco sometime in 1991 that a loyalty card program could change the supermarket chain's relationship to its customers.
-Clive Humby's academic publications are hard to track down.
+The main narrative of Patel's conversion narrative concerned the shift from a well-established loyalty-card based data-mining model developed in the 1990s to a predictive, probabilistic, 'personal relevance' model that would  append items to the shopping list in almost-realtime.
+While the changes Patel described are configured in Tesco-specific ways by DunnHumby, they are also broadly typical of a big data conversion event.
+Analytics service providers such as DunnHumby attempt to convert their customers by stories of conversion.[^65.1] Patel's presentation was part of this effort.
+Tesco is the largest supermarket chain in UK and a notable success story for DunnHumby.
+Tesco's  customer loyalty and targeted marketing programme known as 'Tesco Clubcard' started in 1991.
+DunnHumby -- founded by operations researchers Edwina Dunn and Clive Humby -- is said to have convinced the CEO of Tesco sometime in 1991 that a loyalty card program could change the supermarket chain's relationship to its customers.[^65.2]  Clive Humby's academic publications are hard to track down.
 An early paper given at the _Conference of Young Operational Researchers_ in Nottingham in 1984 suggests the direction that he, the company Dunn and Humby formed (DunnHumby) and later Tesco would take in constructing lists [@OKeefe_1984].
 The abstract for Humby's presentation pre-figures an ongoing trajectory for data mining techniques aimed at eliciting detailed information on individual customer references.
+
+[^65.1] The web pages of the big data analytic companies providing products to various sectors of industry, government and media are stocked with images, graphics and statements centred on this conversion experience.
+[@Amoore_2015] analyses some of these.
+
+[^65.2]: Tesco changed its relations to customers in 1984 when they launched a very online grocery shopping system [@Winterman_2013].
 
@@
 
-In  [@Humby_1989], Humby argues for the need to add lifestage data to neighbourhood data in marketing research.
+In  [@Humby_1989], Humby highlights the need to add lifestage data to neighbourhood data in marketing research.
 While Tesco  succeeded in data-mining its customers using demographic segmentation,  and perhaps became the UK's biggest supermarket with the help of data-mining in the 1990s, the shopping environment  in 2017 is markedly different [@Turow_2017].
@@
 A now-classic approach to 'market basket analysis,' it was listed as one of the top ten data mining algorithms in a survey conducted amongst data miners [@Wu_2008] and usually attracts a chapter in data-mining and machine learning textbooks (e.g.
-[@Hastie_2008]).
+[@Hastie_2009]).
 The interest of `apriori` for our purposes is that it begins to address the problem of understanding large numbers of shopping transactions as a matter of conditional probability.
@@
 In this sense, is still oriented by a notion of absolute probability, inflected by some elements of conditional probability.
-Commonly occurring sets are expressed as 'association rules.' For instance, applied to a dataset of generic, unbranded grocery purchases, the `apriori` algorithm counts frequencies of purchase in the overall set of all items purchased in a supermarket (the `Groceries` dataset  was acquired from a 'local German supermarket' [@Hahsler_2006a]).
+Commonly occurring sets are expressed as 'association rules.' For instance, applied to a dataset of generic, unbranded grocery purchases, the `apriori` algorithm counts frequencies of purchase in the overall set of all items purchased in a supermarket (the `Groceries` dataset  was acquired from a 'local German supermarket' [@Hahsler_2006]).
 Figure \@ref(fig:arules-ex) shows how often the most common items appear.
@@
 `Hadoop` and its legion of 'big data' variants (`mahout`, `spark`, `hive`, `pig`, `yarn`) operationalise repetition at an infrastructural rather than analytic scale.
-Patel's quick gloss of the infrastructural deployment of DunnHumby's relevancy model is  primary to the conversion event: the logistic regression model at the heart of the recommender system is no longer an analytical device but an operational one because it seeks to revise recommendations as situations change.
-[TBA: this paragraph is promising but doesn't deliver yet]
+Patel's quick gloss of the infrastructural deployment of DunnHumby's relevance model is  primary to the conversion event: the logistic regression model at the heart of the recommender system is no longer an analytical device but an operational one because it seeks to revise recommendations as situations change.
 
@@
 
-Like the accounts of laboratory practice provided by science studies [@Latour_1097; @Rheinberger, @Knorr-Cetina; @Lynch_1994], Popper understands experiments as stabilizing the conditions so that probabilities become less variable.
+Like the accounts of laboratory practice provided by science studies [@Latour_1986;  @Lynch_1993], Popper understands experiments as stabilizing the conditions so that probabilities become less variable.
 Experiments play an important role in conversion events, or in the process of bring new predictive models to bear on shopping practice.
@@
 |Control A|Control B|
----------------------
+|---------------------
 
@@
 Effectively an experiment in creating micrologically different worlds, the randomised control trial  sets up a control mechanism that connects  the predictive model (the logistic regression), and the conversion event more broadly.
-With this experimentalisation, the conversion event narrative lacks grounding.[@TBA: not happy with this ending]
+With this experimentalisation, the conversion event narrative lacks grounding.[TBA: not happy with this ending]
 
@@
 
-I suggest that we might understand what precedes and exceeds personalization as probabiliization.
-If, as Popper argues, probabilities are being realized in the same way that 'fields' were realized in modern physics, then we should see the compute  platforms, databases, software libraries, web interfaces and global supply chain logistics as part of this conversion event.
-When we observe what is practically done to construct a personal relevance model and apply it to a shopping list, we see conditional probabilities assembled in predictive models, the rhythms of  platform and infrastructural configurations in gathering and processing data,  and the open-ended construction of features in data.
+I suggest that we might understand what precedes and exceeds personalization as probabilization.
+If, as Popper argues, a propensity-based account of probabilities 'amounts to generalizing and extending the idea of forces again' [@Popper_1990,14], then we should see the compute  platforms, databases, software libraries, web interfaces and global supply chain logistics as part of this conversion event, as a tendency in the process of realization.
+When we track what is practically done to construct a personal relevance model and apply it to a shopping list, we see conditional probabilities assembled in predictive models, the rhythms of  platform and infrastructural configurations in gathering and processing data,  and the open-ended construction of features in data.
 In various ways, they shift and complicate the  associations between people and things, people and people, and things with things.
-If social order is made of associations, if to be social is to associate, then big data conversion events operationalize association in matrices of propensity.
+The framing of this conversion as a 'flat difference of degree, such that it appears as though everything is calculable' [@Amoore_2015, 361] risks misunderstanding the different propensities tending to realization.
+If social order is made of propensities to associate, if to be social is a propensity to associate, then big data conversion events operationalize association in matrices of propensity.
 

### 2017-10-02 render_tufte.r

@@ -4,6 +4,6 @@ library(tint)
 library(tufte)
 #render('lists_conversion_experience.rmd',output_format=tufte_handout())
 #render('lists_conversion_experience.rmd',output_format=ioslides_presentation)
-render('reconstruction.rmd',output_format=tintPdf())
+#render('reconstruction.rmd',output_format=tintPdf())
 render('reconstruction.rmd',output_format=word_document2())
 

### 2017-10-04 log.rmd
doing todos

@@
 - have got most of the references in; need to get Christian_2012 in there. 
-- TODO: go abstract and intro again;
+- TODO:
+     -DONE go abstract and intro again;
     - go through conclusion again;
-    - redo TBAs
-    - check whether the extra Amoore point about homogeneous calculative space is there
+    - DONE I think: redo TBAs
+    - DONE: check whether the extra Amoore point about homogeneous calculative space is there
+    - go over the revisions letter
+    - generate the docx

### 2017-10-04 reconstruction.rmd
doing todos

@@
 Personalization, with all the risks of discrimination and bias it entails, has been the critical focus in accounts of consumption, government, social media and health.
-This paper argues that personalization is part of a more expansive change in probabilization that, in principle, is not reducible to individual attributes and actions.
+This paper argues that personalization through models using large-scale data is part of a more expansive change in probabilization that, in principle, is not reducible to individual or 'personal' attributes and actions.
 It describes the 'personalization' of an online grocery shopping recommender system to list a small number of grocery items of personal relevance for each of the millions of online grocery shoppers at a major UK supermarket chain.
-Drawing on a theory of probability proposed by the philosopher of science Karl Popper and anthropological work on shopping, it suggests that the attempt to generate  personalized predictions changes relations to others and things.
+Drawing on a theory of probability proposed by the philosopher of science Karl Popper and anthropological work on shopping, it suggests that the attempt to generate  personalized predictions necessarily incorporates impersonal relations to others and things.
 Using a mixture of discourse analysis and code-based reconstruction of key elements of the recommender system, it suggests that personalization is one facet of an open-ended weave of propensities associated with people and things in contemporary big data configurations.
@@
 'Data-driven personalization' relies on predictions, themselves dependent on data practices ranging across acquisition, storage, transformation, exchange, integration, modelling and experimentation.
-The reconstruction of a recommender system and its  personalized predictions that I undertake here centres on the notion of probability and its role in modelling and experimentation.
-Although probabilistic calculations of events have  long-standing importance in many settings (insurance, medicine, experimental and field sciences, operations research, risk analysis, engineering design, public health,  economic modelling, etc.), probabilities have taken on an operational function in digital platforms associated with the web and social media.
-They have become part of infrastructures, transactions, and event-structures.
-In order to take this _operational_ reality into account, we might understand probabilities in quasi-physicalist terms  defined by the philosopher of science Karl Popper as a 'world of propensities, as an unfolding process of realizing possibilities and of unfolding new possibilities' [@Popper_1990,19].
+The reconstruction of an online grocery shopping recommender system and its  personalized predictions that I undertake here centres on the notion of probability and its role in modelling and experimentation.
+Although probabilistic calculations of events have  long-standing importance in many settings (insurance, medicine, experimental and field sciences, operations research, risk analysis, engineering design, public health,  economic modelling, etc.), probabilities have taken on an altered operational function in digital platforms.
+They have become mundane part of the weave of infrastructures, transactions, and practices found in everyday life.
+In order to take the _operational_ reality of probabilities into account, we might understand probabilities in quasi-physicalist terms  defined by the philosopher of science Karl Popper as a 'world of propensities, as an unfolding process of realizing possibilities and of unfolding new possibilities' [@Popper_1990,19].
 This operational-physicalist or sociotechnical notion of probability shifts our understanding of data analytics, especially in association with platforms [@Gillespie_2010], away from personalization.
-Finally,  it  draws  on the philosophy of science to frame an engagement with grocery recommendations in terms of probabilities and their worldly enactments.
+It  draws  on Popper's re-framing of probability as propensity to  engage with grocery recommendations in their worldly enactments.
 
@@
 
-At one of the many industry-meets-academia events occurring in data science-oriented higher education institutions in the UK, speakers from industry, government and commerce described their work with predictive models.[^11.1] Shreena Patel, a PhD graduate in statistics and operations research,  works as a data scientist for DunnHumby, a well-known [customer science company ](https://www.dunnhumby.com/) [@dunnhumby_2017].
+At one of the many industry-meets-academia events occurring in data science-oriented higher education institutions in the UK, speakers from industry, government and commerce described their work with predictive models.[^11.1] Their narrative often followed the 'before-Big Data and after-Big data' conversion form.
+Shreena Patel, a PhD graduate in statistics and operations research,  works as a data scientist for DunnHumby, a well-known [customer science company ](https://www.dunnhumby.com/) [@dunnhumby_2017].
 Her work at DunnHumby focuses on online grocery shopping at the supermarket chain Tesco.
@@
 Patel mentioned many of these steps only fleetingly in the presentation, for they are largely taken for granted as part of predictive analytic practice.
-Second, Patel focused on the renovation and updating of long-standing data-mining practice via a much more explicitly 'big data' and 'machine learning'- oriented implementation.
+Second, Patel focused on the renovation and updating of long-standing data-mining practice via a much more explicitly 'big data' and 'machine learning'-oriented implementation.
 Her presentation concerns a typical big data conversion event  in which a long-standing recommender system was replaced by a 'big data'-style system delivering 'personally relevant' recommendations.
-What stands out from the presentation is the normality of the recommender system: it is part of a long-standing and ongoing transformation of grocery shopping.
+What stands out from the presentation is not any state-of-the-art innovation, but the ongoing life of the recommender system: the new predictive model is part of a long-standing and ongoing transformation of grocery shopping.
 From a sociological standpoint, the interest lies less in specific technical innovations and more on how the ordering work done by recommender systems relates to the role of prediction in social order more generally.
@@
 Effectively an experiment in creating micrologically different worlds, the randomised control trial  sets up a control mechanism that connects  the predictive model (the logistic regression), and the conversion event more broadly.
-With this experimentalisation, the conversion event narrative lacks grounding.[TBA: not happy with this ending]
+Without this experimental connection, the conversion event narrative lacks grounding in a state of affairs in the world.
 
@@
 The Tesco recommender system, starting from its early experiments with demographic data-mining, its later adoption of a rule-based 'market basket' analysis system, and its recent implementation of a machine learning 'personal relevance model' follows the trajectory of personalization.
-Its conversion event largely takes the form of personalization.
-Only because grocery shopping is  personally pre-ordered by lists can a predictive recommendation, provisionally conducive to new conversion events, gain traction.
+Its recent big data conversion event largely takes the form of personalization.
 Individualizing personalization, however,  is ill-fit to the socially complex negotiations of grocery shopping, and tends to obscure other implications.
@@
 They include an ongoing process of adjusting, scaffolding, intervening and configuring orders.
+Only because grocery shopping is  personally pre-ordered by lists can a predictive recommendation, provisionally conducive to new conversion events, gain traction.
 
 I suggest that we might understand what precedes and exceeds personalization as probabilization.
-If, as Popper argues, a propensity-based account of probabilities 'amounts to generalizing and extending the idea of forces again' [@Popper_1990,14], then we should see the compute  platforms, databases, software libraries, web interfaces and global supply chain logistics as part of this conversion event, as a tendency in the process of realization.
-When we track what is practically done to construct a personal relevance model and apply it to a shopping list, we see conditional probabilities assembled in predictive models, the rhythms of  platform and infrastructural configurations in gathering and processing data,  and the open-ended construction of features in data.
-In various ways, they shift and complicate the  associations between people and things, people and people, and things with things.
 The framing of this conversion as a 'flat difference of degree, such that it appears as though everything is calculable' [@Amoore_2015, 361] risks misunderstanding the different propensities tending to realization.
+If, as Popper argues, a propensity-based account of probabilities 'amounts to generalizing and extending the idea of forces again' [@Popper_1990,14], then we should see the compute  platforms, databases, software libraries, web interfaces and global supply chain logistics as part of this conversion event, as  tendencies or propensities in  the process of realization.
+When we track what is practically done to construct a personal relevance model and apply it to a shopping list, we see conditional probabilities assembled in predictive models, but also in the rhythms of  platform and infrastructural configurations and experiments that gather and process data,  and in the open-ended construction of features in data.
+In various ways, these tendencies shift and complicate the  associations between people and things, people and people, and things with things.
+They cut across boundaries between the personal and impersonal.
 If social order is made of propensities to associate, if to be social is a propensity to associate, then big data conversion events operationalize association in matrices of propensity.

### 2017-10-05 reply_to_reviews.md
submitting rewritten version

@@
+Dear Evelyn
+This is the re-written version of my earlier submission. I've pasting my response to the reviews below.
+
+best wishes
+
+Adrian Mackenzie
+
+-----------------------------
+
 Thanks very much to reviewers for their work on what must have seemed like a frustrating and elusive manuscript.
@@
 
-Much of the confusion about the main argument stemmed from its multiple threads.
+Much of the call for clarification concerned the main argument.
+It has too many  threads.
 It was engaging in several different debates at once -- about recommender systems and their exploitative effects on  consumption practices, about recommender systems as examples of how big data practices take root in specific sectors, and recommender systems as symptom of some kind of more ontological transformation in sociotechnical collectives.
-In the re-write, I've disentangled these threads in order to clarify the argument: grocery recommender systems are a mundane example of changes associated with big data; personalization is a powerful ordering technique in big data practice; but personalization cannot account for either the range of practices associated with big data or the broader associated transformations.
-The 'conversion' of the Tesco grocery recommender system to a personal relevancy model helps make this argument: grocery shopping is a complex social negotiation, intimately entwined with social groups (family, household, etc).
+In the re-write, I've mainly sought to clarify the argument: grocery recommender systems are a mundane example of changes associated with big data; personalization is a powerful ordering technique in big data practice; but personalization cannot account for either the range of practices associated with big data or the broader associated transformations.
+The case of  'conversion' of the Tesco grocery recommender system to a personal relevancy model helps make this argument: grocery shopping is a complex social negotiation, intimately entwined with social groups (family, household, etc).
 Data-mining to target grocery shoppers pre-dates by several decades recent growth in predictive analytics.
@@
 It would be hard to list all the changes.
+Many of them respond to specific points in the reviews relating to relevant literature or problems in argumentation.
 The basic idea of the rewrite was remove any thread of discussion that didn't relate to the main argument.
-A lot of the discussion of list practice was distracting so that has been reduced to a minimum.
+A lot of the discussion of list practice and orderings was distracting so that has been reduced to a minimum.
 I wasn't able to provide the more substantial empirical engagement requested by Reviewer 2.
-That would require a new research project.
+That would entail a new research project.
 The case study cannot make any claims to ethnographic thickness, so what the paper does to deepen its argumentative engagement around personalization in order to develop a more genealogical account based  on  research literature, widely used software and some examples of data from the Tesco API.
+
+Main points from reviews:
 

### 2017-10-05 log.rmd
submitting rewritten version

@@
     - generate the docx
+
+## Thu 05 Oct 2017 11:59:08 BST
+- submitting it as a new submission. The old one was BDS-17-0018. Does that mean that by march there were 18 submissions?   
+- reconstruction.docx is the main file. It has been handtweaked in various ways, but most of these have been put back into reconstruction.rmd.
+- figure numbering is still not right. Done that by hand.

### 2017-10-05 reconstruction.rmd
submitting rewritten version

@@
 
-```{r arules-ex, cache=TRUE, fig.cap='Frequency of association betwene items in the `Grocery` dataset', message=FALSE, echo=FALSE, fig.margin=TRUE}
+```{r arules-ex, cache=TRUE, fig.cap='Frequency of items in the `Grocery` dataset', message=FALSE, echo=FALSE, fig.margin=TRUE}
 library(knitr)
@@
 
-```{r arules-ex2}
+```{r arules-ex2, echo=FALSE}
 options(digits=2)
@@
 
-```{r tesco-ex, echo=FALSE, cache=TRUE, fig.cap ='Tesco surplus',  fig.margin=TRUE}
+```{r tesco-ex, echo=FALSE, cache=TRUE, fig.cap ='Tesco grocery items with more than 50 products',  fig.margin=TRUE}
     library(ggplot2)
@@
         aes(x=labels, y=actual, fill=labels)) +
+        xlab('Item count') + ylab('Grocery item') +
         geom_bar(stat='identity') + coord_flip() +
-        ggtitle('Tesco grocery items with more than 50 products') +
         theme(legend.position="none", axis.text=element_text(size=7))
 
+ggsave('figure/tesco_grocery_lines.svg')
+ggsave('figure/tesco_grocery_lines.png')
 ```

### 2017-10-06 title.rmd

@@
     word_document: 
-     reference_docx: reference_docx.docx  
 fontsize: 12pt

### 2018-02-02 log.rmd
starting major revisions

@@
 - figure numbering is still not right. Done that by hand. 
+- then had to delete a dual footnote when looking at the submission proofs.
+
+## Fri 02 Feb 2018 09:55:57 GMT
+- starting revisions of next version; reports on the 2nd version are here -- reports_revised.md 
+- will put revisions directly into reconstruction.rmd
+
+- revision letter
+
+I went through the three reports, making nearly all the changes suggested. Many of these were useful and I was happy to respond to them. 
+Here are the ones I didn't do:
+ - R1: For reasons of word count, I wasn't able to include a detailed discussion of the Tesco user interface.
+Given my focus on propensities and ordering of recommendations, I wasn't sure that was so relevant.
+Footnote 3 has a description of how handwritten lists and writing lists have been understood by anthropologists and literary theorists.
+ - R2:

### 2018-02-02 reconstruction.rmd
starting major revisions

@@
 
-'Data-driven personalization' relies on predictions, themselves dependent on data practices ranging across acquisition, storage, transformation, exchange, integration, modelling and experimentation.
+'Data-driven personalization' relies on predictions, themselves dependent on data practices ranging across acquisition, storage, transformation, exchange, integration, modelling and experimentation [@Goy_2007].
 The reconstruction of an online grocery shopping recommender system and its  personalized predictions that I undertake here centres on the notion of probability and its role in modelling and experimentation.
@@
 It makes use of archaeological approaches, in the sense developed by Michel Foucault [@Foucault_1972] to identify and map functional statements and practices configuring the knowledge and order generated by such systems over time.
-It conducts several small-scale code-based experiments in order to reconstruct, using widely available code resources resources such as API (Application Programmer Interfaces) and software libraries for machine learning, some prototypical elements of the system in question.
+It conducts several small-scale code-based experiments in order to reconstruct, using widely available code resources resources such as APIs (Application Programming Interfaces) and software libraries for machine learning, some prototypical elements of the system in question.
 Like some recent work in science and technology studies,  anthropology and media studies [@Marres_2017; @Marcus_2014; @Bogost_2012], the paper is  shaped by practical encounter with the 'object' -- code, data, web platforms --  it describes.
@@
 Against the background of the sheer number of commodities and their distribution of prices, Patel's presentation presents two opportunities.
-First, by recounting, contextualising and commenting on the main steps in making the 'Have your forgotten?' list, we might follow some of the predictive sense-making done by data scientists and customer analytics teams working with transactional data in a typical commercial setting.
+First, by recounting, contextualising and commenting on the main steps in making the 'Have you forgotten?' list, we might follow some of the predictive sense-making done by data scientists and customer analytics teams working with transactional data in a typical commercial setting.
 Patel mentioned many of these steps only fleetingly in the presentation, for they are largely taken for granted as part of predictive analytic practice.
@@
     library(ggplot2)
+    library(dplyr)
+    library(svglite)
     tesco = read.csv('data/reference_data/tesco_groceries.csv')
@@
     tesco$actual = as.integer(tesco$actual)
-    ggplot(tesco[tesco$actual>50,],
+    tesco_cropped <- tesco %>% select(actual, labels) %>% arrange(desc(actual)) %>% filter(actual> 50) %>%
+            mutate(labels = factor(labels, levels = labels[order(-actual)])) 
+    ggplot(tesco_cropped,
         aes(x=labels, y=actual, fill=labels)) +
+        geom_bar(stat='identity') + coord_flip() +
         xlab('Item count') + ylab('Grocery item') +
-        geom_bar(stat='identity') + coord_flip() +
-        theme(legend.position="none", axis.text=element_text(size=7))
+        theme(legend.position="none", axis.text=element_text(size=6))
 
@@
 
-Predictive models have become central technical elements in many big data conversion events because they offer a way of operationally calculating probabilities that incorporate more of the propensities -- to purchase a magnum of French champagne for instance -- inherent in situations.
+Predictive models have become central technical elements in many big data conversion events because they offer a way of operationally calculating probabilities that incorporate some of the propensities -- to purchase a magnum of French champagne for instance -- inherent in situations.
 In the new recommender system described by Patel, the business goal is to extend the list of the items customers have selected for purchase with  a few recommended items.
@@
 Individualizing personalization, however,  is ill-fit to the socially complex negotiations of grocery shopping, and tends to obscure other implications.
-Conversion events, both in the sense of the purchases made on the basis of recommendation and the narratives of changes in modelling practice related by data scientists, do not easily map to personalization.
+Conversion events,  in both the sense of the purchases made on the basis of recommendation and the narratives of changes in modelling practice related by data scientists do not easily map to personalization.
 They include relations running between people and things in time.
@@
 
+In the context of recommender systems, the constitutive incompleteness of shopping lists and their propensity to expand or change might be more important than their capacity to be personalized.
 I suggest that we might understand what precedes and exceeds personalization as probabilization.

### 2018-02-05 log.rmd
starting R2 changes on paper

@@
 
-I went through the three reports, making nearly all the changes suggested. Many of these were useful and I was happy to respond to them. 
-Here are the ones I didn't do:
- - R1: For reasons of word count, I wasn't able to include a detailed discussion of the Tesco user interface.
+I went through the three reports, making many of the changes suggested. Many of these were useful and I was happy to respond to them.
+
+An overview of changes and discussion of suggestions I wasn't able to implement:
+ - R1: 
+     -  For reasons of word count, I wasn't able to include a detailed discussion of the Tesco user interface.
 Given my focus on propensities and ordering of recommendations, I wasn't sure that was so relevant.
 Footnote 3 has a description of how handwritten lists and writing lists have been understood by anthropologists and literary theorists.
+     - R1 asked for literature on adaptive ecommerce systems to be included, which is fine (I included the references), but I didn't discuss this very much.
+Grocery shopping, as I point in the introduction differs in important ways from many typical e-commerce settings.
  - R2: 
+    - to be honest, this was a difficult report to respond to.
+It occasioned me a lot of effort to deal with what seems a report coming from a very different starting point.
+I have tried, but it would require much more scaffolding to deal with them all.
+    - R2 asked for 'specifics about how the analysis may translate to other personalization contexts, despite this being promised.' Although I did claim the case might be 'sociologically resonant,' I don't think I promised to translate to other personalization contexts.
+(Reading abstract, introduction and conclusion shows no sign of that promise.) It would be a significant new section of the paper to specify the general argument for other concrete contexts.
+    - R2 found the borrowing and repurposing of the term 'conversion event' from the recommender system literature confusing.
+I've added a footnote to explain the motivation for this borrowing, but I probably won't be able to convince R2 of the value of a material-semiotic approach in which such dual-register terms are quite common.
+    - R2 was concerned about extremely confusing methodology.
+I've added a sentence or two in the methodology section saying what the paper does.
+But I won't, I fear, be able to assuage R2's concerns.
+They don't countenance what I'm trying to do: ground a theoretical argument by assembling some empirical engagements -- the ethnographic event, the small experiments with Tesco APIs and recommender system models, the existing sociologies and anthropologies of shopping and lists and the archaeology of recommendation.
+Its not a conventional empirical research project, and the methodology of reconstruction that I use (borrowing from Dewey) is unlikely to satisfy narrowly defined sociological requirements.
+Given the methodological pluralism of the journal, I'm hoping that this is ok.
+    - I've added a footnote with URL for the Github repository for the paper and its scripts. 
+    - R2 wasn't able to see what the ethnographic material adds to analysis.
+They could refer to p3, li 44-48 (in pre-revision manuscript) where I set out how I understand it is as part of a 'big data conversion event.'
+    - R2 couldn't work out the 'intended meaning' of some key sentences. There was a grammar problem in these sentences, which I've fixed.
+
+## Mon 05 Feb 2018 08:07:19 GMT
+- responding to R2 today; 
+     - do I need a footnote on conversion event?
+
+     paragraph setting out the structure of the paper:
+
+     The argument develops in four main steps.
+The first main section introduces the case study -- the Tesco online grocery  recommender system as presented at an industry/academic conference.
+The second section frames grocery shopping in terms of sociological and anthropological accounts of shopping lists and social order, highlighting how both lists and shopping overflow individual identity.
+A brief archaeology of recommendation describes the shift from 1980s census-based data-mining to web-based recommender system at the turn of the century.
+The following section is the main body of the paper and approaches recommendations in terms of the different kinds of probability calculations that both target persons and somewhat overflow them.
+These range across calculations concerning the relation between products, between products and people,

### 2018-02-05 reconstruction.rmd
starting R2 changes on paper

@@
 With some forbearance on the part of readers, however, the argument of the paper should resonate sociologically.
-It concerns the relationality of social life, or the consistency and regularity of relating and acting amid a world in where probabilities and calculations based on probabilities are widespread.
+It concerns the relationality of social life, or the consistency and regularity of relating and acting in a world where probabilities and calculations based on probabilities are widespread.
 
@@
 While the changes Patel described are configured in Tesco-specific ways by DunnHumby, they are also broadly typical of a big data conversion event.
-Analytics service providers such as DunnHumby attempt to convert their customers by stories of conversion.[^65.1] Patel's presentation was part of this effort.
+Analytics service providers such as DunnHumby attempt to convert their customers by stories of conversion.[^6531] Patel's presentation was part of this effort.
 Tesco is the largest supermarket chain in UK and a notable success story for DunnHumby.
@@
 
-[^65.1] The web pages of the big data analytic companies providing products to various sectors of industry, government and media are stocked with images, graphics and statements centred on this conversion experience.
+[^6531] The web pages of the big data analytic companies providing products to various sectors of industry, government and media are stocked with images, graphics and statements centred on this conversion experience.
 [@Amoore_2015] analyses some of these.

### 2018-02-05 log.rmd
writing overview section

@@
 
-     paragraph setting out the structure of the paper:
+     paragraph setting out the structure of the paper: this should on page 2. Just after the Hansen section is re-inserted.
 
      The argument develops in four main steps.
-The first main section introduces the case study -- the Tesco online grocery  recommender system as presented at an industry/academic conference.
-The second section frames grocery shopping in terms of sociological and anthropological accounts of shopping lists and social order, highlighting how both lists and shopping overflow individual identity.
-A brief archaeology of recommendation describes the shift from 1980s census-based data-mining to web-based recommender system at the turn of the century.
+The first main section introduces the case study -- the Tesco online grocery  recommender system as presented at an industry/academic conference -- as an instance of a conversion event.
+The second section frames grocery shopping in terms of sociological and anthropological accounts of shopping lists and social order, highlighting how both the constitutional incompleteness of lists as inscriptive devices and the group-structured relationality of shopping overflow operational notions of personalization.
+The key elements of contemporary recommender systems are introduced by a brief archaeology of recommendation described in terms of a shift from 1980s census-based data-mining to web-based recommender system at the turn of the century.
 The following section is the main body of the paper and approaches recommendations in terms of the different kinds of probability calculations that both target persons and somewhat overflow them.
-These range across calculations concerning the relation between products, between products and people,
+This section of the paper extends Popper's propensity-based account of probability in its contemporary ramifications through predictive models, platform-scale data processing for large numbers of customers, the problems of grocery lists in their combinatorial potentials, and the supra-individual potentials of recommendations.

### 2018-02-06 log.rmd
entered changes in section A

@@
 
-I went through the three reports, making many of the changes suggested. Many of these were useful and I was happy to respond to them.
+I went through the three reports, making many of the changes suggested.
+Many of these were useful and I was happy to respond to them.
+The main change is the overview and summary of the paper.
+A new paragraph towards the end of the introduction sets out the argument step by step, and the steps of the argument are also reiterated in the conclusion.
 
@@
      -  For reasons of word count, I wasn't able to include a detailed discussion of the Tesco user interface.
-Given my focus on propensities and ordering of recommendations, I wasn't sure that was so relevant.
+Given my focus on probabilities, propensities and ordering of recommendations, I wasn't sure that was so relevant.
 Footnote 3 has a description of how handwritten lists and writing lists have been understood by anthropologists and literary theorists.
-     - R1 asked for literature on adaptive ecommerce systems to be included, which is fine (I included the references), but I didn't discuss this very much.
+     - R1 asked for computer science and Web science literature on adaptive ecommerce systems to be included, which is fine as part of the context (I included several new references suggested by R1 in the introduction), but I didn't discuss this very much for two reasons.
 Grocery shopping, as I point in the introduction differs in important ways from many typical e-commerce settings.
+Furthermore, because I was interested in the commercial operation of recommender systems, I wanted to focus on what was presented for discussion by the industry data scientist.
+Her presentation framed the sample of research literature I draw on.
+     - I edited the graphic as suggested
+     - I modified the conclusion as suggested
  - R2: 
     - to be honest, this was a difficult report to respond to.
-It occasioned me a lot of effort to deal with what seems a report coming from a very different starting point.
-I have tried, but it would require much more scaffolding to deal with them all.
+I could sympathise with the reviewers frustration, but I think the reviewer was approaching the manuscript from a quite different episteme.
+I have tried to respond constructively to their difficulties, but it would require much more scaffolding to deal with them all.
     - R2 asked for 'specifics about how the analysis may translate to other personalization contexts, despite this being promised.' Although I did claim the case might be 'sociologically resonant,' I don't think I promised to translate to other personalization contexts.
@@
 But I won't, I fear, be able to assuage R2's concerns.
-They don't countenance what I'm trying to do: ground a theoretical argument by assembling some empirical engagements -- the ethnographic event, the small experiments with Tesco APIs and recommender system models, the existing sociologies and anthropologies of shopping and lists and the archaeology of recommendation.
-Its not a conventional empirical research project, and the methodology of reconstruction that I use (borrowing from Dewey) is unlikely to satisfy narrowly defined sociological requirements.
-Given the methodological pluralism of the journal, I'm hoping that this is ok.
+They don't countenance what I'm trying to do: ground a theoretical argument about changing modes of operational knowledge by assembling some empirical engagements -- the ethnographic event, the small experiments with Tesco APIs and recommender system models, the existing sociologies and anthropologies of shopping and lists and the archaeology of recommendation.
+Its not a conventional empirical research project, and the methodology of reconstruction that I use (borrowing from Dewey) does not conform to narrowly defined sociological requirements.
+Given the methodological pluralism of the journal, I'm hoping that this is viable.
     - I've added a footnote with URL for the Github repository for the paper and its scripts. 
@@
 They could refer to p3, li 44-48 (in pre-revision manuscript) where I set out how I understand it is as part of a 'big data conversion event.'
-    - R2 couldn't work out the 'intended meaning' of some key sentences. There was a grammar problem in these sentences, which I've fixed.
+    - R2 couldn't work out the 'intended meaning' of some key sentences.
+There was a grammar problem in these sentences, which I've fixed.
+I'm hoping that this was the problem with those sentences not the actual content.
+    - I've moved key part of footnote 5 into the introduction. That was a useful suggestion. 
+    - I've shortened the historical-archaeology of recommender system, and explained its purpose in the introductory overview. 
+    - The section on limitations of association rule systems has been re-framed in the introduction/overview, and I've added some sentences at the beginning of the section explaining why attention to limitations is relevant to the main argument about probabilities.
+- R3:
+    - it was a relief to not have to do much in responding to this report. I was happy to expand the sentences about object-oriented research.
 
@@
 - responding to R2 today; 
-     - do I need a footnote on conversion event?
+- DONE: do I need a footnote on conversion event?
 
@@
 
-     The argument develops in four main steps.
+     As it explores the big data conversion event, the argument unfolds in four main steps.
 The first main section introduces the case study -- the Tesco online grocery  recommender system as presented at an industry/academic conference -- as an instance of a conversion event.
@@
 The following section is the main body of the paper and approaches recommendations in terms of the different kinds of probability calculations that both target persons and somewhat overflow them.
-This section of the paper extends Popper's propensity-based account of probability in its contemporary ramifications through predictive models, platform-scale data processing for large numbers of customers, the problems of grocery lists in their combinatorial potentials, and the supra-individual potentials of recommendations.
+This section of the paper extends Popper's propensity-based account of probability.
+It attends to the different kinds of probabilities subjected to calculation, the shifts in modelling practices that include many more propensities,  the different operational conditions under which probability calculations take place, and the implications of lists and shopping for probability.
+
+## Tue 06 Feb 2018 09:07:15 GMT
+- finished work on R2 problems 
+- TODO:  need to write a few sentences on the 'social argument' apropos Marres, Bogost, etc.
+
+> Let's draw a distinction: unlike tools and art, philosophical carpentry _is built with philosophy in mind_: it may serve myriad other productive and aesthetic purposes, breaking with its origins and entering into dissemination like anything else, but it's first constructed as a theory, or an experiment, or a question -- one that can be operated.
+Carpentry is philosophical lab equipment.
+[@Bogost_2012,100]
+
+- TODO:  add overview to the conclusion
+- TODO:  revise abstract to reflect argument
+- added some more quotes on reconstruction to ~/notes/dewey_reconstruction_1957.md
+- read through editors requests -- think I've done most of them
+- entered changes up to Section B

### 2018-02-06 reconstruction.rmd
entered changes in section A

@@
 
-Big data practice often converts an existing state of affairs by adding a model or predictive elements based on the collection and analysis of data.
+Big data practice often converts an existing state of affairs by adding a model or predictive elements based on the extraction and analysis of data.
 This paper recounts the conversion of  an online grocery ordering system  from an older demographically-based recommendation system to a 'personal relevance' model.
 The main argument of the paper concerns how  personalization occurs in big data 'conversion events.' The term 'conversion event' has a dual sense.
-In recommender system design and web ecommerce, it refers to those occasions when a customer, user or site visitor clicks a specific hyperlink or purchases something.
+In recommender system design and web ecommerce, it refers to those occasions when a customer, user or site visitor clicks a specific hyperlink or purchases something that has been suggested to them.
 I also use it to refer to the common before-after narrative forms that proponents, developers and advocates of recommender systems adopt in describing, promoting, explaining or otherwise talking about big data, analytics or predictive modelling practices.
-In recent research and opinion on recommendations, personalization has been seen as the goal of predictive modelling.
-Big data discourse in its promissory mode attributes potency to personalization: 'most important, using big data we hope to identify specific individuals rather than groups' [@Mayer-Schonberger_2013].
+The operational and narrative dimension of conversion events lie at the core of this paper.
+It may be that the two senses of the term -- the technical and the imaginary -- cannot be kept apart.
+
+In recent research and opinion on recommendations, personalization has been seen as the goal of predictive modelling [@ACM_2017].
+Big data discourse in its promissory mode attributes potency to personalization: 'most important, using big data we hope to identify specific individuals rather than groups' [@Mayer-Schonberger_2013], even if personalization has a much longer history in commerce and e-commerce in particular [@Goy_2007;@Sarwar_2000].
 Conversely, current debates about the problems of big data emphasize the need to protect people from personalization.
@@
 Analysis of internet filter bubbles and the growth of predictive platforms in industry and government  [@Pariser_2011; @ONeil_2016a] also centre on personalization.
-In this paper, I agree that a transformation or conversion event is occurring, but one which also, even in the midst of much ongoing personalization,  opens the possibility of conceiving of social ordering processes afresh.
+
+In this paper, I confirm that a transformation or conversion event is occurring, but one which also, even in the midst of much ongoing personalization in the interests of platform capitalism's extraction of profits [@Srnicek_2016],  opens the possibility of conceiving of social ordering processes afresh.
 Like Louise Amoore and Viola Piotukh's call for a new epistemology of population  [@Amoore_2015, 360], I will argue that contemporary recommender systems afford us the opportunity to re-conceptualise long-standing assumptions about social order and structure.
-The average everydayness and familiarity of the example here -- grocery shopping -- allows some first hand engagement with the messiness, the entanglements, and potentials of this conversion event in ways that other interesting settings -- social media for instance -- do not.
+The average everydayness and familiarity of the example I discuss -- online grocery shopping -- allows some first hand engagement with the messiness, the entanglements, and potentials of the conversion event in ways that other interesting settings -- social media for instance -- do not.
 
-'Data-driven personalization' relies on predictions, themselves dependent on data practices ranging across acquisition, storage, transformation, exchange, integration, modelling and experimentation [@Goy_2007].
-The reconstruction of an online grocery shopping recommender system and its  personalized predictions that I undertake here centres on the notion of probability and its role in modelling and experimentation.
+Contemporary 'data-driven personalization' relies on predictions, themselves dependent on data practices ranging across acquisition, storage, transformation, exchange, integration, modelling and experimentation.
+These are complex platform-scale systems, with many different elements.
+I reconstruct certain predictive elements of an online grocery shopping recommender system associated with probabilities and their role in modelling and experimentation.
+The reconstruction is a quasi-empirical philosophical undertaking aiming to construct an argument about transformations in probability rather than characterising the platform situation.
 Although probabilistic calculations of events have  long-standing importance in many settings (insurance, medicine, experimental and field sciences, operations research, risk analysis, engineering design, public health,  economic modelling, etc.), probabilities have taken on an altered operational function in digital platforms.
-They have become mundane part of the weave of infrastructures, transactions, and practices found in everyday life.
-In order to take the _operational_ reality of probabilities into account, we might understand probabilities in quasi-physicalist terms  defined by the philosopher of science Karl Popper as a 'world of propensities, as an unfolding process of realizing possibilities and of unfolding new possibilities' [@Popper_1990,19].
+They have become mundane parts of the weave of infrastructures, transactions, and practices found in everyday life.
+In order to take the _operational_ reality of probabilities into account, I  understand probabilities in quasi-physicalist terms  defined by the philosopher of science Karl Popper as a 'world of propensities, as an unfolding process of realizing possibilities and of unfolding new possibilities' [@Popper_1990,19].
 This operational-physicalist or sociotechnical notion of probability shifts our understanding of data analytics, especially in association with platforms [@Gillespie_2010], away from personalization.
@@
 
-The account is empirical in several respects.
-It departs from an ethnographic moment: being a member of the audience at an industry/academic conference where a data scientist was describing how personalized recommendations were created for online grocery orders.
-The paper  draws on anthropological and sociological research concerning shopping and list-making to situate the recommender system amongst the social ordering practices of shopping and lists.
+The media theorist Mark Hansen has recently used Popper's account to argue  that 'predictive analytics are discoveries of micrological propensities that are not directly correlated with human understanding and affectivity and that do not by themselves cohere into clearly identifiable events' [@Hansen_2015a, 111-112].
+Hansen links data to things via probabilities: 'whatever explanatory and causal value predictive analytics of large datasets have is, I suggest, ultimately rooted in this ontological transformation whereby probabilities are understood to be expressions of the actual propensity of things' [@Hansen_2015a, 120].
+Hansen presents big data as an 'ontological transformation' that deploys probabilities in an evermore closely woven and encompassing expression of animated, eventful, propensities of things.
+Predictive analytics discovers probabilistic calculation  as operating on the propensities or the mutable associative agencies of things.
+The pragmatic and indeed empirical  question for me is whether such transformations or 'conversion' in relations between probabilities and 'the propensities of things'  can be detected and articulated in the prosaic setting of shopping lists and recommender systems.
+
+In exploring  big data conversion events, the argument concerning probabilities unfolds in four main steps.
+The first main section introduces the case study -- the Tesco online grocery  recommender system as presented at an industry/academic conference -- as an instance of the narrative of a conversion event.
+The second section frames grocery shopping in terms of sociological and anthropological accounts of shopping lists and social order, highlighting how both the constitutional incompleteness of lists as inscriptive devices and the group-structured relationality of shopping overflow operational notions of personalization.
+The defining predictive operations of contemporary recommender systems are introduced by a brief archaeology of recommendation described in terms of a shift from 1980s census-based data-mining to web-based recommender system at the turn of the century.
+The last section is the main body of the paper and approaches recommendations in terms of the different kinds of probability calculations that both target persons and somewhat overflow them.
+This section of the paper extends Popper's propensity-based account of probability.
+It attends to the different kinds of probabilities subjected to calculation, the shifts in modelling practices that include many more propensities,  the different operational conditions under which probability calculations take place, and the implications of lists and shopping for probability.
+
+The reconstruction of the conversion event is empirical in several respects.
+Discussion departs from an ethnographic moment: being a member of the audience at an industry/academic conference where a data scientist was describing how personalized recommendations were created for online grocery orders.
+I draw on existing anthropological and sociological research concerning shopping and list-making to situate the recommender system amongst the social ordering practices of shopping and lists.
 It makes use of archaeological approaches, in the sense developed by Michel Foucault [@Foucault_1972] to identify and map functional statements and practices configuring the knowledge and order generated by such systems over time.
-It conducts several small-scale code-based experiments in order to reconstruct, using widely available code resources resources such as APIs (Application Programming Interfaces) and software libraries for machine learning, some prototypical elements of the system in question.
-Like some recent work in science and technology studies,  anthropology and media studies [@Marres_2017; @Marcus_2014; @Bogost_2012], the paper is  shaped by practical encounter with the 'object' -- code, data, web platforms --  it describes.
+It conducts several small-scale code-based experiments in order to reconstruct, using widely available code resources resources such as APIs (Application Programming Interfaces) and software libraries for machine learning, some prototypical elements of the system in question.[^41.1]  Like some recent work in science and technology studies,  anthropology and media studies [@Marres_2017; @Marcus_2014; @Bogost_2012], the paper is  shaped by practical encounter with the 'object' -- code, data, web platforms --  it describes.
+Code experiments, as Ian Bogost writes, can 'act as a theory, or an experiment, or question -- one that can be operated' [@Bogost_2012, 100].
+The motivation is both phenomenological -- to re-activate the sense of relationality at work in the model  -- philosophically reconstructive -- to ground an encounter with the troublesome power of recommendation through what John Dewey terms 'specific inquiries into specific structures' [@Dewey_1957, 199].
 At times, this mode of empirical philosophy may seem overly fixated on the technical minutiae of algorithms and recommender systems and inattentive to lived experience or practices.
-With some forbearance on the part of readers, however, the argument of the paper should resonate sociologically.
-It concerns the relationality of social life, or the consistency and regularity of relating and acting in a world where probabilities and calculations based on probabilities are widespread.
+With some forbearance on the part of readers, however, the argument of the paper should resonate sociologically: it concerns the relationality of social life, or the consistency and regularity of relating and acting in a world where probabilities and calculations based on probabilities are widespread.
+
+[^41.1]: Scripts, code and sample datasets used in the paper can be found at https://github.com/rian39/lists
 
@@
 
-At one of the many industry-meets-academia events occurring in data science-oriented higher education institutions in the UK, speakers from industry, government and commerce described their work with predictive models.[^11.1] Their narrative often followed the 'before-Big Data and after-Big data' conversion form.
-Shreena Patel, a PhD graduate in statistics and operations research,  works as a data scientist for DunnHumby, a well-known [customer science company ](https://www.dunnhumby.com/) [@dunnhumby_2017].
+At one of the many industry-meets-academia events occurring in data science-oriented higher education institutions in the UK, speakers from industry, government and commerce described their work with predictive models.[^11.1] Their narratives often followed the 'before-Big Data and after-Big data' conversion form.
+Shreena Patel, a PhD in statistics and operations research,  works as a data scientist for DunnHumby, a well-known [customer science company ](https://www.dunnhumby.com/) [@dunnhumby_2017].
 Her work at DunnHumby focuses on online grocery shopping at the supermarket chain Tesco.
@@
 While Popper's concept  of probability as propensity might seem remote from the concerns of online grocery shopping, it applies quite well to Patel's presentation of the personal relevance model and the ways in which it seeks to calculate probabilities of purchase.
-The reconstructions of the model I am about to discuss will explore the implications of this alternative account of probability for regular narratives of personalization.[^65.1]
-
-[^65.1]: The media theorist Mark Hansen has recently applied Popper's account to argue  that 'predictive analytics are discoveries of micrological propensities that are not directly correlated with human understanding and affectivity and that do not by themselves cohere into clearly identifiable events' [@Hansen_2015a, 111-112].
-Hansen's account differs from many widely shared views of big data.
-In the face of the so much personalization (and Patel's presentation exemplifies this), Hansen attributes a somewhat impersonalizing force to big data.
-If  much predictive practice attempts to elicit 'micrological propensities' from data, Hansen links data to things via probabilities: 'whatever explanatory and causal value predictive analytics of large datasets have is, I suggest, ultimately rooted in this ontological transformation whereby probabilities are understood to be expressions of the actual propensity of things' [@Hansen_2015a, 120].
-Hansen theorises big data or predictive analytics as an 'ontological transformation' that deploys probabilities in an evermore closely woven and encompassing expression of animated, eventful, propensities of things.
-Predictive analytics links probabilities as calculations to propensities or the mutable associative agencies of things.
-The practical and indeed empirical  question is whether such transformations or 'conversion' in relations between probabilities and 'the propensities of things'  can be detected and articulated in the prosaic setting of shopping lists and recommender systems.
-'Conversion' is a preferable term for these changes.
-Not much hinges on the choice of terms, but 'conversion' happens to be the term used by Patel.
-It highlights  re-orientations in subjects, experience, things, numbers and infrastructures that ranges more widely and are more grounded than Hansen's 'ontological transformation' formulation.
+The reconstructions of the model I am about to discuss will explore the implications of this alternative account of probability for regular narratives of personalization.
 

### 2018-02-07 log.rmd
entered changes in section B,C,D

@@
     - I've shortened the historical-archaeology of recommender system, and explained its purpose in the introductory overview. 
-    - The section on limitations of association rule systems has been re-framed in the introduction/overview, and I've added some sentences at the beginning of the section explaining why attention to limitations is relevant to the main argument about probabilities.
+    - The section on limitations of association rule systems has been re-framed in the introduction/overview, and I've changed sentences at the close of the section explaining why attention to limitations is relevant to the main argument about probabilities.
 - R3:
@@
 - entered changes up to Section B
+
+## Wed 07 Feb 2018 09:53:49 GMT
+- starting section B edits. Not too much to do here.  
+- Section C -- only made a few changes; 
+- Section D -- thinking I should combine this into the main section E? no, I didn't. But did do a bit of reframing
+- Reached Section E (v) - platform experiments

### 2018-02-07 reconstruction.rmd
entered changes in section B,C,D

@@
 
-In recent research and opinion on recommendations, personalization has been seen as the goal of predictive modelling [@ACM_2017].
+In recent research and opinion on recommendations, personalization has been seen as the goal of predictive modelling.
+A glance at the conference proceedings of the annual ACM RecSys conference will show many papers with 'personalization' in the title [@ACM_2017].
 Big data discourse in its promissory mode attributes potency to personalization: 'most important, using big data we hope to identify specific individuals rather than groups' [@Mayer-Schonberger_2013], even if personalization has a much longer history in commerce and e-commerce in particular [@Goy_2007;@Sarwar_2000].
@@
 This operational-physicalist or sociotechnical notion of probability shifts our understanding of data analytics, especially in association with platforms [@Gillespie_2010], away from personalization.
-It  draws  on Popper's re-framing of probability as propensity to  engage with grocery recommendations in their worldly enactments.
+I draw  on Popper's re-framing of probability as propensity to  engage with grocery recommendations in their worldly enactments.
 
@@
 The pragmatic and indeed empirical  question for me is whether such transformations or 'conversion' in relations between probabilities and 'the propensities of things'  can be detected and articulated in the prosaic setting of shopping lists and recommender systems.
+If it can, then the narratives of personalization that frame many big data conversion events might be recounted differently.
 
@@
 The second section frames grocery shopping in terms of sociological and anthropological accounts of shopping lists and social order, highlighting how both the constitutional incompleteness of lists as inscriptive devices and the group-structured relationality of shopping overflow operational notions of personalization.
-The defining predictive operations of contemporary recommender systems are introduced by a brief archaeology of recommendation described in terms of a shift from 1980s census-based data-mining to web-based recommender system at the turn of the century.
-The last section is the main body of the paper and approaches recommendations in terms of the different kinds of probability calculations that both target persons and somewhat overflow them.
+The defining predictive operations of contemporary recommender systems are situated by a brief archaeology of Tesco's  shift from census-based data-mining to web-based recommender system.
+The last section is the main body of the paper and approaches recommendations in terms of the different kinds of probability calculations that both target persons and diffuse beyond them.
 This section of the paper extends Popper's propensity-based account of probability.
@@
 
+The reconstructions of the model will explore the implications of this alternative account of probability for regular narratives of personalization.
 The reconstruction of the conversion event is empirical in several respects.
 Discussion departs from an ethnographic moment: being a member of the audience at an industry/academic conference where a data scientist was describing how personalized recommendations were created for online grocery orders.
-I draw on existing anthropological and sociological research concerning shopping and list-making to situate the recommender system amongst the social ordering practices of shopping and lists.
+It refers to existing anthropological and sociological research concerning shopping and list-making to locate recommender system amongst the social ordering practices of shopping and lists.
 It makes use of archaeological approaches, in the sense developed by Michel Foucault [@Foucault_1972] to identify and map functional statements and practices configuring the knowledge and order generated by such systems over time.
@@
 Code experiments, as Ian Bogost writes, can 'act as a theory, or an experiment, or question -- one that can be operated' [@Bogost_2012, 100].
-The motivation is both phenomenological -- to re-activate the sense of relationality at work in the model  -- philosophically reconstructive -- to ground an encounter with the troublesome power of recommendation through what John Dewey terms 'specific inquiries into specific structures' [@Dewey_1957, 199].
+The motivation is both phenomenological -- to re-activate the sense of relationality at work in the model  -- and philosophically reconstructive -- to ground an encounter with the troublesome power of recommendation in what John Dewey terms 'specific inquiries into specific structures' [@Dewey_1957, 199].
 At times, this mode of empirical philosophy may seem overly fixated on the technical minutiae of algorithms and recommender systems and inattentive to lived experience or practices.
-With some forbearance on the part of readers, however, the argument of the paper should resonate sociologically: it concerns the relationality of social life, or the consistency and regularity of relating and acting in a world where probabilities and calculations based on probabilities are widespread.
+With some forbearance on the part of readers, however, the argument of the paper should resonate sociologically: it concerns the relationality of social life, or the consistency and regularity of relating and acting in a world where probabilities and operations based on probabilities are widespread.
 
@@
 What stands out from the presentation is not any state-of-the-art innovation, but the ongoing life of the recommender system: the new predictive model is part of a long-standing and ongoing transformation of grocery shopping.
-From a sociological standpoint, the interest lies less in specific technical innovations and more on how the ordering work done by recommender systems relates to the role of prediction in social order more generally.
+From a sociological standpoint, the interest lies less in specific technical innovations and more in how the various facets of the recommender systems she describes relate to the role of prediction and probabilities in social order more generally.
 
@@
 
-Grocery shopping is difficult to personalize.
+Compared to music, film, travel, fashion and book purchases, grocery shopping is difficult to personalize.
 The anthropologist Daniel Miller has argued [@Miller_2012] that shopping negotiates discrepancies between normative and actual social order (for instance, between the ideals of health emblematised by organic products and commitments to thrift embodied in lower cost generic products).
-Miller suggests that shopping practices attempt to resolve difference between how people think they should live and how they actually live: 'we have to watch how shopping helps resolve these discrepancies between the normative and the actual, but we also need some ideas of where the normative comes from in the first place' [@Miller_2012, 72].
+Miller suggests that household shopping practices attempt to resolve differences between how people think they should live and how they actually live: 'we have to watch how shopping helps resolve these discrepancies between the normative and the actual, but we also need some ideas of where the normative comes from in the first place' [@Miller_2012, 72].
 More provocatively, Miller claims that 'shopping is largely a technology for the expression of love' (85).
@@
 
-The main narrative of Patel's conversion narrative concerned the shift from a well-established loyalty-card based data-mining model developed in the 1990s to a predictive, probabilistic, 'personal relevance' model that would  append items to the shopping list in almost-realtime.
-While the changes Patel described are configured in Tesco-specific ways by DunnHumby, they are also broadly typical of a big data conversion event.
+Although the changes Patel described are configured in Tesco-specific ways by DunnHumby, they are also broadly typical of a big data conversion event.
 Analytics service providers such as DunnHumby attempt to convert their customers by stories of conversion.[^6531] Patel's presentation was part of this effort.
+The main narrative of Patel's conversion narrative concerned Tesco's shift from a well-established loyalty-card based data-mining model developed in the 1990s to a predictive, probabilistic, 'personal relevance' model that would  append items to the shopping list in almost-realtime.
 Tesco is the largest supermarket chain in UK and a notable success story for DunnHumby.
@@
 In  [@Humby_1989], Humby highlights the need to add lifestage data to neighbourhood data in marketing research.
-While Tesco  succeeded in data-mining its customers using demographic segmentation,  and perhaps became the UK's biggest supermarket with the help of data-mining in the 1990s, the shopping environment  in 2017 is markedly different [@Turow_2017].
+Even if Tesco  succeeded in data-mining its customers using demographic segmentation,  and perhaps became the UK's biggest supermarket with the help of data-mining in the 1990s, the shopping environment  in 2017 is markedly different [@Turow_2017].
 It is no longer organised around campaigns involving special offers or redemption of points for demographically segmented loyal customers (DunnHumby made heavy use of UK Census data).
@@
 Given personalized recommendations stretch back two decades, what in  the newly implemented  recommender system changes?
-The components of the new recommender system -- predictive models and their parameters, infrastructural provisioning to run the models, product similarity matrices --  address the challenges of personally-relevant recommendations -- too many products, changes over time, unstable propensities of things in their associations with people.
-The 'personal relevance model' and personalization more generally have a troubled relation to a social order because it is, as grocery shopping attests, intrinsically social.
+The components of the new recommender system -- predictive models and their parameters, infrastructural provisioning to run the models, and platform-scale deployment --  address the challenges of personally-relevant recommendations at a specific point in time, the moment when a customer is close to finishing their grocery order.
+The recommender system juggles   many products, changes over time, and unstable propensities of things in their associations with people.
+The 'personal relevance model' more generally has a troubled relation to a social order because grocery buying, as accounts of shopping suggests, plurally social.
 The model will need to render grocery shopping calculable in a way that somehow includes the associated social ordering and negotiation.
+
 The model, like many big data practices, predicates probabilities as a way to render the situation calculable.
-
 From the standpoint of probability, recommendations are conditional probabilities or probabilities whose calculation takes into the account the occurrence of other events.
 But what is a probability today?
-In exploring Patel's account of the recommender system, my focus probabilities seeks to re-frame the operation of the recommender system in a way freed from lingering incompatibilities between calculation and social life [@Dewey_1957, 26].
-In order to re-frame probability, I draw directly on the work of Karl Popper.
+In exploring Patel's account of the recommender system, any  account of probabilities needs to re-frame the operation of the recommender system in a way freed from lingering incompatibilities between calculation and social life [@Dewey_1957, 26].
+In re-framing probability, I draw directly on the work of Karl Popper.
 In an essay written towards the end of his career, Popper presents a non-standard account of probabilities as real processes.
-He argues that probabilities have a physical reality equivalent to forces and fields [@Popper_1990].
+He argues that probabilities have a  reality equivalent to forces and fields in physics [@Popper_1990].
 Against standard interpretations,  Popper does not identify probabilities with either degree of belief (likelihood) or frequency of events.
@@
 In his account, probabilities are tendencies towards realization inherent in a situation.
-Probabilities have or indeed are _propensities_, tendencies to realize the event (11).
+Probabilities express or indeed _are_ propensities, tendencies to realize the event (11).
 While Popper's concept  of probability as propensity might seem remote from the concerns of online grocery shopping, it applies quite well to Patel's presentation of the personal relevance model and the ways in which it seeks to calculate probabilities of purchase.
-The reconstructions of the model I am about to discuss will explore the implications of this alternative account of probability for regular narratives of personalization.
 
@@
 Probability has been difficult to work with philosophically according to Popper because  dice rolls, coin tosses, urns with balls and other seemingly random events have occupied centre stage.
-Even dice-rolling and coin-tossing has been enormously productive and transformative in scientific thought and practice, it privileges _absolute_ probability at the expense of _conditional_ probabilities.
+Although dice-rolling and coin-tossing has been enormously productive and transformative in scientific thought and practice, it privileges _absolute_ probability at the expense of _conditional_ probabilities.
 'We need,' Popper urges, '_a calculus of relative or conditional probabilities_ as opposed to _a calculus of absolute probabilities_' [@Popper_1990, 16].
 Relative or conditional probabilities are propensities that depend on other events for their own realization.
-All events require other events, so all probabilities are conditional, even if probability calculations typically abstract or ignore the inevitable physical conditioning of their realization.
+All events require other events, so all probabilities are conditional, even if probability calculations typically abstract or ignore the inevitable physical conditioning of their realization.[^102.1]
+
+[^102.1]: In [@Mackenzie_2014b], I develops an account of conditional probabilities in the context of epidemiologies of contagious disease.
+That paper also describes a transformation in probability, but one that is linked to problems of contagion.
 
@@
 
-The notion of conditional probability at work in `aprior` is relatively simple, and assumes that the frequency of transactions provides the best guide to what shoppers are likely to buy.
+The notion of conditional probability at work in `aprior` is relatively simple, and assumes that the frequency of co-occurring items provides the best guide to what shoppers are likely to buy.
 The `apriori` algorithm finds sets of items that commonly occur together in transactions.
-In this sense, is still oriented by a notion of absolute probability, inflected by some elements of conditional probability.
+In this sense, is still oriented by a notion of absolute probability, inflected by some elements of conditional probability concerning the relations between things.
 Commonly occurring sets are expressed as 'association rules.' For instance, applied to a dataset of generic, unbranded grocery purchases, the `apriori` algorithm counts frequencies of purchase in the overall set of all items purchased in a supermarket (the `Groceries` dataset  was acquired from a 'local German supermarket' [@Hahsler_2006]).
@@
 Although the `apriori` model does not use the language of probability calculus, the 'rules,' a term derived from older decision support literature, rank associations between items.
-The resulting set of rules (shown in Table \@ref(tab:arules-ex2)) indicates that milk and vegetables has a stronger association than milk and buns.
+The resulting set of rules (shown in Table \@ref(tab:arules-ex2)) indicates that milk and vegetables have a stronger association than milk and buns.
 The calculations of `support` (how frequently the association appears in the dataset), `confidence` (how often the rule applies in the dataset)  and   `lift` (a ratio between the `support` and the independent frequencies of the items) in the table attempt to measure the strength of these associations in different ways.
@@
 A simple illustration of the combinatorial problem faced by recommender system can be drawn by bringing the `Grocery` dataset together with the actual list of items that Tesco sells online.
-If we take all the items in the `Grocery` dataset and paste them into the 'shopping list' box on the Tesco grocery website (or as I did, run them as searches on the TescoLab Product API [@Tesco_2016]), each of the 169 generic items in the `Grocery` dataset yields dozens and sometimes thousands of products in the Tesco inventory.
+If we take all the items in the `Grocery` dataset and paste them into the 'shopping list' box on the Tesco grocery website (or as I did, run them as searches on the TescoLab Product API [@Tesco_2016]), each of the 169 generic items in the `Grocery` dataset matches dozens and sometimes thousands of products in the Tesco inventory.
 
@@
 Items in the `Groceries` dataset proliferate into a Tesco's list of branded products.
-The 169 items of the `Groceries` dataset  expand into roughly `r total` Tesco items (see Figure \@ref(fig:tesco-ex)).
+The 169 items of the `Groceries` dataset  expand into roughly `r total` Tesco items (see Figure 3).
 Recommender systems confront, I would suggest, a logistical proliferation of commodities.
@@
 
-More importantly, association rules limit the conditioning of probabilities to counting frequency of co-occurrence of items in shopping baskets.
-Even if the association rules provide recommendable sets of frequency-weighted associations, an `apriori`-based recommender system has no way of individualising its recommendations.
+More importantly, the combinatorial proliferation of association rules suggest why personalization or a 'personal relevance' model might become relevant.
+Even if the association rules provide recommendable sets of frequency-weighted associations, an `apriori`-based recommender system has no way of narrowing its recommendations.
 Its probabilisation of shopping is incomplete since it only works on associations between things.
-The tendency of some milk to find itself in a shopping basket alongside bread attests to an ordinary propensity in certain parts of the world.
+The tendency of  milk to find itself in a shopping basket alongside bread attests to an ordinary propensity in certain parts of the world.
 The conditional probabilities implicit in the association rules  do not, however, include much of the world.
@@
 
-Predictive models have become central technical elements in many big data conversion events because they offer a way of operationally calculating probabilities that incorporate some of the propensities -- to purchase a magnum of French champagne for instance -- inherent in situations.
+Predictive models have become central technical elements in many big data conversion events because they offer a way of operationally calculating probabilities that narrow the propensities -- to purchase a magnum of French champagne for instance -- inherent in situations.
 In the new recommender system described by Patel, the business goal is to extend the list of the items customers have selected for purchase with  a few recommended items.
-In order to do this, the list of items selected for purchase will be extended by recommendations that have, according to a predictive model, the most chance of 'conversion' or actual purchase.
+In order to achieve this seemingly modest goal, the list of items selected for purchase will be extended by recommendations that have, according to a predictive model, the most chance of 'conversion' or actual purchase.
 Most recommender system designers assume that modelling 'personal relevance' is the best way to do this.
@@
 
+Persons are strangely remote from such models.
 Patel introduced the new 'personal relevance model' with a data graphic familiar to machine learners and statistical modellers.
-(A sketch of her graph appears in Figure \@ref(fig:models).)   Patel assumed that the audience understood the working of the logistic regression model, and the bulk of her presentation concerned the obstacles and problems that arise in trying to personalize recommendations in ways that lead to the much-desired 'conversion events' or sales.
+(A sketch of her graph appears in Figure 4.)   Patel assumed that the audience of data scientists understood the working of  logistic regression, association rules and random forest models, and the bulk of her presentation concerned the obstacles and problems that arise in trying to personalize recommendations in ways that lead to the much-desired 'conversion events' or sales.
 The graph plots the _precision_ -- the proportion of the recommended products that customers actually purchase -- for several different statistical models.
@@
 Patel used the graph to compare the previous rule-based recommender systems with some of the 'personal relevance model' alternatives -- logistic regression, random forests, gradient boosting and a few others -- in terms of their predictions and how those predictions turned out.
-Patel dismissed most of the models quite quickly and focused only on one, the logistic regression model, which did as well or slight other than other alternatives.
+Patel dismissed most of the models quite quickly and focused only on one, the logistic regression model, which did as well or slight better than the alternatives.
 
@@
 A probability greater than `0.5` typically would be `recommended.`  The logistic regression model generates probabilities of purchase for each product for each customer.
-The fundamental  shift  here is that the classifier model extends the reach of  the recommendations generated for a customer to not only  the 250,000 items in the Tesco inventory but to any item, relationship, similarity or event that can be constructed as a variable in the classifier model.
-In this sense, the recommendations shown to a customer in the 'Have You Forgotten' list derive from an intricate conditional probability statement, woven from, as we will soon see, an open-ended field of relations.
+The fundamental  shift  from the rules-based model is that the classifier model extends the reach of  the recommendations generated for a customer to not only  the 250,000 items in the Tesco inventory but to any item, relationship, similarity or event that can be constructed as a variable in the classifier model.
+In this sense, the recommendations shown to a customer in the 'Have You Forgotten' list derive from a much more extended conditional probability statement, woven from, as we will soon see, an open-ended field of relations ranging beyond associations between grocery items.
 
@@
 
-I will not discuss all the dimensions of probabilisation implicit in the architecture of the new model but highlight those elements that trouble the notion of 'personal relevance' and personalization more generally.
-These include the problem of repetition and temporality of calculation,  experiments on the platform,  and attempts to create new views of the data.
+I will not discuss all the dimensions of probabilisation implicit in the architecture of the new model but highlight those elements that appear as part of the description of 'personal relevance' yet remain irreducible to personalization.
+These include the problem of repetition and temporality of calculation,  the platform as experimental site,  and attempts to create new views of the data.
 
@@
 Popper's notion of probabilities as tendencies to realization inherent in a situation implies this.
-As Popper observes, 'the propensities cannot be measured because the relevant situation changes and cannot be repeated' [@Popper_1990, 17].
+In many cases, as Popper observes, 'the propensities cannot be measured because the relevant situation changes and cannot be repeated' [@Popper_1990, 17].
 Measuring propensities and expressing these measurements as numbers between `0` and `1` becomes difficult because as a situation changes,  the propensities themselves change.
@@
 
-One way in which recommender systems attempt to address the problem of the relational dynamics of propensities is by assuming that purchases will be repeated.
+One way in which recommender systems might attempt to address the problem of the relational dynamics of propensities is by assuming that purchases will be repeated.
 Patel mentioned that the new model uses '52 weeks of data' for each customer.
 In including this data in the model, the assumption is that the probability that a specific customer will buy a specific product will be increased by previous purchases of that product.
-The history of previous purchases constitutes forms of repetition that imply a stabilisation of propensities (high or low; e.g.
+The history of previous purchases constitutes forms of repetition that imply a stabilisation of propensities (e.g.
 a vegan customer will never have purchased chicken products, so the measured propensities for any of the 1000 or so chicken products sold be Tesco will remain close to `0`).
-But the changing situation of the customer only figures here in the accumulated year of purchase data.
+But the changing situation of the customer only figures here in the accumulated year of purchase data, and that past purchases might precisely what does not need to recommended.
 
@@
 Patel's quick gloss of the infrastructural deployment of DunnHumby's relevance model is  primary to the conversion event: the logistic regression model at the heart of the recommender system is no longer an analytical device but an operational one because it seeks to revise recommendations as situations change.
+
+All of these considerations -- the problem of a changing customer, the need to undersample 'unbalanced data' for the predictive model to work, the energy and computational time costs of running models, as well as the technical complexity of staging predictions for an online platform -- form part of big data conversion events, as they concretely and operationally take place.
+They do not belong to the data or predictions as such, but to the situation in which recommendations might become purchases.
 

### 2018-02-08 log.rmd

@@
 A new paragraph towards the end of the introduction sets out the argument step by step, and the steps of the argument are also reiterated in the conclusion.
+The main  section of the paper -- E -- now has a summary at the end, as well as various signposts along the way.
+Throughout these revisions, I've tried to be precise and focused about the argument of the paper: the nature of conversion events.
 
@@
      -  For reasons of word count, I wasn't able to include a detailed discussion of the Tesco user interface.
+This would be slightly tricky given web and app variations.
 Given my focus on probabilities, propensities and ordering of recommendations, I wasn't sure that was so relevant.
@@
  - R2: 
-    - to be honest, this was a difficult report to respond to.
+    - to be honest, this was a difficult report to fully deal with.
 I could sympathise with the reviewers frustration, but I think the reviewer was approaching the manuscript from a quite different episteme.
@@
 I've added a footnote to explain the motivation for this borrowing, but I probably won't be able to convince R2 of the value of a material-semiotic approach in which such dual-register terms are quite common.
-    - R2 was concerned about extremely confusing methodology.
+    - R2 was concerned about the confusing methodology.
 I've added a sentence or two in the methodology section saying what the paper does.
-But I won't, I fear, be able to assuage R2's concerns.
+But I won't, I fear, be able to assuage all R2's concerns.
 They don't countenance what I'm trying to do: ground a theoretical argument about changing modes of operational knowledge by assembling some empirical engagements -- the ethnographic event, the small experiments with Tesco APIs and recommender system models, the existing sociologies and anthropologies of shopping and lists and the archaeology of recommendation.
-Its not a conventional empirical research project, and the methodology of reconstruction that I use (borrowing from Dewey) does not conform to narrowly defined sociological requirements.
+It's not a conventional empirical research project, and the methodology of reconstruction that I use (borrowing from Dewey) does not conform to narrowly defined sociological requirements.
 Given the methodological pluralism of the journal, I'm hoping that this is viable.
@@
 - R3:
-    - it was a relief to not have to do much in responding to this report. I was happy to expand the sentences about object-oriented research.
+    - it was a relief to not have to do much in responding to this report. I was happy to expand the sentences about object-oriented research as suggested in the methods section.
 
@@
 - Reached Section E (v) - platform experiments
+
+## Thu 08 Feb 2018 11:24:54 GMT
+- will do rest of Section E and conclusion today. Submit tomorrow?  
+- wrote new summary for section E:
+
+>The different encounters with the data staged in the personal relevance model cannot be reduced to any simple probabilistic calculation.
+Rather, as Patel's presentation of the different models, the logistics of running the models in a platform setting, the experimental validation of the model in A/B testing, and the construction of new features that sought to increase complementarities and reduce similarities suggests, the expression of propensities occurs in a changing weave of infrastructure, mathematics, history and logistics.
+None of these has a particularly strong relation to personhood or individual experience, even if personalization the main way this weave is figured.
+
+- revised conclusion

### 2018-02-08 extract_slides.sh
cleaning up master branch

@@ -1,2 +1,0 @@
-#!/bin/bash
-grep '\(#\{1,2\}\)\|!\[\|> ' lists_conversion_experience.md > slides.md 

### 2018-02-08 reconstruction.rmd
rendering gh-pages version of major revision

@@
 ```{r humby-1984, fig.cap = 'Abstract from a Clive Humby presentation (Humby,1984)'}
+        library(knitr)
         include_graphics('figure/2_humby_young_org_1994.png')
@@
 
-In his account of probabilities as physical propensities, Popper emphasises why laboratory experiments are important to probability:
+In his account of probabilities as physical propensities, Popper emphasises why laboratory experiments are important:
 
@@
 A crucial consideration for  a recommender system designer will be how to experimentalize the systems recommendations so that they are not disturbed by other propensities.
-This is an inordinately difficult challenge, since, as we have seen, grocery shopping takes place at the intersection of normative and actual social orders, individual and group belongings, and logistically global economic processes.
+This is an inordinately difficult challenge in some settings, since, as we have seen, grocery shopping takes place at the intersection of normative and actual social orders, individual and group belongings, and logistically global economic processes.
+In such situations, propensities are inherently labile.
 
@@
 
-Experimentalization, the practice of creating conditions that reduce disturbing propensities, runs deep in the conversion event.
+Experimentalization, the practice of creating conditions that reduce disturbing propensities, runs deep in big data conversion events.
 The predictions of the recommender system themselves are the subject of experiment.
@@
 
-Whatever disturbing factors or interferences the experimental trials of different predictive models reduce, another stream of propensities runs away from the goal of personal relevance.
-Popper suggest that the realization of tendencies is open-ended:
+Whatever disturbing factors or interferences the experimental trials of different predictive models reduce, another stream of propensities runs off the terrain of personal relevance.
+Popper suggest that the realization of tendencies is inevitably open-ended:
 
@@
 
-The main efforts that DunnHumby made to  construct good features were not closely focused on individuals  but sought to include  relations between things in the predictive model.
+The main efforts that DunnHumby made to  construct good features were not closely focused on individuals  but sought to address relations between things in the predictive model.
 Patel, for instance, described the problem of 'basket similarity.' The recommender system should not recommend items that are too similar to groceries already in a customer's basket.
@@
 
+In a world of propensities, the different encounters with the data staged in the personal relevance model cannot be reduced to any simple probabilistic calculation.
+Rather, as Patel's presentation of the different models, the logistics of running the models in a platform setting, the experimental validation of the model in A/B testing, and the construction of new features that sought to increase complementarities and reduce similarities suggests, the expression of propensities occurs in a changing weave of infrastructure, mathematics, history and logistics.
+None of these has a particularly strong relation to personhood or individual experience, even if personalization the main way this weave is figured.
+
 [^215]: The proliferation of variables or features in the model itself introduces new instabilities, new 'possibilities trying to realize themselves' and the DunnyHumby data scientists used  '`L1 regularization` to drive coefficients down to zero', or, put more simply, to exclude variables there were not contributing much to the predictions produced by the model.
@@
 The personalization of recommendations has been a distinctive feature of big data conversion narratives and practice.
-The Tesco recommender system, starting from its early experiments with demographic data-mining, its later adoption of a rule-based 'market basket' analysis system, and its recent implementation of a machine learning 'personal relevance model' follows the trajectory of personalization.
+The Tesco recommender system, starting from its early experiments with demographic data-mining, its later adoption of a rule-based 'market basket' analysis system, and its recent implementation of a machine learning 'personal relevance model' documents the trajectory of a personalizing conversion event.
 Its recent big data conversion event largely takes the form of personalization.
@@
 They include relations running between people and things in time.
-They include an ongoing process of adjusting, scaffolding, intervening and configuring orders.
-Only because grocery shopping is  personally pre-ordered by lists can a predictive recommendation, provisionally conducive to new conversion events, gain traction.
+They include an ongoing process of adjusting, scaffolding, intervening and configuring orders -- in several senses of that term -- that range across supply chain management.
 
-In the context of recommender systems, the constitutive incompleteness of shopping lists and their propensity to expand or change might be more important than their capacity to be personalized.
-I suggest that we might understand what precedes and exceeds personalization as probabilization.
-The framing of this conversion as a 'flat difference of degree, such that it appears as though everything is calculable' [@Amoore_2015, 361] risks misunderstanding the different propensities tending to realization.
-If, as Popper argues, a propensity-based account of probabilities 'amounts to generalizing and extending the idea of forces again' [@Popper_1990,14], then we should see the compute  platforms, databases, software libraries, web interfaces and global supply chain logistics as part of this conversion event, as  tendencies or propensities in  the process of realization.
+I have suggested that we might understand recommender system as part of the ongoing operationalization of propensities described by Popper.
+The conversion described in this case is a propensity in actualisation, but a propensity that has the character of expressing propensities through predictions and recommendations.
+From the perspective of a world of propensities, the constitutive incompleteness of shopping lists and their propensity to expand or change might be more important than their capacity to be personalized.
+Only because grocery shopping is  personally pre-ordered by lists can a predictive recommendation, somewhat conducive to conversion events, gain traction.
+
+I suggest that we might understand what surrounds and exceeds personalization as probabilization.
+The framing of prediction as a 'flat difference of degree, such that it appears as though everything is calculable' [@Amoore_2015, 361] affirms the importance of calculation but risks misunderstanding the how propensities tend towards realization, and how models express those tendencies.
+If, as Popper argues, a propensity-based account of probabilities 'amounts to generalizing and extending the idea of forces again' [@Popper_1990,14], then we should see the compute  platforms, databases, software libraries, various predictive models, web interfaces, apps and global supply chain logistics as part of this conversion event, as  tendencies or propensities in  the process of realization.
 When we track what is practically done to construct a personal relevance model and apply it to a shopping list, we see conditional probabilities assembled in predictive models, but also in the rhythms of  platform and infrastructural configurations and experiments that gather and process data,  and in the open-ended construction of features in data.

### 2018-02-08 render_tufte.r

@@ -1,4 +1,5 @@
 library(rmarkdown)
 library(bookdown)
-library(tufte)
-render('lists_conversion_experience.rmd',output_file='index.html', output_format=html_document2(toc=TRUE, toc_depth = 1, toc_float=FALSE, theme='journal'))
+library(tint)
+render('reconstruction.rmd',output_format=html_document(), 'index.html')
+

### 2018-04-16 reconstruction.rmd
submitting final revision with minor revisions requested

@@
 The main argument of the paper concerns how  personalization occurs in big data 'conversion events.' The term 'conversion event' has a dual sense.
-In recommender system design and web ecommerce, it refers to those occasions when a customer, user or site visitor clicks a specific hyperlink or purchases something that has been suggested to them.
+In recommender system design and web e-commerce, it refers to those occasions when a customer, user or site visitor clicks a specific hyperlink or purchases items displayed to them.
 I also use it to refer to the common before-after narrative forms that proponents, developers and advocates of recommender systems adopt in describing, promoting, explaining or otherwise talking about big data, analytics or predictive modelling practices.
 The operational and narrative dimension of conversion events lie at the core of this paper.
-It may be that the two senses of the term -- the technical and the imaginary -- cannot be kept apart.
+It may be that the two senses of the term -- the technical and the figurative -- cannot be kept apart.
 
@@
 What stands out from the presentation is not any state-of-the-art innovation, but the ongoing life of the recommender system: the new predictive model is part of a long-standing and ongoing transformation of grocery shopping.
+Her presentation is the basis for this discussion: it shapes the choice of technical literature, code and data samples, as well as the consideration of underlying philosophical issues.
 From a sociological standpoint, the interest lies less in specific technical innovations and more in how the various facets of the recommender systems she describes relate to the role of prediction and probabilities in social order more generally.
@@
 
-Compared to music, film, travel, fashion and book purchases, grocery shopping is difficult to personalize.
-The anthropologist Daniel Miller has argued [@Miller_2012] that shopping negotiates discrepancies between normative and actual social order (for instance, between the ideals of health emblematised by organic products and commitments to thrift embodied in lower cost generic products).
+Compared to music, film, travel, fashion and book purchases, online grocery shopping is difficult to personalize.
+The anthropologist Daniel Miller has argued [@Miller_2012] that all shopping negotiates discrepancies between normative and actual social order (for instance, between the ideals of health emblematised by organic products and commitments to thrift embodied in lower cost generic products).
 Miller suggests that household shopping practices attempt to resolve differences between how people think they should live and how they actually live: 'we have to watch how shopping helps resolve these discrepancies between the normative and the actual, but we also need some ideas of where the normative comes from in the first place' [@Miller_2012, 72].
@@
 
-Shopping lists provisionally stabilise the complex social ordering of grocery shopping.
-They provide important clues as to how local social order is constituted, maintained, and repaired in shopping.
-As people shop, either by  trawling along aisles packed with thousands of products, or scrolling down screens or searching for particular brands amidst search results, lists filter or reduce the excessive abundance, claims on attention, dazzle and distraction of commodities to the practicalities of domestic economy.
-The shopping list, whether written on the back of an envelope, saved as a list in an online grocery shopping system, or remembered, lies at the intersection of logistical flows, infrastructural orderings, and  lively negotiations around actual and normative social orders.[^30.1]  Shopping lists  are intersectional ordering devices that encapsulate a universe of possible references, and a teeming multitude of propensities with an actual local order.[^1.456]
+In contrast to many other forms of online and offline shopping, shopping lists distinctively, albeit provisionally, stabilise the complex social ordering of grocery shopping.
+Shopping lists provide important clues as to how local social order is constituted, maintained, and repaired in grocery shopping.
+As people shop, either by  trawling along aisles packed with thousands of products, or scrolling down screens or searching for particular brands amidst search results, lists may filter or reduce the excessive abundance, claims on attention, dazzle and distraction of commodities to the practical social order of domestic economy.
+The shopping list, whether written on the back of an envelope, saved as a list in an online grocery shopping system, or memorised, lies at the intersection of logistical flows, infrastructural orderings, and lively negotiations around actual and normative social orders.[^30.1]  Shopping lists  are intersectional ordering devices that encapsulate a universe of possible references, and a teeming multitude of propensities with an actual local order.[^1.456] In contrast to much e-commerce where each purchase concerns small numbers of items, the numerous items on a grocery shopping list suggest that online grocery shopping, and attempts to personalize it, will both provide many opportunities for recommendation (and hence conversion events) as well as many complexities in addressing normative and actual social orders.
 
@@
 Hand-written shopping lists have ongoing practical importance and mixed ordering practices (see the montage of handwritten shopping lists at [Grocery List](http://www.grocerylist.org)).
-Online shopping lists, by contrast,  reside at the intersection of web and internet infrastructures, supply chain logistics, individualised practices and *habitus,* and increasingly, the predictive operations of  recommender systems.
+Online grocery shopping lists, by contrast,  reside at the intersection of web and internet infrastructures, supply chain logistics, individualised practices and *habitus,* and increasingly, the predictive operations of  recommender systems.
 Whereas the aisles and shelves of a supermarket present a densely-woven mesh of objects competing for visual attention by offering distinctions of taste, thrift, expedience, novelty, indulgence, health, online shopping recommender systems generate lists that seek to align people to products that they otherwise might have little relation to (see [@Turow_2015] for an overview of the development of these systems).
@@
 From both  anthropological and literary perspectives, shopping lists have excellent reasons to exist: they are deeply rooted in organisational life and infrastructures.
-Lists weave together people, infrastructures, things, and places.While lists intersect strongly with other meaning and sense-making practices in everyday life and popular culture, they can be refractory to discourse  and textual analysis methods built around models of language or speech, with its rules of syntax and grammar.
+Lists weave together people, infrastructures, things, and places.
+While lists intersect strongly with other meaning and sense-making practices in everyday life and popular culture, they can be refractory to discourse  and textual analysis methods built around models of language or speech, with its rules of syntax and grammar.
 On the one hand, lists are highly associative.
@@
 
-[^6531] The web pages of the big data analytic companies providing products to various sectors of industry, government and media are stocked with images, graphics and statements centred on this conversion experience.
+[^6531]: The web pages of the big data analytic companies providing products to various sectors of industry, government and media are stocked with images, graphics and statements centred on this conversion experience.
 [@Amoore_2015] analyses some of these.

### 2018-04-16 log.rmd
start chapter branch

@@
 - revised conclusion
+
+## Mon 16 Apr 2018 09:05:53 BST
+- received reports back:
+
+    From:	Big Data & Society <onbehalfof@manuscriptcentral.com>
+    Reply-to:	e.s.ruppert@mac.com
+    To:	a.mackenzie@lancaster.ac.uk
+    Subject:	Big Data & Society - Decision on Manuscript ID BDS-17-0074.R1
+    Date:	Sat, 31 Mar 2018 11:05:35 -0400 (31/03/18 16:05:35)
+    31-Mar-2018
+
+    Dear Adrian:
+
+    We hope you are well!  Good to see your submission move along. Here is the standard decision letter for minor revisions.....
+
+    Manuscript ID BDS-17-0074.R1 entitled "Personalization and probabilities: impersonal propensities in online grocery shopping" which you submitted to Big Data & Society, has been reviewed.
+The comments of the reviewer(s) are included at the bottom of this letter.
+
+    The reviewer(s) and co-editor have recommended publication subject to the completion of some minor revisions to your paper.
+Therefore, I invite you to respond to their comments and revise your paper.
+
+    To revise your paper, log into https://mc.manuscriptcentral.com/bdas and enter your Author Center, where you will find your paper title listed under "Manuscripts with Decisions."  Under "Actions," click on "Create a Revision."  Your paper number has been appended to denote a revision.
+
+    You may also click the below link to start the revision process (or continue the process if you have already started your revision) for your paper.
+If you use the below link you will not be required to login to ScholarOne Manuscripts.
+
+    *** PLEASE NOTE: This is a two-step process. After clicking on the link, you will be directed to a webpage to confirm. ***
+
+    https://mc.manuscriptcentral.com/bdas?URL_MASK=da35635913af44a5a940c7c2973881ca
+
+    You will be unable to make your revisions on the originally submitted version of the paper.
+Instead, revise your paper using a word processing program and save it on your computer.
+Please also highlight the changes to your paper within the document by using bold or colored text.
+
+    Once the revised paper is prepared, you can upload it and submit it through your Author Center.
+
+    When submitting your revised paper, you will be able to respond to the comments made by the reviewer(s) in the space provided.
+You can use this space to document any changes you make to the original paper.
+In order to expedite the processing of the revised paper, please be as specific as possible in your response to the reviewer(s) as they will be forwarded on your responses.
+Please also ensure that any response you make to reviewer comments remain anonymous.
+
+    IMPORTANT:  Your original files are available to you when you upload your revised paper.  Please delete any redundant files before completing the submission.
+
+    Because we are trying to facilitate timely publication of papers submitted to Big Data & Society, your revised paper should be uploaded as soon as possible and no later than four weeks of receipt of this letter.
+If it is not possible for you to submit your revision in this time, please let me know otherwise we may have to consider your paper as a new submission.
+
+    Once again, thank you for submitting your paper to Big Data & Society and I look forward to receiving your revision.
+
+    Sincerely,
+    Evelyn
+    Professor Evelyn Ruppert
+    Editor in Chief, Big Data & Society
+    e.s.ruppert@mac.com
+
+    Co-Editor(s)' Comments to Author:
+
+    Co-Editor: Gabrys, Jennifer
+    Comments to the Author:
+    Based on the reviewers' comments, we are recommending minor revisions for your manuscript.
+
+    There are a few additional points suggested in the reviewers’ comments, which should be addressed when making your revisions, including:
+
+    * Per R1, make more explicit the industry data scientist presentation as the basis for your material, and per R2 ensure this material is addressed both in terms of method and ethnographic contribution.
+
+    *Per R2 and R1, briefly contextualise online grocery shopping in relation to other forms of e-commerce and other personalization concepts.
+
+    *Per R1, address footnote formatting issues, and per R2 ensure all footnotes are added as indicated, and clarify usage of the ‘conversion event’ throughout.
+
+    There are additional detailed comments to be found in the reviewers’ reports, which you can address at your discretion.
+
+    We look forward to receiving your revised manuscript taking the above notes specifically into account, as well as the detailed comments provided in the reviews.
+
+    Reviewer(s)' Comments to Author:
+
+    Reviewer: 1
+
+    Comments to the Author
+    Thank you for making changes to this paper.
+The justification that what wanted to focus on was what was presented by the industry data scientist, and that her presentation framed the sample of literature you drew upon, should be explicitly in the text.
+Similarly, it would be useful to see a more explicit flag to how grocery shopping differs from traditional e-commerce, as you said.
+
+    There's also a footnote error that looks like it comes from markdown on pages 14 (line 20) and 13 (line 28).
+
+    Beyond putting those simple clarifications you make in the review into the text, I am now happy with the changes you have made to this work and recommend it for publication.
+
+    Reviewer: 2
+
+    Comments to the Author
+    Given the dismissive nature of the author's responses to the majority of my report, I will refer directly to these responses here, and explain the motivation for some of my comments which remains unchanged with the new version of the paper.
+
+    My frustration with the paper was not based on some fundamental disagreement with the methodology employed or theoretical foundation of the paper, but rather its imprecision.
+While I can tell I do not come from the same disciplinary background as the author, I should nonetheless be able to understand the thesis of the paper and how each component or section (or in this case, each piece of the mixed methodology) contributes to this thesis.
+This was my chief criticism of the first version of the paper: the various components are not sufficiently contextualised to explain their contribution to the overall thesis of the paper.
+In some cases, it felt as these gaps are papered over with needlessly complex language and construction.
+The author is correct that I was left frustrated by the paper, but this frustration was borne more of issues with the structure and presentation of the paper than any fundamental disciplinary, methodological or epistemological disagreements.
+This frustration remains with the new draft of the paper.
+For a journal that embraces methodological pluralism as openly as Big Data & Society, surely contributions should be comprehensible for readers not sharing the author's particular disciplinary background?
+
+    Responses to the author's responses:
+
+    "- R2 asked for 'specifics about how the analysis may translate to other personalization contexts, despite
+    this being promised.' Although I did claim the case might be 'sociologically resonant,' I don't think I promised
+    to translate to other personalization contexts. (Reading abstract, introduction and conclusion shows no sign
+    of that promise.) It would be a significant new section of the paper to specify the general argument for other
+    concrete contexts."
+
+    From the abstract: "This paper argues that personalization through models using large-scale data is part of a more expansive change in probabilization that, in principle, is not reducible to individual or ‘personal’ attributes and actions.
+It describes the ‘personalization’ of an online grocery shopping recommender system to list a small number of grocery items of personal relevance for each of the millions of online grocery shoppers at a major UK supermarket chain." To me this clearly says that you will use personalization in online grocery shopping to make a broader point about personalization in large-scale data.
+Perhaps I am reading too much into this.
+I think what you mean to say is that you demonstrate that elements of a broader, previously described trend can be found in online grocery shopping.
+
+    "- R2 found the borrowing and repurposing of the term 'conversion event' from the recommender system
+    literature confusing. I've added a footnote to explain the motivation for this borrowing, but I probably won't be
+    able to convince R2 of the value of a material-semiotic approach in which such dual-register terms are quite
+    common."
+
+    I cannot find the footnote referenced here, can the author specify which footnote specifically by number?
+My point was about clarity of language for the reader.
+Is the term always used in its dual sense throughout the paper?
+If not, at a minimum each usage should refer to which dimension is relevant.
+
+    "- R2 was concerned about the confusing methodology. I've added a sentence or two in the methodology
+    section saying what the paper does. But I won't, I fear, be able to assuage all R2's concerns. They don't
+    countenance what I'm trying to do: ground a theoretical argument about changing modes of operational
+    knowledge by assembling some empirical engagements -- the ethnographic event, the small experiments
+    with Tesco APIs and recommender system models, the existing sociologies and anthropologies of shopping
+    and lists and the archaeology of recommendation. It's not a conventional empirical research project, and the
+    methodology of reconstruction that I use (borrowing from Dewey) does not conform to narrowly defined
+    sociological requirements. Given the methodological pluralism of the journal, I'm hoping that this is viable."
+
+    I do not disagree with the methodology used by the author.
+My concern was that the contribution of each element to the paper's thesis was not sufficiently explained or justified.
+In the original draft, the methodology came across as a convenience sample of methods, rather than a deliberately planned study.
+This concern has been partially addressed by changes in the new draft.
+
+    "- R2 wasn't able to see what the ethnographic material adds to analysis. They could refer to p3, li 44-48 (in
+    pre-revision manuscript) where I set out how I understand it is as part of a 'big data conversion event.'"
+
+    The lines referred to explain the intended contribution of the ethnographic material.
+My criticism was about what the actual material tells us that is relevant to the paper's overall thesis.
+I understand the intended contribution of the material perfectly well; I am not convinced that the material is sufficiently well discussed to actually contribute what has been proposed.
+I gave an example from the dunnhumby presentation where the relevance of a particular detail the author chose to highlight was never explained to the reader.
+
+    "- R2 couldn't work out the 'intended meaning' of some key sentences. There was a grammar problem in
+    these sentences, which I've fixed. I'm hoping that this was the problem with those sentences not the actual
+    content.
+    - I've moved key part of footnote 5 into the introduction. That was a useful suggestion.
+    - I've shortened the historical-archaeology of recommender system, and explained its purpose in the
+    introductory overview.
+    - The section on limitations of association rule systems has been re-framed in the introduction/overview,
+    and I've changed sentences at the close of the section explaining why attention to limitations is relevant to
+    the main argument about probabilities."
+
+    These were all welcome changes.
+
+-------------------------------------------------
+
+- starting to do the minor revisions; I'm not reading the detail reports, just the editor requests with my replies, as pasted in ScholarOne:
+    * Per R1, make more explicit the industry data scientist presentation as the basis for your material, and per R2 ensure this material is addressed both in terms of method and ethnographic contribution.
+    DONE: added a new sentence in Section B, Para 1, which I hope addresses this.
+
+    *Per R2 and R1, briefly contextualise online grocery shopping in relation to other forms of e-commerce and other personalization concepts.
+    DONE: Rewrote much of the third paragraph of the 'shopping lists' to contrast online grocery shopping with other forms of shopping. New sentences added to the end of this paragraph.
+
+    *Per R1, address footnote formatting issues, and per R2 ensure all footnotes are added as indicated, and clarify usage of the ‘conversion event’ throughout.
+    DONE: apologies to R2 -- I said I defined 'conversion event' in a footnote.
+I had done that, but then decided the definition should be in the opening paragraph of the manuscript.
+It's there.
+I also  fixed the footnote formatting issue mentioned by R1
+
+- will do this in reconstruction.rmd
+- submitting to the Mansuscript-central. Will tag this revision as bdas-final-revision
+- had to do the bibliography using 12 Feb 2018 ref_bibs commit; I've tried to tag it as bdas-references

### 2018-05-16 reconstruction.rmd
starting proofs

@@
 a vegan customer will never have purchased chicken products, so the measured propensities for any of the 1000 or so chicken products sold be Tesco will remain close to `0`).
-But the changing situation of the customer only figures here in the accumulated year of purchase data, and that past purchases might precisely what does not need to recommended.
+But the changing situation of the customer only figures here in the accumulated year of purchase data.
+Past purchases might precisely be  what does not need to be recommended.
 

### 2018-05-17 reconstruction.rmd
fixing references for proofs
sending proofs with figures

@@
 library(tint)
-knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tint'))
+knitr::opts_chunk$set(tidy = FALSE, dev='svg',  cache.extra = packageVersion('tint'))
 options(htmltools.dir.version = FALSE)
@@
 
-```{r humby-1984, fig.cap = 'Abstract from a Clive Humby presentation (Humby,1984)'}
+```{r humby-1984, fig.cap = 'Abstract from a Clive Humby presentation in 1984 [@O'Keefe,1984]'}
         library(knitr)
@@
 
-```{r arules-ex, cache=TRUE, fig.cap='Frequency of items in the `Grocery` dataset', message=FALSE, echo=FALSE, fig.margin=TRUE}
+```{r arules-ex, cache=TRUE, fig.cap='Frequency of items in the `Grocery` dataset', message=FALSE, echo=FALSE, fig.margin=TRUE, dev='svg'}
 library(knitr)
@@
 Given almost 2500 `sauces` and 1200 `rice` products listed by the Tesco API,  a tremendous number of associations between sauce and rice are possible.
-The propensities of any given sauce and rice product to find themselves together in a shopping basket greatly.
+The propensities of any given sauce and rice product to find themselves together in a shopping basket vary greatly.
 The proliferation of things on the shelves of supermarket or grocery warehouse produces a combinatorial problem for data-mining machine learning approaches such as association rules.

### 2018-05-22 log.rmd
fixing references for proofs
fixing proofs

@@
 - had to do the bibliography using 12 Feb 2018 ref_bibs commit; I've tried to tag it as bdas-references
+
+## Wed 16 May 2018 19:09:07 BST
+- doing proofs  BDS778310.pdf 
+- TODO:  
+    - look up Sarwar conference location
+    - Ardissono editors, etc.Goy, Anna, Liliana Ardissono, and Giovanna Petrone.
+‘Personalization in E-Commerce Applications’.
+In The Adaptive Web, edited by P.
+Brusilovsky, A.
+Kobsa, and W Nejdl, 485–520.
+Springer, 2007.
+
+## Tue 22 May 2018 14:29:45 BST
+- doing proofs again -- there were some more typos, etc.
